### 2020-10-15 README.Rmd
add readme

---
title: "Road Segment Prioritization for Bicycle Infrastructure"
output: github_document
fontsize: 10pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Missing Data

There are a couple of files that cannot be synced to github due to their size. These files are neseccary for the scripts to run. Below are links to where you can download them, and instructions on where to place them in the repo file structure

Flow Data (2011 Census Origin-Destination Data):

* Source: https://www.nomisweb.co.uk/census/2011/bulk/rOD1  ---> Choose File **"WU03EW"**
* Location in Repo: data-raw/flow_data.csv

Middle Layer Super Output Areas (December 2011) Boundaries:

* Source: http://geoportal.statistics.gov.uk/datasets/826dc85fb600440889480f4d9dbb1a24_0
* Location in Repo: data-raw/MSOA_2011_Boundaries/[Add files here]

***

## Scripts

The scripts should be run in the order they are numbered in (and listed in here). The only exception is _x_dodgr_weighting_profiles.R.

Check the readme [here](https://github.com/Hussein-Mahfouz/Bicycle-Network-Optimization) for detailed info on each script

### 2020-10-17 Paper/journal-article/journal-article.Rmd
rticles package elsevier template

---
title: Short Paper
author:
  - name: Alice Anonymous
    email: alice@example.com
    affiliation: Some Institute of Technology
    footnote: 1
  - name: Bob Security
    email: bob@example.com
    affiliation: Another University
  - name: Cat Memes
    email: cat@example.com
    affiliation: Another University
    footnote: 2
  - name: Derek Zoolander
    email: derek@example.com
    affiliation: Some Institute of Technology
    footnote: 2
address:
  - code: Some Institute of Technology
    address: Department, Street, City, State, Zip
  - code: Another University
    address: Department, Street, City, State, Zip
footnote:
  - code: 1
    text: "Corresponding Author"
  - code: 2
    text: "Equal contribution"
abstract: |
  This is the abstract.

  It consists of two paragraphs.

journal: "An awesome journal"
date: "`r Sys.Date()`"
bibliography: mybibfile.bib
#linenumbers: true
numbersections: true
csl: elsevier-harvard.csl
output: rticles::elsevier_article
---

Introduction
============

What Affects the Decision To Cycle
==================================

Planning Cycling Networks
=========================

Potential Cycling Demand
========================

\begin{align}\label{eq:pcteqn}
     logit(C_{p}) = & -4.018 - 0.6369d +  1.988\sqrt{d} + 0.008775d^2\\ & - 0.2555s + 0.00206ds -0.1234\sqrt{d}s\nonumber 
\end{align}

Routing
=======

Road Segment Prioritization
===========================

Analysis of Methodology
=======================

Conclusions
===========

References {#references .unnumbered}
==========

### 2020-10-17 Paper/journal-article/journal-article.tex
rticles package elsevier template

\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
\usepackage[hyphens]{url}

  \journal{An awesome journal} % Sets Journal name

\usepackage{lineno} % add
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{graphicx}
\usepackage{booktabs} % book-quality tables
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\bibliographystyle{elsarticle-harv}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Short Paper},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\setcounter{secnumdepth}{5}
% Pandoc toggle for numbering sections (defaults to be off)

% Pandoc header

\begin{document}
\begin{frontmatter}

  \title{Short Paper}
    \author[Some Institute of Technology]{Alice Anonymous\corref{1}}
   \ead{alice@example.com} 
    \author[Another University]{Bob Security}
   \ead{bob@example.com} 
    \author[Another University]{Cat Memes\corref{2}}
   \ead{cat@example.com} 
    \author[Some Institute of Technology]{Derek Zoolander\corref{2}}
   \ead{derek@example.com} 
      \address[Some Institute of Technology]{Department, Street, City, State, Zip}
    \address[Another University]{Department, Street, City, State, Zip}
      \cortext[1]{Corresponding Author}
    \cortext[2]{Equal contribution}

  \begin{abstract}
  This is the abstract.

  It consists of two paragraphs.
  \end{abstract}

 \end{frontmatter}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(dplyr)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
## Attaching package: 'dplyr'
\end{verbatim}

\begin{verbatim}
## The following objects are masked from 'package:stats':
## 
##     filter, lag
\end{verbatim}

\begin{verbatim}
## The following objects are masked from 'package:base':
## 
##     intersect, setdiff, setequal, union
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(readr)}
\KeywordTok{library}\NormalTok{(tidyr)}
\KeywordTok{library}\NormalTok{(ggplot2)}
\KeywordTok{library}\NormalTok{(ggrepel)}

\CommentTok{# this is a lookup table matching MSOAs to major towns and cities}
\NormalTok{city_names <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{'../data-raw/Middle_Layer_Super_Output_Area__2011__to_Major_Towns_and_Cities__December_2015__Lookup_in_England_and_Wales.csv'}\NormalTok{) }
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Parsed with column specification:
## cols(
##   MSOA11CD = col_character(),
##   MSOA11NM = col_character(),
##   TCITY15CD = col_character(),
##   TCITY15NM = col_character(),
##   FID = col_double()
## )
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# change column name}
\NormalTok{city_names <-}\StringTok{ }\NormalTok{city_names }\OperatorTok{%>%}\StringTok{ }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{city =}\NormalTok{ TCITY15NM)}

\CommentTok{#unique cities}
\KeywordTok{unique}\NormalTok{(city_names}\OperatorTok{$}\NormalTok{city)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##   [1] "London"               NA                     "Bolton"              
##   [4] "Stockport"            "Rochdale"             "Salford"             
##   [7] "Manchester"           "Oldham"               "Bury"                
##  [10] "Wigan"                "Rotherham"            "Liverpool"           
##  [13] "Doncaster"            "Sheffield"            "St Helens"           
##  [16] "Barnsley"             "Southport"            "Birkenhead"          
##  [19] "Birmingham"           "Coventry"             "Dudley"              
##  [22] "Newcastle upon Tyne"  "Sunderland"           "Sutton Coldfield"    
##  [25] "South Shields"        "West Bromwich"        "Gateshead"           
##  [28] "Leeds"                "Wolverhampton"        "Bradford"            
##  [31] "Walsall"              "Huddersfield"         "Wakefield"           
##  [34] "Solihull"             "Halifax"              "Leicester"           
##  [37] "Nottingham"           "Hartlepool"           "Derby"               
##  [40] "Blackpool"            "Grimsby"              "Warrington"          
##  [43] "Scunthorpe"           "Kingston upon Hull"   "York"                
##  [46] "Blackburn"            "Middlesbrough"        "Stockton-on-Tees"    
##  [49] "Darlington"           "Telford"              "Stoke-on-Trent"      
##  [52] "Swindon"              "Bristol"              "Peterborough"        
##  [55] "Luton"                "Southend-on-Sea"      "Gillingham"          
##  [58] "Plymouth"             "Weston-Super-Mare"    "Bath"                
##  [61] "Bournemouth"          "Poole"                "Portsmouth"          
##  [64] "Milton Keynes"        "Southampton"          "Bedford"             
##  [67] "Reading"              "Chatham"              "Bracknell"           
##  [70] "Slough"               "High Wycombe"         "Brighton and Hove"   
##  [73] "Cambridge"            "Carlisle"             "Chester"             
##  [76] "Chesterfield"         "Chelmsford"           "Basildon"            
##  [79] "Eastbourne"           "Hastings"             "Exeter"              
##  [82] "Colchester"           "Harlow"               "Cheltenham"          
##  [85] "Gloucester"           "Hemel Hempstead"      "St Albans"           
##  [88] "Stevenage"            "Basingstoke"          "Burnley"             
##  [91] "Watford"              "Maidstone"            "Preston"             
##  [94] "Norwich"              "Lincoln"              "Harrogate"           
##  [97] "Northampton"          "Newcastle-under-Lyme" "Oxford"              
## [100] "Mansfield"            "Shrewsbury"           "Burton upon Trent"   
## [103] "Woking"               "Guildford"            "Ipswich"             
## [106] "Crawley"              "Nuneaton"             "Worthing"            
## [109] "Redditch"             "Worcester"            "Swansea"             
## [112] "Cardiff"              "Newport"
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# number of MSOAs in each city }
\NormalTok{no_msoas <-}\StringTok{ }\NormalTok{city_names }\OperatorTok{%>%}\StringTok{ }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{group_by}\NormalTok{(city) }\OperatorTok{%>%}\StringTok{ }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{tally}\NormalTok{()}

\CommentTok{##### CHOOSE YOU CITY }
\NormalTok{chosen_city <-}\StringTok{ "Manchester"}
\CommentTok{#create a directory to store data related to this city (does nothing if directory already exists)}
\KeywordTok{dir.create}\NormalTok{(}\KeywordTok{paste0}\NormalTok{(}\StringTok{"../data/"}\NormalTok{, chosen_city), }\DataTypeTok{showWarnings =} \OtherTok{FALSE}\NormalTok{)}
\CommentTok{# create sub-directory to save plots as well}
\KeywordTok{dir.create}\NormalTok{(}\KeywordTok{paste0}\NormalTok{(}\StringTok{"../data/"}\NormalTok{, chosen_city,}\StringTok{"/Plots"}\NormalTok{), }\DataTypeTok{showWarnings =} \OtherTok{FALSE}\NormalTok{)}

\CommentTok{##### CHOOSE YOU CITY }

\CommentTok{# flow data from the 2011 census https://www.nomisweb.co.uk/census/2011/bulk/rOD1}
\NormalTok{flows <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{'../data-raw/flow_data.csv'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Parsed with column specification:
## cols(
##   `Area of residence` = col_character(),
##   `Area of workplace` = col_character(),
##   `All categories: Method of travel to work` = col_double(),
##   `Work mainly at or from home` = col_double(),
##   `Underground, metro, light rail, tram` = col_double(),
##   Train = col_double(),
##   `Bus, minibus or coach` = col_double(),
##   Taxi = col_double(),
##   `Motorcycle, scooter or moped` = col_double(),
##   `Driving a car or van` = col_double(),
##   `Passenger in a car or van` = col_double(),
##   Bicycle = col_double(),
##   `On foot` = col_double(),
##   `Other method of travel to work` = col_double()
## )
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{###############}
\CommentTok{# MERGING NAMES WITH FLOW DATA (TO GET INTERNAL FLOWS IN ANY CITY)}
\CommentTok{###############}

\CommentTok{# add a column with the city name corresponding to each Residence MSOA}
\NormalTok{flows <-}\StringTok{ }\NormalTok{flows }\OperatorTok{%>%}\StringTok{ }\KeywordTok{left_join}\NormalTok{(city_names[,}\KeywordTok{c}\NormalTok{(}\StringTok{"MSOA11CD"}\NormalTok{, }\StringTok{"city"}\NormalTok{)],}
                             \DataTypeTok{by =} \KeywordTok{c}\NormalTok{(}\StringTok{"Area of residence"}\NormalTok{ =}\StringTok{ "MSOA11CD"}\NormalTok{)) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{city_origin =}\NormalTok{ city) }\CommentTok{# rename column so that we know it is referring to the 'Area of residence'}

\CommentTok{# add a column with the city name corresponding to each Workplace MSOA}
\NormalTok{flows <-}\StringTok{ }\NormalTok{flows }\OperatorTok{%>%}\StringTok{ }\KeywordTok{left_join}\NormalTok{(city_names[,}\KeywordTok{c}\NormalTok{(}\StringTok{"MSOA11CD"}\NormalTok{, }\StringTok{"city"}\NormalTok{)],}
                             \DataTypeTok{by =} \KeywordTok{c}\NormalTok{(}\StringTok{"Area of workplace"}\NormalTok{ =}\StringTok{ "MSOA11CD"}\NormalTok{)) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{city_dest =}\NormalTok{ city) }\CommentTok{# rename column so that we know it is referring to the 'Area of workplace'}

\CommentTok{# get mode share for all cities - just for report}
\NormalTok{cycle_mode_share <-}\StringTok{ }\NormalTok{flows }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{filter}\NormalTok{(city_origin }\OperatorTok{==}\StringTok{ }\NormalTok{city_dest) }\OperatorTok{%>%}\StringTok{   }\CommentTok{# only internal flows}
\StringTok{  }\KeywordTok{group_by}\NormalTok{(city_origin, city_dest) }\OperatorTok{%>%}\StringTok{   }
\StringTok{  }\KeywordTok{summarize}\NormalTok{(}\DataTypeTok{mode_share =}\NormalTok{ (}\KeywordTok{sum}\NormalTok{(Bicycle) }\OperatorTok{/}\StringTok{ }\KeywordTok{sum}\NormalTok{(}\StringTok{`}\DataTypeTok{All categories: Method of travel to work}\StringTok{`}\NormalTok{)) }\OperatorTok{*}\DecValTok{100}\NormalTok{,}
            \DataTypeTok{all =} \KeywordTok{sum}\NormalTok{(}\StringTok{`}\DataTypeTok{All categories: Method of travel to work}\StringTok{`}\NormalTok{)) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{arrange}\NormalTok{(mode_share)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## `summarise()` regrouping output by 'city_origin' (override with `.groups` argument)
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# plot }
\NormalTok{cycle_mode_share }\OperatorTok{%>%}\StringTok{ }\CommentTok{#filter(city_origin != 'London') %>%}
\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(all, mode_share)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_point}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{alpha =}\NormalTok{ mode_share), }\DataTypeTok{show.legend =} \OtherTok{FALSE}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\CommentTok{# some filtering labels for aesthetic purposes. Add some cities explicitly as they are in my case study}
\StringTok{  }\KeywordTok{geom_label_repel}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{label =} \KeywordTok{ifelse}\NormalTok{(mode_share}\OperatorTok{>}\DecValTok{12} \OperatorTok{|}\StringTok{ }\NormalTok{all}\OperatorTok{>}\StringTok{ }\NormalTok{(}\FloatTok{2.3}\OperatorTok{*}\KeywordTok{mean}\NormalTok{(all)) }\OperatorTok{|}\StringTok{ }\NormalTok{city_origin }\OperatorTok{%in%}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{'Manchester'}\NormalTok{, }\StringTok{'Nottingham'}\NormalTok{, }\StringTok{'Birmingham'}\NormalTok{) , }\KeywordTok{as.character}\NormalTok{(city_origin),}\StringTok{''}\NormalTok{)), }
                       \DataTypeTok{size =}\FloatTok{2.5}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{x=}\StringTok{"Total Number of Commuters"}\NormalTok{, }\DataTypeTok{y =} \StringTok{"Cycling Mode Share (%)"}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{scale_x_continuous}\NormalTok{(}\DataTypeTok{trans=}\StringTok{'log10'}\NormalTok{, }\DataTypeTok{labels =}\NormalTok{ scales}\OperatorTok{::}\NormalTok{comma) }\OperatorTok{+}
\StringTok{  }\KeywordTok{theme_minimal}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\includegraphics{journal-article_files/figure-latex/unnamed-chunk-1-1.pdf}

\hypertarget{what-affects-the-decision-to-cycle}{%
\section{What Affects the Decision To
Cycle}\label{what-affects-the-decision-to-cycle}}

\hypertarget{planning-cycling-networks}{%
\section{Planning Cycling Networks}\label{planning-cycling-networks}}

\hypertarget{potential-cycling-demand}{%
\section{Potential Cycling Demand}\label{potential-cycling-demand}}

\begin{align}\label{eq:pcteqn}
     logit(C_{p}) = & -4.018 - 0.6369d +  1.988\sqrt{d} + 0.008775d^2\\ & - 0.2555s + 0.00206ds -0.1234\sqrt{d}s\nonumber 
\end{align}

\hypertarget{routing}{%
\section{Routing}\label{routing}}

\hypertarget{road-segment-prioritization}{%
\section{Road Segment
Prioritization}\label{road-segment-prioritization}}

\hypertarget{analysis-of-methodology}{%
\section{Analysis of Methodology}\label{analysis-of-methodology}}

\hypertarget{conclusions}{%
\section{Conclusions}\label{conclusions}}

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\end{document}

### 2020-10-23 Paper/journal-article/journal-article.Rmd
progress on paper

@@
 date: "`r Sys.Date()`"
+geometry: margin=2cm
 bibliography: mybibfile.bib
@@
 
-What Affects the Decision To Cycle
-==================================
+Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
+Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
+Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
+Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
+Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
+Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
+Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
 
-Planning Cycling Networks
-=========================
+Background
+==========
 
-Potential Cycling Demand
-========================
+## What Affects the Decision To Cycle
+
+Segregated cycling infrastructure\footnote{Segregated cycling infrastructure refers to road space that is allocated to cyclists only, with physical separation to protect cyclists from other modes of transport.} has been shown to increase cycling uptake [@aldred2019impacts; @goodman2014new; @marques2015infrastructure], with the separation from motorized vehicles being key [@winters2011motivators].
+Revealed preference of cyclists shows that they are willing to deviate from the most efficient routes in order to commute on safer roads [@crane2017longitudinal].
+However, such deviations are only considered if they do not considerably increase route circuitry; behaviour studies have found that the probability of choosing a route decreases in proportion to its length relative to the shortest route [@broach2011bicycle; @winters2010far].
+Another defining feature for cycling infrastructure is how well connected it is.
+Cyclists prefer cohesive infrastructure, particularly when cycling on arterial roads with high levels of motorized traffic [@stinson2003commuter], and the lack of well-connected cycling infrastructure is one of the main obstacles to increasing cycling uptake [@caulfield2012determining].
+While direct and cohesive cycling networks have been shown to positively impact cycling rates, density\footnote{making an area's bicycle network denser means adding more cycling routes in the area and thereby giving cyclists more route options} of the cycling network is also vital [@schoner2014missing].
+
+## Planning Cycling Networks
+
+*Optimization* techniques have been used to propose improvements to cycling networks.
+@mesbah2012bilevel propose a bi-level formulation to optimize allocation of cycling lanes to the network without exceeding a set budget.
+The upper level is the proposed interventions and the lower level is the route choices made by users in reaction to changes in the network.
+The problem accounts for the effect of cycling lanes on car traffic, and attempts to maximize utilization of said lanes with minimal impact on car travel times.
+To improve cohesion of the suggested network, a constraint is added so that each link with a bike lane should be connected to at least one destination.
+Car usage is not considered by @mauttone2017bicycle, who develop an optimization framework that aims to minimize the total user cost of cycling on the network.
+The aggregate flow on links is obtained by using shortest paths to route existing cycling demand onto the road network, and the solution is a proposed set of links where cycling infrastructure should be added in order to minimize the overall travel cost of cyclists across the network.
+The cost of traversing a link is given as a function of its length and whether or not it has cycling infrastructure, and a discontinuity penalty is also added to prioritize connected road segments.
+The problem has also been solved by attempting to find the minimum cost of improving roadway links to meet a desired level of service (LOS) [@duthie2014optimization].
+In this formulation, all OD pairs need to be connected by roads that meet the desired LOS, and a directness constraint is added so that paths between OD pairs do not exceed a certain multiple of the shortest path.
+
+These problem formulations do not explicitly solve for continuity, which is dealt with using a either (a) a constraint specifying that each link with a bike lane should be connected to at least one destination [@mesbah2012bilevel], (b) a constraint on deviation from shortest paths [@duthie2014optimization], or (c) a discontinuity penalty [@mauttone2017bicycle].
+To solve for continuity, the graph-theoretic concept of *connected components*, has been used.
+@natera2019data study the existing cycling network in terms of its disconnected components and introduce two different algorithms to connect these components by their most critical links\footnote{\textit{link} refers to a road segment throughout this research} and, in doing so, measure the size of the growth of the largest connected component as a function of the kilometers of network added.
+They observe that small investments at strategic points have a large impact on connectivity in most cases.
+The concept of connected components is also at the core of the methodology proposed by @olmos2020data.
+After routing the cycling demand onto the network links, they use percolation theory to filter out the links based on the aggregate flow\footnote{\textit{flow} is used throughout this research to refer to the cycling demand when it is routed onto the road network.
+The flow on any road segment is the cumulative demand on it, resulting from cyclists commuting between various OD pairs} passing through them.
+They vary the flow threshold for filtering to identify the minimum flow at which the whole city is connected by a giant component.
+The results show a cycling network that connects the entire city, and subtracting links intersecting with current cycling infrastructure identifies links proposed for intervention.
+
+Data and Geographical Scale of Analysis
+=======================================
+
+The analysis is heavily dependant on Origin-Destination census data (commuter data).
+Commuter data in the UK is publicly available at the Middle layer Super Output Area (MSOA) level; the average MSOA has a population of 8209 [@ofn2018population].
+@iacono2010measuring note that such large travel zones are not ideal for understanding route choice behaviour of cyclists and pedestrians.
+They also give rise to an `ecological fallacy' whereby average characteristics are assumed to apply to all residents of the aggregated geographical area.
+Given that more granular data is not publicly available, the study uses MSOA-level commuter data.
+The methodology is however applicable to more granular commuter data should it become available.
+
+Calculating Potential Cycling Demand
+====================================
+
+The Propensity to Cycle Tool (PCT) [@lovelace2017propensity] is used to estimate the proportion of cyclists ($\boldsymbol{C_{p}}$) for each MSOA pair should the government achieve its target of doubling cycling by 2025.
+The PCT uses the following logistic regression model to calculate  $\boldsymbol{C_{p}}$:
 
@@
 
+\noindent where $\boldsymbol{d}$ and $\boldsymbol{s}$ are the distance and slope respectively for the OD pair.
+The authors use square and square-root distance terms ``to capture the non-linear impact of distance on the likelihood of cycling", and interaction terms to capture the combined effect of slope and distance [@lovelace2017propensity].
+
+The potential demand calculations show that the current and potential number of cyclists both follow a bell-shaped distribution, with the number of trips peaking around the 3-5km commuting distance and then going back down for longer distances (See Figure \ref{fig:pot_dem_histograms}).
+
+\begin{figure}[h!]
+    \centering % puts the image in the horizontal centre of the page
+    \includegraphics[width=.6\textwidth]{../../data/Manchester/Plots/desire_facet_cycling.png} 
+    \caption{Current and Potential Cycling Demand} 
+    \label{fig:desire_facet_cycling} 
+\end{figure}
+
 Routing
@@
 
+The next step is to route the potential cycling demand ($\boldsymbol{C_{p}}$) between all OD pairs onto the road network.
+
+To conduct routing, the following is considered:
+
+1. **Cyclist Preference**:
+     Work done by @dill2013four on examining cyclist typologies determined that around 60\% of Portland residents fit under the *interested but concerned* category.
+These were people that enjoyed cycling but avoided it due safety concerns.
+The key to encouraging this group was to create a low-stress cycling network, not only though segregated infrastructure but also by planning routes that passed through residential streets.
+2. **Low-Traffic Neighborhoods**:
+    The UK Department for Transport is allocating funding to local authorities to invest in Active Transport, partially through the creation of LTNs [@departmentgearchange2020].
+This includes closing off residential streets to motorized traffic
+3. **Existing Cycling Infrastructure**:
+    Utilizing existing cycling infrastructure makes economic sense, as small investments may lead to large connectivity gains as the disconnected cycling infrastructure gets joined together.
+
+The weighting profiles are therefore adjusted to favor less-stressful streets (based on information from Table \ref{table:osmroadtypes}), and roads with existing cycling infrastructure.
+This is also in line with the creation of LTNs, as residential streets are those where motorized traffic is most likely to be banned in the creation of LTNs.
+
 Road Segment Prioritization
 ===========================
-
-Analysis of Methodology
-=======================
 
@@
 ==========
+
+<!-- to fix indentation: https://github.com/crsh/papaja/issues/37#issuecomment-104185288 -->
+\setlength{\parindent}{-0.5in}
+\setlength{\leftskip}{0.5in}
+\setlength{\parskip}{8pt}

### 2020-10-23 Paper/journal-article/journal-article.tex
progress on paper

@@
 \IfFileExists{microtype.sty}{\usepackage{microtype}}{}
+\usepackage[margin=2cm]{geometry}
 \bibliographystyle{elsarticle-harv}
-\usepackage{color}
-\usepackage{fancyvrb}
-\newcommand{\VerbBar}{|}
-\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
-\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
-% Add ',fontsize=\small' for more characters per line
-\usepackage{framed}
-\definecolor{shadecolor}{RGB}{248,248,248}
-\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
-\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\BuiltInTok}[1]{#1}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
-\newcommand{\ExtensionTok}[1]{#1}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ImportTok}[1]{#1}
-\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\NormalTok}[1]{#1}
-\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
-\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\RegionMarkerTok}[1]{#1}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\usepackage{graphicx}
 \ifxetex
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{library}\NormalTok{(dplyr)}
-\end{Highlighting}
-\end{Shaded}
+Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod
+tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim
+veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea
+commodo consequat. Duis aute irure dolor in reprehenderit in voluptate
+velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint
+occaecat cupidatat non proident, sunt in culpa qui officia deserunt
+mollit anim id est laborum.Lorem ipsum dolor sit amet, consectetur
+adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore
+magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco
+laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor
+in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
+pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa
+qui officia deserunt mollit anim id est laborum.Lorem ipsum dolor sit
+amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut
+labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud
+exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
+Duis aute irure dolor in reprehenderit in voluptate velit esse cillum
+dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non
+proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
 
-\begin{verbatim}
-## 
-## Attaching package: 'dplyr'
-\end{verbatim}
-
-\begin{verbatim}
-## The following objects are masked from 'package:stats':
-## 
-##     filter, lag
-\end{verbatim}
-
-\begin{verbatim}
-## The following objects are masked from 'package:base':
-## 
-##     intersect, setdiff, setequal, union
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{library}\NormalTok{(readr)}
-\KeywordTok{library}\NormalTok{(tidyr)}
-\KeywordTok{library}\NormalTok{(ggplot2)}
-\KeywordTok{library}\NormalTok{(ggrepel)}
-
-\CommentTok{# this is a lookup table matching MSOAs to major towns and cities}
-\NormalTok{city_names <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{'../data-raw/Middle_Layer_Super_Output_Area__2011__to_Major_Towns_and_Cities__December_2015__Lookup_in_England_and_Wales.csv'}\NormalTok{) }
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-## Parsed with column specification:
-## cols(
-##   MSOA11CD = col_character(),
-##   MSOA11NM = col_character(),
-##   TCITY15CD = col_character(),
-##   TCITY15NM = col_character(),
-##   FID = col_double()
-## )
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{# change column name}
-\NormalTok{city_names <-}\StringTok{ }\NormalTok{city_names }\OperatorTok{%>%}\StringTok{ }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{city =}\NormalTok{ TCITY15NM)}
-
-\CommentTok{#unique cities}
-\KeywordTok{unique}\NormalTok{(city_names}\OperatorTok{$}\NormalTok{city)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-##   [1] "London"               NA                     "Bolton"              
-##   [4] "Stockport"            "Rochdale"             "Salford"             
-##   [7] "Manchester"           "Oldham"               "Bury"                
-##  [10] "Wigan"                "Rotherham"            "Liverpool"           
-##  [13] "Doncaster"            "Sheffield"            "St Helens"           
-##  [16] "Barnsley"             "Southport"            "Birkenhead"          
-##  [19] "Birmingham"           "Coventry"             "Dudley"              
-##  [22] "Newcastle upon Tyne"  "Sunderland"           "Sutton Coldfield"    
-##  [25] "South Shields"        "West Bromwich"        "Gateshead"           
-##  [28] "Leeds"                "Wolverhampton"        "Bradford"            
-##  [31] "Walsall"              "Huddersfield"         "Wakefield"           
-##  [34] "Solihull"             "Halifax"              "Leicester"           
-##  [37] "Nottingham"           "Hartlepool"           "Derby"               
-##  [40] "Blackpool"            "Grimsby"              "Warrington"          
-##  [43] "Scunthorpe"           "Kingston upon Hull"   "York"                
-##  [46] "Blackburn"            "Middlesbrough"        "Stockton-on-Tees"    
-##  [49] "Darlington"           "Telford"              "Stoke-on-Trent"      
-##  [52] "Swindon"              "Bristol"              "Peterborough"        
-##  [55] "Luton"                "Southend-on-Sea"      "Gillingham"          
-##  [58] "Plymouth"             "Weston-Super-Mare"    "Bath"                
-##  [61] "Bournemouth"          "Poole"                "Portsmouth"          
-##  [64] "Milton Keynes"        "Southampton"          "Bedford"             
-##  [67] "Reading"              "Chatham"              "Bracknell"           
-##  [70] "Slough"               "High Wycombe"         "Brighton and Hove"   
-##  [73] "Cambridge"            "Carlisle"             "Chester"             
-##  [76] "Chesterfield"         "Chelmsford"           "Basildon"            
-##  [79] "Eastbourne"           "Hastings"             "Exeter"              
-##  [82] "Colchester"           "Harlow"               "Cheltenham"          
-##  [85] "Gloucester"           "Hemel Hempstead"      "St Albans"           
-##  [88] "Stevenage"            "Basingstoke"          "Burnley"             
-##  [91] "Watford"              "Maidstone"            "Preston"             
-##  [94] "Norwich"              "Lincoln"              "Harrogate"           
-##  [97] "Northampton"          "Newcastle-under-Lyme" "Oxford"              
-## [100] "Mansfield"            "Shrewsbury"           "Burton upon Trent"   
-## [103] "Woking"               "Guildford"            "Ipswich"             
-## [106] "Crawley"              "Nuneaton"             "Worthing"            
-## [109] "Redditch"             "Worcester"            "Swansea"             
-## [112] "Cardiff"              "Newport"
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{# number of MSOAs in each city }
-\NormalTok{no_msoas <-}\StringTok{ }\NormalTok{city_names }\OperatorTok{%>%}\StringTok{ }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{group_by}\NormalTok{(city) }\OperatorTok{%>%}\StringTok{ }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{tally}\NormalTok{()}
-
-\CommentTok{##### CHOOSE YOU CITY }
-\NormalTok{chosen_city <-}\StringTok{ "Manchester"}
-\CommentTok{#create a directory to store data related to this city (does nothing if directory already exists)}
-\KeywordTok{dir.create}\NormalTok{(}\KeywordTok{paste0}\NormalTok{(}\StringTok{"../data/"}\NormalTok{, chosen_city), }\DataTypeTok{showWarnings =} \OtherTok{FALSE}\NormalTok{)}
-\CommentTok{# create sub-directory to save plots as well}
-\KeywordTok{dir.create}\NormalTok{(}\KeywordTok{paste0}\NormalTok{(}\StringTok{"../data/"}\NormalTok{, chosen_city,}\StringTok{"/Plots"}\NormalTok{), }\DataTypeTok{showWarnings =} \OtherTok{FALSE}\NormalTok{)}
-
-\CommentTok{##### CHOOSE YOU CITY }
-
-\CommentTok{# flow data from the 2011 census https://www.nomisweb.co.uk/census/2011/bulk/rOD1}
-\NormalTok{flows <-}\StringTok{ }\KeywordTok{read_csv}\NormalTok{(}\StringTok{'../data-raw/flow_data.csv'}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-## Parsed with column specification:
-## cols(
-##   `Area of residence` = col_character(),
-##   `Area of workplace` = col_character(),
-##   `All categories: Method of travel to work` = col_double(),
-##   `Work mainly at or from home` = col_double(),
-##   `Underground, metro, light rail, tram` = col_double(),
-##   Train = col_double(),
-##   `Bus, minibus or coach` = col_double(),
-##   Taxi = col_double(),
-##   `Motorcycle, scooter or moped` = col_double(),
-##   `Driving a car or van` = col_double(),
-##   `Passenger in a car or van` = col_double(),
-##   Bicycle = col_double(),
-##   `On foot` = col_double(),
-##   `Other method of travel to work` = col_double()
-## )
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{###############}
-\CommentTok{# MERGING NAMES WITH FLOW DATA (TO GET INTERNAL FLOWS IN ANY CITY)}
-\CommentTok{###############}
-
-\CommentTok{# add a column with the city name corresponding to each Residence MSOA}
-\NormalTok{flows <-}\StringTok{ }\NormalTok{flows }\OperatorTok{%>%}\StringTok{ }\KeywordTok{left_join}\NormalTok{(city_names[,}\KeywordTok{c}\NormalTok{(}\StringTok{"MSOA11CD"}\NormalTok{, }\StringTok{"city"}\NormalTok{)],}
-                             \DataTypeTok{by =} \KeywordTok{c}\NormalTok{(}\StringTok{"Area of residence"}\NormalTok{ =}\StringTok{ "MSOA11CD"}\NormalTok{)) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{city_origin =}\NormalTok{ city) }\CommentTok{# rename column so that we know it is referring to the 'Area of residence'}
-
-\CommentTok{# add a column with the city name corresponding to each Workplace MSOA}
-\NormalTok{flows <-}\StringTok{ }\NormalTok{flows }\OperatorTok{%>%}\StringTok{ }\KeywordTok{left_join}\NormalTok{(city_names[,}\KeywordTok{c}\NormalTok{(}\StringTok{"MSOA11CD"}\NormalTok{, }\StringTok{"city"}\NormalTok{)],}
-                             \DataTypeTok{by =} \KeywordTok{c}\NormalTok{(}\StringTok{"Area of workplace"}\NormalTok{ =}\StringTok{ "MSOA11CD"}\NormalTok{)) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{rename}\NormalTok{(}\DataTypeTok{city_dest =}\NormalTok{ city) }\CommentTok{# rename column so that we know it is referring to the 'Area of workplace'}
-
-\CommentTok{# get mode share for all cities - just for report}
-\NormalTok{cycle_mode_share <-}\StringTok{ }\NormalTok{flows }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\KeywordTok{filter}\NormalTok{(city_origin }\OperatorTok{==}\StringTok{ }\NormalTok{city_dest) }\OperatorTok{%>%}\StringTok{   }\CommentTok{# only internal flows}
-\StringTok{  }\KeywordTok{group_by}\NormalTok{(city_origin, city_dest) }\OperatorTok{%>%}\StringTok{   }
-\StringTok{  }\KeywordTok{summarize}\NormalTok{(}\DataTypeTok{mode_share =}\NormalTok{ (}\KeywordTok{sum}\NormalTok{(Bicycle) }\OperatorTok{/}\StringTok{ }\KeywordTok{sum}\NormalTok{(}\StringTok{`}\DataTypeTok{All categories: Method of travel to work}\StringTok{`}\NormalTok{)) }\OperatorTok{*}\DecValTok{100}\NormalTok{,}
-            \DataTypeTok{all =} \KeywordTok{sum}\NormalTok{(}\StringTok{`}\DataTypeTok{All categories: Method of travel to work}\StringTok{`}\NormalTok{)) }\OperatorTok{%>%}
-\StringTok{  }\KeywordTok{arrange}\NormalTok{(mode_share)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-## `summarise()` regrouping output by 'city_origin' (override with `.groups` argument)
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{# plot }
-\NormalTok{cycle_mode_share }\OperatorTok{%>%}\StringTok{ }\CommentTok{#filter(city_origin != 'London') %>%}
-\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(all, mode_share)) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{geom_point}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{alpha =}\NormalTok{ mode_share), }\DataTypeTok{show.legend =} \OtherTok{FALSE}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\CommentTok{# some filtering labels for aesthetic purposes. Add some cities explicitly as they are in my case study}
-\StringTok{  }\KeywordTok{geom_label_repel}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{label =} \KeywordTok{ifelse}\NormalTok{(mode_share}\OperatorTok{>}\DecValTok{12} \OperatorTok{|}\StringTok{ }\NormalTok{all}\OperatorTok{>}\StringTok{ }\NormalTok{(}\FloatTok{2.3}\OperatorTok{*}\KeywordTok{mean}\NormalTok{(all)) }\OperatorTok{|}\StringTok{ }\NormalTok{city_origin }\OperatorTok{%in%}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{'Manchester'}\NormalTok{, }\StringTok{'Nottingham'}\NormalTok{, }\StringTok{'Birmingham'}\NormalTok{) , }\KeywordTok{as.character}\NormalTok{(city_origin),}\StringTok{''}\NormalTok{)), }
-                       \DataTypeTok{size =}\FloatTok{2.5}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{x=}\StringTok{"Total Number of Commuters"}\NormalTok{, }\DataTypeTok{y =} \StringTok{"Cycling Mode Share (%)"}\NormalTok{) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{scale_x_continuous}\NormalTok{(}\DataTypeTok{trans=}\StringTok{'log10'}\NormalTok{, }\DataTypeTok{labels =}\NormalTok{ scales}\OperatorTok{::}\NormalTok{comma) }\OperatorTok{+}
-\StringTok{  }\KeywordTok{theme_minimal}\NormalTok{()}
-\end{Highlighting}
-\end{Shaded}
-
-\includegraphics{journal-article_files/figure-latex/unnamed-chunk-1-1.pdf}
+\hypertarget{background}{%
+\section{Background}\label{background}}
 
 \hypertarget{what-affects-the-decision-to-cycle}{%
-\section{What Affects the Decision To
+\subsection{What Affects the Decision To
 Cycle}\label{what-affects-the-decision-to-cycle}}
 
+Segregated cycling
+infrastructure\footnote{Segregated cycling infrastructure refers to road space that is allocated to cyclists only, with physical separation to protect cyclists from other modes of transport.}
+has been shown to increase cycling uptake (Aldred et al., 2019; Goodman
+et al., 2014; Marqués et al., 2015), with the separation from motorized
+vehicles being key (Winters et al., 2011). Revealed preference of
+cyclists shows that they are willing to deviate from the most efficient
+routes in order to commute on safer roads (Crane et al., 2017). However,
+such deviations are only considered if they do not considerably increase
+route circuitry; behaviour studies have found that the probability of
+choosing a route decreases in proportion to its length relative to the
+shortest route (Broach et al., 2011; Winters et al., 2010). Another
+defining feature for cycling infrastructure is how well connected it is.
+Cyclists prefer cohesive infrastructure, particularly when cycling on
+arterial roads with high levels of motorized traffic (Stinson and Bhat,
+2003), and the lack of well-connected cycling infrastructure is one of
+the main obstacles to increasing cycling uptake (Caulfield et al.,
+2012). While direct and cohesive cycling networks have been shown to
+positively impact cycling rates,
+density\footnote{making an area's bicycle network denser means adding more cycling routes in the area and thereby giving cyclists more route options}
+of the cycling network is also vital (Schoner and Levinson, 2014).
+
 \hypertarget{planning-cycling-networks}{%
-\section{Planning Cycling Networks}\label{planning-cycling-networks}}
+\subsection{Planning Cycling Networks}\label{planning-cycling-networks}}
 
-\hypertarget{potential-cycling-demand}{%
-\section{Potential Cycling Demand}\label{potential-cycling-demand}}
+\emph{Optimization} techniques have been used to propose improvements to
+cycling networks. Mesbah et al. (2012) propose a bi-level formulation to
+optimize allocation of cycling lanes to the network without exceeding a
+set budget. The upper level is the proposed interventions and the lower
+level is the route choices made by users in reaction to changes in the
+network. The problem accounts for the effect of cycling lanes on car
+traffic, and attempts to maximize utilization of said lanes with minimal
+impact on car travel times. To improve cohesion of the suggested
+network, a constraint is added so that each link with a bike lane should
+be connected to at least one destination. Car usage is not considered by
+Mauttone et al. (2017), who develop an optimization framework that aims
+to minimize the total user cost of cycling on the network. The aggregate
+flow on links is obtained by using shortest paths to route existing
+cycling demand onto the road network, and the solution is a proposed set
+of links where cycling infrastructure should be added in order to
+minimize the overall travel cost of cyclists across the network. The
+cost of traversing a link is given as a function of its length and
+whether or not it has cycling infrastructure, and a discontinuity
+penalty is also added to prioritize connected road segments. The problem
+has also been solved by attempting to find the minimum cost of improving
+roadway links to meet a desired level of service (LOS) (Duthie and
+Unnikrishnan, 2014). In this formulation, all OD pairs need to be
+connected by roads that meet the desired LOS, and a directness
+constraint is added so that paths between OD pairs do not exceed a
+certain multiple of the shortest path.
+
+These problem formulations do not explicitly solve for continuity, which
+is dealt with using a either (a) a constraint specifying that each link
+with a bike lane should be connected to at least one destination (Mesbah
+et al., 2012), (b) a constraint on deviation from shortest paths (Duthie
+and Unnikrishnan, 2014), or (c) a discontinuity penalty (Mauttone et
+al., 2017). To solve for continuity, the graph-theoretic concept of
+\emph{connected components}, has been used. Natera et al. (2019) study
+the existing cycling network in terms of its disconnected components and
+introduce two different algorithms to connect these components by their
+most critical
+links\footnote{\textit{link} refers to a road segment throughout this research}
+and, in doing so, measure the size of the growth of the largest
+connected component as a function of the kilometers of network added.
+They observe that small investments at strategic points have a large
+impact on connectivity in most cases. The concept of connected
+components is also at the core of the methodology proposed by Olmos et
+al. (2020). After routing the cycling demand onto the network links,
+they use percolation theory to filter out the links based on the
+aggregate
+flow\footnote{\textit{flow} is used throughout this research to refer to the cycling demand when it is routed onto the road network.
+The flow on any road segment is the cumulative demand on it, resulting from cyclists commuting between various OD pairs}
+passing through them. They vary the flow threshold for filtering to
+identify the minimum flow at which the whole city is connected by a
+giant component. The results show a cycling network that connects the
+entire city, and subtracting links intersecting with current cycling
+infrastructure identifies links proposed for intervention.
+
+\hypertarget{data-and-geographical-scale-of-analysis}{%
+\section{Data and Geographical Scale of
+Analysis}\label{data-and-geographical-scale-of-analysis}}
+
+The analysis is heavily dependant on Origin-Destination census data
+(commuter data). Commuter data in the UK is publicly available at the
+Middle layer Super Output Area (MSOA) level; the average MSOA has a
+population of 8209 (ONS, 2018). Iacono et al. (2010) note that such
+large travel zones are not ideal for understanding route choice
+behaviour of cyclists and pedestrians. They also give rise to an
+`ecological fallacy' whereby average characteristics are assumed to
+apply to all residents of the aggregated geographical area. Given that
+more granular data is not publicly available, the study uses MSOA-level
+commuter data. The methodology is however applicable to more granular
+commuter data should it become available.
+
+\hypertarget{calculating-potential-cycling-demand}{%
+\section{Calculating Potential Cycling
+Demand}\label{calculating-potential-cycling-demand}}
+
+The Propensity to Cycle Tool (PCT) (Lovelace et al., 2017) is used to
+estimate the proportion of cyclists (\(\boldsymbol{C_{p}}\)) for each
+MSOA pair should the government achieve its target of doubling cycling
+by 2025. The PCT uses the following logistic regression model to
+calculate \(\boldsymbol{C_{p}}\):
 
@@
 
+\noindent where \(\boldsymbol{d}\) and \(\boldsymbol{s}\) are the
+distance and slope respectively for the OD pair. The authors use square
+and square-root distance terms ``to capture the non-linear impact of
+distance on the likelihood of cycling", and interaction terms to capture
+the combined effect of slope and distance (Lovelace et al., 2017).
+
+The potential demand calculations show that the current and potential
+number of cyclists both follow a bell-shaped distribution, with the
+number of trips peaking around the 3-5km commuting distance and then
+going back down for longer distances (See Figure
+\ref{fig:pot_dem_histograms}).
+
+\begin{figure}[h!]
+    \centering % puts the image in the horizontal centre of the page
+    \includegraphics[width=.6\textwidth]{../../data/Manchester/Plots/desire_facet_cycling.png} 
+    \caption{Current and Potential Cycling Demand} 
+    \label{fig:desire_facet_cycling} 
+\end{figure}
+
 \hypertarget{routing}{%
 \section{Routing}\label{routing}}
+
+The next step is to route the potential cycling demand
+(\(\boldsymbol{C_{p}}\)) between all OD pairs onto the road network.
+
+To conduct routing, the following is considered:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  \textbf{Cyclist Preference}: Work done by Dill and McNeil (2013) on
+  examining cyclist typologies determined that around 60\% of Portland
+  residents fit under the \emph{interested but concerned} category.
+  These were people that enjoyed cycling but avoided it due safety
+  concerns. The key to encouraging this group was to create a low-stress
+  cycling network, not only though segregated infrastructure but also by
+  planning routes that passed through residential streets.
+\item
+  \textbf{Low-Traffic Neighborhoods}: The UK Department for Transport is
+  allocating funding to local authorities to invest in Active Transport,
+  partially through the creation of LTNs (DfT, 2020). This includes
+  closing off residential streets to motorized traffic
+\item
+  \textbf{Existing Cycling Infrastructure}: Utilizing existing cycling
+  infrastructure makes economic sense, as small investments may lead to
+  large connectivity gains as the disconnected cycling infrastructure
+  gets joined together.
+\end{enumerate}
+
+The weighting profiles are therefore adjusted to favor less-stressful
+streets (based on information from Table \ref{table:osmroadtypes}), and
+roads with existing cycling infrastructure. This is also in line with
+the creation of LTNs, as residential streets are those where motorized
+traffic is most likely to be banned in the creation of LTNs.
 
@@
 Prioritization}\label{road-segment-prioritization}}
-
-\hypertarget{analysis-of-methodology}{%
-\section{Analysis of Methodology}\label{analysis-of-methodology}}
 
@@
 
+\setlength{\parindent}{-0.5in}
+\setlength{\leftskip}{0.5in}
+\setlength{\parskip}{8pt}
+
+\hypertarget{refs}{}
+\leavevmode\hypertarget{ref-aldred2019impacts}{}%
+Aldred, R., Croft, J., Goodman, A., 2019. Impacts of an active travel
+intervention with a cycling focus in a suburban context: One-year
+findings from an evaluation of london's in-progress mini-hollands
+programme. Transportation research part A: policy and practice 123,
+147--169.
+
+\leavevmode\hypertarget{ref-broach2011bicycle}{}%
+Broach, J., Gliebe, J., Dill, J., 2011. Bicycle route choice model
+developed using revealed preference gps data, in: 90th Annual Meeting of
+the Transportation Research Board, Washington, Dc.
+
+\leavevmode\hypertarget{ref-caulfield2012determining}{}%
+Caulfield, B., Brick, E., McCarthy, O.T., 2012. Determining bicycle
+infrastructure preferences--a case study of dublin. Transportation
+research part D: transport and environment 17, 413--417.
+
+\leavevmode\hypertarget{ref-crane2017longitudinal}{}%
+Crane, M., Rissel, C., Standen, C., Ellison, A., Ellison, R., Wen, L.M.,
+Greaves, S., 2017. Longitudinal evaluation of travel and health outcomes
+in relation to new bicycle infrastructure, sydney, australia. Journal of
+Transport \& Health 6, 386--395.
+
+\leavevmode\hypertarget{ref-departmentgearchange2020}{}%
+DfT, 2020. Gear change: A bold vision for cycling and walking.
+
+\leavevmode\hypertarget{ref-dill2013four}{}%
+Dill, J., McNeil, N., 2013. Four types of cyclists? Examination of
+typology for better understanding of bicycling behavior and potential.
+Transportation Research Record 2387, 129--138.
+
+\leavevmode\hypertarget{ref-duthie2014optimization}{}%
+Duthie, J., Unnikrishnan, A., 2014. Optimization framework for bicycle
+network design. Journal of Transportation engineering 140, 04014028.
+
+\leavevmode\hypertarget{ref-goodman2014new}{}%
+Goodman, A., Sahlqvist, S., Ogilvie, D., Consortium, 2014. New walking
+and cycling routes and increased physical activity: One-and 2-year
+findings from the uk iConnect study. American journal of public health
+104, e38--e46.
+
+\leavevmode\hypertarget{ref-iacono2010measuring}{}%
+Iacono, M., Krizek, K.J., El-Geneidy, A., 2010. Measuring non-motorized
+accessibility: Issues, alternatives, and execution. Journal of Transport
+Geography 18, 133--140.
+
+\leavevmode\hypertarget{ref-lovelace2017propensity}{}%
+Lovelace, R., Goodman, A., Aldred, R., Berkoff, N., Abbas, A., Woodcock,
+J., 2017. The propensity to cycle tool: An open source online system for
+sustainable transport planning. Journal of transport and land use 10,
+505--528.
+
+\leavevmode\hypertarget{ref-marques2015infrastructure}{}%
+Marqués, R., Hernández-Herrador, V., Calvo-Salazar, M., García-Cebrián,
+J., 2015. How infrastructure can promote cycling in cities: Lessons from
+seville. Research in Transportation Economics 53, 31--44.
+
+\leavevmode\hypertarget{ref-mauttone2017bicycle}{}%
+Mauttone, A., Mercadante, G., Rabaza, M., Toledo, F., 2017. Bicycle
+network design: Model and solution algorithm. Transportation research
+procedia 27, 969--976.
+
+\leavevmode\hypertarget{ref-mesbah2012bilevel}{}%
+Mesbah, M., Thompson, R., Moridpour, S., 2012. Bilevel optimization
+approach to design of network of bike lanes. Transportation research
+record 2284, 21--28.
+
+\leavevmode\hypertarget{ref-natera2019data}{}%
+Natera, L., Battiston, F., Iñiguez, G., Szell, M., 2019. Data-driven
+strategies for optimal bicycle network growth. arXiv preprint
+arXiv:1907.07080.
+
+\leavevmode\hypertarget{ref-olmos2020data}{}%
+Olmos, L.E., Tadeo, M.S., Vlachogiannis, D., Alhasoun, F., Alegre, X.E.,
+Ochoa, C., Targa, F., González, M.C., 2020. A data science framework for
+planning the growth of bicycle infrastructures. Transportation Research
+Part C: Emerging Technologies 115, 102640.
+
+\leavevmode\hypertarget{ref-ofn2018population}{}%
+ONS, 2018. Population estimates for the uk, england and wales, scotland
+and northern ireland: Mid-2017. Hampshire: Office for National
+Statistics.
+
+\leavevmode\hypertarget{ref-schoner2014missing}{}%
+Schoner, J.E., Levinson, D.M., 2014. The missing link: Bicycle
+infrastructure networks and ridership in 74 us cities. Transportation
+41, 1187--1204.
+
+\leavevmode\hypertarget{ref-stinson2003commuter}{}%
+Stinson, M.A., Bhat, C.R., 2003. Commuter bicyclist route choice:
+Analysis using a stated preference survey. Transportation Research
+Record 1828, 107--115.
+
+\leavevmode\hypertarget{ref-winters2011motivators}{}%
+Winters, M., Davidson, G., Kao, D., Teschke, K., 2011. Motivators and
+deterrents of bicycling: Comparing influences on decisions to ride.
+Transportation 38, 153--168.
+
+\leavevmode\hypertarget{ref-winters2010far}{}%
+Winters, M., Teschke, K., Grant, M., Setton, E.M., Brauer, M., 2010. How
+far out of the way will we travel? Built environment influences on route
+selection for bicycle and car travel. Transportation Research Record
+2190, 1--10.
+
 \end{document}

### 2020-10-23 R/_3.0_potential_demand.R

@@ -7,128 +7,73 @@ flow <- readr::read_csv(paste0("../data/",chosen_city,"/flows_dist_elev_for_pote
 # replace NA values with the mean slope
 flow$slope[is.na(flow$slope)] <- mean(na.omit(flow$slope))
 
+
+# ---------------------------------------------------------------------------------------------------- #
+
 ###############
-# OPTION 3: my own distance decay function
+# OPTION 1: Use a distance decay function from pct package https://itsleeds.github.io/pct/reference/uptake_pct_godutch.html
 ###############
+
+# create a copy of the df 
 uptake_decay <- flow
 
 # What to do with dist= NA
 #opt 1. Remove these rows
 uptake_decay <- uptake_decay %>% filter(!is.na(dist))
 
-#opt  2.replace NA values by column mean 
-#uptake_decay$dist[is.na(uptake_decay$dist)] <- max(uptake_decay$dist[!is.na(uptake_decay$dist)])
-
 # get % of cyclists
 uptake_decay$perc_cycle <- uptake_decay$Bicycle / uptake_decay$`All categories: Method of travel to work`
-# use this df for glm, as intra flows are all assigned distance 0 and so affect the results
-uptake_no_intra <- uptake_decay %>% dplyr::filter(`Area of residence` != `Area of workplace`)
 
+# get % increase in cycling as distance decay 
+uptake_decay$prob_cycle = pct::uptake_pct_govtarget_2020(distance = uptake_decay$dist, gradient = uptake_decay$slope)
 
-##### GLM TO CALCULATE PROBABILITY OF CYCLING BASED ON PHYSICAL GEOGRAPHY (Distance, Slope) #####
-
-# sqrt to get bell shape!  https://itsleeds.github.io/pct/reference/uptake_pct_govtarget.html
-# Should I add all_travel as a predictor???
-glm1 <- glm(perc_cycle ~ dist + sqrt(dist) + slope, data = uptake_no_intra, family = "quasibinomial")
-
-# If coefficient (logit) is positive, the effect of this predictor on cycling is positive and vice versa
-#coeff <- coef(glm1) %>% as.data.frame() 
-summary(glm1)
-# predict cycling probability on all OD pairs
-uptake_decay$prob_cycle <- predict(glm1, data.frame(dist = uptake_decay$dist, slope = uptake_decay$slope), type = "response")
+ggplot(uptake_decay) +
+  geom_point(aes(dist, prob_cycle)) + 
+  labs( x="Commuting Distance (km)", y = "Uptake (%)")
 
[…]

### 2020-10-23 R/_3.2_plot_od_comparison.R

@@ -127,7 +127,7 @@ city_geom <- city_geom %>% dplyr::filter(msoa11cd %in% city_msoas$MSOA11CD)
 
 # get straight line geometry of all OD pairs. Added as a geometry column
 uptake_decay <- stplanr::od2line(uptake_decay, city_centroids)
-uptake_decay <- lwgeom::st_make_valid(uptake_decay)
+uptake_decay <- sf::st_make_valid(uptake_decay)
 
 # keep only od pairs <10km
 uptake_plot <- uptake_decay %>% filter(dist <= 10)

### 2020-10-23 R/_3.3_plot_desire_lines_current_vs_potential.R

@@ -41,7 +41,7 @@ city_od$potential_demand[is.na(city_od$potential_demand)] <- 0
 
 # get straight line geometry of all OD pairs in city_od
 desire_lines <- stplanr::od2line(city_od, city_centroids)
-desire_lines <- lwgeom::st_make_valid(desire_lines)
+desire_lines <- sf::st_make_valid(desire_lines)
 
 # filter out all OD pairs with flow less than this:
 flow_threshold_all <- 50

### 2020-10-23 Paper/journal-article/journal-article.Rmd
updates to paper

@@
 
-  It consists of two paragraphs.
-
 journal: "An awesome journal"
@@
 ---
+
+<!-- Keywords -->
+<!-- ======== -->
+
+<!-- Highlights -->
+<!-- ======== -->
 
@@
 
+\vspace{0.3cm}
+
 These problem formulations do not explicitly solve for continuity, which is dealt with using a either (a) a constraint specifying that each link with a bike lane should be connected to at least one destination [@mesbah2012bilevel], (b) a constraint on deviation from shortest paths [@duthie2014optimization], or (c) a discontinuity penalty [@mauttone2017bicycle].
@@
 The results show a cycling network that connects the entire city, and subtracting links intersecting with current cycling infrastructure identifies links proposed for intervention.
+
+\vspace{0.3cm}
+
+The problem formulations outlined above look at the network as a whole when attempting to improve it.
+An alternative approach is to identify the different sub-networks that exist within the larger network, and work on improving each separately.
+Trip patterns in a city are not uniformly distributed geographically, and *community finding* methods have been used to partition study areas into localized areas that experience a disproportionate number of trips within them.
+@akbarzadeh2018designing use a modularity maximization approach [@blondel2008fast] on taxi trip data to identify 7 different communities in the city of Isfahan, Iran.
+An optimization problem is then formulated to connect nodes within each community with cycling infrastructure.
+The emphasis is on connectivity within the communities, not between them.
+@bao2017planning adopt a similar methodology, but use hierarchical clustering to specify the desired number of clusters.
+They use a greedy network expansion algorithm, where the link with the highest benefit/cost ratio in each cluster is selected, and the network is grown by adding neighboring links to the solution until a budget limit is met.
+The benefit is the flow on the link, and each link is assigned a cost based on current road conditions.
+
+## Underlying Ethical Principles
+
+The methodologies in Section **How do I cite a section??** are underpinned by different ethical principles, even though these principles are not explicitly acknowledged by the authors.
+This is important since different ethical principles constitute different problem formulations and targets.
+Broadly speaking, transport appraisal can be based on either utilitarian or egalitarian principles.
+The former seeks to maximize the overall benefit, while the latter is concerned with a fair distribution of benefits [@jafino2020transport].
+@nahmias2017integrating criticize the utilitarian approach that has been historically popular in the evaluation of transport investments, explaining how the maximization of overall benefit fails to account for the distribution of that benefit among communities or individuals.
+@lucas2016method explain how transport studies have traditionally looked at the bigger picture without studying the distribution of investments on the different parts of the study area, and go on to propose an egalitarian approach that ensures the dis-aggregation of transport policy benefits across the study area.
+@pereira2017distributive also emphasize the need for a more egalitarian approach to transport planning.
+They highlight accessibility as a cornerstone of distributive justice, and contend that policies should aim to distribute investments in a way that minimizes spatial variations in accessibility.
+This research attempts to provide a methodology that is grounded in egalitarian principles.
+**Write some more here**
 
@@
 
+Overarching Policies
+====================
+
+While segregated, connected, and direct cycling infrastructure is key to achieving high levels of cycling, research has shown that it cannot exist in a vacuum.
+@wardman2007factors developed a mode choice model for the UK and their results showed that improved cycling infrastructure on its own only had modest impacts on mode shift, and even the unlikely scenario of all urban routes being serviced by segregated bike lanes was forecast to increase cycling mode share by only 3\%.
+However, cities that invest in more comprehensive cycling projects show a more significant increase in the number of cyclists as well as the cycling mode share [@pucher2010infrastructure].
+These cities do not just focus on infrastructure, but on general policies as well as restricting car use.
+Evaluation of policies in Denmark and Germany and the Netherlands has shown that their high cycling mode share is down to a broader set of policies that also include traffic calming, cycling rights of way, bike parking, integration with the public transport network, and making driving cars both expensive and inconvenient [@pucher2008making].
+While these policies are outside the scope of this research, it is important to recognize their key role in bringing about an increase in levels of cycling.
+
 Conclusions

### 2020-10-23 Paper/journal-article/journal-article.tex
updates to paper

@@
   This is the abstract.
-
-  It consists of two paragraphs.
   \end{abstract}
@@
 
+\vspace{0.3cm}
+
 These problem formulations do not explicitly solve for continuity, which
@@
 infrastructure identifies links proposed for intervention.
+
+\vspace{0.3cm}
+
+The problem formulations outlined above look at the network as a whole
+when attempting to improve it. An alternative approach is to identify
+the different sub-networks that exist within the larger network, and
+work on improving each separately. Trip patterns in a city are not
+uniformly distributed geographically, and \emph{community finding}
+methods have been used to partition study areas into localized areas
+that experience a disproportionate number of trips within them.
+Akbarzadeh et al. (2018) use a modularity maximization approach (Blondel
+et al., 2008) on taxi trip data to identify 7 different communities in
+the city of Isfahan, Iran. An optimization problem is then formulated to
+connect nodes within each community with cycling infrastructure. The
+emphasis is on connectivity within the communities, not between them.
+Bao et al. (2017) adopt a similar methodology, but use hierarchical
+clustering to specify the desired number of clusters. They use a greedy
+network expansion algorithm, where the link with the highest
+benefit/cost ratio in each cluster is selected, and the network is grown
+by adding neighboring links to the solution until a budget limit is met.
+The benefit is the flow on the link, and each link is assigned a cost
+based on current road conditions.
+
+\hypertarget{underlying-ethical-principles}{%
+\subsection{Underlying Ethical
+Principles}\label{underlying-ethical-principles}}
+
+The methodologies in Section \textbf{How do I cite a section??} are
+underpinned by different ethical principles, even though these
+principles are not explicitly acknowledged by the authors. This is
+important since different ethical principles constitute different
+problem formulations and targets. Broadly speaking, transport appraisal
+can be based on either utilitarian or egalitarian principles. The former
+seeks to maximize the overall benefit, while the latter is concerned
+with a fair distribution of benefits (Jafino et al., 2020).
+Nahmias-Biran et al. (2017) criticize the utilitarian approach that has
+been historically popular in the evaluation of transport investments,
+explaining how the maximization of overall benefit fails to account for
+the distribution of that benefit among communities or individuals. Lucas
+et al. (2016) explain how transport studies have traditionally looked at
+the bigger picture without studying the distribution of investments on
+the different parts of the study area, and go on to propose an
+egalitarian approach that ensures the dis-aggregation of transport
+policy benefits across the study area. Pereira et al. (2017) also
+emphasize the need for a more egalitarian approach to transport
+planning. They highlight accessibility as a cornerstone of distributive
+justice, and contend that policies should aim to distribute investments
+in a way that minimizes spatial variations in accessibility. This
+research attempts to provide a methodology that is grounded in
+egalitarian principles. \textbf{Write some more here}
 
@@
 
+\hypertarget{overarching-policies}{%
+\section{Overarching Policies}\label{overarching-policies}}
+
+While segregated, connected, and direct cycling infrastructure is key to
+achieving high levels of cycling, research has shown that it cannot
+exist in a vacuum. Wardman et al. (2007) developed a mode choice model
+for the UK and their results showed that improved cycling infrastructure
+on its own only had modest impacts on mode shift, and even the unlikely
+scenario of all urban routes being serviced by segregated bike lanes was
+forecast to increase cycling mode share by only 3\%. However, cities
+that invest in more comprehensive cycling projects show a more
+significant increase in the number of cyclists as well as the cycling
+mode share (Pucher et al., 2010). These cities do not just focus on
+infrastructure, but on general policies as well as restricting car use.
+Evaluation of policies in Denmark and Germany and the Netherlands has
+shown that their high cycling mode share is down to a broader set of
+policies that also include traffic calming, cycling rights of way, bike
+parking, integration with the public transport network, and making
+driving cars both expensive and inconvenient (Pucher and Buehler, 2008).
+While these policies are outside the scope of this research, it is
+important to recognize their key role in bringing about an increase in
+levels of cycling.
+
 \hypertarget{conclusions}{%
@@
 \hypertarget{refs}{}
+\leavevmode\hypertarget{ref-akbarzadeh2018designing}{}%
+Akbarzadeh, M., Mohri, S.S., Yazdian, E., 2018. Designing bike networks
+using the concept of network clusters. Applied network science 3, 12.
+
 \leavevmode\hypertarget{ref-aldred2019impacts}{}%
@@
 147--169.
+
+\leavevmode\hypertarget{ref-bao2017planning}{}%
+Bao, J., He, T., Ruan, S., Li, Y., Zheng, Y., 2017. Planning bike lanes
+based on sharing-bikes' trajectories, in: Proceedings of the 23rd Acm
+Sigkdd International Conference on Knowledge Discovery and Data Mining.
+pp. 1377--1386.
+
+\leavevmode\hypertarget{ref-blondel2008fast}{}%
+Blondel, V.D., Guillaume, J.-L., Lambiotte, R., Lefebvre, E., 2008. Fast
+unfolding of communities in large networks. Journal of statistical
+mechanics: theory and experiment 2008, P10008.
 
@@
 
+\leavevmode\hypertarget{ref-jafino2020transport}{}%
+Jafino, B.A., Kwakkel, J., Verbraeck, A., 2020. Transport network
+criticality metrics: A comparative analysis and a guideline for
+selection. Transport Reviews 40, 241--264.
+
 \leavevmode\hypertarget{ref-lovelace2017propensity}{}%
@@
 505--528.
+
+\leavevmode\hypertarget{ref-lucas2016method}{}%
+Lucas, K., Van Wee, B., Maat, K., 2016. A method to evaluate equitable
+accessibility: Combining ethical theories and accessibility-based
+approaches. Transportation 43, 473--490.
 
@@
 record 2284, 21--28.
+
+\leavevmode\hypertarget{ref-nahmias2017integrating}{}%
+Nahmias-Biran, B.-h., Martens, K., Shiftan, Y., 2017. Integrating equity
+in transportation project assessment: A philosophical exploration and
+its practical implications. Transport reviews 37, 192--210.
 
@@
 
+\leavevmode\hypertarget{ref-pereira2017distributive}{}%
+Pereira, R.H., Schwanen, T., Banister, D., 2017. Distributive justice
+and equity in transportation. Transport reviews 37, 170--191.
+
+\leavevmode\hypertarget{ref-pucher2008making}{}%
+Pucher, J., Buehler, R., 2008. Making cycling irresistible: Lessons from
+the netherlands, denmark and germany. Transport reviews 28, 495--528.
+
+\leavevmode\hypertarget{ref-pucher2010infrastructure}{}%
+Pucher, J., Dill, J., Handy, S., 2010. Infrastructure, programs, and
+policies to increase bicycling: An international review. Preventive
+medicine 50, S106--S125.
+
 \leavevmode\hypertarget{ref-schoner2014missing}{}%
@@
 Record 1828, 107--115.
+
+\leavevmode\hypertarget{ref-wardman2007factors}{}%
+Wardman, M., Tight, M., Page, M., 2007. Factors influencing the
+propensity to cycle to work. Transportation Research Part A: Policy and
+Practice 41, 339--350.
 

### 2020-10-30 Paper/journal-article/journal-article.Rmd
Update journal-article.Rmd
update paper again

@@
 csl: elsevier-harvard.csl
-output: rticles::elsevier_article
+#output: rticles::elsevier_article
+output:
+  pdf_document:
+    extra_dependencies: ["subcaption", "longtable"]
 ---
@@
 
+\begin{figure} [h!]
+\centering
+\captionsetup{font=footnotesize,labelfont=footnotesize} % size of captions
+\begin{subfigure}{.45\textwidth}
+  \centering
+  \includegraphics[width=1\linewidth]{../../data/Manchester/Plots/Growth_Results/growth_existing_infra_satisfied_km_all_flow_column.png}
+  \caption{Alg 1 (Utilitarian)}
+  \label{fig:growth_utilitarian_satisfied_all}
+\end{subfigure}
+\begin{subfigure}{.45\textwidth}
+  \centering
+  \includegraphics[width=1\linewidth]{../../data/Manchester/Plots/Growth_Results/growth_community_4_satisfied_km_all_flow_column.png}
+  \caption{Alg 2 (Egalitarian)}
+  \label{fig:growth_egalitarian_satisfied_all}
+\end{subfigure}
+\caption{Comparing Overall Person-Km Satisfied (Manchester)}
+\label{fig:growth_existing_infra_satisfied}
+\end{figure}
+
 Overarching Policies

### 2020-10-30 Paper/journal-article/journal-article.tex
update paper again

@@
-\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
-%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
-\usepackage[hyphens]{url}
-
-  \journal{An awesome journal} % Sets Journal name
-
-\usepackage{lineno} % add
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-
-\usepackage{graphicx}
-\usepackage{booktabs} % book-quality tables
-%%%%%%%%%%%%%%%% end my additions to header
-
-\usepackage[T1]{fontenc}
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+%
+\documentclass[
+]{article}
 \usepackage{lmodern}
@@
 \usepackage{ifxetex,ifluatex}
-\usepackage{fixltx2e} % provides \textsubscript
-% use upquote if available, for straight quotes in verbatim environments
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+% Use upquote if available, for straight quotes in verbatim environments
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[utf8]{inputenc}
-\else % if luatex or xelatex
-  \usepackage{fontspec}
-  \ifxetex
-    \usepackage{xltxtra,xunicode}
-  \fi
-  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
-  \newcommand{\euro}{€}
-\fi
-% use microtype if available
-\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
+  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\hypersetup{
+  pdftitle={Short Paper},
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
+\urlstyle{same} % disable monospaced font for URLs
 \usepackage[margin=2cm]{geometry}
-\bibliographystyle{elsarticle-harv}
-\ifxetex
-  \usepackage[setpagesize=false, % page size defined by xetex
-              unicode=false, % unicode breaks when used with xetex
-              xetex]{hyperref}
-\else
-  \usepackage[unicode=true]{hyperref}
-\fi
-\hypersetup{breaklinks=true,
-            bookmarks=true,
-            pdfauthor={},
-            pdftitle={Short Paper},
-            colorlinks=false,
-            urlcolor=blue,
-            linkcolor=magenta,
-            pdfborder={0 0 0}}
-\urlstyle{same}  % don't use monospace font for urls
+\usepackage{graphicx,grffile}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
+\setlength{\emergencystretch}{3em} % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{5}
+\usepackage{subcaption}
+\usepackage{longtable}
 
-\setcounter{secnumdepth}{5}
-% Pandoc toggle for numbering sections (defaults to be off)
-
-% Pandoc header
+\title{Short Paper}
+\author{true \and true \and true \and true}
+\date{2020-10-30}
 
 \begin{document}
-\begin{frontmatter}
-
-  \title{Short Paper}
-    \author[Some Institute of Technology]{Alice Anonymous\corref{1}}
-   \ead{alice@example.com} 
-    \author[Another University]{Bob Security}
-   \ead{bob@example.com} 
-    \author[Another University]{Cat Memes\corref{2}}
-   \ead{cat@example.com} 
-    \author[Some Institute of Technology]{Derek Zoolander\corref{2}}
-   \ead{derek@example.com} 
-      \address[Some Institute of Technology]{Department, Street, City, State, Zip}
-    \address[Another University]{Department, Street, City, State, Zip}
-      \cortext[1]{Corresponding Author}
-    \cortext[2]{Equal contribution}
-
+\maketitle
 \begin{abstract}
@@
 \end{abstract}
-
- \end{frontmatter}
 
@@
 
+\begin{figure} [h!]
+\centering
+\captionsetup{font=footnotesize,labelfont=footnotesize} % size of captions
+\begin{subfigure}{.45\textwidth}
+  \centering
+  \includegraphics[width=1\linewidth]{../../data/Manchester/Plots/Growth_Results/growth_existing_infra_satisfied_km_all_flow_column.png}
+  \caption{Alg 1 (Utilitarian)}
+  \label{fig:growth_utilitarian_satisfied_all}
+\end{subfigure}
+\begin{subfigure}{.45\textwidth}
+  \centering
+  \includegraphics[width=1\linewidth]{../../data/Manchester/Plots/Growth_Results/growth_community_4_satisfied_km_all_flow_column.png}
+  \caption{Alg 2 (Egalitarian)}
+  \label{fig:growth_egalitarian_satisfied_all}
+\end{subfigure}
+\caption{Comparing Overall Person-Km Satisfied (Manchester)}
+\label{fig:growth_existing_infra_satisfied}
+\end{figure}
+
 \hypertarget{overarching-policies}{%

### 2020-10-30 Paper/journal-article/journal-article.Rmd
Add paper to readme

--- 2020-10-23 Paper/journal-article/journal-article.Rmd
@@
 csl: elsevier-harvard.csl
-output: rticles::elsevier_article
+#output: rticles::elsevier_article
+output:
+  bookdown::pdf_document2
+    # extra_dependencies: ["subcaption", "longtable"]
 ---
@@
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Intro
+ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
 Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
@@
 Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+
+and these will be the conclusions
 
@@
 
-The potential demand calculations show that the current and potential number of cyclists both follow a bell-shaped distribution, with the number of trips peaking around the 3-5km commuting distance and then going back down for longer distances (See Figure \ref{fig:pot_dem_histograms}).
+The potential demand calculations show that the current and potential number of cyclists both follow a bell-shaped distribution, with the number of trips peaking around the 3-5km commuting distance and then going back down for longer distances (See Figure @\ref(fig:potdemhistograms)).
 
-\begin{figure}[h!]
-    \centering % puts the image in the horizontal centre of the page
-    \includegraphics[width=.6\textwidth]{../../data/Manchester/Plots/desire_facet_cycling.png} 
-    \caption{Current and Potential Cycling Demand} 
-    \label{fig:desire_facet_cycling} 
-\end{figure}
+```{r potdemhistograms, fig.cap="Current and Potential Cycling Demand", out.width="80%"}
+knitr::include_graphics("../../data/Manchester/Plots/desire_facet_cycling.png")
+```
+
+<!-- \begin{figure}[h!] -->
+<!--     \centering % puts the image in the horizontal centre of the page -->
+<!--     \includegraphics[width=.6\textwidth]{../../data/Manchester/Plots/desire_facet_cycling.png}  -->
+<!--     \caption{Current and Potential Cycling Demand}  -->
+<!--     \label{fig:desire_facet_cycling}  -->
+<!-- \end{figure} -->
 
@@
 
+<!-- \begin{figure} [h!] -->
+<!-- \centering -->
+<!-- \captionsetup{font=footnotesize,labelfont=footnotesize} % size of captions -->
+<!-- \begin{subfigure}{.45\textwidth} -->
+<!--   \centering -->
+<!--   \includegraphics[width=1\linewidth]{../../data/Manchester/Plots/Growth_Results/growth_existing_infra_satisfied_km_all_flow_column.png} -->
+<!--   \caption{Alg 1 (Utilitarian)} -->
+<!--   \label{fig:growth_utilitarian_satisfied_all} -->
+<!-- \end{subfigure} -->
+<!-- \begin{subfigure}{.45\textwidth} -->
+<!--   \centering -->
+<!--   \includegraphics[width=1\linewidth]{../../data/Manchester/Plots/Growth_Results/growth_community_4_satisfied_km_all_flow_column.png} -->
+<!--   \caption{Alg 2 (Egalitarian)} -->
+<!--   \label{fig:growth_egalitarian_satisfied_all} -->
+<!-- \end{subfigure} -->
+<!-- \caption{Comparing Overall Person-Km Satisfied (Manchester)} -->
+<!-- \label{fig:growth_existing_infra_satisfied} -->
+<!-- \end{figure} -->
+
 Overarching Policies

### 2020-10-30 README.Rmd
Add paper to readme

@@
 ---
-title: "Road Segment Prioritization for Bicycle Infrastructure"
-output: github_document
-fontsize: 10pt
+title: Road Segment Prioritization for Bicycle Infrastructure
+date: "`r Sys.Date()`"
+geometry: margin=2cm
+bibliography: mybibfile.bib
+#linenumbers: true
+numbersections: true
+#output: rticles::elsevier_article
+output:
+  # bookdown::github_document2
+  bookdown::pdf_document2
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+```{r, echo=FALSE, eval=FALSE}
+file.rename("README.pdf", "cycle-network-paper.pdf")
 ```
+
+# Ideas and discussion
+
+Road Segment Prioritization for Bicycle Infrastructure
+================
 
@@
 
-There are a couple of files that cannot be synced to github due to their size.
-These files are neseccary for the scripts to run.
-Below are links to where you can download them, and instructions on where to place them in the repo file structure
+There are a couple of files that cannot be synced to github due to their
+size. These files are neseccary for the scripts to run. Below are links
+to where you can download them, and instructions on where to place them
+in the repo file structure
 
@@
 
-* Source: https://www.nomisweb.co.uk/census/2011/bulk/rOD1  ---> Choose File **"WU03EW"**
-* Location in Repo: data-raw/flow_data.csv
+  - Source: <https://www.nomisweb.co.uk/census/2011/bulk/rOD1> —\>
+    Choose File **“WU03EW”**
+  - Location in Repo: data-raw/flow\_data.csv
 
@@
 
-* Source: http://geoportal.statistics.gov.uk/datasets/826dc85fb600440889480f4d9dbb1a24_0
-* Location in Repo: data-raw/MSOA_2011_Boundaries/[Add files here]
+  - Source:
+    <http://geoportal.statistics.gov.uk/datasets/826dc85fb600440889480f4d9dbb1a24_0>
+  - Location in Repo: data-raw/MSOA\_2011\_Boundaries/\[Add files here\]
 
-***
+-----
 
@@
 
-The scripts should be run in the order they are numbered in (and listed in here). The only exception is _x_dodgr_weighting_profiles.R.
+The scripts should be run in the order they are numbered in (and listed
+in here). The only exception is \_x\_dodgr\_weighting\_profiles.R.
 
-Check the readme [here](https://github.com/Hussein-Mahfouz/Bicycle-Network-Optimization) for detailed info on each script
+Check the readme
+[here](https://github.com/Hussein-Mahfouz/Bicycle-Network-Optimization)
+for detailed info on each script
+
+<!-- Keywords -->
+
+<!-- ======== -->
+
+<!-- Highlights -->
+
+<!-- ======== -->
+
+# Introduction
+
+Intro ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
+Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
+Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
+Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
+Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
+Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
+Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+
+and these will be the conclusions
+
+# Background
+
+## What Affects the Decision To Cycle
+
+Segregated cycling infrastructure\footnote{Segregated cycling infrastructure refers to road space that is allocated to cyclists only, with physical separation to protect cyclists from other modes of transport.} has been shown to increase cycling uptake [@aldred2019impacts; @goodman2014new; @marques2015infrastructure], with the separation from motorized vehicles being key [@winters2011motivators].
+Revealed preference of cyclists shows that they are willing to deviate from the most efficient routes in order to commute on safer roads [@crane2017longitudinal].
+However, such deviations are only considered if they do not considerably increase route circuitry; behaviour studies have found that the probability of choosing a route decreases in proportion to its length relative to the shortest route [@broach2011bicycle; @winters2010far].
+Another defining feature for cycling infrastructure is how well connected it is.
+Cyclists prefer cohesive infrastructure, particularly when cycling on arterial roads with high levels of motorized traffic [@stinson2003commuter], and the lack of well-connected cycling infrastructure is one of the main obstacles to increasing cycling uptake [@caulfield2012determining].
+While direct and cohesive cycling networks have been shown to positively impact cycling rates, density\footnote{making an area's bicycle network denser means adding more cycling routes in the area and thereby giving cyclists more route options} of the cycling network is also vital [@schoner2014missing].
+
+## Planning Cycling Networks
+
+*Optimization* techniques have been used to propose improvements to cycling networks.
+@mesbah2012bilevel propose a bi-level formulation to optimize allocation of cycling lanes to the network without exceeding a set budget.
+The upper level is the proposed interventions and the lower level is the route choices made by users in reaction to changes in the network.
+The problem accounts for the effect of cycling lanes on car traffic, and attempts to maximize utilization of said lanes with minimal impact on car travel times.
+To improve cohesion of the suggested network, a constraint is added so that each link with a bike lane should be connected to at least one destination.
+Car usage is not considered by @mauttone2017bicycle, who develop an optimization framework that aims to minimize the total user cost of cycling on the network.
+The aggregate flow on links is obtained by using shortest paths to route existing cycling demand onto the road network, and the solution is a proposed set of links where cycling infrastructure should be added in order to minimize the overall travel cost of cyclists across the network.
+The cost of traversing a link is given as a function of its length and whether or not it has cycling infrastructure, and a discontinuity penalty is also added to prioritize connected road segments.
+The problem has also been solved by attempting to find the minimum cost of improving roadway links to meet a desired level of service (LOS) [@duthie2014optimization].
+In this formulation, all OD pairs need to be connected by roads that meet the desired LOS, and a directness constraint is added so that paths between OD pairs do not exceed a certain multiple of the shortest path.
+
+\vspace{0.3cm}
+
+These problem formulations do not explicitly solve for continuity, which is dealt with using a either (a) a constraint specifying that each link with a bike lane should be connected to at least one destination [@mesbah2012bilevel], (b) a constraint on deviation from shortest paths [@duthie2014optimization], or (c) a discontinuity penalty [@mauttone2017bicycle].
+To solve for continuity, the graph-theoretic concept of *connected components*, has been used.
+@natera2019data study the existing cycling network in terms of its disconnected components and introduce two different algorithms to connect these components by their most critical links\footnote{\textit{link} refers to a road segment throughout this research} and, in doing so, measure the size of the growth of the largest connected component as a function of the kilometers of network added.
+They observe that small investments at strategic points have a large impact on connectivity in most cases.
+The concept of connected components is also at the core of the methodology proposed by @olmos2020data.
+After routing the cycling demand onto the network links, they use percolation theory to filter out the links based on the aggregate flow\footnote{\textit{flow} is used throughout this research to refer to the cycling demand when it is routed onto the road network.
+The flow on any road segment is the cumulative demand on it, resulting from cyclists commuting between various OD pairs} passing through them.
+They vary the flow threshold for filtering to identify the minimum flow at which the whole city is connected by a giant component.
+The results show a cycling network that connects the entire city, and subtracting links intersecting with current cycling infrastructure identifies links proposed for intervention.
+
+\vspace{0.3cm}
+
+The problem formulations outlined above look at the network as a whole when attempting to improve it.
+An alternative approach is to identify the different sub-networks that exist within the larger network, and work on improving each separately.
+Trip patterns in a city are not uniformly distributed geographically, and *community finding* methods have been used to partition study areas into localized areas that experience a disproportionate number of trips within them.
+@akbarzadeh2018designing use a modularity maximization approach [@blondel2008fast] on taxi trip data to identify 7 different communities in the city of Isfahan, Iran.
+An optimization problem is then formulated to connect nodes within each community with cycling infrastructure.
+The emphasis is on connectivity within the communities, not between them.
+@bao2017planning adopt a similar methodology, but use hierarchical clustering to specify the desired number of clusters.
+They use a greedy network expansion algorithm, where the link with the highest benefit/cost ratio in each cluster is selected, and the network is grown by adding neighboring links to the solution until a budget limit is met.
+The benefit is the flow on the link, and each link is assigned a cost based on current road conditions.
+
+## Underlying Ethical Principles
+
+The methodologies in Section **How do I cite a section??** are underpinned by different ethical principles, even though these principles are not explicitly acknowledged by the authors.
+This is important since different ethical principles constitute different problem formulations and targets.
+Broadly speaking, transport appraisal can be based on either utilitarian or egalitarian principles.
+The former seeks to maximize the overall benefit, while the latter is concerned with a fair distribution of benefits [@jafino2020transport].
+@nahmias2017integrating criticize the utilitarian approach that has been historically popular in the evaluation of transport investments, explaining how the maximization of overall benefit fails to account for the distribution of that benefit among communities or individuals.
+@lucas2016method explain how transport studies have traditionally looked at the bigger picture without studying the distribution of investments on the different parts of the study area, and go on to propose an egalitarian approach that ensures the dis-aggregation of transport policy benefits across the study area.
+@pereira2017distributive also emphasize the need for a more egalitarian approach to transport planning.
+They highlight accessibility as a cornerstone of distributive justice, and contend that policies should aim to distribute investments in a way that minimizes spatial variations in accessibility.
+This research attempts to provide a methodology that is grounded in egalitarian principles.
+**Write some more here**
+
+# Data and Geographical Scale of Analysis
+
+The analysis is heavily dependant on Origin-Destination census data (commuter data).
+Commuter data in the UK is publicly available at the Middle layer Super Output Area (MSOA) level; the average MSOA has a population of 8209 [@ofn2018population].
+@iacono2010measuring note that such large travel zones are not ideal for understanding route choice behaviour of cyclists and pedestrians.
+They also give rise to an \`ecological fallacy' whereby average characteristics are assumed to apply to all residents of the aggregated geographical area.
+Given that more granular data is not publicly available, the study uses MSOA-level commuter data.
+The methodology is however applicable to more granular commuter data should it become available.
+
+# Calculating Potential Cycling Demand
+
+The Propensity to Cycle Tool (PCT) [@lovelace2017propensity] is used to estimate the proportion of cyclists ($\boldsymbol{C_{p}}$) for each MSOA pair should the government achieve its target of doubling cycling by 2025.
+The PCT uses the following logistic regression model to calculate $\boldsymbol{C_{p}}$:
+
+```{=tex}
+\begin{align}\label{eq:pcteqn}
+     logit(C_{p}) = & -4.018 - 0.6369d +  1.988\sqrt{d} + 0.008775d^2\\ & - 0.2555s + 0.00206ds -0.1234\sqrt{d}s\nonumber 
+\end{align}
+```
+\noindent where $\boldsymbol{d}$ and $\boldsymbol{s}$ are the distance and slope respectively for the OD pair.
+The authors use square and square-root distance terms \`\`to capture the non-linear impact of distance on the likelihood of cycling", and interaction terms to capture the combined effect of slope and distance [@lovelace2017propensity].
+
+@simini2012
+
+The potential demand calculations show that the current and potential number of cyclists both follow a bell-shaped distribution, with the number of trips peaking around the 3-5km commuting distance and then going back down for longer distances (see Figure \@ref(fig:potdemhistograms).
+
+```{r potdemhistograms, fig.cap="Current and Potential Cycling Demand", out.width="80%", echo=FALSE}
+knitr::include_graphics("data/Manchester/Plots/desire_facet_cycling.png")
+```
+
+# Routing
+
+The next step is to route the potential cycling demand ($\boldsymbol{C_{p}}$) between all OD pairs onto the road network.
+
+To conduct routing, the following is considered:
+
+1.
+**Cyclist Preference**: Work done by @dill2013four on examining cyclist typologies determined that around 60% of Portland residents fit under the *interested but concerned* category.
+These were people that enjoyed cycling but avoided it due safety concerns.
+The key to encouraging this group was to create a low-stress cycling network, not only though segregated infrastructure but also by planning routes that passed through residential streets.
+2.
+**Low-Traffic Neighborhoods**: The UK Department for Transport is allocating funding to local authorities to invest in Active Transport, partially through the creation of LTNs [@departmentgearchange2020].
+This includes closing off residential streets to motorized traffic
+3.
+**Existing Cycling Infrastructure**: Utilizing existing cycling infrastructure makes economic sense, as small investments may lead to large connectivity gains as the disconnected cycling infrastructure gets joined together.
+
+The weighting profiles are therefore adjusted to favor less-stressful streets (based on information from Table \ref{table:osmroadtypes}), and roads with existing cycling infrastructure.
+This is also in line with the creation of LTNs, as residential streets are those where motorized traffic is most likely to be banned in the creation of LTNs.
+
+# Road Segment Prioritization
+
+<!-- \begin{figure} [h!] -->
+
+<!-- \centering -->
+
+<!-- \captionsetup{font=footnotesize,labelfont=footnotesize} % size of captions -->
+
+<!-- \begin{subfigure}{.45\textwidth} -->
+
+<!--   \centering -->
+
+<!--   \includegraphics[width=1\linewidth]{../../data/Manchester/Plots/Growth_Results/growth_existing_infra_satisfied_km_all_flow_column.png} -->
+
+<!--   \caption{Alg 1 (Utilitarian)} -->
+
+<!--   \label{fig:growth_utilitarian_satisfied_all} -->
+
+<!-- \end{subfigure} -->
+
+<!-- \begin{subfigure}{.45\textwidth} -->
+
+<!--   \centering -->
+
+<!--   \includegraphics[width=1\linewidth]{../../data/Manchester/Plots/Growth_Results/growth_community_4_satisfied_km_all_flow_column.png} -->
+
+<!--   \caption{Alg 2 (Egalitarian)} -->
+
+<!--   \label{fig:growth_egalitarian_satisfied_all} -->
+
+<!-- \end{subfigure} -->
+
+<!-- \caption{Comparing Overall Person-Km Satisfied (Manchester)} -->
+
+<!-- \label{fig:growth_existing_infra_satisfied} -->
+
+<!-- \end{figure} -->
+
+# Overarching Policies
+
+While segregated, connected, and direct cycling infrastructure is key to achieving high levels of cycling, research has shown that it cannot exist in a vacuum.
+@wardman2007factors developed a mode choice model for the UK and their results showed that improved cycling infrastructure on its own only had modest impacts on mode shift, and even the unlikely scenario of all urban routes being serviced by segregated bike lanes was forecast to increase cycling mode share by only 3%.
+However, cities that invest in more comprehensive cycling projects show a more significant increase in the number of cyclists as well as the cycling mode share [@pucher2010infrastructure].
+These cities do not just focus on infrastructure, but on general policies as well as restricting car use.
+Evaluation of policies in Denmark and Germany and the Netherlands has shown that their high cycling mode share is down to a broader set of policies that also include traffic calming, cycling rights of way, bike parking, integration with the public transport network, and making driving cars both expensive and inconvenient [@pucher2008making].
+While these policies are outside the scope of this research, it is important to recognize their key role in bringing about an increase in levels of cycling.
+
+# Conclusions
+
+# References {#references .unnumbered}
+
+<!-- to fix indentation: https://github.com/crsh/papaja/issues/37#issuecomment-104185288 -->
+
+```{=tex}
+\setlength{\parindent}{-0.5in}
+\setlength{\leftskip}{0.5in}
+\setlength{\parskip}{8pt}
+```

### 2020-10-30 README.Rmd
Update paper, show Figure of growth in readme

@@
 output:
-  # bookdown::github_document2
-  bookdown::pdf_document2
+  bookdown::github_document2
+  # bookdown::pdf_document2:
+  #   extra_dependencies: ["subcaption", "longtable"]
 ---
@@
 
+```{r growth, fig.cap="teswt", out.width="48%", fig.show='hold'}
+knitr::include_graphics(c(
+  "data/Manchester/Plots/Growth_Results/growth_existing_infra_satisfied_km_all_flow_column.png",
+  "data/Manchester/Plots/Growth_Results/growth_community_4_satisfied_km_all_flow_column.png"
+))
+```
+
+<!-- Todo: uncomment for final submission -->
 <!-- \begin{figure} [h!] -->
@@
 
-<!--   \includegraphics[width=1\linewidth]{../../data/Manchester/Plots/Growth_Results/growth_existing_infra_satisfied_km_all_flow_column.png} -->
+<!--   \includegraphics[width=1\linewidth]{data/Manchester/Plots/Growth_Results/growth_existing_infra_satisfied_km_all_flow_column.png} -->
 
@@
 
-<!--   \includegraphics[width=1\linewidth]{../../data/Manchester/Plots/Growth_Results/growth_community_4_satisfied_km_all_flow_column.png} -->
+<!--   \includegraphics[width=1\linewidth]{data/Manchester/Plots/Growth_Results/growth_community_4_satisfied_km_all_flow_column.png} -->
 

### 2020-10-30 README.Rmd

@@
 
-The methodologies in Section **How do I cite a section??** are underpinned by different ethical principles, even though these principles are not explicitly acknowledged by the authors.
+The methodologies in Section \@ref(planning-cycling-networks) are underpinned by different ethical principles, even though these principles are not explicitly acknowledged by the authors.
 This is important since different ethical principles constitute different problem formulations and targets.

### 2020-10-30 README.Rmd
commented out reproducibility instructions in readme

@@
 
-Road Segment Prioritization for Bicycle Infrastructure
-================
+- Community Detection
+- Other Cities
 
-## Missing Data
+<!-- ## Missing Data -->
 
-There are a couple of files that cannot be synced to github due to their
-size. These files are neseccary for the scripts to run. Below are links
-to where you can download them, and instructions on where to place them
-in the repo file structure
+<!-- There are a couple of files that cannot be synced to github due to their -->
+<!-- size. These files are neseccary for the scripts to run. Below are links -->
+<!-- to where you can download them, and instructions on where to place them -->
+<!-- in the repo file structure -->
 
-Flow Data (2011 Census Origin-Destination Data):
+<!-- Flow Data (2011 Census Origin-Destination Data): -->
 
-  - Source: <https://www.nomisweb.co.uk/census/2011/bulk/rOD1> —\>
-    Choose File **“WU03EW”**
-  - Location in Repo: data-raw/flow\_data.csv
+<!--   - Source: <https://www.nomisweb.co.uk/census/2011/bulk/rOD1> —\> -->
+<!--     Choose File **“WU03EW”** -->
+<!--   - Location in Repo: data-raw/flow\_data.csv -->
 
-Middle Layer Super Output Areas (December 2011) Boundaries:
+<!-- Middle Layer Super Output Areas (December 2011) Boundaries: -->
 
-  - Source:
-    <http://geoportal.statistics.gov.uk/datasets/826dc85fb600440889480f4d9dbb1a24_0>
-  - Location in Repo: data-raw/MSOA\_2011\_Boundaries/\[Add files here\]
+<!--   - Source: -->
+<!--     <http://geoportal.statistics.gov.uk/datasets/826dc85fb600440889480f4d9dbb1a24_0> -->
+<!--   - Location in Repo: data-raw/MSOA\_2011\_Boundaries/\[Add files here\] -->
 
------
+<!-- ----- -->
 
-## Scripts
+<!-- ## Scripts -->
 
-The scripts should be run in the order they are numbered in (and listed
-in here). The only exception is \_x\_dodgr\_weighting\_profiles.R.
+<!-- The scripts should be run in the order they are numbered in (and listed -->
+<!-- in here). The only exception is \_x\_dodgr\_weighting\_profiles.R. -->
 
-Check the readme
-[here](https://github.com/Hussein-Mahfouz/Bicycle-Network-Optimization)
-for detailed info on each script
+<!-- Check the readme -->
+<!-- [here](https://github.com/Hussein-Mahfouz/Bicycle-Network-Optimization) -->
+<!-- for detailed info on each script -->
 
@@
 <!-- ======== -->
+
+### Hello
 
@@
 Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
-
-and these will be the conclusions
 

### 2020-10-30 README.Rmd
paper updates and formatting from latex -> rmd

@@
 
-### Hello
-
 # Introduction
@@
 
-Segregated cycling infrastructure\footnote{Segregated cycling infrastructure refers to road space that is allocated to cyclists only, with physical separation to protect cyclists from other modes of transport.} has been shown to increase cycling uptake [@aldred2019impacts; @goodman2014new; @marques2015infrastructure], with the separation from motorized vehicles being key [@winters2011motivators].
+Segregated cycling infrastructure^[Segregated cycling infrastructure refers to road space that is allocated to cyclists only, with physical separation to protect cyclists from other modes of transport.] has been shown to increase cycling uptake [@aldred2019impacts; @goodman2014new; @marques2015infrastructure], with the separation from motorized vehicles being key [@winters2011motivators].
 Revealed preference of cyclists shows that they are willing to deviate from the most efficient routes in order to commute on safer roads [@crane2017longitudinal].
@@
 Cyclists prefer cohesive infrastructure, particularly when cycling on arterial roads with high levels of motorized traffic [@stinson2003commuter], and the lack of well-connected cycling infrastructure is one of the main obstacles to increasing cycling uptake [@caulfield2012determining].
-While direct and cohesive cycling networks have been shown to positively impact cycling rates, density\footnote{making an area's bicycle network denser means adding more cycling routes in the area and thereby giving cyclists more route options} of the cycling network is also vital [@schoner2014missing].
+While direct and cohesive cycling networks have been shown to positively impact cycling rates, density^[making an area's bicycle network denser means adding more cycling routes in the area and thereby giving cyclists more route options} of the cycling network is also vital [@schoner2014missing]].
 
@@
 To solve for continuity, the graph-theoretic concept of *connected components*, has been used.
-@natera2019data study the existing cycling network in terms of its disconnected components and introduce two different algorithms to connect these components by their most critical links\footnote{\textit{link} refers to a road segment throughout this research} and, in doing so, measure the size of the growth of the largest connected component as a function of the kilometers of network added.
+@natera2019data study the existing cycling network in terms of its disconnected components and introduce two different algorithms to connect these components by their most critical links^[*link* refers to a road segment throughout this research] and, in doing so, measure the size of the growth of the largest connected component as a function of the kilometers of network added.
 They observe that small investments at strategic points have a large impact on connectivity in most cases.
 The concept of connected components is also at the core of the methodology proposed by @olmos2020data.
-After routing the cycling demand onto the network links, they use percolation theory to filter out the links based on the aggregate flow\footnote{\textit{flow} is used throughout this research to refer to the cycling demand when it is routed onto the road network.
-The flow on any road segment is the cumulative demand on it, resulting from cyclists commuting between various OD pairs} passing through them.
+After routing the cycling demand onto the network links, they use percolation theory to filter out the links based on the aggregate flow^[*flow* is used throughout this research to refer to the cycling demand when it is routed onto the road network.
+The flow on any road segment is the cumulative demand on it, resulting from cyclists commuting between various OD pairs] passing through them.
 They vary the flow threshold for filtering to identify the minimum flow at which the whole city is connected by a giant component.
@@
 
-@simini2012
-
 The potential demand calculations show that the current and potential number of cyclists both follow a bell-shaped distribution, with the number of trips peaking around the 3-5km commuting distance and then going back down for longer distances (see Figure \@ref(fig:potdemhistograms).
 
-```{r potdemhistograms, fig.cap="Current and Potential Cycling Demand", out.width="80%", echo=FALSE}
+```{r potdemhistograms, fig.cap="Distribution of Potential Cycling Demand", out.width="32%", fig.show='hold', echo=FALSE}
+knitr::include_graphics(c(
+  "data/Manchester/Plots/histogram_distance_all_vs_cycling.png",
+  "data/Manchester/Plots/histogram_distance_all_vs_cycling_potential.png",
+  "data/Manchester/Plots/histogram_distance_cycling_potential_vs_current.png"
+))
+```
+
+```{r desirefacetcycling, fig.cap="Current and Potential Cycling Demand", out.width="80%", echo=FALSE}
 knitr::include_graphics("data/Manchester/Plots/desire_facet_cycling.png")
 ```
+
+It should be noted that the calculations assume a future that is constrained by physical geography; i.e.
+we consider cycling in the traditional sense.
+Recently there have been various micro-mobility solutions, including e-bikes, that allow commuters to traverse longer distances and hillier roads with less effort than traditional bicycles.
+While these modes would probably be associated with less geographical impedance, it is beyond the scope of this work to integrate that into the analysis.
+Doing so is partially restricted by the lack of data on the proliferation of these modes, which raises the point that perhaps the census data category of `Bicycle' is too vague, and should be further dis-aggregated to distinguish between traditional bicycles and other forms of micro-mobility.
 
@@
 The next step is to route the potential cycling demand ($\boldsymbol{C_{p}}$) between all OD pairs onto the road network.
+
+**ADD TABLE**
 
@@
 
-The weighting profiles are therefore adjusted to favor less-stressful streets (based on information from Table \ref{table:osmroadtypes}), and roads with existing cycling infrastructure.
+The weighting profiles are therefore adjusted to favor less-stressful streets (based on information from Table **REFERENCE THE TABLE**), and roads with existing cycling infrastructure.
 This is also in line with the creation of LTNs, as residential streets are those where motorized traffic is most likely to be banned in the creation of LTNs.
+
+**ADD TABLE**
+
+\noindent A weighted distance $\boldsymbol{d_{w}}$ for each road segment is calculated as following ^[The **dodgr** r package [@padgham2019dodgr] is used to route cycling demand onto the road network.
+The package uses the OpenStreetMaps (OSM) road network and allows the user to assign weights to roads based on their type.
+The routing is done based on weighted shortest paths, with the distance along each road segment being divided by a factor to obtain the weighted distance for routing.
+It is more intuitive to multiply when weighting a network, but the dodgr package divides by numbers between 0 and 1, which achieves the same result.
+For the sake of reproducibility, we stick to the convention used in the package.]:
+
+\begin{equation}\label{eq:weight_distance}
+    d_{w} = d_{unw}  \:/\:  W
+\end{equation}
+
+\noindent where $\boldsymbol{d_{unw}}$ is the unweighted distance and $\boldsymbol{W}$ is the weight from Table \ref{table:weightprofiles}.
+
+All weights are between 0 and 1, and the values in the \textit{Weighted} profile are chosen so as to be inversely proportional to the stress level experienced by cyclists on them.
+The \textit{Unweighted} weighting profile is used to compare increases in route length resulting from two different approaches:
+
+1. **Weighted**: Relatively high impedance on Primary and Trunk roads (to minimize cycling on them).
+2. **Weighted\_2**: Avoiding Primary and Trunk Roads completely.
+
+Comparing the cycling demand routed on the weighted and unweighted road network allows us to get a better understanding of the importance of different road types.
+In the case of Manchester, trunk roads bisect the city and are a major part of unweighted shortest paths (Figure \@ref(fig:flowsfacetunweighted)).
+On the other hand, cycleways are not part of unweighted shortest paths, and so very little of the cycling demand is routed through them.
+In the weighted network, cycleways are much better utilized, and the majority of the cycling demand passes through tertiary roads, as expected.
+
+```{r flowsfacetunweighted, fig.cap="Flow Results Based on **Unweighted** Shortest Paths (Manchester)", out.width="90%", echo=FALSE}
+knitr::include_graphics("data/Manchester/Plots/flows_facet_unweighted_Manchester.png")
+```
+
+```{r flowsfacetweighted, fig.cap="Flow Results Based on **Weighted** Shortest Paths (Manchester)", out.width="90%", echo=FALSE}
+knitr::include_graphics("data/Manchester/Plots/flows_facet_weighted_Manchester.png")
+```
+
+<!-- The difference between aggregate flow on weighted and unweighted networks is dependant on the road network of the city.
+Comparing Manchester to Nottingham, we see that trunk roads are much more important in the former, as over 25\% of flow on the unweighted road network passes through them.
+For Nottingham, less than 10\% of the flow on the unweighted network passes through trunk roads, but almost 25\% of the flow passes through tertiary roads (Figure \ref{fig:perc_person-km}).
+-->
+
+The results of routing potential cycling demand on the weighted and unweighted networks are understandably quite different.
+From Figure \@ref(fig:flowsfacetunweighted) we can see that trunk and primary roads are the most efficient means of traversing the road network of Manchester.
+Both of these road types are classified as Primary A roads according to the UK Department for Transport's road classification (**Reference the table**), and are therefore part of the Primary Route Network (PRN) [@department2012guidance].
+The PRN has the widest, most direct roads on the network, and carries most of the through traffic.
+This includes freight, with all roads in the PRN being required by law to provide unrestricted access to trucks up to 40 tonnes [@department2012guidance].
+
+We choose to avoid routing the potential cycling demand on Primary A Roads for the following 2 reasons:
+
+1. **Logistical Difficulty**: Changes on these roads need to be agreed upon by all affected authorities [@department2012guidance], which may prove to be difficult. 
+2.
+**Low Traffic Neighborhoods (LTNs)**: The UK government is aiming to restrict access to motorized vehicles on residential roads to create LTNs [@departmentgearchange2020].
+This is part of a policy to prevent automobile rat-running and make streets more accessible to cyclists and pedestrians.
+Under such a policy, Primary A roads would become even more essential for motorized traffic and it would be more difficult to reallocate road space on these roads to cyclists.
+
+Routing potential cycling demand on a weighted network is more in line with government policy to create LTNs.
+Figure \@ref(fig:flowsfacetweighted) shows that routing on the weighted network significantly reduces flow on the trunk and primary roads, but does not eliminate it completely.
+This is intentional, as the impedance on these roads is only slightly higher than remaining road types (See Table **Reference the table**).
+Potential cycling demand is only routed on these roads if there are no routes through other roads that offer comparable directness.
+
+Banning cycling flow completely on trunk and primary roads may result in excessively circuitous paths, as seen in Figure \ref{fig:boxplotcircuity}.
+When routing using the weighting profile in Table **Reference Table**, we see that shortest paths increase by less than 5\% on average from unweighted shortest paths, with the largest increases still below 30\%.
+When routing on primary and trunk roads is banned, the average increase relative to unweighted shortest paths rises to 10\%, with certain locations experiencing more significant negative effects on accessibility.
+
+```{r boxplotcircuity, fig.cap="Effect of Banning Cyclists from Trunk and Primary Roads for all OD Pairs (Manchester)", out.width="35%", echo=FALSE}
+knitr::include_graphics("data/Manchester/Plots/boxplot_weighted_unweighted_distances.png")
+```
+
+Given that cyclists will only deviate from shortest paths by a certain amount to access better cycling infrastructure (as explained in Section \@ref(what-affects-the-decision-to-cycle), allowing flow on some stretches of trunk and primary roads is necessary to insure cycling uptake and equitable access to cycling infrastructure.
+In its new vision for walking and cycling, the Department for Transport acknowledges that minimal segregated stretches of bicycle lanes on main roads will be necessary to avoid circuitous cycling networks [@departmentgearchange2020].
+
+Weighting the road network also allows us to better utilize existing cycling infrastructure, as can be seen by the higher flow on cycleways in Figure \@ref(fig:flowsfacetweighted).
+Again, the small differences in impedance between cycleways and other road types mean that cycleways that require significant deviation are not routed on.
+
+It should be reiterated that the weighting profile used for routing has been developed for the purposes of this study.
+It creates a hierarchy of road preference that is grounded in cyclist preference and government plans to create LTNs.
+Sensitivity analysis should be done to determine an optimal weighting profile, but given the variation in city road networks , <!-- (Figure \ref{fig:perc_person-km}) -->, these would probably require calibration to the specific city.
+More accurate routing could be carried out given the availability of road-level data.
+In such cases we would add additional impedance to specific roads, giving more useful routing results than the current methodology which considers all roads of the same type to be equivalent.
+
+One use-case of such granular data would be to identify roads that serve schools.
+The Department of Transport notes that the number of school children being driven to school has trebled over the past 40 years [@departmentgearchange2020], and so having cycling infrastructure serving schools is key to achieving the government target of getting more children to cycle.
+This would not be difficult, as over 75\% of children in the UK live within a 15 minute cycle from their school [@departmentcycleinfradesign2020].
+@goodman2019scenarios show that if dutch levels of cycling were achieved in the UK, the \% of children cycling to school could increase from 1.8\% to 41\%.
+
+In their typology of cyclists, @dill2013four found that a majority of people who say they would never cycle had never cycled to school, whereas confident cyclists were those most likely to have cycled to school.
+Getting people to cycle from a young age is therefore key to achieving societal change in commuting habits.
 

### 2020-10-31 README.Rmd
paper: community detection section

@@
 
-Banning cycling flow completely on trunk and primary roads may result in excessively circuitous paths, as seen in Figure \ref{fig:boxplotcircuity}.
+Banning cycling flow completely on trunk and primary roads may result in excessively circuitous paths, as seen in Figure \@ref(fig:boxplotcircuity).
 When routing using the weighting profile in Table **Reference Table**, we see that shortest paths increase by less than 5\% on average from unweighted shortest paths, with the largest increases still below 30\%.
@@
 Getting people to cycle from a young age is therefore key to achieving societal change in commuting habits.
+
+# Community Detection
+
+One of the main aims of this research is to incorporate egalitarian principles by fairly distributing investments in cycling infrastructure.
+One way of quantifying this is to split up the city into smaller geospatial areas and target equal investment in each of those areas.
+Community detection offers us a way to delineate such a split; cyclists are limited in their commuting distance (see Figure \@ref(fig:cyclinghistmanc)), and so trip attractors are more likely to have a local catchment area of cyclists.
+
+```{r cyclinghistmanc, fig.cap="Cycling Commuting Distance - Manchester (2011 Census Data)", out.width="45%", echo=FALSE}
+knitr::include_graphics("data/Manchester/Plots/histogram_distance_cycling.png")
+```
+
+In our case, the network is the city; the nodes are the population-weighted MSOA centroids and the links connecting each MSOA pair are weighted by the potential cycling demand between them.
+The Louvian method  [@blondel2008fast] is used to separate MSOAs into communities.
+Potential cycling demand is used since we assume that this is what the cycling demand will be once the cycling infrastructure is added.
+To assign road links to communities, the following steps are carried out:
+
+1. Create links between MSOA centroids and weigh these links by potential cycling demand between them.
+2. Use Louvian method to determine optimal number of communities and assign each MSOA centroid to a community.
+3. Assign each road link to the same community as the closest MSOA centroid to it.
+
+The results show that Manchester can be split into three large communities and one small one (Figure \@ref(fig:communitiesmanchester)).
+
+```{r communitiesmanchester, fig.cap="Communities Based on Potential Cycling Demand (Manchester)", out.width="75%", echo=FALSE}
+knitr::include_graphics("data/Manchester/Plots/communities_alternative_Manchester.png")
+```
 

### 2020-11-02 README.Rmd
cleaned up script algorithm scripts (8.0/8.1/8.2)

@@
 output:
-  bookdown::github_document2
-  # bookdown::pdf_document2:
-  #   extra_dependencies: ["subcaption", "longtable"]
+  # bookdown::github_document2
+  bookdown::pdf_document2:
+    extra_dependencies: ["subcaption", "longtable"]
 ---
@@
 Cyclists prefer cohesive infrastructure, particularly when cycling on arterial roads with high levels of motorized traffic [@stinson2003commuter], and the lack of well-connected cycling infrastructure is one of the main obstacles to increasing cycling uptake [@caulfield2012determining].
-While direct and cohesive cycling networks have been shown to positively impact cycling rates, density^[making an area's bicycle network denser means adding more cycling routes in the area and thereby giving cyclists more route options} of the cycling network is also vital [@schoner2014missing]].
+While direct and cohesive cycling networks have been shown to positively impact cycling rates, density^[making an area's bicycle network denser means adding more cycling routes in the area and thereby giving cyclists more route options] of the cycling network is also vital [@schoner2014missing]].
 
@@
 
-```{r desirefacetcycling, fig.cap="Current and Potential Cycling Demand", out.width="80%", echo=FALSE}
+```{r desirefacetcycling, fig.cap="Current and Potential Cycling Demand", out.width="80%", fig.show='hold', echo=FALSE}
 knitr::include_graphics("data/Manchester/Plots/desire_facet_cycling.png")
@@
 
-```{r flowsfacetunweighted, fig.cap="Flow Results Based on **Unweighted** Shortest Paths (Manchester)", out.width="90%", echo=FALSE}
+```{r flowsfacetunweighted, fig.cap="Flow Results Based on **Unweighted** Shortest Paths (Manchester)", out.width="90%", fig.show='hold', echo=FALSE}
 knitr::include_graphics("data/Manchester/Plots/flows_facet_unweighted_Manchester.png")
@@
 
-```{r flowsfacetweighted, fig.cap="Flow Results Based on **Weighted** Shortest Paths (Manchester)", out.width="90%", echo=FALSE}
+```{r flowsfacetweighted, fig.cap="Flow Results Based on **Weighted** Shortest Paths (Manchester)", out.width="90%", fig.show='hold', echo=FALSE}
 knitr::include_graphics("data/Manchester/Plots/flows_facet_weighted_Manchester.png")
@@
 
-```{r boxplotcircuity, fig.cap="Effect of Banning Cyclists from Trunk and Primary Roads for all OD Pairs (Manchester)", out.width="35%", echo=FALSE}
+```{r boxplotcircuity, fig.cap="Effect of Banning Cyclists from Trunk and Primary Roads for all OD Pairs (Manchester)", out.width="35%", fig.show='hold', echo=FALSE}
 knitr::include_graphics("data/Manchester/Plots/boxplot_weighted_unweighted_distances.png")
@@
 
-```{r cyclinghistmanc, fig.cap="Cycling Commuting Distance - Manchester (2011 Census Data)", out.width="45%", echo=FALSE}
+```{r cyclinghistmanc, fig.cap="Cycling Commuting Distance - Manchester (2011 Census Data)", out.width="45%", fig.show='hold', echo=FALSE}
 knitr::include_graphics("data/Manchester/Plots/histogram_distance_cycling.png")
@@
 
-```{r communitiesmanchester, fig.cap="Communities Based on Potential Cycling Demand (Manchester)", out.width="75%", echo=FALSE}
+```{r communitiesmanchester, fig.cap="Communities Based on Potential Cycling Demand (Manchester)", out.width="75%", fig.show='hold', echo=FALSE}
 knitr::include_graphics("data/Manchester/Plots/communities_alternative_Manchester.png")
@@
 
-```{r growth, fig.cap="teswt", out.width="48%", fig.show='hold'}
+After routing the potential cycling demand onto the road network using weighted shortest paths, we have estimates for the cumulative potential cycling demand passing through all road segments.
+This cumulative demand (referred to as *flow* is then used as a basis for determining where best to invest in segregated cycling infrastructure.
+In doing so, we must account for the motivations and deterrents for cycling identified in Section \@ref(what-affects-the-decision-to-cycle), namely direct and well connected routes.
+
+For this purpose, two algorithms are proposed.
+Both utilize existing infrastructure from the beginning and allow us to compare a solution that focuses on utilitarianism to one that focuses on egalitarianism.
+In both algorithms, links^[*link* refers to a road segment] are selected iteratively and the iteration at which each link is added to the solution is recorded.
+Investments in cycling infrastructure can be limited by budget constraints, so it can be useful to see where best to allocate a defined length of segregated infrastructure.
+
+## Algorithm 1: Utilitarian Growth
+
+1. Identify all links that have segregated cycling infrastructure and add them to the initial solution
+2. Identify all links that neighbor links in the current solution
+3. Select neighboring link with highest flow and add it to the solution
+4. Repeat steps 2 \& 3 until all flow is satisfied or investment threshold is met
+
+This algorithm ensures that the resulting network is connected.
+It also satisfies the directness criteria, since links on the weighted shortest paths are those that have the highest flow passing through them (this is a result of the routing in Section \@ref(what-affects-the-decision-to-cycle).
+
+Algorithm 2: Egalitarian Growth (Focus on Fair Distribution of Resources)
+
+The first algorithm focuses on connectivity and directness, but not on fairly distributing investment.
+The latter is not a requirement for increasing cycling uptake, but it is fundamental for spatial equity, as explained in Section \@ref(underlying-ethical-principles).
+This algorithm incorporates the ideal of fair distribution by using community detection to partition the road network.
+
+The algorithm uses the following logic to ensure fair distribution between communities:
+
+1. Identify all links that have segregated cycling infrastructure and add them to the initial solution
+2. Identify all links that neighbor links in the current solution
+3. Select *from each community* one neighboring link with highest flow and add it to the solution
+4. If there are no more neighboring links in a community, select the link with the highest flow in that community, regardless of connectivity, and add it to the solution
+5. Repeat steps 2, 3 \& 4 until all flow is satisfied or investment threshold is met
+
+Even though we may end up with a more disconnected network, we will have separate connected networks in each community.
+Given that communities are defined by having more internal flow than external flow, this is a satisfactory solution.
+
+```{r growthtotal, fig.cap="Comparing Overall Person-Km Satisfied (Manchester)", out.width="45%", fig.show='hold', echo=FALSE}
 knitr::include_graphics(c(
@@
 
-<!-- Todo: uncomment for final submission -->
-<!-- \begin{figure} [h!] -->
+```{r growthcommunities, fig.cap="Comparing Community Level Person-Km Satisfied (Manchester)", out.width="45%", fig.show='hold', echo=FALSE}
+knitr::include_graphics(c(
+  "data/Manchester/Plots/Growth_Results/growth_existing_infra_satisfied_km_community_flow_column.png",
+  "data/Manchester/Plots/Growth_Results/growth_community_4_satisfied_km_community_flow_column.png"
+))
+```
 
-<!-- \centering -->
+```{r growth3MapandBar, fig.cap="Results of Alg. 2 (Manchester)", out.width="45%", fig.show='hold', echo=FALSE}
+knitr::include_graphics(c(
+  "data/Manchester/Plots/Growth_Results/growth_community_4_priority_all_FLOW.png",
+  "data/Manchester/Plots/Growth_Results/growth_community_4_investment_highways_flow.png"
+))
+```
 
-<!-- \captionsetup{font=footnotesize,labelfont=footnotesize} % size of captions -->
+## Connectivity
 
-<!-- \begin{subfigure}{.45\textwidth} -->
+Existing cycling infrastructure is made up of many disconnected components.
+Both Algorithm 1 and 2 start with all existing segregated cycling infrastructure and aim to create an efficient, connected network.
+Figure \@ref(fig:componentsandGCC) compares the performance of both Algorithms in terms of connectivity gains.
+We can see that, for all three cities analyzed, the number of components is gradually reduced as more infrastructure is added (this is true for both Algorithms).
+However these connectivity gains require considerable investment; both algorithms take almost 150km of investment to reduce the number of components by half.
 
-<!--   \centering -->
+Consistent growth can also be seen for the size of the Largest Connected Component in the proposed bicycle network (Figure \@ref(fig:componentsandGCC)).
+Again, we can see from the results of the three different cities that there is little difference between both Algorithms.
 
-<!--   \includegraphics[width=1\linewidth]{data/Manchester/Plots/Growth_Results/growth_existing_infra_satisfied_km_all_flow_column.png} -->
-
-<!--   \caption{Alg 1 (Utilitarian)} -->
-
-<!--   \label{fig:growth_utilitarian_satisfied_all} -->
-
-<!-- \end{subfigure} -->
-
-<!-- \begin{subfigure}{.45\textwidth} -->
-
-<!--   \centering -->
-
-<!--   \includegraphics[width=1\linewidth]{data/Manchester/Plots/Growth_Results/growth_community_4_satisfied_km_all_flow_column.png} -->
-
-<!--   \caption{Alg 2 (Egalitarian)} -->
-
-<!--   \label{fig:growth_egalitarian_satisfied_all} -->
-
-<!-- \end{subfigure} -->
-
-<!-- \caption{Comparing Overall Person-Km Satisfied (Manchester)} -->
-
-<!-- \label{fig:growth_existing_infra_satisfied} -->
-
-<!-- \end{figure} -->
+```{r componentsandGCC, fig.cap="Network Characteristics", out.width="45%", fig.show='hold', echo=FALSE}
+knitr::include_graphics(c(
+  "data/Manchester/Plots/Growth_Results/growth_existing_infra_components_number_comparisonManchester.png",
+  "data/Manchester/Plots/Growth_Results/growth_existing_infra_components_gcc_comparisonManchester.png"
+))
+```
 

### 2020-11-02 R/_8.0_growing_a_network.R

@@ -19,79 +19,8 @@ check_km_value <- function(graph, km) {
   }
 }
 
-          ################ FUNCTIONS THAT DON'T DEPEND ON COMMUNITY DETECTION ################
+# ----------------------------------- ALGORITHM 1: UTILITARIAN GROWTH ----------------------------------------- #
 
-################################################ FUNCTION 1 ################################################
-###########################################################################################################  
-### THIS FUNCTION IDENTIFIES THE EDGE WITH THE HIGHEST FLOW IN THE WHOLE GRAPH. IT THEN PROCEEDS TO ADD ###
-### NEIGHBORING EDGES INCREMENTALLY UNTIL THE INVESTMENT LENGTH IS MET ###
-###########################################################################################################  
-
-# 1. select investment length (km)
-# 2. Identify edge with highest flow 
-# 3. Add edge to solution
-# 4. Identify all edges that are connected to the current solution
-# 5. Select edge with highest flow and append it to the solution
-# 6. Repeat steps 4 & 5 until the length of the edges in the solution reaches the investment length
-
-# col_name: the column that you are choosing segments based on, passed inside "". Could be something other than "flow"
-growth_one_seed <- function(graph, km, col_name) {
-  
-  ### check if km chosen is a reasonable number (using a predefined function)
-  check_km_value(graph, km)
-  # add an empty sequence column
-  graph <- graph %>% mutate(sequen= NA)
-  #get edge_id of edge with highest flow
-  edge_sel <- graph$edge_id[which.max(graph[[col_name]])]
-  # assign a sequence  to the selected edge
-  #graph <- within(graph, sequen[edge_id == edge_sel] <- 0)
-  graph$sequen[graph$edge_id == edge_sel] <- 0
-  # i keeps track of which iteration a chosen edge was added in
-  i <- 1
-  # we keep track of the edges in the solution, so that we can identify the edges that neighbor them
-  # edges in the solution are those that have been assigned a 'sequen' value
-  chosen <- graph %>% filter(!(is.na(sequen)))
-  # j counts km added. We don't count segments that already have cycling infrastructure
-  j <- sum(chosen$d) - sum(chosen$cycle_infra * chosen$d)
-  #while length of chosen segments is less than specified length 
-  while (j/1000 < km){
[…]

### 2020-11-02 R/_8.1_plot_network_growth.R

@@ -19,7 +19,7 @@ graph_sf %>%
   group_by(Community) %>% 
   summarize(total_person_km = sum(person_km) / 1000) %>% 
   mutate(perc_person_km = (total_person_km / sum(total_person_km) * 100)) 
-########## GGPLOTS SHOWING FLOW/PERSON_KM SATISFIED AT THE NETWORK LEVEL AND AT THE COMMUNITY LEVEL #############
+
 
 # get percentage contibution of each edge to the network (distance, flow, person_km)
 graph_sf <- graph_sf %>%
@@ -42,109 +42,21 @@ no_cycle_infra <- round((sum(graph_sf$d) - sum(cycle_infra$d))  / 1000)
 no_cycle_infra <- no_cycle_infra - 5
 ################
 
-########################################### FUNCTION 1: GROWTH FROM ONE SEED/EDGE ###########################################
 
-# let's grow the network based on the flow column
-grow_flow_1_seed <- growth_one_seed(graph = graph_sf, km = no_cycle_infra, col_name = "flow")
-# save
-saveRDS(grow_flow_1_seed, file = paste0("../data/", chosen_city, "/growth_fun_1.Rds"))
-# read in 
-#grow_flow_1_seed <- readRDS(paste0("../data/", chosen_city,"/growth_fun_1.Rds"))
+# ----------------------------------- ALGORITHM 1: UTILITARIAN GROWTH ----------------------------------------- #
 
-
-# prepare a dataframe for plotting the results
-grow_flow_1_seed_c <- grow_flow_1_seed %>% 
-  arrange(sequen) %>%  # make sure it is arranged by sequen (Otherwise cumulative sum values will be wrong)
-  ungroup %>%   # not sure why it is a grouped df. This only has an effect on the select argument
-  #st_drop_geometry() %>% 
-  dplyr::select(Community, d, flow, cycle_infra, sequen, perc_dist, perc_flow,
-                perc_person_km, perc_dist_comm, perc_flow_comm, perc_person_km_comm, highway)
-
-# cumsum is cumulative sum. We see how much of person_km has been satisfied after each iteration 
-grow_flow_1_seed_c <- grow_flow_1_seed_c %>%
-  mutate(dist_c = cumsum(d/1000),
-         perc_dist_c = cumsum(perc_dist),
-         perc_person_km_c = cumsum(perc_person_km)) %>%
-  # groupby so that you can apply cumsum by community 
[…]

### 2020-11-02 R/_8.2_plot_network_growth_community.R

@@ -8,9 +8,8 @@ graph_sf <- readRDS(paste0("../data/", chosen_city, "/graph_with_flows_weighted_
 # we weigh the flow on each edge by its distance. We can then get how much of the commuter km are satisfied
 graph_sf$person_km <- graph_sf$flow * graph_sf$d
 
-########## GGPLOTS SHOWING FLOW/PERSON_KM SATISFIED AT THE NETWORK LEVEL AND AT THE COMMUNITY LEVEL #############
 
-# get percentage contibution of each edge to the network (distance, flow, person_km)
+# get percentage contribution of each edge to the network (distance, flow, person_km)
 graph_sf <- graph_sf %>%
   mutate(perc_dist = (d/sum(d))  *100,      # edge length as % of total network length
          perc_flow = (flow/sum(flow))  *100, # flow on edge as % of total
@@ -22,434 +21,10 @@ graph_sf <- graph_sf %>%
          perc_person_km_comm = (person_km/sum(person_km))  *100) %>%
   ungroup()
 
-########################################### FUNCTION 3: GROWTH WITH ONE SEED FROM EACH COMMUNITY ###########################################
-###########################################################################################################  
-### THIS FUNCTION IDENTIFIES THE EDGE WITH THE HIGHEST FLOW IN EACH COMMUNITY. THESE EDGES ACT AS ###
-### SEEDS SO THAT THE GRAPH CAN GROW FROM MULTIPLE LOCATIONS, NOT JUST ONE INITIAL EDGE. WHILE ONE ###
-### SEED IS CHOSEN FROM EACH COMMUNITY, THERE IS NO REQUIREMENT FOR THE EDGES APPENDED LATER TO BELONG ###
-### TO A SPECIFIC COMMUNITY. WE DO NOT CONTROL FOR THE NUMBER OF EDGES ADDED FROM EACH COMMUNITY. ###
-########################################################################################################### 
 
-growth_community_seeds <- growth_community(graph = graph_sf, km = no_cycle_infra, col_name = "flow")
-# save
-saveRDS(growth_community_seeds, file = paste0("../data/", chosen_city, "/growth_fun_3.Rds"))
-# read in 
-#growth_community_seeds <- readRDS(paste0("../data/", chosen_city,"/growth_fun_3.Rds"))
 
-# prepare a dataframe for plotting the results
-growth_community_seeds_c <- growth_community_seeds %>% 
[…]

### 2020-11-06 README.Rmd
paper: alg writeup + code cleaning

@@
 
-- Community Detection
+- Lit Review: I can also organize this thematically and trim down **Planning Cycling Networks** substantially. I would then add some literature in:
+  - Calculating Potential Demand
+  - Routing
+  - Road Segment Prioritization
+- Community Detection: Where in the document should this section be?
 - Other Cities
@@
 
-The results show that Manchester can be split into three large communities and one small one (Figure \@ref(fig:communitiesmanchester)).
+The results show that Manchester can be split into four large communities and one small one (Figure \@ref(fig:communitiesmanchester)).
 
@@
 
-Algorithm 2: Egalitarian Growth (Focus on Fair Distribution of Resources)
+## Algorithm 2: Egalitarian Growth (Focus on Fair Distribution of Resources)
 
@@
 
-```{r growthtotal, fig.cap="Comparing Overall Person-Km Satisfied (Manchester)", out.width="45%", fig.show='hold', echo=FALSE}
+The results of the community detection are used to evaluate the algorithms.
+This is done by looking at the *person-km satisfied* as cycling infrastructure is added.
+Person-km is a measure of the total km cycled on a road segment, so it is the product of the number of potential commuters cycling on that road segment ($flow$) and the length of the segment in km ($l$).
+For each road segment, the person-km is equal to $flow * l$.
+In the case of Manchester, Table (**Reference the table**) shows that almost half of the person-km is in community 1 , while only 0.5\% of total person-km on the network is in community 5.
+
+```{r personkmtable, echo=FALSE, message = FALSE}
+
+library(tidyverse)
+library(sf)
+
+table <- readRDS("data/Manchester/graph_with_flows_weighted_communities.RDS")
+# we weigh the flow on each edge by its distance. We can then get how much of the commuter km are satisfied
+table$person_km <- table$flow * table$d
+# get % of total person km on each link
+table$person_km_perc <- (table$person_km / sum(table$person_km)) * 100
+
+table <- table %>% sf::st_drop_geometry() %>% 
+  dplyr::group_by(Community) %>%
+  summarize(`Person-Km (Total)` = round(sum(person_km),0),
+            `Person-Km (%)` = round(sum(person_km_perc),1))
+
+knitr::kable(table,
+             caption = "Total Person-Km in Different Communities (Manchester)")
+
+```
+
+Looking at the person-km satisfied (Figure \@ref(fig:growthtotal)), we see that the incremental addition of cycling infrastructure is better distributed between communities using Algorithm 2; equal distribution of investment results in the gain in % of person km satisfied in each community being inversely correlated with the size of the community.
+In addition, we find that the restrictions imposed by Algorithm 2 on the network growth do not seem to have a noticeable effect on the city-wide \% of person-km satisfied.
+Comparing both algorithms, we can see that Algorithm 1 provides only marginally quicker city-wide gains than Algorithm 2.
+
+```{r growthtotal, fig.cap="Comparing Overall (Dashed) and Community Level Person-Km Satisfied (Manchester)", out.width="45%", fig.show='hold', echo=FALSE}
 knitr::include_graphics(c(
-  "data/Manchester/Plots/Growth_Results/growth_existing_infra_satisfied_km_all_flow_column.png",
-  "data/Manchester/Plots/Growth_Results/growth_community_4_satisfied_km_all_flow_column.png"
+  "data/Manchester/Plots/Growth_Results/growth_utilitarian_satisfied_km_both_flow_column.png",
+  "data/Manchester/Plots/Growth_Results/growth_egalitarian_satisfied_km_both_flow_column.png"
 ))
@@
 
-```{r growthcommunities, fig.cap="Comparing Community Level Person-Km Satisfied (Manchester)", out.width="45%", fig.show='hold', echo=FALSE}
-knitr::include_graphics(c(
-  "data/Manchester/Plots/Growth_Results/growth_existing_infra_satisfied_km_community_flow_column.png",
-  "data/Manchester/Plots/Growth_Results/growth_community_4_satisfied_km_community_flow_column.png"
-))
-```
+Figure \@ref(fig:growth3MapandBar) gives us a geographic representation of the results from Algorithm 2; it shows when each link was added to the solution (first 100km, second 100km, etc).
+We can see that, generally, road segments around cycling infrastructure are prioritized, except for those neighboring cycling infrastructure on the very periphery.
+The first 100km is also spatially distributed across the city, with no apparent bias towards a particular area.
+
+It is also important to understand how the different highway types contribute to the proposed network.
+Figure \@ref(fig:growth3MapandBar) shows that most of the flow will be on residential and tertiary roads, as expected from the weighting profile defined in Table **reference the table**.
 
@@
 knitr::include_graphics(c(
-  "data/Manchester/Plots/Growth_Results/growth_community_4_priority_all_FLOW.png",
-  "data/Manchester/Plots/Growth_Results/growth_community_4_investment_highways_flow.png"
+  "data/Manchester/Plots/Growth_Results/growth_egalitarian_priority_all_FLOW.png",
+  "data/Manchester/Plots/Growth_Results/growth_egalitarian_investment_highways_flow.png"
 ))
@@
 Both Algorithm 1 and 2 start with all existing segregated cycling infrastructure and aim to create an efficient, connected network.
-Figure \@ref(fig:componentsandGCC) compares the performance of both Algorithms in terms of connectivity gains.
-We can see that, for all three cities analyzed, the number of components is gradually reduced as more infrastructure is added (this is true for both Algorithms).
-However these connectivity gains require considerable investment; both algorithms take almost 150km of investment to reduce the number of components by half.
+Figure \@ref(fig:componentsandGCC) shows that both algorithms gradually reduce the number of components as more infrastructure is added, but Algorithm 2 is able to provide better connectivity with less investment.
 
 Consistent growth can also be seen for the size of the Largest Connected Component in the proposed bicycle network (Figure \@ref(fig:componentsandGCC)).
-Again, we can see from the results of the three different cities that there is little difference between both Algorithms.
+Here however, we find that there is little difference between both Algorithms.
 
@@
 knitr::include_graphics(c(
-  "data/Manchester/Plots/Growth_Results/growth_existing_infra_components_number_comparisonManchester.png",
-  "data/Manchester/Plots/Growth_Results/growth_existing_infra_components_gcc_comparisonManchester.png"
+  "data/Manchester/Plots/Growth_Results/growth_util_egal_components_number_comparisonManchester.png",
+  "data/Manchester/Plots/Growth_Results/growth_util_egal_components_gcc_comparisonManchester.png"
 ))

### 2020-11-06 R/_4.0_aggregating_flows.R

@@ -1,6 +1,7 @@
 library(tidyverse)
 library(sf)
 library(dodgr)
+library(geodist)
 
 
 # read in MSOA centroids
@@ -39,7 +40,7 @@ to <- colnames(od_flow_matrix) %>% as.data.frame() %>% left_join(lon_lat, by = c
 # od_flow_matrix_2 <- od_flow_matrix %>% as.data.frame() %>% slice(1:10) %>% select(1:10) %>% as.matrix()
 
 # ROUTE
-# load in graph save in script 2
+# load in graph saved in script 2
 #graph <- readRDS(paste0("../data/",chosen_city,"/city_graph.Rds"))
 streetnet_sc <- readRDS(paste0("../data/",chosen_city,"/unweighted_streetnet.Rds"))
 ######
@@ -47,18 +48,13 @@ streetnet_sc <- readRDS(paste0("../data/",chosen_city,"/unweighted_streetnet.Rds
 # CREATE GRAPHS WITH DIFFERENT WEIGHTING PROFILES. THE WEIGHTING PROFILES ARE EDITED BY
 # DOWNLOADING THE JSON IN -x-dodgr_weight_profiles.R AND EDITING IN A TEXT EDITOR
 
+
 # this graph has weights so that low stress roads (secondary, tertiary) are preferred
 graph_weighted <- weight_streetnet(streetnet_sc, 
                                   wt_profile_file = "../data/weight_profile_weighted.json")
 # all road types are weighted equally here. We use these weights to get the absolute shortest paths
 graph_unweighted <- weight_streetnet(streetnet_sc, 
                                      wt_profile_file = "../data/weight_profile_unweighted.json")
-
-# Here trunk roads are changed from 0.3 to 0.7 (equal to primary). The idea is that segregated 
-# bicycle lanes can easily be built on trunk roads (park lane, euston road), so this helps determine 
-# which trunk routes would be useful for cyclists
-# graph_trunk <- weight_streetnet(streetnet_sc, 
-#                                 wt_profile_file = "../data/weight_profile_trunk.json")
 
 
 # # add flows to road segments # norm_sums isn't a parameter apparently. Tried updating dodgr...
@@ -110,13 +106,10 @@ aggregate_flows <- function(graph, from, to, flows){
 graph_sf_weighted <- aggregate_flows(graph=graph_weighted, from=from, to=to, flows=od_flow_matrix)
 # unweighted graph
[…]

### 2020-11-06 R/_8.1_plot_network_growth.R

@@ -1,250 +1,0 @@
-library(sf)
-library(sfnetworks)
-library(tidyverse)
-library(ggtext)
-library(igraph)
-library(tmap)
-
-#create a directory to store data related to this city (does nothing if directory already exists)
-dir.create(paste0("../data/", chosen_city, "/Plots/Growth_Results"), showWarnings = FALSE)
-
-graph_sf <- readRDS(paste0("../data/", chosen_city, "/graph_with_flows_weighted_communities.RDS"))
-
-# we weigh the flow on each edge by its distance. We can then get how much of the commuter km are satisfied
-graph_sf$person_km <- graph_sf$flow * graph_sf$d
-
-# check how many person-km in each Community 
-graph_sf %>% 
-  st_drop_geometry %>% 
-  group_by(Community) %>% 
-  summarize(total_person_km = sum(person_km) / 1000) %>% 
-  mutate(perc_person_km = (total_person_km / sum(total_person_km) * 100)) 
-
-
-# get percentage contibution of each edge to the network (distance, flow, person_km)
-graph_sf <- graph_sf %>%
-  mutate(perc_dist = (d/sum(d))  *100,      # edge length as % of total network length
-         perc_flow = (flow/sum(flow))  *100, # flow on edge as % of total
-         perc_person_km = (person_km/sum(person_km))  *100) %>% # % of person_km satisfied
-  # get the same % for each community as a % of the community totals
-  group_by(Community) %>%
-  mutate(perc_dist_comm = (d/sum(d)) * 100,  
-         perc_flow_comm = (flow/sum(flow))  *100,
-         perc_person_km_comm = (person_km/sum(person_km))  *100) %>%
-  ungroup()
-
-################
-# get no. of km of road network without cycling infrastructure (to add as a variable to the growth functions)
-cycle_infra <- graph_sf %>% dplyr::filter(cycle_infra == 1) 
-# total length of edges without cycling infrastructure
-no_cycle_infra <- round((sum(graph_sf$d) - sum(cycle_infra$d))  / 1000)
-# subtract 5km just to be safe. I don't want the growth functions to iterate exactly to the network length
-no_cycle_infra <- no_cycle_infra - 5
-################
-
-
-# ----------------------------------- ALGORITHM 1: UTILITARIAN GROWTH ----------------------------------------- #
-
[…]

### 2020-11-06 R/_8.2_plot_network_growth_community.R

@@ -1,331 +1,0 @@
-library(sf)
-library(tidyverse)
-library(ggtext)
-library(tmap)
-
-graph_sf <- readRDS(paste0("../data/", chosen_city, "/graph_with_flows_weighted_communities.RDS"))
-
-# we weigh the flow on each edge by its distance. We can then get how much of the commuter km are satisfied
-graph_sf$person_km <- graph_sf$flow * graph_sf$d
-
-
-# get percentage contribution of each edge to the network (distance, flow, person_km)
-graph_sf <- graph_sf %>%
-  mutate(perc_dist = (d/sum(d))  *100,      # edge length as % of total network length
-         perc_flow = (flow/sum(flow))  *100, # flow on edge as % of total
-         perc_person_km = (person_km/sum(person_km))  *100) %>% # % of person_km satisfied
-  # get the same % for each community as a % of the community totals
-  group_by(Community) %>%
-  mutate(perc_dist_comm = (d/sum(d)) * 100,  
-         perc_flow_comm = (flow/sum(flow))  *100,
-         perc_person_km_comm = (person_km/sum(person_km))  *100) %>%
-  ungroup()
-
-
-
-# ----------------------------------- ALGORITHM 2: EGALITARIAN GROWTH ----------------------------------------- #
-
-###########################################################################################################  
-### THIS FUNCTION IS ALMOST IDENTICAL TO FUNCTION 5 WITH ONE EXCEPTION. WE START WITH ALL EDGES THAT ALREADY ##
-### HAVE CYCLING INFRASTRUCTURE. THESE ARE THE BEGINNING OF THE SOLUTION. TO ENSURE THAT THERE IS AT LEAST ##
-### ONE EDGE FROM EACH COMMUNITY, WE GET THE EDGE WITH THE HIGHEST FLOW IN EACH COMMUNITY AND APPEND IT TO ##
-### THE SOLUTION AT SEQUENCE 0. 
-###########################################################################################################  
-                              #########################################
-# growth_egal function starts with all edges that already have cycling infrastructure. It may be more 
-# practical in real life. It also shows the no of components and the size of the gcc at each iteration
-                              #########################################
-
-# let's grow the network based on the flow column
[…]

### 2020-11-06 README.Rmd
update growth figures

@@
 output:
-  # bookdown::github_document2
-  bookdown::pdf_document2:
-    extra_dependencies: ["subcaption", "longtable"]
+  bookdown::github_document2
+  # bookdown::pdf_document2:
+  #   extra_dependencies: ["subcaption", "longtable"]
 ---

### 2020-11-06 R/_8.1_growth_utilitarian.R
add modified scripts

@@ -1,0 +1,263 @@
+library(sf)
+library(sfnetworks)
+library(tidyverse)
+library(ggtext)
+library(igraph)
+library(tmap)
+
+#create a directory to store data related to this city (does nothing if directory already exists)
+dir.create(paste0("../data/", chosen_city, "/Plots/Growth_Results"), showWarnings = FALSE)
+
+graph_sf <- readRDS(paste0("../data/", chosen_city, "/graph_with_flows_weighted_communities.RDS"))
+
+# we weigh the flow on each edge by its distance. We can then get how much of the commuter km are satisfied
+graph_sf$person_km <- graph_sf$flow * graph_sf$d
+
+# check how many person-km in each Community 
+graph_sf %>% 
+  st_drop_geometry %>% 
+  group_by(Community) %>% 
+  summarize(total_person_km = sum(person_km) / 1000) %>% 
+  mutate(perc_person_km = (total_person_km / sum(total_person_km) * 100)) 
+
+
+# get percentage contibution of each edge to the network (distance, flow, person_km)
+graph_sf <- graph_sf %>%
+  mutate(perc_dist = (d/sum(d))  *100,      # edge length as % of total network length
+         perc_flow = (flow/sum(flow))  *100, # flow on edge as % of total
+         perc_person_km = (person_km/sum(person_km))  *100) %>% # % of person_km satisfied
+  # get the same % for each community as a % of the community totals
+  group_by(Community) %>%
+  mutate(perc_dist_comm = (d/sum(d)) * 100,  
+         perc_flow_comm = (flow/sum(flow))  *100,
+         perc_person_km_comm = (person_km/sum(person_km))  *100) %>%
+  ungroup()
+
+################
+# get no. of km of road network without cycling infrastructure (to add as a variable to the growth functions)
+cycle_infra <- graph_sf %>% dplyr::filter(cycle_infra == 1) 
+# total length of edges without cycling infrastructure
+no_cycle_infra <- round((sum(graph_sf$d) - sum(cycle_infra$d))  / 1000)
+# subtract 5km just to be safe. I don't want the growth functions to iterate exactly to the network length
+no_cycle_infra <- no_cycle_infra - 5
+################
+
+
+# ----------------------------------- ALGORITHM 1: UTILITARIAN GROWTH ----------------------------------------- #
+
[…]

### 2020-11-06 R/_8.2_growth_egalitarian.R

@@ -1,0 +1,345 @@
+library(sf)
+library(tidyverse)
+library(ggtext)
+library(tmap)
+
+graph_sf <- readRDS(paste0("../data/", chosen_city, "/graph_with_flows_weighted_communities.RDS"))
+
+# we weigh the flow on each edge by its distance. We can then get how much of the commuter km are satisfied
+graph_sf$person_km <- graph_sf$flow * graph_sf$d
+
+
+# get percentage contribution of each edge to the network (distance, flow, person_km)
+graph_sf <- graph_sf %>%
+  mutate(perc_dist = (d/sum(d))  *100,      # edge length as % of total network length
+         perc_flow = (flow/sum(flow))  *100, # flow on edge as % of total
+         perc_person_km = (person_km/sum(person_km))  *100) %>% # % of person_km satisfied
+  # get the same % for each community as a % of the community totals
+  group_by(Community) %>%
+  mutate(perc_dist_comm = (d/sum(d)) * 100,  
+         perc_flow_comm = (flow/sum(flow))  *100,
+         perc_person_km_comm = (person_km/sum(person_km))  *100) %>%
+  ungroup()
+
+
+
+# ----------------------------------- ALGORITHM 2: EGALITARIAN GROWTH ----------------------------------------- #
+
+###########################################################################################################  
+### THIS FUNCTION IS ALMOST IDENTICAL TO FUNCTION 5 WITH ONE EXCEPTION. WE START WITH ALL EDGES THAT ALREADY ##
+### HAVE CYCLING INFRASTRUCTURE. THESE ARE THE BEGINNING OF THE SOLUTION. TO ENSURE THAT THERE IS AT LEAST ##
+### ONE EDGE FROM EACH COMMUNITY, WE GET THE EDGE WITH THE HIGHEST FLOW IN EACH COMMUNITY AND APPEND IT TO ##
+### THE SOLUTION AT SEQUENCE 0. 
+###########################################################################################################  
+                              #########################################
+# growth_egal function starts with all edges that already have cycling infrastructure. It may be more 
+# practical in real life. It also shows the no of components and the size of the gcc at each iteration
+                              #########################################
+
+# let's grow the network based on the flow column
[…]

### 2020-11-06 R/reprex/dodgr_bug.R

@@ -1,0 +1,17 @@
+library(dodgr)
+packageVersion("dodgr")
+library(dplyr)
+
+#sf format 
+hampi <- dodgr_streetnet("hampi india")
+hampi_weighted<- weight_streetnet(hampi, wt_profile = "bicycle") 
+# check...all good
+head(hampi_weighted) 
+
+#sc format
+hampi_sc <- dodgr_streetnet_sc("hampi india")
+hampi_sc_weighted <- weight_streetnet(hampi_sc, wt_profile = "bicycle") 
+
+
+remotes::install_github("hypertidy/geodist") 
+remotes::install_github("atfutures/dodgr")

### 2020-11-07 README.Rmd
paper abstract and pot demand

--- 2020-11-06 README.Rmd
@@
 numbersections: true
-#output: rticles::elsevier_article
+abstract: "Understanding the motivators and deterrents to cycling is essential for creating infrastructure that is successful in getting more people to adopt cycling as a mode of transport.
+Therefore, the aim of this study is to develop a methodology that determines where cycling infrastructure should be added by accounting for cyclist preference and UK government policies to create Low Traffic Neighborhoods.
+Distance decay, routing, and network analysis techniques are used to examine where future cycling demand in a city will come from, how such demand should be routed on the street network, and how to  consider fair distribution of cycling infrastructure that is in line with egalitarian principles.
+The city of Manchester is used as a case study, but the methodology is generic enough to be applicable to any city.
+For routing, it is found that deviating from shortest paths to avoid high-stress road segments can result in acceptable travel time increases for cyclists.
+Community detection is used to partition the city and evaluate results from utilitarian to egalitarian algorithms on distributing cycling infrastructure.
+The results show that the egalitarian algorithm provides comparable city-wide results to the utilitarian one.
+\\par\\textbf{Keywords:} cycling network, routing, transport equity"
+# keywords: "cycling, potential demand, routing"
 output:
@@
   bookdown::pdf_document2:
+      toc: false
       extra_dependencies: ["subcaption", "longtable"]
@@
 
-The Propensity to Cycle Tool (PCT) [@lovelace2017propensity] is used to estimate the proportion of cyclists ($\boldsymbol{C_{p}}$) for each MSOA pair should the government achieve its target of doubling cycling by 2025.
-The PCT uses the following logistic regression model to calculate $\boldsymbol{C_{p}}$:
+Using existing cycling demand to inform decisions on where cycling infrastructure should be added reinforces existing cycling patterns and ignores potential cycling demand that could be satisfied by a connected network.
+To avoid this issue, @duthie2014optimization choose to ignore demand completely, and focus on creating a network that connects the entire study area.
+@olmos2020data obtain the distance distribution of cyclists using a smartphone-based bicycle GPS data.
+They then use a rejection-sampling algorithm on the OD data of the study area to match the potential demand distribution to the distribution obtained from GPS data.
+
+For our purposes, we use a logistic regression model to calculate potential cycling demand.
+The model is adopted directly from the Propensity to Cycle Tool (PCT) [@lovelace2017propensity].
+The PCT estimates  the proportion of cyclists ($\boldsymbol{C_{p}}$) for each MSOA pair should the government achieve its target of doubling cycling by 2025.
+The logistic regression model used to calculate $\boldsymbol{C_{p}}$ has the following parameters:
 
@@
 ```
-\noindent where $\boldsymbol{d}$ and $\boldsymbol{s}$ are the distance and slope respectively for the OD pair.
-The authors use square and square-root distance terms \`\`to capture the non-linear impact of distance on the likelihood of cycling", and interaction terms to capture the combined effect of slope and distance [@lovelace2017propensity].
 
-The potential demand calculations show that the current and potential number of cyclists both follow a bell-shaped distribution, with the number of trips peaking around the 3-5km commuting distance and then going back down for longer distances (see Figure \@ref(fig:potdemhistograms).
+\noindent where **d** and **s** are the distance and slope respectively for the OD pair.
+The authors use square and square-root distance terms "to capture the non-linear impact of distance on the likelihood of cycling", and interaction terms to capture the combined effect of slope and distance [@lovelace2017propensity].
+
+The potential demand calculations show that the current and potential number of cyclists both follow a bell-shaped distribution, with the number of trips peaking around the 3-5km commuting distance and then going back down for longer distances (see Figure \@ref(fig:potdemhistograms)).
 

### 2020-11-13 README.Rmd
add pdf
add pdf and rmd
paper intro done

@@
 \\par\\textbf{Keywords:} cycling network, routing, transport equity"
-# keywords: "cycling, potential demand, routing"
+# keywords: "cycling, low-traffic neighborhoods, potential demand, routing"
 output:
@@
 - Community Detection: Where in the document should this section be?
-- Other Cities
+- Should we use other cities for the sake of comparison
+- How to add tables
+- Footnotes are sometimes split between pages. How do I keep them on the same page?
 
@@
 
-Intro ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
-Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
-Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
-Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
-Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
-Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
-Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
-Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
-Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
-Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+The 2015 Paris agreement [@agreement2015paris] was an acknowledgment that fundamental changes to our lives and economies are necessary to reduce the impacts of climate change.
+The transport sector is one of many sectors facing transformation, with innovation and policy measures leading to the introduction of ‘greener’ private vehicles.
+But as demand for transport rises, resulting in increased congestion, public and active modes of transport are gaining more recognition.
+The benefits of the latter extend beyond congestion and the environment, as it promises to help alleviate what is referred to as the pandemic of global inactivity; Physical inactivity is on the rise and has become the 4th highest cause of death globally [@kohl2012pandemic].
+Various studies have documented the association between active transport and lower risk of disease, including cancer and cardiovascular disease [@celis2017association; @jarrett2012effect; @patterson2020associations].
+In the wake of the Covid-19 pandemic, and the resulting reduced capacity of public transport, the UK government has pledged to invest billions of pounds to improve walking and cycling infrastructure across the country.
+While this unprecedented sum is an opportunity to reshape cities in a way that improves the well-being of citizens, it does come with a warning:
 
-# Background
+> "Inadequate cycling infrastructure discourages cycling and wastes public money.
+Much cycling infrastructure in this country is inadequate.
+It reflects a belief, con- scious or otherwise, that hardly anyone cycles, that cycling is unimportant and that cycles must take no meaningful space from more important road users, such as motor vehicles and pedestrians - [@departmentgearchange2020]"
 
-## What Affects the Decision To Cycle
-
+The funding on its own is therefore no guarantee of a change in commuting across the country; it must be used to design adequate cycling infrastructure that is based on motivators and deterrents to cycling.
+A considerable amount of research has been done towards that end.
 Segregated cycling infrastructure^[Segregated cycling infrastructure refers to road space that is allocated to cyclists only, with physical separation to protect cyclists from other modes of transport.] has been shown to increase cycling uptake [@aldred2019impacts; @goodman2014new; @marques2015infrastructure], with the separation from motorized vehicles being key [@winters2011motivators].
@@
 Cyclists prefer cohesive infrastructure, particularly when cycling on arterial roads with high levels of motorized traffic [@stinson2003commuter], and the lack of well-connected cycling infrastructure is one of the main obstacles to increasing cycling uptake [@caulfield2012determining].
-While direct and cohesive cycling networks have been shown to positively impact cycling rates, density^[making an area's bicycle network denser means adding more cycling routes in the area and thereby giving cyclists more route options] of the cycling network is also vital [@schoner2014missing]].
+While direct and cohesive cycling networks have been shown to positively impact cycling rates, density^[making an area's bicycle network denser means adding more cycling routes in the area and thereby giving cyclists more route options] of the cycling network is also vital [@schoner2014missing].
 
-## Planning Cycling Networks
-
+As noted by @buehler2016bikeway, it is only recently that research has pivoted from focusing on localized, street or intersection level impact towards studying the whole cycling network.
+They emphasize the promise of this shift in capturing the network-wide effect of street-level interventions.
 *Optimization* techniques have been used to propose improvements to cycling networks.
@@
 
-\vspace{0.3cm}
-
 These problem formulations do not explicitly solve for continuity, which is dealt with using a either (a) a constraint specifying that each link with a bike lane should be connected to at least one destination [@mesbah2012bilevel], (b) a constraint on deviation from shortest paths [@duthie2014optimization], or (c) a discontinuity penalty [@mauttone2017bicycle].
@@
 After routing the cycling demand onto the network links, they use percolation theory to filter out the links based on the aggregate flow^[*flow* is used throughout this research to refer to the cycling demand when it is routed onto the road network.
-The flow on any road segment is the cumulative demand on it, resulting from cyclists commuting between various OD pairs] passing through them.
-They vary the flow threshold for filtering to identify the minimum flow at which the whole city is connected by a giant component.
+The flow on any road segment is the cumulative demand on it, resulting from cyclists commuting between various OD pairs] passing through them, varying the flow threshold for filtering to identify the minimum flow at which the whole city is connected by a giant component.
 The results show a cycling network that connects the entire city, and subtracting links intersecting with current cycling infrastructure identifies links proposed for intervention.
-
-\vspace{0.3cm}
 
@@
 
-## Underlying Ethical Principles
-
-The methodologies in Section \@ref(planning-cycling-networks) are underpinned by different ethical principles, even though these principles are not explicitly acknowledged by the authors.
+All of these network-level methodologies are underpinned by different ethical principles, even though these principles are not explicitly acknowledged by the authors.
 This is important since different ethical principles constitute different problem formulations and targets.
@@
 They highlight accessibility as a cornerstone of distributive justice, and contend that policies should aim to distribute investments in a way that minimizes spatial variations in accessibility.
-This research attempts to provide a methodology that is grounded in egalitarian principles.
+This research attempts to compare the two principles, and in doing so determine whether a methodology formulated based on an egalitarian approach can be feasible in designing a cycling network that aligns with motivators and deterrents to cycling.
 **Write some more here**
@@
 
-The analysis is heavily dependant on Origin-Destination census data (commuter data).
+The analysis is heavily dependent on commuter data^[*commuter data* refers to publicly available Origin-Destination (OD) data, normally obtained from a census.
+This study uses data from the UK census [@ONS2011flowdata], which contains aggregate statistics on number of commuters between MSOA pairs, by mode of travel.].
 Commuter data in the UK is publicly available at the Middle layer Super Output Area (MSOA) level; the average MSOA has a population of 8209 [@ofn2018population].
@@
 
-# Conclusions
+<!-- # Conclusions -->
 

### 2020-11-14 README.Rmd
added latex tables

@@
       toc: false
-      extra_dependencies: ["subcaption", "longtable"]
+      extra_dependencies: ["subcaption", "longtable", "float", "multirow", "placeins", "multicol"]
 ---
@@
 
-- Lit Review: I can also organize this thematically and trim down **Planning Cycling Networks** substantially. I would then add some literature in:
-  - Calculating Potential Demand
-  - Routing
-  - Road Segment Prioritization
+- Intro: trim down substantially then add some literature in: Calculating Potential Demand / Routing / Road Segment Prioritization
 - Community Detection: Where in the document should this section be?
 - Should we use other cities for the sake of comparison
-- How to add tables
+- How to add tables properly
 - Footnotes are sometimes split between pages. How do I keep them on the same page?
@@
 The transport sector is one of many sectors facing transformation, with innovation and policy measures leading to the introduction of ‘greener’ private vehicles.
-But as demand for transport rises, resulting in increased congestion, public and active modes of transport are gaining more recognition.
-The benefits of the latter extend beyond congestion and the environment, as it promises to help alleviate what is referred to as the pandemic of global inactivity; Physical inactivity is on the rise and has become the 4th highest cause of death globally [@kohl2012pandemic].
+But as demand for transport rises, resulting in increased congestion, public and active modes of transport are also gaining more recognition.
+The benefits of the latter extend beyond congestion and the environment, as it promises to help alleviate what is referred to as the pandemic of global inactivity; physical inactivity is on the rise and has become the 4th highest cause of death globally [@kohl2012pandemic].
 Various studies have documented the association between active transport and lower risk of disease, including cancer and cardiovascular disease [@celis2017association; @jarrett2012effect; @patterson2020associations].
@@
 The funding on its own is therefore no guarantee of a change in commuting across the country; it must be used to design adequate cycling infrastructure that is based on motivators and deterrents to cycling.
+
 A considerable amount of research has been done towards that end.
@@
 
-As noted by @buehler2016bikeway, it is only recently that research has pivoted from focusing on localized, street or intersection level impact towards studying the whole cycling network.
+As noted by @buehler2016bikeway, it is only recently that research has pivoted from focusing on localized, street or intersection level impact towards studying the cycling network as a whole.
 They emphasize the promise of this shift in capturing the network-wide effect of street-level interventions.
@@
 The problem accounts for the effect of cycling lanes on car traffic, and attempts to maximize utilization of said lanes with minimal impact on car travel times.
-To improve cohesion of the suggested network, a constraint is added so that each link with a bike lane should be connected to at least one destination.
+To improve cohesion of the suggested network, a constraint is added so that each link^[*link* refers to a road segment throughout this research] with a bike lane should be connected to at least one destination.
 Car usage is not considered by @mauttone2017bicycle, who develop an optimization framework that aims to minimize the total user cost of cycling on the network.
@@
 To solve for continuity, the graph-theoretic concept of *connected components*, has been used.
-@natera2019data study the existing cycling network in terms of its disconnected components and introduce two different algorithms to connect these components by their most critical links^[*link* refers to a road segment throughout this research] and, in doing so, measure the size of the growth of the largest connected component as a function of the kilometers of network added.
+@natera2019data study the existing cycling network in terms of its disconnected components and introduce two different algorithms to connect these components by their most critical links and, in doing so, measure the size of the growth of the largest connected component as a function of the kilometers of network added.
 They observe that small investments at strategic points have a large impact on connectivity in most cases.
@@
 @bao2017planning adopt a similar methodology, but use hierarchical clustering to specify the desired number of clusters.
-They use a greedy network expansion algorithm, where the link with the highest benefit/cost ratio in each cluster is selected, and the network is grown by adding neighboring links to the solution until a budget limit is met.
+They use a greedy network expansion algorithm, where the link with the highest benefit-cost ratio in each cluster is selected, and the network is grown by adding neighboring links to the solution until a budget limit is met.
 The benefit is the flow on the link, and each link is assigned a cost based on current road conditions.
@@
 To avoid this issue, @duthie2014optimization choose to ignore demand completely, and focus on creating a network that connects the entire study area.
-@olmos2020data obtain the distance distribution of cyclists using a smartphone-based bicycle GPS data.
-They then use a rejection-sampling algorithm on the OD data of the study area to match the potential demand distribution to the distribution obtained from GPS data.
+@olmos2020data obtain the distance distribution of cyclists using a smartphone-based bicycle GPS data, and then use a rejection-sampling algorithm on the OD data of the study area to match the potential demand distribution to the distribution obtained from GPS data.
 
@@
 
-```{r desirefacetcycling, fig.cap="Current and Potential Cycling Demand", out.width="80%", fig.show='hold', echo=FALSE}
+```{r desirefacetcycling, fig.cap="Current and Potential Cycling Demand", out.width="75%", fig.show='hold', echo=FALSE}
 knitr::include_graphics("data/Manchester/Plots/desire_facet_cycling.png")
@@
 
-The next step is to route the potential cycling demand ($\boldsymbol{C_{p}}$) between all OD pairs onto the road network.
+The next step is to route the potential cycling demand ($\boldsymbol{C_{p}}$) between all OD pairs onto the road network. **TALK ABOUT WEIGHTED ROUTING IN OTHER STUDIES **
+<!-- This expands on the work of @mauttone2017bicycle, by going beyond simply favoring roads with existing cycling infrastructure to creating a hierarchy of road preference. -->
 
-**ADD TABLE**
+\begin{longtable}{lll}
+\caption{OSM Road Types}
+\label{table:osmroadtypes}\\
+\textbf{OSM Road Type } & \textbf{Description}                                                                                                                                            & \textbf{UK Equivalent}                                                                                                   \endfirsthead
+\hline
+Motorway                & \begin{tabular}[c]{@{}l@{}}Road open to high-speed \\vehicles only\end{tabular}                                                                                 & Motorway                                                                                                                 \\*
+\hline
+Trunk                   & \multirow{5}{*}{\begin{tabular}[c]{@{}l@{}}Roads that don't \\meet motorway criteria, \\in descending order \\of importance and \\through-traffic\end{tabular}} & A-Roads with Primary Status                                                                                              \\*
+\cline{3-3}
+Primary                 &                                                                                                                                                                 & A-Roads with Non-Primary Status                                                                                          \\*
+\cline{3-3}
+Secondary               &                                                                                                                                                                 & B-roads                                                                                                                  \\*
+\cline{3-3}
+Tertiary                &                                                                                                                                                                 & \begin{tabular}[c]{@{}l@{}}Classified unnumbered roads \\OR unclassified busy through roads\end{tabular}                 \\*
+\cline{3-3}
+Unclassified            &                                                                                                                                                                 & \multirow{2}{*}{\begin{tabular}[c]{@{}l@{}}Unclassified (Intended for \\local traffic - 60\% of UK roads)\end{tabular}}  \\*
+\cline{1-2}
+Residential             & Function is purely residential                                                                                                                                  &                                                                                                                          \\
+\hline
+Service                 & \begin{tabular}[c]{@{}l@{}}Road that provides \\access to a facility\end{tabular}                                                                               &                                                                                                                          \\
+\hline
+Track                   & \begin{tabular}[c]{@{}l@{}}Unpaved - suitable for \\two track vehicles (mostly rural)\end{tabular}                                                              &                                                                                                                          \\
+\hline
+Cycleway                & \begin{tabular}[c]{@{}l@{}}Designated cycleway - \\open to cyclists only\end{tabular}                                                                           &                                                                                                                          \\
+\hline
+Path                    & \begin{tabular}[c]{@{}l@{}}Unpaved - open to \\non-motorized traffic only\end{tabular}                                                                          &                                                                                                                          \\
+\hline
+\end{longtable}
 
@@
 **Low-Traffic Neighborhoods**: The UK Department for Transport is allocating funding to local authorities to invest in Active Transport, partially through the creation of LTNs [@departmentgearchange2020].
-This includes closing off residential streets to motorized traffic
+This includes closing off residential streets to motorized traffic.
 3.
@@
 
-The weighting profiles are therefore adjusted to favor less-stressful streets (based on information from Table **REFERENCE THE TABLE**), and roads with existing cycling infrastructure.
+The weighting profiles are therefore adjusted to favor less stressful streets (based on information from Table  \@ref(table:osmroadtypes)), and roads with existing cycling infrastructure.
 This is also in line with the creation of LTNs, as residential streets are those where motorized traffic is most likely to be banned in the creation of LTNs.
 
-**ADD TABLE**
+<!-- **ADD TABLE - THIS IS BASIC** -->
+<!-- ```{r, echo=FALSE, message = FALSE} -->
+<!-- weight_profiles <- readxl::read_excel("Paper/paper_tables.xlsx", sheet = "Weighting Profile") -->
+
+<!-- knitr::kable(weight_profiles, -->
+<!--              caption = "Weighting Profiles") -->
+<!-- ``` -->
+\begin{longtable}{lrrr}
+\caption{Weighting Profiles}
+\label{table:weightprofiles}\\
+\multirow{2}{*}{\textbf{OSM Road Type}} & \multicolumn{3}{c}{\textbf{Weighting Profile}}                                                 \\*
+                                        & \multicolumn{1}{l}{\textit{Unweighted}} & \multicolumn{1}{l}{\textit{Weighted}} & \multicolumn{1}{l}{\textit{Weighted\_2}}  \endfirsthead
+Cycleway                                & 1          & 1                                     & 1                                         \\
+Path                                    & 1          & 0.9                                   & 0.9                                       \\
+Residential                             & 1          & 0.9                                   & 0.9                                       \\
+Service                                 & 1          & 0.9                                   & 0.9                                       \\
+Tertiary                                & 1          & 0.9                                   & 0.9                                       \\
+Track                                   & 1          & 0.9                                   & 0.9                                       \\
+Unclassified                            & 1          & 0.9                                   & 0.9                                       \\
+Secondary                               & 1          & 0.8                                   & 0.8                                       \\
+Primary                                 & 1          & 0.7                                   & \textcolor{red}{0}                        \\
+Trunk                                   & 1          & 0.6                                   & \textcolor{red}{0}                        \\
+Motorway                                & 1          & \textcolor{red}{0}                    & \textcolor{red}{0}                       
+\end{longtable}
 
@@
 From Figure \@ref(fig:flowsfacetunweighted) we can see that trunk and primary roads are the most efficient means of traversing the road network of Manchester.
-Both of these road types are classified as Primary A roads according to the UK Department for Transport's road classification (**Reference the table**), and are therefore part of the Primary Route Network (PRN) [@department2012guidance].
+Both of these road types are classified as Primary A roads according to the UK Department for Transport's road classification (Table \@ref(table:osmroadtypes)), and are therefore part of the Primary Route Network (PRN) [@department2012guidance].
 The PRN has the widest, most direct roads on the network, and carries most of the through traffic.
@@
 Figure \@ref(fig:flowsfacetweighted) shows that routing on the weighted network significantly reduces flow on the trunk and primary roads, but does not eliminate it completely.
-This is intentional, as the impedance on these roads is only slightly higher than remaining road types (See Table **Reference the table**).
+This is intentional, as the impedance on these roads is only slightly higher than remaining road types (See Table \@ref(table:weightprofiles)).
 Potential cycling demand is only routed on these roads if there are no routes through other roads that offer comparable directness.
@@
 Banning cycling flow completely on trunk and primary roads may result in excessively circuitous paths, as seen in Figure \@ref(fig:boxplotcircuity).
-When routing using the weighting profile in Table **Reference Table**, we see that shortest paths increase by less than 5\% on average from unweighted shortest paths, with the largest increases still below 30\%.
-When routing on primary and trunk roads is banned, the average increase relative to unweighted shortest paths rises to 10\%, with certain locations experiencing more significant negative effects on accessibility.
+When routing using the weighting profile in Table \@ref(table:weightprofiles), we see that shortest paths increase by less than 5\% on average from unweighted shortest paths, with the largest increases still below 30\%.
+When routing on primary and trunk roads is banned (weighted_2 profile in Table \@ref(table:weightprofiles)), the average increase relative to unweighted shortest paths rises to 10\%, with certain locations experiencing more significant negative effects on accessibility.
 
@@
 
-Given that cyclists will only deviate from shortest paths by a certain amount to access better cycling infrastructure (as explained in Section \@ref(what-affects-the-decision-to-cycle), allowing flow on some stretches of trunk and primary roads is necessary to insure cycling uptake and equitable access to cycling infrastructure.
+Given that cyclists will only deviate from shortest paths by a certain amount to access better cycling infrastructure (as explained in Section \@ref(introduction), allowing flow on some stretches of trunk and primary roads is necessary to insure cycling uptake and equitable access to cycling infrastructure.
 In its new vision for walking and cycling, the Department for Transport acknowledges that minimal segregated stretches of bicycle lanes on main roads will be necessary to avoid circuitous cycling networks [@departmentgearchange2020].
@@
 It creates a hierarchy of road preference that is grounded in cyclist preference and government plans to create LTNs.
-Sensitivity analysis should be done to determine an optimal weighting profile, but given the variation in city road networks , <!-- (Figure \ref{fig:perc_person-km}) -->, these would probably require calibration to the specific city.
+Sensitivity analysis should be done to determine an optimal weighting profile, but given the variation in city road networks<!-- (Figure \ref{fig:perc_person-km}) -->, these would probably require calibration to the specific city.
 More accurate routing could be carried out given the availability of road-level data.
@@
 
-```{r cyclinghistmanc, fig.cap="Cycling Commuting Distance - Manchester (2011 Census Data)", out.width="45%", fig.show='hold', echo=FALSE}
+```{r cyclinghistmanc, fig.cap="Cycling Commuting Distance - Manchester (2011 Census Data)", out.width="30%", fig.show='hold', echo=FALSE}
 knitr::include_graphics("data/Manchester/Plots/histogram_distance_cycling.png")
@@
 This cumulative demand (referred to as *flow* is then used as a basis for determining where best to invest in segregated cycling infrastructure.
-In doing so, we must account for the motivations and deterrents for cycling identified in Section \@ref(what-affects-the-decision-to-cycle), namely direct and well connected routes.
+In doing so, we must account for the motivations and deterrents for cycling identified in Section \@ref(introduction), namely direct and well connected routes.
 
@@
 Both utilize existing infrastructure from the beginning and allow us to compare a solution that focuses on utilitarianism to one that focuses on egalitarianism.
-In both algorithms, links^[*link* refers to a road segment] are selected iteratively and the iteration at which each link is added to the solution is recorded.
+In both algorithms, links are selected iteratively and the iteration at which each link is added to the solution is recorded.
 Investments in cycling infrastructure can be limited by budget constraints, so it can be useful to see where best to allocate a defined length of segregated infrastructure.
@@
 This algorithm ensures that the resulting network is connected.
-It also satisfies the directness criteria, since links on the weighted shortest paths are those that have the highest flow passing through them (this is a result of the routing in Section \@ref(what-affects-the-decision-to-cycle).
+It also satisfies the directness criteria, since links on the weighted shortest paths are those that have the highest flow passing through them (this is a result of the routing in Section \@ref(introduction).
 
@@
 The first algorithm focuses on connectivity and directness, but not on fairly distributing investment.
-The latter is not a requirement for increasing cycling uptake, but it is fundamental for spatial equity, as explained in Section \@ref(underlying-ethical-principles).
+The latter is not a requirement for increasing cycling uptake, but it is fundamental for spatial equity, as explained in Section \@ref(introduction).
 This algorithm incorporates the ideal of fair distribution by using community detection to partition the road network.
@@
 It is also important to understand how the different highway types contribute to the proposed network.
-Figure \@ref(fig:growth3MapandBar) shows that most of the flow will be on residential and tertiary roads, as expected from the weighting profile defined in Table **reference the table**.
+Figure \@ref(fig:growth3MapandBar) shows that most of the flow will be on residential and tertiary roads, as expected from the weighting profile defined in Table \@ref(table:weightprofiles).
 
@@
 
-<!-- # Conclusions -->
+# Conclusions
+
+The research set out to provide a methodology for designing a cycling network that respects both the needs of the users and the ambitions of the government.
+The methodology can be used as a starting point for introducing a direct, connected, and low-stress network for any city by leveraging available open data.
+The only data necessary for reproducing it is the road network (from OSM), its topography (from satellite imagery), and commuter data (from the census).
+It should be noted that the recommendations are only as good as the granularity of the available data.
+A core part of the methodology is determining where there is high potential for cycling, and using this as a basis for recommendations on road space reallocation in order to unlock potential cycling demand.
+A routing engine was used to study the effects of limiting access on different road types, and it was found that reducing cyclist flow on roads with high through-traffic resulted in acceptable increases in commuting distances.
+As a result, a hierarchy of road preference was used to route potential cycling demand.
+Algorithms were developed to determine where investments in cycling infrastructure should be prioritized.
+These algorithms, based on connectivity and egalitarian principles of resource distribution, ensured that whatever the level of investment, the resulting cycling network would improve connectivity of existing cycling infrastructure.
+
+The research hopes to aid in the current shift towards active transport that is being promoted by the UK government.
+The success of such a shift is contingent on the appeal of the cycling network for prospective cyclists, a fact that was used as the basis for the methodology outlined here.
 

### 2020-11-16 README.Rmd
update pdf and markdown files

@@
 
-- Intro: trim down substantially then add some literature in: Calculating Potential Demand / Routing / Road Segment Prioritization
-- Community Detection: Where in the document should this section be?
-- Should we use other cities for the sake of comparison
-- How to add tables properly
+- Intro: possible trim down substantially then add some literature in (a) Calculating Potential Demand / (b) Routing / (c) Road Segment Prioritization
+- Community Detection: Where in the document should this section be? It is now before Road Segment Prioritization, but I can merge them
+- Should we use other cities for the sake of comparison (compare communities, road types etc)
+- How to add tables properly (they are now added using latex code which only shows up when outputting to pdf)
 - Footnotes are sometimes split between pages. How do I keep them on the same page?
+- What is the best way to write pseudocode (Steps for Algorithm 1 and 2)
 
@@
 
-## Algorithm 1: Utilitarian Growth
+## Algorithm 1: Utilitarian Expansion
 
@@
 
-## Algorithm 2: Egalitarian Growth (Focus on Fair Distribution of Resources)
+## Algorithm 2: Egalitarian Expansion (Focus on Fair Distribution of Resources)
 
@@
 Looking at the person-km satisfied (Figure \@ref(fig:growthtotal)), we see that the incremental addition of cycling infrastructure is better distributed between communities using Algorithm 2; equal distribution of investment results in the gain in % of person km satisfied in each community being inversely correlated with the size of the community.
-In addition, we find that the restrictions imposed by Algorithm 2 on the network growth do not seem to have a noticeable effect on the city-wide \% of person-km satisfied.
+In addition, we find that the restrictions imposed by Algorithm 2 on the network expansion do not seem to have a noticeable effect on the city-wide \% of person-km satisfied.
 Comparing both algorithms, we can see that Algorithm 1 provides only marginally quicker city-wide gains than Algorithm 2.
@@
 
-The research set out to provide a methodology for designing a cycling network that respects both the needs of the users and the ambitions of the government.
+The research provides a methodology for designing a cycling network that respects both the needs of the users and the ambitions of the government.
 The methodology can be used as a starting point for introducing a direct, connected, and low-stress network for any city by leveraging available open data.
-The only data necessary for reproducing it is the road network (from OSM), its topography (from satellite imagery), and commuter data (from the census).
+The only data necessary for reproducing it is the road network (from OSM), its topography (from satellite imagery), and commuter data (from the national census).
 It should be noted that the recommendations are only as good as the granularity of the available data.

### 2020-11-18 README.Rmd
edited Routing section in paper

@@
 
-The next step is to route the potential cycling demand ($\boldsymbol{C_{p}}$) between all OD pairs onto the road network. **TALK ABOUT WEIGHTED ROUTING IN OTHER STUDIES **
+The next step is to route the potential cycling demand ($\boldsymbol{C_{p}}$) between all OD pairs onto the road network. 
 <!-- This expands on the work of @mauttone2017bicycle, by going beyond simply favoring roads with existing cycling infrastructure to creating a hierarchy of road preference. -->
@@
 
-The weighting profiles are therefore adjusted to favor less stressful streets (based on information from Table  \@ref(table:osmroadtypes)), and roads with existing cycling infrastructure.
-This is also in line with the creation of LTNs, as residential streets are those where motorized traffic is most likely to be banned in the creation of LTNs.
+The above points are accounted for by using a weighted road network for routing.
+This has previously been done by multiplying all road segments without cycling infrastructure by an impedance factor [@mauttone2017bicycle], or by assigning a weight to the road segment proportional to the investment cost of bringing it to an acceptable level of stress for cycling  [@duthie2014optimization].
+Our approach is similar to the latter, as we create a weighting profile that is adjusted to favor less stressful streets (based on information from Table  \@ref(table:osmroadtypes)), and roads with existing cycling infrastructure.
+We believe this to be more practical  than the approach adopted by @mauttone2017bicycle, as it goes beyond simply favoring roads with existing cycling infrastructure to creating a hierarchy of road preference based on perceived stress levels.
+The approach is also in line with the creation of LTNs, as residential streets are those where motorized traffic is most likely to be banned in the creation of LTNs.
 
@@
 
-Routing potential cycling demand on a weighted network is more in line with government policy to create LTNs.
 Figure \@ref(fig:flowsfacetweighted) shows that routing on the weighted network significantly reduces flow on the trunk and primary roads, but does not eliminate it completely.
@@
 
-Given that cyclists will only deviate from shortest paths by a certain amount to access better cycling infrastructure (as explained in Section \@ref(introduction), allowing flow on some stretches of trunk and primary roads is necessary to insure cycling uptake and equitable access to cycling infrastructure.
+Given that cyclists will only deviate from shortest paths by a certain amount to access better cycling infrastructure (as explained in Section \@ref(introduction)), allowing flow on some stretches of trunk and primary roads is necessary to insure cycling uptake and equitable access to cycling infrastructure.
 In its new vision for walking and cycling, the Department for Transport acknowledges that minimal segregated stretches of bicycle lanes on main roads will be necessary to avoid circuitous cycling networks [@departmentgearchange2020].

### 2020-11-20 README.Rmd
updated draft paper

@@
 Community detection is used to partition the city and evaluate results from utilitarian to egalitarian algorithms on distributing cycling infrastructure.
-The results show that the egalitarian algorithm provides comparable city-wide results to the utilitarian one.
-\\par\\textbf{Keywords:} cycling network, routing, transport equity"
+The results show that the egalitarian algorithm provides comparable city-wide results to the utilitarian one, demostrating it's potential suitability in the planning of cycling networks.
+\\par\\textbf{Keywords:} cycling networks, routing, transport equity"
 # keywords: "cycling, low-traffic neighborhoods, potential demand, routing"
@@
 
-- Intro: possible trim down substantially then add some literature in (a) Calculating Potential Demand / (b) Routing / (c) Road Segment Prioritization
+- Intro: possible trim down substantially then add some literature in (a) Calculating Potential Demand [DONE]/ (b) Routing [DONE]/ (c) Road Segment Prioritization
 - Community Detection: Where in the document should this section be? It is now before Road Segment Prioritization, but I can merge them
@@
 This research attempts to compare the two principles, and in doing so determine whether a methodology formulated based on an egalitarian approach can be feasible in designing a cycling network that aligns with motivators and deterrents to cycling.
-**Write some more here**
+**Write some more here?**
 
@@
 Using existing cycling demand to inform decisions on where cycling infrastructure should be added reinforces existing cycling patterns and ignores potential cycling demand that could be satisfied by a connected network.
-To avoid this issue, @duthie2014optimization choose to ignore demand completely, and focus on creating a network that connects the entire study area.
+To avoid this issue, @duthie2014optimization choose to ignore existing demand completely, and focus on creating a network that connects the entire study area.
 @olmos2020data obtain the distance distribution of cyclists using a smartphone-based bicycle GPS data, and then use a rejection-sampling algorithm on the OD data of the study area to match the potential demand distribution to the distribution obtained from GPS data.
@@
 
-The potential demand calculations show that the current and potential number of cyclists both follow a bell-shaped distribution, with the number of trips peaking around the 3-5km commuting distance and then going back down for longer distances (see Figure \@ref(fig:potdemhistograms)).
+The potential demand calculations show that the current and potential number of cyclists both follow a bell-shaped distribution, with the number of trips peaking around the 3-4km commuting distance and then going back down for longer distances (see Figure \@ref(fig:potdemhistograms)).
+
+```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE}
+# this doesn't work because R project for code is in a different path to R project for paper (relative paths inside code won't work)
+# source("R/_1.0_get_flow_data.R")
+```
 
@@
 Banning cycling flow completely on trunk and primary roads may result in excessively circuitous paths, as seen in Figure \@ref(fig:boxplotcircuity).
-When routing using the weighting profile in Table \@ref(table:weightprofiles), we see that shortest paths increase by less than 5\% on average from unweighted shortest paths, with the largest increases still below 30\%.
-When routing on primary and trunk roads is banned (weighted_2 profile in Table \@ref(table:weightprofiles)), the average increase relative to unweighted shortest paths rises to 10\%, with certain locations experiencing more significant negative effects on accessibility.
+When routing using the *weighted* weighting profile in Table \@ref(table:weightprofiles), we see that shortest paths increase by less than 5\% on average from unweighted shortest paths, with the largest increases still below 30\%.
+When routing on primary and trunk roads is banned (*weighted_2* profile in Table \@ref(table:weightprofiles)), the average increase relative to unweighted shortest paths rises to 10\%, with certain locations experiencing more significant negative effects on accessibility.
 
@@
 @goodman2019scenarios show that if dutch levels of cycling were achieved in the UK, the \% of children cycling to school could increase from 1.8\% to 41\%.
-
 In their typology of cyclists, @dill2013four found that a majority of people who say they would never cycle had never cycled to school, whereas confident cyclists were those most likely to have cycled to school.
@@
 After routing the potential cycling demand onto the road network using weighted shortest paths, we have estimates for the cumulative potential cycling demand passing through all road segments.
-This cumulative demand (referred to as *flow* is then used as a basis for determining where best to invest in segregated cycling infrastructure.
+This cumulative demand (referred to as *flow*) is then used as a basis for determining where best to invest in segregated cycling infrastructure.
 In doing so, we must account for the motivations and deterrents for cycling identified in Section \@ref(introduction), namely direct and well connected routes.

### 2020-11-20 R/_1.0_get_flow_data.R

@@ -80,3 +80,4 @@ write_csv(flows_city, path = paste0("../data/", chosen_city, "/flows_city.csv"))
 
 # remove variables from global environment
 rm(flows, flows_city, flows_internal, cycle_mode_share, no_msoas)
+

### 2020-11-20 README.Rmd
Update readme, reproduce
Changes for reproducibility

@@
 <!--   - Source: -->
-<!--     <http://geoportal.statistics.gov.uk/datasets/826dc85fb600440889480f4d9dbb1a24_0> -->
+<!--     <http://geoportal.statistics.gov.datasets/826dc85fb600440889480f4d9dbb1a24_0> -->
 <!--   - Location in Repo: data-raw/MSOA\_2011\_Boundaries/\[Add files here\] -->
@@
 ```{r desirefacetcycling, fig.cap="Current and Potential Cycling Demand", out.width="75%", fig.show='hold', echo=FALSE}
-knitr::include_graphics("data/Manchester/Plots/desire_facet_cycling.png")
+knitr::include_graphics("data/Manchester/Plodesire_facet_cycling.png")
 ```
@@
 
-table <- readRDS("data/Manchester/graph_with_flows_weighted_communities.RDS")
+table <- readRDS("data/Manchester/graph_with_flows_weighted_communities.Rds")
 # we weigh the flow on each edge by its distance. We can then get how much of the commuter km are satisfied

### 2020-11-20 R/_1.0_get_flow_data.R

@@ -5,7 +5,7 @@ library(ggplot2)
 library(ggrepel)
 
 # this is a lookup table matching MSOAs to major towns and cities
-city_names <- read_csv('../data-raw/Middle_Layer_Super_Output_Area__2011__to_Major_Towns_and_Cities__December_2015__Lookup_in_England_and_Wales.csv') 
+city_names <- read_csv('data-raw/Middle_Layer_Super_Output_Area__2011__to_Major_Towns_and_Cities__December_2015__Lookup_in_England_and_Wales.csv') 
 # change column name
 city_names <- city_names %>% rename(city = TCITY15NM)
 
@@ -18,14 +18,22 @@ no_msoas <- city_names %>% dplyr::group_by(city) %>% dplyr::tally()
 ##### CHOOSE YOU CITY 
 chosen_city <- "Manchester"
 #create a directory to store data related to this city (does nothing if directory already exists)
-dir.create(paste0("../data/", chosen_city), showWarnings = FALSE)
+dir.create(paste0("data/", chosen_city), showWarnings = FALSE)
 # create sub-directory to save plots as well
-dir.create(paste0("../data/", chosen_city,"/Plots"), showWarnings = FALSE)
+dir.create(paste0("data/", chosen_city,"/Plots"), showWarnings = FALSE)
 
 ##### CHOOSE YOU CITY 
 
 # flow data from the 2011 census https://www.nomisweb.co.uk/census/2011/bulk/rOD1
-flows <- read_csv('../data-raw/flow_data.csv')
+u = "https://www.nomisweb.co.uk/output/census/2011/wu03ew_msoa.zip"
+f = 'data-raw/flow_data.csv'
+f_zip = file.path(tempdir(), basename(u))
+if(!file.exists(f)) {
+  download.file(url = u, destfile = f_zip)
+  unzip(zipfile = f_zip, exdir = "data-raw")
+  file.rename("data-raw/wu03ew_msoa.csv", f)
+}
+flows <- read_csv(f)
 
 ###############
 # MERGING NAMES WITH FLOW DATA (TO GET INTERNAL FLOWS IN ANY CITY)
@@ -60,7 +68,7 @@ cycle_mode_share %>% #filter(city_origin != 'London') %>%
   scale_x_continuous(trans='log10', labels = scales::comma) +
   theme_minimal()
 
-ggsave("../data/uk_cities_mode_share.png")
+ggsave("data/uk_cities_mode_share.png")
 
 # Subset flows to keep only those that are within a specific city
[…]

### 2020-11-20 R/_2.0_distance_and_elevation.R

@@ -17,16 +17,16 @@ msoas_in_city <- function(df, city_name) {
 # get MSOAs in chosen_city. chosen_city is chosen in script 1
 msoas_city <- msoas_in_city(city_names, chosen_city)
 #save it for plotting later
-write_csv(msoas_city, path = paste0("../data/",chosen_city,"/msoa_codes_city.csv"))
+write_csv(msoas_city, path = paste0("data/",chosen_city,"/msoa_codes_city.csv"))
 
 
 # get geometry of msoas_city. We nned this geometry for plotting later
-city_geom <- sf::st_read("../data-raw/MSOA_2011_Boundaries/Middle_Layer_Super_Output_Areas__December_2011__Boundaries.shp") %>%
+city_geom <- sf::st_read("data-raw/MSOA_2011_Boundaries/Middle_Layer_Super_Output_Areas__December_2011__Boundaries.shp") %>%
   st_transform(4326)
 # filter only MSOAs in the msoas_city df
 city_geom <- city_geom %>% dplyr::filter(msoa11cd %in% msoas_city$MSOA11CD)
 # save to load in when plotting
-st_write(city_geom, paste0("../data/",chosen_city,"/msoas_geometry.shp"), append=FALSE)
+st_write(city_geom, paste0("data/",chosen_city,"/msoas_geometry.shp"), append=FALSE)
 
 
 # get population weighted centroids from pct and change crs (default is northing)
@@ -92,11 +92,11 @@ msoa_centroids_snapped %>%
   cbind(st_coordinates(.)) %>%  #split geometry into X and Y columns
   rename(lon = X, lat = Y) %>%  # rename x and y
   dplyr::select(-c(msoa11nm)) %>% 
-  #st_write("../data/alt_city/msoa_lon_lat.shp", append=FALSE) # save as shp file. append=FALSE to overwrite existing layer
-  st_write(paste0("../data/",chosen_city,"/msoa_lon_lat.shp"), append=FALSE)
+  #st_write("data/alt_city/msoa_lon_lat.shp", append=FALSE) # save as shp file. append=FALSE to overwrite existing layer
+  st_write(paste0("data/",chosen_city,"/msoa_lon_lat.shp"), append=FALSE)
 
 # read it in for dodgr_dist calculations (below)
-msoa_lon_lat <- st_read(paste0("../data/",chosen_city,"/msoa_lon_lat.shp")) %>%
[…]

### 2020-11-20 R/_3.3_plot_desire_lines_current_vs_potential.R

@@ -10,11 +10,11 @@ library(lwgeom)
 
 # FLOW DATA
 # get the flow for the chosen city: data retrieved in Script 1 _1_get_flow_data
-city_od <- readr::read_csv(paste0("../data/", chosen_city, "/flows_city.csv"))
+city_od <- readr::read_csv(paste0("data/", chosen_city, "/flows_city.csv"))
 
 # MSOA CODES
 # get the MSOA codes of MSOAs in the chosen city. Data retrieved from _2_distance_and_elevation
-city_msoas <- readr::read_csv(paste0("../data/",chosen_city,"/msoa_codes_city.csv"))
+city_msoas <- readr::read_csv(paste0("data/",chosen_city,"/msoa_codes_city.csv"))
 
 # MSOA CENTROIDS
 # get population weighted centroids from pct and change crs (default is northing)
@@ -24,13 +24,13 @@ city_centroids <- city_centroids %>% dplyr::filter(msoa11cd %in% city_msoas$MSOA
 
 # MSOA BOUNDARIES
 #get msoa boundaries for plotting 
-city_geom <- sf::st_read("../data-raw/MSOA_2011_Boundaries/Middle_Layer_Super_Output_Areas__December_2011__Boundaries.shp") %>%
+city_geom <- sf::st_read("data-raw/MSOA_2011_Boundaries/Middle_Layer_Super_Output_Areas__December_2011__Boundaries.shp") %>%
   st_transform(4326)
 # filter only MSOAs in the city_msoas df
 city_geom <- city_geom %>% dplyr::filter(msoa11cd %in% city_msoas$MSOA11CD)
 
 #POTENTIAL FLOW (Script 3)
-city_potential_cycling <- read_csv(paste0("../data/",chosen_city, "/flows_for_aggregated_routing_opt_3.csv"))
+city_potential_cycling <- read_csv(paste0("data/",chosen_city, "/flows_for_aggregated_routing_opt_3.csv"))
 # add potential flow column to city_od df
 city_od <- city_od %>% left_join(city_potential_cycling, 
                                  by = c("Area of residence" = "Area of residence", 
@@ -125,7 +125,7 @@ tm_shape(desire_cycling_long) +
   tm_scale_bar(color.dark = "gray60") -> p
 
 #save
-tmap_save(tm = p, filename = paste0("../data/", chosen_city,"/Plots/desire_facet_cycling.png"), 
[…]

### 2020-11-20 R/_5.0_identifying_cycle_infastructure_from_osm_tags.R

@@ -11,7 +11,7 @@ library(tmap)
 ########
 
 # centroids for bounding box of chosen city
-msoa_centroids <- st_read(paste0("../data/", chosen_city,"/msoa_lon_lat.shp"))
+msoa_centroids <- st_read(paste0("data/", chosen_city,"/msoa_lon_lat.shp"))
 
 # bounding box
 pts <- st_coordinates (msoa_centroids)
@@ -68,7 +68,7 @@ maxspeed <- streetnet2 %>%
 ########
 
 # Read in the graph with the road network and flows
-graph_sf <- readRDS(paste0("../data/", chosen_city,"/graph_with_flows_weighted.RDS"))
+graph_sf <- readRDS(paste0("data/", chosen_city,"/graph_with_flows_weighted.RDS"))
 
 # this is all road segments with bicycle == 'designated'
 cycle_designated <- streetnet2 %>% 
@@ -195,7 +195,7 @@ tm_shape(graph_sf) +
 p <- tmap_arrange(p1, p2, p4, nrow = 1)
 
 
-tmap_save(tm = p, filename = paste0("../data/", chosen_city,"/Plots/OSM_identifying_cycle_infrastructure.png"),
+tmap_save(tm = p, filename = paste0("data/", chosen_city,"/Plots/OSM_identifying_cycle_infrastructure.png"),
           width=10, height=5)
 
 
@@ -235,7 +235,7 @@ graph_sf <- dplyr::left_join(graph_sf, graph_sf_cycle, by = "edge_id")
 graph_sf$cycle_infra[is.na(graph_sf$cycle_infra)] <- 0
 
 # save it as an RDS
-saveRDS(graph_sf, file = paste0("../data/", chosen_city, "/graph_with_flows_weighted.Rds"))
+saveRDS(graph_sf, file = paste0("data/", chosen_city, "/graph_with_flows_weighted.Rds"))
 
 ###### FUNCTION TO ADD BINARY COLUMN INDICATING THE PRESENCE OF CYCLING INFRASTRUCTURE #####
 
@@ -273,14 +273,14 @@ infra_exists <- function(graph, network){
 }
 
 # Read in the other graphs, add column for cycling infrastrucure, then overwrite
-graph_sf_unweight <- readRDS(paste0("../data/", chosen_city,"/graph_with_flows_unweighted.RDS"))
[…]

### 2020-11-20 R/_6.0_comparing_weighting_profiles.R

@@ -3,9 +3,9 @@ library(sf)
 library(tmap)
 
 
-graph_sf_weighted <- readRDS(paste0("../data/", chosen_city,"/graph_with_flows_weighted.RDS"))
-graph_sf_unweight <- readRDS(paste0("../data/", chosen_city,"/graph_with_flows_unweighted.RDS"))
-#graph_sf_trunk <- readRDS(paste0("../data/", chosen_city,"/graph_with_flows_trunk.RDS"))
+graph_sf_weighted <- readRDS(paste0("data/", chosen_city,"/graph_with_flows_weighted.RDS"))
+graph_sf_unweight <- readRDS(paste0("data/", chosen_city,"/graph_with_flows_unweighted.RDS"))
+#graph_sf_trunk <- readRDS(paste0("data/", chosen_city,"/graph_with_flows_trunk.RDS"))
 
 
 plot(graph_sf_weighted['flow'])
@@ -67,7 +67,7 @@ ggplot(data=person_km , aes(x=highway, y=dist_perc, group=factor(weighting), fil
     coord_flip() -> p
 p
 
-ggsave(path = paste0("../data/", chosen_city,"/Plots"), 
+ggsave(path = paste0("data/", chosen_city,"/Plots"), 
        file=paste0("perc_person-km-per-highway-type_", chosen_city, ".png"), p, width = 10, height = 6)
 
 
@@ -85,7 +85,7 @@ ggplot(data=person_km , aes(x=highway, y=dist, group=factor(weighting), fill=fac
 
 p
 
-ggsave(path = paste0("../data/", chosen_city,"/Plots"), 
+ggsave(path = paste0("data/", chosen_city,"/Plots"), 
        file=paste0("person-km-per-highway-type_", chosen_city, ".png"), p, width = 10, height = 6)
 
 
@@ -124,7 +124,7 @@ tm_shape(x) +
             #legend.position = c("right", "bottom"),
             frame = FALSE) -> p
 
-tmap_save(tm = p, filename = paste0("../data/", chosen_city,"/Plots/osm_road_types_all.png"))
+tmap_save(tm = p, filename = paste0("data/", chosen_city,"/Plots/osm_road_types_all.png"))
 
 
 #### Roads with Routed Flow Only 
@@ -150,7 +150,7 @@ tm_shape(x) +
             #legend.position = c("right", "bottom"),
             frame = FALSE) -> p
 
[…]

### 2020-11-20 R/_8.0_growing_a_network.R

@@ -3,7 +3,7 @@ library(sf)
 library(sfnetworks)
 library(tidyverse)
 
-# graph_sf <- readRDS(paste0("../data/", chosen_city, "/graph_with_flows_default_communities.RDS"))
+# graph_sf <- readRDS(paste0("data/", chosen_city, "/graph_with_flows_default_communities.RDS"))
 
 
 ##### FUNCTION TO CHECK IF THE INVESTMENT LENGTH CHOSEN IS REASONABLE. IT SHOULD BE LESS THAN THE TOTAL KM
@@ -132,7 +132,7 @@ growth_egalitarian <- function(graph, km, col_name) {
                             no_components = NA,
                             gcc_size = NA)
   # Group by community and get the edge with the highest flow in each group
-  # to pass column name in function (specific to dplyr): https://stackoverflow.com/questions/48062213/dplyr-using-column-names-as-function-arguments
+  # to pass column name in function (specific to dplyr): https://stackoverflow.com/questions/480622dplyr-using-column-names-as-function-arguments
   # we get max flow per group, and then get max distance on result in case of ties (so that we end up with 1 row per group)
   x <- graph %>% group_by(Community) %>% 
     slice_max(order_by = !! sym(col_name)) %>% 

### 2020-11-20 R/_8.1_growth_utilitarian.R

@@ -6,9 +6,9 @@ library(igraph)
 library(tmap)
 
 #create a directory to store data related to this city (does nothing if directory already exists)
-dir.create(paste0("../data/", chosen_city, "/Plots/Growth_Results"), showWarnings = FALSE)
+dir.create(paste0("data/", chosen_city, "/Plots/Growth_Results"), showWarnings = FALSE)
 
-graph_sf <- readRDS(paste0("../data/", chosen_city, "/graph_with_flows_weighted_communities.RDS"))
+graph_sf <- readRDS(paste0("data/", chosen_city, "/graph_with_flows_weighted_communities.RDS"))
 
 # we weigh the flow on each edge by its distance. We can then get how much of the commuter km are satisfied
 graph_sf$person_km <- graph_sf$flow * graph_sf$d
@@ -48,9 +48,9 @@ no_cycle_infra <- no_cycle_infra - 5
 # let's grow the network from the existing infrastructure
 grow_util <- growth_utilitarian(graph = graph_sf, km = no_cycle_infra, col_name = "flow")
 #save
-saveRDS(grow_util, file = paste0("../data/", chosen_city, "/growth_util.Rds"))
+saveRDS(grow_util, file = paste0("data/", chosen_city, "/growth_util.Rds"))
 # read in 
-#grow_util <- readRDS(paste0("../data/", chosen_city,"/growth_util.Rds"))
+#grow_util <- readRDS(paste0("data/", chosen_city,"/growth_util.Rds"))
 
 
 # get % of edges in gcc
@@ -116,7 +116,7 @@ ggplot(data=grow_util_c , aes(x=dist_c, y=perc_person_km_c)) +
   scale_x_continuous(expand = c(0, 0), limits = c(0, NA)) + 
   scale_y_continuous(expand = c(0, 0), limits = c(0, 100))
 
-ggsave(paste0("../data/", chosen_city,"/Plots/Growth_Results/growth_utilitarian_satisfied_km_all_flow_column.png"))
+ggsave(paste0("data/", chosen_city,"/Plots/Growth_Results/growth_utilitarian_satisfied_km_all_flow_column.png"))
 
 # community level plots
 ggplot(data=grow_util_c, 
@@ -131,7 +131,7 @@ ggplot(data=grow_util_c,
   scale_x_continuous(expand = c(0, 0), limits = c(0, NA)) + 
   scale_y_continuous(expand = c(0, 0), limits = c(0, 100))
 
[…]

### 2020-11-20 R/_8.2_growth_egalitarian.R

@@ -3,7 +3,7 @@ library(tidyverse)
 library(ggtext)
 library(tmap)
 
-graph_sf <- readRDS(paste0("../data/", chosen_city, "/graph_with_flows_weighted_communities.RDS"))
+graph_sf <- readRDS(paste0("data/", chosen_city, "/graph_with_flows_weighted_communities.RDS"))
 
 # we weigh the flow on each edge by its distance. We can then get how much of the commuter km are satisfied
 graph_sf$person_km <- graph_sf$flow * graph_sf$d
@@ -40,9 +40,9 @@ graph_sf <- graph_sf %>%
 grow_egal <- growth_egalitarian(graph = graph_sf, km = no_cycle_infra, col_name = "flow")
 
 # save
-saveRDS(grow_egal, file = paste0("../data/", chosen_city, "/growth_egal.Rds"))
+saveRDS(grow_egal, file = paste0("data/", chosen_city, "/growth_egal.Rds"))
 # read in 
-#grow_egal <- readRDS(paste0("../data/", chosen_city,"/growth_egal.Rds"))
+#grow_egal <- readRDS(paste0("data/", chosen_city,"/growth_egal.Rds"))
 
 # get % of edges in gcc
 grow_egal$gcc_size_perc <- (grow_egal$gcc_size / nrow(graph_sf)) * 100
@@ -107,7 +107,7 @@ ggplot(data=grow_egal_c , aes(x=dist_c, y=perc_person_km_c)) +
   #scale_x_continuous(expand = c(0, 0), limits = c(0, NA)) + 
   scale_y_continuous(expand = c(0, 0), limits = c(0, 100))
 
-ggsave(paste0("../data/", chosen_city,"/Plots/Growth_Results/growth_egalitarian_satisfied_km_all_flow_column.png"))
+ggsave(paste0("data/", chosen_city,"/Plots/Growth_Results/growth_egalitarian_satisfied_km_all_flow_column.png"))
 
 # community level plots
 ggplot(data=grow_egal_c, 
@@ -122,7 +122,7 @@ ggplot(data=grow_egal_c,
   scale_x_continuous(expand = c(0, 0), limits = c(0, NA)) + 
   scale_y_continuous(expand = c(0, 0), limits = c(0, 100))
 
-ggsave(paste0("../data/", chosen_city,"/Plots/Growth_Results/growth_egalitarian_satisfied_km_community_flow_column.png"))
[…]

### 2020-11-20 R/_x_dodgr_weighting_profiles.R

@@ -14,5 +14,5 @@ penalties <- dodgr::weighting_profiles$penalties
 weights_bicycle <- dodgr::weighting_profiles$weighting_profiles %>% 
   filter(name == 'bicycle')
 
-write_dodgr_wt_profile(file = "../data/weight_profile_default")
+write_dodgr_wt_profile(file = "data/weight_profile_default")
 

### 2020-11-20 R/archived/archived_8_growing_a_network.R

@@ -3,7 +3,7 @@ library(sf)
 library(sfnetworks)
 library(tidyverse)
 
-# graph_sf <- readRDS(paste0("../data/", chosen_city, "/graph_with_flows_default_communities.RDS"))
+# graph_sf <- readRDS(paste0("data/", chosen_city, "/graph_with_flows_default_communities.RDS"))
 
 
 ##### FUNCTION TO CHECK IF THE INVESTMENT LENGTH CHOSEN IS REASONABLE. IT SHOULD BE LESS THAN THE TOTAL KM
@@ -173,7 +173,7 @@ growth_community_OLD <- function(graph, km, col_name) {
   # copy of graph to edit
   x <- graph
   # Group by community and get the edge with the highest flow in each group
-  # to pass column name in function (specific to dplyr): https://stackoverflow.com/questions/48062213/dplyr-using-column-names-as-function-arguments
+  # to pass column name in function (specific to dplyr): https://stackoverflow.com/questions/480622dplyr-using-column-names-as-function-arguments
   x <-  x %>% group_by(Community) %>% top_n(1, !! sym(col_name)) %>% ungroup()
   # above might return more than one edge per group (edges tied for highest flow), so here we group the 
   # result by Community and select the longer edge

### 2020-11-20 R/benchmarking_functions.R

@@ -4,7 +4,7 @@ library(sfnetworks)
 library(dplyr)
 library(rbenchmark)
 
-test <- readRDS(paste0("../data/", chosen_city, "/graph_with_flows_default_communities.RDS"))
+test <- readRDS(paste0("data/", chosen_city, "/graph_with_flows_default_communities.RDS"))
 
 
 growth_one_seed_modified <- function(graph, km, col_name) {
@@ -134,7 +134,7 @@ growth_community_modified <- function(graph, km, col_name) {
   # add sequen column
   graph$sequen <- NA
   # Group by community and get the edge with the highest flow in each group
-  # to pass column name in function (specific to dplyr): https://stackoverflow.com/questions/48062213/dplyr-using-column-names-as-function-arguments
+  # to pass column name in function (specific to dplyr): https://stackoverflow.com/questions/480622dplyr-using-column-names-as-function-arguments
   # we get max flow per group, and then get max distance on result in case of ties (so that we end up with 1 row per group)
   x <- graph %>% group_by(Community) %>% 
     slice_max(order_by = !! sym(col_name)) %>% 
@@ -203,7 +203,7 @@ growth_community_2_modified <- function(graph, km, col_name) {
   # add sequence column
   graph$sequen <- NA
   # Group by community and get the edge with the highest flow in each group
-  # to pass column name in function (specific to dplyr): https://stackoverflow.com/questions/48062213/dplyr-using-column-names-as-function-arguments
+  # to pass column name in function (specific to dplyr): https://stackoverflow.com/questions/480622dplyr-using-column-names-as-function-arguments
   # we get max flow per group, and then get max distance on result in case of ties (so that we end up with 1 row per group)
   x <- graph %>% group_by(Community) %>% 
     slice_max(order_by = !! sym(col_name)) %>% 
@@ -297,7 +297,7 @@ growth_community_3_modified <- function(graph, km, col_name) {
   # add sequence column
   graph$sequen <- NA
   # Group by community and get the edge with the highest flow in each group
-  # to pass column name in function (specific to dplyr): https://stackoverflow.com/questions/48062213/dplyr-using-column-names-as-function-arguments
+  # to pass column name in function (specific to dplyr): https://stackoverflow.com/questions/480622dplyr-using-column-names-as-function-arguments
   # we get max flow per group, and then get max distance on result in case of ties (so that we end up with 1 row per group)
   x <- graph %>% group_by(Community) %>% 
     slice_max(order_by = !! sym(col_name)) %>% 
[…]

### 2020-11-20 R/functions.R

@@ -358,7 +358,7 @@ growth_community <- function(graph, km, col_name) {
   # add sequen column
   graph$sequen <- NA
   # Group by community and get the edge with the highest flow in each group
-  # to pass column name in function (specific to dplyr): https://stackoverflow.com/questions/48062213/dplyr-using-column-names-as-function-arguments
+  # to pass column name in function (specific to dplyr): https://stackoverflow.com/questions/480622dplyr-using-column-names-as-function-arguments
   # we get max flow per group, and then get max distance on result in case of ties (so that we end up with 1 row per group)
   x <- graph %>% group_by(Community) %>% 
     slice_max(order_by = !! sym(col_name)) %>% 
@@ -453,7 +453,7 @@ growth_community_2 <- function(graph, km, col_name) {
   # add sequence column
   graph$sequen <- NA
   # Group by community and get the edge with the highest flow in each group
-  # to pass column name in function (specific to dplyr): https://stackoverflow.com/questions/48062213/dplyr-using-column-names-as-function-arguments
+  # to pass column name in function (specific to dplyr): https://stackoverflow.com/questions/480622dplyr-using-column-names-as-function-arguments
   # we get max flow per group, and then get max distance on result in case of ties (so that we end up with 1 row per group)
   x <- graph %>% group_by(Community) %>% 
     slice_max(order_by = !! sym(col_name)) %>% 
@@ -574,7 +574,7 @@ growth_community_3 <- function(graph, km, col_name) {
   # add sequence column
   graph$sequen <- NA
   # Group by community and get the edge with the highest flow in each group
-  # to pass column name in function (specific to dplyr): https://stackoverflow.com/questions/48062213/dplyr-using-column-names-as-function-arguments
+  # to pass column name in function (specific to dplyr): https://stackoverflow.com/questions/480622dplyr-using-column-names-as-function-arguments
   # we get max flow per group, and then get max distance on result in case of ties (so that we end up with 1 row per group)
   x <- graph %>% group_by(Community) %>% 
     slice_max(order_by = !! sym(col_name)) %>% 
@@ -690,7 +690,7 @@ growth_community_4 <- function(graph, km, col_name) {
                             no_components = NA,
                             gcc_size = NA)
   # Group by community and get the edge with the highest flow in each group
-  # to pass column name in function (specific to dplyr): https://stackoverflow.com/questions/48062213/dplyr-using-column-names-as-function-arguments
[…]

### 2020-11-20 R/reprex/dodgr_bug.R

@@ -14,4 +14,4 @@ hampi_sc_weighted <- weight_streetnet(hampi_sc, wt_profile = "bicycle")
 
 
 remotes::install_github("hypertidy/geodist") 
-remotes::install_github("atfutures/dodgr")
+remotes::install_github("atfuturdodgr")

### 2020-11-20 README.Rmd
Minor typo fix

@@
 # this doesn't work because R project for code is in a different path to R project for paper (relative paths inside code won't work)
-# source("R/_1.0_get_flow_data.R")
+source("R/_1.0_get_flow_data.R")
 ```
@@
 ```{r desirefacetcycling, fig.cap="Current and Potential Cycling Demand", out.width="75%", fig.show='hold', echo=FALSE}
-knitr::include_graphics("data/Manchester/Plodesire_facet_cycling.png")
+knitr::include_graphics("data/Manchester/Plots/desire_facet_cycling.png")
 ```

### 2020-11-20 README.Rmd
Update md version of paper

@@
 output:
-  # bookdown::github_document2
+  bookdown::github_document2:
   bookdown::pdf_document2:

### 2020-11-20 README.Rmd
Update abstract

@@
 abstract: "Understanding the motivators and deterrents to cycling is essential for creating infrastructure that is successful in getting more people to adopt cycling as a mode of transport.
-Therefore, the aim of this study is to develop a methodology that determines where cycling infrastructure should be added by accounting for cyclist preference and UK government policies to create Low Traffic Neighborhoods.
-Distance decay, routing, and network analysis techniques are used to examine where future cycling demand in a city will come from, how such demand should be routed on the street network, and how to  consider fair distribution of cycling infrastructure that is in line with egalitarian principles.
-The city of Manchester is used as a case study, but the methodology is generic enough to be applicable to any city.
-For routing, it is found that deviating from shortest paths to avoid high-stress road segments can result in acceptable travel time increases for cyclists.
-Community detection is used to partition the city and evaluate results from utilitarian to egalitarian algorithms on distributing cycling infrastructure.
-The results show that the egalitarian algorithm provides comparable city-wide results to the utilitarian one, demostrating it's potential suitability in the planning of cycling networks.
+The aim of this paper is to develop methods where to prioritise cycling infrastructure, accounting for cyclist preference and UK government policies to create Low Traffic Neighborhoods.
+Distance decay, routing, and network analysis techniques are used to examine where future cycling demand is most likely to arise, how such demand could be accomodated by the street network, and how to ensure a fair distribution of investment.
+The city of Manchester, which has substantial cycling ambitions, is used to demonstrate the methods which, we demonstrate, can be applied to any city.
+Flexible routing profiles that are adjustable to local needs can avoid high-stress road segments, enabling the identification of networks that balance directness against the need for safe and stress-free routes.
+Community detection is used to partition the city and evaluate results from 'utilitarian' and 'egalitarian' policy perspectives.
+We conclude that the 'egalitarian' model prioritisation approach demonstrated in this paper has great potential to support cost-effective strategic cycle network planning internationally.
 \\par\\textbf{Keywords:} cycling networks, routing, transport equity"

### 2020-11-20 README.Rmd
Add brand 2020 reference

@@
 geometry: margin=2cm
-bibliography: mybibfile.bib
+bibliography:
+  - mybibfile.bib
+  - rlbibfile.bib
 #linenumbers: true
@@
 ---
+
+```{r, echo=FALSE, eval=FALSE}
+# aim: add citations efficiently to the paper (can tidy up after)
+citr::tidy_bib_file(rmd_file = "README.Rmd", messy_bibliography = "~/uaf/allrefs.bib", file = "rlbibfile.bib")
+```
 
@@
 
-The 2015 Paris agreement [@agreement2015paris] was an acknowledgment that fundamental changes to our lives and economies are necessary to reduce the impacts of climate change.
-The transport sector is one of many sectors facing transformation, with innovation and policy measures leading to the introduction of ‘greener’ private vehicles.
-But as demand for transport rises, resulting in increased congestion, public and active modes of transport are also gaining more recognition.
+The 2015 Paris agreement [@agreement2015paris] acknowledged that fundamental changes to societies and economies are necessary to mitigate climate change.
+Like other sectors, transport is under substantial pressure to decarbonise resulting in a number of technical innovations including electric vehicles.
+But new vehicle technologies can only go so far and do not tackle parallel problems such as congestion, road traffic casualties and physical inactivity [@brand_climate_2020].
+
+In this context, interest and investment in active modes is growing.
 The benefits of the latter extend beyond congestion and the environment, as it promises to help alleviate what is referred to as the pandemic of global inactivity; physical inactivity is on the rise and has become the 4th highest cause of death globally [@kohl2012pandemic].

### 2020-11-21 README.Rmd
merge robin updates and fix minor typos

@@
   bookdown::github_document2:
-  bookdown::pdf_document2:
-      toc: false
-      extra_dependencies: ["subcaption", "longtable", "float", "multirow", "placeins", "multicol"]
+  # bookdown::pdf_document2:
+  #     toc: false
+  #     extra_dependencies: ["subcaption", "longtable", "float", "multirow", "placeins", "multicol"]
 ---
@@
 <!--   - Source: -->
-<!--     <http://geoportal.statistics.gov.datasets/826dc85fb600440889480f4d9dbb1a24_0> -->
+<!--     <http://geoportal.statistics.gov.uk/datasets/826dc85fb600440889480f4d9dbb1a24_0> -->
 <!--   - Location in Repo: data-raw/MSOA\_2011\_Boundaries/\[Add files here\] -->
@@
 
+```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE}
+# scripts for potential demand
+
+# source("R/_1.0_get_flow_data.R")
+# source("R/_2.0_distance_and_elevation.R")
+# source("R/_3.0_potential_demand.R")
+# source("R/_3.1_plot_mode_shares.R")
+# source("R/_3.2_plot_od_comparison.R")
+# source("R/_3.3_plot_desire_lines_current_vs_potential.R")
+```
+
 Using existing cycling demand to inform decisions on where cycling infrastructure should be added reinforces existing cycling patterns and ignores potential cycling demand that could be satisfied by a connected network.
@@
 
-```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE}
-# this doesn't work because R project for code is in a different path to R project for paper (relative paths inside code won't work)
-source("R/_1.0_get_flow_data.R")
-```
-
 ```{r potdemhistograms, fig.cap="Distribution of Potential Cycling Demand", out.width="32%", fig.show='hold', echo=FALSE}
@@
 # Routing
+
+```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE}
+# scripts for routing
+
+# source("R/_4.0_aggregating_flows.R")
+# source("R/_5.0_identifying_cycle_infastructure_from_osm_tags.R")
+# source("R/_6.0_comparing_weighting_profiles.R")
+```
 
@@
 # Community Detection
+
+```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE}
+# scripts for cycle network algorithms
+
+# source("R/_7.0_community_detection.R")
+# source("R/_8.0_growing_a_network.R")
+# source("R/_8.1_growth_utilitarian.R")
+# source("R/_8.2_growth_egalitarian.R")
+```
 

### 2020-11-21 R/_8.0_growing_a_network.R

@@ -132,7 +132,7 @@ growth_egalitarian <- function(graph, km, col_name) {
                             no_components = NA,
                             gcc_size = NA)
   # Group by community and get the edge with the highest flow in each group
-  # to pass column name in function (specific to dplyr): https://stackoverflow.com/questions/480622dplyr-using-column-names-as-function-arguments
+  # to pass column name in function (specific to dplyr): https://stackoverflow.com/questions/48062213/dplyr-using-column-names-as-function-arguments
   # we get max flow per group, and then get max distance on result in case of ties (so that we end up with 1 row per group)
   x <- graph %>% group_by(Community) %>% 
     slice_max(order_by = !! sym(col_name)) %>% 

### 2020-11-21 R/reprex/dodgr_bug.R

@@ -14,4 +14,4 @@ hampi_sc_weighted <- weight_streetnet(hampi_sc, wt_profile = "bicycle")
 
 
 remotes::install_github("hypertidy/geodist") 
-remotes::install_github("atfuturdodgr")
+remotes::install_github("atfutures/dodgr")

### 2020-11-23 README.Rmd
deleted Rproj inside R folder

@@
 abstract: "Understanding the motivators and deterrents to cycling is essential for creating infrastructure that is successful in getting more people to adopt cycling as a mode of transport.
-The aim of this paper is to develop methods where to prioritise cycling infrastructure, accounting for cyclist preference and UK government policies to create Low Traffic Neighborhoods.
+The aim of this paper is to develop methods to prioritise cycling infrastructure, accounting for cyclist preference and UK government policies to create Low Traffic Neighborhoods.
 Distance decay, routing, and network analysis techniques are used to examine where future cycling demand is most likely to arise, how such demand could be accomodated by the street network, and how to ensure a fair distribution of investment.
@@
 output:
-  bookdown::github_document2:
-  # bookdown::pdf_document2:
-  #     toc: false
-  #     extra_dependencies: ["subcaption", "longtable", "float", "multirow", "placeins", "multicol"]
+  # bookdown::github_document2:
+  bookdown::pdf_document2:
+      toc: false
+      extra_dependencies: ["subcaption", "longtable", "float", "multirow", "placeins", "multicol"]
 ---

### 2020-11-23 R/_3.3_plot_desire_lines_current_vs_potential.R

@@ -125,7 +125,7 @@ tm_shape(desire_cycling_long) +
   tm_scale_bar(color.dark = "gray60") -> p
 
 #save
-tmap_save(tm = p, filename = paste0("data/", chosen_city,"/Plodesire_facet_cycling.png"), 
+tmap_save(tm = p, filename = paste0("data/", chosen_city,"/Plots/desire_facet_cycling.png"), 
           width=8, height=6)
 
 # clear environment

### 2020-12-11 README.Rmd
Mor updates

@@
 output:
-  # bookdown::github_document2:
-  bookdown::pdf_document2:
+  bookdown::github_document2:
+  # bookdown::pdf_document2:
       toc: false
@@
 
-As noted by @buehler2016bikeway, it is only recently that research has pivoted from focusing on localized, street or intersection level impact towards studying the cycling network as a whole.
+As noted by @buehler2016bikeway, there has been shift in focus from localized (street or intersection) level planning towards planning based on studying the cycling network as a whole since the turn of the 21^st^ century.
 They emphasize the promise of this shift in capturing the network-wide effect of street-level interventions.
@@
 
-The research provides a methodology for designing a cycling network that respects both the needs of the users and the ambitions of the government.
+<!-- I think this section should start with a short (one sentence) re-statement of what the paper has done. -->
+
+The research presented in this paper provides a methodology for designing a cycling network that respects both the needs of the users and the ambitions of transport planners and local/regional level government stakeholders.
 The methodology can be used as a starting point for introducing a direct, connected, and low-stress network for any city by leveraging available open data.
@@
 It should be noted that the recommendations are only as good as the granularity of the available data.
+
 A core part of the methodology is determining where there is high potential for cycling, and using this as a basis for recommendations on road space reallocation in order to unlock potential cycling demand.
@@
 
-The research hopes to aid in the current shift towards active transport that is being promoted by the UK government.
-The success of such a shift is contingent on the appeal of the cycling network for prospective cyclists, a fact that was used as the basis for the methodology outlined here.
+The methods can support the ongoing shift in transport planning towards active transport that is being promoted by the UK government and other authorities at city, regional and national levels worldwide.
+The success of such a shift depends on the appeal of the cycling network for prospective cyclists, a fact that was used as the basis for the methodology outlined here.
+
+<!-- I think it would be good to say how others could reproduce the methods in other cities (RL) -->
+
+<!-- There should also be links to the literature. -->
 

### 2020-12-20 README.Rmd
add pdf

@@
 output:
-  bookdown::github_document2:
-  # bookdown::pdf_document2:
+  # bookdown::github_document2:
+  bookdown::pdf_document2:
       toc: false
@@
 
-As noted by @buehler2016bikeway, there has been shift in focus from localized (street or intersection) level planning towards planning based on studying the cycling network as a whole since the turn of the 21^st^ century.
+As noted by @buehler2016bikeway, there has been a shift in focus from localized (street or intersection) level planning towards planning based on studying the cycling network as a whole since the turn of the 21^st^ century.
 They emphasize the promise of this shift in capturing the network-wide effect of street-level interventions.
@@
 
-1. Identify all links that have segregated cycling infrastructure and add them to the initial solution
+<!-- 1. Identify all links that have segregated cycling infrastructure and add them to the initial solution
 2. Identify all links that neighbor links in the current solution
 3. Select neighboring link with highest flow and add it to the solution
-4. Repeat steps 2 \& 3 until all flow is satisfied or investment threshold is met
+4. Repeat steps 2 \& 3 until all flow is satisfied or investment threshold is met -->
+
+```{r, eval = FALSE}
+1. Identify all links that have segregated cycling infrastructure and add them to the 
+   initial solution
+2. Identify all links that neighbor links in the current solution
+3. Select neighboring link with highest flow and add it to the solution
+4. Repeat steps 2 and 3 until all flow is satisfied or investment threshold is met
+```
 
@@
 
-1. Identify all links that have segregated cycling infrastructure and add them to the initial solution
+<!-- 1. Identify all links that have segregated cycling infrastructure and add them to the initial solution
 2. Identify all links that neighbor links in the current solution
@@
 4. If there are no more neighboring links in a community, select the link with the highest flow in that community, regardless of connectivity, and add it to the solution
-5. Repeat steps 2, 3 \& 4 until all flow is satisfied or investment threshold is met
+5. Repeat steps 2, 3 \& 4 until all flow is satisfied or investment threshold is met -->
+
+```{r, eval = FALSE}
+1. Identify all links that have segregated cycling infrastructure and add them to the initial 
+   solution
+2. Identify all links that neighbor links in the current solution
+3. Select from each community one neighboring link with highest flow and add it to the 
+   solution
+4. If there are no more neighboring links in a community, select the link with the highest 
+   flow in that community, regardless of connectivity, and add it to the solution
+5. Repeat steps 2, 3 and 4 until all flow is satisfied or investment threshold is met
+
+```
 
@@
 For each road segment, the person-km is equal to $flow * l$.
-In the case of Manchester, Table (**Reference the table**) shows that almost half of the person-km is in community 1 , while only 0.5\% of total person-km on the network is in community 5.
+In the case of Manchester, Table \@ref(tab:personkmtable) shows that almost half of the person-km is in community 1 , while only 0.5\% of total person-km on the network is in community 5.
 
@@
 <!-- I think it would be good to say how others could reproduce the methods in other cities (RL) -->
+<!--  HUSSEIN TO DO!-->
 

### 2020-12-29 README.Rmd
conclusion edits

@@
 We conclude that the 'egalitarian' model prioritisation approach demonstrated in this paper has great potential to support cost-effective strategic cycle network planning internationally.
-\\par\\textbf{Keywords:} cycling networks, routing, transport equity"
+\\par\\textbf{Keywords:} cycling networks, low-traffic neighborhoods, routing, transport equity"
 # keywords: "cycling, low-traffic neighborhoods, potential demand, routing"
@@
 
-<!-- There are a couple of files that cannot be synced to github due to their -->
-<!-- size. These files are neseccary for the scripts to run. Below are links -->
-<!-- to where you can download them, and instructions on where to place them -->
+<!-- The following file cannot be synced to github due to its size. -->
+<!-- The file is neseccary for the scripts to run. Below is a link -->
+<!-- to where you can download it, and instructions on where to place it -->
 <!-- in the repo file structure -->
-
-<!-- Flow Data (2011 Census Origin-Destination Data): -->
-
-<!--   - Source: <https://www.nomisweb.co.uk/census/2011/bulk/rOD1> —\> -->
-<!--     Choose File **“WU03EW”** -->
-<!--   - Location in Repo: data-raw/flow\_data.csv -->
 
@@
 Much cycling infrastructure in this country is inadequate.
-It reflects a belief, con- scious or otherwise, that hardly anyone cycles, that cycling is unimportant and that cycles must take no meaningful space from more important road users, such as motor vehicles and pedestrians - [@departmentgearchange2020]"
+It reflects a belief, conscious or otherwise, that hardly anyone cycles, that cycling is unimportant and that cycles must take no meaningful space from more important road users, such as motor vehicles and pedestrians - [@departmentgearchange2020]"
 
@@
 A considerable amount of research has been done towards that end.
-Segregated cycling infrastructure^[Segregated cycling infrastructure refers to road space that is allocated to cyclists only, with physical separation to protect cyclists from other modes of transport.] has been shown to increase cycling uptake [@aldred2019impacts; @goodman2014new; @marques2015infrastructure], with the separation from motorized vehicles being key [@winters2011motivators].
+Segregated cycling infrastructure^[*Segregated cycling infrastructure* refers to road space that is allocated to cyclists only, with physical separation to protect cyclists from other modes of transport.] has been shown to increase cycling uptake [@aldred2019impacts; @goodman2014new; @marques2015infrastructure], with the separation from motorized vehicles being key [@winters2011motivators].
 Revealed preference of cyclists shows that they are willing to deviate from the most efficient routes in order to commute on safer roads [@crane2017longitudinal].
@@
 It should be reiterated that the weighting profile used for routing has been developed for the purposes of this study.
-It creates a hierarchy of road preference that is grounded in cyclist preference and government plans to create LTNs.
+It creates a hierarchy of road preference that is grounded in cyclist preferences and government plans to create LTNs.
 Sensitivity analysis should be done to determine an optimal weighting profile, but given the variation in city road networks<!-- (Figure \ref{fig:perc_person-km}) -->, these would probably require calibration to the specific city.
@@
 
+<!-- 1. Create links between MSOA centroids and weigh these links by potential cycling demand between them.
+2. Use Louvian method to determine optimal number of communities and assign each MSOA centroid to a community.
+3. Assign each road link to the same community as the closest MSOA centroid to it. -->
+```{r, eval = FALSE}
 1. Create links between MSOA centroids and weigh these links by potential cycling demand between them.
@@
 3. Assign each road link to the same community as the closest MSOA centroid to it.
+```
 
@@
   dplyr::group_by(Community) %>%
-  summarize(`Person-Km (Total)` = round(sum(person_km),0),
+  summarize(`Person-Km (Total)` = format(round(sum(person_km),0), big.mark=","),
             `Person-Km (%)` = round(sum(person_km_perc),1))
@@
 <!-- I think this section should start with a short (one sentence) re-statement of what the paper has done. -->
+<!-- (HM) Answer: done --> 
+This paper provides a methodology for prioritizing road segments for investments in cycling infrastructure, with the output being a proposed cycling network for the city under study.
 
-The research presented in this paper provides a methodology for designing a cycling network that respects both the needs of the users and the ambitions of transport planners and local/regional level government stakeholders.
-The methodology can be used as a starting point for introducing a direct, connected, and low-stress network for any city by leveraging available open data.
-The only data necessary for reproducing it is the road network (from OSM), its topography (from satellite imagery), and commuter data (from the national census).
+The methodology is designed so as to respect both the needs of the users and the ambitions of transport planners and local/regional level government stakeholders.
+It can can be used as a starting point for introducing a direct, connected, and low-stress network for any city by leveraging available open data.
+The only data necessary for reproducing it is the road network (from OSM), its topography (from satellite imagery), and commuter data (from the national census)^[The results are easily reproducible for all UK cities, and can also be reproduced for cities elsewhere given the availability of commuter data.
+Instructions for reproducing the results are detailed in this [Github repository](https://github.com/Hussein-Mahfouz/cycle-networks)].
 It should be noted that the recommendations are only as good as the granularity of the available data.
@@
 <!-- I think it would be good to say how others could reproduce the methods in other cities (RL) -->
-<!--  HUSSEIN TO DO!-->
+<!-- (HM) Answer: Done. I added a footnote to the first paragraph in the conclusion linking to the github repo -->
 
 <!-- There should also be links to the literature. -->
+<!-- (HM) Answer: There are links to the literature at the end of each thematic section (Calculating Potential Cycling Demand / Routing / Road Segment Prioritization).
+There is a discussion paragraph in each of them.
+What additional links are necessary here?
+-->
 

### 2021-01-03 README.Rmd
Build docx version

@@
   # bookdown::github_document2:
-  bookdown::pdf_document2:
+  # bookdown::pdf_document2:
+  bookdown::word_document2:
       toc: false

### 2021-01-03 README.Rmd
Tweak abstract (again)

@@
 numbersections: true
-abstract: "Understanding the motivators and deterrents to cycling is essential for creating infrastructure that is successful in getting more people to adopt cycling as a mode of transport.
-The aim of this paper is to develop methods to prioritise cycling infrastructure, accounting for cyclist preference and UK government policies to create Low Traffic Neighborhoods.
-Distance decay, routing, and network analysis techniques are used to examine where future cycling demand is most likely to arise, how such demand could be accomodated by the street network, and how to ensure a fair distribution of investment.
-The city of Manchester, which has substantial cycling ambitions, is used to demonstrate the methods which, we demonstrate, can be applied to any city.
-Flexible routing profiles that are adjustable to local needs can avoid high-stress road segments, enabling the identification of networks that balance directness against the need for safe and stress-free routes.
+abstract: "Understanding the motivators and deterrents to cycling is essential for creating infrastructure that gets more people to adopt cycling as a mode of transport.
+The aim of this paper is to develop methods to support the prioritisation of cycling infrastructure and cycling network design, accounting for cyclist preferences and the growing emphasis on 'filtered permeability' and 'Low Traffic Neighborhood' interventions internationally.
+Distance decay, routing, and network analysis techniques are used to examine where future cycling demand is most likely to arise, how such demand could be accomodated within existing street networks, and how to ensure a fair distribution of investment.
+The city of Manchester is used as a case study to demonstrate the methods which, as demonstrated in reproducible code and open access data accompanying the paper, can be applied to any city.
+We find that flexible routing profiles that are adjustable to local needs can avoid high-stress road segments, enabling the identification of networks that balance directness against the need for safe and stress-free routes.
 Community detection is used to partition the city and evaluate results from 'utilitarian' and 'egalitarian' policy perspectives.
-We conclude that the 'egalitarian' model prioritisation approach demonstrated in this paper has great potential to support cost-effective strategic cycle network planning internationally.
+Based on discussion of the findings from a policy perspective, we conclude that the 'egalitarian' model prioritisation approach demonstrated in this paper has great potential to support cost-effective strategic cycle network planning internationally.
 \\par\\textbf{Keywords:} cycling networks, low-traffic neighborhoods, routing, transport equity"

### 2021-01-04 README.Rmd

@@
   # bookdown::github_document2:
-  # bookdown::pdf_document2:
-  bookdown::word_document2:
+  bookdown::pdf_document2:
+  # bookdown::word_document2:
       toc: false
@@
 
-The analysis is heavily dependent on commuter data^[*commuter data* refers to publicly available Origin-Destination (OD) data, normally obtained from a census.
-This study uses data from the UK census [@ONS2011flowdata], which contains aggregate statistics on number of commuters between MSOA pairs, by mode of travel.].
-Commuter data in the UK is publicly available at the Middle layer Super Output Area (MSOA) level; the average MSOA has a population of 8209 [@ofn2018population].
-@iacono2010measuring note that such large travel zones are not ideal for understanding route choice behaviour of cyclists and pedestrians.
-They also give rise to an \`ecological fallacy' whereby average characteristics are assumed to apply to all residents of the aggregated geographical area.
-Given that more granular data is not publicly available, the study uses MSOA-level commuter data.
-The methodology is however applicable to more granular commuter data should it become available.
+The approach is based on origin-destination (OD), which can be obtained from many sources.
+In this paper we use open access data from the UK census [@ONS2011flowdata], which contains aggregate statistics on number of commuters between Middle layer Super Output Area (MSOA) zones, which have an average just over 8000 [@ofn2018population].
 
@@
 
-# Conclusions
+# Disussion and conclusions
 
-<!-- I think this section should start with a short (one sentence) re-statement of what the paper has done. -->
-<!-- (HM) Answer: done --> 
 This paper provides a methodology for prioritizing road segments for investments in cycling infrastructure, with the output being a proposed cycling network for the city under study. 
-
-The methodology is designed so as to respect both the needs of the users and the ambitions of transport planners and local/regional level government stakeholders.
-It can can be used as a starting point for introducing a direct, connected, and low-stress network for any city by leveraging available open data.
+The approach aims to respect both the needs of the users and the ambitions of stakeholders working at local/regional levels.
+The results, primarily detailed route network maps based on current travel behaviour derived from origin-destination data, can starting point for introducing a direct, connected, and low-stress network for any city by leveraging available open data.
 The only data necessary for reproducing it is the road network (from OSM), its topography (from satellite imagery), and commuter data (from the national census)^[The results are easily reproducible for all UK cities, and can also be reproduced for cities elsewhere given the availability of commuter data.
 Instructions for reproducing the results are detailed in this [Github repository](https://github.com/Hussein-Mahfouz/cycle-networks)].
-It should be noted that the recommendations are only as good as the granularity of the available data.
+
+The approach is not without limitations.
+The level of detail is only as good as the granularity of the available data (in this case relatively coarse zones).
+@iacono2010measuring note that such large travel zones are not ideal for understanding route choice behaviour of cyclists and pedestrians.
+They also give rise to an \`ecological fallacy' whereby average characteristics are assumed to apply to all residents of the aggregated geographical area, suggesting a need to applying the methods to more granular origin-destination data (and for governments and other data-collecting organisations to make origin-destination more readily available).
 

### 2021-01-04 R/README.Rmd
minor typos + added README inside R directory with reproducibility instructions

---
title: "Road Segment Prioritization for Bicycle Infrastructure"
author: "Hussein Mahfouz"
date: "18/8/2020"
output: github_document
fontsize: 10pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This repository contains the code used for the paper. Below is an explanation of the scripts used, and how the analysis can be reproduced.

## Paper

The paper can be found in the repo,[**ADD LINK TO PDF**]

***

## Missing Data

There are a couple of files that cannot be synced to github due to their size. These files are neseccary for the scripts to run. Below are links to where you can download them, and instructions on where to place them in the repo file structure

Middle Layer Super Output Areas (December 2011) Boundaries:

* Source: http://geoportal.statistics.gov.uk/datasets/826dc85fb600440889480f4d9dbb1a24_0
* Location in Repo: data-raw/MSOA_2011_Boundaries/[Add files here]

***

## Scripts

The scripts should be run in the order they are numbered in (and listed in here). The only exception is _x_dodgr_weighting_profiles.R.

#### ___1.0_get_flow_data.R_

This script matches MSOA to major towns and cities using data from [here](http://geoportal.statistics.gov.uk/datasets/78ff27e752e44c3194617017f3f15929). 
It then matches the results to the Census flow data so that all OD pairs have an Origin City and Destination City (MSOAs in rural areas are not matched to a city)

In this script, you choose which city you wish to run the analysis from this list of available towns and cities:

```{r, warning=FALSE, echo=FALSE, message=FALSE}
library(tidyverse)
library(gt)
# this is a lookup table matching MSOAs to major towns and cities
city_names <- read_csv('../data-raw/Middle_Layer_Super_Output_Area__2011__to_Major_Towns_and_Cities__December_2015__Lookup_in_England_and_Wales.csv') 
# change column name
city_names <- city_names %>% rename(city = TCITY15NM)

unique(city_names$city) %>% sort()

```

This is done in line 17. For example:

```{r}
chosen_city <- "Manchester"
```

The script then filters all flow data where both the Origin MSOA **AND** the destination MSOA are in the chosen city

If you wish to run the analysis on London, then make sure you have a computer that is up to the task. I didn't :(

#### ___2.0_distance_and_elevation.R_

This script is used to get the distance and slope between each OD pair

* Distance: The routed distance using the [dodgr](https://atfutures.github.io/dodgr/) package.
* Slope: The average slope along the route seperating the OD pair, using the [slopes](https://itsleeds.github.io/slopes/) package

#### ___3.0_potential_demand.R_

**EDIT THIS!!**

This script is used to estimate where additional cycling demand will come from. Let's say that the target for Manchester is a 10% increase in cyling mode share, how many additional cyclists do we assign to each OD pair to reach that target mode share

1. Predict Probability of Cycling Between Each OD Pair Based On Geography
      1. A glm is used to predict probability of cycling based on distance and slope

2. Accounting for Existing Mode Share
      1. Look at performance of each OD pair and assign additional cyclists accordingly. OD pairs that have a            low cycling mode share are allocated more cyclists than OD pairs that already have a high cycling mode          share. This is because OD pairs with low cycling mode share have more potential (latent demand) than OD          pairs with high cycling mode share

3. Scaling Results To Match Mode Share Target
      1. Specify target mode share increase (default is 10% but this is unreasonable for a city like Cambridge           that already has a cycling mode share of 40%)
      2. Scale potential cycling demand up or down so that it matches the target % increase

#### ___3.1_plot_mode_shares.R_ , __3.2_plot_od_comparison.R_ , __3.3_plot_desire_lines_current_vs_potential.R_

These three scripts plot the results of ___3.0_potential_demand.R_ .

* Compare the distance distribution of existing cycling mode share and potential cycling mode share:

<p float="left">
  <img src="../data/Manchester/Plots/histogram_distance_all_vs_cycling.png" width="250" />
  <img src="../data/Manchester/Plots/histogram_distance_all_vs_cycling_potential.png" width="250" /> 
  <img src="../data/Manchester/Plots/histogram_distance_cycling_potential_vs_current.png" width="250" />
</p>

* Vizualize Existing and Potential Cycling Flow as Desire Lines

![desire lines](../data/Manchester/Plots/desire_facet_cycling.png)

* Examine where potential cycling demand is assigned.

The methodology in ___3.0_potential_demand.R_ insures that OD pairs that have a low cycling mode share are allocated more cyclists than OD pairs that already have a high cycling mode share. In the figure below, the x axis is a ratio of the cycling mode share of the OD pair to its expected cycling mode share. The expected cycling mode share is obtained from a glm where distance, sqrt(distance), and slope are used as predictors. Looking at the resulting cycling mode share, we see that OD pairs between 2-8km have the highest mode share (consistent with bell-shaped distribution of cycling vs distance), and that mode share increase is highest for OD pairs that have lower than expected cycling mode shares.

<p float="center">
  <img src="../data/Manchester/Plots/mode_share_increase_vs_performance_smooth_Manchester.png" width="400" />
</p>

#### ___x_dodgr_weighting_profiles.R_

The [dodgr](https://atfutures.github.io/dodgr/) package is used to route the cycling demand (flow) onto the road network. This is done using different weighting profiles, as explained in the [documentation](https://atfutures.github.io/dodgr/reference/weighting_profiles.html) of the package. This script is used to download a json file of the weight profile and edit the 'bicycle' entries. Weights are assigned to all OSM road types (for example, we assign a weight of 0 to make sure that no cycling routes utilize them). The weighting profiles used are explained in the methodology.

The weighting profiles used in the analysis are in the data file of the repo. These are:
* weight_profile_unweighted.json: unweighted shortest paths
* weight_profile_weighted.json: weighted shortest paths (weighting profile explained in methodology)
* weight_profile_no_primary_trunk.json: weighted shortest paths with cycling banned on primary and trunk roads

These weighting profiles are used in script ___4.0_aggregating_flows.R_

#### ___4.0_aggregating_flows.R_

This script uses the The [dodgr](https://atfutures.github.io/dodgr/) package is used to route potential cycling demand onto the road network. This is done for the different weighting profiles used in the analysis

#### ___5.0_identifying_cycle_infastructure_from_osm_tags.R_

This script is used to identify all road segments that have segregated cycling lanes. This includes all roads that match any of the 3 following tags:

* highway = cycleway  
* cycleway = track 
* bicycle = designated

#### ___6.0_comparing_weighting_profiles.R_

Here we analyze the street network configuration of the city by comparing the unweighted shortest paths to the weighted shortest paths (check methodology for explanation of weighted shortest paths). The aggregated flow shows us which road types are used, and it is clear that cycleways are not utilized unless the road network is weighted to create a hierarchy of road type preference.

![Unweighted Routing](../data/Manchester/Plots/flows_facet_unweighted.png)

![Weighted Routing](../data/Manchester/Plots/flows_facet_weighted.png)

#### ___7.0_community_detection.R_

Using the potential cycling demand between OD pairs, we are able to define communities in the network. The nodes are the population-weighted MSOA centroids (location obtained from the [pct](https://itsleeds.github.io/pct/) package) and the links between them are weighted by the potential cycling demand between them. The Louvian algorithm is used to assign each MSOA centroid to a community, and then each road segment on the network is assigned to the same community as the MSOA centroid closest to it. The results for Manchester are shown below

![Community Detection](../data/Manchester/Plots/communities_alternative_Manchester.png)

#### ___8.0_growing_a_network.R_

This script contains all the functions for prioritizing road segments for dedicated infrastructure. It is necessary to run this script before 8.1 and 8.2. The speed of the functions is inversely proportional to the size of the city being analyzed (Script 8.2 takes almost 2 hours for Birmingham on a  2.7 GHz Intel Core i5 laptop with 8GB of RAM)

#### ___8.1_plot_network_growth.R_

Here we obtain results for the utilitarian growth function (Algorithms 1 and 2 in the paper)

##### Algorithm 1 (Utilitarian Growth)

Logic:

1.  Identify all links **that have dedicated cycling infrastructure** and add them to the initialsolution
2.  Identify all links that neighbor links in the current solution
3.  Select neighboring link with highest flow and add it to the solution
4.  Repeat steps 2 & 3 until all flow is satisfied or investment threshold is met

Results:

<p float="left">
  <img src="../data/Manchester/Plots/Growth_Results/growth_utilitarian_satisfied_km_all_flow_column.png" width="350" />
  <img src="../data/Manchester/Plots/Growth_Results/growth_utilitarian_satisfied_km_community_flow_column.png" width="350" /> 
</p>

The results show the priority of each road segment (Roads are grouped into 100km groups for vizualization purposes)

<p float="center">
  <img src="../data/Manchester/Plots/Growth_Results/growth_utilitarian_priority_all_FLOW.png" width="450" />
</p>

#### ___8.2_plot_network_growth_community.R_

Here we obtain results for the egalitarian growth function (Algorithms 2 in the paper). We also compare the connectivity of the network proposed by both algorithms

##### Algorithm 2 (Egalitarian Growth)

Logic:

1.  Identify all links that have dedicated cycling infrastructure and add them to the initial solution
2.  Identify all links that neighbor links in the current solution
3.  Select **from each community** one neighboring link with highest flow and add it to thesolution
4.  If there are no more neighboring links in a community, select the link with the highest flow in the community, regardless of connectivity, and add it to the solution
5.  Repeat steps 2, 3 & 4 until all flow is satisfied or investment threshold is met

Results:

<p float="left">
  <img src="../data/Manchester/Plots/Growth_Results/growth_egalitarian_satisfied_km_all_flow_column.png" width="350" />
  <img src="../data/Manchester/Plots/Growth_Results/growth_egalitarian_satisfied_km_community_flow_column.png" width="350" /> 
</p>

Priority of each road segment, and utilization of different OSM road types:

<p float="left">
  <img src="../data/Manchester/Plots/Growth_Results/growth_egalitarian_priority_all_FLOW.png" width="300" />
  <img src="../data/Manchester/Plots/Growth_Results/growth_egalitarian_investment_highways_flow.png" width="300" /> 
</p>

##### Comparing Connectivity of Algorithm 2 and 3

We check the number of connected components and the size of the Largest Connected Component as road segments are added to the solution (the components are the road segments). The initial number of components depends on the existing bicycle network of the city. For Manchester, we can see that the existing bicycle network has over 120 disconnected components (Remember we are only looking at segregated bicycle infrastructure, not painted bicycle lanes).

<p float="left">
  <img src="../data/Manchester/Plots/Growth_Results/growth_util_egal_components_gcc_comparisonManchester.png" width="350" />
  <img src="../data/Manchester/Plots/Growth_Results/growth_util_egal_components_number_comparisonManchester.png" width="350" /> 
</p>

The algorithms seem to provide comparable connectivity gains.

### 2021-01-04 README.Rmd
minor typos + added README inside R directory with reproducibility instructions

@@
 The approach is based on origin-destination (OD), which can be obtained from many sources.
-In this paper we use open access data from the UK census [@ONS2011flowdata], which contains aggregate statistics on number of commuters between Middle layer Super Output Area (MSOA) zones, which have an average just over 8000 [@ofn2018population].
+In this paper we use open access data from the UK census [@ONS2011flowdata], which contains aggregate statistics on number of commuters between Middle layer Super Output Area (MSOA) zones, which have an average population of just over 8000 [@ofn2018population].
 
@@
 The approach aims to respect both the needs of the users and the ambitions of stakeholders working at local/regional levels.
-The results, primarily detailed route network maps based on current travel behaviour derived from origin-destination data, can starting point for introducing a direct, connected, and low-stress network for any city by leveraging available open data.
+The results, primarily detailed route network maps based on current travel behaviour derived from origin-destination data, can be a starting point for introducing a direct, connected, and low-stress network for any city by leveraging available open data.
 The only data necessary for reproducing it is the road network (from OSM), its topography (from satellite imagery), and commuter data (from the national census)^[The results are easily reproducible for all UK cities, and can also be reproduced for cities elsewhere given the availability of commuter data.
@@
 @iacono2010measuring note that such large travel zones are not ideal for understanding route choice behaviour of cyclists and pedestrians.
-They also give rise to an \`ecological fallacy' whereby average characteristics are assumed to apply to all residents of the aggregated geographical area, suggesting a need to applying the methods to more granular origin-destination data (and for governments and other data-collecting organisations to make origin-destination more readily available).
+They also give rise to an \`ecological fallacy' whereby average characteristics are assumed to apply to all residents of the aggregated geographical area, suggesting a need for applying the methods to more granular origin-destination data (and for governments and other data-collecting organisations to make origin-destination data more readily available).
 

### 2021-01-04 README.Rmd
update md

@@
   bookdown::pdf_document2:
-  # bookdown::word_document2:
+  # # bookdown::word_document2:
       toc: false

### 2021-01-09 R/README.Rmd
update md
edit section 4 & merge section 3 into section 2

@@
 ---
-title: "Road Segment Prioritization for Bicycle Infrastructure"
-author: "Hussein Mahfouz"
+title: "Reproducibility Instructions"
+#author: "Hussein Mahfouz"
 date: "18/8/2020"
@@
 
-There are a couple of files that cannot be synced to github due to their size.
-These files are neseccary for the scripts to run.
-Below are links to where you can download them, and instructions on where to place them in the repo file structure
+There is one file that cannot be synced to github due to its size. This file is neseccary for the scripts to run:
 
@@
 
-The script then filters all flow data where both the Origin MSOA **AND** the destination MSOA are in the chosen city
-
-If you wish to run the analysis on London, then make sure you have a computer that is up to the task. I didn't :(
+The script then filters all flow data where both the Origin MSOA **AND** the destination MSOA are in the chosen city.
+If you wish to run the analysis on London, then make sure you have a computer that is up to the task.
 
@@
 
-This script is used to estimate where additional cycling demand will come from.
-Let's say that the target for Manchester is a 10% increase in cyling mode share, how many additional cyclists do we assign to each OD pair to reach that target mode share
-
-1. Predict Probability of Cycling Between Each OD Pair Based On Geography
-      1. A glm is used to predict probability of cycling based on distance and slope
-
-2. Accounting for Existing Mode Share
-      1.
-Look at performance of each OD pair and assign additional cyclists accordingly.
-OD pairs that have a            low cycling mode share are allocated more cyclists than OD pairs that already have a high cycling mode          share.
-This is because OD pairs with low cycling mode share have more potential (latent demand) than OD          pairs with high cycling mode share
-
-3. Scaling Results To Match Mode Share Target
-      1. Specify target mode share increase (default is 10% but this is unreasonable for a city like Cambridge           that already has a cycling mode share of 40%)
-      2. Scale potential cycling demand up or down so that it matches the target % increase
+This script is used to estimate where additional cycling demand will come from. It uses an uptake function from the Propensity to Cycle Tool ([pct](https://itsleeds.github.io/pct/)).
 
@@
 
-* Compare the distance distribution of existing cycling mode share and potential cycling mode share:
+<!-- * Compare the distance distribution of existing cycling mode share and potential cycling mode share: -->
 
-<p float="left">
-  <img src="../data/Manchester/Plots/histogram_distance_all_vs_cycling.png" width="250" />
-  <img src="../data/Manchester/Plots/histogram_distance_all_vs_cycling_potential.png" width="250" /> 
-  <img src="../data/Manchester/Plots/histogram_distance_cycling_potential_vs_current.png" width="250" />
-</p>
+<!-- <p float="left"> -->
+<!--   <img src="../data/Manchester/Plots/histogram_distance_all_vs_cycling.png" width="250" /> -->
+<!--   <img src="../data/Manchester/Plots/histogram_distance_all_vs_cycling_potential.png" width="250" />  -->
+<!--   <img src="../data/Manchester/Plots/histogram_distance_cycling_potential_vs_current.png" width="250" /> -->
+<!-- </p> -->
 
-* Vizualize Existing and Potential Cycling Flow as Desire Lines
+<!-- * Vizualize Existing and Potential Cycling Flow as Desire Lines -->
 
-![desire lines](../data/Manchester/Plots/desire_facet_cycling.png)
+<!-- ![desire lines](../data/Manchester/Plots/desire_facet_cycling.png) -->
 
-* Examine where potential cycling demand is assigned.
+<!-- * Examine where potential cycling demand is assigned.  -->
 
-The methodology in ___3.0_potential_demand.R_ insures that OD pairs that have a low cycling mode share are allocated more cyclists than OD pairs that already have a high cycling mode share.
-In the figure below, the x axis is a ratio of the cycling mode share of the OD pair to its expected cycling mode share.
-The expected cycling mode share is obtained from a glm where distance, sqrt(distance), and slope are used as predictors.
+<!-- In the figure below, the x axis is a ratio of the cycling mode share of the OD pair to its expected cycling mode share.
 Looking at the resulting cycling mode share, we see that OD pairs between 2-8km have the highest mode share (consistent with bell-shaped distribution of cycling vs distance), and that mode share increase is highest for OD pairs that have lower than expected cycling mode shares.
+-->
 
-<p float="center">
-  <img src="../data/Manchester/Plots/mode_share_increase_vs_performance_smooth_Manchester.png" width="400" />
-</p>
+<!-- <p float="center"> -->
+<!--   <img src="../data/Manchester/Plots/mode_share_increase_vs_performance_smooth_Manchester.png" width="400" /> -->
+<!-- </p> -->
 
@@
 
-These weighting profiles are used in script ___4.0_aggregating_flows.R_
+These weighting profiles are used in script ___2.0_distance_and_elevation.R_ AND ___4.0_aggregating_flows.R_
 
@@
 Here we analyze the street network configuration of the city by comparing the unweighted shortest paths to the weighted shortest paths (check methodology for explanation of weighted shortest paths). 
-The aggregated flow shows us which road types are used, and it is clear that cycleways are not utilized unless the road network is weighted to create a hierarchy of road type preference.
+<!-- The aggregated flow shows us which road types are used, and it is clear that cycleways are not utilized unless the road network is weighted to create a hierarchy of road type preference.  -->
 
-![Unweighted Routing](../data/Manchester/Plots/flows_facet_unweighted.png)
+<!-- ![Unweighted Routing](../data/Manchester/Plots/flows_facet_unweighted.png) -->
 
-![Weighted Routing](../data/Manchester/Plots/flows_facet_weighted.png)
+<!-- ![Weighted Routing](../data/Manchester/Plots/flows_facet_weighted.png) -->
 
@@
 The Louvian algorithm is used to assign each MSOA centroid to a community, and then each road segment on the network is assigned to the same community as the MSOA centroid closest to it.
-The results for Manchester are shown below
 
-![Community Detection](../data/Manchester/Plots/communities_alternative_Manchester.png)
+<!-- The results for Manchester are shown below  -->
+
+<!-- ![Community Detection](../data/Manchester/Plots/communities_alternative_Manchester.png) -->
 
@@
 
-Results:
+<!-- Results:  -->
 
-<p float="left">
-  <img src="../data/Manchester/Plots/Growth_Results/growth_utilitarian_satisfied_km_all_flow_column.png" width="350" />
-  <img src="../data/Manchester/Plots/Growth_Results/growth_utilitarian_satisfied_km_community_flow_column.png" width="350" /> 
-</p>
+<!-- <p float="left"> -->
+<!--   <img src="../data/Manchester/Plots/Growth_Results/growth_utilitarian_satisfied_km_all_flow_column.png" width="350" /> -->
+<!--   <img src="../data/Manchester/Plots/Growth_Results/growth_utilitarian_satisfied_km_community_flow_column.png" width="350" />  -->
+<!-- </p> -->
 
-The results show the priority of each road segment (Roads are grouped into 100km groups for vizualization purposes)
+<!-- The results show the priority of each road segment (Roads are grouped into 100km groups for vizualization purposes) -->
 
-<p float="center">
-  <img src="../data/Manchester/Plots/Growth_Results/growth_utilitarian_priority_all_FLOW.png" width="450" />
-</p>
+<!-- <p float="center"> -->
+<!--   <img src="../data/Manchester/Plots/Growth_Results/growth_utilitarian_priority_all_FLOW.png" width="450" /> -->
+<!-- </p> -->
 
@@
 
-Results:
+<!-- Results:  -->
 
-<p float="left">
-  <img src="../data/Manchester/Plots/Growth_Results/growth_egalitarian_satisfied_km_all_flow_column.png" width="350" />
-  <img src="../data/Manchester/Plots/Growth_Results/growth_egalitarian_satisfied_km_community_flow_column.png" width="350" /> 
-</p>
+<!-- <p float="left"> -->
+<!--   <img src="../data/Manchester/Plots/Growth_Results/growth_egalitarian_satisfied_km_all_flow_column.png" width="350" /> -->
+<!--   <img src="../data/Manchester/Plots/Growth_Results/growth_egalitarian_satisfied_km_community_flow_column.png" width="350" />  -->
+<!-- </p> -->
 
-Priority of each road segment, and utilization of different OSM road types:
+<!-- Priority of each road segment, and utilization of different OSM road types: -->
 
-<p float="left">
-  <img src="../data/Manchester/Plots/Growth_Results/growth_egalitarian_priority_all_FLOW.png" width="300" />
-  <img src="../data/Manchester/Plots/Growth_Results/growth_egalitarian_investment_highways_flow.png" width="300" /> 
-</p>
+<!-- <p float="left"> -->
+<!--   <img src="../data/Manchester/Plots/Growth_Results/growth_egalitarian_priority_all_FLOW.png" width="300" /> -->
+<!--   <img src="../data/Manchester/Plots/Growth_Results/growth_egalitarian_investment_highways_flow.png" width="300" />  -->
+<!-- </p> -->
 
-##### Comparing Connectivity of Algorithm 2 and 3
+<!-- ##### Comparing Connectivity of Algorithm 2 and 3 -->
 
-We check the number of connected components and the size of the Largest Connected Component as road segments are added to the solution (the components are the road segments).
+<!-- We check the number of connected components and the size of the Largest Connected Component as road segments are added to the solution (the components are the road segments).
 The initial number of components depends on the existing bicycle network of the city.
 For Manchester, we can see that the existing bicycle network has over 120 disconnected components (Remember we are only looking at segregated bicycle infrastructure, not painted bicycle lanes).
+-->
 
-<p float="left">
-  <img src="../data/Manchester/Plots/Growth_Results/growth_util_egal_components_gcc_comparisonManchester.png" width="350" />
-  <img src="../data/Manchester/Plots/Growth_Results/growth_util_egal_components_number_comparisonManchester.png" width="350" /> 
-</p>
+<!-- <p float="left"> -->
+<!--   <img src="../data/Manchester/Plots/Growth_Results/growth_util_egal_components_gcc_comparisonManchester.png" width="350" /> -->
+<!--   <img src="../data/Manchester/Plots/Growth_Results/growth_util_egal_components_number_comparisonManchester.png" width="350" />  -->
+<!-- </p> -->
 
-The algorithms seem to provide comparable connectivity gains.
+<!-- The algorithms seem to provide comparable connectivity gains.  -->

### 2021-01-09 README.Rmd
edit section 5

@@
 
-# Data and Geographical Scale of Analysis
+<!-- # Data and Geographical Scale of Analysis -->
 
@@
 
-Using existing cycling demand to inform decisions on where cycling infrastructure should be added reinforces existing cycling patterns and ignores potential cycling demand that could be satisfied by a connected network.
+Relying only on existing cycling activity to prioritize cycling infrastructure can reinforce existing cycling patterns and ignores potential cycling demand that could be satisfied by a connected network.
 To avoid this issue, @duthie2014optimization choose to ignore existing demand completely, and focus on creating a network that connects the entire study area.
@@
 For our purposes, we use a logistic regression model to calculate potential cycling demand.
-The model is adopted directly from the Propensity to Cycle Tool (PCT) [@lovelace2017propensity].
-The PCT estimates  the proportion of cyclists ($\boldsymbol{C_{p}}$) for each MSOA pair should the government achieve its target of doubling cycling by 2025.
+The model was adopted from the Propensity to Cycle Tool (PCT), which estimates the proportion of trips ($\boldsymbol{C_{p}}$) for each origin-destination pair that would cycle under different scenarios of change (we used the Government Target scenario, representing a nationwide target of doubling cycling by 2025) as a function of distance and hilliness [@lovelace2017propensity].
 The logistic regression model used to calculate $\boldsymbol{C_{p}}$ has the following parameters:
@@
 
-The next step is to route the potential cycling demand ($\boldsymbol{C_{p}}$) between all OD pairs onto the road network. 
+The next step in our approach is to route the potential cycling demand ($\boldsymbol{C_{p}}$) between all OD pairs onto the road network.
+We choose not to use the PCT approach for routing as it is reliant on external and therefore inflexible routing services.
 <!-- This expands on the work of @mauttone2017bicycle, by going beyond simply favoring roads with existing cycling infrastructure to creating a hierarchy of road preference. -->
@@
 The above points are accounted for by using a weighted road network for routing.
-This has previously been done by multiplying all road segments without cycling infrastructure by an impedance factor [@mauttone2017bicycle], or by assigning a weight to the road segment proportional to the investment cost of bringing it to an acceptable level of stress for cycling  [@duthie2014optimization].
+This has previously been done by multiplying all road segments without cycling infrastructure by an impedance factor [@mauttone2017bicycle], or by assigning different weights to the road segments proportional to the investment cost of bringing them to an acceptable level of stress for cycling  [@duthie2014optimization].
 Our approach is similar to the latter, as we create a weighting profile that is adjusted to favor less stressful streets (based on information from Table  \@ref(table:osmroadtypes)), and roads with existing cycling infrastructure.
@@
 
-All weights are between 0 and 1, and the values in the \textit{Weighted} profile are chosen so as to be inversely proportional to the stress level experienced by cyclists on them.
+All weights are between 0 and 1, and the values in the \textit{Weighted} profile are chosen to be inversely proportional to the stress level experienced by cyclists on them.
 The \textit{Unweighted} weighting profile is used to compare increases in route length resulting from two different approaches:

### 2021-01-09 README.Rmd
edit section 6 + bibliography

@@
 
-One of the main aims of this research is to incorporate egalitarian principles by fairly distributing investments in cycling infrastructure.
-One way of quantifying this is to split up the city into smaller geospatial areas and target equal investment in each of those areas.
+A major challenge facing ‘top-down’ planning approaches is how to incorporate egalitarian principles by fairly distributing investments in cycling infrastructure.
+One way of quantifying this is to split up the city into smaller geospatial areas and target equal investment in each of those areas, , which could also help ensure that on-the-ground surveys are made by local stakeholders, an important component of the planning process  [@parkin2018designing].
 Community detection offers us a way to delineate such a split; cyclists are limited in their commuting distance (see Figure \@ref(fig:cyclinghistmanc)), and so trip attractors are more likely to have a local catchment area of cyclists.

### 2021-01-09 README.Rmd
edit section 7 + 8

@@
 This research attempts to compare the two principles, and in doing so determine whether a methodology formulated based on an egalitarian approach can be feasible in designing a cycling network that aligns with motivators and deterrents to cycling.
-**Write some more here?**
 
@@
 After routing the potential cycling demand onto the road network using weighted shortest paths, we have estimates for the cumulative potential cycling demand passing through all road segments.
-This cumulative demand (referred to as *flow*) is then used as a basis for determining where best to invest in segregated cycling infrastructure.
-In doing so, we must account for the motivations and deterrents for cycling identified in Section \@ref(introduction), namely direct and well connected routes.
+This cumulative demand (referred to as *flow*) is then used as a basis for identifying segments that are in most need of investment in segregated cycling infrastructure.
+In doing so, we must account for the motivations and deterrents for cycling identified in Section \@ref(introduction), namely direct and well-connected routes.
 
-For this purpose, two algorithms are proposed.
+A range of algorithms could be used for prioritisation.
+Because policy priorities vary, we present two algorithms.
 Both utilize existing infrastructure from the beginning and allow us to compare a solution that focuses on utilitarianism to one that focuses on egalitarianism.
@@
 @wardman2007factors developed a mode choice model for the UK and their results showed that improved cycling infrastructure on its own only had modest impacts on mode shift, and even the unlikely scenario of all urban routes being serviced by segregated bike lanes was forecast to increase cycling mode share by only 3%.
-However, cities that invest in more comprehensive cycling projects show a more significant increase in the number of cyclists as well as the cycling mode share [@pucher2010infrastructure].
+International research shows that cities that invest in more comprehensive cycling projects show a more significant increase in the number of cyclists as well as the cycling mode share [@parkin2018designing; @pucher2010infrastructure].
 These cities do not just focus on infrastructure, but on general policies as well as restricting car use.
-Evaluation of policies in Denmark and Germany and the Netherlands has shown that their high cycling mode share is down to a broader set of policies that also include traffic calming, cycling rights of way, bike parking, integration with the public transport network, and making driving cars both expensive and inconvenient [@pucher2008making].
+Evaluation of policies in Denmark and Germany and the Netherlands has shown that their high cycling mode share is down to a broader set of policies that also include traffic calming, filtered permeability interventions, cycling rights of way, bike parking, integration with the public transport network, and making driving cars both expensive and inconvenient [@pucher2008making].
 While these policies are outside the scope of this research, it is important to recognize their key role in bringing about an increase in levels of cycling.
@@
 The approach aims to respect both the needs of the users and the ambitions of stakeholders working at local/regional levels.
-The results, primarily detailed route network maps based on current travel behaviour derived from origin-destination data, can be a starting point for introducing a direct, connected, and low-stress network for any city by leveraging available open data.
-The only data necessary for reproducing it is the road network (from OSM), its topography (from satellite imagery), and commuter data (from the national census)^[The results are easily reproducible for all UK cities, and can also be reproduced for cities elsewhere given the availability of commuter data.
+The results, primarily detailed route network maps based on current travel behaviour derived from origin-destination data, can provide an evidence-based starting point for introducing a direct, connected, and low-stress network for any city by leveraging available open data.
+Given that cycle network development is seen as the “most essential activity entailed in the design of cycle-friendly infrastructure” [@parkin2018designing], we believe that the methods have great potential to inform investment in cities such as Manchester where there is political will to invest in cycling long-term.
+A benefit of the approach is that it has relatively modest data requirements: only the road network (from OSM), its topography (from satellite imagery), and origin-destination data (from the national census in this case) are needed, opening up the possibility of deploying the approach in other cities^[The results are easily reproducible for all UK cities, and can also be reproduced for cities elsewhere given the availability of commuter data.
+For instructions on reproducing the results shown in this paper, see the **README** file in the R folder of the .zip file that accompanies this paper (to be released as open source software when the paper is published).].
+
+<!-- For after publishing
+
+^[The results are easily reproducible for all UK cities, and can also be reproduced for cities elsewhere given the availability of commuter data.
 Instructions for reproducing the results are detailed in this [Github repository](https://github.com/Hussein-Mahfouz/cycle-networks)].
+-->
 

### 2021-01-19 README.Rmd
updated paper based on comments

@@
 The aim of this paper is to develop methods to support the prioritisation of cycling infrastructure and cycling network design, accounting for cyclist preferences and the growing emphasis on 'filtered permeability' and 'Low Traffic Neighborhood' interventions internationally.
-Distance decay, routing, and network analysis techniques are used to examine where future cycling demand is most likely to arise, how such demand could be accomodated within existing street networks, and how to ensure a fair distribution of investment.
-The city of Manchester is used as a case study to demonstrate the methods which, as demonstrated in reproducible code and open access data accompanying the paper, can be applied to any city.
-We find that flexible routing profiles that are adjustable to local needs can avoid high-stress road segments, enabling the identification of networks that balance directness against the need for safe and stress-free routes.
-Community detection is used to partition the city and evaluate results from 'utilitarian' and 'egalitarian' policy perspectives.
-Based on discussion of the findings from a policy perspective, we conclude that the 'egalitarian' model prioritisation approach demonstrated in this paper has great potential to support cost-effective strategic cycle network planning internationally.
+Distance decay, routing, and network analysis techniques are used to examine where future cycling demand is most likely to arise, how such demand could be accommodated within existing street networks, and how to ensure a fair distribution of investment.
+The city of Manchester is used as a case study, although the methods can be applied to any city, as demonstrated in a reproducible code and open access data accompanying the paper.
+We find that flexible routing profiles that are adjustable to local needs can avoid road segments that are stressful to cycle on, enabling the identification of networks that balance directness against the need for safe and stress-free routes.
+We assess two different policy interventions , 'utilitarian' and 'egalitarian' by analyzing the connectivity of the city through community detection.
+From a policy perspective, our findings indicate that the 'egalitarian' prioritisation model approach is the more adequate to support cost-effective strategic cycle network planning.
 \\par\\textbf{Keywords:} cycling networks, low-traffic neighborhoods, routing, transport equity"
@@
 
-- Intro: possible trim down substantially then add some literature in (a) Calculating Potential Demand [DONE]/ (b) Routing [DONE]/ (c) Road Segment Prioritization
+<!-- - Intro: possible trim down substantially then add some literature in (a) Calculating Potential Demand [DONE]/ (b) Routing [DONE]/ (c) Road Segment Prioritization -->
 - Community Detection: Where in the document should this section be? It is now before Road Segment Prioritization, but I can merge them
-- Should we use other cities for the sake of comparison (compare communities, road types etc)
+<!-- - Should we use other cities for the sake of comparison (compare communities, road types etc) -->
 - How to add tables properly (they are now added using latex code which only shows up when outputting to pdf)
 - Footnotes are sometimes split between pages. How do I keep them on the same page?
-- What is the best way to write pseudocode (Steps for Algorithm 1 and 2)
+<!-- - What is the best way to write pseudocode (Steps for Algorithm 1 and 2) -->
 
@@
 
-In this context, interest and investment in active modes is growing.
-The benefits of the latter extend beyond congestion and the environment, as it promises to help alleviate what is referred to as the pandemic of global inactivity; physical inactivity is on the rise and has become the 4th highest cause of death globally [@kohl2012pandemic].
+In this context, interest and investment in active modes are growing.
+The benefits extend beyond congestion and the environment, as active travel can also help alleviate what is referred to as the pandemic of global inactivity; physical inactivity is on the rise and has become the 4th highest cause of death globally [@kohl2012pandemic].
 Various studies have documented the association between active transport and lower risk of disease, including cancer and cardiovascular disease [@celis2017association; @jarrett2012effect; @patterson2020associations].
@@
 
+## Motivators and Deterrents to Cycling
+
 A considerable amount of research has been done towards that end.
@@
 Another defining feature for cycling infrastructure is how well connected it is.
-Cyclists prefer cohesive infrastructure, particularly when cycling on arterial roads with high levels of motorized traffic [@stinson2003commuter], and the lack of well-connected cycling infrastructure is one of the main obstacles to increasing cycling uptake [@caulfield2012determining].
+Cyclists prefer cohesive infrastructure, particularly when cycling on arterial roads with high levels of motorized traffic [@stinson2003commuter].
+The lack of well-connected cycling infrastructure is one of the main obstacles to increasing cycling uptake [@caulfield2012determining].
 While direct and cohesive cycling networks have been shown to positively impact cycling rates, density^[making an area's bicycle network denser means adding more cycling routes in the area and thereby giving cyclists more route options] of the cycling network is also vital [@schoner2014missing].
 
-As noted by @buehler2016bikeway, there has been a shift in focus from localized (street or intersection) level planning towards planning based on studying the cycling network as a whole since the turn of the 21^st^ century.
-They emphasize the promise of this shift in capturing the network-wide effect of street-level interventions.
+## Network-Level Approaches
+
+The studies outlined above lay out the fundamentals for designing cycling networks that generate significant cycling uptake, but they do not propose network-level interventions.
+In this section we outline techniques used in past studies, namely optimization and network analysis techniques, such as connected components and community detection, and examine how they are leveraged to suggest cycling network designs.
+We compare the effectiveness of these network-level studies in incorporating the fundamentals outlined above.
+The methodology of this research is inspired by these techniques, but it attempts to add to them by ensuring that all of the outlined fundamentals are accounted for.
+It also goes further by attempting to factor in ethical considerations relating to distribution of investment.
+
 *Optimization* techniques have been used to propose improvements to cycling networks.
 @mesbah2012bilevel propose a bi-level formulation to optimize allocation of cycling lanes to the network without exceeding a set budget.
-The upper level is the proposed interventions and the lower level is the route choices made by users in reaction to changes in the network.
-The problem accounts for the effect of cycling lanes on car traffic, and attempts to maximize utilization of said lanes with minimal impact on car travel times.
-To improve cohesion of the suggested network, a constraint is added so that each link^[*link* refers to a road segment throughout this research] with a bike lane should be connected to at least one destination.
+They account for the effect of cycling lanes on car traffic, and attempt to maximize utilization of said lanes with minimal impact on car travel times.
 Car usage is not considered by @mauttone2017bicycle, who develop an optimization framework that aims to minimize the total user cost of cycling on the network.
-The aggregate flow on links is obtained by using shortest paths to route existing cycling demand onto the road network, and the solution is a proposed set of links where cycling infrastructure should be added in order to minimize the overall travel cost of cyclists across the network.
-The cost of traversing a link is given as a function of its length and whether or not it has cycling infrastructure, and a discontinuity penalty is also added to prioritize connected road segments.
+The aggregate flow on links^[*link* refers to a road segment throughout this research] is obtained by using shortest paths to route existing cycling demand onto the road network, and the solution is a proposed set of links where cycling infrastructure should be added in order to minimize the overall travel cost of cyclists across the network.
+The cost of traversing a link is given as a function of its length and whether or not it has cycling infrastructure.
 The problem has also been solved by attempting to find the minimum cost of improving roadway links to meet a desired level of service (LOS) [@duthie2014optimization].
-In this formulation, all OD pairs need to be connected by roads that meet the desired LOS, and a directness constraint is added so that paths between OD pairs do not exceed a certain multiple of the shortest path.
+In this formulation, all origin-destination (OD)^[*Origin-Destination (OD)* data is normally obtained from a census.
+In this paper we use open access data from the UK census [@ONS2011flowdata], which contains aggregate statistics on number of commuters between Middle layer Super Output Area (MSOA) zones, by mode of travel.
+MSOAs have an average population of just over 8000 [@ofn2018population].] pairs need to be connected by roads that meet the desired LOS.
+A limitation of these problem formulations do not explicitly solve for continuity, which is dealt with using either a constraint specifying that each link with a bike lane should be connected to at least one destination [@mesbah2012bilevel], a constraint on maximum deviation from shortest paths [@duthie2014optimization], or a discontinuity penalty to prioritize connected road segments [@mauttone2017bicycle].
 
-These problem formulations do not explicitly solve for continuity, which is dealt with using a either (a) a constraint specifying that each link with a bike lane should be connected to at least one destination [@mesbah2012bilevel], (b) a constraint on deviation from shortest paths [@duthie2014optimization], or (c) a discontinuity penalty [@mauttone2017bicycle].
-To solve for continuity, the graph-theoretic concept of *connected components*, has been used.
-@natera2019data study the existing cycling network in terms of its disconnected components and introduce two different algorithms to connect these components by their most critical links and, in doing so, measure the size of the growth of the largest connected component as a function of the kilometers of network added.
-They observe that small investments at strategic points have a large impact on connectivity in most cases.
+In this paper, continuity is analysed by looking at the connectivity of the network through the graph-theoretic concept of *connected components*.
+Other academics have taken this approach; @natera2019data study the existing cycling network in terms of its disconnected components and introduce two different algorithms to connect these components by their most critical links and, in doing so, measure the size of the growth of the largest connected component as a function of the kilometers of network added.
 The concept of connected components is also at the core of the methodology proposed by @olmos2020data.
@@
 The flow on any road segment is the cumulative demand on it, resulting from cyclists commuting between various OD pairs] passing through them, varying the flow threshold for filtering to identify the minimum flow at which the whole city is connected by a giant component.
-The results show a cycling network that connects the entire city, and subtracting links intersecting with current cycling infrastructure identifies links proposed for intervention.
+While these approaches deal with continuity better, they look at the network as a whole when attempting to improve it, and in doing so fail to account for equitable distribution of infrastructure.
 
-The problem formulations outlined above look at the network as a whole when attempting to improve it.
-An alternative approach is to identify the different sub-networks that exist within the larger network, and work on improving each separately.
-Trip patterns in a city are not uniformly distributed geographically, and *community finding* methods have been used to partition study areas into localized areas that experience a disproportionate number of trips within them.
-@akbarzadeh2018designing use a modularity maximization approach [@blondel2008fast] on taxi trip data to identify 7 different communities in the city of Isfahan, Iran.
-An optimization problem is then formulated to connect nodes within each community with cycling infrastructure.
-The emphasis is on connectivity within the communities, not between them.
-@bao2017planning adopt a similar methodology, but use hierarchical clustering to specify the desired number of clusters.
-They use a greedy network expansion algorithm, where the link with the highest benefit-cost ratio in each cluster is selected, and the network is grown by adding neighboring links to the solution until a budget limit is met.
-The benefit is the flow on the link, and each link is assigned a cost based on current road conditions.
+## Ethical Underpinnings and Proposed Approach
 
-All of these network-level methodologies are underpinned by different ethical principles, even though these principles are not explicitly acknowledged by the authors.
+All of these network-level methodologies are underpinned by ethical principles, even though these principles are not explicitly acknowledged by the authors.
 This is important since different ethical principles constitute different problem formulations and targets.
@@
 @nahmias2017integrating criticize the utilitarian approach that has been historically popular in the evaluation of transport investments, explaining how the maximization of overall benefit fails to account for the distribution of that benefit among communities or individuals.
-@lucas2016method explain how transport studies have traditionally looked at the bigger picture without studying the distribution of investments on the different parts of the study area, and go on to propose an egalitarian approach that ensures the dis-aggregation of transport policy benefits across the study area.
+@lucas2016method explain how transport studies have traditionally looked at the bigger picture without studying the distribution of investments on the different parts of the study area, and go on to propose an egalitarian approach that ensures the disaggregation of transport policy benefits across the study area.
 @pereira2017distributive also emphasize the need for a more egalitarian approach to transport planning.
 They highlight accessibility as a cornerstone of distributive justice, and contend that policies should aim to distribute investments in a way that minimizes spatial variations in accessibility.
-This research attempts to compare the two principles, and in doing so determine whether a methodology formulated based on an egalitarian approach can be feasible in designing a cycling network that aligns with motivators and deterrents to cycling.
 
-<!-- # Data and Geographical Scale of Analysis -->
+This research attempts to propose an egalitarian framework for cycling network design.
+This is done by identifying the different sub-networks that exist within the larger network, and ensuring they all get a fair share of investment.
+Trip patterns in a city are not uniformly distributed geographically, and *community finding* methods have been used to partition study areas into localized areas that experience a disproportionate number of trips within them.
+@akbarzadeh2018designing use a modularity maximization approach [@blondel2008fast] on taxi trip data to identify 7 different communities in the city of Isfahan, Iran.
+An optimization problem is then formulated to connect nodes within each community with cycling infrastructure, with the emphasis being on connectivity within the communities, not between them.
+@bao2017planning adopt a similar methodology, first identifying communities and then using a greedy network expansion algorithm to simultaneously add links to each community.
+The link with the highest benefit-cost ratio in each community is selected, and the network is grown by adding neighboring links to the solution until a budget limit is met.
+The benefit is the flow on the link, and each link is assigned a cost based on current road conditions.
 
-The approach is based on origin-destination (OD), which can be obtained from many sources.
-In this paper we use open access data from the UK census [@ONS2011flowdata], which contains aggregate statistics on number of commuters between Middle layer Super Output Area (MSOA) zones, which have an average population of just over 8000 [@ofn2018population].
+<!-- EDIT! Add more to limitations -->
+Our work builds on these *community finding* approaches by proposing a similar greedy network expansion algorithm for cycle network expansion within communities.
+We incorporate community finding methods for study area partitioning with weighted routing to avoid links that are stressful to cycle on.
+In doing so, we propose a methodology that accounts for motivators and deterrents to cycling.
+We propose three sub-methodologies that expand on some of the limitations of previous studies.
+These limitations include (a) bias inherent when basing network design solely on existing cycling demand, (b) proposing routes that may not correspond to studies on cyclist preference and government policies, and (3) insufficient consideration of the ethical principles underlying the analysis.
+Section \@ref(calculating-potential-cycling-demand) focuses on calculating potential cycling demand.
+Section \@ref(routing) focuses on routing the demand onto the road network while accounting for cyclist preferences and government priorities.
+Section \@ref(community-detection) outlines a methodology for partitioning the study area based on a community finding algorithm and routed cycling demand.
+Section \@ref(road-segment-prioritization) introduces the network expansion algorithms, and compares an approach grounded in ‘egalitarianism' to one grounded in ‘utilitarianism'.
 
@@
 
-Relying only on existing cycling activity to prioritize cycling infrastructure can reinforce existing cycling patterns and ignores potential cycling demand that could be satisfied by a connected network.
+Relying only on existing cycling activity to prioritise cycling infrastructure can reinforce existing cycling patterns and ignores potential cycling demand that could be satisfied by a connected network.
 To avoid this issue, @duthie2014optimization choose to ignore existing demand completely, and focus on creating a network that connects the entire study area.
@@
 For our purposes, we use a logistic regression model to calculate potential cycling demand.
-The model was adopted from the Propensity to Cycle Tool (PCT), which estimates the proportion of trips ($\boldsymbol{C_{p}}$) for each origin-destination pair that would cycle under different scenarios of change (we used the Government Target scenario, representing a nationwide target of doubling cycling by 2025) as a function of distance and hilliness [@lovelace2017propensity].
+The model was adapted from the Propensity to Cycle Tool (PCT), which estimates the proportion of trips ($\boldsymbol{C_{p}}$) for each OD pair that would cycle under different scenarios of change (we used the Government Target scenario, representing a nationwide target of doubling cycling by 2025) as a function of distance and hilliness [@lovelace2017propensity].
 The logistic regression model used to calculate $\boldsymbol{C_{p}}$ has the following parameters:
@@
 
-It should be noted that the calculations assume a future that is constrained by physical geography; i.e.
-we consider cycling in the traditional sense.
+It should be noted that the calculations assume cycling in the traditional sense, i.e.
+the topography affects the decision to cycle.
 Recently there have been various micro-mobility solutions, including e-bikes, that allow commuters to traverse longer distances and hillier roads with less effort than traditional bicycles.
-While these modes would probably be associated with less geographical impedance, it is beyond the scope of this work to integrate that into the analysis.
-Doing so is partially restricted by the lack of data on the proliferation of these modes, which raises the point that perhaps the census data category of `Bicycle' is too vague, and should be further dis-aggregated to distinguish between traditional bicycles and other forms of micro-mobility.
+While these modes would probably be associated with less geographical impedance, it is beyond the scope of this work to integrate those modes into the analysis.
+This is partially restricted by the lack of data on the proliferation of these modes, raising the point that perhaps the census data category of `Bicycle' is too vague, and should be further disaggregated to distinguish between traditional bicycles and other forms of micro-mobility.
 
@@
 
+A limitation of the PCT approach outlined in the previous section is that it is reliant on external and therefore inflexible routing services.
 The next step in our approach is to route the potential cycling demand ($\boldsymbol{C_{p}}$) between all OD pairs onto the road network.
@@
 
-\noindent A weighted distance $\boldsymbol{d_{w}}$ for each road segment is calculated as following ^[The **dodgr** r package [@padgham2019dodgr] is used to route cycling demand onto the road network.
+\noindent A weighted distance $\boldsymbol{d_{w}}$ for each road segment is calculated as follows ^[The **dodgr** r package [@padgham2019dodgr] is used to route cycling demand onto the road network.
 The package uses the OpenStreetMaps (OSM) road network and allows the user to assign weights to roads based on their type.
@@
 
-```{r flowsfacetunweighted, fig.cap="Flow Results Based on **Unweighted** Shortest Paths (Manchester)", out.width="90%", fig.show='hold', echo=FALSE}
+```{r flowsfacetunweighted, fig.cap="Flow Results Based on Unweighted Shortest Paths (Manchester)", out.width="90%", fig.show='hold', echo=FALSE}
 knitr::include_graphics("data/Manchester/Plots/flows_facet_unweighted_Manchester.png")
@@
 
-```{r flowsfacetweighted, fig.cap="Flow Results Based on **Weighted** Shortest Paths (Manchester)", out.width="90%", fig.show='hold', echo=FALSE}
+```{r flowsfacetweighted, fig.cap="Flow Results Based on Weighted Shortest Paths (Manchester)", out.width="90%", fig.show='hold', echo=FALSE}
 knitr::include_graphics("data/Manchester/Plots/flows_facet_weighted_Manchester.png")
@@
 A major challenge facing ‘top-down’ planning approaches is how to incorporate egalitarian principles by fairly distributing investments in cycling infrastructure.
-One way of quantifying this is to split up the city into smaller geospatial areas and target equal investment in each of those areas, , which could also help ensure that on-the-ground surveys are made by local stakeholders, an important component of the planning process  [@parkin2018designing].
+One way of quantifying this is to split up the city into smaller geospatial areas and target equal investment in each of those areas, which could also help ensure that on-the-ground surveys are made by local stakeholders, an important component of the planning process  [@parkin2018designing].
 Community detection offers us a way to delineate such a split; cyclists are limited in their commuting distance (see Figure \@ref(fig:cyclinghistmanc)), and so trip attractors are more likely to have a local catchment area of cyclists.
@@
 In our case, the network is the city; the nodes are the population-weighted MSOA centroids and the links connecting each MSOA pair are weighted by the potential cycling demand between them.
-The Louvian method  [@blondel2008fast] is used to separate MSOAs into communities.
+The Louvain method  [@blondel2008fast] is used to separate MSOAs into communities.
 Potential cycling demand is used since we assume that this is what the cycling demand will be once the cycling infrastructure is added.
@@
 ```{r, eval = FALSE}
-1. Create links between MSOA centroids and weigh these links by potential cycling demand between them.
-2. Use Louvian method to determine optimal number of communities and assign each MSOA centroid to a community.
+1. Create links between MSOA centroids and weigh these links by potential cycling demand 
+   between them.
+2. Use the Louvain method to determine the optimal number of communities and assign each MSOA 
+   centroid to a community.
 3. Assign each road link to the same community as the closest MSOA centroid to it.
@@
 
-```{r growthtotal, fig.cap="Comparing Overall (Dashed) and Community Level Person-Km Satisfied (Manchester)", out.width="45%", fig.show='hold', echo=FALSE}
+```{r growthtotal, fig.cap="Comparing Overall (Dashed) and Community Level Person-Km Satisfied (Manchester)", out.width="35%", fig.show='hold', echo=FALSE}
 knitr::include_graphics(c(
@@
 
-```{r growth3MapandBar, fig.cap="Results of Alg. 2 (Manchester)", out.width="45%", fig.show='hold', echo=FALSE}
+```{r growth3MapandBar, fig.cap="Results of Alg. 2 (Manchester)", out.width="35%", fig.show='hold', echo=FALSE}
 knitr::include_graphics(c(
@@
 
-```{r componentsandGCC, fig.cap="Network Characteristics", out.width="45%", fig.show='hold', echo=FALSE}
+```{r componentsandGCC, fig.cap="Network Characteristics", out.width="35%", fig.show='hold', echo=FALSE}
 knitr::include_graphics(c(
@@
 @wardman2007factors developed a mode choice model for the UK and their results showed that improved cycling infrastructure on its own only had modest impacts on mode shift, and even the unlikely scenario of all urban routes being serviced by segregated bike lanes was forecast to increase cycling mode share by only 3%.
-International research shows that cities that invest in more comprehensive cycling projects show a more significant increase in the number of cyclists as well as the cycling mode share [@parkin2018designing; @pucher2010infrastructure].
+International research shows that cities that invest in more comprehensive cycling projects have a more significant increase in the number of cyclists as well as the cycling mode share [@parkin2018designing; @pucher2010infrastructure].
 These cities do not just focus on infrastructure, but on general policies as well as restricting car use.
@@
 The approach aims to respect both the needs of the users and the ambitions of stakeholders working at local/regional levels.
-The results, primarily detailed route network maps based on current travel behaviour derived from origin-destination data, can provide an evidence-based starting point for introducing a direct, connected, and low-stress network for any city by leveraging available open data.
-Given that cycle network development is seen as the “most essential activity entailed in the design of cycle-friendly infrastructure” [@parkin2018designing], we believe that the methods have great potential to inform investment in cities such as Manchester where there is political will to invest in cycling long-term.
+The results, primarily detailed route network maps based on current travel behaviour derived from OD data, can provide an evidence-based starting point for introducing a direct, connected, and low-stress network for any city by leveraging available open data.
+Given that the “most essential activity entailed in the design of cycle-friendly infrastructure is developing a cycle network” [@parkin2018designing], we believe that the methods have great potential to inform investment in cities such as Manchester where there is political will to invest in cycling long-term.
 A benefit of the approach is that it has relatively modest data requirements: only the road network (from OSM), its topography (from satellite imagery), and origin-destination data (from the national census in this case) are needed, opening up the possibility of deploying the approach in other cities^[The results are easily reproducible for all UK cities, and can also be reproduced for cities elsewhere given the availability of commuter data.
@@
 
-A core part of the methodology is determining where there is high potential for cycling, and using this as a basis for recommendations on road space reallocation in order to unlock potential cycling demand.
-A routing engine was used to study the effects of limiting access on different road types, and it was found that reducing cyclist flow on roads with high through-traffic resulted in acceptable increases in commuting distances.
-As a result, a hierarchy of road preference was used to route potential cycling demand.
-Algorithms were developed to determine where investments in cycling infrastructure should be prioritized.
+A core part of the methodology is determining where there is a high potential for cycling, and using this as a basis for recommendations on road space reallocation in order to unlock potential cycling demand.
+The framework provides a routing engine to study the effects of limiting access on different road types, finding that reducing cyclist flow on roads with high through-traffic results in acceptable increases in commuting distances.
+As a result, a hierarchy of road preference needs to be defined to route potential cycling demand.
+The framework provides different algorithms to determine where investments in cycling infrastructure should be prioritized.
 These algorithms, based on connectivity and egalitarian principles of resource distribution, ensured that whatever the level of investment, the resulting cycling network would improve connectivity of existing cycling infrastructure.
 
-The methods can support the ongoing shift in transport planning towards active transport that is being promoted by the UK government and other authorities at city, regional and national levels worldwide.
-The success of such a shift depends on the appeal of the cycling network for prospective cyclists, a fact that was used as the basis for the methodology outlined here.
+The methods can support the ongoing shift towards active transport, which is one of the priority areas of investments worldwide in transport planning.
+This necessary shift predates the pandemic and is driven by ambitions to improve the health and well-being of the citizens worldwide.
+The success of such a shift depends on the appeal of the cycling network for prospective cyclists, together with policies and interventions deterring the usage of the private car.
 

### 2021-01-19 README.Rmd
Updates from meeting

@@
 <!-- - Intro: possible trim down substantially then add some literature in (a) Calculating Potential Demand [DONE]/ (b) Routing [DONE]/ (c) Road Segment Prioritization -->
-- Community Detection: Where in the document should this section be? It is now before Road Segment Prioritization, but I can merge them
+<!-- - Community Detection: Where in the document should this section be? It is now before Road Segment Prioritization, but I can merge them -->
+<!-- Can go into road segment section. -->
 <!-- - Should we use other cities for the sake of comparison (compare communities, road types etc) -->
@@
 The problem has also been solved by attempting to find the minimum cost of improving roadway links to meet a desired level of service (LOS) [@duthie2014optimization].
-In this formulation, all origin-destination (OD)^[*Origin-Destination (OD)* data is normally obtained from a census.
-In this paper we use open access data from the UK census [@ONS2011flowdata], which contains aggregate statistics on number of commuters between Middle layer Super Output Area (MSOA) zones, by mode of travel.
-MSOAs have an average population of just over 8000 [@ofn2018population].] pairs need to be connected by roads that meet the desired LOS.
+In this formulation, all origin-destination (OD) pairs need to be connected by roads that meet the desired LOS.
 A limitation of these problem formulations do not explicitly solve for continuity, which is dealt with using either a constraint specifying that each link with a bike lane should be connected to at least one destination [@mesbah2012bilevel], a constraint on maximum deviation from shortest paths [@duthie2014optimization], or a discontinuity penalty to prioritize connected road segments [@mauttone2017bicycle].
@@
 
+*Origin-Destination (OD)* data can be obtained from a range of sources, including GPS data, household travels surveys and Census data on work locations.
+In areas where observed OD data is unavailable, modelling techniques such as spatial interaction models can be used to estimate travel volumes between zones [@ref?].
+In this paper we use open access data from the UK census [@ONS2011flowdata], which contains aggregate statistics on number of commuters between administrative zones --- Middle layer Super Output Areas (MSOA) --- by mode of travel.
+MSOAs have an average population of just over 8000 [@ofn2018population].^[
+See https://wicid.ukdataservice.ac.uk/ for open access to the OD data.
+]
+Figures \@ref(fig:potdemhistograms) and \@ref(fig:desirefacetcycling) illustrates the proportion of trips cycled by distance and the geographic extent of the input OD dataset used in this paper.
+
 For our purposes, we use a logistic regression model to calculate potential cycling demand.
@@
 The authors use square and square-root distance terms "to capture the non-linear impact of distance on the likelihood of cycling", and interaction terms to capture the combined effect of slope and distance [@lovelace2017propensity].
+Alternative cycling uptake models could be used in different contexts or scenarios of change.
 
-The potential demand calculations show that the current and potential number of cyclists both follow a bell-shaped distribution, with the number of trips peaking around the 3-4km commuting distance and then going back down for longer distances (see Figure \@ref(fig:potdemhistograms)).
+The potential demand calculations show that the current and potential number of cyclists both follow a bell-shaped distribution, with the number of trips peaking around the 3-4km commuting distance and then going back down for longer distances (see Figures \@ref(fig:potdemhistograms) and \@ref(fig:desirefacetcycling)).
+
+<!-- note: could be a joint table/map showing sample - if reviewers request -->
 
@@
 
-```{r desirefacetcycling, fig.cap="Current and Potential Cycling Demand", out.width="75%", fig.show='hold', echo=FALSE}
+```{r desirefacetcycling, fig.cap="Current and Potential Cycling Demand", out.width="100%", fig.show='hold', echo=FALSE}
 knitr::include_graphics("data/Manchester/Plots/desire_facet_cycling.png")
@@
 The above points are accounted for by using a weighted road network for routing.
-This has previously been done by multiplying all road segments without cycling infrastructure by an impedance factor [@mauttone2017bicycle], or by assigning different weights to the road segments proportional to the investment cost of bringing them to an acceptable level of stress for cycling  [@duthie2014optimization].
-Our approach is similar to the latter, as we create a weighting profile that is adjusted to favor less stressful streets (based on information from Table  \@ref(table:osmroadtypes)), and roads with existing cycling infrastructure.
-We believe this to be more practical  than the approach adopted by @mauttone2017bicycle, as it goes beyond simply favoring roads with existing cycling infrastructure to creating a hierarchy of road preference based on perceived stress levels.
-The approach is also in line with the creation of LTNs, as residential streets are those where motorized traffic is most likely to be banned in the creation of LTNs.
+This has previously been done by multiplying all road segments without cycling infrastructure by a fixed impedance factor [@mauttone2017bicycle], or by assigning different weights to the road segments proportional to the investment cost of bringing them to an acceptable level of stress for cycling  [@duthie2014optimization].
+
+For the purposes of this paper, created a weighting profile that is adjusted to favor less stressful streets (based on information from Table  \@ref(table:osmroadtypes)), and roads with existing cycling infrastructure.
+We believe this to be more practical  than the simplistic approach adopted by @mauttone2017bicycle, as it goes beyond simply favoring roads with existing cycling infrastructure enabling a hierarchy of road preference based on perceived stress levels.
+A wide range of weighting profiles could be used, to represent different types of cyclists and road environments, as described in @ref.
+<!-- The approach is also in line with the creation of LTNs, as residential streets are those where motorized traffic is most likely to be banned in the creation of LTNs.  -->
 

### 2021-01-28 README.Rmd
Update README.Rmd
routing section edits

@@
 
-# Ideas and discussion
+<!-- # Ideas and discussion -->
 
@@
 <!-- - Should we use other cities for the sake of comparison (compare communities, road types etc) -->
-- How to add tables properly (they are now added using latex code which only shows up when outputting to pdf)
-- Footnotes are sometimes split between pages. How do I keep them on the same page?
+<!-- - How to add tables properly (they are now added using latex code which only shows up when outputting to pdf) -->
+<!-- - Footnotes are sometimes split between pages. How do I keep them on the same page? -->
 <!-- - What is the best way to write pseudocode (Steps for Algorithm 1 and 2) -->
-
-<!-- ## Missing Data -->
-
-<!-- The following file cannot be synced to github due to its size. -->
-<!-- The file is neseccary for the scripts to run. Below is a link -->
-<!-- to where you can download it, and instructions on where to place it -->
-<!-- in the repo file structure -->
-
-<!-- Middle Layer Super Output Areas (December 2011) Boundaries: -->
-
-<!--   - Source: -->
-<!--     <http://geoportal.statistics.gov.uk/datasets/826dc85fb600440889480f4d9dbb1a24_0> -->
-<!--   - Location in Repo: data-raw/MSOA\_2011\_Boundaries/\[Add files here\] -->
-
-<!-- ----- -->
-
-<!-- ## Scripts -->
-
-<!-- The scripts should be run in the order they are numbered in (and listed -->
-<!-- in here). The only exception is \_x\_dodgr\_weighting\_profiles.R. -->
-
-<!-- Check the readme -->
-<!-- [here](https://github.com/Hussein-Mahfouz/Bicycle-Network-Optimization) -->
-<!-- for detailed info on each script -->
 
@@
 *Origin-Destination (OD)* data can be obtained from a range of sources, including GPS data, household travels surveys and Census data on work locations.
-In areas where observed OD data is unavailable, modelling techniques such as spatial interaction models can be used to estimate travel volumes between zones [@ref?].
+In areas where observed OD data is unavailable, modelling techniques such as spatial interaction models can be used to estimate travel volumes between zones [@black1995spatial; @martinez2013new; @wilson1971family].
 In this paper we use open access data from the UK census [@ONS2011flowdata], which contains aggregate statistics on number of commuters between administrative zones --- Middle layer Super Output Areas (MSOA) --- by mode of travel.
@@
 
-A limitation of the PCT approach outlined in the previous section is that it is reliant on external and therefore inflexible routing services.
 The next step in our approach is to route the potential cycling demand ($\boldsymbol{C_{p}}$) between all OD pairs onto the road network.
@@
 This has previously been done by multiplying all road segments without cycling infrastructure by a fixed impedance factor [@mauttone2017bicycle], or by assigning different weights to the road segments proportional to the investment cost of bringing them to an acceptable level of stress for cycling  [@duthie2014optimization].
+Perceived stress has been found to be a function of vehicular traffic volume and speed [@sorton1994bicycle], both of which vary predictably with road type. 
+@gehrke2020cycling build on this information to use routing impedance factors for road segments that are based on road type and the existence of cycling infrastructure.
 
-For the purposes of this paper, created a weighting profile that is adjusted to favor less stressful streets (based on information from Table  \@ref(table:osmroadtypes)), and roads with existing cycling infrastructure.
+For the purposes of this paper, we created a similar weighting profile that is adjusted to favor less stressful roads (based on the definition by @sorton1994bicycle and information from Table  \@ref(table:osmroadtypes)), and roads with existing cycling infrastructure.
 We believe this to be more practical than the simplistic approach adopted by @mauttone2017bicycle, as it goes beyond simply favoring roads with existing cycling infrastructure enabling a hierarchy of road preference based on perceived stress levels.
-A wide range of weighting profiles could be used, to represent different types of cyclists and road environments, as described in @ref.
+
 <!-- The approach is also in line with the creation of LTNs, as residential streets are those where motorized traffic is most likely to be banned in the creation of LTNs.  -->
@@
 When routing on primary and trunk roads is banned (*weighted_2* profile in Table \@ref(table:weightprofiles)), the average increase relative to unweighted shortest paths rises to 10\%, with certain locations experiencing more significant negative effects on accessibility.
+Given that cyclists will only deviate from shortest paths by a certain amount to access better cycling infrastructure (as explained in Section \@ref(introduction)), allowing flow on some stretches of trunk and primary roads is necessary to insure cycling uptake and equitable access to cycling infrastructure.
+In its new vision for walking and cycling, the Department for Transport acknowledges that minimal segregated stretches of bicycle lanes on main roads will be necessary to avoid circuitous cycling networks [@departmentgearchange2020].
 
@@
 ```
-
-Given that cyclists will only deviate from shortest paths by a certain amount to access better cycling infrastructure (as explained in Section \@ref(introduction)), allowing flow on some stretches of trunk and primary roads is necessary to insure cycling uptake and equitable access to cycling infrastructure.
-In its new vision for walking and cycling, the Department for Transport acknowledges that minimal segregated stretches of bicycle lanes on main roads will be necessary to avoid circuitous cycling networks [@departmentgearchange2020].
 
@@
 It creates a hierarchy of road preference that is grounded in cyclist preferences and government plans to create LTNs.
-Sensitivity analysis should be done to determine an optimal weighting profile, but given the variation in city road networks<!-- (Figure \ref{fig:perc_person-km}) -->, these would probably require calibration to the specific city.
+A wide range of weighting profiles could be used, to represent different types of cyclists and road environments, as described in @gehrke2020cycling and @furth2016network.
+Sensitivity analysis could be done to determine an optimal weighting profile, but given the variation in city road networks<!-- (Figure \ref{fig:perc_person-km}) -->, this would probably require calibration to the specific city.
 More accurate routing could be carried out given the availability of road-level data.
@@
 
-# Community Detection
+# Road Segment Prioritization
 
@@
 
+After routing the potential cycling demand onto the road network using weighted shortest paths, we have estimates for the cumulative potential cycling demand passing through all road segments.
+This cumulative demand (referred to as *flow*) is then used as a basis for identifying segments that are in most need of investment in segregated cycling infrastructure.
+In doing so, we must account for the motivations and deterrents for cycling identified in Section \@ref(introduction), namely direct and well-connected routes.
+
+A range of algorithms could be used for prioritisation.
+Because policy priorities vary, we present two algorithms.
+Both utilize existing infrastructure from the beginning and allow us to compare a solution that focuses on utilitarianism to one that focuses on egalitarianism.
+In both algorithms, links are selected iteratively and the iteration at which each link is added to the solution is recorded.
+Investments in cycling infrastructure can be limited by budget constraints, so it can be useful to see where best to allocate a defined length of segregated infrastructure.
+In order to incorporate egalitarian principles in our approach, we use community detection to partition the study area and evaluate the distribution of investment over the different subdivisions.
+
+## Community Detection
+
 A major challenge facing ‘top-down’ planning approaches is how to incorporate egalitarian principles by fairly distributing investments in cycling infrastructure. 
-One way of quantifying this is to split up the city into smaller geospatial areas and target equal investment in each of those areas, which could also help ensure that on-the-ground surveys are made by local stakeholders, an important component of the planning process  [@parkin2018designing].
+One way of quantifying this is to split up the city into smaller geospatial areas and target equal investment in each of those areas.
+This approach could also help ensure that on-the-ground surveys are made by local stakeholders, an important component of the planning process  [@parkin2018designing].
 Community detection offers us a way to delineate such a split; cyclists are limited in their commuting distance (see Figure \@ref(fig:cyclinghistmanc)), and so trip attractors are more likely to have a local catchment area of cyclists.
@@
 ```
-
-# Road Segment Prioritization
-
-After routing the potential cycling demand onto the road network using weighted shortest paths, we have estimates for the cumulative potential cycling demand passing through all road segments.
-This cumulative demand (referred to as *flow*) is then used as a basis for identifying segments that are in most need of investment in segregated cycling infrastructure.
-In doing so, we must account for the motivations and deterrents for cycling identified in Section \@ref(introduction), namely direct and well-connected routes.
-
-A range of algorithms could be used for prioritisation.
-Because policy priorities vary, we present two algorithms.
-Both utilize existing infrastructure from the beginning and allow us to compare a solution that focuses on utilitarianism to one that focuses on egalitarianism.
-In both algorithms, links are selected iteratively and the iteration at which each link is added to the solution is recorded.
-Investments in cycling infrastructure can be limited by budget constraints, so it can be useful to see where best to allocate a defined length of segregated infrastructure.
 
@@
 
-# Overarching Policies
-
-While segregated, connected, and direct cycling infrastructure is key to achieving high levels of cycling, research has shown that it cannot exist in a vacuum.
-@wardman2007factors developed a mode choice model for the UK and their results showed that improved cycling infrastructure on its own only had modest impacts on mode shift, and even the unlikely scenario of all urban routes being serviced by segregated bike lanes was forecast to increase cycling mode share by only 3%.
-International research shows that cities that invest in more comprehensive cycling projects have a more significant increase in the number of cyclists as well as the cycling mode share [@parkin2018designing; @pucher2010infrastructure].
-These cities do not just focus on infrastructure, but on general policies as well as restricting car use.
-Evaluation of policies in Denmark and Germany and the Netherlands has shown that their high cycling mode share is down to a broader set of policies that also include traffic calming, filtered permeability interventions, cycling rights of way, bike parking, integration with the public transport network, and making driving cars both expensive and inconvenient [@pucher2008making].
-While these policies are outside the scope of this research, it is important to recognize their key role in bringing about an increase in levels of cycling.
-
 # Disussion and conclusions
@@
 This paper provides a methodology for prioritizing road segments for investments in cycling infrastructure, with the output being a proposed cycling network for the city under study. 
-The approach aims to respect both the needs of the users and the ambitions of stakeholders working at local/regional levels.
+The approach aims to respect both the needs of the users and the ambitions of stakeholders working at local or regional levels.
 The results, primarily detailed route network maps based on current travel behaviour derived from OD data, can provide an evidence-based starting point for introducing a direct, connected, and low-stress network for any city by leveraging available open data.
@@
 -->
+A core part of the methodology is determining where there is a high potential for cycling, and using this as a basis for recommendations on road space reallocation in order to unlock potential cycling demand.
+The framework provides a routing engine to study the effects of limiting access on different road types, finding that reducing cyclist flow on roads with high through-traffic results in acceptable increases in commuting distances.
+As a result, a hierarchy of road preference needs to be defined to route potential cycling demand.
+The framework provides different algorithms to determine where investments in cycling infrastructure should be prioritized.
+These algorithms, based on connectivity and egalitarian principles of resource distribution, ensured that whatever the level of investment, the resulting cycling network would improve connectivity of existing cycling infrastructure.
 
@@
 
-A core part of the methodology is determining where there is a high potential for cycling, and using this as a basis for recommendations on road space reallocation in order to unlock potential cycling demand.
-The framework provides a routing engine to study the effects of limiting access on different road types, finding that reducing cyclist flow on roads with high through-traffic results in acceptable increases in commuting distances.
-As a result, a hierarchy of road preference needs to be defined to route potential cycling demand.
-The framework provides different algorithms to determine where investments in cycling infrastructure should be prioritized.
-These algorithms, based on connectivity and egalitarian principles of resource distribution, ensured that whatever the level of investment, the resulting cycling network would improve connectivity of existing cycling infrastructure.
+The approach is also focused solely on the allocation of cycling infrastructure, and does not consider the larger political and regulatory environment that needs to exist to promote cycling; while segregated, connected, and direct cycling infrastructure is key to achieving high levels of cycling, research has shown that it cannot exist in a vacuum.
+@wardman2007factors developed a mode choice model for the UK and their results showed that improved cycling infrastructure on its own only had modest impacts on mode shift, and even the unlikely scenario of all urban routes being serviced by segregated bike lanes was forecast to increase cycling mode share by only 3%.
+International research shows that cities that invest in more comprehensive cycling projects have a more significant increase in the number of cyclists as well as the cycling mode share [@parkin2018designing; @pucher2010infrastructure].
+These cities do not just focus on infrastructure, but on general policies as well as restricting car use.
+Evaluation of policies in Denmark and Germany and the Netherlands has shown that their high cycling mode share is down to a broader set of policies that also include traffic calming, filtered permeability interventions, cycling rights of way, bike parking, integration with the public transport network, and making driving cars both expensive and inconvenient [@pucher2008making].
+While these policies are outside the scope of this research, it is important to recognize their key role in bringing about an increase in levels of cycling.
 

### 2021-02-02 README.Rmd
transport geog citations

@@
 We find that flexible routing profiles that are adjustable to local needs can avoid road segments that are stressful to cycle on, enabling the identification of networks that balance directness against the need for safe and stress-free routes.
-We assess two different policy interventions , 'utilitarian' and 'egalitarian' by analyzing the connectivity of the city through community detection.
+We assess two different policy interventions , 'utilitarian' and 'egalitarian', by analyzing the connectivity of the city through community detection.
 From a policy perspective, our findings indicate that the 'egalitarian' prioritisation model approach is the more adequate to support cost-effective strategic cycle network planning.
@@
       extra_dependencies: ["subcaption", "longtable", "float", "multirow", "placeins", "multicol"]
+editor_options: 
+  markdown: 
+    wrap: 72
 ---
@@
 <!-- - Intro: possible trim down substantially then add some literature in (a) Calculating Potential Demand [DONE]/ (b) Routing [DONE]/ (c) Road Segment Prioritization -->
+
 <!-- - Community Detection: Where in the document should this section be? It is now before Road Segment Prioritization, but I can merge them -->
+
 <!-- Can go into road segment section. -->
+
 <!-- - Should we use other cities for the sake of comparison (compare communities, road types etc) -->
+
 <!-- - How to add tables properly (they are now added using latex code which only shows up when outputting to pdf) -->
+
 <!-- - Footnotes are sometimes split between pages. How do I keep them on the same page? -->
+
 <!-- - What is the best way to write pseudocode (Steps for Algorithm 1 and 2) -->
@@
 
-The 2015 Paris agreement [@agreement2015paris] acknowledged that fundamental changes to societies and economies are necessary to mitigate climate change.
-Like other sectors, transport is under substantial pressure to decarbonise resulting in a number of technical innovations including electric vehicles.
-But new vehicle technologies can only go so far and do not tackle parallel problems such as congestion, road traffic casualties and physical inactivity [@brand_climate_2020].
+The 2015 Paris agreement [@agreement2015paris] acknowledged that
+fundamental changes to societies and economies are necessary to mitigate
+climate change. Like other sectors, transport is under substantial
+pressure to decarbonise resulting in a number of technical innovations
+including electric vehicles. But new vehicle technologies can only go so
+far and do not tackle parallel problems such as congestion, road traffic
+casualties and physical inactivity [@brand_climate_2020].
 
 In this context, interest and investment in active modes are growing.
-The benefits extend beyond congestion and the environment, as active travel can also help alleviate what is referred to as the pandemic of global inactivity; physical inactivity is on the rise and has become the 4th highest cause of death globally [@kohl2012pandemic].
-Various studies have documented the association between active transport and lower risk of disease, including cancer and cardiovascular disease [@celis2017association; @jarrett2012effect; @patterson2020associations].
-In the wake of the Covid-19 pandemic, and the resulting reduced capacity of public transport, the UK government has pledged to invest billions of pounds to improve walking and cycling infrastructure across the country.
-While this unprecedented sum is an opportunity to reshape cities in a way that improves the well-being of citizens, it does come with a warning:
+The benefits extend beyond congestion and the environment, as active
+travel can also help alleviate what is referred to as the pandemic of
+global inactivity; physical inactivity is on the rise and has become the
+4th highest cause of death globally [@kohl2012pandemic]. Various
+studies have documented the association between active transport and
+lower risk of disease, including cancer and cardiovascular disease
+[@celis2017association; @jarrett2012effect;
+@patterson2020associations]. In the wake of the Covid-19 pandemic, and
+the resulting reduced capacity of public transport, the UK government
+has pledged to invest billions of pounds to improve walking and cycling
+infrastructure across the country. While this unprecedented sum is an
+opportunity to reshape cities in a way that improves the well-being of
+citizens, it does come with a warning:
 
-> "Inadequate cycling infrastructure discourages cycling and wastes public money.
-Much cycling infrastructure in this country is inadequate.
-It reflects a belief, conscious or otherwise, that hardly anyone cycles, that cycling is unimportant and that cycles must take no meaningful space from more important road users, such as motor vehicles and pedestrians - [@departmentgearchange2020]"
+> "Inadequate cycling infrastructure discourages cycling and wastes
+> public money. Much cycling infrastructure in this country is
+> inadequate. It reflects a belief, conscious or otherwise, that hardly
+> anyone cycles, that cycling is unimportant and that cycles must take
+> no meaningful space from more important road users, such as motor
+> vehicles and pedestrians - [@departmentgearchange2020]"
 
-The funding on its own is therefore no guarantee of a change in commuting across the country; it must be used to design adequate cycling infrastructure that is based on motivators and deterrents to cycling.
+The funding on its own is therefore no guarantee of a change in
+commuting across the country; it must be used to design adequate cycling
+infrastructure that is based on motivators and deterrents to cycling.
 
@@
 
-A considerable amount of research has been done towards that end.
-Segregated cycling infrastructure^[*Segregated cycling infrastructure* refers to road space that is allocated to cyclists only, with physical separation to protect cyclists from other modes of transport.] has been shown to increase cycling uptake [@aldred2019impacts; @goodman2014new; @marques2015infrastructure], with the separation from motorized vehicles being key [@winters2011motivators].
-Revealed preference of cyclists shows that they are willing to deviate from the most efficient routes in order to commute on safer roads [@crane2017longitudinal].
-However, such deviations are only considered if they do not considerably increase route circuitry; behaviour studies have found that the probability of choosing a route decreases in proportion to its length relative to the shortest route [@broach2011bicycle; @winters2010far].
-Another defining feature for cycling infrastructure is how well connected it is.
-Cyclists prefer cohesive infrastructure, particularly when cycling on arterial roads with high levels of motorized traffic [@stinson2003commuter].
-The lack of well-connected cycling infrastructure is one of the main obstacles to increasing cycling uptake [@caulfield2012determining].
-While direct and cohesive cycling networks have been shown to positively impact cycling rates, density^[making an area's bicycle network denser means adding more cycling routes in the area and thereby giving cyclists more route options] of the cycling network is also vital [@schoner2014missing].
+A considerable amount of research has been done to understand what gets people to cycle.
+Segregated cycling infrastructure^[*Segregated cycling infrastructure*
+refers to road space that is allocated to cyclists only, with physical
+separation to protect cyclists from other modes of transport.] has been
+shown to increase cycling uptake [@aldred2019impacts; @goodman2014new;
+@marques2015infrastructure], with the separation from motorized
+vehicles being key [@winters2011motivators]. Revealed preference of
+cyclists shows that they are willing to deviate from the most efficient
+routes in order to commute on safer roads [@crane2017longitudinal].
+However, such deviations are only considered if they do not considerably
+increase route circuitry; behaviour studies have found that the
+probability of choosing a route decreases in proportion to its length
+relative to the shortest route [@broach2011bicycle; @winters2010far].
+Another defining feature for cycling infrastructure is how well
+connected it is. Cyclists prefer cohesive infrastructure, particularly
+when cycling on arterial roads with high levels of motorized traffic
+[@stinson2003commuter]. The lack of well-connected cycling
+infrastructure is one of the main obstacles to increasing cycling uptake
+[@caulfield2012determining]. While direct and cohesive cycling networks
+have been shown to positively impact cycling rates, density^[making an
+area's bicycle network denser means adding more cycling routes in the
+area and thereby giving cyclists more route options] of the cycling
+network is also vital [@schoner2014missing].
 
@@
 
-The studies outlined above lay out the fundamentals for designing cycling networks that generate significant cycling uptake, but they do not propose network-level interventions.
-In this section we outline techniques used in past studies, namely optimization and network analysis techniques, such as connected components and community detection, and examine how they are leveraged to suggest cycling network designs.
-We compare the effectiveness of these network-level studies in incorporating the fundamentals outlined above.
-The methodology of this research is inspired by these techniques, but it attempts to add to them by ensuring that all of the outlined fundamentals are accounted for.
-It also goes further by attempting to factor in ethical considerations relating to distribution of investment.
+The studies outlined above lay out the fundamentals for designing
+cycling networks that generate significant cycling uptake, but they do
+not propose network-level interventions. In this section we outline
+techniques used in past studies, namely optimization and network
+analysis techniques, such as connected components and community
+detection, and examine how they are leveraged to suggest cycling network
+designs. We compare the effectiveness of these network-level studies in
+incorporating the fundamentals outlined above. The methodology of this
+research is inspired by these techniques, but it attempts to add to them
+by ensuring that all of the outlined fundamentals are accounted for. It
+also goes further by attempting to factor in ethical considerations
+relating to distribution of investment.
 
-*Optimization* techniques have been used to propose improvements to cycling networks.
-@mesbah2012bilevel propose a bi-level formulation to optimize allocation of cycling lanes to the network without exceeding a set budget.
-They account for the effect of cycling lanes on car traffic, and attempt to maximize utilization of said lanes with minimal impact on car travel times.
-Car usage is not considered by @mauttone2017bicycle, who develop an optimization framework that aims to minimize the total user cost of cycling on the network.
-The aggregate flow on links^[*link* refers to a road segment throughout this research] is obtained by using shortest paths to route existing cycling demand onto the road network, and the solution is a proposed set of links where cycling infrastructure should be added in order to minimize the overall travel cost of cyclists across the network.
-The cost of traversing a link is given as a function of its length and whether or not it has cycling infrastructure.
-The problem has also been solved by attempting to find the minimum cost of improving roadway links to meet a desired level of service (LOS) [@duthie2014optimization].
-In this formulation, all origin-destination (OD) pairs need to be connected by roads that meet the desired LOS.
-A limitation of these problem formulations do not explicitly solve for continuity, which is dealt with using either a constraint specifying that each link with a bike lane should be connected to at least one destination [@mesbah2012bilevel], a constraint on maximum deviation from shortest paths [@duthie2014optimization], or a discontinuity penalty to prioritize connected road segments [@mauttone2017bicycle].
+*Optimization* techniques have been used to propose improvements to
+cycling networks. @mesbah2012bilevel propose a bi-level formulation to
+optimize allocation of cycling lanes to the network without exceeding a
+set budget. They account for the effect of cycling lanes on car traffic,
+and attempt to maximize utilization of said lanes with minimal impact on
+car travel times.
+Safeguarding against increased car traffic may be counter-productive if the goal is to create a mode-shift, as research has shown that reducing road space for cars leads to less cars on the road, a phenomenon referred to as "traffic evaporation" [@nello2020environmental].
+Car usage is not considered by @mauttone2017bicycle,
+who develop an optimization framework that aims to minimize the total
+user cost of cycling on the network. The aggregate flow^[*flow* is
+used throughout this research to refer to the cycling demand when it is
+routed onto the road network. The flow on any road segment is the
+cumulative demand on it, resulting from cyclists commuting between
+various OD pairs] on links^[*link*
+refers to a road segment throughout this research] is obtained by using
+shortest paths to route existing cycling demand onto the road network,
+and the solution is a proposed set of links where cycling infrastructure
+should be added in order to minimize the overall travel cost of cyclists
+across the network. The cost of traversing a link is given as a function
+of its length and whether or not it has cycling infrastructure. The
+problem has also been solved by attempting to find the minimum cost of
+improving roadway links to meet a desired level of service (LOS)
+[@duthie2014optimization]. In this formulation, all origin-destination
+(OD) pairs need to be connected by roads that meet the desired LOS. A
+limitation of these problem formulations do not explicitly solve for
+continuity, which is dealt with using either a constraint specifying
+that each link with a bike lane should be connected to at least one
+destination [@mesbah2012bilevel], a constraint on maximum deviation
+from shortest paths [@duthie2014optimization], or a discontinuity
+penalty to prioritize connected road segments [@mauttone2017bicycle].
 
-In this paper, continuity is analysed by looking at the connectivity of the network through the graph-theoretic concept of *connected components*.
-Other academics have taken this approach; @natera2019data study the existing cycling network in terms of its disconnected components and introduce two different algorithms to connect these components by their most critical links and, in doing so, measure the size of the growth of the largest connected component as a function of the kilometers of network added.
-The concept of connected components is also at the core of the methodology proposed by @olmos2020data.
-After routing the cycling demand onto the network links, they use percolation theory to filter out the links based on the aggregate flow^[*flow* is used throughout this research to refer to the cycling demand when it is routed onto the road network.
-The flow on any road segment is the cumulative demand on it, resulting from cyclists commuting between various OD pairs] passing through them, varying the flow threshold for filtering to identify the minimum flow at which the whole city is connected by a giant component.
-While these approaches deal with continuity better, they look at the network as a whole when attempting to improve it, and in doing so fail to account for equitable distribution of infrastructure.
+In this paper, continuity is analysed by looking at the connectivity of
+the network through the graph-theoretic concept of *connected
+components*. Other academics have taken this approach; @natera2019data
+study the existing cycling network in terms of its disconnected
+components and introduce two different algorithms to connect these
+components by their most critical links and, in doing so, measure the
+size of the growth of the largest connected component as a function of
+the kilometers of network added. The concept of connected components is
+also at the core of the methodology proposed by @olmos2020data. After
+routing the cycling demand onto the network links, they use percolation
+theory to filter out the links based on the aggregate flow passing through them, varying the flow threshold for
+filtering to identify the minimum flow at which the whole city is
+connected by a giant component. While these approaches deal with
+continuity better, they look at the network as a whole when attempting
+to improve it, and in doing so fail to account for equitable
+distribution of infrastructure.
 
@@
 
-All of these network-level methodologies are underpinned by ethical principles, even though these principles are not explicitly acknowledged by the authors.
-This is important since different ethical principles constitute different problem formulations and targets.
-Broadly speaking, transport appraisal can be based on either utilitarian or egalitarian principles.
-The former seeks to maximize the overall benefit, while the latter is concerned with a fair distribution of benefits [@jafino2020transport].
-@nahmias2017integrating criticize the utilitarian approach that has been historically popular in the evaluation of transport investments, explaining how the maximization of overall benefit fails to account for the distribution of that benefit among communities or individuals.
-@lucas2016method explain how transport studies have traditionally looked at the bigger picture without studying the distribution of investments on the different parts of the study area, and go on to propose an egalitarian approach that ensures the disaggregation of transport policy benefits across the study area.
-@pereira2017distributive also emphasize the need for a more egalitarian approach to transport planning.
-They highlight accessibility as a cornerstone of distributive justice, and contend that policies should aim to distribute investments in a way that minimizes spatial variations in accessibility.
+All of these network-level methodologies are underpinned by ethical
+principles, even though these principles are not explicitly acknowledged
+by the authors. This is important since different ethical principles
+constitute different problem formulations and targets. Broadly speaking,
+transport appraisal can be based on either utilitarian or egalitarian
+principles. The former seeks to maximize the overall benefit, while the
+latter is concerned with a fair distribution of benefits
+[@jafino2020transport]. @nahmias2017integrating criticize the
+utilitarian approach that has been historically popular in the
+evaluation of transport investments, explaining how the maximization of
+overall benefit fails to account for the distribution of that benefit
+among communities or individuals. @lucas2016method explain how
+transport studies have traditionally looked at the bigger picture
+without studying the distribution of investments on the different parts
+of the study area, and go on to propose an egalitarian approach that
+ensures the disaggregation of transport policy benefits across the study
+area. @pereira2017distributive also emphasize the need for a more
+egalitarian approach to transport planning. They highlight accessibility
+as a cornerstone of distributive justice, and contend that policies
+should aim to distribute investments in a way that minimizes spatial
+variations in accessibility.
 
-This research attempts to propose an egalitarian framework for cycling network design.
-This is done by identifying the different sub-networks that exist within the larger network, and ensuring they all get a fair share of investment.
-Trip patterns in a city are not uniformly distributed geographically, and *community finding* methods have been used to partition study areas into localized areas that experience a disproportionate number of trips within them.
-@akbarzadeh2018designing use a modularity maximization approach [@blondel2008fast] on taxi trip data to identify 7 different communities in the city of Isfahan, Iran.
-An optimization problem is then formulated to connect nodes within each community with cycling infrastructure, with the emphasis being on connectivity within the communities, not between them.
-@bao2017planning adopt a similar methodology, first identifying communities and then using a greedy network expansion algorithm to simultaneously add links to each community.
-The link with the highest benefit-cost ratio in each community is selected, and the network is grown by adding neighboring links to the solution until a budget limit is met.
-The benefit is the flow on the link, and each link is assigned a cost based on current road conditions.
+This research attempts to propose an egalitarian framework for cycling
+network design. This is done by identifying the different sub-networks
+that exist within the larger network, and ensuring that each gets a fair
+share of investment. Trip patterns in a city are not uniformly
+distributed geographically, and *community finding* methods have been
+used to partition study areas into localized areas that experience a
+disproportionate number of trips within them. @akbarzadeh2018designing
+use a modularity maximization approach [@blondel2008fast] on taxi trip
+data to identify 7 different communities in the city of Isfahan, Iran.
+An optimization problem is then formulated to connect nodes within each
+community with cycling infrastructure, with the emphasis being on
+connectivity within the communities, not between them. @bao2017planning
+adopt a similar methodology, first identifying communities and then
+using a greedy network expansion algorithm to simultaneously add links
+to each community. The link with the highest benefit-cost ratio in each
+community is selected, and the network is grown by adding neighboring
+links to the solution until a budget limit is met. The benefit is the
+flow on the link, and each link is assigned a cost based on current road
+conditions.
 
 <!-- EDIT! Add more to limitations -->
-Our work builds on these *community finding* approaches by proposing a similar greedy network expansion algorithm for cycle network expansion within communities.
-We incorporate community finding methods for study area partitioning with weighted routing to avoid links that are stressful to cycle on.
-In doing so, we propose a methodology that accounts for motivators and deterrents to cycling.
-We propose three sub-methodologies that expand on some of the limitations of previous studies.
-These limitations include (a) bias inherent when basing network design solely on existing cycling demand, (b) proposing routes that may not correspond to studies on cyclist preference and government policies, and (3) insufficient consideration of the ethical principles underlying the analysis.
-Section \@ref(calculating-potential-cycling-demand) focuses on calculating potential cycling demand.
-Section \@ref(routing) focuses on routing the demand onto the road network while accounting for cyclist preferences and government priorities.
-Section \@ref(community-detection) outlines a methodology for partitioning the study area based on a community finding algorithm and routed cycling demand.
-Section \@ref(road-segment-prioritization) introduces the network expansion algorithms, and compares an approach grounded in ‘egalitarianism' to one grounded in ‘utilitarianism'.
+
+Our work builds on these *community finding* approaches by proposing a
+similar greedy network expansion algorithm for cycle network expansion
+within communities. We incorporate community finding methods for study
+area partitioning with weighted routing to avoid links that are
+stressful to cycle on. In doing so, we propose a methodology that
+accounts for motivators and deterrents to cycling. We propose three
+sub-methodologies that expand on some of the limitations of previous
+studies. These limitations include (a) bias inherent when basing network
+design solely on existing cycling demand, (b) proposing routes that may
+not correspond to studies on cyclist preference and government policies,
+and (3) insufficient consideration of the ethical principles underlying
+the analysis. Section \@ref(calculating-potential-cycling-demand)
+focuses on calculating potential cycling demand. Section \@ref(routing)
+focuses on routing the demand onto the road network while accounting for
+cyclist preferences and government priorities. Section
+\@ref(road-segment-prioritisation) outlines a methodology for partitioning the
+study area based on a community finding algorithm and routed cycling
+demand. It then introduces the network expansion algorithms, and compares an approach grounded in
+'egalitarianism' to one grounded in 'utilitarianism'.
 
@@
 
-Relying only on existing cycling activity to prioritise cycling infrastructure can reinforce existing cycling patterns and ignores potential cycling demand that could be satisfied by a connected network.
-To avoid this issue, @duthie2014optimization choose to ignore existing demand completely, and focus on creating a network that connects the entire study area.
-@olmos2020data obtain the distance distribution of cyclists using a smartphone-based bicycle GPS data, and then use a rejection-sampling algorithm on the OD data of the study area to match the potential demand distribution to the distribution obtained from GPS data.
+Relying only on existing cycling activity to prioritise cycling
+infrastructure can reinforce existing cycling patterns and ignores
+potential cycling demand that could be satisfied by a connected network.
+To avoid this issue, @duthie2014optimization choose to ignore existing
+demand completely, and focus on creating a network that connects the
+entire study area. @olmos2020data obtain the distance distribution of
+cyclists using a smartphone-based bicycle GPS data, and then use a
+rejection-sampling algorithm on the OD data of the study area to match
+the potential demand distribution to the distribution obtained from GPS
+data.
 
-*Origin-Destination (OD)* data can be obtained from a range of sources, including GPS data, household travels surveys and Census data on work locations.
-In areas where observed OD data is unavailable, modelling techniques such as spatial interaction models can be used to estimate travel volumes between zones [@black1995spatial; @martinez2013new; @wilson1971family].
-In this paper we use open access data from the UK census [@ONS2011flowdata], which contains aggregate statistics on number of commuters between administrative zones --- Middle layer Super Output Areas (MSOA) --- by mode of travel.
-MSOAs have an average population of just over 8000 [@ofn2018population].^[
-See https://wicid.ukdataservice.ac.uk/ for open access to the OD data.
-]
-Figures \@ref(fig:potdemhistograms) and \@ref(fig:desirefacetcycling) illustrates the proportion of trips cycled by distance and the geographic extent of the input OD dataset used in this paper.
+*Origin-Destination (OD)* data can be obtained from a range of sources,
+including GPS data, household travels surveys and Census data on work
+locations. In areas where observed OD data is unavailable, modelling
+techniques such as spatial interaction models can be used to estimate
+travel volumes between zones [@black1995spatial; @martinez2013new;
+@wilson1971family]. In this paper we use open access data from the UK
+census [@ONS2011flowdata], which contains aggregate statistics on
+number of commuters between administrative zones --- Middle layer Super
+Output Areas (MSOA) --- by mode of travel. MSOAs have an average
+population of just over 8000 [@ofn2018population].^[ See
+<https://wicid.ukdataservice.ac.uk/> for open access to the OD data.]
+Figures \@ref(fig:potdemhistograms) and \@ref(fig:desirefacetcycling)
+illustrates the proportion of trips cycled by distance and the
+geographic extent of the input OD dataset used in this paper.
 
-For our purposes, we use a logistic regression model to calculate potential cycling demand.
-The model was adapted from the Propensity to Cycle Tool (PCT), which estimates the proportion of trips ($\boldsymbol{C_{p}}$) for each OD pair that would cycle under different scenarios of change (we used the Government Target scenario, representing a nationwide target of doubling cycling by 2025) as a function of distance and hilliness [@lovelace2017propensity].
-The logistic regression model used to calculate $\boldsymbol{C_{p}}$ has the following parameters:
+For our purposes, we use a logistic regression model to calculate
+potential cycling demand. The model was adapted from the Propensity to
+Cycle Tool (PCT), which estimates the proportion of trips
+($\boldsymbol{C_{p}}$) for each OD pair that would cycle under
+different scenarios of change (we used the Government Target scenario,
+representing a nationwide target of doubling cycling by 2025) as a
+function of distance and hilliness [@lovelace2017propensity]. The
+logistic regression model used to calculate $\boldsymbol{C_{p}}$ has
+the following parameters:
 
@@
 
-\noindent where **d** and **s** are the distance and slope respectively for the OD pair.
-The authors use square and square-root distance terms "to capture the non-linear impact of distance on the likelihood of cycling", and interaction terms to capture the combined effect of slope and distance [@lovelace2017propensity].
-Alternative cycling uptake models could be used in different contexts or scenarios of change.
+\noindent where **d** and **s** are the distance and slope respectively
+for the OD pair. The authors use square and square-root distance terms
+"to capture the non-linear impact of distance on the likelihood of
+cycling", and interaction terms to capture the combined effect of slope
+and distance [@lovelace2017propensity]. Alternative cycling uptake
+models could be used in different contexts or scenarios of change.
 
-The potential demand calculations show that the current and potential number of cyclists both follow a bell-shaped distribution, with the number of trips peaking around the 3-4km commuting distance and then going back down for longer distances (see Figures \@ref(fig:potdemhistograms) and \@ref(fig:desirefacetcycling)).
+The potential demand calculations show that the current and potential
+number of cyclists both follow a bell-shaped distribution, with the
+number of trips peaking around the 3-4km commuting distance and then
+going back down for longer distances (see Figures
+\@ref(fig:potdemhistograms) and \@ref(fig:desirefacetcycling)).
 
@@
 
-```{r desirefacetcycling, fig.cap="Current and Potential Cycling Demand", out.width="100%", fig.show='hold', echo=FALSE}
+```{r desirefacetcycling, fig.cap="Current and Potential Cycling Demand", out.width="65%", fig.pos="H", fig.align="center", echo=FALSE}
 knitr::include_graphics("data/Manchester/Plots/desire_facet_cycling.png")
@@
 
-It should be noted that the calculations assume cycling in the traditional sense, i.e.
-the topography affects the decision to cycle.
-Recently there have been various micro-mobility solutions, including e-bikes, that allow commuters to traverse longer distances and hillier roads with less effort than traditional bicycles.
-While these modes would probably be associated with less geographical impedance, it is beyond the scope of this work to integrate those modes into the analysis.
-This is partially restricted by the lack of data on the proliferation of these modes, raising the point that perhaps the census data category of `Bicycle' is too vague, and should be further disaggregated to distinguish between traditional bicycles and other forms of micro-mobility.
+It should be noted that the calculations assume cycling in the
+traditional sense, i.e. the topography affects the decision to cycle.
+Recently there have been various micro-mobility solutions, including
+e-bikes, that allow commuters to traverse longer distances and hillier
+roads with less effort than traditional bicycles. While these modes
+would probably be associated with less geographical impedance, it is
+beyond the scope of this work to integrate those modes into the
+analysis. This is partially restricted by the lack of data on the
+proliferation of these modes, raising the point that perhaps the census
+data category of `Bicycle' is too vague, and should be further
+disaggregated to distinguish between traditional bicycles and other
+forms of micro-mobility.
 
@@
 
-The next step in our approach is to route the potential cycling demand ($\boldsymbol{C_{p}}$) between all OD pairs onto the road network.
-We choose not to use the PCT approach for routing as it is reliant on external and therefore inflexible routing services.
+The next step in our approach is to route the potential cycling demand
+($\boldsymbol{C_{p}}$) between all OD pairs onto the road network. We
+choose not to use the PCT approach for routing as it is reliant on
+external and therefore inflexible routing services.
 <!-- This expands on the work of @mauttone2017bicycle, by going beyond simply favoring roads with existing cycling infrastructure to creating a hierarchy of road preference. -->
@@
 
-1.
-**Cyclist Preference**: Work done by @dill2013four on examining cyclist typologies determined that around 60% of Portland residents fit under the *interested but concerned* category.
-These were people that enjoyed cycling but avoided it due safety concerns.
-The key to encouraging this group was to create a low-stress cycling network, not only though segregated infrastructure but also by planning routes that passed through residential streets.
-2.
-**Low-Traffic Neighborhoods**: The UK Department for Transport is allocating funding to local authorities to invest in Active Transport, partially through the creation of LTNs [@departmentgearchange2020].
-This includes closing off residential streets to motorized traffic.
-3.
-**Existing Cycling Infrastructure**: Utilizing existing cycling infrastructure makes economic sense, as small investments may lead to large connectivity gains as the disconnected cycling infrastructure gets joined together.
+1.  **Cyclist Preference**: Work done by @dill2013four on examining
+    cyclist typologies determined that around 60% of Portland residents
+    fit under the *interested but concerned* category. These were people
+    that enjoyed cycling but avoided it due safety concerns. The key to
+    encouraging this group was to create a low-stress cycling network,
+    not only though segregated infrastructure but also by planning
+    routes that passed through residential streets.
+2.  **Low-Traffic Neighborhoods**: The UK Department for Transport is
+    allocating funding to local authorities to invest in Active
+    Transport, partially through the creation of LTNs
+    [@departmentgearchange2020]. This includes closing off residential
+    streets to motorized traffic.
+3.  **Existing Cycling Infrastructure**: Utilizing existing cycling
+    infrastructure makes economic sense, as small investments may lead
+    to large connectivity gains as the disconnected cycling
+    infrastructure gets joined together.
 
-The above points are accounted for by using a weighted road network for routing.
-This has previously been done by multiplying all road segments without cycling infrastructure by a fixed impedance factor [@mauttone2017bicycle], or by assigning different weights to the road segments proportional to the investment cost of bringing them to an acceptable level of stress for cycling  [@duthie2014optimization].
-Perceived stress has been found to be a function of vehicular traffic volume and speed [@sorton1994bicycle], both of which vary predictably with road type. 
-@gehrke2020cycling build on this information to use routing impedance factors for road segments that are based on road type and the existence of cycling infrastructure.
+The above points are accounted for by using a weighted road network for
+routing. This has previously been done by multiplying all road segments
+without cycling infrastructure by a fixed impedance factor
+[@mauttone2017bicycle], or by assigning different weights to the road
+segments proportional to the investment cost of bringing them to an
+acceptable level of stress for cycling [@duthie2014optimization].
+Perceived stress for cyclists has been found to be a function of vehicular traffic
+volume and speed [@sorton1994bicycle], both of which vary predictably
+with road type. @gehrke2020cycling build on this information to use
+routing impedance factors for road segments that are based on road type
+and the existence of cycling infrastructure.
 
-For the purposes of this paper, we created a similar weighting profile that is adjusted to favor less stressful roads (based on the definition by @sorton1994bicycle and information from Table  \@ref(table:osmroadtypes)), and roads with existing cycling infrastructure.
-We believe this to be more practical than the simplistic approach adopted by @mauttone2017bicycle, as it goes beyond simply favoring roads with existing cycling infrastructure enabling a hierarchy of road preference based on perceived stress levels.
+For the purposes of this paper, we created a similar weighting profile
+that is adjusted to favor less stressful roads (based on the definition
+by @sorton1994bicycle and information from Table
+\@ref(table:osmroadtypes)), and roads with existing cycling
+infrastructure. We believe this to be more practical than the simplistic
+approach adopted by @mauttone2017bicycle, as it goes beyond simply
+favoring roads with existing cycling infrastructure enabling a hierarchy
+of road preference based on perceived stress levels.
 
@@
 <!-- **ADD TABLE - THIS IS BASIC** -->
+
 <!-- ```{r, echo=FALSE, message = FALSE} -->
+
 <!-- weight_profiles <- readxl::read_excel("Paper/paper_tables.xlsx", sheet = "Weighting Profile") -->
@@
 <!-- knitr::kable(weight_profiles, -->
+
 <!--              caption = "Weighting Profiles") -->
+
 <!-- ``` -->
+
 \begin{longtable}{lrrr}
@@
 
-\noindent A weighted distance $\boldsymbol{d_{w}}$ for each road segment is calculated as follows ^[The **dodgr** r package [@padgham2019dodgr] is used to route cycling demand onto the road network.
-The package uses the OpenStreetMaps (OSM) road network and allows the user to assign weights to roads based on their type.
-The routing is done based on weighted shortest paths, with the distance along each road segment being divided by a factor to obtain the weighted distance for routing.
-It is more intuitive to multiply when weighting a network, but the dodgr package divides by numbers between 0 and 1, which achieves the same result.
-For the sake of reproducibility, we stick to the convention used in the package.]:
+\noindent A weighted distance $\boldsymbol{d_{w}}$ for each road
+segment is calculated as follows ^[The **dodgr** r package
+[@padgham2019dodgr] is used to route cycling demand onto the road
+network. The package uses the OpenStreetMaps (OSM) road network and
+allows the user to assign weights to roads based on their type. The
+routing is done based on weighted shortest paths, with the distance
+along each road segment being divided by a factor to obtain the weighted
+distance for routing. It is more intuitive to multiply when weighting a
+network, but the dodgr package divides by numbers between 0 and 1, which
+achieves the same result. For the sake of reproducibility, we stick to
+the convention used in the package.]:
 
@@
 
-\noindent where $\boldsymbol{d_{unw}}$ is the unweighted distance and $\boldsymbol{W}$ is the weight from Table \ref{table:weightprofiles}.
+\noindent where $\boldsymbol{d_{unw}}$ is the unweighted distance and
+$\boldsymbol{W}$ is the weight from Table \ref{table:weightprofiles}.
 
-All weights are between 0 and 1, and the values in the \textit{Weighted} profile are chosen to be inversely proportional to the stress level experienced by cyclists on them.
-The \textit{Unweighted} weighting profile is used to compare increases in route length resulting from two different approaches:
+All weights are between 0 and 1, and the values in the
+\textit{Weighted} profile are chosen to be inversely proportional to
+the stress level experienced by cyclists on them. The
+\textit{Unweighted} weighting profile is used to compare increases in
+route length resulting from two different approaches:
 
-1. **Weighted**: Relatively high impedance on Primary and Trunk roads (to minimize cycling on them).
-2. **Weighted\_2**: Avoiding Primary and Trunk Roads completely.
+1.  **Weighted**: Relatively high impedance on Primary and Trunk roads
+    (to minimize cycling on them).
+2.  **Weighted_2**: Avoiding Primary and Trunk Roads completely.
 
-Comparing the cycling demand routed on the weighted and unweighted road network allows us to get a better understanding of the importance of different road types.
-In the case of Manchester, trunk roads bisect the city and are a major part of unweighted shortest paths (Figure \@ref(fig:flowsfacetunweighted)).
-On the other hand, cycleways are not part of unweighted shortest paths, and so very little of the cycling demand is routed through them.
-In the weighted network, cycleways are much better utilized, and the majority of the cycling demand passes through tertiary roads, as expected.
+Comparing the cycling demand routed on the weighted and unweighted road
+network allows us to get a better understanding of the importance of
+different road types. In the case of Manchester, trunk roads bisect the
+city and are a major part of unweighted shortest paths (Figure
+\@ref(fig:flowsfacetunweighted)). On the other hand, cycleways are not
+part of unweighted shortest paths, and so very little of the cycling
+demand is routed through them. In the weighted network, cycleways are
+much better utilized, and the majority of the cycling demand passes
+through tertiary roads, as expected.
 
@@
 
-The results of routing potential cycling demand on the weighted and unweighted networks are understandably quite different.
-From Figure \@ref(fig:flowsfacetunweighted) we can see that trunk and primary roads are the most efficient means of traversing the road network of Manchester.
-Both of these road types are classified as Primary A roads according to the UK Department for Transport's road classification (Table \@ref(table:osmroadtypes)), and are therefore part of the Primary Route Network (PRN) [@department2012guidance].
-The PRN has the widest, most direct roads on the network, and carries most of the through traffic.
-This includes freight, with all roads in the PRN being required by law to provide unrestricted access to trucks up to 40 tonnes [@department2012guidance].
+The results of routing potential cycling demand on the weighted and
+unweighted networks are understandably quite different. From Figure
+\@ref(fig:flowsfacetunweighted) we can see that trunk and primary roads
+are the most efficient means of traversing the road network of
+Manchester. Both of these road types are classified as Primary A roads
+according to the UK Department for Transport's road classification
+(Table \@ref(table:osmroadtypes)), and are therefore part of the Primary
+Route Network (PRN) [@department2012guidance]. The PRN has the widest,
+most direct roads on the network, and carries most of the through
+traffic. This includes freight, with all roads in the PRN being required
+by law to provide unrestricted access to trucks up to 40 tonnes
+[@department2012guidance].
 
-We choose to avoid routing the potential cycling demand on Primary A Roads for the following 2 reasons:
+We choose to avoid routing the potential cycling demand on Primary A
+Roads for the following 2 reasons:
 
-1. **Logistical Difficulty**: Changes on these roads need to be agreed upon by all affected authorities [@department2012guidance], which may prove to be difficult. 
-2.
-**Low Traffic Neighborhoods (LTNs)**: The UK government is aiming to restrict access to motorized vehicles on residential roads to create LTNs [@departmentgearchange2020].
-This is part of a policy to prevent automobile rat-running and make streets more accessible to cyclists and pedestrians.
-Under such a policy, Primary A roads would become even more essential for motorized traffic and it would be more difficult to reallocate road space on these roads to cyclists.
+1.  **Logistical Difficulty**: Changes on these roads need to be agreed
+    upon by all affected authorities [@department2012guidance], which
+    may prove to be difficult.
+2.  **Low Traffic Neighborhoods (LTNs)**: The UK government is aiming to
+    restrict access to motorized vehicles on residential roads to create
+    LTNs [@departmentgearchange2020]. This is part of a policy to
+    prevent automobile rat-running and make streets more accessible to
+    cyclists and pedestrians. Under such a policy, Primary A roads would
+    become even more essential for motorized traffic and it would be
+    more difficult to reallocate road space on these roads to cyclists.
 
-Figure \@ref(fig:flowsfacetweighted) shows that routing on the weighted network significantly reduces flow on the trunk and primary roads, but does not eliminate it completely.
-This is intentional, as the impedance on these roads is only slightly higher than remaining road types (See Table \@ref(table:weightprofiles)).
-Potential cycling demand is only routed on these roads if there are no routes through other roads that offer comparable directness.
+Figure \@ref(fig:flowsfacetweighted) shows that routing on the weighted
+network significantly reduces flow on the trunk and primary roads, but
+does not eliminate it completely. This is intentional, as the impedance
+on these roads is only slightly higher than remaining road types (See
+Table \@ref(table:weightprofiles)). Potential cycling demand is only
+routed on these roads if there are no routes through other roads that
+offer comparable directness.
 
-Banning cycling flow completely on trunk and primary roads may result in excessively circuitous paths, as seen in Figure \@ref(fig:boxplotcircuity).
-When routing using the *weighted* weighting profile in Table \@ref(table:weightprofiles), we see that shortest paths increase by less than 5\% on average from unweighted shortest paths, with the largest increases still below 30\%.
-When routing on primary and trunk roads is banned (*weighted_2* profile in Table \@ref(table:weightprofiles)), the average increase relative to unweighted shortest paths rises to 10\%, with certain locations experiencing more significant negative effects on accessibility.
-Given that cyclists will only deviate from shortest paths by a certain amount to access better cycling infrastructure (as explained in Section \@ref(introduction)), allowing flow on some stretches of trunk and primary roads is necessary to insure cycling uptake and equitable access to cycling infrastructure.
-In its new vision for walking and cycling, the Department for Transport acknowledges that minimal segregated stretches of bicycle lanes on main roads will be necessary to avoid circuitous cycling networks [@departmentgearchange2020].
+Banning cycling flow completely on trunk and primary roads may result in
+excessively circuitous paths, as seen in Figure
+\@ref(fig:boxplotcircuity). When routing using the *weighted* weighting
+profile in Table \@ref(table:weightprofiles), we see that shortest paths
+increase by less than 5% on average from unweighted shortest paths, with
+the largest increases still below 30%. When routing on primary and trunk
+roads is banned (*weighted_2* profile in Table
+\@ref(table:weightprofiles)), the average increase relative to
+unweighted shortest paths rises to 10%, with certain locations
+experiencing more significant negative effects on accessibility. Given
+that cyclists will only deviate from shortest paths by a certain amount
+to access better cycling infrastructure (as explained in Section
+\@ref(introduction)), allowing flow on some stretches of trunk and
+primary roads is necessary to insure cycling uptake and equitable access
+to cycling infrastructure. In its new vision for walking and cycling,
+the Department for Transport acknowledges that minimal segregated
+stretches of bicycle lanes on main roads will be necessary to avoid
+circuitous cycling networks [@departmentgearchange2020].
 
@@
 
-Weighting the road network also allows us to better utilize existing cycling infrastructure, as can be seen by the higher flow on cycleways in Figure \@ref(fig:flowsfacetweighted).
-Again, the small differences in impedance between cycleways and other road types mean that cycleways that require significant deviation are not routed on.
+Weighting the road network also allows us to better utilize existing
+cycling infrastructure, as can be seen by the higher flow on cycleways
+in Figure \@ref(fig:flowsfacetweighted). Again, the small differences in
+impedance between cycleways and other road types mean that cycleways
+that require significant deviation are not routed on.
 
-It should be reiterated that the weighting profile used for routing has been developed for the purposes of this study.
-It creates a hierarchy of road preference that is grounded in cyclist preferences and government plans to create LTNs.
-A wide range of weighting profiles could be used, to represent different types of cyclists and road environments, as described in @gehrke2020cycling and @furth2016network.
-Sensitivity analysis could be done to determine an optimal weighting profile, but given the variation in city road networks<!-- (Figure \ref{fig:perc_person-km}) -->, this would probably require calibration to the specific city.
-More accurate routing could be carried out given the availability of road-level data.
-In such cases we would add additional impedance to specific roads, giving more useful routing results than the current methodology which considers all roads of the same type to be equivalent.
+It should be reiterated that the weighting profile used for routing has
+been developed for the purposes of this study. It creates a hierarchy of
+road preference that is grounded in cyclist preferences and government
+plans to create LTNs. A wide range of weighting profiles could be used
+to represent different types of cyclists and road environments, as
+described in @gehrke2020cycling and @furth2016network. Sensitivity
+analysis could be done to determine an optimal weighting profile, but
+given the variation in city road
+networks<!-- (Figure \ref{fig:perc_person-km}) -->, this would probably
+require calibration to the specific city. More accurate routing could be
+carried out given the availability of road-level data. In such cases we
+would add additional impedance to specific roads, giving more useful
+routing results than the current methodology which considers all roads
+of the same type to be equivalent.
 
-One use-case of such granular data would be to identify roads that serve schools.
-The Department of Transport notes that the number of school children being driven to school has trebled over the past 40 years [@departmentgearchange2020], and so having cycling infrastructure serving schools is key to achieving the government target of getting more children to cycle.
-This would not be difficult, as over 75\% of children in the UK live within a 15 minute cycle from their school [@departmentcycleinfradesign2020].
-@goodman2019scenarios show that if dutch levels of cycling were achieved in the UK, the \% of children cycling to school could increase from 1.8\% to 41\%.
-In their typology of cyclists, @dill2013four found that a majority of people who say they would never cycle had never cycled to school, whereas confident cyclists were those most likely to have cycled to school.
-Getting people to cycle from a young age is therefore key to achieving societal change in commuting habits.
+One use-case of such granular data would be to identify roads that serve
+schools. The Department of Transport notes that the number of school
+children being driven to school has trebled over the past 40 years
+[@departmentgearchange2020], and so having cycling infrastructure
+serving schools is key to achieving the government target of getting
+more children to cycle. This would not be difficult, as over 75% of
+children in the UK live within a 15 minute cycle from their school
+[@departmentcycleinfradesign2020]. @goodman2019scenarios show that if
+dutch levels of cycling were achieved in the UK, the % of children
+cycling to school could increase from 1.8% to 41%. In their typology of
+cyclists, @dill2013four found that a majority of people who say they
+would never cycle had never cycled to school, whereas confident cyclists
+were those most likely to have cycled to school. Getting people to cycle
+from a young age is therefore key to achieving societal change in
+commuting habits.
 
-# Road Segment Prioritization
+# Road Segment Prioritisation
 
@@
 
-After routing the potential cycling demand onto the road network using weighted shortest paths, we have estimates for the cumulative potential cycling demand passing through all road segments.
-This cumulative demand (referred to as *flow*) is then used as a basis for identifying segments that are in most need of investment in segregated cycling infrastructure.
-In doing so, we must account for the motivations and deterrents for cycling identified in Section \@ref(introduction), namely direct and well-connected routes.
+After routing the potential cycling demand onto the road network using
+weighted shortest paths, we have estimates for the cumulative potential
+cycling demand passing through all road segments. This cumulative demand
+(referred to as *flow*) is then used as a basis for identifying segments
+that are in most need of investment in segregated cycling
+infrastructure. In doing so, we must account for the motivations and
+deterrents for cycling identified in Section \@ref(introduction), namely
+direct and well-connected routes.
 
-A range of algorithms could be used for prioritisation.
-Because policy priorities vary, we present two algorithms.
-Both utilize existing infrastructure from the beginning and allow us to compare a solution that focuses on utilitarianism to one that focuses on egalitarianism.
-In both algorithms, links are selected iteratively and the iteration at which each link is added to the solution is recorded.
-Investments in cycling infrastructure can be limited by budget constraints, so it can be useful to see where best to allocate a defined length of segregated infrastructure.
-In order to incorporate egalitarian principles in our approach, we use community detection to partition the study area and evaluate the distribution of investment over the different subdivisions.
+A range of algorithms could be used for prioritisation. Because policy
+priorities vary, we present two algorithms. Both utilize existing
+infrastructure from the beginning and allow us to compare a solution
+that focuses on utilitarianism to one that focuses on egalitarianism. In
+both algorithms, links are selected iteratively and the iteration at
+which each link is added to the solution is recorded. Investments in
+cycling infrastructure can be limited by budget constraints, so it can
+be useful to see where best to allocate a defined length of segregated
+infrastructure. In order to incorporate egalitarian principles in our
+approach, we use community detection to partition the study area and
+evaluate the distribution of investment over the different subdivisions.
 
@@
 
-A major challenge facing ‘top-down’ planning approaches is how to incorporate egalitarian principles by fairly distributing investments in cycling infrastructure. 
-One way of quantifying this is to split up the city into smaller geospatial areas and target equal investment in each of those areas.
-This approach could also help ensure that on-the-ground surveys are made by local stakeholders, an important component of the planning process  [@parkin2018designing].
-Community detection offers us a way to delineate such a split; cyclists are limited in their commuting distance (see Figure \@ref(fig:cyclinghistmanc)), and so trip attractors are more likely to have a local catchment area of cyclists.
+As explained in Section
+\@ref(ethical-underpinnings-and-proposed-approach), a major challenge
+facing 'top-down' planning approaches is how to incorporate egalitarian
+principles by fairly distributing investments in cycling infrastructure.
+One way of quantifying this is to split up the city into smaller
+geospatial areas and target equal investment in each of those areas.
+This approach could also help ensure that on-the-ground surveys are made
+by local stakeholders, an important component of the planning process
+[@parkin2018designing]. Community detection offers us a way to
+delineate such a split; cyclists are limited in their commuting distance
+(see Figure \@ref(fig:cyclinghistmanc)), and so trip attractors are more
+likely to have a local catchment area of cyclists.
 
@@
 
-In our case, the network is the city; the nodes are the population-weighted MSOA centroids and the links connecting each MSOA pair are weighted by the potential cycling demand between them.
-The Louvain method  [@blondel2008fast] is used to separate MSOAs into communities.
-Potential cycling demand is used since we assume that this is what the cycling demand will be once the cycling infrastructure is added.
-To assign road links to communities, the following steps are carried out:
+In our case, the network is the city; the nodes are the
+population-weighted MSOA centroids and the links connecting each MSOA
+pair are weighted by the potential cycling demand between them. The
+Louvain method [@blondel2008fast] is used to separate MSOAs into
+communities. Potential cycling demand is used since we assume that this
+is what the cycling demand will be once the cycling infrastructure is
+added. To assign road links to communities, the following steps are
+carried out:
 
@@
 3. Assign each road link to the same community as the closest MSOA centroid to it. -->
+
 ```{r, eval = FALSE}
@@
 
-The results show that Manchester can be split into four large communities and one small one (Figure \@ref(fig:communitiesmanchester)).
+The results show that Manchester can be split into four large
+communities and one small one (Figure \@ref(fig:communitiesmanchester)).
 
-```{r communitiesmanchester, fig.cap="Communities Based on Potential Cycling Demand (Manchester)", out.width="75%", fig.show='hold', echo=FALSE}
+```{r communitiesmanchester, fig.cap="Communities Based on Potential Cycling Demand (Manchester)", out.width="75%", fig.show='hold', fig.align = "center", echo=FALSE}
 knitr::include_graphics("data/Manchester/Plots/communities_alternative_Manchester.png")
@@
 
-This algorithm ensures that the resulting network is connected.
-It also satisfies the directness criteria, since links on the weighted shortest paths are those that have the highest flow passing through them (this is a result of the routing in Section \@ref(introduction).
+This algorithm ensures that the resulting network is connected. It also
+satisfies the directness criteria, since links on the weighted shortest
+paths are those that have the highest flow passing through them (this is
+a result of the routing in Section \@ref(introduction).
 
-## Algorithm 2: Egalitarian Expansion (Focus on Fair Distribution of Resources)
+## Algorithm 2: Egalitarian Expansion
 
-The first algorithm focuses on connectivity and directness, but not on fairly distributing investment.
-The latter is not a requirement for increasing cycling uptake, but it is fundamental for spatial equity, as explained in Section \@ref(introduction).
-This algorithm incorporates the ideal of fair distribution by using community detection to partition the road network.
+The first algorithm focuses on connectivity and directness, but not on
+fairly distributing investment. The latter is not a requirement for
+increasing cycling uptake, but it is fundamental for spatial equity, as
+explained in Section \@ref(ethical-underpinnings-and-proposed-approach). This algorithm incorporates
+the ideal of fair distribution by using community detection to partition
+the road network.
 
-The algorithm uses the following logic to ensure fair distribution between communities:
+The algorithm uses the following logic to ensure fair distribution
+between communities:
 
@@
 
-Even though we may end up with a more disconnected network, we will have separate connected networks in each community.
-Given that communities are defined by having more internal flow than external flow, this is a satisfactory solution.
+Even though we may end up with a more disconnected network, we will have
+separate connected networks in each community. Given that communities
+are defined by having more internal flow than external flow, this is a
+satisfactory solution.
 
-The results of the community detection are used to evaluate the algorithms.
-This is done by looking at the *person-km satisfied* as cycling infrastructure is added.
-Person-km is a measure of the total km cycled on a road segment, so it is the product of the number of potential commuters cycling on that road segment ($flow$) and the length of the segment in km ($l$).
-For each road segment, the person-km is equal to $flow * l$.
-In the case of Manchester, Table \@ref(tab:personkmtable) shows that almost half of the person-km is in community 1 , while only 0.5\% of total person-km on the network is in community 5.
+The results of the community detection are used to evaluate the
+algorithms. This is done by looking at the *person-km satisfied* as
+cycling infrastructure is added. Person-km is a measure of the total km
+cycled on a road segment, so it is the product of the number of
+potential commuters cycling on that road segment ($flow$) and the length
+of the segment in km ($l$). For each road segment, the person-km is
+equal to $flow * l$. In the case of Manchester, Table
+\@ref(tab:personkmtable) shows that almost half of the person-km is in
+community 1 , while only 0.5% of total person-km on the network is in
+community 5.
 
@@
 
-Looking at the person-km satisfied (Figure \@ref(fig:growthtotal)), we see that the incremental addition of cycling infrastructure is better distributed between communities using Algorithm 2; equal distribution of investment results in the gain in % of person km satisfied in each community being inversely correlated with the size of the community.
-In addition, we find that the restrictions imposed by Algorithm 2 on the network expansion do not seem to have a noticeable effect on the city-wide \% of person-km satisfied.
-Comparing both algorithms, we can see that Algorithm 1 provides only marginally quicker city-wide gains than Algorithm 2.
+Looking at the person-km satisfied (Figure \@ref(fig:growthtotal)), we
+see that the incremental addition of cycling infrastructure is better
+distributed between communities using Algorithm 2; equal distribution of
+investment results in the gain in % of person km satisfied in each
+community being inversely correlated with the size of the community. In
+addition, we find that the restrictions imposed by Algorithm 2 on the
+network expansion do not seem to have a noticeable effect on the
+city-wide % of person-km satisfied. Comparing both algorithms, we can
+see that Algorithm 1 provides only marginally quicker city-wide gains
+than Algorithm 2.
 
-```{r growthtotal, fig.cap="Comparing Overall (Dashed) and Community Level Person-Km Satisfied (Manchester)", out.width="35%", fig.show='hold', echo=FALSE}
+```{r growthtotal, fig.cap="Comparing Overall (Dashed) and Community Level Person-Km Satisfied (Manchester)", out.width="35%", fig.show='hold', fig.align = "center", echo=FALSE}
 knitr::include_graphics(c(
@@
 
-Figure \@ref(fig:growth3MapandBar) gives us a geographic representation of the results from Algorithm 2; it shows when each link was added to the solution (first 100km, second 100km, etc).
-We can see that, generally, road segments around cycling infrastructure are prioritized, except for those neighboring cycling infrastructure on the very periphery.
-The first 100km is also spatially distributed across the city, with no apparent bias towards a particular area.
+Figure \@ref(fig:growth3MapandBar) gives us a geographic representation
+of the results from Algorithm 2; it shows when each link was added to
+the solution (first 100km, second 100km, etc). We can see that,
+generally, road segments around cycling infrastructure are prioritized,
+except for those neighboring cycling infrastructure on the very
+periphery. The first 100km is also spatially distributed across the
+city, with no apparent bias towards a particular area.
 
-It is also important to understand how the different highway types contribute to the proposed network.
-Figure \@ref(fig:growth3MapandBar) shows that most of the flow will be on residential and tertiary roads, as expected from the weighting profile defined in Table \@ref(table:weightprofiles).
+It is also important to understand how the different highway types
+contribute to the proposed network. Figure \@ref(fig:growth3MapandBar)
+shows that most of the flow will be on residential and tertiary roads,
+as expected from the weighting profile defined in Table
+\@ref(table:weightprofiles).
 
-```{r growth3MapandBar, fig.cap="Results of Alg. 2 (Manchester)", out.width="35%", fig.show='hold', echo=FALSE}
+```{r growth3MapandBar, fig.cap="Road Segment Priority (left), disaggregated by road type (right) - Egalitarian Growth", out.width="35%", fig.show='hold', fig.align="center", echo=FALSE}
 knitr::include_graphics(c(
@@
 
-Existing cycling infrastructure is made up of many disconnected components.
-Both Algorithm 1 and 2 start with all existing segregated cycling infrastructure and aim to create an efficient, connected network.
-Figure \@ref(fig:componentsandGCC) shows that both algorithms gradually reduce the number of components as more infrastructure is added, but Algorithm 2 is able to provide better connectivity with less investment.
+Existing cycling infrastructure is made up of many disconnected
+components. Both Algorithm 1 and 2 start with all existing segregated
+cycling infrastructure and aim to create an efficient, connected
+network. Figure \@ref(fig:componentsandGCC) shows that both algorithms
+gradually reduce the number of components as more infrastructure is
+added, but Algorithm 2 is able to provide better connectivity with less
+investment.
 
-Consistent growth can also be seen for the size of the Largest Connected Component in the proposed bicycle network (Figure \@ref(fig:componentsandGCC)).
-Here however, we find that there is little difference between both Algorithms.
+Consistent growth can also be seen for the size of the Largest Connected
+Component in the proposed bicycle network (Figure
+\@ref(fig:componentsandGCC)). Here however, we find that there is little
+difference between both Algorithms.
 
-```{r componentsandGCC, fig.cap="Network Characteristics", out.width="35%", fig.show='hold', echo=FALSE}
+```{r componentsandGCC, fig.cap="Network Characteristics", out.width="30%", fig.show='hold', fig.pos="H", fig.align = "center", echo=FALSE}
 knitr::include_graphics(c(
@@
 
-# Disussion and conclusions
+# Discussion and conclusions
 
-This paper provides a methodology for prioritizing road segments for investments in cycling infrastructure, with the output being a proposed cycling network for the city under study. 
-The approach aims to respect both the needs of the users and the ambitions of stakeholders working at local or regional levels.
-The results, primarily detailed route network maps based on current travel behaviour derived from OD data, can provide an evidence-based starting point for introducing a direct, connected, and low-stress network for any city by leveraging available open data.
-Given that the “most essential activity entailed in the design of cycle-friendly infrastructure is developing a cycle network” [@parkin2018designing], we believe that the methods have great potential to inform investment in cities such as Manchester where there is political will to invest in cycling long-term.
-A benefit of the approach is that it has relatively modest data requirements: only the road network (from OSM), its topography (from satellite imagery), and origin-destination data (from the national census in this case) are needed, opening up the possibility of deploying the approach in other cities^[The results are easily reproducible for all UK cities, and can also be reproduced for cities elsewhere given the availability of commuter data.
-For instructions on reproducing the results shown in this paper, see the **README** file in the R folder of the .zip file that accompanies this paper (to be released as open source software when the paper is published).].
+This paper provides a methodology for prioritizing road segments for
+investments in cycling infrastructure, with the output being a proposed
+cycling network for the city under study. The approach aims to respect
+both the needs of the users and the ambitions of stakeholders working at
+local or regional levels. The results, primarily detailed route network
+maps based on current travel behaviour derived from OD data, can provide
+an evidence-based starting point for introducing a direct, connected,
+and low-stress network for any city by leveraging available open data.
+Given that the "most essential activity entailed in the design of
+cycle-friendly infrastructure is developing a cycle network"
+[@parkin2018designing], we believe that the methods have great
+potential to inform investment in cities such as Manchester where there
+is political will to invest in cycling long-term. A benefit of the
+approach is that it has relatively modest data requirements: only the
+road network (from OSM), its topography (from satellite imagery), and
+OD data (from the national census in this case) are
+needed, opening up the possibility of deploying the approach in other
+cities^[The results are easily reproducible for all UK cities, and can
+also be reproduced for cities elsewhere given the availability of
+commuter data. For instructions on reproducing the results shown in this
+paper, see the **README** file in the R folder of the .zip file that
+accompanies this paper (to be released as open source software when the
+paper is published).].
 
@@
 -->
-A core part of the methodology is determining where there is a high potential for cycling, and using this as a basis for recommendations on road space reallocation in order to unlock potential cycling demand.
-The framework provides a routing engine to study the effects of limiting access on different road types, finding that reducing cyclist flow on roads with high through-traffic results in acceptable increases in commuting distances.
-As a result, a hierarchy of road preference needs to be defined to route potential cycling demand.
-The framework provides different algorithms to determine where investments in cycling infrastructure should be prioritized.
-These algorithms, based on connectivity and egalitarian principles of resource distribution, ensured that whatever the level of investment, the resulting cycling network would improve connectivity of existing cycling infrastructure.
 
-The approach is not without limitations.
-The level of detail is only as good as the granularity of the available data (in this case relatively coarse zones).
-@iacono2010measuring note that such large travel zones are not ideal for understanding route choice behaviour of cyclists and pedestrians.
-They also give rise to an \`ecological fallacy' whereby average characteristics are assumed to apply to all residents of the aggregated geographical area, suggesting a need for applying the methods to more granular origin-destination data (and for governments and other data-collecting organisations to make origin-destination data more readily available).
+A core part of the methodology is determining where there is a high
+potential for cycling, and using this as a basis for recommendations on
+road space reallocation in order to unlock potential cycling demand. The
+framework provides a routing engine to study the effects of limiting
+access on different road types, finding that reducing cyclist flow on
+roads with high through-traffic results in acceptable increases in
+commuting distances. As a result, a hierarchy of road preference needs
+to be defined to route potential cycling demand. The framework provides
+different algorithms to determine where investments in cycling
+infrastructure should be prioritized. These algorithms, based on
+connectivity and egalitarian principles of resource distribution,
+ensured that whatever the level of investment, the resulting cycling
+network would improve connectivity of existing cycling infrastructure.
 
-The approach is also focused solely on the allocation of cycling infrastructure, and does not consider the larger political and regulatory environment that needs to exist to promote cycling; while segregated, connected, and direct cycling infrastructure is key to achieving high levels of cycling, research has shown that it cannot exist in a vacuum.
-@wardman2007factors developed a mode choice model for the UK and their results showed that improved cycling infrastructure on its own only had modest impacts on mode shift, and even the unlikely scenario of all urban routes being serviced by segregated bike lanes was forecast to increase cycling mode share by only 3%.
-International research shows that cities that invest in more comprehensive cycling projects have a more significant increase in the number of cyclists as well as the cycling mode share [@parkin2018designing; @pucher2010infrastructure].
-These cities do not just focus on infrastructure, but on general policies as well as restricting car use.
-Evaluation of policies in Denmark and Germany and the Netherlands has shown that their high cycling mode share is down to a broader set of policies that also include traffic calming, filtered permeability interventions, cycling rights of way, bike parking, integration with the public transport network, and making driving cars both expensive and inconvenient [@pucher2008making].
-While these policies are outside the scope of this research, it is important to recognize their key role in bringing about an increase in levels of cycling.
+The approach is not without limitations. The level of detail is only as
+good as the granularity of the available data (in this case relatively
+coarse zones). @iacono2010measuring note that such large travel zones
+are not ideal for understanding route choice behaviour of cyclists and
+pedestrians. They also give rise to an \`ecological fallacy' whereby
+average characteristics are assumed to apply to all residents of the
+aggregated geographical area, suggesting a need for applying the methods
+to more granular OD data (and for governments and other
+data-collecting organisations to make OD data more
+readily available).
 
-The methods can support the ongoing shift towards active transport, which is one of the priority areas of investments worldwide in transport planning.
-This necessary shift predates the pandemic and is driven by ambitions to improve the health and well-being of the citizens worldwide.
-The success of such a shift depends on the appeal of the cycling network for prospective cyclists, together with policies and interventions deterring the usage of the private car.
+The approach is also focused solely on the allocation of cycling
+infrastructure, and does not consider the larger political and
+regulatory environment that needs to exist to promote cycling; while
+segregated, connected, and direct cycling infrastructure is key to
+achieving high levels of cycling, research has shown that it cannot
+exist in a vacuum. @wardman2007factors developed a mode choice model
+for the UK and their results showed that improved cycling infrastructure
+on its own only had modest impacts on mode shift, and even the unlikely
+scenario of all urban routes being serviced by segregated bike lanes was
+forecast to increase cycling mode share by only 3%. International
+research shows that cities that invest in more comprehensive cycling
+projects have a more significant increase in the number of cyclists as
+well as the cycling mode share [@parkin2018designing;
+@pucher2010infrastructure]. These cities do not just focus on
+infrastructure, but on general policies as well as restricting car use.
+Evaluation of policies in Denmark and Germany and the Netherlands has
+shown that their high cycling mode share is down to a broader set of
+soft and hard policies. Hard policies include traffic calming, filtered permeability
+interventions, cycling rights of way, bike parking, integration with the
+public transport network, and making driving cars both expensive and
+inconvenient, while soft policies include marketing and awareness compaigns [@gossling2013urban; @pucher2008making].
+While these policies are outside the scope of this research, it is important to recognize their key role in
+bringing about an increase in levels of cycling.
+
+The methods can support the ongoing shift towards active transport,
+which is one of the priority areas of investments worldwide in transport
+planning. This necessary shift predates the pandemic and is driven by
+ambitions to improve the health and well-being of the citizens
+worldwide. The success of such a shift depends on the appeal of the
+cycling network for prospective cyclists, together with policies and
+interventions deterring the usage of the private car.
 
 <!-- I think it would be good to say how others could reproduce the methods in other cities (RL) -->
+
 <!-- (HM) Answer: Done. I added a footnote to the first paragraph in the conclusion linking to the github repo -->
@@
 <!-- There should also be links to the literature. -->
+
 <!-- (HM) Answer: There are links to the literature at the end of each thematic section (Calculating Potential Cycling Demand / Routing / Road Segment Prioritization).

### 2021-02-07 README.Rmd
text edits - potential demand

@@
 
-Relying only on existing cycling activity to prioritise cycling
-infrastructure can reinforce existing cycling patterns and ignores
-potential cycling demand that could be satisfied by a connected network.
-To avoid this issue, @duthie2014optimization choose to ignore existing
-demand completely, and focus on creating a network that connects the
-entire study area. @olmos2020data obtain the distance distribution of
-cyclists using a smartphone-based bicycle GPS data, and then use a
-rejection-sampling algorithm on the OD data of the study area to match
-the potential demand distribution to the distribution obtained from GPS
-data.
+Many of the cycling network studies mentioned above use demand data for cycling as a starting point. 
+Some use existing cycling demand, some calculate potential cycling demand, and others ignore demand completely. 
+@duthie2014optimization note that relying only on existing cycling activity to prioritise cycling
+infrastructure can reinforce existing cycling patterns and ignores potential cycling demand that could be satisfied by a connected network.
+To avoid this issue they choose to ignore existing demand completely, and focus on creating a network that connects the
+entire study area. 
+@olmos2020data opt to calculate potential demand instead; they obtain the distance distribution of cyclists using a smartphone-based bicycle GPS data, and then use a rejection-sampling algorithm on the OD data of the study area to match
+the potential demand distribution to the distribution obtained from GPS data.
 
-*Origin-Destination (OD)* data can be obtained from a range of sources,
+*OD* data can be obtained from a range of sources,
 including GPS data, household travels surveys and Census data on work

### 2021-02-26 README.Rmd
text edits

@@
 (OD) pairs need to be connected by roads that meet the desired LOS. A
-limitation of these problem formulations do not explicitly solve for
+limitation of these problem formulations is that they do not explicitly solve for
 continuity, which is dealt with using either a constraint specifying
@@
 latter is concerned with a fair distribution of benefits
-[@jafino2020transport]. @nahmias2017integrating criticize the
-utilitarian approach that has been historically popular in the
-evaluation of transport investments, explaining how the maximization of
-overall benefit fails to account for the distribution of that benefit
-among communities or individuals. @lucas2016method explain how
-transport studies have traditionally looked at the bigger picture
-without studying the distribution of investments on the different parts
-of the study area, and go on to propose an egalitarian approach that
-ensures the disaggregation of transport policy benefits across the study
-area. @pereira2017distributive also emphasize the need for a more
+[@jafino2020transport]. 
+<!-- @nahmias2017integrating criticize the -->
+<!-- utilitarian approach that has been historically popular in the -->
+<!-- evaluation of transport investments, explaining how the maximization of -->
+<!-- overall benefit fails to account for the distribution of that benefit -->
+<!-- among communities or individuals. @lucas2016method explain how -->
+<!-- transport studies have traditionally looked at the bigger picture -->
+<!-- without studying the distribution of investments on the different parts -->
+<!-- of the study area, and go on to propose an egalitarian approach that -->
+<!-- ensures the disaggregation of transport policy benefits across the study -->
+<!-- area.  -->
+The utilitarian approach, historically popular in transport planning, has been criticised for focusing on the bigger picture and failing to account for the distribution of investments on the different communities of the study area [@lucas2016method; @nahmias2017integrating].
+@pereira2017distributive emphasize the need for a more
 egalitarian approach to transport planning. They highlight accessibility
@@
 
-```{r potdemhistograms, fig.cap="Distribution of Potential Cycling Demand", out.width="32%", fig.show='hold', echo=FALSE}
+```{r potdemhistograms, fig.cap="Distribution of Potential Cycling Demand", out.width="30%", fig.show='hold', echo=FALSE}
 knitr::include_graphics(c(
@@
 
-```{r desirefacetcycling, fig.cap="Current and Potential Cycling Demand", out.width="65%", fig.pos="H", fig.align="center", echo=FALSE}
+```{r desirefacetcycling, fig.cap="Current and Potential Cycling Demand", out.width="60%", fig.pos="H", fig.align="center", echo=FALSE}
 knitr::include_graphics("data/Manchester/Plots/desire_facet_cycling.png")
@@
 does not eliminate it completely. This is intentional, as the impedance
-on these roads is only slightly higher than remaining road types (See
+on these roads is only slightly higher than the remaining road types (See
 Table \@ref(table:weightprofiles)). Potential cycling demand is only
@@
 
+The algorithm logic is as follows:
+
 <!-- 1. Identify all links that have segregated cycling infrastructure and add them to the initial solution
@@
 explained in Section \@ref(ethical-underpinnings-and-proposed-approach). This algorithm incorporates
-the ideal of fair distribution by using community detection to partition
-the road network.
-
-The algorithm uses the following logic to ensure fair distribution
-between communities:
+the ideal of fair distribution by ensuring that investment is distributed between the defined communities. 
+This is done using the following logic:
 
@@
 library(sf)
+library(kableExtra)
 
@@
 knitr::kable(table,
-             caption = "Total Person-Km in Different Communities (Manchester)")
+             caption = "Total Person-Km in Different Communities (Manchester)") %>%
+  kable_styling(latex_options = "hold_position")
 
@@
 
+Overall, the improved distribution of investment shown by Algorithm 2 does not seem to come at the cost 
+of poorer connectivity or less city-wide gains when compared to Algorithm 1. 
+The results therefore advocate for the incorporation of egalitarian principles in bicycle network planning.
+
 # Discussion and conclusions
@@
 to be defined to route potential cycling demand. The framework provides
-different algorithms to determine where investments in cycling
-infrastructure should be prioritized. These algorithms, based on
+an algorithm to determine where investments in cycling
+infrastructure should be prioritized. This algorithm, based on
 connectivity and egalitarian principles of resource distribution,
@@
 network would improve connectivity of existing cycling infrastructure.
+The algorithm was also shown to provide similar city-wide results as well better distribution of investment when compared to one based on utilitarian principles.
 
@@
 readily available).
+The potential demand calculation is also based on cycling in the traditional sense, and does not consider the 
+effect of micro-mobility on reducing topology-related impedance to cycling.
+Given that we are proposing a methodological framework which can accommodate any cycling uptake functions, this is an acceptable limitation.
 

### 2021-03-06 README.Rmd
Tweak abstract

@@
 abstract: "Understanding the motivators and deterrents to cycling is essential for creating infrastructure that gets more people to adopt cycling as a mode of transport.
-The aim of this paper is to develop methods to support the prioritisation of cycling infrastructure and cycling network design, accounting for cyclist preferences and the growing emphasis on 'filtered permeability' and 'Low Traffic Neighborhood' interventions internationally.
-Distance decay, routing, and network analysis techniques are used to examine where future cycling demand is most likely to arise, how such demand could be accommodated within existing street networks, and how to ensure a fair distribution of investment.
-The city of Manchester is used as a case study, although the methods can be applied to any city, as demonstrated in a reproducible code and open access data accompanying the paper.
-We find that flexible routing profiles that are adjustable to local needs can avoid road segments that are stressful to cycle on, enabling the identification of networks that balance directness against the need for safe and stress-free routes.
-We assess two different policy interventions , 'utilitarian' and 'egalitarian', by analyzing the connectivity of the city through community detection.
-From a policy perspective, our findings indicate that the 'egalitarian' prioritisation model approach is the more adequate to support cost-effective strategic cycle network planning.
+This paper demonstrates new methods to support the prioritisation of cycling infrastructure and cycling network design, accounting for cyclist preferences and the growing emphasis on 'filtered permeability' and 'Low Traffic Neighborhood' interventions internationally.
+The approach combines distance decay, route calculation, and network analysis techniques to examine where future cycling demand is most likely to arise, how such demand could be accommodated within existing street networks, and how to ensure a fair distribution of investment.
+Although each of these approaches have been applied to cycling infrastructure prioritisation in previous research, this is the first time that they have been combined, creating an integrated road segment prioritisation approach.
+The approach is demonstrated on with a case study of Manchester and its scalability is demonstrated in an appendix which demonstrates its applicability to other cities using reproducible code and open access data accompanying the paper.
+The approach enables networks that balance directness against the need for safe and stress-free routes and the assessment of investment scenarios.
+From a policy perspective, the method's key benefit is its ability to support egalitarian and cost-effective strategic cycle network planning.
 \\par\\textbf{Keywords:} cycling networks, low-traffic neighborhoods, routing, transport equity"

### 2021-03-06 README.Rmd
Fix typo in and shorten abstract

@@
 Although each of these approaches have been applied to cycling infrastructure prioritisation in previous research, this is the first time that they have been combined, creating an integrated road segment prioritisation approach.
-The approach is demonstrated on with a case study of Manchester and its scalability is demonstrated in an appendix which demonstrates its applicability to other cities using reproducible code and open access data accompanying the paper.
-The approach enables networks that balance directness against the need for safe and stress-free routes and the assessment of investment scenarios.
-From a policy perspective, the method's key benefit is its ability to support egalitarian and cost-effective strategic cycle network planning.
+The approach is demonstrated in a case study of Manchester and other cities using reproducible code presented in the appendix, resulting in networks that balance directness against the need for safe and stress-free routes under different investment scenarios.
+A key benefit of the approach from a policy perspective is its ability to support egalitarian and cost-effective strategic cycle network planning.
 \\par\\textbf{Keywords:} cycling networks, low-traffic neighborhoods, routing, transport equity"

### 2021-03-06 README.Rmd
Remove quote markes from extended quote

@@
 
-> "Inadequate cycling infrastructure discourages cycling and wastes
-> public money. Much cycling infrastructure in this country is
-> inadequate. It reflects a belief, conscious or otherwise, that hardly
-> anyone cycles, that cycling is unimportant and that cycles must take
-> no meaningful space from more important road users, such as motor
-> vehicles and pedestrians - [@departmentgearchange2020]"
+> Inadequate cycling infrastructure discourages cycling and wastes
+ public money. Much cycling infrastructure in this country is
+ inadequate. It reflects a belief, conscious or otherwise, that hardly
+ anyone cycles, that cycling is unimportant and that cycles must take
+ no meaningful space from more important road users, such as motor
+ vehicles and pedestrians [@departmentgearchange2020]
 

### 2021-03-06 README.Rmd
expand -> address limitations of past research

@@
 
-A considerable amount of research has been done to understand what gets people to cycle.
+Much research has been done to understand what gets people to cycle.
 Segregated cycling infrastructure^[*Segregated cycling infrastructure*
@@
 accounts for motivators and deterrents to cycling. We propose three
-sub-methodologies that expand on some of the limitations of previous
+sub-methodologies that address on some of the limitations of previous
 studies. These limitations include (a) bias inherent when basing network

### 2021-03-06 README.Rmd
Update PCT formula part

@@
 Figures \@ref(fig:potdemhistograms) and \@ref(fig:desirefacetcycling)
-illustrates the proportion of trips cycled by distance and the
+illustrate the proportion of trips cycled by distance and the
 geographic extent of the input OD dataset used in this paper.
@@
 \begin{align}\label{eq:pcteqn}
-     logit(C_{p}) = & -4.018 - 0.6369d +  1.988\sqrt{d} + 0.008775d^2\\ & - 0.2555s + 0.00206ds -0.1234\sqrt{d}s\nonumber 
+     logit(C_{p}) = & \alpha + \beta_1 d + \beta_2 \sqrt{d} + \beta_3 d^2 + \beta_4 s + \beta_5 ds + \beta_6 \sqrt{d}s\nonumber 
 \end{align}
@@
 
-\noindent where **d** and **s** are the distance and slope respectively
-for the OD pair. The authors use square and square-root distance terms
-"to capture the non-linear impact of distance on the likelihood of
+\noindent where **d** and **s** are the distance and slope associated with each
+OD pair and $\alpha$ and $\beta$ values are parameters calculated by a regression
+model on training data. The square and square-root distance terms
+"capture the non-linear impact of distance on the likelihood of
 cycling", and interaction terms to capture the combined effect of slope
 and distance [@lovelace2017propensity]. Alternative cycling uptake
-models could be used in different contexts or scenarios of change.
+models could be 'plugged in' to our approach for different
+different contexts or scenarios of change.
 

### 2021-03-06 README.Rmd
Update final section, new reference

@@
 
-It should be noted that the calculations assume cycling in the
-traditional sense, i.e. the topography affects the decision to cycle.
-Recently there have been various micro-mobility solutions, including
-e-bikes, that allow commuters to traverse longer distances and hillier
-roads with less effort than traditional bicycles. While these modes
-would probably be associated with less geographical impedance, it is
-beyond the scope of this work to integrate those modes into the
-analysis. This is partially restricted by the lack of data on the
-proliferation of these modes, raising the point that perhaps the census
-data category of `Bicycle' is too vague, and should be further
-disaggregated to distinguish between traditional bicycles and other
-forms of micro-mobility.
+The uptake model used in this paper assumes uptake of traditional bicycles,
+which are affected by topography and distance due to physical effort.
+As discussed in Section \@ref(discussion-and-conclusions), using adapted uptake
+models could enable the approach to help plan for solutions such as e-bikes
+that enable trips covering longer distances and hillier roads.
 
@@
 
-This paper provides a methodology for prioritizing road segments for
-investments in cycling infrastructure, with the output being a proposed
-cycling network for the city under study. The approach aims to respect
+This paper demonstrates an approach for prioritizing
+investments in city-wide cycleway networks. The approach aims to respect
 both the needs of the users and the ambitions of stakeholders working at
-local or regional levels. The results, primarily detailed route network
+local or regional levels. The results, including detailed route network
 maps based on current travel behaviour derived from OD data, can provide
-an evidence-based starting point for introducing a direct, connected,
-and low-stress network for any city by leveraging available open data.
+an evidence-base for designing direct, connected,
+and low-stress networks.
+
 Given that the "most essential activity entailed in the design of
@@
 
-A core part of the methodology is determining where there is a high
-potential for cycling, and using this as a basis for recommendations on
-road space reallocation in order to unlock potential cycling demand. The
-framework provides a routing engine to study the effects of limiting
-access on different road types, finding that reducing cyclist flow on
-roads with high through-traffic results in acceptable increases in
-commuting distances. As a result, a hierarchy of road preference needs
-to be defined to route potential cycling demand. The framework provides
-an algorithm to determine where investments in cycling
-infrastructure should be prioritized. This algorithm, based on
-connectivity and egalitarian principles of resource distribution,
-ensured that whatever the level of investment, the resulting cycling
-network would improve connectivity of existing cycling infrastructure.
-The algorithm was also shown to provide similar city-wide results as well better distribution of investment when compared to one based on utilitarian principles.
+The approach can identify not only where there is high
+potential for cycling but also trade-offs between
+quietness (the absence of motor traffic) and directness.
+The results can therefore be used as a basis for recommendations on
+road space reallocation *and* new infrastructure to unlock potential cycling demand.
+The approach encourages consideration of a wider range of preferences
+and needs that previous methods that focus only on absolute potential.
+Moreover, the inclusion of egalitarian principles in scenarios of change
+encourages investment in cycling infrastructure to
+increase the connectivity of existing cycling infrastructure
+*and* investment that addresses geographical and social inequalities.
+This ability to address inequalities in network prioritisation is particularly
+important given research research showing substantial *inequalities*
+around transport provision in general and cycling uptake and investment
+in particular [@lucas2016method; @vidaltortosa_infrastructure_2020].
 

### 2021-03-06 README.Rmd
Complete edits

@@
 public transport network, and making driving cars both expensive and
-inconvenient, while soft policies include marketing and awareness compaigns [@gossling2013urban; @pucher2008making].
+inconvenient, while soft policies include marketing and awareness campaigns [@gossling2013urban; @pucher2008making].
 While these policies are outside the scope of this research, it is important to recognize their key role in
@@
 
-The methods can support the ongoing shift towards active transport,
-which is one of the priority areas of investments worldwide in transport
-planning. This necessary shift predates the pandemic and is driven by
-ambitions to improve the health and well-being of the citizens
-worldwide. The success of such a shift depends on the appeal of the
-cycling network for prospective cyclists, together with policies and
-interventions deterring the usage of the private car.
+Consideration of these limitations suggest future directions of research.
+New datasets and continuously evolving computational (hardware and software)
+capabilities should enable the data related issues to be overcome as open
+datasets, and our ability to process them, improve.
+Plugging-in alternative uptake models could help address the relatively narrow
+definition of 'potential' used in this paper, to consider broader social factors.
+A promising area of future research in this direction could be 
+to integrate a wider range of modes, including e-bikes and e-scooters, into the
+analysis. Currently scope for progress in this direction is
+partially restricted by the lack of data on the
+proliferation of these modes, raising the point that surveys could
+replace the broad category of `Bicycle' a range of small modes such as
+'pedal cycles' (including tricycles and recumbent cycles), e-bikes
+and other forms of micro-mobility.
+
+Despite the rapid growth of these alternative modes, there is little doubt that
+the humble bicycle is a key ingredient in the ongoing shift towards active transport.
+Recognition of the importance of this shift for improved the health and well-being
+the citizens, and the wider challenge to decarbonise the global economy,
+has grown with pressures on health and public transport systems during the pandemic.
+The success of policies to accelerate cycling uptake depends on a range of
+factors including, vitally, the design of the
+cycling network for potential cyclists.
+The approach presented in this paper provides a strong evidence base,
+that considers both cycling potential and social equity,
+for designing joined-up and cost-effective strategic cycle networks.
+<!-- In tandem with policies to reduce excessive reliance on private cars, evidence-based investment in strategic cycle networks is key to. -->
 

### 2021-03-12 README.Rmd
Update README.Rmd

@@
 Although each of these approaches have been applied to cycling infrastructure prioritisation in previous research, this is the first time that they have been combined, creating an integrated road segment prioritisation approach.
-The approach is demonstrated in a case study of Manchester and other cities using reproducible code presented in the appendix, resulting in networks that balance directness against the need for safe and stress-free routes under different investment scenarios.
+The approach, which can be applied to other cities, as shown in the Appendix, is demonstrated in a case study of Manchester, resulting in cycling networks that balance directness against the need for safe and stress-free routes under different investment scenarios.
 A key benefit of the approach from a policy perspective is its ability to support egalitarian and cost-effective strategic cycle network planning.

### 2021-03-12 README.Rmd

@@
 potential for cycling but also trade-offs between
-quietness (the absence of motor traffic) and directness.
+stress levels (related to motor traffic) and directness.
 The results can therefore be used as a basis for recommendations on

### 2021-03-12 README.Rmd
change methodology -> approach

@@
 abstract: "Understanding the motivators and deterrents to cycling is essential for creating infrastructure that gets more people to adopt cycling as a mode of transport.
-This paper demonstrates new methods to support the prioritisation of cycling infrastructure and cycling network design, accounting for cyclist preferences and the growing emphasis on 'filtered permeability' and 'Low Traffic Neighborhood' interventions internationally.
-The approach combines distance decay, route calculation, and network analysis techniques to examine where future cycling demand is most likely to arise, how such demand could be accommodated within existing street networks, and how to ensure a fair distribution of investment.
-Although each of these approaches have been applied to cycling infrastructure prioritisation in previous research, this is the first time that they have been combined, creating an integrated road segment prioritisation approach.
+This paper demonstrates a new approach to support the prioritisation of cycling infrastructure and cycling network design, accounting for cyclist preferences and the growing emphasis on 'filtered permeability' and 'Low Traffic Neighborhood' interventions internationally.
+The approach combines distance decay, route calculation, and network analysis methods to examine where future cycling demand is most likely to arise, how such demand could be accommodated within existing street networks, and how to ensure a fair distribution of investment.
+Although each of these methods have been applied to cycling infrastructure prioritisation in previous research, this is the first time that they have been combined, creating an integrated road segment prioritisation approach.
 The approach, which can be applied to other cities, as shown in the Appendix, is demonstrated in a case study of Manchester, resulting in cycling networks that balance directness against the need for safe and stress-free routes under different investment scenarios.
@@
 not propose network-level interventions. In this section we outline
-techniques used in past studies, namely optimization and network
-analysis techniques, such as connected components and community
+methods used in past studies, namely optimization and network
+analysis methods, such as connected components and community
 detection, and examine how they are leveraged to suggest cycling network
 designs. We compare the effectiveness of these network-level studies in
-incorporating the fundamentals outlined above. The methodology of this
-research is inspired by these techniques, but it attempts to add to them
+incorporating the fundamentals outlined above. 
+Our proposed approach is inspired by these methods, but it attempts to add to them
 by ensuring that all of the outlined fundamentals are accounted for. It
@@
 
-*Optimization* techniques have been used to propose improvements to
+*Optimization* methods have been used to propose improvements to
 cycling networks. @mesbah2012bilevel propose a bi-level formulation to
@@
 area partitioning with weighted routing to avoid links that are
-stressful to cycle on. In doing so, we propose a methodology that
+stressful to cycle on. In doing so, we propose an approach that
 accounts for motivators and deterrents to cycling. We propose three
-sub-methodologies that address on some of the limitations of previous
+sub-methods that address on some of the limitations of previous
 studies. These limitations include (a) bias inherent when basing network
@@
 cyclist preferences and government priorities. Section
-\@ref(road-segment-prioritisation) outlines a methodology for partitioning the
+\@ref(road-segment-prioritisation) outlines a method for partitioning the
 study area based on a community finding algorithm and routed cycling
@@
 and distance [@lovelace2017propensity]. Alternative cycling uptake
-models could be 'plugged in' to our approach for different
-different contexts or scenarios of change.
+models could be 'plugged in' to our approach for different contexts or scenarios of change.
 
@@
 would add additional impedance to specific roads, giving more useful
-routing results than the current methodology which considers all roads
+routing results than the current approach which considers all roads
 of the same type to be equivalent.
@@
 cycle-friendly infrastructure is developing a cycle network"
-[@parkin2018designing], we believe that the methods have great
+[@parkin2018designing], we believe that the approach has great
 potential to inform investment in cities such as Manchester where there
@@
 The approach encourages consideration of a wider range of preferences
-and needs that previous methods that focus only on absolute potential.
+and needs than previous approaches that focus only on absolute potential.
 Moreover, the inclusion of egalitarian principles in scenarios of change
@@
 effect of micro-mobility on reducing topology-related impedance to cycling.
-Given that we are proposing a methodological framework which can accommodate any cycling uptake functions, this is an acceptable limitation.
+Given that we are proposing an approach which can accommodate any cycling uptake functions, this is an acceptable limitation.
 

### 2021-03-13 R/_1.0_get_flow_data.R
Add build file, reproducibility fixes

@@ -16,7 +16,7 @@ unique(city_names$city)
 no_msoas <- city_names %>% dplyr::group_by(city) %>% dplyr::tally()
 
 ##### CHOOSE YOU CITY 
-chosen_city <- "Manchester"
+if(!exists("chosen_city")) chosen_city <- "Manchester"
 #create a directory to store data related to this city (does nothing if directory already exists)
 dir.create(paste0("data/", chosen_city), showWarnings = FALSE)
 # create sub-directory to save plots as well

### 2021-03-13 R/_5.0_identifying_cycle_infastructure_from_osm_tags.R

@@ -68,7 +68,7 @@ maxspeed <- streetnet2 %>%
 ########
 
 # Read in the graph with the road network and flows
-graph_sf <- readRDS(paste0("data/", chosen_city,"/graph_with_flows_weighted.RDS"))
+graph_sf <- readRDS(paste0("data/", chosen_city,"/graph_with_flows_weighted.Rds"))
 
 # this is all road segments with bicycle == 'designated'
 cycle_designated <- streetnet2 %>% 
@@ -273,7 +273,7 @@ infra_exists <- function(graph, network){
 }
 
 # Read in the other graphs, add column for cycling infrastrucure, then overwrite
-graph_sf_unweight <- readRDS(paste0("data/", chosen_city,"/graph_with_flows_unweighted.RDS"))
+graph_sf_unweight <- readRDS(paste0("data/", chosen_city,"/graph_with_flows_unweighted.Rds"))
 graph_sf_unweight <- infra_exists(graph=graph_sf_unweight, network=streetnet2)
 saveRDS(graph_sf_unweight, file = paste0("data/", chosen_city, "/graph_with_flows_unweighted.Rds"))
 

### 2021-03-13 R/build.R

@@ -1,0 +1,37 @@
+# Aim: reproduce results for other UK cities
+
+chosen_city <- "Leeds"
+
+source("R/_1.0_get_flow_data.R")
+
+# Get open data needed for 2nd script. Warning: 100 MB file
+piggyback::pb_download("MSOA_2011_Boundaries.zip", tag = "1")
+unzip("MSOA_2011_Boundaries.zip", exdir = "data-raw")
+
+# file.edit("R/_2.0_distance_and_elevation.R")
+source("R/_2.0_distance_and_elevation.R")
+
+# file.edit("R/_3.0_potential_demand.R")
+source("R/_3.0_potential_demand.R")
+# check if figures have been created
+magick::image_read("data/Leeds/Plots/mode_share_increase_vs_performance_smooth_Leeds.png")
+
+remotes::install_cran("waffle")
+remotes::install_cran("latex2exp")
+
+source("R/_3.1_plot_mode_shares.R")
+source("R/_3.2_plot_od_comparison.R")
+source("R/_3.3_plot_desire_lines_current_vs_potential.R")
+
+
+source("R/_4.0_aggregating_flows.R")
+
+file.exists("data/Leeds/msoa_lon_lat.shp")
+file.exists("data/Leeds/graph_with_flows_unweighted.Rds")
+file.edit("R/_5.0_identifying_cycle_infastructure_from_osm_tags.R")
+source("R/_5.0_identifying_cycle_infastructure_from_osm_tags.R")
+
+
+source("R/_3.1_plot_mode_shares.R")
+
+

### 2021-03-13 R/_6.0_comparing_weighting_profiles.R
Reproducibility updates

@@ -3,9 +3,9 @@ library(sf)
 library(tmap)
 
 
-graph_sf_weighted <- readRDS(paste0("data/", chosen_city,"/graph_with_flows_weighted.RDS"))
-graph_sf_unweight <- readRDS(paste0("data/", chosen_city,"/graph_with_flows_unweighted.RDS"))
-#graph_sf_trunk <- readRDS(paste0("data/", chosen_city,"/graph_with_flows_trunk.RDS"))
+graph_sf_weighted <- readRDS(paste0("data/", chosen_city,"/graph_with_flows_weighted.Rds"))
+graph_sf_unweight <- readRDS(paste0("data/", chosen_city,"/graph_with_flows_unweighted.Rds"))
+#graph_sf_trunk <- readRDS(paste0("data/", chosen_city,"/graph_with_flows_trunk.Rds"))
 
 
 plot(graph_sf_weighted['flow'])

### 2021-03-13 R/_7.0_community_detection.R

@@ -41,7 +41,7 @@ plot(nodes['group'], add = TRUE)
 
 
 # read in road edges with aggregated flow data 
-road_segments <- readRDS(paste0("data/", chosen_city,"/graph_with_flows_weighted.RDS"))
+road_segments <- readRDS(paste0("data/", chosen_city,"/graph_with_flows_weighted.Rds"))
 #road_segments <- road_segments %>% dplyr::select(flow)
 
 # plot

### 2021-03-13 R/build.R

@@ -28,8 +28,21 @@ source("R/_4.0_aggregating_flows.R")
 
 file.exists("data/Leeds/msoa_lon_lat.shp")
 file.exists("data/Leeds/graph_with_flows_unweighted.Rds")
-file.edit("R/_5.0_identifying_cycle_infastructure_from_osm_tags.R")
+# file.edit("R/_5.0_identifying_cycle_infastructure_from_osm_tags.R")
+# Warning: takes several minutes to run
 source("R/_5.0_identifying_cycle_infastructure_from_osm_tags.R")
+
+# file.edit("R/_6.0_comparing_weighting_profiles.R")
+source("R/_6.0_comparing_weighting_profiles.R")
+
+file.edit("R/_7.0_community_detection.R")
+source("R/_7.0_community_detection.R")
+
+source("R/_8.0_growing_a_network.R")
+source("R/_8.1_growth_utilitarian.R")
+source("R/_8.2_growth_egalitarian.R")
+
+source("R/_")
 
 
 source("R/_3.1_plot_mode_shares.R")

### 2021-03-13 appendix.Rmd
Final changes

---
title: "Appendix"
author: "Hussein Mahfouz, Robin Lovelace, Elsa Arcaute"
date: "13/03/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2021-03-13 R/_8.1_growth_utilitarian.R

@@ -8,7 +8,7 @@ library(tmap)
 #create a directory to store data related to this city (does nothing if directory already exists)
 dir.create(paste0("data/", chosen_city, "/Plots/Growth_Results"), showWarnings = FALSE)
 
-graph_sf <- readRDS(paste0("data/", chosen_city, "/graph_with_flows_weighted_communities.RDS"))
+graph_sf <- readRDS(paste0("data/", chosen_city, "/graph_with_flows_weighted_communities.Rds"))
 
 # we weigh the flow on each edge by its distance. We can then get how much of the commuter km are satisfied
 graph_sf$person_km <- graph_sf$flow * graph_sf$d

### 2021-03-13 R/_8.2_growth_egalitarian.R

@@ -3,7 +3,7 @@ library(tidyverse)
 library(ggtext)
 library(tmap)
 
-graph_sf <- readRDS(paste0("data/", chosen_city, "/graph_with_flows_weighted_communities.RDS"))
+graph_sf <- readRDS(paste0("data/", chosen_city, "/graph_with_flows_weighted_communities.Rds"))
 
 # we weigh the flow on each edge by its distance. We can then get how much of the commuter km are satisfied
 graph_sf$person_km <- graph_sf$flow * graph_sf$d

### 2021-03-13 R/build.R

@@ -35,7 +35,7 @@ source("R/_5.0_identifying_cycle_infastructure_from_osm_tags.R")
 # file.edit("R/_6.0_comparing_weighting_profiles.R")
 source("R/_6.0_comparing_weighting_profiles.R")
 
-file.edit("R/_7.0_community_detection.R")
+# file.edit("R/_7.0_community_detection.R")
 source("R/_7.0_community_detection.R")
 
 source("R/_8.0_growing_a_network.R")

### 2021-03-13 appendix.Rmd
Update renv

@@
 
-## R Markdown
+# Reproducing the results for other cities
 
-This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
+To reproduce the results presented for Manchester for other cities there are a few computing pre-requisites:
 
-When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
-You can embed an R code chunk like this:
+- A reasonably powerful computer, e.g. with 8 GB RAM
+- The statistical programming language R installed (we recommend running it in the RStudio editor)
+- The source code of the paper repository, which can be download from `url TBC`
 
-```{r cars}
-summary(cars)
+```{r, eval=FALSE, echo=FALSE}
+required_packages = renv::dependencies()
+renv::snapshot()
 ```
 
-## Including Plots
+# Birmingham
 
-You can also embed plots, for example:
+# Leeds
 
-```{r pressure, echo=FALSE}
-plot(pressure)
-```
-
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
+# Nottingham

### 2021-03-13 R/build.R

@@ -40,11 +40,6 @@ source("R/_7.0_community_detection.R")
 
 source("R/_8.0_growing_a_network.R")
 source("R/_8.1_growth_utilitarian.R")
+
 source("R/_8.2_growth_egalitarian.R")
-
-source("R/_")
-
-
-source("R/_3.1_plot_mode_shares.R")
-
 

### 2021-03-13 appendix.Rmd
Update appendix - working

@@
 date: "13/03/2021"
-output: pdf_document
+output:
+  pdf_document:
+    extra_dependencies: ["subcaption", "longtable", "float", "multirow", "placeins", "multicol"]
 ---
@@
 ```{r, eval=FALSE, echo=FALSE}
+# run these lines to update the snapshot
 required_packages = renv::dependencies()
@@
 
+To ensure you have the necessary packages installed, open the `cycle-networks.Rproj` file in RStudio and run the following commands to ensure you have the necessary packages installed:
+
+```{r, eval=FALSE}
+install.packages("renv")
+renv::restore()
+```
+
+The `build.R` script contains all of the necessary steps to reproduce the results for any city in the UK.
+Open this file with the following command and run the code line-by-line.
+
+```{r, eval=FALSE}
+file.edit("R/build.R")
+```
+
+<!-- Todo: make the .zip download work for anyone, e.g. by uploading the data to the osf.io website alongside the preprint. -->
+
+You can select a city to build by executing the following command:
+
+```{r}
+chosen_city <- "Leeds"
+```
+
+Replacing `Leeds` with any other city will make the script work for that city.
+Note: this only works for Local Authority names in England and Wales listed below.
+
+In this script, you choose which city you wish to run the analysis from
+this list of available towns and cities:
+
+    ##   [1] "Barnsley"             "Basildon"             "Basingstoke"         
+    ##   [4] "Bath"                 "Bedford"              "Birkenhead"          
+    ##   [7] "Birmingham"           "Blackburn"            "Blackpool"           
+    ##  [10] "Bolton"               "Bournemouth"          "Bracknell"           
+    ##  [13] "Bradford"             "Brighton and Hove"    "Bristol"             
+    ##  [16] "Burnley"              "Burton upon Trent"    "Bury"                
+    ##  [19] "Cambridge"            "Cardiff"              "Carlisle"            
+    ##  [22] "Chatham"              "Chelmsford"           "Cheltenham"          
+    ##  [25] "Chester"              "Chesterfield"         "Colchester"          
+    ##  [28] "Coventry"             "Crawley"              "Darlington"          
+    ##  [31] "Derby"                "Doncaster"            "Dudley"              
+    ##  [34] "Eastbourne"           "Exeter"               "Gateshead"           
+    ##  [37] "Gillingham"           "Gloucester"           "Grimsby"             
+    ##  [40] "Guildford"            "Halifax"              "Harlow"              
+    ##  [43] "Harrogate"            "Hartlepool"           "Hastings"            
+    ##  [46] "Hemel Hempstead"      "High Wycombe"         "Huddersfield"        
+    ##  [49] "Ipswich"              "Kingston upon Hull"   "Leeds"               
+    ##  [52] "Leicester"            "Lincoln"              "Liverpool"           
+    ##  [55] "London"               "Luton"                "Maidstone"           
+    ##  [58] "Manchester"           "Mansfield"            "Middlesbrough"       
+    ##  [61] "Milton Keynes"        "Newcastle upon Tyne"  "Newcastle-under-Lyme"
+    ##  [64] "Newport"              "Northampton"          "Norwich"             
+    ##  [67] "Nottingham"           "Nuneaton"             "Oldham"              
+    ##  [70] "Oxford"               "Peterborough"         "Plymouth"            
+    ##  [73] "Poole"                "Portsmouth"           "Preston"             
+    ##  [76] "Reading"              "Redditch"             "Rochdale"            
+    ##  [79] "Rotherham"            "Salford"              "Scunthorpe"          
+    ##  [82] "Sheffield"            "Shrewsbury"           "Slough"              
+    ##  [85] "Solihull"             "South Shields"        "Southampton"         
+    ##  [88] "Southend-on-Sea"      "Southport"            "St Albans"           
+    ##  [91] "St Helens"            "Stevenage"            "Stockport"           
+    ##  [94] "Stockton-on-Tees"     "Stoke-on-Trent"       "Sunderland"          
+    ##  [97] "Sutton Coldfield"     "Swansea"              "Swindon"             
+    ## [100] "Telford"              "Wakefield"            "Walsall"             
+    ## [103] "Warrington"           "Watford"              "West Bromwich"       
+    ## [106] "Weston-Super-Mare"    "Wigan"                "Woking"              
+    ## [109] "Wolverhampton"        "Worcester"            "Worthing"            
+    ## [112] "York"
+
 # Birmingham
+
+\subsubsection{Potential Demand}
+
+\begin{figure} [H] 
+    \centering % puts the image in the horizontal centre of the page
+    \captionsetup{font=footnotesize,labelfont=footnotesize}
+    \includegraphics[width=.4\textwidth]{data/Birmingham/Plots/mode_share_increase_vs_performance_smooth_Birmingham} 
+    \caption*{Distribution of Cycling Demand Relative to OD Pair Performance} 
+\end{figure}
+
+\begin{figure} [H] 
+    \centering % puts the image in the horizontal centre of the page
+    \captionsetup{font=footnotesize,labelfont=footnotesize}
+    \includegraphics[width=.85\textwidth]{data/Birmingham/Plots/desire_facet_cycling_Birmingham} 
+    \caption*{Current and Potential Cycling Flow (Birmingham)} 
+\end{figure}
+
+\newpage
+
+\subsubsection{Routing Cycling Flows}
+
+\begin{figure} [H] 
+    \centering % puts the image in the horizontal centre of the page
+    \captionsetup{font=footnotesize,labelfont=footnotesize}
+    \includegraphics[width=.9\textwidth]{data/Birmingham/Plots/flows_facet_unweighted_Birmingham} 
+    \caption*{Flow Results Based on \textbf{Unweighted} Shortest Paths (Birmingham)} 
+\end{figure}
+
+\begin{figure} [H] 
+    \centering % puts the image in the horizontal centre of the page
+    \captionsetup{font=footnotesize,labelfont=footnotesize}
+    \includegraphics[width=.9\textwidth]{data/Birmingham/Plots/flows_facet_weighted_Birmingham} 
+    \caption*{Flow Results Based on \textbf{Weighted} Shortest Paths (Birmingham)} 
+\end{figure}
+
+\subsubsection{Algorithm 2 (Utilitarian)}
+
+\begin{figure} [H] 
+\centering
+\captionsetup{font=footnotesize,labelfont=footnotesize} % size of captions
+\begin{subfigure}{.4\textwidth}
+  \centering
+  \includegraphics[width=1\linewidth]{data/Birmingham/Plots/Growth_Results/growth_existing_infra_satisfied_km_all_flow_column_Birmingham}
+  \caption{\% of Person-Km Satisfied}
+\end{subfigure}
+\begin{subfigure}{.4\textwidth}
+  \centering
+  \includegraphics[width=1\linewidth]{data/Birmingham/Plots/Growth_Results/growth_existing_infra_satisfied_km_community_flow_column_Birmingham}
+  \caption{\% of Person-Km Satisfied in Each Community}
+\end{subfigure}
+\caption*{Effectiveness of Alg. 2 on Overall and Community-Wide Person-Km Satisfied (Birmingham)}
+\end{figure}
+
+\subsubsection{Algorithm 3 (Egalitarian)}
+
+\begin{figure} [H] 
+\centering
+\captionsetup{font=footnotesize,labelfont=footnotesize} % size of captions
+\begin{subfigure}{.4\textwidth}
+  \centering
+  \includegraphics[width=1\linewidth]{data/Birmingham/Plots/Growth_Results/growth_community_4_satisfied_km_all_flow_column_Birmingham}
+  \caption{\% of Person-Km Satisfied}
+\end{subfigure}
+\begin{subfigure}{.4\textwidth}
+  \centering
+  \includegraphics[width=1\linewidth]{data/Birmingham/Plots/Growth_Results/growth_community_4_satisfied_km_community_flow_column_Birmingham}
+  \caption{\% of Person-Km Satisfied in Each Community}
+\end{subfigure}
+\caption*{Effectiveness of Alg. 3 on Overall and Community-Wide Person-Km Satisfied (Birmingham)}
+\end{figure}
+
+\begin{figure} [H] 
+\centering
+\captionsetup{font=footnotesize,labelfont=footnotesize} % size of captions
+\includegraphics[width=1\linewidth]{data/Birmingham/Plots/Growth_Results/growth_community_4_priority_all_FLOW_Birmingham}
+\caption*{Results of Alg. 3 (Birmingham)}
+\end{figure}
 
@@
 
+\subsubsection{Potential Demand}
+
+\begin{figure} [H] 
+    \centering % puts the image in the horizontal centre of the page
+    \captionsetup{font=footnotesize,labelfont=footnotesize}
+    \includegraphics[width=.4\textwidth]{data/Leeds/Plots/mode_share_increase_vs_performance_smooth_Leeds} 
+    \caption*{Distribution of Cycling Demand Relative to OD Pair Performance} 
+\end{figure}
+
+<!-- \begin{figure} [H]  -->
+<!--     \centering % puts the image in the horizontal centre of the page -->
+<!--     \captionsetup{font=footnotesize,labelfont=footnotesize} -->
+<!--     \includegraphics[width=.85\textwidth]{data/Leeds/Plots/desire_facet_cycling_Leeds}  -->
+<!--     \caption*{Current and Potential Cycling Flow (Leeds)}  -->
+<!-- \end{figure} -->
+
+\newpage
+
+\subsubsection{Routing Cycling Flows}
+
+\begin{figure} [H] 
+    \centering % puts the image in the horizontal centre of the page
+    \captionsetup{font=footnotesize,labelfont=footnotesize}
+    \includegraphics[width=.9\textwidth]{data/Leeds/Plots/flows_facet_unweighted_Leeds} 
+    \caption*{Flow Results Based on \textbf{Unweighted} Shortest Paths (Leeds)} 
+\end{figure}
+
+\begin{figure} [H] 
+    \centering % puts the image in the horizontal centre of the page
+    \captionsetup{font=footnotesize,labelfont=footnotesize}
+    \includegraphics[width=.9\textwidth]{data/Leeds/Plots/flows_facet_weighted_Leeds} 
+    \caption*{Flow Results Based on \textbf{Weighted} Shortest Paths (Leeds)} 
+\end{figure}
+
+\subsubsection{Algorithm 2 (Utilitarian)}
+
+<!-- \begin{figure} [H]  -->
+<!-- \centering -->
+<!-- \captionsetup{font=footnotesize,labelfont=footnotesize} % size of captions -->
+<!-- \begin{subfigure}{.4\textwidth} -->
+<!--   \centering -->
+<!--   \includegraphics[width=1\linewidth]{data/Leeds/Plots/Growth_Results/growth_existing_infra_satisfied_km_all_flow_column_Leeds} -->
+<!--   \caption{\% of Person-Km Satisfied} -->
+<!-- \end{subfigure} -->
+<!-- \begin{subfigure}{.4\textwidth} -->
+<!--   \centering -->
+<!--   \includegraphics[width=1\linewidth]{data/Leeds/Plots/Growth_Results/growth_existing_infra_satisfied_km_community_flow_column_Leeds} -->
+<!--   \caption{\% of Person-Km Satisfied in Each Community} -->
+<!-- \end{subfigure} -->
+<!-- \caption*{Effectiveness of Alg. 2 on Overall and Community-Wide Person-Km Satisfied (Leeds)} -->
+<!-- \end{figure} -->
+
+\subsubsection{Algorithm 3 (Egalitarian)}
+
+\begin{figure} [H]
+\centering
+\captionsetup{font=footnotesize,labelfont=footnotesize} % size of captions
+\begin{subfigure}{.4\textwidth}
+  \centering
+  \includegraphics[width=1\linewidth]{data/Leeds/Plots/Growth_Results/growth_utilitarian_satisfied_km_all_flow_column}
+  \caption{\% of Person-Km Satisfied}
+\end{subfigure}
+\begin{subfigure}{.4\textwidth}
+  \centering
+  \includegraphics[width=1\linewidth]{data/Leeds/Plots/Growth_Results/growth_utilitarian_satisfied_km_community_flow_column}
+  \caption{\% of Person-Km Satisfied in Each Community}
+\end{subfigure}
+\caption*{Effectiveness of Alg. 3 on Overall and Community-Wide Person-Km Satisfied (Leeds)}
+\end{figure}
+
+\begin{figure} [H] 
+\centering
+\captionsetup{font=footnotesize,labelfont=footnotesize} % size of captions
+\includegraphics[width=1\linewidth]{data/Leeds/Plots/Growth_Results/growth_egalitarian_satisfied_km_community_flow_column}
+\caption*{Results of Alg. 3 (Leeds)}
+\end{figure}
+
 # Nottingham

### 2021-03-16 appendix.Rmd
appendix

--- 2021-03-13 appendix.Rmd
@@
 date: "13/03/2021"
-output: pdf_document
+geometry: margin=1.5cm
+output:
+  pdf_document:
+    extra_dependencies: ["subcaption", "float"]
 ---
@@
 
-To reproduce the results presented for Manchester for other cities there are a few computing pre-requisites:
+To reproduce the results presented for other cities in the UK there are a few computing pre-requisites:
 
@@
 
+To ensure you have the necessary packages installed, open the `cycle-networks.Rproj` file in RStudio and run the following commands to ensure you have the necessary packages installed:
+
+```{r, eval=FALSE}
+install.packages("renv")
+renv::restore()
+```
+
+The `build.R` script contains all of the necessary steps to reproduce the results for any city in the UK.
+Open this file with the following command and run the code line-by-line.
+
+```{r, eval=FALSE}
+file.edit("R/build.R")
+```
+
+<!-- Todo: make the .zip download work for anyone, e.g. by uploading the data to the osf.io website alongside the preprint. -->
+
+You can select a city to build by executing the following command:
+
+```{r}
+chosen_city <- "Leeds"
+```
+
+Replacing `Leeds` with any other city will make the script work for that city.
+Note: this only works for Local Authority names in England and Wales listed below.
+
+In this script, you choose which city you wish to run the analysis from
+this list of available towns and cities:
+
+    ##   [1] "Barnsley"             "Basildon"             "Basingstoke"         
+    ##   [4] "Bath"                 "Bedford"              "Birkenhead"          
+    ##   [7] "Birmingham"           "Blackburn"            "Blackpool"           
+    ##  [10] "Bolton"               "Bournemouth"          "Bracknell"           
+    ##  [13] "Bradford"             "Brighton and Hove"    "Bristol"             
+    ##  [16] "Burnley"              "Burton upon Trent"    "Bury"                
+    ##  [19] "Cambridge"            "Cardiff"              "Carlisle"            
+    ##  [22] "Chatham"              "Chelmsford"           "Cheltenham"          
+    ##  [25] "Chester"              "Chesterfield"         "Colchester"          
+    ##  [28] "Coventry"             "Crawley"              "Darlington"          
+    ##  [31] "Derby"                "Doncaster"            "Dudley"              
+    ##  [34] "Eastbourne"           "Exeter"               "Gateshead"           
+    ##  [37] "Gillingham"           "Gloucester"           "Grimsby"             
+    ##  [40] "Guildford"            "Halifax"              "Harlow"              
+    ##  [43] "Harrogate"            "Hartlepool"           "Hastings"            
+    ##  [46] "Hemel Hempstead"      "High Wycombe"         "Huddersfield"        
+    ##  [49] "Ipswich"              "Kingston upon Hull"   "Leeds"               
+    ##  [52] "Leicester"            "Lincoln"              "Liverpool"           
+    ##  [55] "London"               "Luton"                "Maidstone"           
+    ##  [58] "Manchester"           "Mansfield"            "Middlesbrough"       
+    ##  [61] "Milton Keynes"        "Newcastle upon Tyne"  "Newcastle-under-Lyme"
+    ##  [64] "Newport"              "Northampton"          "Norwich"             
+    ##  [67] "Nottingham"           "Nuneaton"             "Oldham"              
+    ##  [70] "Oxford"               "Peterborough"         "Plymouth"            
+    ##  [73] "Poole"                "Portsmouth"           "Preston"             
+    ##  [76] "Reading"              "Redditch"             "Rochdale"            
+    ##  [79] "Rotherham"            "Salford"              "Scunthorpe"          
+    ##  [82] "Sheffield"            "Shrewsbury"           "Slough"              
+    ##  [85] "Solihull"             "South Shields"        "Southampton"         
+    ##  [88] "Southend-on-Sea"      "Southport"            "St Albans"           
+    ##  [91] "St Helens"            "Stevenage"            "Stockport"           
+    ##  [94] "Stockton-on-Tees"     "Stoke-on-Trent"       "Sunderland"          
+    ##  [97] "Sutton Coldfield"     "Swansea"              "Swindon"             
+    ## [100] "Telford"              "Wakefield"            "Walsall"             
+    ## [103] "Warrington"           "Watford"              "West Bromwich"       
+    ## [106] "Weston-Super-Mare"    "Wigan"                "Woking"              
+    ## [109] "Wolverhampton"        "Worcester"            "Worthing"            
+    ## [112] "York"
+
+\newpage
+
 # Birmingham
+
+\subsubsection{Potential Demand}
+
+```{r potdemhistogramsBirmingham, fig.cap="Distribution of Potential Cycling Demand", out.width="30%", fig.show='hold', echo=FALSE}
+knitr::include_graphics(c(
+  "data/Birmingham/Plots/histogram_distance_all_vs_cycling.png",
+  "data/Birmingham/Plots/histogram_distance_all_vs_cycling_potential.png",
+  "data/Birmingham/Plots/histogram_distance_cycling_potential_vs_current.png"
+))
+```
+
+```{r desirefacetcyclingBirmingham, fig.cap="Current and Potential Cycling Demand", out.width="80%", fig.pos="H", fig.align="center", echo=FALSE}
+knitr::include_graphics("data/Birmingham/Plots/desire_facet_cycling.png")
+```
+
+\clearpage
+
+\subsubsection{Routing Cycling Flows}
+
+```{r flowsfacetunweightedBirmingham, fig.cap="Flow Results Based on Unweighted Shortest Paths (Birmingham)", out.width="75%", fig.show='hold', fig.align = "center", echo=FALSE}
+knitr::include_graphics("data/Birmingham/Plots/flows_facet_unweighted_Birmingham.png")
+```
+
+```{r flowsfacetweightedBirmingham, fig.cap="Flow Results Based on Weighted Shortest Paths (Birmingham)", out.width="75%", fig.show='hold', fig.align = "center", echo=FALSE}
+knitr::include_graphics("data/Birmingham/Plots/flows_facet_weighted_Birmingham.png")
+```
+
+\clearpage
+
+\subsubsection{Community Detection}
+
+```{r communitiesBirmingham, fig.cap="Communities Based on Potential Cycling Demand (Birmingham)", out.width="90%", fig.show='hold', fig.align = "center", echo=FALSE}
+knitr::include_graphics("data/Birmingham/Plots/communities_alternative_Birmingham.png")
+```
+
+\subsubsection{Network Expansion Algorithms}
+
+```{r growthtotal, fig.cap="Comparing Overall (Dashed) and Community Level Person-Km Satisfied (Birmingham)", out.width="45%", fig.show='hold', fig.pos="H", fig.align = "center", echo=FALSE}
+knitr::include_graphics(c(
+  "data/Birmingham/Plots/Growth_Results/growth_utilitarian_satisfied_km_both_flow_column.png",
+  "data/Birmingham/Plots/Growth_Results/growth_egalitarian_satisfied_km_both_flow_column.png"
+))
+```
+
+```{r componentsandGCCBirmingham, fig.cap="Network Characteristics", out.width="45%", fig.show='hold', fig.pos="H", fig.align = "center", echo=FALSE}
+knitr::include_graphics(c(
+  "data/Birmingham/Plots/Growth_Results/growth_util_egal_components_number_comparisonBirmingham.png",
+  "data/Birmingham/Plots/Growth_Results/growth_util_egal_components_gcc_comparisonBirmingham.png"
+))
+```
+
+```{r growth3MapandBarBirmingham, fig.cap="Road Segment Priority (left), disaggregated by road type (right) - Egalitarian Growth", out.width="45%", fig.show='hold', fig.align="center", echo=FALSE}
+knitr::include_graphics(c(
+  "data/Birmingham/Plots/Growth_Results/growth_egalitarian_priority_all_FLOW.png"#,
+ # "data/Birmingham/Plots/Growth_Results/growth_egalitarian_investment_highways_flow.png"
+))
+```
+
+\clearpage
 
@@
 
+\subsubsection{Potential Demand}
+
+```{r potdemhistogramsLeeds, fig.cap="Distribution of Potential Cycling Demand", out.width="30%", fig.show='hold', echo=FALSE}
+knitr::include_graphics(c(
+  "data/Leeds/Plots/histogram_distance_all_vs_cycling.png",
+  "data/Leeds/Plots/histogram_distance_all_vs_cycling_potential.png",
+  "data/Leeds/Plots/histogram_distance_cycling_potential_vs_current.png"
+))
+```
+
+```{r desirefacetcyclingLeeds, fig.cap="Current and Potential Cycling Demand", out.width="80%", fig.pos="H", fig.align="center", echo=FALSE}
+knitr::include_graphics("data/Leeds/Plots/desire_facet_cycling.png")
+```
+
+\clearpage
+
+\subsubsection{Routing Cycling Flows}
+
+```{r flowsfacetunweightedLeeds, fig.cap="Flow Results Based on Unweighted Shortest Paths (Leeds)", out.width="75%", fig.show='hold', fig.align = "center", echo=FALSE}
+knitr::include_graphics("data/Leeds/Plots/flows_facet_unweighted_Leeds.png")
+```
+
+```{r flowsfacetweightedLeeds, fig.cap="Flow Results Based on Weighted Shortest Paths (Leeds)", out.width="75%", fig.show='hold', fig.align = "center", echo=FALSE}
+knitr::include_graphics("data/Leeds/Plots/flows_facet_weighted_Leeds.png")
+```
+
+\clearpage
+
+\subsubsection{Community Detection}
+
+```{r communitiesLeeds, fig.cap="Communities Based on Potential Cycling Demand (Leeds)", out.width="90%", fig.show='hold', fig.align = "center", echo=FALSE}
+knitr::include_graphics("data/Leeds/Plots/communities_alternative_Leeds.png")
+```
+
+\subsubsection{Network Expansion Algorithms}
+
+```{r growthtotalLeeds, fig.cap="Comparing Overall (Dashed) and Community Level Person-Km Satisfied (Leeds)", out.width="45%", fig.show='hold', fig.pos="H", fig.align = "center", echo=FALSE}
+knitr::include_graphics(c(
+  "data/Leeds/Plots/Growth_Results/growth_utilitarian_satisfied_km_both_flow_column.png",
+  "data/Leeds/Plots/Growth_Results/growth_egalitarian_satisfied_km_both_flow_column.png"
+))
+```
+
+```{r componentsandGCCLeeds, fig.cap="Network Characteristics", out.width="45%", fig.show='hold', fig.pos="H", fig.align = "center", echo=FALSE}
+knitr::include_graphics(c(
+  "data/Leeds/Plots/Growth_Results/growth_util_egal_components_number_comparisonLeeds.png",
+  "data/Leeds/Plots/Growth_Results/growth_util_egal_components_gcc_comparisonLeeds.png"
+))
+```
+
+```{r growth3MapandBarLeeds, fig.cap="Road Segment Priority (left), disaggregated by road type (right) - Egalitarian Growth", out.width="45%", fig.show='hold', fig.align="center", echo=FALSE}
+knitr::include_graphics(c(
+  "data/Leeds/Plots/Growth_Results/growth_egalitarian_priority_all_FLOW.png",
+  "data/Leeds/Plots/Growth_Results/growth_egalitarian_investment_highways_flow.png"
+))
+```
+
+\clearpage
+
 # Nottingham
+
+\subsubsection{Potential Demand}
+
+```{r potdemhistogramsNottingham, fig.cap="Distribution of Potential Cycling Demand", out.width="30%", fig.show='hold', echo=FALSE}
+knitr::include_graphics(c(
+  "data/Nottingham/Plots/histogram_distance_all_vs_cycling.png",
+  "data/Nottingham/Plots/histogram_distance_all_vs_cycling_potential.png",
+  "data/Nottingham/Plots/histogram_distance_cycling_potential_vs_current.png"
+))
+```
+
+```{r desirefacetcyclingNottingham, fig.cap="Current and Potential Cycling Demand", out.width="80%", fig.pos="H", fig.align="center", echo=FALSE}
+knitr::include_graphics("data/Nottingham/Plots/desire_facet_cycling.png")
+```
+
+\clearpage
+
+\subsubsection{Routing Cycling Flows}
+
+```{r flowsfacetunweightedNottingham, fig.cap="Flow Results Based on Unweighted Shortest Paths (Nottingham)", out.width="75%", fig.show='hold', fig.align = "center", echo=FALSE}
+knitr::include_graphics("data/Nottingham/Plots/flows_facet_unweighted_Nottingham.png")
+```
+
+```{r flowsfacetweightedNottingham, fig.cap="Flow Results Based on Weighted Shortest Paths (Nottingham)", out.width="75%", fig.show='hold', fig.align = "center", echo=FALSE}
+knitr::include_graphics("data/Nottingham/Plots/flows_facet_weighted_Nottingham.png")
+```
+
+\clearpage
+
+\subsubsection{Community Detection}
+
+```{r communitiesNottingham, fig.cap="Communities Based on Potential Cycling Demand (Nottingham)", out.width="90%", fig.show='hold', fig.align = "center", echo=FALSE}
+knitr::include_graphics("data/Nottingham/Plots/communities_alternative_Nottingham.png")
+```
+
+\subsubsection{Network Expansion Algorithms}
+
+```{r growthtotalNottingham, fig.cap="Comparing Overall (Dashed) and Community Level Person-Km Satisfied (Nottingham)", out.width="45%", fig.show='hold', fig.pos="H", bifig.align = "center", echo=FALSE}
+knitr::include_graphics(c(
+  "data/Nottingham/Plots/Growth_Results/growth_utilitarian_satisfied_km_both_flow_column.png",
+  "data/Nottingham/Plots/Growth_Results/growth_egalitarian_satisfied_km_both_flow_column.png"
+))
+```
+
+```{r componentsandGCCNottingham, fig.cap="Network Characteristics", out.width="45%", fig.show='hold', fig.pos="H", fig.align = "center", echo=FALSE}
+knitr::include_graphics(c(
+  "data/Nottingham/Plots/Growth_Results/growth_util_egal_components_number_comparisonNottingham.png",
+  "data/Nottingham/Plots/Growth_Results/growth_util_egal_components_gcc_comparisonNottingham.png"
+))
+```
+
+```{r growth3MapandBarNottingham, fig.cap="Road Segment Priority (left), disaggregated by road type (right) - Egalitarian Growth", out.width="45%", fig.show='hold', fig.align="center", echo=FALSE}
+knitr::include_graphics(c(
+  "data/Nottingham/Plots/Growth_Results/growth_egalitarian_priority_all_FLOW.png",
+  "data/Nottingham/Plots/Growth_Results/growth_egalitarian_investment_highways_flow.png"
+))
+```
+
+\clearpage

### 2021-03-16 R/build.R

@@ -1,12 +1,12 @@
 # Aim: reproduce results for other UK cities
 
-chosen_city <- "Leeds"
+chosen_city <- "Birmingham"
 
 source("R/_1.0_get_flow_data.R")
 
-# Get open data needed for 2nd script. Warning: 100 MB file
-piggyback::pb_download("MSOA_2011_Boundaries.zip", tag = "1")
-unzip("MSOA_2011_Boundaries.zip", exdir = "data-raw")
+# # Get open data needed for 2nd script. Warning: 100 MB file  (HM : Uncomment Later)
+# piggyback::pb_download("MSOA_2011_Boundaries.zip", tag = "1")
+# unzip("MSOA_2011_Boundaries.zip", exdir = "data-raw")
 
 # file.edit("R/_2.0_distance_and_elevation.R")
 source("R/_2.0_distance_and_elevation.R")
@@ -14,10 +14,7 @@ source("R/_2.0_distance_and_elevation.R")
 # file.edit("R/_3.0_potential_demand.R")
 source("R/_3.0_potential_demand.R")
 # check if figures have been created
-magick::image_read("data/Leeds/Plots/mode_share_increase_vs_performance_smooth_Leeds.png")
 
-remotes::install_cran("waffle")
-remotes::install_cran("latex2exp")
 
 source("R/_3.1_plot_mode_shares.R")
 source("R/_3.2_plot_od_comparison.R")
@@ -26,9 +23,9 @@ source("R/_3.3_plot_desire_lines_current_vs_potential.R")
 
 source("R/_4.0_aggregating_flows.R")
 
-file.exists("data/Leeds/msoa_lon_lat.shp")
-file.exists("data/Leeds/graph_with_flows_unweighted.Rds")
-# file.edit("R/_5.0_identifying_cycle_infastructure_from_osm_tags.R")
+# file.exists("data/Leeds/msoa_lon_lat.shp")
+# file.exists("data/Leeds/graph_with_flows_unweighted.Rds")
+
 # Warning: takes several minutes to run
 source("R/_5.0_identifying_cycle_infastructure_from_osm_tags.R")
 

### 2021-04-01 README.Rmd
Update README.Rmd

@@
 The approach combines distance decay, route calculation, and network analysis methods to examine where future cycling demand is most likely to arise, how such demand could be accommodated within existing street networks, and how to ensure a fair distribution of investment.
-Although each of these methods have been applied to cycling infrastructure prioritisation in previous research, this is the first time that they have been combined, creating an integrated road segment prioritisation approach.
+Although each of these methods has been applied to cycling infrastructure prioritisation in previous research, this is the first time that they have been combined, creating an integrated road segment prioritisation approach.
 The approach, which can be applied to other cities, as shown in the Appendix, is demonstrated in a case study of Manchester, resulting in cycling networks that balance directness against the need for safe and stress-free routes under different investment scenarios.
@@
 climate change. Like other sectors, transport is under substantial
-pressure to decarbonise resulting in a number of technical innovations
+pressure to decarbonise, resulting in a number of technical innovations
 including electric vehicles. But new vehicle technologies can only go so
@@
 
-> Inadequate cycling infrastructure discourages cycling and wastes
+> "Inadequate cycling infrastructure discourages cycling and wastes
  public money. Much cycling infrastructure in this country is
@@
  no meaningful space from more important road users, such as motor
- vehicles and pedestrians [@departmentgearchange2020]
+ vehicles and pedestrians" [@departmentgearchange2020].
 
@@
 area's bicycle network denser means adding more cycling routes in the
-area and thereby giving cyclists more route options] of the cycling
+area and thereby giving cyclists more route options.] of the cycling
 network is also vital [@schoner2014missing].
@@
 also goes further by attempting to factor in ethical considerations
-relating to distribution of investment.
+related to the distribution of investment.
 
@@
 Safeguarding against increased car traffic may be counter-productive if the goal is to create a mode-shift, as research has shown that reducing road space for cars leads to less cars on the road, a phenomenon referred to as "traffic evaporation" [@nello2020environmental].
-Car usage is not considered by @mauttone2017bicycle,
-who develop an optimization framework that aims to minimize the total
-user cost of cycling on the network. The aggregate flow^[*flow* is
+On the other hand, @mauttone2017bicycle developed an optimization framework aiming at minimizing the total
+user cost of cycling on the network without considering car usage. The aggregate flow^[*flow* is
 used throughout this research to refer to the cycling demand when it is
@@
 cumulative demand on it, resulting from cyclists commuting between
-various OD pairs] on links^[*link*
-refers to a road segment throughout this research] is obtained by using
-shortest paths to route existing cycling demand onto the road network,
-and the solution is a proposed set of links where cycling infrastructure
+various Origin-Destination (OD) pairs.] on road segments (links) is obtained by using
+shortest paths to route existing cycling demand onto the road network. The solution is a proposed set of links where cycling infrastructure
 should be added in order to minimize the overall travel cost of cyclists
@@
 (OD) pairs need to be connected by roads that meet the desired LOS. A
-limitation of these problem formulations is that they do not explicitly solve for
-continuity, which is dealt with using either a constraint specifying
+limitation of these approaches is that they do not explicitly solve for
+continuity. The latter is addressed using either a constraint specifying
 that each link with a bike lane should be connected to at least one
@@
 the network through the graph-theoretic concept of *connected
-components*. Other academics have taken this approach; @natera2019data
+components*. @natera2019data for example,
 study the existing cycling network in terms of its disconnected
-components and introduce two different algorithms to connect these
+components. They propose two different algorithms to connect disconnected
 components by their most critical links and, in doing so, measure the
@@
 accounts for motivators and deterrents to cycling. We propose three
-sub-methods that address on some of the limitations of previous
-studies. These limitations include (a) bias inherent when basing network
-design solely on existing cycling demand, (b) proposing routes that may
+sub-methods that address some of the limitations of previous
+studies. These limitations include (a) an inherent bias of basing network
+design solely on existing cycling demand, (b) the proposal of routes that may
 not correspond to studies on cyclist preference and government policies,
-and (3) insufficient consideration of the ethical principles underlying
-the analysis. Section \@ref(calculating-potential-cycling-demand)
+and (c) an insufficient consideration of the ethical principles underlying
+the analysis. The current work addresses these limitations through the following organisation. Section \@ref(calculating-potential-cycling-demand)
 focuses on calculating potential cycling demand. Section \@ref(routing)
@@
 @duthie2014optimization note that relying only on existing cycling activity to prioritise cycling
-infrastructure can reinforce existing cycling patterns and ignores potential cycling demand that could be satisfied by a connected network.
+infrastructure can reinforce existing cycling patterns and ignore potential cycling demand that could be satisfied by a connected network.
 To avoid this issue they choose to ignore existing demand completely, and focus on creating a network that connects the
@@
 ($\boldsymbol{C_{p}}$) for each OD pair that would cycle under
-different scenarios of change (we used the Government Target scenario,
-representing a nationwide target of doubling cycling by 2025) as a
-function of distance and hilliness [@lovelace2017propensity]. The
+different scenarios of change as a
+function of distance and hilliness [@lovelace2017propensity]. We used the Government Target scenario,
+indicating a nationwide target of doubling cycling by 2025. The
 logistic regression model used to calculate $\boldsymbol{C_{p}}$ has
@@
 \begin{align}\label{eq:pcteqn}
-     logit(C_{p}) = & \alpha + \beta_1 d + \beta_2 \sqrt{d} + \beta_3 d^2 + \beta_4 s + \beta_5 ds + \beta_6 \sqrt{d}s\nonumber 
+     \logit(C_{p}) = & \alpha + \beta_1 d + \beta_2 \sqrt{d} + \beta_3 d^2 + \beta_4 s + \beta_5 ds + \beta_6 \sqrt{d}s 
 \end{align}
@@
 
-\noindent where **d** and **s** are the distance and slope associated with each
-OD pair and $\alpha$ and $\beta$ values are parameters calculated by a regression
-model on training data. The square and square-root distance terms
+\noindent where $d$ and $s$ are the distance and slope associated with each
+OD pair, and $\alpha$ and all $\beta$s, are parameters calculated by a regression
+model on the training data. The square and square-root distance terms
 "capture the non-linear impact of distance on the likelihood of
@@
     routes that passed through residential streets.
-2.  **Low-Traffic Neighborhoods**: The UK Department for Transport is
+2.  **Low-Traffic Neighbourhoods (LTNs)**: The UK Department for Transport is
     allocating funding to local authorities to invest in Active
@@
 The above points are accounted for by using a weighted road network for
-routing. This has previously been done by multiplying all road segments
+routing. This has been previously done by multiplying all road segments
 without cycling infrastructure by a fixed impedance factor
@@
 \@ref(table:osmroadtypes)), and roads with existing cycling
-infrastructure. We believe this to be more practical than the simplistic
-approach adopted by @mauttone2017bicycle, as it goes beyond simply
-favoring roads with existing cycling infrastructure enabling a hierarchy
-of road preference based on perceived stress levels.
+infrastructure. We believe this to be more appropriate than the simplistic
+approach adopted by @mauttone2017bicycle, as it makes use of a hierarchy
+of road preference based on perceived stress levels, going beyond simply
+favoring roads with existing cycling infrastructure.
 
@@
 
-\noindent A weighted distance $\boldsymbol{d_{w}}$ for each road
+\noindent A weighted distance $d_{w}$ for each road
 segment is calculated as follows ^[The **dodgr** r package
@@
 \begin{equation}\label{eq:weight_distance}
-    d_{w} = d_{unw}  \:/\:  W
+    d_{w} = \frac{d_{unw}}{W}
 \end{equation}
 
-\noindent where $\boldsymbol{d_{unw}}$ is the unweighted distance and
-$\boldsymbol{W}$ is the weight from Table \ref{table:weightprofiles}.
+\noindent where $d_{unw}$ is the unweighted distance and
+$W$ is the weight from Table \ref{table:weightprofiles}.
 
@@
 
-```{r boxplotcircuity, fig.cap="Effect of Banning Cyclists from Trunk and Primary Roads for all OD Pairs (Manchester)", out.width="35%", fig.show='hold', echo=FALSE}
+```{r boxplotcircuity, fig.cap="Effect of banning cyclists from trunk and primary roads for all OD pairs (Manchester)", out.width="35%", fig.show='hold', echo=FALSE}
 knitr::include_graphics("data/Manchester/Plots/boxplot_weighted_unweighted_distances.png")
@@
 to represent different types of cyclists and road environments, as
-described in @gehrke2020cycling and @furth2016network. Sensitivity
+described in @gehrke2020cycling and @furth2016network. A sensitivity
 analysis could be done to determine an optimal weighting profile, but
@@
 
-```{r communitiesmanchester, fig.cap="Communities Based on Potential Cycling Demand (Manchester)", out.width="75%", fig.show='hold', fig.align = "center", echo=FALSE}
+```{r communitiesmanchester, fig.cap="Communities based on potential cycling demand between MSOAs (Manchester)", out.width="75%", fig.show='hold', fig.align = "center", echo=FALSE}
 knitr::include_graphics("data/Manchester/Plots/communities_alternative_Manchester.png")
@@
 \@ref(tab:personkmtable) shows that almost half of the person-km is in
-community 1 , while only 0.5% of total person-km on the network is in
+community 1, while only 0.5% of total person-km on the network is in
 community 5.
@@
 This ability to address inequalities in network prioritisation is particularly
-important given research research showing substantial *inequalities*
+important given research showing substantial *inequalities*
 around transport provision in general and cycling uptake and investment

### 2021-04-10 README.Rmd
paper and figure edits

@@
 ---
-title: Road Segment Prioritization for Bicycle Infrastructure
+title: A Road Segment Prioritization Approach for Cycling Infrastructure
 date: "`r Sys.Date()`"
@@
 \\par\\textbf{Keywords:} cycling networks, low-traffic neighborhoods, routing, transport equity"
-# keywords: "cycling, low-traffic neighborhoods, potential demand, routing"
+# keywords: "cycling networks, low-traffic neighborhoods, potential demand, routing"
 output:
@@
   # # bookdown::word_document2:
+      highlight: null   # to remove code highlighting in r chunks
       toc: false
@@
     wrap: 72
+# elsevier style for referencing (from https://github.com/citation-style-language/styles)
+csl: elsevier-harvard.csl
 ---
@@
 study the existing cycling network in terms of its disconnected
-components. They propose two different algorithms to connect disconnected
-components by their most critical links and, in doing so, measure the
-size of the growth of the largest connected component as a function of
-the kilometers of network added. The concept of connected components is
+components.
+They propose two different algorithms to connect disconnected components; one connects the two largest components, and the other connects the largest component to its nearest neighbor.
+By measuring the growth of the largest connected component as a function of the kilometers of network added, they determine that these approaches are more effective at improving cycling network connectivity than random allocation of cycling infrastructure.
+The concept of connected components is
 also at the core of the methodology proposed by @olmos2020data. After
@@
 \begin{align}\label{eq:pcteqn}
-     \logit(C_{p}) = & \alpha + \beta_1 d + \beta_2 \sqrt{d} + \beta_3 d^2 + \beta_4 s + \beta_5 ds + \beta_6 \sqrt{d}s 
+     logit(C_{p}) = & \alpha + \beta_1 d + \beta_2 \sqrt{d} + \beta_3 d^2 + \beta_4 s + \beta_5 ds + \beta_6 \sqrt{d}s 
 \end{align}
@@
 
-```{r potdemhistograms, fig.cap="Distribution of Potential Cycling Demand", out.width="30%", fig.show='hold', echo=FALSE}
-knitr::include_graphics(c(
-  "data/Manchester/Plots/histogram_distance_all_vs_cycling.png",
-  "data/Manchester/Plots/histogram_distance_all_vs_cycling_potential.png",
-  "data/Manchester/Plots/histogram_distance_cycling_potential_vs_current.png"
-))
+<!-- ```{r potdemhistograms, fig.cap="Distribution of potential cycling demand", out.width="30%", fig.show='hold', echo=FALSE} -->
+<!-- knitr::include_graphics(c( -->
+<!--   "data/Manchester/Plots/histogram_distance_all_vs_cycling.png", -->
+<!--   "data/Manchester/Plots/histogram_distance_all_vs_cycling_potential.png", -->
+<!--   "data/Manchester/Plots/histogram_distance_cycling_potential_vs_current.png" -->
+<!-- )) -->
+<!-- ``` -->
+
+```{r potdemhistograms, fig.cap="Distribution of potential cycling demand", out.width="100%", fig.show='hold', fig.align="center", echo=FALSE}
+knitr::include_graphics("data/Manchester/Plots/histogram_distance_cycling_potential_current_all.png")
 ```
 
-```{r desirefacetcycling, fig.cap="Current and Potential Cycling Demand", out.width="60%", fig.pos="H", fig.align="center", echo=FALSE}
+```{r desirefacetcycling, fig.cap="Current and potential cycling demand", out.width="60%", fig.pos="H", fig.align="center", echo=FALSE}
 knitr::include_graphics("data/Manchester/Plots/desire_facet_cycling.png")
@@
 \begin{longtable}{lll}
-\caption{OSM Road Types}
+\caption{OSM road types}
 \label{table:osmroadtypes}\\
@@
 \begin{longtable}{lrrr}
-\caption{Weighting Profiles}
+\caption{Weighting profiles}
 \label{table:weightprofiles}\\
@@
 network, but the dodgr package divides by numbers between 0 and 1, which
-achieves the same result. For the sake of reproducibility, we stick to
+achieves the same result (road types with a weight of 0 are not routed on). 
+For the sake of reproducibility, we stick to
 the convention used in the package.]:
@@
 
-```{r flowsfacetunweighted, fig.cap="Flow Results Based on Unweighted Shortest Paths (Manchester)", out.width="90%", fig.show='hold', echo=FALSE}
+```{r flowsfacetunweighted, fig.cap="Flow results based on unweighted shortest paths (Manchester)", out.width="90%", fig.show='hold', echo=FALSE}
 knitr::include_graphics("data/Manchester/Plots/flows_facet_unweighted_Manchester.png")
@@
 
-```{r flowsfacetweighted, fig.cap="Flow Results Based on Weighted Shortest Paths (Manchester)", out.width="90%", fig.show='hold', echo=FALSE}
+```{r flowsfacetweighted, fig.cap="Flow results based on weighted shortest paths (Manchester)", out.width="90%", fig.show='hold', echo=FALSE}
 knitr::include_graphics("data/Manchester/Plots/flows_facet_weighted_Manchester.png")
@@
 
-```{r boxplotcircuity, fig.cap="Effect of banning cyclists from trunk and primary roads for all OD pairs (Manchester)", out.width="35%", fig.show='hold', echo=FALSE}
+```{r boxplotcircuity, fig.cap="Effect of banning cyclists from trunk and primary roads for all OD pairs (Manchester)", out.width="35%", fig.show='hold', fig.align="center", echo=FALSE}
 knitr::include_graphics("data/Manchester/Plots/boxplot_weighted_unweighted_distances.png")
@@
 
-```{r cyclinghistmanc, fig.cap="Cycling Commuting Distance - Manchester (2011 Census Data)", out.width="30%", fig.show='hold', echo=FALSE}
+```{r cyclinghistmanc, fig.cap="Cycling commuting distance - Manchester (2011 census data)", out.width="30%", fig.show='hold', fig.align = "center", echo=FALSE}
 knitr::include_graphics("data/Manchester/Plots/histogram_distance_cycling.png")
@@
 ```
+\  
+<!-- this prevents text from floating above figure -->
 
@@
 
-```{r, eval = FALSE}
+```{r, eval = FALSE, results='asis'}
 1. Identify all links that have segregated cycling infrastructure and add them to the 
@@
 
-<!-- 1. Identify all links that have segregated cycling infrastructure and add them to the initial solution
-2. Identify all links that neighbor links in the current solution
-3. Select *from each community* one neighboring link with highest flow and add it to the solution
-4. If there are no more neighboring links in a community, select the link with the highest flow in that community, regardless of connectivity, and add it to the solution
-5. Repeat steps 2, 3 \& 4 until all flow is satisfied or investment threshold is met -->
+<!-- 1. Identify all links that have segregated cycling infrastructure and add them to the initial solution -->
+<!-- 2. Identify all links that neighbor links in the current solution -->
+<!-- 3. Select *from each community* one neighboring link with highest flow and add it to the solution -->
+<!-- 4. If there are no more neighboring links in a community, select the link with the highest flow in that community, regardless of connectivity, and add it to the solution -->
+<!-- 5. Repeat steps 2, 3 \& 4 until all flow is satisfied or investment threshold is met  -->
 
@@
 knitr::kable(table,
-             caption = "Total Person-Km in Different Communities (Manchester)") %>%
+             caption = "Total person-km in different communities (Manchester)") %>%
   kable_styling(latex_options = "hold_position")
@@
 
-```{r growthtotal, fig.cap="Comparing Overall (Dashed) and Community Level Person-Km Satisfied (Manchester)", out.width="35%", fig.show='hold', fig.align = "center", echo=FALSE}
+```{r growthtotal, fig.cap="Comparing overall (dashed) and community level person-km satisfied (Manchester)", out.width="35%", fig.show='hold', fig.align = "center", echo=FALSE}
 knitr::include_graphics(c(
@@
 
-```{r growth3MapandBar, fig.cap="Road Segment Priority (left), disaggregated by road type (right) - Egalitarian Growth", out.width="35%", fig.show='hold', fig.align="center", echo=FALSE}
+```{r growth3MapandBar, fig.cap="Road segment priority (left), disaggregated by road type (right) - egalitarian growth", out.width="35%", fig.show='hold', fig.align="center", echo=FALSE}
 knitr::include_graphics(c(
@@
 
-```{r componentsandGCC, fig.cap="Network Characteristics", out.width="30%", fig.show='hold', fig.pos="H", fig.align = "center", echo=FALSE}
+```{r componentsandGCC, fig.cap="Network characteristics", out.width="30%", fig.show='hold', fig.pos="H", fig.align = "center", echo=FALSE}
 knitr::include_graphics(c(
@@
 also be reproduced for cities elsewhere given the availability of
-commuter data. For instructions on reproducing the results shown in this
+commuter data. The appendix shows the results for 3 other UK cities: Birmingham, Leeds and Nottingham. 
+For instructions on reproducing the results shown in this
 paper, see the **README** file in the R folder of the .zip file that

### 2021-04-10 R/_3.3_plot_desire_lines_current_vs_potential.R

@@ -1,8 +1,9 @@
-library(dplyr)
+library(tidyverse)
 library(pct)
 library(stplanr)
 library(tmap)
 library(lwgeom)
+library(sf)
 
 # This script depends on data from Script 1, 2 and 3
 
@@ -122,7 +123,12 @@ tm_shape(desire_cycling_long) +
             legend.outside = TRUE,
             legend.outside.position = 'bottom',
             frame = FALSE) +
-  tm_scale_bar(color.dark = "gray60") -> p
+  tm_scale_bar(color.dark = "gray60") + 
+  tm_compass(type = "arrow",
+             size = 1,
+             show.labels = 0,
+             position = c("right","top"),
+             color.dark = "gray60") -> p
 
 #save
 tmap_save(tm = p, filename = paste0("data/", chosen_city,"/Plots/desire_facet_cycling.png"), 

### 2021-04-10 R/_7.0_community_detection.R

@@ -155,7 +155,12 @@ tm_shape(road_segments) +
   tm_layout(fontfamily = 'Georgia',
             legend.show =FALSE,
             frame = FALSE) +
-  tm_scale_bar(color.dark = "gray60") -> tm2
+  tm_scale_bar(color.dark = "gray60") + 
+  tm_compass(type = "arrow",
+             size = 1,
+             show.labels = 0,
+             position = c("right","top"),
+             color.dark = "gray60") -> tm2
 
 # 3.3  get legend only for facet map
 tm_shape(road_segments) +

### 2021-04-10 R/_8.2_growth_egalitarian.R

@@ -218,7 +218,12 @@ tm_shape(grow_egal_c) +
             fontfamily = 'Georgia',
             #legend.position = c("right", "bottom"),
             frame = FALSE) +
-  tm_scale_bar(color.dark = "gray60") +
+  tm_scale_bar(color.dark = "gray60") + 
+  tm_compass(type = "arrow",
+             size = 1,
+             show.labels = 0,
+             position = c("left","bottom"),
+             color.dark = "gray60") +
   # add legend for the existing cycling infrastructure
   tm_add_legend(type = "line", labels = 'Existing Cycling Infrastructure', col = 'firebrick2', lwd = 2) -> p
 

### 2021-04-17 R/README.Rmd
update R readme

@@
 #author: "Hussein Mahfouz"
-date: "18/8/2020"
+#date: "18/8/2020"
 output: github_document
@@
 
-This repository contains the code used for the paper. Below is an explanation of the scripts used, and how the analysis can be reproduced.
+This repository contains the code used for the paper. Below is an explanation of the scripts used, and how the analysis can be reproduced for other cities.
 
@@
 
-The paper can be found in the repo,[**ADD LINK TO PDF**]
+The paper can be found in the repo, [here](https://github.com/Hussein-Mahfouz/cycle-networks/blob/main/README.pdf)
 
@@
 
-There is one file that cannot be synced to github due to its size. This file is neseccary for the scripts to run:
+There is one file that cannot be synced to github due to its size. This file is necessary for the scripts to run:
 
@@
 
+## Reproducing the Results for Other Cities
+
+To reproduce the results presented for other cities in the UK there are a few computing pre-requisites:
+
+- A reasonably powerful computer, e.g. with 8 GB RAM
+- The statistical programming language R installed (we recommend running it in the RStudio editor)
+
+To ensure you have the necessary packages installed, open the `cycle-networks.Rproj` file in RStudio and run the following commands:
+
+```{r, eval=FALSE}
+install.packages("renv")
+renv::restore()
+```
+
+The `build.R` script contains all of the necessary steps to reproduce the results for any city in the UK.
+Open this file with the following command and run the code line-by-line.
+
+```{r, eval=FALSE}
+file.edit("R/build.R")
+```
+
+<!-- Todo: make the .zip download work for anyone, e.g. by uploading the data to the osf.io website alongside the preprint. -->
+
+You can select a city to build by executing the following command:
+
+```{r}
+chosen_city <- "Leeds"
+```
+
+Replacing `Leeds` with any other city will make the script work for that city.
+Note: this only works for Local Authority names in England and Wales listed below.
+
 ## Scripts
 
-The scripts should be run in the order they are numbered in (and listed in here). The only exception is _x_dodgr_weighting_profiles.R.
+Below is a description of all the scripts used to run the analysis.
+They are all sourced in the `build.R` script, but a description of all the scripts/components should allow any part of the analysis to be improved
+
+The scripts are run in the order they are numbered in (and listed in here). The only exception is _x_dodgr_weighting_profiles.R.
 
@@
 
-This script uses the The [dodgr](https://atfutures.github.io/dodgr/) package is used to route potential cycling demand onto the road network.
-This is done for the different weighting profiles used in the analysis
+This script uses the [dodgr](https://atfutures.github.io/dodgr/) package to route potential cycling demand onto the road network. This is done for the different weighting profiles used in the analysis
 

### 2021-04-17 appendix.Rmd
update R readme

@@
 title: "Appendix"
-author: "Hussein Mahfouz, Robin Lovelace, Elsa Arcaute"
+#author: "Hussein Mahfouz, Robin Lovelace, Elsa Arcaute"
 date: "13/03/2021"

### 2021-04-17 R/_3.1_plot_mode_shares.R

@@ -1,5 +1,6 @@
 library(tidyverse)
 library(waffle)
+library(patchwork)
 
 # read in the data
 flow <- readr::read_csv(paste0("data/",chosen_city,"/flows_for_desire_lines.csv"))
@@ -203,12 +204,81 @@ ggsave(paste0("data/", chosen_city,"/Plots/histogram_distance_cycling_potential_
 
 
 
+
+# --- Create a Facet Plot Using Patchwork 
+
+
+### current cycling trips vs all commuter trips
+cols <- c("All" = "grey60", "Bicycle (Current)" = "darkred")
+ggplot(histogram, aes(x=dist)) + 
+  geom_histogram(data=flow_long_all, color = 'grey50', aes(fill = "All")) +
+  geom_histogram(data= flow_long_bike, color = 'grey50', aes(fill = "Bicycle (Current)")) +
+  scale_fill_manual(name = "Mode", values = cols) +
+  labs(title = "Current Cycling Trips", 
+       # keep only the y axis (this is the left-most plot)
+       x="", y = "No. of Commuters", color = "Legend") + 
+  theme(axis.title=element_text(size=15)) -> a
+
+
+### potential cycling trips vs all commuter trips
+cols <- c("All" = "grey60", "Bicycle (Potential)" = "darkgreen")
+ggplot(histogram, aes(x=dist)) + 
+  geom_histogram(data=flow_long_all, color = 'grey50', aes(fill = "All")) +
+  geom_histogram(data= flow_long_potential, color = 'grey50', aes(fill = "Bicycle (Potential)")) +
+  scale_fill_manual(name = "Mode", values = cols) +
+  labs(title = "Potential Cycling Trips", 
+       # keep only the x axis (this is the center plot)
+       x="Commuting Distance (km)", y = "", color = "Legend")  + 
+  theme(axis.title=element_text(size=15)) -> b
+
+
+### current cycling trips vs potential cycling trips
+cols <- c("Bicycle (Potential)" = "darkgreen", "Bicycle (Current)" = "darkred")
+ggplot(histogram, aes(x=dist)) + 
+  geom_histogram(data=flow_long_potential, color = 'grey50', aes(fill = "Bicycle (Potential)")) +
+  geom_histogram(data= flow_long_bike, color = 'grey50', aes(fill = "Bicycle (Current)")) +
+  scale_fill_manual(name = "", values = cols) +
+  labs(title = "Current vs Potential Cycling Trips", 
+       # remove both axis labels
[…]

### 2021-04-19 README.Rmd
increase figure label sizes

@@
 title: A Road Segment Prioritization Approach for Cycling Infrastructure
-date: "`r Sys.Date()`"
+#date: "`r Sys.Date()`"
 geometry: margin=2cm
@@
 
-```{r desirefacetcycling, fig.cap="Current and potential cycling demand", out.width="60%", fig.pos="H", fig.align="center", echo=FALSE}
+```{r desirefacetcycling, fig.cap="Current and potential cycling demand", out.width="65%", fig.pos="H", fig.align="center", echo=FALSE}
 knitr::include_graphics("data/Manchester/Plots/desire_facet_cycling.png")
@@
 
-```{r boxplotcircuity, fig.cap="Effect of banning cyclists from trunk and primary roads for all OD pairs (Manchester)", out.width="35%", fig.show='hold', fig.align="center", echo=FALSE}
+```{r boxplotcircuity, fig.cap="Effect of banning cyclists from trunk and primary roads for all OD pairs (Manchester)", out.width="50%", fig.show='hold', fig.align="center", echo=FALSE}
 knitr::include_graphics("data/Manchester/Plots/boxplot_weighted_unweighted_distances.png")
@@
 
-```{r cyclinghistmanc, fig.cap="Cycling commuting distance - Manchester (2011 census data)", out.width="30%", fig.show='hold', fig.align = "center", echo=FALSE}
+```{r cyclinghistmanc, fig.cap="Cycling commuting distance - Manchester (2011 census data)", out.width="40%", fig.show='hold', fig.align = "center", echo=FALSE}
 knitr::include_graphics("data/Manchester/Plots/histogram_distance_cycling.png")
@@
 
-```{r communitiesmanchester, fig.cap="Communities based on potential cycling demand between MSOAs (Manchester)", out.width="75%", fig.show='hold', fig.align = "center", echo=FALSE}
+```{r communitiesmanchester, fig.cap="Communities based on potential cycling demand between MSOAs (Manchester)", out.width="80%", fig.show='hold', fig.align = "center", echo=FALSE}
 knitr::include_graphics("data/Manchester/Plots/communities_alternative_Manchester.png")
@@
 
-```{r growthtotal, fig.cap="Comparing overall (dashed) and community level person-km satisfied (Manchester)", out.width="35%", fig.show='hold', fig.align = "center", echo=FALSE}
+```{r growthtotal, fig.cap="Comparing overall (dashed) and community level person-km satisfied (Manchester)", out.width="48%", fig.show='hold', fig.align = "center", echo=FALSE}
 knitr::include_graphics(c(
@@
 
-```{r growth3MapandBar, fig.cap="Road segment priority (left), disaggregated by road type (right) - egalitarian growth", out.width="35%", fig.show='hold', fig.align="center", echo=FALSE}
+```{r growth3MapandBar, fig.cap="Road segment priority (left), disaggregated by road type (right) - egalitarian growth", out.width="45%", fig.show='hold', fig.align="center", echo=FALSE}
 knitr::include_graphics(c(
@@
 ))
+
 ```
@@
 
-```{r componentsandGCC, fig.cap="Network characteristics", out.width="30%", fig.show='hold', fig.pos="H", fig.align = "center", echo=FALSE}
-knitr::include_graphics(c(
-  "data/Manchester/Plots/Growth_Results/growth_util_egal_components_number_comparisonManchester.png",
-  "data/Manchester/Plots/Growth_Results/growth_util_egal_components_gcc_comparisonManchester.png"
-))
+```{r componentsandGCC, fig.cap="Network characteristics", out.width="85%", fig.show='hold', fig.pos="H", fig.align = "center", echo=FALSE}
+# knitr::include_graphics(c(
+#   "data/Manchester/Plots/Growth_Results/growth_util_egal_components_number_comparisonManchester.png",
+#   "data/Manchester/Plots/Growth_Results/growth_util_egal_components_gcc_comparisonManchester.png"
+# ))
+knitr::include_graphics("data/Manchester/Plots/Growth_Results/growth_util_egal_components_gcc_components_together_Manchester.png")
+
 ```

### 2021-04-19 R/_2.0_distance_and_elevation.R

@@ -287,12 +287,12 @@ dist_mat_all_box %>%
   filter(value <= 3) %>%   # because there are a few very large values that ruin the scale of the plot
 ggplot(aes(value, circuity)) + 
   geom_boxplot(outlier.alpha = 0.5) +
-  scale_x_continuous(name = 'Distance Relative to Unweighted Shortest Path (Ratio)', 
+  scale_x_continuous(name = 'Distance Relative to Unweighted \nShortest Path (Ratio)', 
                      breaks = seq(1, 3, by = 0.2)) +  # seq(1, max(dist_mat_all_box$value, , by = 0.2)
   scale_y_discrete(name = '', labels=c("Weighted", "Weighted \n(No Trunk \n or Primary)")) +
   #xlim(NA, 3) +
   #ggtitle("Comparing Weighted to Unweighted Shortest Paths") +
-  theme_minimal()
+  theme_minimal(base_size = 13)
 
 ggsave(paste0("data/", chosen_city,"/Plots/boxplot_weighted_unweighted_distances.png"))
 

### 2021-04-19 R/_3.1_plot_mode_shares.R

@@ -100,7 +100,7 @@ flow_long_bike$dist <- flow_long_bike$dist / 1000
 
 ggplot(flow_long_bike, aes(x = dist)) + 
   geom_histogram(color = "black", alpha = 0.4, binwidth = 1) +
-  theme_minimal() +
+  theme_minimal(base_size = 13) +
   labs(title = "Trips Made By Bicycle", x="Commuting Distance (km)", y = "No. of Commuters")
 
 ggsave(paste0("data/", chosen_city,"/Plots/histogram_distance_cycling.png"))

### 2021-04-19 R/_8.1_growth_utilitarian.R

@@ -142,7 +142,7 @@ ggplot(data=grow_util_c, mapping=aes(x=dist_c, y = perc_person_km_comm_c)) +
   ggtitle("Algorithm 1 (Utilitarian)") +
   labs(x = "Total Length of Investment (km)", y = "% of person km satisfied within community",
        subtitle="Segments Prioritized Based On **Flow**") +
-  theme_minimal() +
+  theme_minimal(base_size = 13) +
   theme(plot.title = element_text(size = 14)) +
   theme(plot.subtitle = element_markdown(size = 10)) +
   scale_x_continuous(expand = c(0, 0), limits = c(0, NA)) + 

### 2021-04-19 R/_8.2_growth_egalitarian.R

@@ -1,7 +1,9 @@
 library(sf)
 library(tidyverse)
 library(ggtext)
+library(patchwork)
 library(tmap)
+
 
 graph_sf <- readRDS(paste0("data/", chosen_city, "/graph_with_flows_weighted_communities.Rds"))
 
@@ -132,7 +134,7 @@ ggplot(data=grow_egal_c, mapping=aes(x=dist_c, y = perc_person_km_comm_c)) +
   ggtitle("Algorithm 2 (Egalitarian)") +
   labs(x = "Total Length of Investment (km)", y = "% of person km satisfied within community",
        subtitle="Segments Prioritized Based On **Flow**") +
-  theme_minimal() +
+  theme_minimal(base_size = 13) +
   theme(plot.title = element_text(size = 14)) +
   theme(plot.subtitle = element_markdown(size = 10)) +
   scale_x_continuous(expand = c(0, 0), limits = c(0, NA)) + 
@@ -179,14 +181,15 @@ grow_egal_c %>% st_drop_geometry() %>%
 ggplot(data=p , aes(x=highway, y=dist, fill = distance_groups)) +
   geom_col(position = position_stack(reverse = TRUE)) +
   scale_fill_brewer(palette = "Blues", direction=-1) +
-  ggtitle("Investment on Different Highway Types") +
-  labs(x = "Highway Type", y = "Length (km)", fill = "Investment Priority \n(km groups)") +
+  ggtitle("Investment on Different Road Types") +
+  labs(x = "Road Type", y = "Length (km)", fill = "Investment Priority \n(km groups)") +
   theme_minimal() +
   # edit angle of text, hjust argument so that text stays below plot AND center plot title
   theme(axis.text.x = element_text(angle=50, hjust=1), plot.title = element_text(hjust = 0.5)) +
   coord_flip()
 
-ggsave(paste0("data/", chosen_city,"/Plots/Growth_Results/growth_egalitarian_investment_highways_flow.png"))
+ggsave(paste0("data/", chosen_city,"/Plots/Growth_Results/growth_egalitarian_investment_highways_flow.png"),
+       width = 15, height = 18, units = "cm")
 
 ###### MAP #####
 
@@ -322,10 +325,10 @@ ggplot(data=components_util_egal,
        title = "Connectivity",
        #subtitle = chosen_city,
        color = "") +
-  theme_minimal() +
+  theme_minimal(base_size = 13) +
[…]

### 2021-05-07 README.Rmd
update readme.md

@@
 abstract: "Understanding the motivators and deterrents to cycling is essential for creating infrastructure that gets more people to adopt cycling as a mode of transport.
-This paper demonstrates a new approach to support the prioritisation of cycling infrastructure and cycling network design, accounting for cyclist preferences and the growing emphasis on 'filtered permeability' and 'Low Traffic Neighborhood' interventions internationally.
+This paper demonstrates a new approach to support the prioritization of cycling infrastructure and cycling network design, accounting for cyclist preferences and the growing emphasis on 'filtered permeability' and 'Low Traffic Neighborhood' interventions internationally.
 The approach combines distance decay, route calculation, and network analysis methods to examine where future cycling demand is most likely to arise, how such demand could be accommodated within existing street networks, and how to ensure a fair distribution of investment.
-Although each of these methods has been applied to cycling infrastructure prioritisation in previous research, this is the first time that they have been combined, creating an integrated road segment prioritisation approach.
+Although each of these methods has been applied to cycling infrastructure prioritization in previous research, this is the first time that they have been combined, creating an integrated road segment prioritization approach.
 The approach, which can be applied to other cities, as shown in the Appendix, is demonstrated in a case study of Manchester, resulting in cycling networks that balance directness against the need for safe and stress-free routes under different investment scenarios.
@@
    bookdown::pdf_document2:
-  # # bookdown::word_document2:
+ # bookdown::word_document2:
       highlight: null   # to remove code highlighting in r chunks
@@
 csl: elsevier-harvard.csl
+always_allow_html: yes
 ---
@@
 ```
-
-<!-- # Ideas and discussion -->
-
-<!-- - Intro: possible trim down substantially then add some literature in (a) Calculating Potential Demand [DONE]/ (b) Routing [DONE]/ (c) Road Segment Prioritization -->
-
-<!-- - Community Detection: Where in the document should this section be? It is now before Road Segment Prioritization, but I can merge them -->
-
-<!-- Can go into road segment section. -->
-
-<!-- - Should we use other cities for the sake of comparison (compare communities, road types etc) -->
-
-<!-- - How to add tables properly (they are now added using latex code which only shows up when outputting to pdf) -->
-
-<!-- - Footnotes are sometimes split between pages. How do I keep them on the same page? -->
-
-<!-- - What is the best way to write pseudocode (Steps for Algorithm 1 and 2) -->
-
-<!-- Keywords -->
-
-<!-- ======== -->
 
@@
 cyclist preferences and government priorities. Section
-\@ref(road-segment-prioritisation) outlines a method for partitioning the
+\@ref(road-segment-prioritization) outlines a method for partitioning the
 study area based on a community finding algorithm and routed cycling
@@
 Some use existing cycling demand, some calculate potential cycling demand, and others ignore demand completely. 
-@duthie2014optimization note that relying only on existing cycling activity to prioritise cycling
+@duthie2014optimization note that relying only on existing cycling activity to prioritize cycling
 infrastructure can reinforce existing cycling patterns and ignore potential cycling demand that could be satisfied by a connected network.
@@
 
-# Road Segment Prioritisation
+# Road Segment Prioritization
 
@@
 
-A range of algorithms could be used for prioritisation. Because policy
+A range of algorithms could be used for prioritization. Because policy
 priorities vary, we present two algorithms. Both utilize existing
@@
 needed, opening up the possibility of deploying the approach in other
-cities^[The results are easily reproducible for all UK cities, and can
-also be reproduced for cities elsewhere given the availability of
-commuter data. The appendix shows the results for 3 other UK cities: Birmingham, Leeds and Nottingham. 
-For instructions on reproducing the results shown in this
-paper, see the **README** file in the R folder of the .zip file that
-accompanies this paper (to be released as open source software when the
-paper is published).].
+cities^[The results are easily reproducible for all UK cities, and can also be reproduced for cities elsewhere given the availability of commuter data.
+Instructions for reproducing the results are detailed in the **README** file in the R folder of this [Github repository](https://github.com/Hussein-Mahfouz/cycle-networks)]
+
+<!-- [The results are easily reproducible for all UK cities, and can -->
+<!-- also be reproduced for cities elsewhere given the availability of -->
+<!-- commuter data. The appendix shows the results for 3 other UK cities: Birmingham, Leeds and Nottingham.  -->
+<!-- For instructions on reproducing the results shown in this -->
+<!-- paper, see the **README** file in the R folder of the .zip file that -->
+<!-- accompanies this paper (to be released as open source software when the -->
+<!-- paper is published).]. -->
 
@@
 *and* investment that addresses geographical and social inequalities.
-This ability to address inequalities in network prioritisation is particularly
+This ability to address inequalities in network prioritization is particularly
 important given research showing substantial *inequalities*

### 2021-05-07 appendix.Rmd
update readme.md

--- 2021-03-16 appendix.Rmd
@@
 title: "Appendix"
-author: "Hussein Mahfouz, Robin Lovelace, Elsa Arcaute"
-date: "13/03/2021"
+#author: "Hussein Mahfouz, Robin Lovelace, Elsa Arcaute"
+#date: "13/03/2021"
 geometry: margin=1.5cm

### 2021-05-07 R/_1.0_get_flow_data.R

@@ -5,7 +5,7 @@ library(ggplot2)
 library(ggrepel)
 
 # this is a lookup table matching MSOAs to major towns and cities
-city_names <- read_csv('data-raw/Middle_Layer_Super_Output_Area__2011__to_Major_Towns_and_Cities__December_2015__Lookup_in_England_and_Wales.csv') 
+city_names <- read_csv('data-raw/MSOA_2011_to_Major_Towns_and_Cities.csv') 
 # change column name
 city_names <- city_names %>% rename(city = TCITY15NM)
 

### 2021-05-07 R/_2.0_distance_and_elevation.R

@@ -184,7 +184,7 @@ flows_slope <- flows_slope %>% left_join(msoa_centroids_snapped[,c('msoa11cd' ,'
 net <- dodgr::dodgr_streetnet(pts = pts, expand = 0.1)
 
 # load in elevation data
-uk_elev <- raster::raster('data/uk_elev.tif')
+#elev <- raster::raster('data/uk_elev.tif')
 
 # 1. get geometries
 route <- function(df, net){
@@ -208,7 +208,8 @@ slope <- function(df, elev){
 }
 
 # get routes column then slope column (function 1 then function 2)
-flows_slope <- flows_slope %>% route(net = net) %>% slope(elev = uk_elev)
+#flows_slope <- flows_slope %>% route(net = net) %>% slope(elev = uk_elev)
+flows_slope <- flows_slope %>% route(net = net) %>% slope(elev = city_elev)
 
 flows_slope %>% 
   dplyr::select(-c(cent_orig, cent_dest, route)) %>%

### 2021-05-08 arxiv.Rmd
arxiv tex file

--- 2021-05-07 README.Rmd
@@
 title: A Road Segment Prioritization Approach for Cycling Infrastructure
-#date: "`r Sys.Date()`"
+author:
+  - Hussein Mahfouz:
+      email: h.mahfouz.19@ucl.ac.uk
+      institute: casa
+      correspondence: true
+  - name: Elsa Arcaute
+    institute: casa
+  - name: Robin Lovelace
+    institute: its
+institute:
+  - casa: The Centre for Advanced Spatial Analysis, University College London, London, UK
+  - its: Institute of Transport Studies, University of Leeds, Leeds, UK
+
 geometry: margin=2cm
@@
       extra_dependencies: ["subcaption", "longtable", "float", "multirow", "placeins", "multicol"]
+      # https://stackoverflow.com/questions/52918716/authors-and-affiliations-in-the-yaml-of-rmarkdown
+      pandoc_args:
+      - '--lua-filter=extras/scholarly-metadata.lua'
+      - '--lua-filter=extras/author-info-blocks.lua'
+      keep_tex:  true
 editor_options:
@@
 # elsevier style for referencing (from https://github.com/citation-style-language/styles)
-csl: elsevier-harvard.csl
+csl: extras/elsevier-harvard.csl
 always_allow_html: yes
@@
 citr::tidy_bib_file(rmd_file = "README.Rmd", messy_bibliography = "~/uaf/allrefs.bib", file = "rlbibfile.bib")
-```
-
-```{r, echo=FALSE, eval=FALSE}
-file.rename("README.pdf", "cycle-network-paper.pdf")
 ```
@@
 Another defining feature for cycling infrastructure is how well
-connected it is. Cyclists prefer cohesive infrastructure, particularly
+connected it is; cyclists prefer cohesive infrastructure, particularly
 when cycling on arterial roads with high levels of motorized traffic
@@
 \noindent where $d$ and $s$ are the distance and slope associated with each
-OD pair, and $\alpha$ and all $\beta$s, are parameters calculated by a regression
+OD pair, and $\alpha$ and all $\beta$s are parameters calculated by a regression
 model on the training data. The square and square-root distance terms
@@
 
+```{r flowsfacetunweighted, fig.cap="Flow results based on unweighted shortest paths (Manchester)", out.width="90%", fig.show='hold', echo=FALSE}
+knitr::include_graphics("data/Manchester/Plots/flows_facet_unweighted_Manchester.png")
+```
+
+```{r flowsfacetweighted, fig.cap="Flow results based on weighted shortest paths (Manchester)", out.width="90%", fig.show='hold', echo=FALSE}
+knitr::include_graphics("data/Manchester/Plots/flows_facet_weighted_Manchester.png")
+```
+
 Comparing the cycling demand routed on the weighted and unweighted road
@@
 through tertiary roads, as expected.
-
-```{r flowsfacetunweighted, fig.cap="Flow results based on unweighted shortest paths (Manchester)", out.width="90%", fig.show='hold', echo=FALSE}
-knitr::include_graphics("data/Manchester/Plots/flows_facet_unweighted_Manchester.png")
-```
-
-```{r flowsfacetweighted, fig.cap="Flow results based on weighted shortest paths (Manchester)", out.width="90%", fig.show='hold', echo=FALSE}
-knitr::include_graphics("data/Manchester/Plots/flows_facet_weighted_Manchester.png")
-```
-
-<!-- The difference between aggregate flow on weighted and unweighted networks is dependant on the road network of the city.
-Comparing Manchester to Nottingham, we see that trunk roads are much more important in the former, as over 25\% of flow on the unweighted road network passes through them.
-For Nottingham, less than 10\% of the flow on the unweighted network passes through trunk roads, but almost 25\% of the flow passes through tertiary roads (Figure \ref{fig:perc_person-km}).
--->
 
@@
 
-```{r boxplotcircuity, fig.cap="Effect of banning cyclists from trunk and primary roads for all OD pairs (Manchester)", out.width="50%", fig.show='hold', fig.align="center", echo=FALSE}
-knitr::include_graphics("data/Manchester/Plots/boxplot_weighted_unweighted_distances.png")
-```
-
 Weighting the road network also allows us to better utilize existing
@@
 
+```{r boxplotcircuity, fig.cap="Effect of banning cyclists from trunk and primary roads for all OD pairs (Manchester)", out.width="50%", fig.show='hold', fig.align="center", echo=FALSE}
+knitr::include_graphics("data/Manchester/Plots/boxplot_weighted_unweighted_distances.png")
+```
+
 # Road Segment Prioritization
@@
 paths are those that have the highest flow passing through them (this is
-a result of the routing in Section \@ref(introduction).
+a result of the routing in Section \@ref(routing).
 
@@
 needed, opening up the possibility of deploying the approach in other
-cities^[The results are easily reproducible for all UK cities, and can also be reproduced for cities elsewhere given the availability of commuter data.
-Instructions for reproducing the results are detailed in the **README** file in the R folder of this [Github repository](https://github.com/Hussein-Mahfouz/cycle-networks)]
+cities.^[The results are easily reproducible for all UK cities, and can also be reproduced for cities elsewhere given the availability of commuter data.
+Instructions for reproducing the results are detailed in the **README** file in the R folder of this [Github repository](https://github.com/Hussein-Mahfouz/cycle-networks)
 
@@
 <!-- paper is published).]. -->
-
-<!-- For after publishing
-
-^[The results are easily reproducible for all UK cities, and can also be reproduced for cities elsewhere given the availability of commuter data.
-Instructions for reproducing the results are detailed in this [Github repository](https://github.com/Hussein-Mahfouz/cycle-networks)].
--->
 
@@
 proliferation of these modes, raising the point that surveys could
-replace the broad category of `Bicycle' a range of small modes such as
+replace the broad category of `Bicycle' with a range of small modes such as
 'pedal cycles' (including tricycles and recumbent cycles), e-bikes
@@
 the humble bicycle is a key ingredient in the ongoing shift towards active transport.
-Recognition of the importance of this shift for improved the health and well-being
-the citizens, and the wider challenge to decarbonise the global economy,
+Recognition of the importance of this shift for improved health and well-being
+of citizens, and the wider challenge to decarbonise the global economy,
 has grown with pressures on health and public transport systems during the pandemic.
@@
 for designing joined-up and cost-effective strategic cycle networks.
-<!-- In tandem with policies to reduce excessive reliance on private cars, evidence-based investment in strategic cycle networks is key to. -->
-
-<!-- I think it would be good to say how others could reproduce the methods in other cities (RL) -->
-
-<!-- (HM) Answer: Done. I added a footnote to the first paragraph in the conclusion linking to the github repo -->
-
-<!-- There should also be links to the literature. -->
-
-<!-- (HM) Answer: There are links to the literature at the end of each thematic section (Calculating Potential Cycling Demand / Routing / Road Segment Prioritization).
-There is a discussion paragraph in each of them.
-What additional links are necessary here?
--->
+
+\clearpage
 
@@
 ```
+
+<div id="refs"></div>
+
+# (APPENDIX) Appendix {-} 
+\clearpage
+# Appendix: Reproducing the results for other cities
+
+To reproduce the results presented for other cities in the UK there are a few computing pre-requisites:
+
+- A reasonably powerful computer, e.g. with 8 GB RAM
+- The statistical programming language R installed (we recommend running it in the RStudio editor)
+- The source code of the paper repository, which can be download from this [Github repository](https://github.com/Hussein-Mahfouz/cycle-networks)]
+
+```{r, eval=FALSE, echo=FALSE}
+# run these lines to update the snapshot
+required_packages = renv::dependencies()
+renv::snapshot()
+```
+
+To ensure you have the necessary packages installed, open the `cycle-networks.Rproj` file in RStudio and run the following commands to ensure you have the necessary packages installed:
+
+```{r, eval=FALSE}
+install.packages("renv")
+renv::restore()
+```
+
+The `build.R` script contains all of the necessary steps to reproduce the results for any city in the UK.
+Open this file with the following command and run the code line-by-line.
+
+```{r, eval=FALSE}
+file.edit("R/build.R")
+```
+
+<!-- Todo: make the .zip download work for anyone, e.g. by uploading the data to the osf.io website alongside the preprint. -->
+
+You can select a city to build by executing the following command:
+
+```{r}
+chosen_city <- "Leeds"
+```
+
+Replacing `Leeds` with any other city will make the script work for that city.
+Note: this only works for Local Authority names in England and Wales listed below.
+
+In this script, you choose which city you wish to run the analysis from
+this list of available towns and cities:
+
+    ##   [1] "Barnsley"             "Basildon"             "Basingstoke"
+    ##   [4] "Bath"                 "Bedford"              "Birkenhead"
+    ##   [7] "Birmingham"           "Blackburn"            "Blackpool"
+    ##  [10] "Bolton"               "Bournemouth"          "Bracknell"
+    ##  [13] "Bradford"             "Brighton and Hove"    "Bristol"
+    ##  [16] "Burnley"              "Burton upon Trent"    "Bury"
+    ##  [19] "Cambridge"            "Cardiff"              "Carlisle"
+    ##  [22] "Chatham"              "Chelmsford"           "Cheltenham"
+    ##  [25] "Chester"              "Chesterfield"         "Colchester"
+    ##  [28] "Coventry"             "Crawley"              "Darlington"
+    ##  [31] "Derby"                "Doncaster"            "Dudley"
+    ##  [34] "Eastbourne"           "Exeter"               "Gateshead"
+    ##  [37] "Gillingham"           "Gloucester"           "Grimsby"
+    ##  [40] "Guildford"            "Halifax"              "Harlow"
+    ##  [43] "Harrogate"            "Hartlepool"           "Hastings"
+    ##  [46] "Hemel Hempstead"      "High Wycombe"         "Huddersfield"
+    ##  [49] "Ipswich"              "Kingston upon Hull"   "Leeds"
+    ##  [52] "Leicester"            "Lincoln"              "Liverpool"
+    ##  [55] "London"               "Luton"                "Maidstone"
+    ##  [58] "Manchester"           "Mansfield"            "Middlesbrough"
+    ##  [61] "Milton Keynes"        "Newcastle upon Tyne"  "Newcastle-under-Lyme"
+    ##  [64] "Newport"              "Northampton"          "Norwich"
+    ##  [67] "Nottingham"           "Nuneaton"             "Oldham"
+    ##  [70] "Oxford"               "Peterborough"         "Plymouth"
+    ##  [73] "Poole"                "Portsmouth"           "Preston"
+    ##  [76] "Reading"              "Redditch"             "Rochdale"
+    ##  [79] "Rotherham"            "Salford"              "Scunthorpe"
+    ##  [82] "Sheffield"            "Shrewsbury"           "Slough"
+    ##  [85] "Solihull"             "South Shields"        "Southampton"
+    ##  [88] "Southend-on-Sea"      "Southport"            "St Albans"
+    ##  [91] "St Helens"            "Stevenage"            "Stockport"
+    ##  [94] "Stockton-on-Tees"     "Stoke-on-Trent"       "Sunderland"
+    ##  [97] "Sutton Coldfield"     "Swansea"              "Swindon"
+    ## [100] "Telford"              "Wakefield"            "Walsall"
+    ## [103] "Warrington"           "Watford"              "West Bromwich"
+    ## [106] "Weston-Super-Mare"    "Wigan"                "Woking"
+    ## [109] "Wolverhampton"        "Worcester"            "Worthing"
+    ## [112] "York"
+
+\newpage
+
+## Birmingham
+
+\subsubsection{Potential Demand}
+
+```{r potdemhistogramsBirmingham, fig.cap="Distribution of Potential Cycling Demand", out.width="30%", fig.show='hold', echo=FALSE}
+knitr::include_graphics(c(
+  "data/Birmingham/Plots/histogram_distance_all_vs_cycling.png",
+  "data/Birmingham/Plots/histogram_distance_all_vs_cycling_potential.png",
+  "data/Birmingham/Plots/histogram_distance_cycling_potential_vs_current.png"
+))
+```
+
+```{r desirefacetcyclingBirmingham, fig.cap="Current and Potential Cycling Demand", out.width="80%", fig.pos="H", fig.align="center", echo=FALSE}
+knitr::include_graphics("data/Birmingham/Plots/desire_facet_cycling.png")
+```
+
+\clearpage
+
+\subsubsection{Routing Cycling Flows}
+
+```{r flowsfacetunweightedBirmingham, fig.cap="Flow Results Based on Unweighted Shortest Paths (Birmingham)", out.width="75%", fig.show='hold', fig.align = "center", echo=FALSE}
+knitr::include_graphics("data/Birmingham/Plots/flows_facet_unweighted_Birmingham.png")
+```
+
+```{r flowsfacetweightedBirmingham, fig.cap="Flow Results Based on Weighted Shortest Paths (Birmingham)", out.width="75%", fig.show='hold', fig.align = "center", echo=FALSE}
+knitr::include_graphics("data/Birmingham/Plots/flows_facet_weighted_Birmingham.png")
+```
+
+\clearpage
+
+\subsubsection{Community Detection}
+
+```{r communitiesBirmingham, fig.cap="Communities Based on Potential Cycling Demand (Birmingham)", out.width="90%", fig.show='hold', fig.align = "center", echo=FALSE}
+knitr::include_graphics("data/Birmingham/Plots/communities_alternative_Birmingham.png")
+```
+
+\subsubsection{Network Expansion Algorithms}
+
+```{r growthtotalBirmingham, fig.cap="Comparing Overall (Dashed) and Community Level Person-Km Satisfied (Birmingham)", out.width="45%", fig.show='hold', fig.pos="H", fig.align = "center", echo=FALSE}
+knitr::include_graphics(c(
+  "data/Birmingham/Plots/Growth_Results/growth_utilitarian_satisfied_km_both_flow_column.png",
+  "data/Birmingham/Plots/Growth_Results/growth_egalitarian_satisfied_km_both_flow_column.png"
+))
+```
+
+```{r componentsandGCCBirmingham, fig.cap="Network Characteristics", out.width="45%", fig.show='hold', fig.pos="H", fig.align = "center", echo=FALSE}
+knitr::include_graphics(c(
+  "data/Birmingham/Plots/Growth_Results/growth_util_egal_components_number_comparisonBirmingham.png",
+  "data/Birmingham/Plots/Growth_Results/growth_util_egal_components_gcc_comparisonBirmingham.png"
+))
+```
+
+```{r growth3MapandBarBirmingham, fig.cap="Road Segment Priority (left), disaggregated by road type (right) - Egalitarian Growth", out.width="45%", fig.show='hold', fig.align="center", echo=FALSE}
+knitr::include_graphics(c(
+  "data/Birmingham/Plots/Growth_Results/growth_egalitarian_priority_all_FLOW.png"#,
+ # "data/Birmingham/Plots/Growth_Results/growth_egalitarian_investment_highways_flow.png"
+))
+```
+
+\clearpage
+
+## Leeds
+
+\subsubsection{Potential Demand}
+
+```{r potdemhistogramsLeeds, fig.cap="Distribution of Potential Cycling Demand", out.width="30%", fig.show='hold', echo=FALSE}
+knitr::include_graphics(c(
+  "data/Leeds/Plots/histogram_distance_all_vs_cycling.png",
+  "data/Leeds/Plots/histogram_distance_all_vs_cycling_potential.png",
+  "data/Leeds/Plots/histogram_distance_cycling_potential_vs_current.png"
+))
+```
+
+```{r desirefacetcyclingLeeds, fig.cap="Current and Potential Cycling Demand", out.width="80%", fig.pos="H", fig.align="center", echo=FALSE}
+knitr::include_graphics("data/Leeds/Plots/desire_facet_cycling.png")
+```
+
+\clearpage
+
+\subsubsection{Routing Cycling Flows}
+
+```{r flowsfacetunweightedLeeds, fig.cap="Flow Results Based on Unweighted Shortest Paths (Leeds)", out.width="75%", fig.show='hold', fig.align = "center", echo=FALSE}
+knitr::include_graphics("data/Leeds/Plots/flows_facet_unweighted_Leeds.png")
+```
+
+```{r flowsfacetweightedLeeds, fig.cap="Flow Results Based on Weighted Shortest Paths (Leeds)", out.width="75%", fig.show='hold', fig.align = "center", echo=FALSE}
+knitr::include_graphics("data/Leeds/Plots/flows_facet_weighted_Leeds.png")
+```
+
+\clearpage
+
+\subsubsection{Community Detection}
+
+```{r communitiesLeeds, fig.cap="Communities Based on Potential Cycling Demand (Leeds)", out.width="90%", fig.show='hold', fig.align = "center", echo=FALSE}
+knitr::include_graphics("data/Leeds/Plots/communities_alternative_Leeds.png")
+```
+
+\subsubsection{Network Expansion Algorithms}
+
+```{r growthtotalLeeds, fig.cap="Comparing Overall (Dashed) and Community Level Person-Km Satisfied (Leeds)", out.width="45%", fig.show='hold', fig.pos="H", fig.align = "center", echo=FALSE}
+knitr::include_graphics(c(
+  "data/Leeds/Plots/Growth_Results/growth_utilitarian_satisfied_km_both_flow_column.png",
+  "data/Leeds/Plots/Growth_Results/growth_egalitarian_satisfied_km_both_flow_column.png"
+))
+```
+
+```{r componentsandGCCLeeds, fig.cap="Network Characteristics", out.width="45%", fig.show='hold', fig.pos="H", fig.align = "center", echo=FALSE}
+knitr::include_graphics(c(
+  "data/Leeds/Plots/Growth_Results/growth_util_egal_components_number_comparisonLeeds.png",
+  "data/Leeds/Plots/Growth_Results/growth_util_egal_components_gcc_comparisonLeeds.png"
+))
+```
+
+```{r growth3MapandBarLeeds, fig.cap="Road Segment Priority (left), disaggregated by road type (right) - Egalitarian Growth", out.width="45%", fig.show='hold', fig.align="center", echo=FALSE}
+knitr::include_graphics(c(
+  "data/Leeds/Plots/Growth_Results/growth_egalitarian_priority_all_FLOW.png",
+  "data/Leeds/Plots/Growth_Results/growth_egalitarian_investment_highways_flow.png"
+))
+```
+
+\clearpage
+
+## Nottingham
+
+\subsubsection{Potential Demand}
+
+```{r potdemhistogramsNottingham, fig.cap="Distribution of Potential Cycling Demand", out.width="30%", fig.show='hold', echo=FALSE}
+knitr::include_graphics(c(
+  "data/Nottingham/Plots/histogram_distance_all_vs_cycling.png",
+  "data/Nottingham/Plots/histogram_distance_all_vs_cycling_potential.png",
+  "data/Nottingham/Plots/histogram_distance_cycling_potential_vs_current.png"
+))
+```
+
+```{r desirefacetcyclingNottingham, fig.cap="Current and Potential Cycling Demand", out.width="80%", fig.pos="H", fig.align="center", echo=FALSE}
+knitr::include_graphics("data/Nottingham/Plots/desire_facet_cycling.png")
+```
+
+\clearpage
+
+\subsubsection{Routing Cycling Flows}
+
+```{r flowsfacetunweightedNottingham, fig.cap="Flow Results Based on Unweighted Shortest Paths (Nottingham)", out.width="75%", fig.show='hold', fig.align = "center", echo=FALSE}
+knitr::include_graphics("data/Nottingham/Plots/flows_facet_unweighted_Nottingham.png")
+```
+
+```{r flowsfacetweightedNottingham, fig.cap="Flow Results Based on Weighted Shortest Paths (Nottingham)", out.width="75%", fig.show='hold', fig.align = "center", echo=FALSE}
+knitr::include_graphics("data/Nottingham/Plots/flows_facet_weighted_Nottingham.png")
+```
+
+\clearpage
+
+\subsubsection{Community Detection}
+
+```{r communitiesNottingham, fig.cap="Communities Based on Potential Cycling Demand (Nottingham)", out.width="90%", fig.show='hold', fig.align = "center", echo=FALSE}
+knitr::include_graphics("data/Nottingham/Plots/communities_alternative_Nottingham.png")
+```
+
+\subsubsection{Network Expansion Algorithms}
+
+```{r growthtotalNottingham, fig.cap="Comparing Overall (Dashed) and Community Level Person-Km Satisfied (Nottingham)", out.width="45%", fig.show='hold', fig.pos="H", bifig.align = "center", echo=FALSE}
+knitr::include_graphics(c(
+  "data/Nottingham/Plots/Growth_Results/growth_utilitarian_satisfied_km_both_flow_column.png",
+  "data/Nottingham/Plots/Growth_Results/growth_egalitarian_satisfied_km_both_flow_column.png"
+))
+```
+
+```{r componentsandGCCNottingham, fig.cap="Network Characteristics", out.width="45%", fig.show='hold', fig.pos="H", fig.align = "center", echo=FALSE}
+knitr::include_graphics(c(
+  "data/Nottingham/Plots/Growth_Results/growth_util_egal_components_number_comparisonNottingham.png",
+  "data/Nottingham/Plots/Growth_Results/growth_util_egal_components_gcc_comparisonNottingham.png"
+))
+```
+
+```{r growth3MapandBarNottingham, fig.cap="Road Segment Priority (left), disaggregated by road type (right) - Egalitarian Growth", out.width="45%", fig.show='hold', fig.align="center", echo=FALSE}
+knitr::include_graphics(c(
+  "data/Nottingham/Plots/Growth_Results/growth_egalitarian_priority_all_FLOW.png",
+  "data/Nottingham/Plots/Growth_Results/growth_egalitarian_investment_highways_flow.png"
+))
+```
+
+\clearpage

### 2021-05-08 arxiv.tex
arxiv tex file

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
  \usepackage{amssymb}
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={A Road Segment Prioritization Approach for Cycling Infrastructure},
  pdfauthor={Hussein Mahfouz1,; Elsa Arcaute1; Robin Lovelace2},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=2cm]{geometry}
\usepackage{longtable,booktabs}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{subcaption}
\usepackage{longtable}
\usepackage{float}
\usepackage{multirow}
\usepackage{placeins}
\usepackage{multicol}
\usepackage{booktabs}
\usepackage{array}
\usepackage{wrapfig}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\title{A Road Segment Prioritization Approach for Cycling Infrastructure}
\author{Hussein Mahfouz\textsuperscript{1,*} \and Elsa Arcaute\textsuperscript{1} \and Robin Lovelace\textsuperscript{2}}
\date{}

\begin{document}
\maketitle
\begin{abstract}
Understanding the motivators and deterrents to cycling is essential for creating infrastructure that gets more people to adopt cycling as a mode of transport. This paper demonstrates a new approach to support the prioritization of cycling infrastructure and cycling network design, accounting for cyclist preferences and the growing emphasis on `filtered permeability' and `Low Traffic Neighborhood' interventions internationally. The approach combines distance decay, route calculation, and network analysis methods to examine where future cycling demand is most likely to arise, how such demand could be accommodated within existing street networks, and how to ensure a fair distribution of investment. Although each of these methods has been applied to cycling infrastructure prioritization in previous research, this is the first time that they have been combined, creating an integrated road segment prioritization approach. The approach, which can be applied to other cities, as shown in the Appendix, is demonstrated in a case study of Manchester, resulting in cycling networks that balance directness against the need for safe and stress-free routes under different investment scenarios. A key benefit of the approach from a policy perspective is its ability to support egalitarian and cost-effective strategic cycle network planning.

\par

\textbf{Keywords:} cycling networks, low-traffic neighborhoods, routing, transport equity
\end{abstract}

\textsuperscript{1} The Centre for Advanced Spatial Analysis, University College London, London, UK\\
\textsuperscript{2} Institute of Transport Studies, University of Leeds, Leeds, UK

\textsuperscript{*} Correspondence: \href{mailto:h.mahfouz.19@ucl.ac.uk}{Hussein Mahfouz \textless{}\href{mailto:h.mahfouz.19@ucl.ac.uk}{\nolinkurl{h.mahfouz.19@ucl.ac.uk}}\textgreater{}}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

The 2015 Paris agreement (UN, 2015) acknowledged that
fundamental changes to societies and economies are necessary to mitigate
climate change. Like other sectors, transport is under substantial
pressure to decarbonise, resulting in a number of technical innovations
including electric vehicles. But new vehicle technologies can only go so
far and do not tackle parallel problems such as congestion, road traffic
casualties and physical inactivity (Brand et al., 2020).

In this context, interest and investment in active modes are growing.
The benefits extend beyond congestion and the environment, as active
travel can also help alleviate what is referred to as the pandemic of
global inactivity; physical inactivity is on the rise and has become the
4th highest cause of death globally (Kohl 3rd et al., 2012). Various
studies have documented the association between active transport and
lower risk of disease, including cancer and cardiovascular disease
(Celis-Morales et al., 2017; Jarrett et al., 2012; Patterson et al., 2020). In the wake of the Covid-19 pandemic, and
the resulting reduced capacity of public transport, the UK government
has pledged to invest billions of pounds to improve walking and cycling
infrastructure across the country. While this unprecedented sum is an
opportunity to reshape cities in a way that improves the well-being of
citizens, it does come with a warning:

\begin{quote}
``Inadequate cycling infrastructure discourages cycling and wastes
public money. Much cycling infrastructure in this country is
inadequate. It reflects a belief, conscious or otherwise, that hardly
anyone cycles, that cycling is unimportant and that cycles must take
no meaningful space from more important road users, such as motor
vehicles and pedestrians'' (DfT, 2020a).
\end{quote}

The funding on its own is therefore no guarantee of a change in
commuting across the country; it must be used to design adequate cycling
infrastructure that is based on motivators and deterrents to cycling.

\hypertarget{motivators-and-deterrents-to-cycling}{%
\subsection{Motivators and Deterrents to Cycling}\label{motivators-and-deterrents-to-cycling}}

Much research has been done to understand what gets people to cycle.
Segregated cycling infrastructure\footnote{\emph{Segregated cycling infrastructure}
  refers to road space that is allocated to cyclists only, with physical
  separation to protect cyclists from other modes of transport.} has been
shown to increase cycling uptake (Aldred et al., 2019; Goodman et al., 2014; Marqués et al., 2015), with the separation from motorized
vehicles being key (Winters et al., 2011). Revealed preference of
cyclists shows that they are willing to deviate from the most efficient
routes in order to commute on safer roads (Crane et al., 2017).
However, such deviations are only considered if they do not considerably
increase route circuitry; behaviour studies have found that the
probability of choosing a route decreases in proportion to its length
relative to the shortest route (Broach et al., 2011; Winters et al., 2010).
Another defining feature for cycling infrastructure is how well
connected it is; cyclists prefer cohesive infrastructure, particularly
when cycling on arterial roads with high levels of motorized traffic
(Stinson and Bhat, 2003). The lack of well-connected cycling
infrastructure is one of the main obstacles to increasing cycling uptake
(Caulfield et al., 2012). While direct and cohesive cycling networks
have been shown to positively impact cycling rates, density\footnote{making an
  area's bicycle network denser means adding more cycling routes in the
  area and thereby giving cyclists more route options.} of the cycling
network is also vital (Schoner and Levinson, 2014).

\hypertarget{network-level-approaches}{%
\subsection{Network-Level Approaches}\label{network-level-approaches}}

The studies outlined above lay out the fundamentals for designing
cycling networks that generate significant cycling uptake, but they do
not propose network-level interventions. In this section we outline
methods used in past studies, namely optimization and network
analysis methods, such as connected components and community
detection, and examine how they are leveraged to suggest cycling network
designs. We compare the effectiveness of these network-level studies in
incorporating the fundamentals outlined above.
Our proposed approach is inspired by these methods, but it attempts to add to them
by ensuring that all of the outlined fundamentals are accounted for. It
also goes further by attempting to factor in ethical considerations
related to the distribution of investment.

\emph{Optimization} methods have been used to propose improvements to
cycling networks. Mesbah et al. (2012) propose a bi-level formulation to
optimize allocation of cycling lanes to the network without exceeding a
set budget. They account for the effect of cycling lanes on car traffic,
and attempt to maximize utilization of said lanes with minimal impact on
car travel times. Safeguarding against increased car traffic may be counter-productive if the goal is to create a mode-shift, as research has shown that reducing road space for cars leads to less cars on the road, a phenomenon referred to as ``traffic evaporation'' (Nello-Deakin, 2020).
On the other hand, Mauttone et al. (2017) developed an optimization framework aiming at minimizing the total
user cost of cycling on the network without considering car usage. The aggregate flow\footnote{\emph{flow} is
  used throughout this research to refer to the cycling demand when it is
  routed onto the road network. The flow on any road segment is the
  cumulative demand on it, resulting from cyclists commuting between
  various Origin-Destination (OD) pairs.} on road segments (links) is obtained by using
shortest paths to route existing cycling demand onto the road network. The solution is a proposed set of links where cycling infrastructure
should be added in order to minimize the overall travel cost of cyclists
across the network. The cost of traversing a link is given as a function
of its length and whether or not it has cycling infrastructure. The
problem has also been solved by attempting to find the minimum cost of
improving roadway links to meet a desired level of service (LOS)
(Duthie and Unnikrishnan, 2014). In this formulation, all origin-destination
(OD) pairs need to be connected by roads that meet the desired LOS. A
limitation of these approaches is that they do not explicitly solve for
continuity. The latter is addressed using either a constraint specifying
that each link with a bike lane should be connected to at least one
destination (Mesbah et al., 2012), a constraint on maximum deviation
from shortest paths (Duthie and Unnikrishnan, 2014), or a discontinuity
penalty to prioritize connected road segments (Mauttone et al., 2017).

In this paper, continuity is analysed by looking at the connectivity of
the network through the graph-theoretic concept of \emph{connected
components}. Natera et al. (2019) for example,
study the existing cycling network in terms of its disconnected
components. They propose two different algorithms to connect disconnected components; one connects the two largest components, and the other connects the largest component to its nearest neighbor. By measuring the growth of the largest connected component as a function of the kilometers of network added, they determine that these approaches are more effective at improving cycling network connectivity than random allocation of cycling infrastructure.
The concept of connected components is
also at the core of the methodology proposed by Olmos et al. (2020). After
routing the cycling demand onto the network links, they use percolation
theory to filter out the links based on the aggregate flow passing through them, varying the flow threshold for
filtering to identify the minimum flow at which the whole city is
connected by a giant component. While these approaches deal with
continuity better, they look at the network as a whole when attempting
to improve it, and in doing so fail to account for equitable
distribution of infrastructure.

\hypertarget{ethical-underpinnings-and-proposed-approach}{%
\subsection{Ethical Underpinnings and Proposed Approach}\label{ethical-underpinnings-and-proposed-approach}}

All of these network-level methodologies are underpinned by ethical
principles, even though these principles are not explicitly acknowledged
by the authors. This is important since different ethical principles
constitute different problem formulations and targets. Broadly speaking,
transport appraisal can be based on either utilitarian or egalitarian
principles. The former seeks to maximize the overall benefit, while the
latter is concerned with a fair distribution of benefits
(Jafino et al., 2020).
The utilitarian approach, historically popular in transport planning, has been criticised for focusing on the bigger picture and failing to account for the distribution of investments on the different communities of the study area (Lucas et al., 2016; Nahmias-Biran et al., 2017).
Pereira et al. (2017) emphasize the need for a more
egalitarian approach to transport planning. They highlight accessibility
as a cornerstone of distributive justice, and contend that policies
should aim to distribute investments in a way that minimizes spatial
variations in accessibility.

This research attempts to propose an egalitarian framework for cycling
network design. This is done by identifying the different sub-networks
that exist within the larger network, and ensuring that each gets a fair
share of investment. Trip patterns in a city are not uniformly
distributed geographically, and \emph{community finding} methods have been
used to partition study areas into localized areas that experience a
disproportionate number of trips within them. Akbarzadeh et al. (2018)
use a modularity maximization approach (Blondel et al., 2008) on taxi trip
data to identify 7 different communities in the city of Isfahan, Iran.
An optimization problem is then formulated to connect nodes within each
community with cycling infrastructure, with the emphasis being on
connectivity within the communities, not between them. Bao et al. (2017)
adopt a similar methodology, first identifying communities and then
using a greedy network expansion algorithm to simultaneously add links
to each community. The link with the highest benefit-cost ratio in each
community is selected, and the network is grown by adding neighboring
links to the solution until a budget limit is met. The benefit is the
flow on the link, and each link is assigned a cost based on current road
conditions.

Our work builds on these \emph{community finding} approaches by proposing a
similar greedy network expansion algorithm for cycle network expansion
within communities. We incorporate community finding methods for study
area partitioning with weighted routing to avoid links that are
stressful to cycle on. In doing so, we propose an approach that
accounts for motivators and deterrents to cycling. We propose three
sub-methods that address some of the limitations of previous
studies. These limitations include (a) an inherent bias of basing network
design solely on existing cycling demand, (b) the proposal of routes that may
not correspond to studies on cyclist preference and government policies,
and (c) an insufficient consideration of the ethical principles underlying
the analysis. The current work addresses these limitations through the following organisation. Section \ref{calculating-potential-cycling-demand}
focuses on calculating potential cycling demand. Section \ref{routing}
focuses on routing the demand onto the road network while accounting for
cyclist preferences and government priorities. Section
\ref{road-segment-prioritization} outlines a method for partitioning the
study area based on a community finding algorithm and routed cycling
demand. It then introduces the network expansion algorithms, and compares an approach grounded in
`egalitarianism' to one grounded in `utilitarianism.'

\hypertarget{calculating-potential-cycling-demand}{%
\section{Calculating Potential Cycling Demand}\label{calculating-potential-cycling-demand}}

Many of the cycling network studies mentioned above use demand data for cycling as a starting point.
Some use existing cycling demand, some calculate potential cycling demand, and others ignore demand completely.
Duthie and Unnikrishnan (2014) note that relying only on existing cycling activity to prioritize cycling
infrastructure can reinforce existing cycling patterns and ignore potential cycling demand that could be satisfied by a connected network.
To avoid this issue they choose to ignore existing demand completely, and focus on creating a network that connects the
entire study area.
Olmos et al. (2020) opt to calculate potential demand instead; they obtain the distance distribution of cyclists using a smartphone-based bicycle GPS data, and then use a rejection-sampling algorithm on the OD data of the study area to match
the potential demand distribution to the distribution obtained from GPS data.

\emph{OD} data can be obtained from a range of sources,
including GPS data, household travels surveys and Census data on work
locations. In areas where observed OD data is unavailable, modelling
techniques such as spatial interaction models can be used to estimate
travel volumes between zones (Black, 1995; Martínez and Viegas, 2013; Wilson, 1971). In this paper we use open access data from the UK
census (ONS, 2011), which contains aggregate statistics on
number of commuters between administrative zones --- Middle layer Super
Output Areas (MSOA) --- by mode of travel. MSOAs have an average
population of just over 8000 (ONS, 2018).\footnote{ See
  \url{https://wicid.ukdataservice.ac.uk/} for open access to the OD data.}
Figures \ref{fig:potdemhistograms} and \ref{fig:desirefacetcycling}
illustrate the proportion of trips cycled by distance and the
geographic extent of the input OD dataset used in this paper.

For our purposes, we use a logistic regression model to calculate
potential cycling demand. The model was adapted from the Propensity to
Cycle Tool (PCT), which estimates the proportion of trips
(\(\boldsymbol{C_{p}}\)) for each OD pair that would cycle under
different scenarios of change as a
function of distance and hilliness (Lovelace et al., 2017). We used the Government Target scenario,
indicating a nationwide target of doubling cycling by 2025. The
logistic regression model used to calculate \(\boldsymbol{C_{p}}\) has
the following parameters:

\begin{align}\label{eq:pcteqn}
     logit(C_{p}) = & \alpha + \beta_1 d + \beta_2 \sqrt{d} + \beta_3 d^2 + \beta_4 s + \beta_5 ds + \beta_6 \sqrt{d}s 
\end{align}

\noindent where \(d\) and \(s\) are the distance and slope associated with each
OD pair, and \(\alpha\) and all \(\beta\)s are parameters calculated by a regression
model on the training data. The square and square-root distance terms
``capture the non-linear impact of distance on the likelihood of
cycling,'' and interaction terms to capture the combined effect of slope
and distance (Lovelace et al., 2017). Alternative cycling uptake
models could be `plugged in' to our approach for different contexts or scenarios of change.

The potential demand calculations show that the current and potential
number of cyclists both follow a bell-shaped distribution, with the
number of trips peaking around the 3-4km commuting distance and then
going back down for longer distances (see Figures
\ref{fig:potdemhistograms} and \ref{fig:desirefacetcycling}).

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{data/Manchester/Plots/histogram_distance_cycling_potential_current_all}

}

\caption{Distribution of potential cycling demand}\label{fig:potdemhistograms}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=0.65\linewidth]{data/Manchester/Plots/desire_facet_cycling}

}

\caption{Current and potential cycling demand}\label{fig:desirefacetcycling}
\end{figure}

The uptake model used in this paper assumes uptake of traditional bicycles,
which are affected by topography and distance due to physical effort.
As discussed in Section \ref{discussion-and-conclusions}, using adapted uptake
models could enable the approach to help plan for solutions such as e-bikes
that enable trips covering longer distances and hillier roads.

\hypertarget{routing}{%
\section{Routing}\label{routing}}

The next step in our approach is to route the potential cycling demand
(\(\boldsymbol{C_{p}}\)) between all OD pairs onto the road network. We
choose not to use the PCT approach for routing as it is reliant on
external and therefore inflexible routing services.

\begin{longtable}{lll}
\caption{OSM road types}
\label{table:osmroadtypes}\\
\textbf{OSM Road Type } & \textbf{Description}                                                                                                                                            & \textbf{UK Equivalent}                                                                                                   \endfirsthead 
\hline
Motorway                & \begin{tabular}[c]{@{}l@{}}Road open to high-speed \\vehicles only\end{tabular}                                                                                 & Motorway                                                                                                                 \\* 
\hline
Trunk                   & \multirow{5}{*}{\begin{tabular}[c]{@{}l@{}}Roads that don't \\meet motorway criteria, \\in descending order \\of importance and \\through-traffic\end{tabular}} & A-Roads with Primary Status                                                                                              \\* 
\cline{3-3}
Primary                 &                                                                                                                                                                 & A-Roads with Non-Primary Status                                                                                          \\* 
\cline{3-3}
Secondary               &                                                                                                                                                                 & B-roads                                                                                                                  \\* 
\cline{3-3}
Tertiary                &                                                                                                                                                                 & \begin{tabular}[c]{@{}l@{}}Classified unnumbered roads \\OR unclassified busy through roads\end{tabular}                 \\* 
\cline{3-3}
Unclassified            &                                                                                                                                                                 & \multirow{2}{*}{\begin{tabular}[c]{@{}l@{}}Unclassified (Intended for \\local traffic - 60\% of UK roads)\end{tabular}}  \\* 
\cline{1-2}
Residential             & Function is purely residential                                                                                                                                  &                                                                                                                          \\ 
\hline
Service                 & \begin{tabular}[c]{@{}l@{}}Road that provides \\access to a facility\end{tabular}                                                                               &                                                                                                                          \\ 
\hline
Track                   & \begin{tabular}[c]{@{}l@{}}Unpaved - suitable for \\two track vehicles (mostly rural)\end{tabular}                                                              &                                                                                                                          \\ 
\hline
Cycleway                & \begin{tabular}[c]{@{}l@{}}Designated cycleway - \\open to cyclists only\end{tabular}                                                                           &                                                                                                                          \\ 
\hline
Path                    & \begin{tabular}[c]{@{}l@{}}Unpaved - open to \\non-motorized traffic only\end{tabular}                                                                          &                                                                                                                          \\
\hline
\end{longtable}

To conduct routing, the following is considered:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Cyclist Preference}: Work done by Dill and McNeil (2013) on examining
  cyclist typologies determined that around 60\% of Portland residents
  fit under the \emph{interested but concerned} category. These were people
  that enjoyed cycling but avoided it due safety concerns. The key to
  encouraging this group was to create a low-stress cycling network,
  not only though segregated infrastructure but also by planning
  routes that passed through residential streets.
\item
  \textbf{Low-Traffic Neighbourhoods (LTNs)}: The UK Department for Transport is
  allocating funding to local authorities to invest in Active
  Transport, partially through the creation of LTNs
  (DfT, 2020a). This includes closing off residential
  streets to motorized traffic.
\item
  \textbf{Existing Cycling Infrastructure}: Utilizing existing cycling
  infrastructure makes economic sense, as small investments may lead
  to large connectivity gains as the disconnected cycling
  infrastructure gets joined together.
\end{enumerate}

The above points are accounted for by using a weighted road network for
routing. This has been previously done by multiplying all road segments
without cycling infrastructure by a fixed impedance factor
(Mauttone et al., 2017), or by assigning different weights to the road
segments proportional to the investment cost of bringing them to an
acceptable level of stress for cycling (Duthie and Unnikrishnan, 2014).
Perceived stress for cyclists has been found to be a function of vehicular traffic
volume and speed (Sorton and Walsh, 1994), both of which vary predictably
with road type. Gehrke et al. (2020) build on this information to use
routing impedance factors for road segments that are based on road type
and the existence of cycling infrastructure.

For the purposes of this paper, we created a similar weighting profile
that is adjusted to favor less stressful roads (based on the definition
by Sorton and Walsh (1994) and information from Table
\ref{table:osmroadtypes}), and roads with existing cycling
infrastructure. We believe this to be more appropriate than the simplistic
approach adopted by Mauttone et al. (2017), as it makes use of a hierarchy
of road preference based on perceived stress levels, going beyond simply
favoring roads with existing cycling infrastructure.

\begin{longtable}{lrrr}
\caption{Weighting profiles}
\label{table:weightprofiles}\\
\multirow{2}{*}{\textbf{OSM Road Type}} & \multicolumn{3}{c}{\textbf{Weighting Profile}}                                                 \\*
                                        & \multicolumn{1}{l}{\textit{Unweighted}} & \multicolumn{1}{l}{\textit{Weighted}} & \multicolumn{1}{l}{\textit{Weighted\_2}}  \endfirsthead
Cycleway                                & 1          & 1                                     & 1                                         \\
Path                                    & 1          & 0.9                                   & 0.9                                       \\
Residential                             & 1          & 0.9                                   & 0.9                                       \\
Service                                 & 1          & 0.9                                   & 0.9                                       \\
Tertiary                                & 1          & 0.9                                   & 0.9                                       \\
Track                                   & 1          & 0.9                                   & 0.9                                       \\
Unclassified                            & 1          & 0.9                                   & 0.9                                       \\
Secondary                               & 1          & 0.8                                   & 0.8                                       \\
Primary                                 & 1          & 0.7                                   & \textcolor{red}{0}                        \\
Trunk                                   & 1          & 0.6                                   & \textcolor{red}{0}                        \\
Motorway                                & 1          & \textcolor{red}{0}                    & \textcolor{red}{0}                       
\end{longtable}

\noindent A weighted distance \(d_{w}\) for each road
segment is calculated as follows:\footnote{The \textbf{dodgr} r package
  (Padgham, 2019) is used to route cycling demand onto the road
  network. The package uses the OpenStreetMaps (OSM) road network and
  allows the user to assign weights to roads based on their type. The
  routing is done based on weighted shortest paths, with the distance
  along each road segment being divided by a factor to obtain the weighted
  distance for routing. It is more intuitive to multiply when weighting a
  network, but the dodgr package divides by numbers between 0 and 1, which
  achieves the same result (road types with a weight of 0 are not routed on).
  For the sake of reproducibility, we stick to
  the convention used in the package.}

\begin{equation}\label{eq:weight_distance}
    d_{w} = \frac{d_{unw}}{W}
\end{equation}

\noindent where \(d_{unw}\) is the unweighted distance and
\(W\) is the weight from Table \ref{table:weightprofiles}.

All weights are between 0 and 1, and the values in the
\textit{Weighted} profile are chosen to be inversely proportional to
the stress level experienced by cyclists on them. The
\textit{Unweighted} weighting profile is used to compare increases in
route length resulting from two different approaches:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Weighted}: Relatively high impedance on Primary and Trunk roads
  (to minimize cycling on them).
\item
  \textbf{Weighted\_2}: Avoiding Primary and Trunk Roads completely.
\end{enumerate}

\begin{figure}
\includegraphics[width=0.9\linewidth]{data/Manchester/Plots/flows_facet_unweighted_Manchester} \caption{Flow results based on unweighted shortest paths (Manchester)}\label{fig:flowsfacetunweighted}
\end{figure}

\begin{figure}
\includegraphics[width=0.9\linewidth]{data/Manchester/Plots/flows_facet_weighted_Manchester} \caption{Flow results based on weighted shortest paths (Manchester)}\label{fig:flowsfacetweighted}
\end{figure}

Comparing the cycling demand routed on the weighted and unweighted road
network allows us to get a better understanding of the importance of
different road types. In the case of Manchester, trunk roads bisect the
city and are a major part of unweighted shortest paths (Figure
\ref{fig:flowsfacetunweighted}). On the other hand, cycleways are not
part of unweighted shortest paths, and so very little of the cycling
demand is routed through them. In the weighted network, cycleways are
much better utilized, and the majority of the cycling demand passes
through tertiary roads, as expected.

The results of routing potential cycling demand on the weighted and
unweighted networks are understandably quite different. From Figure
\ref{fig:flowsfacetunweighted} we can see that trunk and primary roads
are the most efficient means of traversing the road network of
Manchester. Both of these road types are classified as Primary A roads
according to the UK Department for Transport's road classification
(Table \ref{table:osmroadtypes}), and are therefore part of the Primary
Route Network (PRN) (DfT, 2012). The PRN has the widest,
most direct roads on the network, and carries most of the through
traffic. This includes freight, with all roads in the PRN being required
by law to provide unrestricted access to trucks up to 40 tonnes
(DfT, 2012).

We choose to avoid routing the potential cycling demand on Primary A
Roads for the following 2 reasons:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Logistical Difficulty}: Changes on these roads need to be agreed
  upon by all affected authorities (DfT, 2012), which
  may prove to be difficult.
\item
  \textbf{Low Traffic Neighborhoods (LTNs)}: The UK government is aiming to
  restrict access to motorized vehicles on residential roads to create
  LTNs (DfT, 2020a). This is part of a policy to
  prevent automobile rat-running and make streets more accessible to
  cyclists and pedestrians. Under such a policy, Primary A roads would
  become even more essential for motorized traffic and it would be
  more difficult to reallocate road space on these roads to cyclists.
\end{enumerate}

Figure \ref{fig:flowsfacetweighted} shows that routing on the weighted
network significantly reduces flow on the trunk and primary roads, but
does not eliminate it completely. This is intentional, as the impedance
on these roads is only slightly higher than the remaining road types (See
Table \ref{table:weightprofiles}). Potential cycling demand is only
routed on these roads if there are no routes through other roads that
offer comparable directness.

Banning cycling flow completely on trunk and primary roads may result in
excessively circuitous paths, as seen in Figure
\ref{fig:boxplotcircuity}. When routing using the \emph{weighted} weighting
profile in Table \ref{table:weightprofiles}, we see that shortest paths
increase by less than 5\% on average from unweighted shortest paths, with
the largest increases still below 30\%. When routing on primary and trunk
roads is banned (\emph{weighted\_2} profile in Table
\ref{table:weightprofiles}), the average increase relative to
unweighted shortest paths rises to 10\%, with certain locations
experiencing more significant negative effects on accessibility. Given
that cyclists will only deviate from shortest paths by a certain amount
to access better cycling infrastructure (as explained in Section
\ref{introduction}), allowing flow on some stretches of trunk and
primary roads is necessary to insure cycling uptake and equitable access
to cycling infrastructure. In its new vision for walking and cycling,
the Department for Transport acknowledges that minimal segregated
stretches of bicycle lanes on main roads will be necessary to avoid
circuitous cycling networks (DfT, 2020a).

Weighting the road network also allows us to better utilize existing
cycling infrastructure, as can be seen by the higher flow on cycleways
in Figure \ref{fig:flowsfacetweighted}. Again, the small differences in
impedance between cycleways and other road types mean that cycleways
that require significant deviation are not routed on.

It should be reiterated that the weighting profile used for routing has
been developed for the purposes of this study. It creates a hierarchy of
road preference that is grounded in cyclist preferences and government
plans to create LTNs. A wide range of weighting profiles could be used
to represent different types of cyclists and road environments, as
described in Gehrke et al. (2020) and Furth et al. (2016). A sensitivity
analysis could be done to determine an optimal weighting profile, but
given the variation in city road
networks, this would probably
require calibration to the specific city. More accurate routing could be
carried out given the availability of road-level data. In such cases we
would add additional impedance to specific roads, giving more useful
routing results than the current approach which considers all roads
of the same type to be equivalent.

One use-case of such granular data would be to identify roads that serve
schools. The Department of Transport notes that the number of school
children being driven to school has trebled over the past 40 years
(DfT, 2020a), and so having cycling infrastructure
serving schools is key to achieving the government target of getting
more children to cycle. This would not be difficult, as over 75\% of
children in the UK live within a 15 minute cycle from their school
(DfT, 2020b). Goodman et al. (2019) show that if
dutch levels of cycling were achieved in the UK, the \% of children
cycling to school could increase from 1.8\% to 41\%. In their typology of
cyclists, Dill and McNeil (2013) found that a majority of people who say they
would never cycle had never cycled to school, whereas confident cyclists
were those most likely to have cycled to school. Getting people to cycle
from a young age is therefore key to achieving societal change in
commuting habits.

\begin{figure}

{\centering \includegraphics[width=0.5\linewidth]{data/Manchester/Plots/boxplot_weighted_unweighted_distances}

}

\caption{Effect of banning cyclists from trunk and primary roads for all OD pairs (Manchester)}\label{fig:boxplotcircuity}
\end{figure}

\hypertarget{road-segment-prioritization}{%
\section{Road Segment Prioritization}\label{road-segment-prioritization}}

After routing the potential cycling demand onto the road network using
weighted shortest paths, we have estimates for the cumulative potential
cycling demand passing through all road segments. This cumulative demand
(referred to as \emph{flow}) is then used as a basis for identifying segments
that are in most need of investment in segregated cycling
infrastructure. In doing so, we must account for the motivations and
deterrents for cycling identified in Section \ref{introduction}, namely
direct and well-connected routes.

A range of algorithms could be used for prioritization. Because policy
priorities vary, we present two algorithms. Both utilize existing
infrastructure from the beginning and allow us to compare a solution
that focuses on utilitarianism to one that focuses on egalitarianism. In
both algorithms, links are selected iteratively and the iteration at
which each link is added to the solution is recorded. Investments in
cycling infrastructure can be limited by budget constraints, so it can
be useful to see where best to allocate a defined length of segregated
infrastructure. In order to incorporate egalitarian principles in our
approach, we use community detection to partition the study area and
evaluate the distribution of investment over the different subdivisions.

\hypertarget{community-detection}{%
\subsection{Community Detection}\label{community-detection}}

As explained in Section
\ref{ethical-underpinnings-and-proposed-approach}, a major challenge
facing `top-down' planning approaches is how to incorporate egalitarian
principles by fairly distributing investments in cycling infrastructure.
One way of quantifying this is to split up the city into smaller
geospatial areas and target equal investment in each of those areas.
This approach could also help ensure that on-the-ground surveys are made
by local stakeholders, an important component of the planning process
(Parkin, 2018). Community detection offers us a way to
delineate such a split; cyclists are limited in their commuting distance
(see Figure \ref{fig:cyclinghistmanc}), and so trip attractors are more
likely to have a local catchment area of cyclists.

\begin{figure}

{\centering \includegraphics[width=0.4\linewidth]{data/Manchester/Plots/histogram_distance_cycling}

}

\caption{Cycling commuting distance - Manchester (2011 census data)}\label{fig:cyclinghistmanc}
\end{figure}

In our case, the network is the city; the nodes are the
population-weighted MSOA centroids and the links connecting each MSOA
pair are weighted by the potential cycling demand between them. The
Louvain method (Blondel et al., 2008) is used to separate MSOAs into
communities. Potential cycling demand is used since we assume that this
is what the cycling demand will be once the cycling infrastructure is
added. To assign road links to communities, the following steps are
carried out:

\begin{verbatim}
1. Create links between MSOA centroids and weigh these links by potential cycling demand 
   between them.
2. Use the Louvain method to determine the optimal number of communities and assign each MSOA 
   centroid to a community.
3. Assign each road link to the same community as the closest MSOA centroid to it.
\end{verbatim}

The results show that Manchester can be split into four large
communities and one small one (Figure \ref{fig:communitiesmanchester}).

\begin{figure}

{\centering \includegraphics[width=0.8\linewidth]{data/Manchester/Plots/communities_alternative_Manchester}

}

\caption{Communities based on potential cycling demand between MSOAs (Manchester)}\label{fig:communitiesmanchester}
\end{figure}

~

\hypertarget{algorithm-1-utilitarian-expansion}{%
\subsection{Algorithm 1: Utilitarian Expansion}\label{algorithm-1-utilitarian-expansion}}

The algorithm logic is as follows:

\begin{verbatim}
1. Identify all links that have segregated cycling infrastructure and add them to the 
   initial solution
2. Identify all links that neighbor links in the current solution
3. Select neighboring link with highest flow and add it to the solution
4. Repeat steps 2 and 3 until all flow is satisfied or investment threshold is met
\end{verbatim}

This algorithm ensures that the resulting network is connected. It also
satisfies the directness criteria, since links on the weighted shortest
paths are those that have the highest flow passing through them (this is
a result of the routing in Section \ref{routing}.

\hypertarget{algorithm-2-egalitarian-expansion}{%
\subsection{Algorithm 2: Egalitarian Expansion}\label{algorithm-2-egalitarian-expansion}}

The first algorithm focuses on connectivity and directness, but not on
fairly distributing investment. The latter is not a requirement for
increasing cycling uptake, but it is fundamental for spatial equity, as
explained in Section \ref{ethical-underpinnings-and-proposed-approach}. This algorithm incorporates
the ideal of fair distribution by ensuring that investment is distributed between the defined communities.
This is done using the following logic:

\begin{verbatim}
1. Identify all links that have segregated cycling infrastructure and add them to the initial
   solution
2. Identify all links that neighbor links in the current solution
3. Select from each community one neighboring link with highest flow and add it to the
   solution
4. If there are no more neighboring links in a community, select the link with the highest
   flow in that community, regardless of connectivity, and add it to the solution
5. Repeat steps 2, 3 and 4 until all flow is satisfied or investment threshold is met
\end{verbatim}

Even though we may end up with a more disconnected network, we will have
separate connected networks in each community. Given that communities
are defined by having more internal flow than external flow, this is a
satisfactory solution.

The results of the community detection are used to evaluate the
algorithms. This is done by looking at the \emph{person-km satisfied} as
cycling infrastructure is added. Person-km is a measure of the total km
cycled on a road segment, so it is the product of the number of
potential commuters cycling on that road segment (\(flow\)) and the length
of the segment in km (\(l\)). For each road segment, the person-km is
equal to \(flow * l\). In the case of Manchester, Table
\ref{tab:personkmtable} shows that almost half of the person-km is in
community 1, while only 0.5\% of total person-km on the network is in
community 5.

\begin{table}[!h]

\caption{\label{tab:personkmtable}Total person-km in different communities (Manchester)}
\centering
\begin{tabular}[t]{l|l|r}
\hline
Community & Person-Km (Total) & Person-Km (\%)\\
\hline
1 & 284,458 & 44.4\\
\hline
2 & 163,877 & 25.6\\
\hline
3 & 79,218 & 12.4\\
\hline
4 & 109,635 & 17.1\\
\hline
5 & 3,317 & 0.5\\
\hline
\end{tabular}
\end{table}

Looking at the person-km satisfied (Figure \ref{fig:growthtotal}), we
see that the incremental addition of cycling infrastructure is better
distributed between communities using Algorithm 2; equal distribution of
investment results in the gain in \% of person km satisfied in each
community being inversely correlated with the size of the community. In
addition, we find that the restrictions imposed by Algorithm 2 on the
network expansion do not seem to have a noticeable effect on the
city-wide \% of person-km satisfied. Comparing both algorithms, we can
see that Algorithm 1 provides only marginally quicker city-wide gains
than Algorithm 2.

\begin{figure}

{\centering \includegraphics[width=0.48\linewidth]{data/Manchester/Plots/Growth_Results/growth_utilitarian_satisfied_km_both_flow_column} \includegraphics[width=0.48\linewidth]{data/Manchester/Plots/Growth_Results/growth_egalitarian_satisfied_km_both_flow_column}

}

\caption{Comparing overall (dashed) and community level person-km satisfied (Manchester)}\label{fig:growthtotal}
\end{figure}

Figure \ref{fig:growth3MapandBar} gives us a geographic representation
of the results from Algorithm 2; it shows when each link was added to
the solution (first 100km, second 100km, etc). We can see that,
generally, road segments around cycling infrastructure are prioritized,
except for those neighboring cycling infrastructure on the very
periphery. The first 100km is also spatially distributed across the
city, with no apparent bias towards a particular area.

It is also important to understand how the different highway types
contribute to the proposed network. Figure \ref{fig:growth3MapandBar}
shows that most of the flow will be on residential and tertiary roads,
as expected from the weighting profile defined in Table
\ref{table:weightprofiles}.

\begin{figure}

{\centering \includegraphics[width=0.45\linewidth]{data/Manchester/Plots/Growth_Results/growth_egalitarian_priority_all_FLOW} \includegraphics[width=0.45\linewidth]{data/Manchester/Plots/Growth_Results/growth_egalitarian_investment_highways_flow}

}

\caption{Road segment priority (left), disaggregated by road type (right) - egalitarian growth}\label{fig:growth3MapandBar}
\end{figure}

\hypertarget{connectivity}{%
\subsection{Connectivity}\label{connectivity}}

Existing cycling infrastructure is made up of many disconnected
components. Both Algorithm 1 and 2 start with all existing segregated
cycling infrastructure and aim to create an efficient, connected
network. Figure \ref{fig:componentsandGCC} shows that both algorithms
gradually reduce the number of components as more infrastructure is
added, but Algorithm 2 is able to provide better connectivity with less
investment.

Consistent growth can also be seen for the size of the Largest Connected
Component in the proposed bicycle network (Figure
\ref{fig:componentsandGCC}). Here however, we find that there is little
difference between both Algorithms.

\begin{figure}[H]

{\centering \includegraphics[width=0.85\linewidth]{data/Manchester/Plots/Growth_Results/growth_util_egal_components_gcc_components_together_Manchester}

}

\caption{Network characteristics}\label{fig:componentsandGCC}
\end{figure}

Overall, the improved distribution of investment shown by Algorithm 2 does not seem to come at the cost
of poorer connectivity or less city-wide gains when compared to Algorithm 1.
The results therefore advocate for the incorporation of egalitarian principles in bicycle network planning.

\hypertarget{discussion-and-conclusions}{%
\section{Discussion and conclusions}\label{discussion-and-conclusions}}

This paper demonstrates an approach for prioritizing
investments in city-wide cycleway networks. The approach aims to respect
both the needs of the users and the ambitions of stakeholders working at
local or regional levels. The results, including detailed route network
maps based on current travel behaviour derived from OD data, can provide
an evidence-base for designing direct, connected,
and low-stress networks.

Given that the ``most essential activity entailed in the design of
cycle-friendly infrastructure is developing a cycle network''
(Parkin, 2018), we believe that the approach has great
potential to inform investment in cities such as Manchester where there
is political will to invest in cycling long-term. A benefit of the
approach is that it has relatively modest data requirements: only the
road network (from OSM), its topography (from satellite imagery), and
OD data (from the national census in this case) are
needed, opening up the possibility of deploying the approach in other
cities.\^{}{[}The results are easily reproducible for all UK cities, and can also be reproduced for cities elsewhere given the availability of commuter data. Instructions for reproducing the results are detailed in the \textbf{README} file in the R folder of this \href{https://github.com/Hussein-Mahfouz/cycle-networks}{Github repository}

The approach can identify not only where there is high
potential for cycling but also trade-offs between
stress levels (related to motor traffic) and directness.
The results can therefore be used as a basis for recommendations on
road space reallocation \emph{and} new infrastructure to unlock potential cycling demand.
The approach encourages consideration of a wider range of preferences
and needs than previous approaches that focus only on absolute potential.
Moreover, the inclusion of egalitarian principles in scenarios of change
encourages investment in cycling infrastructure to
increase the connectivity of existing cycling infrastructure
\emph{and} investment that addresses geographical and social inequalities.
This ability to address inequalities in network prioritization is particularly
important given research showing substantial \emph{inequalities}
around transport provision in general and cycling uptake and investment
in particular (Lucas et al., 2016; Vidal Tortosa et al., 2021).

The approach is not without limitations. The level of detail is only as
good as the granularity of the available data (in this case relatively
coarse zones). Iacono et al. (2010) note that such large travel zones
are not ideal for understanding route choice behaviour of cyclists and
pedestrians. They also give rise to an `ecological fallacy' whereby
average characteristics are assumed to apply to all residents of the
aggregated geographical area, suggesting a need for applying the methods
to more granular OD data (and for governments and other
data-collecting organisations to make OD data more
readily available).
The potential demand calculation is also based on cycling in the traditional sense, and does not consider the
effect of micro-mobility on reducing topology-related impedance to cycling. Given that we are proposing an approach which can accommodate any cycling uptake functions, this is an acceptable limitation.

The approach is also focused solely on the allocation of cycling
infrastructure, and does not consider the larger political and
regulatory environment that needs to exist to promote cycling; while
segregated, connected, and direct cycling infrastructure is key to
achieving high levels of cycling, research has shown that it cannot
exist in a vacuum. Wardman et al. (2007) developed a mode choice model
for the UK and their results showed that improved cycling infrastructure
on its own only had modest impacts on mode shift, and even the unlikely
scenario of all urban routes being serviced by segregated bike lanes was
forecast to increase cycling mode share by only 3\%. International
research shows that cities that invest in more comprehensive cycling
projects have a more significant increase in the number of cyclists as
well as the cycling mode share (Parkin, 2018; Pucher et al., 2010). These cities do not just focus on
infrastructure, but on general policies as well as restricting car use.
Evaluation of policies in Denmark and Germany and the Netherlands has
shown that their high cycling mode share is down to a broader set of
soft and hard policies. Hard policies include traffic calming, filtered permeability
interventions, cycling rights of way, bike parking, integration with the
public transport network, and making driving cars both expensive and
inconvenient, while soft policies include marketing and awareness campaigns (Gössling, 2013; Pucher and Buehler, 2008). While these policies are outside the scope of this research, it is important to recognize their key role in
bringing about an increase in levels of cycling.

Consideration of these limitations suggest future directions of research.
New datasets and continuously evolving computational (hardware and software)
capabilities should enable the data related issues to be overcome as open
datasets, and our ability to process them, improve.
Plugging-in alternative uptake models could help address the relatively narrow
definition of `potential' used in this paper, to consider broader social factors.
A promising area of future research in this direction could be
to integrate a wider range of modes, including e-bikes and e-scooters, into the
analysis. Currently scope for progress in this direction is
partially restricted by the lack of data on the
proliferation of these modes, raising the point that surveys could
replace the broad category of `Bicycle' with a range of small modes such as
`pedal cycles' (including tricycles and recumbent cycles), e-bikes
and other forms of micro-mobility.

Despite the rapid growth of these alternative modes, there is little doubt that
the humble bicycle is a key ingredient in the ongoing shift towards active transport.
Recognition of the importance of this shift for improved health and well-being
of citizens, and the wider challenge to decarbonise the global economy,
has grown with pressures on health and public transport systems during the pandemic.
The success of policies to accelerate cycling uptake depends on a range of
factors including, vitally, the design of the
cycling network for potential cyclists.
The approach presented in this paper provides a strong evidence base,
that considers both cycling potential and social equity,
for designing joined-up and cost-effective strategic cycle networks.

\clearpage

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}
\setlength{\parskip}{8pt}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-akbarzadeh2018designing}{}%
Akbarzadeh, M., Mohri, S.S., Yazdian, E., 2018. Designing bike networks using the concept of network clusters. Applied network science 3, 12.

\leavevmode\hypertarget{ref-aldred2019impacts}{}%
Aldred, R., Croft, J., Goodman, A., 2019. Impacts of an active travel intervention with a cycling focus in a suburban context: One-year findings from an evaluation of london's in-progress mini-hollands programme. Transportation research part A: policy and practice 123, 147--169.

\leavevmode\hypertarget{ref-bao2017planning}{}%
Bao, J., He, T., Ruan, S., Li, Y., Zheng, Y., 2017. Planning bike lanes based on sharing-bikes' trajectories, in: Proceedings of the 23rd ACM SIGKDD International Conference on Knowledge Discovery and Data Mining. pp. 1377--1386.

\leavevmode\hypertarget{ref-black1995spatial}{}%
Black, W.R., 1995. Spatial interaction modeling using artificial neural networks. Journal of Transport Geography 3, 159--166.

\leavevmode\hypertarget{ref-blondel2008fast}{}%
Blondel, V.D., Guillaume, J.-L., Lambiotte, R., Lefebvre, E., 2008. Fast unfolding of communities in large networks. Journal of statistical mechanics: theory and experiment 2008, P10008.

\leavevmode\hypertarget{ref-brand_climate_2020}{}%
Brand, C., Dons, E., Anaya-Boig, E., Avila-Palencia, I., Clark, A., de Nazelle, A., Gascon, M., Gaupp-Berghausen, M., Gerike, R., Gotschi, T., 2020. The climate change mitigation effects of active travel. Preprint: researchsquare.com.

\leavevmode\hypertarget{ref-broach2011bicycle}{}%
Broach, J., Gliebe, J., Dill, J., 2011. Bicycle route choice model developed using revealed preference GPS data, in: 90th Annual Meeting of the Transportation Research Board, Washington, DC.

\leavevmode\hypertarget{ref-caulfield2012determining}{}%
Caulfield, B., Brick, E., McCarthy, O.T., 2012. Determining bicycle infrastructure preferences--a case study of dublin. Transportation research part D: transport and environment 17, 413--417.

\leavevmode\hypertarget{ref-celis2017association}{}%
Celis-Morales, C.A., Lyall, D.M., Welsh, P., Anderson, J., Steell, L., Guo, Y., Maldonado, R., Mackay, D.F., Pell, J.P., Sattar, N., others, 2017. Association between active commuting and incident cardiovascular disease, cancer, and mortality: Prospective cohort study. bmj 357, j1456.

\leavevmode\hypertarget{ref-crane2017longitudinal}{}%
Crane, M., Rissel, C., Standen, C., Ellison, A., Ellison, R., Wen, L.M., Greaves, S., 2017. Longitudinal evaluation of travel and health outcomes in relation to new bicycle infrastructure, sydney, australia. Journal of Transport \& Health 6, 386--395.

\leavevmode\hypertarget{ref-departmentgearchange2020}{}%
DfT, 2020a. Gear change: A bold vision for cycling and walking.

\leavevmode\hypertarget{ref-departmentcycleinfradesign2020}{}%
DfT, 2020b. Cycling infrastructure design.

\leavevmode\hypertarget{ref-department2012guidance}{}%
DfT, 2012. Guidance on road classification and the primary route network.

\leavevmode\hypertarget{ref-dill2013four}{}%
Dill, J., McNeil, N., 2013. Four types of cyclists? Examination of typology for better understanding of bicycling behavior and potential. Transportation Research Record 2387, 129--138.

\leavevmode\hypertarget{ref-duthie2014optimization}{}%
Duthie, J., Unnikrishnan, A., 2014. Optimization framework for bicycle network design. Journal of Transportation engineering 140, 04014028.

\leavevmode\hypertarget{ref-furth2016network}{}%
Furth, P.G., Mekuria, M.C., Nixon, H., 2016. Network connectivity for low-stress bicycling. Transportation research record 2587, 41--49.

\leavevmode\hypertarget{ref-gehrke2020cycling}{}%
Gehrke, S.R., Akhavan, A., Furth, P.G., Wang, Q., Reardon, T.G., 2020. A cycling-focused accessibility tool to support regional bike network connectivity. Transportation research part D: transport and environment 85, 102388.

\leavevmode\hypertarget{ref-goodman2019scenarios}{}%
Goodman, A., Rojas, I.F., Woodcock, J., Aldred, R., Berkoff, N., Morgan, M., Abbas, A., Lovelace, R., 2019. Scenarios of cycling to school in england, and associated health and carbon impacts: Application of the {`propensity to cycle tool.'} Journal of Transport \& Health 12, 263--278.

\leavevmode\hypertarget{ref-goodman2014new}{}%
Goodman, A., Sahlqvist, S., Ogilvie, D., Consortium, iConnect, 2014. New walking and cycling routes and increased physical activity: One-and 2-year findings from the UK iConnect study. American journal of public health 104, e38--e46.

\leavevmode\hypertarget{ref-gossling2013urban}{}%
Gössling, S., 2013. Urban transport transitions: Copenhagen, city of cyclists. Journal of Transport Geography 33, 196--206.

\leavevmode\hypertarget{ref-iacono2010measuring}{}%
Iacono, M., Krizek, K.J., El-Geneidy, A., 2010. Measuring non-motorized accessibility: Issues, alternatives, and execution. Journal of Transport Geography 18, 133--140.

\leavevmode\hypertarget{ref-jafino2020transport}{}%
Jafino, B.A., Kwakkel, J., Verbraeck, A., 2020. Transport network criticality metrics: A comparative analysis and a guideline for selection. Transport Reviews 40, 241--264.

\leavevmode\hypertarget{ref-jarrett2012effect}{}%
Jarrett, J., Woodcock, J., Griffiths, U.K., Chalabi, Z., Edwards, P., Roberts, I., Haines, A., 2012. Effect of increasing active travel in urban england and wales on costs to the national health service. The Lancet 379, 2198--2205.

\leavevmode\hypertarget{ref-kohl2012pandemic}{}%
Kohl 3rd, H.W., Craig, C.L., Lambert, E.V., Inoue, S., Alkandari, J.R., Leetongin, G., Kahlmeier, S., Group, L.P.A.S.W., others, 2012. The pandemic of physical inactivity: Global action for public health. The lancet 380, 294--305.

\leavevmode\hypertarget{ref-lovelace2017propensity}{}%
Lovelace, R., Goodman, A., Aldred, R., Berkoff, N., Abbas, A., Woodcock, J., 2017. The propensity to cycle tool: An open source online system for sustainable transport planning. Journal of transport and land use 10, 505--528.

\leavevmode\hypertarget{ref-lucas2016method}{}%
Lucas, K., Van Wee, B., Maat, K., 2016. A method to evaluate equitable accessibility: Combining ethical theories and accessibility-based approaches. Transportation 43, 473--490.

\leavevmode\hypertarget{ref-marques2015infrastructure}{}%
Marqués, R., Hernández-Herrador, V., Calvo-Salazar, M., García-Cebrián, J., 2015. How infrastructure can promote cycling in cities: Lessons from seville. Research in Transportation Economics 53, 31--44.

\leavevmode\hypertarget{ref-martinez2013new}{}%
Martínez, L.M., Viegas, J.M., 2013. A new approach to modelling distance-decay functions for accessibility assessment in transport studies. Journal of Transport Geography 26, 87--96.

\leavevmode\hypertarget{ref-mauttone2017bicycle}{}%
Mauttone, A., Mercadante, G., Rabaza, M., Toledo, F., 2017. Bicycle network design: Model and solution algorithm. Transportation research procedia 27, 969--976.

\leavevmode\hypertarget{ref-mesbah2012bilevel}{}%
Mesbah, M., Thompson, R., Moridpour, S., 2012. Bilevel optimization approach to design of network of bike lanes. Transportation research record 2284, 21--28.

\leavevmode\hypertarget{ref-nahmias2017integrating}{}%
Nahmias-Biran, B., Martens, K., Shiftan, Y., 2017. Integrating equity in transportation project assessment: A philosophical exploration and its practical implications. Transport reviews 37, 192--210.

\leavevmode\hypertarget{ref-natera2019data}{}%
Natera, L., Battiston, F., Iñiguez, G., Szell, M., 2019. Data-driven strategies for optimal bicycle network growth. arXiv preprint arXiv:1907.07080.

\leavevmode\hypertarget{ref-nello2020environmental}{}%
Nello-Deakin, S., 2020. Environmental determinants of cycling: Not seeing the forest for the trees? Journal of transport geography 85, 102704.

\leavevmode\hypertarget{ref-olmos2020data}{}%
Olmos, L.E., Tadeo, M.S., Vlachogiannis, D., Alhasoun, F., Alegre, X.E., Ochoa, C., Targa, F., González, M.C., 2020. A data science framework for planning the growth of bicycle infrastructures. Transportation Research Part C: Emerging Technologies 115, 102640.

\leavevmode\hypertarget{ref-ofn2018population}{}%
ONS, 2018. Population estimates for the UK, england and wales, scotland and northern ireland: Mid-2017. Hampshire: Office for National Statistics.

\leavevmode\hypertarget{ref-ONS2011flowdata}{}%
ONS, 2011. 2011 census: Special workplace statistics (united kingdom).

\leavevmode\hypertarget{ref-padgham2019dodgr}{}%
Padgham, M., 2019. Dodgr: An r package for network flow aggregation. Transport Findings. Network Design Lab.

\leavevmode\hypertarget{ref-parkin2018designing}{}%
Parkin, J., 2018. Designing for cycle traffic. ICE Publishing. \url{https://doi.org/10.1680/dfct.63495}

\leavevmode\hypertarget{ref-patterson2020associations}{}%
Patterson, R., Panter, J., Vamos, E.P., Cummins, S., Millett, C., Laverty, A.A., 2020. Associations between commute mode and cardiovascular disease, cancer, and all-cause mortality, and cancer incidence, using linked census data over 25 years in england and wales: A cohort study. The Lancet Planetary Health 4, e186--e194.

\leavevmode\hypertarget{ref-pereira2017distributive}{}%
Pereira, R.H., Schwanen, T., Banister, D., 2017. Distributive justice and equity in transportation. Transport reviews 37, 170--191.

\leavevmode\hypertarget{ref-pucher2008making}{}%
Pucher, J., Buehler, R., 2008. Making cycling irresistible: Lessons from the netherlands, denmark and germany. Transport reviews 28, 495--528.

\leavevmode\hypertarget{ref-pucher2010infrastructure}{}%
Pucher, J., Dill, J., Handy, S., 2010. Infrastructure, programs, and policies to increase bicycling: An international review. Preventive medicine 50, S106--S125.

\leavevmode\hypertarget{ref-schoner2014missing}{}%
Schoner, J.E., Levinson, D.M., 2014. The missing link: Bicycle infrastructure networks and ridership in 74 US cities. Transportation 41, 1187--1204.

\leavevmode\hypertarget{ref-sorton1994bicycle}{}%
Sorton, A., Walsh, T., 1994. Bicycle stress level as a tool to evaluate urban and suburban bicycle compatibility. Transportation Research Record 17--17.

\leavevmode\hypertarget{ref-stinson2003commuter}{}%
Stinson, M.A., Bhat, C.R., 2003. Commuter bicyclist route choice: Analysis using a stated preference survey. Transportation Research Record 1828, 107--115.

\leavevmode\hypertarget{ref-agreement2015paris}{}%
UN, 2015. Paris agreement, in: Report of the Conference of the Parties to the United Nations Framework Convention on Climate Change (21st Session, 2015: Paris). Retrived December. HeinOnline, p. 2017.

\leavevmode\hypertarget{ref-vidaltortosa_infrastructure_2020}{}%
Vidal Tortosa, E., Lovelace, R., Heinen, E., Mann, R.P., 2021. Infrastructure is not enough: Interactions between the environment, socioeconomic disadvantage and cycling participation in {England}. Journal of Transport and Land Use.

\leavevmode\hypertarget{ref-wardman2007factors}{}%
Wardman, M., Tight, M., Page, M., 2007. Factors influencing the propensity to cycle to work. Transportation Research Part A: Policy and Practice 41, 339--350.

\leavevmode\hypertarget{ref-wilson1971family}{}%
Wilson, A.G., 1971. A family of spatial interaction models, and associated developments. Environment and Planning A 3, 1--32.

\leavevmode\hypertarget{ref-winters2011motivators}{}%
Winters, M., Davidson, G., Kao, D., Teschke, K., 2011. Motivators and deterrents of bicycling: Comparing influences on decisions to ride. Transportation 38, 153--168.

\leavevmode\hypertarget{ref-winters2010far}{}%
Winters, M., Teschke, K., Grant, M., Setton, E.M., Brauer, M., 2010. How far out of the way will we travel? Built environment influences on route selection for bicycle and car travel. Transportation Research Record 2190, 1--10.

\end{CSLReferences}

\hypertarget{appendix-appendix}{%
\appendix}

\clearpage

\hypertarget{appendix-reproducing-the-results-for-other-cities}{%
\section{Appendix: Reproducing the results for other cities}\label{appendix-reproducing-the-results-for-other-cities}}

To reproduce the results presented for other cities in the UK there are a few computing pre-requisites:

\begin{itemize}
\tightlist
\item
  A reasonably powerful computer, e.g.~with 8 GB RAM
\item
  The statistical programming language R installed (we recommend running it in the RStudio editor)
\item
  The source code of the paper repository, which can be download from this \href{https://github.com/Hussein-Mahfouz/cycle-networks}{Github repository}{]}
\end{itemize}

To ensure you have the necessary packages installed, open the \texttt{cycle-networks.Rproj} file in RStudio and run the following commands to ensure you have the necessary packages installed:

\begin{verbatim}
install.packages("renv")
renv::restore()
\end{verbatim}

The \texttt{build.R} script contains all of the necessary steps to reproduce the results for any city in the UK.
Open this file with the following command and run the code line-by-line.

\begin{verbatim}
file.edit("R/build.R")
\end{verbatim}

You can select a city to build by executing the following command:

\begin{verbatim}
chosen_city <- "Leeds"
\end{verbatim}

Replacing \texttt{Leeds} with any other city will make the script work for that city.
Note: this only works for Local Authority names in England and Wales listed below.

In this script, you choose which city you wish to run the analysis from
this list of available towns and cities:

\begin{verbatim}
##   [1] "Barnsley"             "Basildon"             "Basingstoke"
##   [4] "Bath"                 "Bedford"              "Birkenhead"
##   [7] "Birmingham"           "Blackburn"            "Blackpool"
##  [10] "Bolton"               "Bournemouth"          "Bracknell"
##  [13] "Bradford"             "Brighton and Hove"    "Bristol"
##  [16] "Burnley"              "Burton upon Trent"    "Bury"
##  [19] "Cambridge"            "Cardiff"              "Carlisle"
##  [22] "Chatham"              "Chelmsford"           "Cheltenham"
##  [25] "Chester"              "Chesterfield"         "Colchester"
##  [28] "Coventry"             "Crawley"              "Darlington"
##  [31] "Derby"                "Doncaster"            "Dudley"
##  [34] "Eastbourne"           "Exeter"               "Gateshead"
##  [37] "Gillingham"           "Gloucester"           "Grimsby"
##  [40] "Guildford"            "Halifax"              "Harlow"
##  [43] "Harrogate"            "Hartlepool"           "Hastings"
##  [46] "Hemel Hempstead"      "High Wycombe"         "Huddersfield"
##  [49] "Ipswich"              "Kingston upon Hull"   "Leeds"
##  [52] "Leicester"            "Lincoln"              "Liverpool"
##  [55] "London"               "Luton"                "Maidstone"
##  [58] "Manchester"           "Mansfield"            "Middlesbrough"
##  [61] "Milton Keynes"        "Newcastle upon Tyne"  "Newcastle-under-Lyme"
##  [64] "Newport"              "Northampton"          "Norwich"
##  [67] "Nottingham"           "Nuneaton"             "Oldham"
##  [70] "Oxford"               "Peterborough"         "Plymouth"
##  [73] "Poole"                "Portsmouth"           "Preston"
##  [76] "Reading"              "Redditch"             "Rochdale"
##  [79] "Rotherham"            "Salford"              "Scunthorpe"
##  [82] "Sheffield"            "Shrewsbury"           "Slough"
##  [85] "Solihull"             "South Shields"        "Southampton"
##  [88] "Southend-on-Sea"      "Southport"            "St Albans"
##  [91] "St Helens"            "Stevenage"            "Stockport"
##  [94] "Stockton-on-Tees"     "Stoke-on-Trent"       "Sunderland"
##  [97] "Sutton Coldfield"     "Swansea"              "Swindon"
## [100] "Telford"              "Wakefield"            "Walsall"
## [103] "Warrington"           "Watford"              "West Bromwich"
## [106] "Weston-Super-Mare"    "Wigan"                "Woking"
## [109] "Wolverhampton"        "Worcester"            "Worthing"
## [112] "York"
\end{verbatim}

\newpage

\hypertarget{birmingham}{%
\subsection{Birmingham}\label{birmingham}}

\subsubsection{Potential Demand}

\begin{figure}
\includegraphics[width=0.3\linewidth]{data/Birmingham/Plots/histogram_distance_all_vs_cycling} \includegraphics[width=0.3\linewidth]{data/Birmingham/Plots/histogram_distance_all_vs_cycling_potential} \includegraphics[width=0.3\linewidth]{data/Birmingham/Plots/histogram_distance_cycling_potential_vs_current} \caption{Distribution of Potential Cycling Demand}\label{fig:potdemhistogramsBirmingham}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=0.8\linewidth]{data/Birmingham/Plots/desire_facet_cycling}

}

\caption{Current and Potential Cycling Demand}\label{fig:desirefacetcyclingBirmingham}
\end{figure}

\clearpage

\subsubsection{Routing Cycling Flows}

\begin{figure}

{\centering \includegraphics[width=0.75\linewidth]{data/Birmingham/Plots/flows_facet_unweighted_Birmingham}

}

\caption{Flow Results Based on Unweighted Shortest Paths (Birmingham)}\label{fig:flowsfacetunweightedBirmingham}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=0.75\linewidth]{data/Birmingham/Plots/flows_facet_weighted_Birmingham}

}

\caption{Flow Results Based on Weighted Shortest Paths (Birmingham)}\label{fig:flowsfacetweightedBirmingham}
\end{figure}

\clearpage

\subsubsection{Community Detection}

\begin{figure}

{\centering \includegraphics[width=0.9\linewidth]{data/Birmingham/Plots/communities_alternative_Birmingham}

}

\caption{Communities Based on Potential Cycling Demand (Birmingham)}\label{fig:communitiesBirmingham}
\end{figure}

\subsubsection{Network Expansion Algorithms}

\begin{figure}[H]

{\centering \includegraphics[width=0.45\linewidth]{data/Birmingham/Plots/Growth_Results/growth_utilitarian_satisfied_km_both_flow_column} \includegraphics[width=0.45\linewidth]{data/Birmingham/Plots/Growth_Results/growth_egalitarian_satisfied_km_both_flow_column}

}

\caption{Comparing Overall (Dashed) and Community Level Person-Km Satisfied (Birmingham)}\label{fig:growthtotalBirmingham}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=0.45\linewidth]{data/Birmingham/Plots/Growth_Results/growth_util_egal_components_number_comparisonBirmingham} \includegraphics[width=0.45\linewidth]{data/Birmingham/Plots/Growth_Results/growth_util_egal_components_gcc_comparisonBirmingham}

}

\caption{Network Characteristics}\label{fig:componentsandGCCBirmingham}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=0.45\linewidth]{data/Birmingham/Plots/Growth_Results/growth_egalitarian_priority_all_FLOW}

}

\caption{Road Segment Priority (left), disaggregated by road type (right) - Egalitarian Growth}\label{fig:growth3MapandBarBirmingham}
\end{figure}

\clearpage

\hypertarget{leeds}{%
\subsection{Leeds}\label{leeds}}

\subsubsection{Potential Demand}

\begin{figure}
\includegraphics[width=0.3\linewidth]{data/Leeds/Plots/histogram_distance_all_vs_cycling} \includegraphics[width=0.3\linewidth]{data/Leeds/Plots/histogram_distance_all_vs_cycling_potential} \includegraphics[width=0.3\linewidth]{data/Leeds/Plots/histogram_distance_cycling_potential_vs_current} \caption{Distribution of Potential Cycling Demand}\label{fig:potdemhistogramsLeeds}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=0.8\linewidth]{data/Leeds/Plots/desire_facet_cycling}

}

\caption{Current and Potential Cycling Demand}\label{fig:desirefacetcyclingLeeds}
\end{figure}

\clearpage

\subsubsection{Routing Cycling Flows}

\begin{figure}

{\centering \includegraphics[width=0.75\linewidth]{data/Leeds/Plots/flows_facet_unweighted_Leeds}

}

\caption{Flow Results Based on Unweighted Shortest Paths (Leeds)}\label{fig:flowsfacetunweightedLeeds}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=0.75\linewidth]{data/Leeds/Plots/flows_facet_weighted_Leeds}

}

\caption{Flow Results Based on Weighted Shortest Paths (Leeds)}\label{fig:flowsfacetweightedLeeds}
\end{figure}

\clearpage

\subsubsection{Community Detection}

\begin{figure}

{\centering \includegraphics[width=0.9\linewidth]{data/Leeds/Plots/communities_alternative_Leeds}

}

\caption{Communities Based on Potential Cycling Demand (Leeds)}\label{fig:communitiesLeeds}
\end{figure}

\subsubsection{Network Expansion Algorithms}

\begin{figure}[H]

{\centering \includegraphics[width=0.45\linewidth]{data/Leeds/Plots/Growth_Results/growth_utilitarian_satisfied_km_both_flow_column} \includegraphics[width=0.45\linewidth]{data/Leeds/Plots/Growth_Results/growth_egalitarian_satisfied_km_both_flow_column}

}

\caption{Comparing Overall (Dashed) and Community Level Person-Km Satisfied (Leeds)}\label{fig:growthtotalLeeds}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=0.45\linewidth]{data/Leeds/Plots/Growth_Results/growth_util_egal_components_number_comparisonLeeds} \includegraphics[width=0.45\linewidth]{data/Leeds/Plots/Growth_Results/growth_util_egal_components_gcc_comparisonLeeds}

}

\caption{Network Characteristics}\label{fig:componentsandGCCLeeds}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=0.45\linewidth]{data/Leeds/Plots/Growth_Results/growth_egalitarian_priority_all_FLOW} \includegraphics[width=0.45\linewidth]{data/Leeds/Plots/Growth_Results/growth_egalitarian_investment_highways_flow}

}

\caption{Road Segment Priority (left), disaggregated by road type (right) - Egalitarian Growth}\label{fig:growth3MapandBarLeeds}
\end{figure}

\clearpage

\hypertarget{nottingham}{%
\subsection{Nottingham}\label{nottingham}}

\subsubsection{Potential Demand}

\begin{figure}
\includegraphics[width=0.3\linewidth]{data/Nottingham/Plots/histogram_distance_all_vs_cycling} \includegraphics[width=0.3\linewidth]{data/Nottingham/Plots/histogram_distance_all_vs_cycling_potential} \includegraphics[width=0.3\linewidth]{data/Nottingham/Plots/histogram_distance_cycling_potential_vs_current} \caption{Distribution of Potential Cycling Demand}\label{fig:potdemhistogramsNottingham}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=0.8\linewidth]{data/Nottingham/Plots/desire_facet_cycling}

}

\caption{Current and Potential Cycling Demand}\label{fig:desirefacetcyclingNottingham}
\end{figure}

\clearpage

\subsubsection{Routing Cycling Flows}

\begin{figure}

{\centering \includegraphics[width=0.75\linewidth]{data/Nottingham/Plots/flows_facet_unweighted_Nottingham}

}

\caption{Flow Results Based on Unweighted Shortest Paths (Nottingham)}\label{fig:flowsfacetunweightedNottingham}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=0.75\linewidth]{data/Nottingham/Plots/flows_facet_weighted_Nottingham}

}

\caption{Flow Results Based on Weighted Shortest Paths (Nottingham)}\label{fig:flowsfacetweightedNottingham}
\end{figure}

\clearpage

\subsubsection{Community Detection}

\begin{figure}

{\centering \includegraphics[width=0.9\linewidth]{data/Nottingham/Plots/communities_alternative_Nottingham}

}

\caption{Communities Based on Potential Cycling Demand (Nottingham)}\label{fig:communitiesNottingham}
\end{figure}

\subsubsection{Network Expansion Algorithms}

\begin{figure}[H]
\includegraphics[width=0.45\linewidth]{data/Nottingham/Plots/Growth_Results/growth_utilitarian_satisfied_km_both_flow_column} \includegraphics[width=0.45\linewidth]{data/Nottingham/Plots/Growth_Results/growth_egalitarian_satisfied_km_both_flow_column} \caption{Comparing Overall (Dashed) and Community Level Person-Km Satisfied (Nottingham)}\label{fig:growthtotalNottingham}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=0.45\linewidth]{data/Nottingham/Plots/Growth_Results/growth_util_egal_components_number_comparisonNottingham} \includegraphics[width=0.45\linewidth]{data/Nottingham/Plots/Growth_Results/growth_util_egal_components_gcc_comparisonNottingham}

}

\caption{Network Characteristics}\label{fig:componentsandGCCNottingham}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=0.45\linewidth]{data/Nottingham/Plots/Growth_Results/growth_egalitarian_priority_all_FLOW} \includegraphics[width=0.45\linewidth]{data/Nottingham/Plots/Growth_Results/growth_egalitarian_investment_highways_flow}

}

\caption{Road Segment Priority (left), disaggregated by road type (right) - Egalitarian Growth}\label{fig:growth3MapandBarNottingham}
\end{figure}

\clearpage

\end{document}

### 2022-08-29 README.Rmd
intro: manchester context

@@
 title: A Road Segment Prioritization Approach for Cycling Infrastructure
-#date: "`r Sys.Date()`"
+author:
+  - Hussein Mahfouz:
+      email: h.mahfouz.19@ucl.ac.uk
+      institute: casa
+      correspondence: true
+  - name: Elsa Arcaute
+    institute: casa
+  - name: Robin Lovelace
+    institute: its
+institute:
+  - casa: Centre for Advanced Spatial Analysis (CASA), University College London, London, UK
+  - its: Institute for Transport Studies, University of Leeds, Leeds, UK
+
 geometry: margin=2cm
@@
 commuting across the country; it must be used to design adequate cycling
-infrastructure that is based on motivators and deterrents to cycling.
+infrastructure that is based on motivators and deterrents to cycling. In this paper we 
+propose an approach for planning a connected cycling network that is meant to 
+be applicable to any city, and we use the city of Manchester as a case study. Greater 
+Manchester aims to become the first region to reach a carbon neutral transport network; 
+Its local authorities have developed plans for a 1800 mile walking and cycling network 
+(Bee Network) that will be delivered over 10 years [@manchestercyclingbee2020]. While schemes 
+have already been put in place for around a quarter of this network, the majority is yet to be built.
+
+# Literature Review
 
@@
 
-## Network-Level Approaches
+## Network-Level Approaches to Bicycle Infrastructure Planning
 
@@
 the network through the graph-theoretic concept of *connected
-components*. @natera2019data for example,
+components*. @orozco2020data for example,
 study the existing cycling network in terms of its disconnected
@@
 
-```{r potdemhistograms, fig.cap="Distribution of potential cycling demand", out.width="100%", fig.show='hold', fig.align="center", echo=FALSE}
+```{r potdemhistograms, fig.cap="Distribution of potential cycling demand", out.width="90%", fig.show='hold', fig.align="center", echo=FALSE}
 knitr::include_graphics("data/Manchester/Plots/histogram_distance_cycling_potential_current_all.png")

### 2022-08-31 response to reviewers.Rmd
highlighting how egalitarian principles were included

---
title: "Response to Reviewers"
author: "Hussein"
date: "8/29/2022"
output:
  pdf_document:
    #extra_dependencies: ["subcaption", "float"]
    #keep_tex: true
---

# Overall Comments

# Reviewer 1

*This paper presents an approach to help planners prioritize investments in cycling infrastructure.*

-   *Now that I have read the full article, I would consider this a methods paper, as it presents and describes a proposed method. This wasn't clear from the outset, and I found it difficult to follow without a clear methods and results section (there are methods and results presented, but I am not sure where they begin and end - they seem to be combined).*

We agree that this is a methods paper, and in fact focus on the **approach**​ (which could be implemented using different methods, we outline one) as highlighted below. We have rewritten relevant parts of the paper to emphasise the methodological advances outlined in the paper. Furthermore, we have demonstrated applying the methods to additional locations in the appendix. The abstract reads as follows (emphasis added to highlight the framing as a methods paper):

"This paper demonstrates a new **approach** to support the prioritization of cycling infrastructure and cycling network design, accounting for cyclist preferences and the growing emphasis on 'filtered permeability' and 'Low Traffic Neighborhood' interventions internationally. The **approach** combines distance decay, route calculation, and network analysis methods to examine where future cycling demand is most likely to arise, how such demand could be accommodated within existing street networks, and how to ensure a fair distribution of investment. ***Although each of these methods has been applied to cycling infrastructure prioritization in previous research, this is the first time that they have been combined, creating an integrated road segment prioritization approach***."

-   *I see the merit of this paper, especially to policymakers, but I do wonder if the Journal of Transport Geography is the best fit for the paper. The contribution to geography is not especially clear. Perhaps Transport Policy, or some other policy-focused journal (or even a methods-focused journal) would be a better fit?*

We think that this paper fits the remit of Transport Geography well (based on the aims and scope of the journal) as it presents a new methodological approach in planning cycling networks that is grounded in both GIS and literature on the impact of infrastructure on cycling uptake.

-   *The inclusion of egalitarian principles in the paper was intriguing, but it was difficult to parse out exactly how this was done, how this approach differs from past work, and the implications/ benefits of using this approach.*

[How this approach differs from past work:]{.ul}

end of section 2.2: "while these approaches deal with continuity better, they look at the network as a whole when attempting to improve it, and in doing so fail to account for equitable distribution of infrastructure

section 2.3: "This research attempts to propose an egalitarian framework for cycling network design. This is done by identifying the different sub-networks that exist within the larger network, and ensuring that each gets a fair share of investment"

[How this was done/ benefits (bold is new text):]{.ul}

"Our work builds on these community finding approaches by proposing a similar greedy network expansion algorithm for cycle network expansion within communities. We incorporate community finding methods for study area partitioning with weighted routing to avoid links that are stressful to cycle on. **In doing so, we propose an approach that ensures a better distribution investments across the study area** while also accounting for motivators and deterrents to cycling."

a major challenge facing 'top-down' planning approaches is how to incorporate egalitarian principles by fairly distributing investments in cycling infrastructure. One way of quantifying this is to split up the city into smaller geospatial areas and target equal investment in each of those areas.....Community detection offers us a way to delineate such a split; cyclists are limited in their commuting distance (see Figure 6), and so trip attractors are more likely to have a local catchment area of cyclists. In our case, the network is the city; the nodes are the population-weighted MSOA centroids and the links connecting each MSOA pair are weighted by the potential cycling demand between them. The Louvain method (Blondel et al., 2008) is used to separate MSOAs into communities. Potential cycling demand is used since we assume that this is what the cycling demand will be once the cycling infrastructure is added"

"the inclusion of egalitarian principles in scenarios of change encourages investment in cycling infrastructure to increase the connectivity of existing cycling infrastructure and investment that addresses geographical and social inequalities. This ability to address inequalities in network prioritization is particularly important given research showing substantial inequalities around transport provision in general and cycling uptake and investment in particular (Lucas et al., 2016; Vidal Tortosa et al., 2021

-   The big picture take aways from the work were also not clear...

# Reviewer 2

This paper presents an interesting approach to network design for cycling infrastructure. I believe the proposed approach is really interesting and valid.

-   *However, the paper still has some flaws which should be solved. I suggest the authors to revise the structure of the paper: in particular, readers may not be familiar with the Manchester case. Instead it is taken for granted by the authors, who introduce the applied methodology in section 2 without ever having presented the case study and its peculiarities; even some results (e.g. those relating to the calculation of the demand potential) are presented in section 2, before having concluded the description of the methodology. Authors should choose from one of the following structures: 1) presentation of the different steps of the methodology (theoretically) and subsequent application to the case study. 2) Presentation of the case study and step-by-step application of the methodology, however including a clear schematization of it and any procedures to ensure its scalability.*

We have chosen option 2 as we think it is better to show the methodology and results for each sub-secion together. Some context was added on Manchester and its recent cycling and walking targets (specifically the Change a Region to Change a Nation report [[link](<https://beeactive.tfgm.com/publications-and-downloads/>)].

We think that the methodology is generic enough to be scalable. The readme in the github page (linked in the paper) has step by step instructions for reproducibility, and the appendix includes examples fr two other UK cities.

-   *I would also suggest that the authors to separate the literature review (currently in section 1.2) from the introduction, in order to better highlight what research gaps the article tries to solve.*

This has been done now

-   *Finally, a careful proofreading should be done, since the article currently has several typos.*

This has been done as well

### 2022-08-31 README.Rmd
highlighting how egalitarian principles were included

@@
 
-Our work builds on these *community finding* approaches by proposing a
+Our work builds on these community finding approaches by proposing a
 similar greedy network expansion algorithm for cycle network expansion
-within communities. We incorporate community finding methods for study
+*within communities*. We incorporate community finding methods for study
 area partitioning with weighted routing to avoid links that are
-stressful to cycle on. In doing so, we propose an approach that
-accounts for motivators and deterrents to cycling. We propose three
-sub-methods that address some of the limitations of previous
+stressful to cycle on. In doing so, we propose an approach that ensures a better distribution 
+investments across the study area while also accounting for motivators and deterrents to cycling. 
+We propose three sub-methods that address some of the limitations of previous
 studies. These limitations include (a) an inherent bias of basing network
@@
 
-```{r personkmtable, echo=FALSE, message = FALSE}
+```{r personkmtable, echo=FALSE, message = FALSE, warning=FALSE}
 
@@
 \@ref(fig:componentsandGCC)). Here however, we find that there is little
-difference between both Algorithms.
+difference between both algorithms.
 
@@
 cities^[The results are easily reproducible for all UK cities, and can also be reproduced for cities elsewhere given the availability of commuter data.
-Instructions for reproducing the results are detailed in the **README** file in the R folder of this [Github repository](https://github.com/Hussein-Mahfouz/cycle-networks)]
+Instructions for reproducing the results are detailed in the **README** file in the R folder of this [Github repository](https://github.com/Hussein-Mahfouz/cycle-networks)].
 

### 2022-08-31 README.Rmd
remove reference to appendix in abstract

@@
 Although each of these methods has been applied to cycling infrastructure prioritization in previous research, this is the first time that they have been combined, creating an integrated road segment prioritization approach.
-The approach, which can be applied to other cities, as shown in the Appendix, is demonstrated in a case study of Manchester, resulting in cycling networks that balance directness against the need for safe and stress-free routes under different investment scenarios.
+The approach, which can be applied to other cities, is demonstrated in a case study of Manchester, resulting in cycling networks that balance directness against the need for safe and stress-free routes under different investment scenarios.
 A key benefit of the approach from a policy perspective is its ability to support egalitarian and cost-effective strategic cycle network planning.

### 2022-08-31 README.Rmd
proofreading

@@
 propose an approach for planning a connected cycling network that is meant to 
-be applicable to any city, and we use the city of Manchester as a case study. Greater 
+be applicable to any city (as shown in the Appendix), and we use the city of Manchester as a case study. Greater 
 Manchester aims to become the first region to reach a carbon neutral transport network; 
-Its local authorities have developed plans for a 1800 mile walking and cycling network 
+its local authorities have developed plans for a 1800 mile walking and cycling network 
 (Bee Network) that will be delivered over 10 years [@manchestercyclingbee2020]. While schemes 
@@
 limitation of these approaches is that they do not explicitly solve for
-continuity. The latter is addressed using either a constraint specifying
+continuity. Continuity is addressed using either a constraint specifying
 that each link with a bike lane should be connected to at least one
@@
 the network through the graph-theoretic concept of *connected
-components*. @orozco2020data for example,
-study the existing cycling network in terms of its disconnected
+components*. @orozco2020data study the existing cycling network in terms of its disconnected
 components.
@@
 use a modularity maximization approach [@blondel2008fast] on taxi trip
-data to identify 7 different communities in the city of Isfahan, Iran.
+data to identify seven different communities in the city of Isfahan, Iran.
 An optimization problem is then formulated to connect nodes within each
@@
 We propose three sub-methods that address some of the limitations of previous
-studies. These limitations include (a) an inherent bias of basing network
+studies. These limitations include (a) an inherent bias when basing network
 design solely on existing cycling demand, (b) the proposal of routes that may
@@
 *OD* data can be obtained from a range of sources,
-including GPS data, household travels surveys and Census data on work
+including GPS data, household travels surveys and census data on work
 locations. In areas where observed OD data is unavailable, modelling
@@
 \noindent where $d$ and $s$ are the distance and slope associated with each
-OD pair, and $\alpha$ and all $\beta$s, are parameters calculated by a regression
+OD pair, and $\alpha$ and all $\beta$s are parameters calculated by a regression
 model on the training data. The square and square-root distance terms
 "capture the non-linear impact of distance on the likelihood of
-cycling", and interaction terms to capture the combined effect of slope
+cycling", and interaction terms capture the combined effect of slope
 and distance [@lovelace2017propensity]. Alternative cycling uptake
@@
 
-For the purposes of this paper, we created a similar weighting profile
+For the purposes of this paper we created a similar weighting profile
 that is adjusted to favor less stressful roads (based on the definition
 by @sorton1994bicycle and information from Table
-\@ref(table:osmroadtypes)), and roads with existing cycling
+\@ref(table:osmroadtypes)) and roads with existing cycling
 infrastructure. We believe this to be more appropriate than the simplistic
-approach adopted by @mauttone2017bicycle, as it makes use of a hierarchy
+approach adopted by @mauttone2017bicycle as it makes use of a hierarchy
 of road preference based on perceived stress levels, going beyond simply
-favoring roads with existing cycling infrastructure.
+favouring roads with existing cycling infrastructure.
 
@@
 approach, we use community detection to partition the study area and
-evaluate the distribution of investment over the different subdivisions.
+distribute investment over the different subdivisions.
 
@@
 Component in the proposed bicycle network (Figure
-\@ref(fig:componentsandGCC)). Here however, we find that there is little
+\@ref(fig:componentsandGCC)). Here, however, we find that there is little
 difference between both algorithms.

### 2022-09-01 README.Rmd
minor conclusion edits to emphasize big picture takeaways

@@
 an evidence-base for designing direct, connected,
-and low-stress networks.
+and low-stress networks. A core contribution of this approach is that it insures 
+that investments in cycling infrastructure are distributed fairly between different 
+communities in a city.
 
@@
 Moreover, the inclusion of egalitarian principles in scenarios of change
-encourages investment in cycling infrastructure to
-increase the connectivity of existing cycling infrastructure
-*and* investment that addresses geographical and social inequalities.
+encourages investment in cycling infrastructure that addresses geographical and 
+social inequalities. This is done by explicitly ensuring that infrastructure improvements 
+are pursued in different parts of the city simultaneously. This approach compares well with one 
+that focuses exclusively on maximizing the number of potential cycling trips that 
+benefit from investment, showing that a fair distribution can be achieved without 
+compromising on overall benefit. 
 This ability to address inequalities in network prioritization is particularly

### 2022-09-03 paper-original-submission.Rmd
Add original submission Rmd and latexdiff code

--- 2021-05-07 README.Rmd
@@
 title: A Road Segment Prioritization Approach for Cycling Infrastructure
-#date: "`r Sys.Date()`"
+author:
+  - Hussein Mahfouz:
+      email: h.mahfouz.19@ucl.ac.uk
+      institute: casa
+      correspondence: true
+  - name: Elsa Arcaute
+    institute: casa
+  - name: Robin Lovelace
+    institute: its
+institute:
+  - casa: Centre for Advanced Spatial Analysis (CASA), University College London, London, UK
+  - its: Institute of Transport Studies, University of Leeds, Leeds, UK
+
 geometry: margin=2cm

### 2022-09-03 latexdiff.R

@@ -1,0 +1,8 @@
+# Aim: generate diff for peer review
+
+
+download.file(
+  "https://github.com/Hussein-Mahfouz/cycle-networks/raw/ce4827fc708d510bead10d4998ce7ba875286d0d/README.Rmd",
+  "paper-original-submission.Rmd"
+  )
+rmarkdown::render("paper-original-submission.Rmd")

### 2022-09-03 R/_5.0_identifying_cycle_infastructure_from_osm_tags.R
Improve reproducibility, fix #13

@@ -232,7 +232,9 @@ graph_sf_cycle <-
 graph_sf <- dplyr::left_join(graph_sf, graph_sf_cycle, by = "edge_id")
 # all NA values in cycle_infra are those that had nothing to join to. It means they have no 
 # cycling infrastructure. We will give them a value of 0 for cycle_infra
-graph_sf$cycle_infra[is.na(graph_sf$cycle_infra)] <- 0
+if(any(is.na(graph_sf$cycle_infra))) {
+  graph_sf$cycle_infra[is.na(graph_sf$cycle_infra)] <- 0
+}
 
 # save it as an RDS
 saveRDS(graph_sf, file = paste0("data/", chosen_city, "/graph_with_flows_weighted.Rds"))

### 2022-09-03 R/build.R

@@ -1,6 +1,7 @@
 # Aim: reproduce results for other UK cities
 
-chosen_city <- "Bradford"
+chosen_city <- "Manchester"
+# chosen_city <- "Bradford" # For a different city
 
 source("R/_1.0_get_flow_data.R")
 
@@ -19,7 +20,6 @@ source("R/_3.0_potential_demand.R")
 source("R/_3.1_plot_mode_shares.R")
 source("R/_3.2_plot_od_comparison.R")
 source("R/_3.3_plot_desire_lines_current_vs_potential.R")
-
 
 source("R/_4.0_aggregating_flows.R")
 

### 2022-09-03 R/_5.0_identifying_cycle_infastructure_from_osm_tags.R
Additional fix #13

@@ -269,7 +269,9 @@ infra_exists <- function(graph, network){
   
   # all NA values in cycle_infra are those that had nothing to join to. It means they have no 
   # cycling infrastructure. We will give them a value of 0 for cycle_infra
-  graph$cycle_infra[is.na(graph$cycle_infra)] <- 0
+  if(any(is.na(graph$cycle_infra))) {
+    graph$cycle_infra[is.na(graph$cycle_infra)] <- 0
+  }  
   
   return(graph)
 }

### 2022-09-03 R/build.R

@@ -27,6 +27,7 @@ source("R/_4.0_aggregating_flows.R")
 # file.exists("data/Leeds/graph_with_flows_unweighted.Rds")
 
 # Warning: takes several minutes to run
+file.edit("R/_5.0_identifying_cycle_infastructure_from_osm_tags.R")
 source("R/_5.0_identifying_cycle_infastructure_from_osm_tags.R")
 
 # file.edit("R/_6.0_comparing_weighting_profiles.R")

### 2022-09-03 README.Rmd
Fully reproduce results

@@
 output:
- # bookdown::github_document2:
+ bookdown::github_document2:
    bookdown::pdf_document2:

### 2022-09-03 R/_8.1_growth_utilitarian.R

@@ -10,6 +10,13 @@ dir.create(paste0("data/", chosen_city, "/Plots/Growth_Results"), showWarnings =
 
 graph_sf <- readRDS(paste0("data/", chosen_city, "/graph_with_flows_weighted_communities.Rds"))
 
+# TODO: prevent this from happening upstream: https://github.com/Hussein-Mahfouz/cycle-networks/issues/13
+if("cycle_infra.x" %in% names(graph_sf)) {
+  graph_sf = graph_sf %>% 
+    rename(cycle_infra = cycle_infra.x) %>% 
+    select(-cycle_infra.y)
+}
+
 # we weigh the flow on each edge by its distance. We can then get how much of the commuter km are satisfied
 graph_sf$person_km <- graph_sf$flow * graph_sf$d
 

### 2022-09-03 R/build.R

@@ -27,7 +27,7 @@ source("R/_4.0_aggregating_flows.R")
 # file.exists("data/Leeds/graph_with_flows_unweighted.Rds")
 
 # Warning: takes several minutes to run
-file.edit("R/_5.0_identifying_cycle_infastructure_from_osm_tags.R")
+# file.edit("R/_5.0_identifying_cycle_infastructure_from_osm_tags.R")
 source("R/_5.0_identifying_cycle_infastructure_from_osm_tags.R")
 
 # file.edit("R/_6.0_comparing_weighting_profiles.R")
@@ -40,4 +40,5 @@ source("R/_8.0_growing_a_network.R")
 source("R/_8.1_growth_utilitarian.R")
 
 source("R/_8.2_growth_egalitarian.R")
+
 

### 2022-09-03 README.Rmd
Add pdf and tex versions

@@
 output:
- bookdown::github_document2:
+ # bookdown::github_document2:
    bookdown::pdf_document2:
@@
       # https://stackoverflow.com/questions/52918716/authors-and-affiliations-in-the-yaml-of-rmarkdown
-      pandoc_args:
-      - '--lua-filter=extras/scholarly-metadata.lua'
-      - '--lua-filter=extras/author-info-blocks.lua'
+      # pandoc_args:
+      # - '--lua-filter=extras/scholarly-metadata.lua'
+      # - '--lua-filter=extras/author-info-blocks.lua'
       keep_tex:  true

### 2022-09-03 README.tex
Add pdf and tex versions

--- 2021-05-08 arxiv.tex
@@
 ]{article}
+\usepackage{amsmath,amssymb}
 \usepackage{lmodern}
-\usepackage{amsmath}
-\usepackage{ifxetex,ifluatex}
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+\usepackage{iftex}
+\ifPDFTeX
   \usepackage[T1]{fontenc}
@@
   \usepackage{textcomp} % provide euro and other symbols
-  \usepackage{amssymb}
 \else % if luatex or xetex
@@
 \usepackage{xcolor}
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
-\hypersetup{
-  pdftitle={A Road Segment Prioritization Approach for Cycling Infrastructure},
-  pdfauthor={Hussein Mahfouz1,; Elsa Arcaute1; Robin Lovelace2},
-  hidelinks,
-  pdfcreator={LaTeX via pandoc}}
-\urlstyle{same} % disable monospaced font for URLs
 \usepackage[margin=2cm]{geometry}
-\usepackage{longtable,booktabs}
+\usepackage{longtable,booktabs,array}
 \usepackage{calc} % for calculating minipage widths
@@
 \setcounter{secnumdepth}{5}
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newlength{\cslentryspacingunit} % times entry-spacing
+\setlength{\cslentryspacingunit}{\parskip}
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1
+  \let\oldpar\par
+  \def\par{\hangindent=\cslhangindent\oldpar}
+  \fi
+  % set entry spacing
+  \setlength{\parskip}{#2\cslentryspacingunit}
+ }%
+ {}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 \usepackage{subcaption}
@@
 \usepackage{makecell}
-\ifluatex
+\usepackage{xcolor}
+\ifLuaTeX
   \usepackage{selnolig}  % disable illegal ligatures
 \fi
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
-  % turn on hanging indent if param 1 is 1
-  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
-  % set entry spacing
-  \ifnum #2 > 0
-  \setlength{\parskip}{#2\baselineskip}
-  \fi
- }%
- {}
-\usepackage{calc}
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\urlstyle{same} % disable monospaced font for URLs
+\hypersetup{
+  pdftitle={A Road Segment Prioritization Approach for Cycling Infrastructure},
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
 
 \title{A Road Segment Prioritization Approach for Cycling Infrastructure}
-\author{Hussein Mahfouz\textsuperscript{1,*} \and Elsa Arcaute\textsuperscript{1} \and Robin Lovelace\textsuperscript{2}}
+\author{true \and true \and true}
 \date{}
@@
 \end{abstract}
-
-\textsuperscript{1} The Centre for Advanced Spatial Analysis, University College London, London, UK\\
-\textsuperscript{2} Institute of Transport Studies, University of Leeds, Leeds, UK
-
-\textsuperscript{*} Correspondence: \href{mailto:h.mahfouz.19@ucl.ac.uk}{Hussein Mahfouz \textless{}\href{mailto:h.mahfouz.19@ucl.ac.uk}{\nolinkurl{h.mahfouz.19@ucl.ac.uk}}\textgreater{}}
 
@@
 commuting across the country; it must be used to design adequate cycling
-infrastructure that is based on motivators and deterrents to cycling.
+infrastructure that is based on motivators and deterrents to cycling. In this paper we
+propose an approach for planning a connected cycling network that is meant to
+be applicable to any city (as shown in the Appendix), and we use the city of Manchester as a case study. Greater
+Manchester aims to become the first region to reach a carbon neutral transport network;
+its local authorities have developed plans for a 1800 mile walking and cycling network
+(Bee Network) that will be delivered over 10 years (Greater Manchester, 2020). While schemes
+have already been put in place for around a quarter of this network, the majority is yet to be built.
+
+\hypertarget{literature-review}{%
+\section{Literature Review}\label{literature-review}}
 
@@
 Another defining feature for cycling infrastructure is how well
-connected it is; cyclists prefer cohesive infrastructure, particularly
+connected it is. Cyclists prefer cohesive infrastructure, particularly
 when cycling on arterial roads with high levels of motorized traffic
@@
 
-\hypertarget{network-level-approaches}{%
-\subsection{Network-Level Approaches}\label{network-level-approaches}}
+\hypertarget{network-level-approaches-to-bicycle-infrastructure-planning}{%
+\subsection{Network-Level Approaches to Bicycle Infrastructure Planning}\label{network-level-approaches-to-bicycle-infrastructure-planning}}
 
@@
 the network through the graph-theoretic concept of \emph{connected
-components}. Natera et al. (2019) for example,
-study the existing cycling network in terms of its disconnected
+components}. Orozco et al. (2020) study the existing cycling network in terms of its disconnected
 components.
@@
 use a modularity maximization approach (Blondel et al., 2008) on taxi trip
-data to identify 7 different communities in the city of Isfahan, Iran.
+data to identify seven different communities in the city of Isfahan, Iran.
 An optimization problem is then formulated to connect nodes within each
@@
 
-Our work builds on these \emph{community finding} approaches by proposing a
+Our work builds on these community finding approaches by proposing a
 similar greedy network expansion algorithm for cycle network expansion
-within communities. We incorporate community finding methods for study
+\emph{within communities}. We incorporate community finding methods for study
 area partitioning with weighted routing to avoid links that are
-stressful to cycle on. In doing so, we propose an approach that
-accounts for motivators and deterrents to cycling. We propose three
-sub-methods that address some of the limitations of previous
-studies. These limitations include (a) an inherent bias of basing network
+stressful to cycle on. In doing so, we propose an approach that ensures a better distribution
+investments across the study area while also accounting for motivators and deterrents to cycling.
+We propose three sub-methods that address some of the limitations of previous
+studies. These limitations include (a) an inherent bias when basing network
 design solely on existing cycling demand, (b) the proposal of routes that may
@@
 demand. It then introduces the network expansion algorithms, and compares an approach grounded in
-`egalitarianism' to one grounded in `utilitarianism.'
+`egalitarianism' to one grounded in `utilitarianism'.
 
@@
 \emph{OD} data can be obtained from a range of sources,
-including GPS data, household travels surveys and Census data on work
+including GPS data, household travels surveys and census data on work
 locations. In areas where observed OD data is unavailable, modelling
@@
 ``capture the non-linear impact of distance on the likelihood of
-cycling,'' and interaction terms to capture the combined effect of slope
+cycling'', and interaction terms capture the combined effect of slope
 and distance (Lovelace et al., 2017). Alternative cycling uptake
@@
 
-{\centering \includegraphics[width=1\linewidth]{data/Manchester/Plots/histogram_distance_cycling_potential_current_all}
+{\centering \includegraphics[width=0.9\linewidth]{data/Manchester/Plots/histogram_distance_cycling_potential_current_all}
 
@@
 
-For the purposes of this paper, we created a similar weighting profile
+For the purposes of this paper we created a similar weighting profile
 that is adjusted to favor less stressful roads (based on the definition
 by Sorton and Walsh (1994) and information from Table
-\ref{table:osmroadtypes}), and roads with existing cycling
+\ref{table:osmroadtypes}) and roads with existing cycling
 infrastructure. We believe this to be more appropriate than the simplistic
-approach adopted by Mauttone et al. (2017), as it makes use of a hierarchy
+approach adopted by Mauttone et al. (2017) as it makes use of a hierarchy
 of road preference based on perceived stress levels, going beyond simply
-favoring roads with existing cycling infrastructure.
+favouring roads with existing cycling infrastructure.
 
@@
 \noindent A weighted distance \(d_{w}\) for each road
-segment is calculated as follows:\footnote{The \textbf{dodgr} r package
+segment is calculated as follows \footnote{The \textbf{dodgr} r package
   (Padgham, 2019) is used to route cycling demand onto the road
@@
   For the sake of reproducibility, we stick to
-  the convention used in the package.}
+  the convention used in the package.}:
 
@@
 
-\begin{figure}
-\includegraphics[width=0.9\linewidth]{data/Manchester/Plots/flows_facet_unweighted_Manchester} \caption{Flow results based on unweighted shortest paths (Manchester)}\label{fig:flowsfacetunweighted}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=0.9\linewidth]{data/Manchester/Plots/flows_facet_weighted_Manchester} \caption{Flow results based on weighted shortest paths (Manchester)}\label{fig:flowsfacetweighted}
-\end{figure}
-
 Comparing the cycling demand routed on the weighted and unweighted road
@@
 through tertiary roads, as expected.
+
+\begin{figure}
+\includegraphics[width=0.9\linewidth]{data/Manchester/Plots/flows_facet_unweighted_Manchester} \caption{Flow results based on unweighted shortest paths (Manchester)}\label{fig:flowsfacetunweighted}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=0.9\linewidth]{data/Manchester/Plots/flows_facet_weighted_Manchester} \caption{Flow results based on weighted shortest paths (Manchester)}\label{fig:flowsfacetweighted}
+\end{figure}
 
@@
 
+\begin{figure}
+
+{\centering \includegraphics[width=0.5\linewidth]{data/Manchester/Plots/boxplot_weighted_unweighted_distances}
+
+}
+
+\caption{Effect of banning cyclists from trunk and primary roads for all OD pairs (Manchester)}\label{fig:boxplotcircuity}
+\end{figure}
+
 Weighting the road network also allows us to better utilize existing
@@
 
-\begin{figure}
-
-{\centering \includegraphics[width=0.5\linewidth]{data/Manchester/Plots/boxplot_weighted_unweighted_distances}
-
-}
-
-\caption{Effect of banning cyclists from trunk and primary roads for all OD pairs (Manchester)}\label{fig:boxplotcircuity}
-\end{figure}
-
 \hypertarget{road-segment-prioritization}{%
@@
 \hline
-1 & 284,458 & 44.4\\
-\hline
-2 & 163,877 & 25.6\\
-\hline
-3 & 79,218 & 12.4\\
-\hline
-4 & 109,635 & 17.1\\
-\hline
-5 & 3,317 & 0.5\\
+1 & 168,895 & 26.4\\
+\hline
+2 & 273,027 & 42.6\\
+\hline
+3 & 87,509 & 13.7\\
+\hline
+4 & 108,293 & 16.9\\
+\hline
+5 & 2,782 & 0.4\\
 \hline
@@
 Component in the proposed bicycle network (Figure
-\ref{fig:componentsandGCC}). Here however, we find that there is little
-difference between both Algorithms.
+\ref{fig:componentsandGCC}). Here, however, we find that there is little
+difference between both algorithms.
 
@@
 an evidence-base for designing direct, connected,
-and low-stress networks.
+and low-stress networks. A core contribution of this approach is that it insures
+that investments in cycling infrastructure are distributed fairly between different
+communities in a city.
 
@@
 needed, opening up the possibility of deploying the approach in other
-cities.\^{}{[}The results are easily reproducible for all UK cities, and can also be reproduced for cities elsewhere given the availability of commuter data.
-Instructions for reproducing the results are detailed in the \textbf{README} file in the R folder of this \href{https://github.com/Hussein-Mahfouz/cycle-networks}{Github repository}
+cities\footnote{The results are easily reproducible for all UK cities, and can also be reproduced for cities elsewhere given the availability of commuter data.
+Instructions for reproducing the results are detailed in the \textbf{README} file in the R folder of this \href{https://github.com/Hussein-Mahfouz/cycle-networks}{Github repository}}.
 
@@
 Moreover, the inclusion of egalitarian principles in scenarios of change
-encourages investment in cycling infrastructure to
-increase the connectivity of existing cycling infrastructure
-\emph{and} investment that addresses geographical and social inequalities.
+encourages investment in cycling infrastructure that addresses geographical and
+social inequalities. This is done by explicitly ensuring that infrastructure improvements
+are pursued in different parts of the city simultaneously. This approach compares well with one
+that focuses exclusively on maximizing the number of potential cycling trips that
+benefit from investment, showing that a fair distribution can be achieved without
+compromising on overall benefit.
 This ability to address inequalities in network prioritization is particularly
@@
 
-\clearpage
-
 \hypertarget{references}{%
@@
 \begin{CSLReferences}{1}{0}
-\leavevmode\hypertarget{ref-akbarzadeh2018designing}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-akbarzadeh2018designing}{}}%
 Akbarzadeh, M., Mohri, S.S., Yazdian, E., 2018. Designing bike networks using the concept of network clusters. Applied network science 3, 12.
 
-\leavevmode\hypertarget{ref-aldred2019impacts}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-aldred2019impacts}{}}%
 Aldred, R., Croft, J., Goodman, A., 2019.
@@
 
-\leavevmode\hypertarget{ref-bao2017planning}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-bao2017planning}{}}%
 Bao, J., He, T., Ruan, S., Li, Y., Zheng, Y., 2017.
@@
 
-\leavevmode\hypertarget{ref-black1995spatial}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-black1995spatial}{}}%
 Black, W.R., 1995. Spatial interaction modeling using artificial neural networks. Journal of Transport Geography 3, 159--166.
 
-\leavevmode\hypertarget{ref-blondel2008fast}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-blondel2008fast}{}}%
 Blondel, V.D., Guillaume, J.-L., Lambiotte, R., Lefebvre, E., 2008. Fast unfolding of communities in large networks. Journal of statistical mechanics: theory and experiment 2008, P10008.
 
-\leavevmode\hypertarget{ref-brand_climate_2020}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-brand_climate_2020}{}}%
 Brand, C., Dons, E., Anaya-Boig, E., Avila-Palencia, I., Clark, A., de Nazelle, A., Gascon, M., Gaupp-Berghausen, M., Gerike, R., Gotschi, T., 2020.
@@
 
-\leavevmode\hypertarget{ref-broach2011bicycle}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-broach2011bicycle}{}}%
 Broach, J., Gliebe, J., Dill, J., 2011. Bicycle route choice model developed using revealed preference GPS data, in: 90th Annual Meeting of the Transportation Research Board, Washington, DC.
 
-\leavevmode\hypertarget{ref-caulfield2012determining}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-caulfield2012determining}{}}%
 Caulfield, B., Brick, E., McCarthy, O.T., 2012. Determining bicycle infrastructure preferences--a case study of dublin. Transportation research part D: transport and environment 17, 413--417.
 
-\leavevmode\hypertarget{ref-celis2017association}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-celis2017association}{}}%
 Celis-Morales, C.A., Lyall, D.M., Welsh, P., Anderson, J., Steell, L., Guo, Y., Maldonado, R., Mackay, D.F., Pell, J.P., Sattar, N., others, 2017.
@@
 
-\leavevmode\hypertarget{ref-crane2017longitudinal}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-crane2017longitudinal}{}}%
 Crane, M., Rissel, C., Standen, C., Ellison, A., Ellison, R., Wen, L.M., Greaves, S., 2017.
@@
 
-\leavevmode\hypertarget{ref-departmentgearchange2020}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-departmentgearchange2020}{}}%
 DfT, 2020a. Gear change: A bold vision for cycling and walking.
 
-\leavevmode\hypertarget{ref-departmentcycleinfradesign2020}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-departmentcycleinfradesign2020}{}}%
 DfT, 2020b. Cycling infrastructure design.
 
-\leavevmode\hypertarget{ref-department2012guidance}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-department2012guidance}{}}%
 DfT, 2012. Guidance on road classification and the primary route network.
 
-\leavevmode\hypertarget{ref-dill2013four}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-dill2013four}{}}%
 Dill, J., McNeil, N., 2013. Four types of cyclists? Examination of typology for better understanding of bicycling behavior and potential. Transportation Research Record 2387, 129--138.
 
-\leavevmode\hypertarget{ref-duthie2014optimization}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-duthie2014optimization}{}}%
 Duthie, J., Unnikrishnan, A., 2014. Optimization framework for bicycle network design. Journal of Transportation engineering 140, 04014028.
 
-\leavevmode\hypertarget{ref-furth2016network}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-furth2016network}{}}%
 Furth, P.G., Mekuria, M.C., Nixon, H., 2016. Network connectivity for low-stress bicycling. Transportation research record 2587, 41--49.
 
-\leavevmode\hypertarget{ref-gehrke2020cycling}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-gehrke2020cycling}{}}%
 Gehrke, S.R., Akhavan, A., Furth, P.G., Wang, Q., Reardon, T.G., 2020.
@@
 
-\leavevmode\hypertarget{ref-goodman2019scenarios}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-goodman2019scenarios}{}}%
 Goodman, A., Rojas, I.F., Woodcock, J., Aldred, R., Berkoff, N., Morgan, M., Abbas, A., Lovelace, R., 2019.
-Scenarios of cycling to school in england, and associated health and carbon impacts: Application of the {`propensity to cycle tool.'} Journal of Transport \& Health 12, 263--278.
-
-\leavevmode\hypertarget{ref-goodman2014new}{}%
+Scenarios of cycling to school in england, and associated health and carbon impacts: Application of the {``propensity to cycle tool.''} Journal of Transport \& Health 12, 263--278.
+
+\leavevmode\vadjust pre{\hypertarget{ref-goodman2014new}{}}%
 Goodman, A., Sahlqvist, S., Ogilvie, D., Consortium, iConnect, 2014.
@@
 
-\leavevmode\hypertarget{ref-gossling2013urban}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-gossling2013urban}{}}%
 Gössling, S., 2013. Urban transport transitions: Copenhagen, city of cyclists. Journal of Transport Geography 33, 196--206.
 
-\leavevmode\hypertarget{ref-iacono2010measuring}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-manchestercyclingbee2020}{}}%
+Greater Manchester, T. for, 2020. Change a region to change a nation - greater manchester's walking and cycling investment plan.
+
+\leavevmode\vadjust pre{\hypertarget{ref-iacono2010measuring}{}}%
 Iacono, M., Krizek, K.J., El-Geneidy, A., 2010. Measuring non-motorized accessibility: Issues, alternatives, and execution. Journal of Transport Geography 18, 133--140.
 
-\leavevmode\hypertarget{ref-jafino2020transport}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-jafino2020transport}{}}%
 Jafino, B.A., Kwakkel, J., Verbraeck, A., 2020. Transport network criticality metrics: A comparative analysis and a guideline for selection. Transport Reviews 40, 241--264.
 
-\leavevmode\hypertarget{ref-jarrett2012effect}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-jarrett2012effect}{}}%
 Jarrett, J., Woodcock, J., Griffiths, U.K., Chalabi, Z., Edwards, P., Roberts, I., Haines, A., 2012.
@@
 
-\leavevmode\hypertarget{ref-kohl2012pandemic}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-kohl2012pandemic}{}}%
 Kohl 3rd, H.W., Craig, C.L., Lambert, E.V., Inoue, S., Alkandari, J.R., Leetongin, G., Kahlmeier, S., Group, L.P.A.S.W., others, 2012.
@@
 
-\leavevmode\hypertarget{ref-lovelace2017propensity}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-lovelace2017propensity}{}}%
 Lovelace, R., Goodman, A., Aldred, R., Berkoff, N., Abbas, A., Woodcock, J., 2017.
@@
 
-\leavevmode\hypertarget{ref-lucas2016method}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-lucas2016method}{}}%
 Lucas, K., Van Wee, B., Maat, K., 2016. A method to evaluate equitable accessibility: Combining ethical theories and accessibility-based approaches. Transportation 43, 473--490.
 
-\leavevmode\hypertarget{ref-marques2015infrastructure}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-marques2015infrastructure}{}}%
 Marqués, R., Hernández-Herrador, V., Calvo-Salazar, M., García-Cebrián, J., 2015.
@@
 
-\leavevmode\hypertarget{ref-martinez2013new}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-martinez2013new}{}}%
 Martínez, L.M., Viegas, J.M., 2013. A new approach to modelling distance-decay functions for accessibility assessment in transport studies. Journal of Transport Geography 26, 87--96.
 
-\leavevmode\hypertarget{ref-mauttone2017bicycle}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-mauttone2017bicycle}{}}%
 Mauttone, A., Mercadante, G., Rabaza, M., Toledo, F., 2017. Bicycle network design: Model and solution algorithm. Transportation research procedia 27, 969--976.
 
-\leavevmode\hypertarget{ref-mesbah2012bilevel}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-mesbah2012bilevel}{}}%
 Mesbah, M., Thompson, R., Moridpour, S., 2012. Bilevel optimization approach to design of network of bike lanes. Transportation research record 2284, 21--28.
 
-\leavevmode\hypertarget{ref-nahmias2017integrating}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-nahmias2017integrating}{}}%
 Nahmias-Biran, B., Martens, K., Shiftan, Y., 2017. Integrating equity in transportation project assessment: A philosophical exploration and its practical implications. Transport reviews 37, 192--210.
 
-\leavevmode\hypertarget{ref-natera2019data}{}%
-Natera, L., Battiston, F., Iñiguez, G., Szell, M., 2019. Data-driven strategies for optimal bicycle network growth. arXiv preprint arXiv:1907.07080.
-
-\leavevmode\hypertarget{ref-nello2020environmental}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-nello2020environmental}{}}%
 Nello-Deakin, S., 2020. Environmental determinants of cycling: Not seeing the forest for the trees? Journal of transport geography 85, 102704.
 
-\leavevmode\hypertarget{ref-olmos2020data}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-olmos2020data}{}}%
 Olmos, L.E., Tadeo, M.S., Vlachogiannis, D., Alhasoun, F., Alegre, X.E., Ochoa, C., Targa, F., González, M.C., 2020.
@@
 
-\leavevmode\hypertarget{ref-ofn2018population}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-ofn2018population}{}}%
 ONS, 2018. Population estimates for the UK, england and wales, scotland and northern ireland: Mid-2017. Hampshire: Office for National Statistics.
 
-\leavevmode\hypertarget{ref-ONS2011flowdata}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-ONS2011flowdata}{}}%
 ONS, 2011. 2011 census: Special workplace statistics (united kingdom).
 
-\leavevmode\hypertarget{ref-padgham2019dodgr}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-orozco2020data}{}}%
+Orozco, L.G.N., Battiston, F., Iñiguez, G., Szell, M., 2020. Data-driven strategies for optimal bicycle network growth. Royal Society open science 7.
+
+\leavevmode\vadjust pre{\hypertarget{ref-padgham2019dodgr}{}}%
 Padgham, M., 2019. Dodgr: An r package for network flow aggregation. Transport Findings. Network Design Lab.
 
-\leavevmode\hypertarget{ref-parkin2018designing}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-parkin2018designing}{}}%
 Parkin, J., 2018. Designing for cycle traffic. ICE Publishing. \url{https://doi.org/10.1680/dfct.63495}
 
-\leavevmode\hypertarget{ref-patterson2020associations}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-patterson2020associations}{}}%
 Patterson, R., Panter, J., Vamos, E.P., Cummins, S., Millett, C., Laverty, A.A., 2020.
@@
 
-\leavevmode\hypertarget{ref-pereira2017distributive}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-pereira2017distributive}{}}%
 Pereira, R.H., Schwanen, T., Banister, D., 2017. Distributive justice and equity in transportation. Transport reviews 37, 170--191.
 
-\leavevmode\hypertarget{ref-pucher2008making}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-pucher2008making}{}}%
 Pucher, J., Buehler, R., 2008. Making cycling irresistible: Lessons from the netherlands, denmark and germany. Transport reviews 28, 495--528.
 
-\leavevmode\hypertarget{ref-pucher2010infrastructure}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-pucher2010infrastructure}{}}%
 Pucher, J., Dill, J., Handy, S., 2010. Infrastructure, programs, and policies to increase bicycling: An international review. Preventive medicine 50, S106--S125.
 
-\leavevmode\hypertarget{ref-schoner2014missing}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-schoner2014missing}{}}%
 Schoner, J.E., Levinson, D.M., 2014. The missing link: Bicycle infrastructure networks and ridership in 74 US cities. Transportation 41, 1187--1204.
 
-\leavevmode\hypertarget{ref-sorton1994bicycle}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-sorton1994bicycle}{}}%
 Sorton, A., Walsh, T., 1994. Bicycle stress level as a tool to evaluate urban and suburban bicycle compatibility. Transportation Research Record 17--17.
 
-\leavevmode\hypertarget{ref-stinson2003commuter}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-stinson2003commuter}{}}%
 Stinson, M.A., Bhat, C.R., 2003. Commuter bicyclist route choice: Analysis using a stated preference survey. Transportation Research Record 1828, 107--115.
 
-\leavevmode\hypertarget{ref-agreement2015paris}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-agreement2015paris}{}}%
 UN, 2015. Paris agreement, in: Report of the Conference of the Parties to the United Nations Framework Convention on Climate Change (21st Session, 2015: Paris). Retrived December. HeinOnline, p. 2017.
 
-\leavevmode\hypertarget{ref-vidaltortosa_infrastructure_2020}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-vidaltortosa_infrastructure_2020}{}}%
 Vidal Tortosa, E., Lovelace, R., Heinen, E., Mann, R.P., 2021.
@@
 
-\leavevmode\hypertarget{ref-wardman2007factors}{}%
-Wardman, M., Tight, M., Page, M., 2007. Factors influencing the propensity to cycle to work. Transportation Research Part A: Policy and Practice 41, 339--350.
-
-\leavevmode\hypertarget{ref-wilson1971family}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-wardman2007factors}{}}%
+Wardman, M., Tight, M., Page, M., 2007.
+\href{https://doi.org/10.1016/j.tra.2006.09.011}{Factors influencing the propensity to cycle to work}.
+Transportation Research Part A: Policy and Practice 41, 339--350.
+
+\leavevmode\vadjust pre{\hypertarget{ref-wilson1971family}{}}%
 Wilson, A.G., 1971. A family of spatial interaction models, and associated developments. Environment and Planning A 3, 1--32.
 
-\leavevmode\hypertarget{ref-winters2011motivators}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-winters2011motivators}{}}%
 Winters, M., Davidson, G., Kao, D., Teschke, K., 2011. Motivators and deterrents of bicycling: Comparing influences on decisions to ride. Transportation 38, 153--168.
 
-\leavevmode\hypertarget{ref-winters2010far}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-winters2010far}{}}%
 Winters, M., Teschke, K., Grant, M., Setton, E.M., Brauer, M., 2010.
@@
 
-\hypertarget{appendix-appendix}{%
-\appendix}
-
-\clearpage
-
-\hypertarget{appendix-reproducing-the-results-for-other-cities}{%
-\section{Appendix: Reproducing the results for other cities}\label{appendix-reproducing-the-results-for-other-cities}}
-
-To reproduce the results presented for other cities in the UK there are a few computing pre-requisites:
-
-\begin{itemize}
-\tightlist
-\item
-  A reasonably powerful computer, e.g.~with 8 GB RAM
-\item
-  The statistical programming language R installed (we recommend running it in the RStudio editor)
-\item
-  The source code of the paper repository, which can be download from this \href{https://github.com/Hussein-Mahfouz/cycle-networks}{Github repository}{]}
-\end{itemize}
-
-To ensure you have the necessary packages installed, open the \texttt{cycle-networks.Rproj} file in RStudio and run the following commands to ensure you have the necessary packages installed:
-
-\begin{verbatim}
-install.packages("renv")
-renv::restore()
-\end{verbatim}
-
-The \texttt{build.R} script contains all of the necessary steps to reproduce the results for any city in the UK.
-Open this file with the following command and run the code line-by-line.
-
-\begin{verbatim}
-file.edit("R/build.R")
-\end{verbatim}
-
-You can select a city to build by executing the following command:
-
-\begin{verbatim}
-chosen_city <- "Leeds"
-\end{verbatim}
-
-Replacing \texttt{Leeds} with any other city will make the script work for that city.
-Note: this only works for Local Authority names in England and Wales listed below.
-
-In this script, you choose which city you wish to run the analysis from
-this list of available towns and cities:
-
-\begin{verbatim}
-##   [1] "Barnsley"             "Basildon"             "Basingstoke"
-##   [4] "Bath"                 "Bedford"              "Birkenhead"
-##   [7] "Birmingham"           "Blackburn"            "Blackpool"
-##  [10] "Bolton"               "Bournemouth"          "Bracknell"
-##  [13] "Bradford"             "Brighton and Hove"    "Bristol"
-##  [16] "Burnley"              "Burton upon Trent"    "Bury"
-##  [19] "Cambridge"            "Cardiff"              "Carlisle"
-##  [22] "Chatham"              "Chelmsford"           "Cheltenham"
-##  [25] "Chester"              "Chesterfield"         "Colchester"
-##  [28] "Coventry"             "Crawley"              "Darlington"
-##  [31] "Derby"                "Doncaster"            "Dudley"
-##  [34] "Eastbourne"           "Exeter"               "Gateshead"
-##  [37] "Gillingham"           "Gloucester"           "Grimsby"
-##  [40] "Guildford"            "Halifax"              "Harlow"
-##  [43] "Harrogate"            "Hartlepool"           "Hastings"
-##  [46] "Hemel Hempstead"      "High Wycombe"         "Huddersfield"
-##  [49] "Ipswich"              "Kingston upon Hull"   "Leeds"
-##  [52] "Leicester"            "Lincoln"              "Liverpool"
-##  [55] "London"               "Luton"                "Maidstone"
-##  [58] "Manchester"           "Mansfield"            "Middlesbrough"
-##  [61] "Milton Keynes"        "Newcastle upon Tyne"  "Newcastle-under-Lyme"
-##  [64] "Newport"              "Northampton"          "Norwich"
-##  [67] "Nottingham"           "Nuneaton"             "Oldham"
-##  [70] "Oxford"               "Peterborough"         "Plymouth"
-##  [73] "Poole"                "Portsmouth"           "Preston"
-##  [76] "Reading"              "Redditch"             "Rochdale"
-##  [79] "Rotherham"            "Salford"              "Scunthorpe"
-##  [82] "Sheffield"            "Shrewsbury"           "Slough"
-##  [85] "Solihull"             "South Shields"        "Southampton"
-##  [88] "Southend-on-Sea"      "Southport"            "St Albans"
-##  [91] "St Helens"            "Stevenage"            "Stockport"
-##  [94] "Stockton-on-Tees"     "Stoke-on-Trent"       "Sunderland"
-##  [97] "Sutton Coldfield"     "Swansea"              "Swindon"
-## [100] "Telford"              "Wakefield"            "Walsall"
-## [103] "Warrington"           "Watford"              "West Bromwich"
-## [106] "Weston-Super-Mare"    "Wigan"                "Woking"
-## [109] "Wolverhampton"        "Worcester"            "Worthing"
-## [112] "York"
-\end{verbatim}
-
-\newpage
-
-\hypertarget{birmingham}{%
-\subsection{Birmingham}\label{birmingham}}
-
-\subsubsection{Potential Demand}
-
-\begin{figure}
-\includegraphics[width=0.3\linewidth]{data/Birmingham/Plots/histogram_distance_all_vs_cycling} \includegraphics[width=0.3\linewidth]{data/Birmingham/Plots/histogram_distance_all_vs_cycling_potential} \includegraphics[width=0.3\linewidth]{data/Birmingham/Plots/histogram_distance_cycling_potential_vs_current} \caption{Distribution of Potential Cycling Demand}\label{fig:potdemhistogramsBirmingham}
-\end{figure}
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.8\linewidth]{data/Birmingham/Plots/desire_facet_cycling}
-
-}
-
-\caption{Current and Potential Cycling Demand}\label{fig:desirefacetcyclingBirmingham}
-\end{figure}
-
-\clearpage
-
-\subsubsection{Routing Cycling Flows}
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.75\linewidth]{data/Birmingham/Plots/flows_facet_unweighted_Birmingham}
-
-}
-
-\caption{Flow Results Based on Unweighted Shortest Paths (Birmingham)}\label{fig:flowsfacetunweightedBirmingham}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.75\linewidth]{data/Birmingham/Plots/flows_facet_weighted_Birmingham}
-
-}
-
-\caption{Flow Results Based on Weighted Shortest Paths (Birmingham)}\label{fig:flowsfacetweightedBirmingham}
-\end{figure}
-
-\clearpage
-
-\subsubsection{Community Detection}
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.9\linewidth]{data/Birmingham/Plots/communities_alternative_Birmingham}
-
-}
-
-\caption{Communities Based on Potential Cycling Demand (Birmingham)}\label{fig:communitiesBirmingham}
-\end{figure}
-
-\subsubsection{Network Expansion Algorithms}
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.45\linewidth]{data/Birmingham/Plots/Growth_Results/growth_utilitarian_satisfied_km_both_flow_column} \includegraphics[width=0.45\linewidth]{data/Birmingham/Plots/Growth_Results/growth_egalitarian_satisfied_km_both_flow_column}
-
-}
-
-\caption{Comparing Overall (Dashed) and Community Level Person-Km Satisfied (Birmingham)}\label{fig:growthtotalBirmingham}
-\end{figure}
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.45\linewidth]{data/Birmingham/Plots/Growth_Results/growth_util_egal_components_number_comparisonBirmingham} \includegraphics[width=0.45\linewidth]{data/Birmingham/Plots/Growth_Results/growth_util_egal_components_gcc_comparisonBirmingham}
-
-}
-
-\caption{Network Characteristics}\label{fig:componentsandGCCBirmingham}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.45\linewidth]{data/Birmingham/Plots/Growth_Results/growth_egalitarian_priority_all_FLOW}
-
-}
-
-\caption{Road Segment Priority (left), disaggregated by road type (right) - Egalitarian Growth}\label{fig:growth3MapandBarBirmingham}
-\end{figure}
-
-\clearpage
-
-\hypertarget{leeds}{%
-\subsection{Leeds}\label{leeds}}
-
-\subsubsection{Potential Demand}
-
-\begin{figure}
-\includegraphics[width=0.3\linewidth]{data/Leeds/Plots/histogram_distance_all_vs_cycling} \includegraphics[width=0.3\linewidth]{data/Leeds/Plots/histogram_distance_all_vs_cycling_potential} \includegraphics[width=0.3\linewidth]{data/Leeds/Plots/histogram_distance_cycling_potential_vs_current} \caption{Distribution of Potential Cycling Demand}\label{fig:potdemhistogramsLeeds}
-\end{figure}
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.8\linewidth]{data/Leeds/Plots/desire_facet_cycling}
-
-}
-
-\caption{Current and Potential Cycling Demand}\label{fig:desirefacetcyclingLeeds}
-\end{figure}
-
-\clearpage
-
-\subsubsection{Routing Cycling Flows}
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.75\linewidth]{data/Leeds/Plots/flows_facet_unweighted_Leeds}
-
-}
-
-\caption{Flow Results Based on Unweighted Shortest Paths (Leeds)}\label{fig:flowsfacetunweightedLeeds}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.75\linewidth]{data/Leeds/Plots/flows_facet_weighted_Leeds}
-
-}
-
-\caption{Flow Results Based on Weighted Shortest Paths (Leeds)}\label{fig:flowsfacetweightedLeeds}
-\end{figure}
-
-\clearpage
-
-\subsubsection{Community Detection}
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.9\linewidth]{data/Leeds/Plots/communities_alternative_Leeds}
-
-}
-
-\caption{Communities Based on Potential Cycling Demand (Leeds)}\label{fig:communitiesLeeds}
-\end{figure}
-
-\subsubsection{Network Expansion Algorithms}
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.45\linewidth]{data/Leeds/Plots/Growth_Results/growth_utilitarian_satisfied_km_both_flow_column} \includegraphics[width=0.45\linewidth]{data/Leeds/Plots/Growth_Results/growth_egalitarian_satisfied_km_both_flow_column}
-
-}
-
-\caption{Comparing Overall (Dashed) and Community Level Person-Km Satisfied (Leeds)}\label{fig:growthtotalLeeds}
-\end{figure}
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.45\linewidth]{data/Leeds/Plots/Growth_Results/growth_util_egal_components_number_comparisonLeeds} \includegraphics[width=0.45\linewidth]{data/Leeds/Plots/Growth_Results/growth_util_egal_components_gcc_comparisonLeeds}
-
-}
-
-\caption{Network Characteristics}\label{fig:componentsandGCCLeeds}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.45\linewidth]{data/Leeds/Plots/Growth_Results/growth_egalitarian_priority_all_FLOW} \includegraphics[width=0.45\linewidth]{data/Leeds/Plots/Growth_Results/growth_egalitarian_investment_highways_flow}
-
-}
-
-\caption{Road Segment Priority (left), disaggregated by road type (right) - Egalitarian Growth}\label{fig:growth3MapandBarLeeds}
-\end{figure}
-
-\clearpage
-
-\hypertarget{nottingham}{%
-\subsection{Nottingham}\label{nottingham}}
-
-\subsubsection{Potential Demand}
-
-\begin{figure}
-\includegraphics[width=0.3\linewidth]{data/Nottingham/Plots/histogram_distance_all_vs_cycling} \includegraphics[width=0.3\linewidth]{data/Nottingham/Plots/histogram_distance_all_vs_cycling_potential} \includegraphics[width=0.3\linewidth]{data/Nottingham/Plots/histogram_distance_cycling_potential_vs_current} \caption{Distribution of Potential Cycling Demand}\label{fig:potdemhistogramsNottingham}
-\end{figure}
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.8\linewidth]{data/Nottingham/Plots/desire_facet_cycling}
-
-}
-
-\caption{Current and Potential Cycling Demand}\label{fig:desirefacetcyclingNottingham}
-\end{figure}
-
-\clearpage
-
-\subsubsection{Routing Cycling Flows}
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.75\linewidth]{data/Nottingham/Plots/flows_facet_unweighted_Nottingham}
-
-}
-
-\caption{Flow Results Based on Unweighted Shortest Paths (Nottingham)}\label{fig:flowsfacetunweightedNottingham}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.75\linewidth]{data/Nottingham/Plots/flows_facet_weighted_Nottingham}
-
-}
-
-\caption{Flow Results Based on Weighted Shortest Paths (Nottingham)}\label{fig:flowsfacetweightedNottingham}
-\end{figure}
-
-\clearpage
-
-\subsubsection{Community Detection}
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.9\linewidth]{data/Nottingham/Plots/communities_alternative_Nottingham}
-
-}
-
-\caption{Communities Based on Potential Cycling Demand (Nottingham)}\label{fig:communitiesNottingham}
-\end{figure}
-
-\subsubsection{Network Expansion Algorithms}
-
-\begin{figure}[H]
-\includegraphics[width=0.45\linewidth]{data/Nottingham/Plots/Growth_Results/growth_utilitarian_satisfied_km_both_flow_column} \includegraphics[width=0.45\linewidth]{data/Nottingham/Plots/Growth_Results/growth_egalitarian_satisfied_km_both_flow_column} \caption{Comparing Overall (Dashed) and Community Level Person-Km Satisfied (Nottingham)}\label{fig:growthtotalNottingham}
-\end{figure}
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.45\linewidth]{data/Nottingham/Plots/Growth_Results/growth_util_egal_components_number_comparisonNottingham} \includegraphics[width=0.45\linewidth]{data/Nottingham/Plots/Growth_Results/growth_util_egal_components_gcc_comparisonNottingham}
-
-}
-
-\caption{Network Characteristics}\label{fig:componentsandGCCNottingham}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.45\linewidth]{data/Nottingham/Plots/Growth_Results/growth_egalitarian_priority_all_FLOW} \includegraphics[width=0.45\linewidth]{data/Nottingham/Plots/Growth_Results/growth_egalitarian_investment_highways_flow}
-
-}
-
-\caption{Road Segment Priority (left), disaggregated by road type (right) - Egalitarian Growth}\label{fig:growth3MapandBarNottingham}
-\end{figure}
-
-\clearpage
-
 \end{document}

### 2022-09-03 paper-original-submission.tex
Add tracked changes, fix #15

--- 2022-09-03 README.tex
@@
 \usepackage{multicol}
-\usepackage{booktabs}
-\usepackage{array}
-\usepackage{wrapfig}
-\usepackage{colortbl}
-\usepackage{pdflscape}
-\usepackage{tabu}
-\usepackage{threeparttable}
-\usepackage{threeparttablex}
-\usepackage[normalem]{ulem}
-\usepackage{makecell}
-\usepackage{xcolor}
 \ifLuaTeX
@@
 \title{A Road Segment Prioritization Approach for Cycling Infrastructure}
-\author{true \and true \and true}
-\date{}
+\author{}
+\date{\vspace{-2.5em}}
 
@@
 Although each of these methods has been applied to cycling infrastructure prioritization in previous research, this is the first time that they have been combined, creating an integrated road segment prioritization approach.
-The approach, which can be applied to other cities, is demonstrated in a case study of Manchester, resulting in cycling networks that balance directness against the need for safe and stress-free routes under different investment scenarios.
+The approach, which can be applied to other cities, as shown in the Appendix, is demonstrated in a case study of Manchester, resulting in cycling networks that balance directness against the need for safe and stress-free routes under different investment scenarios.
 A key benefit of the approach from a policy perspective is its ability to support egalitarian and cost-effective strategic cycle network planning.
@@
 commuting across the country; it must be used to design adequate cycling
-infrastructure that is based on motivators and deterrents to cycling. In this paper we
-propose an approach for planning a connected cycling network that is meant to
-be applicable to any city (as shown in the Appendix), and we use the city of Manchester as a case study. Greater
-Manchester aims to become the first region to reach a carbon neutral transport network;
-its local authorities have developed plans for a 1800 mile walking and cycling network
-(Bee Network) that will be delivered over 10 years (Greater Manchester, 2020). While schemes
-have already been put in place for around a quarter of this network, the majority is yet to be built.
-
-\hypertarget{literature-review}{%
-\section{Literature Review}\label{literature-review}}
+infrastructure that is based on motivators and deterrents to cycling.
 
@@
 
-\hypertarget{network-level-approaches-to-bicycle-infrastructure-planning}{%
-\subsection{Network-Level Approaches to Bicycle Infrastructure Planning}\label{network-level-approaches-to-bicycle-infrastructure-planning}}
+\hypertarget{network-level-approaches}{%
+\subsection{Network-Level Approaches}\label{network-level-approaches}}
 
@@
 limitation of these approaches is that they do not explicitly solve for
-continuity. Continuity is addressed using either a constraint specifying
+continuity. The latter is addressed using either a constraint specifying
 that each link with a bike lane should be connected to at least one
@@
 the network through the graph-theoretic concept of \emph{connected
-components}. Orozco et al. (2020) study the existing cycling network in terms of its disconnected
+components}. (\textbf{natera2019data?}) for example,
+study the existing cycling network in terms of its disconnected
 components.
@@
 use a modularity maximization approach (Blondel et al., 2008) on taxi trip
-data to identify seven different communities in the city of Isfahan, Iran.
+data to identify 7 different communities in the city of Isfahan, Iran.
 An optimization problem is then formulated to connect nodes within each
@@
 
-Our work builds on these community finding approaches by proposing a
+Our work builds on these \emph{community finding} approaches by proposing a
 similar greedy network expansion algorithm for cycle network expansion
-\emph{within communities}. We incorporate community finding methods for study
+within communities. We incorporate community finding methods for study
 area partitioning with weighted routing to avoid links that are
-stressful to cycle on. In doing so, we propose an approach that ensures a better distribution
-investments across the study area while also accounting for motivators and deterrents to cycling.
-We propose three sub-methods that address some of the limitations of previous
-studies. These limitations include (a) an inherent bias when basing network
+stressful to cycle on. In doing so, we propose an approach that
+accounts for motivators and deterrents to cycling. We propose three
+sub-methods that address some of the limitations of previous
+studies. These limitations include (a) an inherent bias of basing network
 design solely on existing cycling demand, (b) the proposal of routes that may
@@
 \emph{OD} data can be obtained from a range of sources,
-including GPS data, household travels surveys and census data on work
+including GPS data, household travels surveys and Census data on work
 locations. In areas where observed OD data is unavailable, modelling
@@
 \noindent where \(d\) and \(s\) are the distance and slope associated with each
-OD pair, and \(\alpha\) and all \(\beta\)s are parameters calculated by a regression
+OD pair, and \(\alpha\) and all \(\beta\)s, are parameters calculated by a regression
 model on the training data. The square and square-root distance terms
 ``capture the non-linear impact of distance on the likelihood of
-cycling'', and interaction terms capture the combined effect of slope
+cycling'', and interaction terms to capture the combined effect of slope
 and distance (Lovelace et al., 2017). Alternative cycling uptake
@@
 
-{\centering \includegraphics[width=0.9\linewidth]{data/Manchester/Plots/histogram_distance_cycling_potential_current_all}
+{\centering \includegraphics[width=1\linewidth]{data/Manchester/Plots/histogram_distance_cycling_potential_current_all}
 
@@
 
-For the purposes of this paper we created a similar weighting profile
+For the purposes of this paper, we created a similar weighting profile
 that is adjusted to favor less stressful roads (based on the definition
 by Sorton and Walsh (1994) and information from Table
-\ref{table:osmroadtypes}) and roads with existing cycling
+\ref{table:osmroadtypes}), and roads with existing cycling
 infrastructure. We believe this to be more appropriate than the simplistic
-approach adopted by Mauttone et al. (2017) as it makes use of a hierarchy
+approach adopted by Mauttone et al. (2017), as it makes use of a hierarchy
 of road preference based on perceived stress levels, going beyond simply
-favouring roads with existing cycling infrastructure.
+favoring roads with existing cycling infrastructure.
 
@@
 approach, we use community detection to partition the study area and
-distribute investment over the different subdivisions.
+evaluate the distribution of investment over the different subdivisions.
 
@@
 Component in the proposed bicycle network (Figure
-\ref{fig:componentsandGCC}). Here, however, we find that there is little
-difference between both algorithms.
+\ref{fig:componentsandGCC}). Here however, we find that there is little
+difference between both Algorithms.
 
@@
 an evidence-base for designing direct, connected,
-and low-stress networks. A core contribution of this approach is that it insures
-that investments in cycling infrastructure are distributed fairly between different
-communities in a city.
+and low-stress networks.
 
@@
 cities\footnote{The results are easily reproducible for all UK cities, and can also be reproduced for cities elsewhere given the availability of commuter data.
-Instructions for reproducing the results are detailed in the \textbf{README} file in the R folder of this \href{https://github.com/Hussein-Mahfouz/cycle-networks}{Github repository}}.
+Instructions for reproducing the results are detailed in the \textbf{README} file in the R folder of this \href{https://github.com/Hussein-Mahfouz/cycle-networks}{Github repository}}
 
@@
 Moreover, the inclusion of egalitarian principles in scenarios of change
-encourages investment in cycling infrastructure that addresses geographical and
-social inequalities. This is done by explicitly ensuring that infrastructure improvements
-are pursued in different parts of the city simultaneously. This approach compares well with one
-that focuses exclusively on maximizing the number of potential cycling trips that
-benefit from investment, showing that a fair distribution can be achieved without
-compromising on overall benefit.
+encourages investment in cycling infrastructure to
+increase the connectivity of existing cycling infrastructure
+\emph{and} investment that addresses geographical and social inequalities.
 This ability to address inequalities in network prioritization is particularly
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-manchestercyclingbee2020}{}}%
-Greater Manchester, T. for, 2020. Change a region to change a nation - greater manchester's walking and cycling investment plan.
-
 \leavevmode\vadjust pre{\hypertarget{ref-iacono2010measuring}{}}%
@@
 ONS, 2011. 2011 census: Special workplace statistics (united kingdom).
-
-\leavevmode\vadjust pre{\hypertarget{ref-orozco2020data}{}}%
-Orozco, L.G.N., Battiston, F., Iñiguez, G., Szell, M., 2020. Data-driven strategies for optimal bicycle network growth. Royal Society open science 7.
 

### 2022-09-03 latexdiff.R

@@ -5,4 +5,8 @@ download.file(
   "https://github.com/Hussein-Mahfouz/cycle-networks/raw/ce4827fc708d510bead10d4998ce7ba875286d0d/README.Rmd",
   "paper-original-submission.Rmd"
   )
+file.edit("paper-original-submission.Rmd")
 rmarkdown::render("paper-original-submission.Rmd")
+system("latexdiff paper-original-submission.tex README.tex > tracked-changes.tex")
+tinytex::pdflatex("tracked-changes.tex")
+browseURL("tracked-changes.pdf")

### 2022-09-04 README.Rmd

@@
 title: A Road Segment Prioritization Approach for Cycling Infrastructure
-author:
-  - Hussein Mahfouz:
-      email: h.mahfouz.19@ucl.ac.uk
-      institute: casa
-      correspondence: true
-  - name: Elsa Arcaute
-    institute: casa
-  - name: Robin Lovelace
-    institute: its
-institute:
-  - casa: Centre for Advanced Spatial Analysis (CASA), University College London, London, UK
-  - its: Institute for Transport Studies, University of Leeds, Leeds, UK
+# author:
+#   - Hussein Mahfouz:
+#       email: h.mahfouz.19@ucl.ac.uk
+#       institute: casa
+#       correspondence: true
+#   - name: Elsa Arcaute
+#     institute: casa
+#   - name: Robin Lovelace
+#     institute: its
+# institute:
+#   - casa: Centre for Advanced Spatial Analysis (CASA), University College London, London, UK
+#   - its: Institute for Transport Studies, University of Leeds, Leeds, UK
 

### 2022-09-04 paper-original-submission.Rmd

@@
 title: A Road Segment Prioritization Approach for Cycling Infrastructure
-author:
-  - Hussein Mahfouz:
-      email: h.mahfouz.19@ucl.ac.uk
-      institute: casa
-      correspondence: true
-  - name: Elsa Arcaute
-    institute: casa
-  - name: Robin Lovelace
-    institute: its
-institute:
-  - casa: Centre for Advanced Spatial Analysis (CASA), University College London, London, UK
-  - its: Institute of Transport Studies, University of Leeds, Leeds, UK
+# author:
+#   - Hussein Mahfouz:
+#       email: h.mahfouz.19@ucl.ac.uk
+#       institute: casa
+#       correspondence: true
+#   - name: Elsa Arcaute
+#     institute: casa
+#   - name: Robin Lovelace
+#     institute: its
+# institute:
+#   - casa: Centre for Advanced Spatial Analysis (CASA), University College London, London, UK
+#   - its: Institute of Transport Studies, University of Leeds, Leeds, UK
 

### 2022-09-04 response to reviewers.Rmd

@@
 ---
-
-# Overall Comments
 
@@
 
--   The big picture take aways from the work were also not clear...
+-   *The big picture take aways from the work were also not clear...*
+
+We have added a couple of lines to the conclusion to emphasise the contribution
 
@@
 We have chosen option 2 as we think it is better to show the methodology and results for each sub-secion together.
-Some context was added on Manchester and its recent cycling and walking targets (specifically the Change a Region to Change a Nation report [[link](<https://beeactive.tfgm.com/publications-and-downloads/>)].
+Some context was added on Manchester and its recent cycling and walking targets (specifically the Change a Region to Change a Nation report [[link](https://beeactive.tfgm.com/publications-and-downloads/)].
 

### 2022-09-14 response to reviewers.Rmd
edit response doc

@@
 title: "Response to Reviewers"
-author: "Hussein"
-date: "8/29/2022"
+#author: "Hussein"
+date: "9/13/2022"
 output:
@@
 We agree that this is a methods paper, and in fact focus on the **approach**​ (which could be implemented using different methods, we outline one) as highlighted below.
-We have rewritten relevant parts of the paper to emphasise the methodological advances outlined in the paper.
-Furthermore, we have demonstrated applying the methods to additional locations in the appendix.
 The abstract reads as follows (emphasis added to highlight the framing as a methods paper):
@@
 ***Although each of these methods has been applied to cycling infrastructure prioritization in previous research, this is the first time that they have been combined, creating an integrated road segment prioritization approach***."
+
+We have also modified the manuscript in order to make this clearer as suggested.
+This has mostly been done in the Discussion and Conclusion section.
+Furthermore, we have demonstrated applying the methods to additional locations in the appendix.
 
@@
 
-end of section 2.2: "while these approaches deal with continuity better, they look at the network as a whole when attempting to improve it, and in doing so fail to account for equitable distribution of infrastructure
+end of section 2.2: "while these approaches deal with continuity better, they look at the network as a whole when attempting to improve it, and in doing so fail to account for equitable distribution of infrastructure"
 
@@
 
-a major challenge facing 'top-down' planning approaches is how to incorporate egalitarian principles by fairly distributing investments in cycling infrastructure.
+"a major challenge facing 'top-down' planning approaches is how to incorporate egalitarian principles by fairly distributing investments in cycling infrastructure.
 One way of quantifying this is to split up the city into smaller geospatial areas and target equal investment in each of those areas.....Community detection offers us a way to delineate such a split; cyclists are limited in their commuting distance (see Figure 6), and so trip attractors are more likely to have a local catchment area of cyclists.
@@
 
-We have added a couple of lines to the conclusion to emphasise the contribution
+We have added a couple of lines to the conclusion to emphasise the contribution (emphasized in bold):
+
+1.  "**A core contribution of this approach is that it insures that investments in cycling infrastructure are distributed fairly between different communities in a city**"
+
+2.
+"The results can therefore be used as a basis for recommendations on road space reallocation \*and\* new infrastructure to unlock potential cycling demand.
+The approach encourages consideration of a wider range of preferences and needs than previous approaches that focus only on absolute potential.
+Moreover, the inclusion of egalitarian principles in scenarios of change encourages investment in cycling infrastructure that addresses geographical and social inequalities.
+**This is done by explicitly ensuring that infrastructure improvements are pursued in different parts of the city simultaneously.
+This approach compares well with one that focuses exclusively on maximizing the number of potential cycling trips that benefit from investment, showing that a fair distribution can be achieved without compromising on overall benefit.**"
 
@@
 
-We have chosen option 2 as we think it is better to show the methodology and results for each sub-secion together.
-Some context was added on Manchester and its recent cycling and walking targets (specifically the Change a Region to Change a Nation report [[link](https://beeactive.tfgm.com/publications-and-downloads/)].
+We have chosen option 2 as we think it is better to show the methodology and results for each subsection together.
+Some context was added on Manchester and its recent cycling and walking targets (specifically the *Change a Region to Change a Nation* report [[link](https://beeactive.tfgm.com/publications-and-downloads/)].
 
-We think that the methodology is generic enough to be scalable.
-The readme in the github page (linked in the paper) has step by step instructions for reproducibility, and the appendix includes examples fr two other UK cities.
+We have created a methodology that is scalable and we have provided the code and instructions in a Github repository.
+The instructions are in the README file in the R folder of the .zip file that accompanies this paper.
+We have now clarified this and added a footnote that says "The results are easily reproducible for all UK cities, and can also be reproduced for cities elsewhere given the availability of commuter data.
+Instructions for reproducing the results are detailed in the **README** file in the R folder of this Github repository".
+A link to the repo will not be added until the paper is accepted to keep the authors anonymous
 
@@
 
-This has been done now
+This has been done now. We have an Introduction section follow by a Literature Review section
 

### 2022-09-14 README.Rmd
edit response doc

@@
 cities^[The results are easily reproducible for all UK cities, and can also be reproduced for cities elsewhere given the availability of commuter data.
+Instructions for reproducing the results are detailed in the **README** file in the R folder of the .zip file that accompanies this paper (Link to public github repo to be added when paper is published)].
+
+<!-- ^[The results are easily reproducible for all UK cities, and can also be reproduced for cities elsewhere given the availability of commuter data.
 Instructions for reproducing the results are detailed in the **README** file in the R folder of this [Github repository](https://github.com/Hussein-Mahfouz/cycle-networks)].
+-->
 

### 2022-09-14 README.tex
edit response doc

@@
 ]{article}
-\usepackage{amsmath,amssymb}
 \usepackage{lmodern}
-\usepackage{iftex}
-\ifPDFTeX
+\usepackage{amsmath}
+\usepackage{ifxetex,ifluatex}
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
   \usepackage[T1]{fontenc}
@@
   \usepackage{textcomp} % provide euro and other symbols
+  \usepackage{amssymb}
 \else % if luatex or xetex
@@
 \usepackage{xcolor}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\hypersetup{
+  pdftitle={A Road Segment Prioritization Approach for Cycling Infrastructure},
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
+\urlstyle{same} % disable monospaced font for URLs
 \usepackage[margin=2cm]{geometry}
-\usepackage{longtable,booktabs,array}
+\usepackage{longtable,booktabs}
 \usepackage{calc} % for calculating minipage widths
@@
 \setcounter{secnumdepth}{5}
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-\newlength{\cslentryspacingunit} % times entry-spacing
-\setlength{\cslentryspacingunit}{\parskip}
-\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
-  % turn on hanging indent if param 1 is 1
-  \ifodd #1
-  \let\oldpar\par
-  \def\par{\hangindent=\cslhangindent\oldpar}
-  \fi
-  % set entry spacing
-  \setlength{\parskip}{#2\cslentryspacingunit}
- }%
- {}
-\usepackage{calc}
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 \usepackage{subcaption}
@@
 \usepackage{xcolor}
-\ifLuaTeX
+\ifluatex
   \usepackage{selnolig}  % disable illegal ligatures
 \fi
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\urlstyle{same} % disable monospaced font for URLs
-\hypersetup{
-  pdftitle={A Road Segment Prioritization Approach for Cycling Infrastructure},
-  hidelinks,
-  pdfcreator={LaTeX via pandoc}}
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
+  % set entry spacing
+  \ifnum #2 > 0
+  \setlength{\parskip}{#2\baselineskip}
+  \fi
+ }%
+ {}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 
 \title{A Road Segment Prioritization Approach for Cycling Infrastructure}
-\author{true \and true \and true}
-\date{}
+\author{}
+\date{\vspace{-2.5em}}
 
@@
 demand. It then introduces the network expansion algorithms, and compares an approach grounded in
-`egalitarianism' to one grounded in `utilitarianism'.
+`egalitarianism' to one grounded in `utilitarianism.'
 
@@
 ``capture the non-linear impact of distance on the likelihood of
-cycling'', and interaction terms capture the combined effect of slope
+cycling,'' and interaction terms capture the combined effect of slope
 and distance (Lovelace et al., 2017). Alternative cycling uptake
@@
 \noindent A weighted distance \(d_{w}\) for each road
-segment is calculated as follows \footnote{The \textbf{dodgr} r package
+segment is calculated as follows:\footnote{The \textbf{dodgr} r package
   (Padgham, 2019) is used to route cycling demand onto the road
@@
   For the sake of reproducibility, we stick to
-  the convention used in the package.}:
+  the convention used in the package.}
 
@@
 cities\footnote{The results are easily reproducible for all UK cities, and can also be reproduced for cities elsewhere given the availability of commuter data.
-Instructions for reproducing the results are detailed in the \textbf{README} file in the R folder of this \href{https://github.com/Hussein-Mahfouz/cycle-networks}{Github repository}}.
+Instructions for reproducing the results are detailed in the \textbf{README} file in the R folder of the .zip file that accompanies this paper (Link to public github repo to be added when paper is published)}.
 
@@
 \begin{CSLReferences}{1}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-akbarzadeh2018designing}{}}%
+\leavevmode\hypertarget{ref-akbarzadeh2018designing}{}%
 Akbarzadeh, M., Mohri, S.S., Yazdian, E., 2018. Designing bike networks using the concept of network clusters. Applied network science 3, 12.
 
-\leavevmode\vadjust pre{\hypertarget{ref-aldred2019impacts}{}}%
+\leavevmode\hypertarget{ref-aldred2019impacts}{}%
 Aldred, R., Croft, J., Goodman, A., 2019.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-bao2017planning}{}}%
+\leavevmode\hypertarget{ref-bao2017planning}{}%
 Bao, J., He, T., Ruan, S., Li, Y., Zheng, Y., 2017.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-black1995spatial}{}}%
+\leavevmode\hypertarget{ref-black1995spatial}{}%
 Black, W.R., 1995. Spatial interaction modeling using artificial neural networks. Journal of Transport Geography 3, 159--166.
 
-\leavevmode\vadjust pre{\hypertarget{ref-blondel2008fast}{}}%
+\leavevmode\hypertarget{ref-blondel2008fast}{}%
 Blondel, V.D., Guillaume, J.-L., Lambiotte, R., Lefebvre, E., 2008. Fast unfolding of communities in large networks. Journal of statistical mechanics: theory and experiment 2008, P10008.
 
-\leavevmode\vadjust pre{\hypertarget{ref-brand_climate_2020}{}}%
+\leavevmode\hypertarget{ref-brand_climate_2020}{}%
 Brand, C., Dons, E., Anaya-Boig, E., Avila-Palencia, I., Clark, A., de Nazelle, A., Gascon, M., Gaupp-Berghausen, M., Gerike, R., Gotschi, T., 2020.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-broach2011bicycle}{}}%
+\leavevmode\hypertarget{ref-broach2011bicycle}{}%
 Broach, J., Gliebe, J., Dill, J., 2011. Bicycle route choice model developed using revealed preference GPS data, in: 90th Annual Meeting of the Transportation Research Board, Washington, DC.
 
-\leavevmode\vadjust pre{\hypertarget{ref-caulfield2012determining}{}}%
+\leavevmode\hypertarget{ref-caulfield2012determining}{}%
 Caulfield, B., Brick, E., McCarthy, O.T., 2012. Determining bicycle infrastructure preferences--a case study of dublin. Transportation research part D: transport and environment 17, 413--417.
 
-\leavevmode\vadjust pre{\hypertarget{ref-celis2017association}{}}%
+\leavevmode\hypertarget{ref-celis2017association}{}%
 Celis-Morales, C.A., Lyall, D.M., Welsh, P., Anderson, J., Steell, L., Guo, Y., Maldonado, R., Mackay, D.F., Pell, J.P., Sattar, N., others, 2017.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-crane2017longitudinal}{}}%
+\leavevmode\hypertarget{ref-crane2017longitudinal}{}%
 Crane, M., Rissel, C., Standen, C., Ellison, A., Ellison, R., Wen, L.M., Greaves, S., 2017.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-departmentgearchange2020}{}}%
+\leavevmode\hypertarget{ref-departmentgearchange2020}{}%
 DfT, 2020a. Gear change: A bold vision for cycling and walking.
 
-\leavevmode\vadjust pre{\hypertarget{ref-departmentcycleinfradesign2020}{}}%
+\leavevmode\hypertarget{ref-departmentcycleinfradesign2020}{}%
 DfT, 2020b. Cycling infrastructure design.
 
-\leavevmode\vadjust pre{\hypertarget{ref-department2012guidance}{}}%
+\leavevmode\hypertarget{ref-department2012guidance}{}%
 DfT, 2012. Guidance on road classification and the primary route network.
 
-\leavevmode\vadjust pre{\hypertarget{ref-dill2013four}{}}%
+\leavevmode\hypertarget{ref-dill2013four}{}%
 Dill, J., McNeil, N., 2013. Four types of cyclists? Examination of typology for better understanding of bicycling behavior and potential. Transportation Research Record 2387, 129--138.
 
-\leavevmode\vadjust pre{\hypertarget{ref-duthie2014optimization}{}}%
+\leavevmode\hypertarget{ref-duthie2014optimization}{}%
 Duthie, J., Unnikrishnan, A., 2014. Optimization framework for bicycle network design. Journal of Transportation engineering 140, 04014028.
 
-\leavevmode\vadjust pre{\hypertarget{ref-furth2016network}{}}%
+\leavevmode\hypertarget{ref-furth2016network}{}%
 Furth, P.G., Mekuria, M.C., Nixon, H., 2016. Network connectivity for low-stress bicycling. Transportation research record 2587, 41--49.
 
-\leavevmode\vadjust pre{\hypertarget{ref-gehrke2020cycling}{}}%
+\leavevmode\hypertarget{ref-gehrke2020cycling}{}%
 Gehrke, S.R., Akhavan, A., Furth, P.G., Wang, Q., Reardon, T.G., 2020.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-goodman2019scenarios}{}}%
+\leavevmode\hypertarget{ref-goodman2019scenarios}{}%
 Goodman, A., Rojas, I.F., Woodcock, J., Aldred, R., Berkoff, N., Morgan, M., Abbas, A., Lovelace, R., 2019.
-Scenarios of cycling to school in england, and associated health and carbon impacts: Application of the {``propensity to cycle tool.''} Journal of Transport \& Health 12, 263--278.
-
-\leavevmode\vadjust pre{\hypertarget{ref-goodman2014new}{}}%
+Scenarios of cycling to school in england, and associated health and carbon impacts: Application of the {`propensity to cycle tool.'} Journal of Transport \& Health 12, 263--278.
+
+\leavevmode\hypertarget{ref-goodman2014new}{}%
 Goodman, A., Sahlqvist, S., Ogilvie, D., Consortium, iConnect, 2014.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-gossling2013urban}{}}%
+\leavevmode\hypertarget{ref-gossling2013urban}{}%
 Gössling, S., 2013. Urban transport transitions: Copenhagen, city of cyclists. Journal of Transport Geography 33, 196--206.
 
-\leavevmode\vadjust pre{\hypertarget{ref-manchestercyclingbee2020}{}}%
+\leavevmode\hypertarget{ref-manchestercyclingbee2020}{}%
 Greater Manchester, T. for, 2020. Change a region to change a nation - greater manchester's walking and cycling investment plan.
 
-\leavevmode\vadjust pre{\hypertarget{ref-iacono2010measuring}{}}%
+\leavevmode\hypertarget{ref-iacono2010measuring}{}%
 Iacono, M., Krizek, K.J., El-Geneidy, A., 2010. Measuring non-motorized accessibility: Issues, alternatives, and execution. Journal of Transport Geography 18, 133--140.
 
-\leavevmode\vadjust pre{\hypertarget{ref-jafino2020transport}{}}%
+\leavevmode\hypertarget{ref-jafino2020transport}{}%
 Jafino, B.A., Kwakkel, J., Verbraeck, A., 2020. Transport network criticality metrics: A comparative analysis and a guideline for selection. Transport Reviews 40, 241--264.
 
-\leavevmode\vadjust pre{\hypertarget{ref-jarrett2012effect}{}}%
+\leavevmode\hypertarget{ref-jarrett2012effect}{}%
 Jarrett, J., Woodcock, J., Griffiths, U.K., Chalabi, Z., Edwards, P., Roberts, I., Haines, A., 2012.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-kohl2012pandemic}{}}%
+\leavevmode\hypertarget{ref-kohl2012pandemic}{}%
 Kohl 3rd, H.W., Craig, C.L., Lambert, E.V., Inoue, S., Alkandari, J.R., Leetongin, G., Kahlmeier, S., Group, L.P.A.S.W., others, 2012.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-lovelace2017propensity}{}}%
+\leavevmode\hypertarget{ref-lovelace2017propensity}{}%
 Lovelace, R., Goodman, A., Aldred, R., Berkoff, N., Abbas, A., Woodcock, J., 2017.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-lucas2016method}{}}%
+\leavevmode\hypertarget{ref-lucas2016method}{}%
 Lucas, K., Van Wee, B., Maat, K., 2016. A method to evaluate equitable accessibility: Combining ethical theories and accessibility-based approaches. Transportation 43, 473--490.
 
-\leavevmode\vadjust pre{\hypertarget{ref-marques2015infrastructure}{}}%
+\leavevmode\hypertarget{ref-marques2015infrastructure}{}%
 Marqués, R., Hernández-Herrador, V., Calvo-Salazar, M., García-Cebrián, J., 2015.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-martinez2013new}{}}%
+\leavevmode\hypertarget{ref-martinez2013new}{}%
 Martínez, L.M., Viegas, J.M., 2013. A new approach to modelling distance-decay functions for accessibility assessment in transport studies. Journal of Transport Geography 26, 87--96.
 
-\leavevmode\vadjust pre{\hypertarget{ref-mauttone2017bicycle}{}}%
+\leavevmode\hypertarget{ref-mauttone2017bicycle}{}%
 Mauttone, A., Mercadante, G., Rabaza, M., Toledo, F., 2017. Bicycle network design: Model and solution algorithm. Transportation research procedia 27, 969--976.
 
-\leavevmode\vadjust pre{\hypertarget{ref-mesbah2012bilevel}{}}%
+\leavevmode\hypertarget{ref-mesbah2012bilevel}{}%
 Mesbah, M., Thompson, R., Moridpour, S., 2012. Bilevel optimization approach to design of network of bike lanes. Transportation research record 2284, 21--28.
 
-\leavevmode\vadjust pre{\hypertarget{ref-nahmias2017integrating}{}}%
+\leavevmode\hypertarget{ref-nahmias2017integrating}{}%
 Nahmias-Biran, B., Martens, K., Shiftan, Y., 2017. Integrating equity in transportation project assessment: A philosophical exploration and its practical implications. Transport reviews 37, 192--210.
 
-\leavevmode\vadjust pre{\hypertarget{ref-nello2020environmental}{}}%
+\leavevmode\hypertarget{ref-nello2020environmental}{}%
 Nello-Deakin, S., 2020. Environmental determinants of cycling: Not seeing the forest for the trees? Journal of transport geography 85, 102704.
 
-\leavevmode\vadjust pre{\hypertarget{ref-olmos2020data}{}}%
+\leavevmode\hypertarget{ref-olmos2020data}{}%
 Olmos, L.E., Tadeo, M.S., Vlachogiannis, D., Alhasoun, F., Alegre, X.E., Ochoa, C., Targa, F., González, M.C., 2020.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-ofn2018population}{}}%
+\leavevmode\hypertarget{ref-ofn2018population}{}%
 ONS, 2018. Population estimates for the UK, england and wales, scotland and northern ireland: Mid-2017. Hampshire: Office for National Statistics.
 
-\leavevmode\vadjust pre{\hypertarget{ref-ONS2011flowdata}{}}%
+\leavevmode\hypertarget{ref-ONS2011flowdata}{}%
 ONS, 2011. 2011 census: Special workplace statistics (united kingdom).
 
-\leavevmode\vadjust pre{\hypertarget{ref-orozco2020data}{}}%
+\leavevmode\hypertarget{ref-orozco2020data}{}%
 Orozco, L.G.N., Battiston, F., Iñiguez, G., Szell, M., 2020. Data-driven strategies for optimal bicycle network growth. Royal Society open science 7.
 
-\leavevmode\vadjust pre{\hypertarget{ref-padgham2019dodgr}{}}%
+\leavevmode\hypertarget{ref-padgham2019dodgr}{}%
 Padgham, M., 2019. Dodgr: An r package for network flow aggregation. Transport Findings. Network Design Lab.
 
-\leavevmode\vadjust pre{\hypertarget{ref-parkin2018designing}{}}%
+\leavevmode\hypertarget{ref-parkin2018designing}{}%
 Parkin, J., 2018. Designing for cycle traffic. ICE Publishing. \url{https://doi.org/10.1680/dfct.63495}
 
-\leavevmode\vadjust pre{\hypertarget{ref-patterson2020associations}{}}%
+\leavevmode\hypertarget{ref-patterson2020associations}{}%
 Patterson, R., Panter, J., Vamos, E.P., Cummins, S., Millett, C., Laverty, A.A., 2020.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-pereira2017distributive}{}}%
+\leavevmode\hypertarget{ref-pereira2017distributive}{}%
 Pereira, R.H., Schwanen, T., Banister, D., 2017. Distributive justice and equity in transportation. Transport reviews 37, 170--191.
 
-\leavevmode\vadjust pre{\hypertarget{ref-pucher2008making}{}}%
+\leavevmode\hypertarget{ref-pucher2008making}{}%
 Pucher, J., Buehler, R., 2008. Making cycling irresistible: Lessons from the netherlands, denmark and germany. Transport reviews 28, 495--528.
 
-\leavevmode\vadjust pre{\hypertarget{ref-pucher2010infrastructure}{}}%
+\leavevmode\hypertarget{ref-pucher2010infrastructure}{}%
 Pucher, J., Dill, J., Handy, S., 2010. Infrastructure, programs, and policies to increase bicycling: An international review. Preventive medicine 50, S106--S125.
 
-\leavevmode\vadjust pre{\hypertarget{ref-schoner2014missing}{}}%
+\leavevmode\hypertarget{ref-schoner2014missing}{}%
 Schoner, J.E., Levinson, D.M., 2014. The missing link: Bicycle infrastructure networks and ridership in 74 US cities. Transportation 41, 1187--1204.
 
-\leavevmode\vadjust pre{\hypertarget{ref-sorton1994bicycle}{}}%
+\leavevmode\hypertarget{ref-sorton1994bicycle}{}%
 Sorton, A., Walsh, T., 1994. Bicycle stress level as a tool to evaluate urban and suburban bicycle compatibility. Transportation Research Record 17--17.
 
-\leavevmode\vadjust pre{\hypertarget{ref-stinson2003commuter}{}}%
+\leavevmode\hypertarget{ref-stinson2003commuter}{}%
 Stinson, M.A., Bhat, C.R., 2003. Commuter bicyclist route choice: Analysis using a stated preference survey. Transportation Research Record 1828, 107--115.
 
-\leavevmode\vadjust pre{\hypertarget{ref-agreement2015paris}{}}%
+\leavevmode\hypertarget{ref-agreement2015paris}{}%
 UN, 2015. Paris agreement, in: Report of the Conference of the Parties to the United Nations Framework Convention on Climate Change (21st Session, 2015: Paris). Retrived December. HeinOnline, p. 2017.
 
-\leavevmode\vadjust pre{\hypertarget{ref-vidaltortosa_infrastructure_2020}{}}%
+\leavevmode\hypertarget{ref-vidaltortosa_infrastructure_2020}{}%
 Vidal Tortosa, E., Lovelace, R., Heinen, E., Mann, R.P., 2021.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-wardman2007factors}{}}%
-Wardman, M., Tight, M., Page, M., 2007.
-\href{https://doi.org/10.1016/j.tra.2006.09.011}{Factors influencing the propensity to cycle to work}.
-Transportation Research Part A: Policy and Practice 41, 339--350.
-
-\leavevmode\vadjust pre{\hypertarget{ref-wilson1971family}{}}%
+\leavevmode\hypertarget{ref-wardman2007factors}{}%
+Wardman, M., Tight, M., Page, M., 2007. Factors influencing the propensity to cycle to work. Transportation Research Part A: Policy and Practice 41, 339--350.
+
+\leavevmode\hypertarget{ref-wilson1971family}{}%
 Wilson, A.G., 1971. A family of spatial interaction models, and associated developments. Environment and Planning A 3, 1--32.
 
-\leavevmode\vadjust pre{\hypertarget{ref-winters2011motivators}{}}%
+\leavevmode\hypertarget{ref-winters2011motivators}{}%
 Winters, M., Davidson, G., Kao, D., Teschke, K., 2011. Motivators and deterrents of bicycling: Comparing influences on decisions to ride. Transportation 38, 153--168.
 
-\leavevmode\vadjust pre{\hypertarget{ref-winters2010far}{}}%
+\leavevmode\hypertarget{ref-winters2010far}{}%
 Winters, M., Teschke, K., Grant, M., Setton, E.M., Brauer, M., 2010.

### 2022-09-14 paper-original-submission.tex
edit response doc

--- 2021-05-08 arxiv.tex
@@
 \usepackage{makecell}
+\usepackage{xcolor}
 \ifluatex
@@
 
-\textsuperscript{1} The Centre for Advanced Spatial Analysis, University College London, London, UK\\
+\textsuperscript{1} Centre for Advanced Spatial Analysis (CASA), University College London, London, UK\\
 \textsuperscript{2} Institute of Transport Studies, University of Leeds, Leeds, UK
@@
 the network through the graph-theoretic concept of \emph{connected
-components}. Natera et al. (2019) for example,
+components}. (\textbf{natera2019data?}) for example,
 study the existing cycling network in terms of its disconnected
@@
 \noindent where \(d\) and \(s\) are the distance and slope associated with each
-OD pair, and \(\alpha\) and all \(\beta\)s are parameters calculated by a regression
+OD pair, and \(\alpha\) and all \(\beta\)s, are parameters calculated by a regression
 model on the training data. The square and square-root distance terms
@@
 
+\begin{verbatim}
+## Warning: package 'tidyr' was built under R version 4.0.5
+\end{verbatim}
+
+\begin{verbatim}
+## Warning: package 'readr' was built under R version 4.0.5
+\end{verbatim}
+
+\begin{verbatim}
+## Warning: package 'dplyr' was built under R version 4.0.5
+\end{verbatim}
+
+\begin{verbatim}
+## Warning: package 'sf' was built under R version 4.0.5
+\end{verbatim}
+
 \begin{table}[!h]
@@
 needed, opening up the possibility of deploying the approach in other
-cities.\^{}{[}The results are easily reproducible for all UK cities, and can also be reproduced for cities elsewhere given the availability of commuter data.
-Instructions for reproducing the results are detailed in the \textbf{README} file in the R folder of this \href{https://github.com/Hussein-Mahfouz/cycle-networks}{Github repository}
+cities\footnote{The results are easily reproducible for all UK cities, and can also be reproduced for cities elsewhere given the availability of commuter data.
+Instructions for reproducing the results are detailed in the \textbf{README} file in the R folder of this \href{https://github.com/Hussein-Mahfouz/cycle-networks}{Github repository}}
 
@@
 
-\leavevmode\hypertarget{ref-natera2019data}{}%
-Natera, L., Battiston, F., Iñiguez, G., Szell, M., 2019. Data-driven strategies for optimal bicycle network growth. arXiv preprint arXiv:1907.07080.
-
 \leavevmode\hypertarget{ref-nello2020environmental}{}%

### 2022-09-15 response to reviewers.Rmd

@@
 
-We agree that this is a methods paper, and in fact focus on the **approach**​ (which could be implemented using different methods, we outline one) as highlighted below.
-The abstract reads as follows (emphasis added to highlight the framing as a methods paper):
+We agree that this is a methods paper, and in fact focus on the **approach**​, which could be implemented using different methods.
+The paper presents an implementation of the approach.
+We have made this clear in the abstract and throughout (emphasis added to highlight the framing as a methods paper):
 
@@
 
-We have also modified the manuscript in order to make this clearer as suggested.
+We have updated the manuscript to make this clearer as suggested.
 This has mostly been done in the Discussion and Conclusion section.
@@
 
-We think that this paper fits the remit of Transport Geography well (based on the aims and scope of the journal) as it presents a new methodological approach in planning cycling networks that is grounded in both GIS and literature on the impact of infrastructure on cycling uptake.
+The paper clearly fits the remit of Transport Geography well (based on the aims and scope of the journal, and previous papers in the area of geographic prioritisation of transport infrastructure and spatial inequalities in transport provision).
+It presents a new methodological approach in planning cycling networks that is grounded in both GIS and literature on the impact of infrastructure on cycling uptake.
 
@@
 
-[How this approach differs from past work:]{.ul}
+[How this approach differs from past work:]{.underline}
+
+We are confident that the approach does different substantially from previous work in this area.
+This is evidenced in the text below.
 
@@
 
-[How this was done/ benefits (bold is new text):]{.ul}
+[How this was done/ benefits (bold is new text):]{.underline}
+
+We have taken care to explain how this was done and the benefits, with reference to the literature.
+We agree that the description of the benefits of the approach could have been made clearer and have taken steps in that direction.
 
@@
 
-This paper presents an interesting approach to network design for cycling infrastructure. I believe the proposed approach is really interesting and valid.
+This paper presents an interesting approach to network design for cycling infrastructure.
+I believe the proposed approach is really interesting and valid.
 
@@
 Instructions for reproducing the results are detailed in the **README** file in the R folder of this Github repository".
-A link to the repo will not be added until the paper is accepted to keep the authors anonymous
+A link to the repo will be added after the paper is accepted.
 
@@
 
-This has been done now. We have an Introduction section follow by a Literature Review section
+We thanks the reviewer for this helpful comment and agree.
+This has been done now.
+We have an Introduction section follow by a Literature Review section
 

### 2023-04-24 README.Rmd
peer review 2: flowchart and method headings

@@
 
+# Method and Application
+
+## Method overview
+
 This research attempts to propose an egalitarian framework for cycling
@@
 
-<!-- EDIT! Add more to limitations -->
-
 Our work builds on these community finding approaches by proposing a
@@
 demand. It then introduces the network expansion algorithms, and compares an approach grounded in
-'egalitarianism' to one grounded in 'utilitarianism'.
-
-# Calculating Potential Cycling Demand
+'egalitarianism' to one grounded in 'utilitarianism'. As mentioned above, Manchester is used as a case study. 
+An overview of the steps involved is shown in Figure \@ref(fig:methodflowchart).
+
+```{r methodflowchart, fig.cap="Flowchart of proposed methodology", out.width="80%", fig.show='hold', fig.align = "center", echo=FALSE}
+knitr::include_graphics("extras/methodology_flowchart.png")
+```
+
+## Calculating Potential Cycling Demand
 
@@
 
-# Routing
+## Routing
 
@@
 
-# Road Segment Prioritization
+## Road Segment Prioritization
 
@@
 
-## Community Detection
+### Community Detection
 
@@
 
-## Algorithm 1: Utilitarian Expansion
+### Algorithm 1: Utilitarian Expansion
 
@@
 
-## Algorithm 2: Egalitarian Expansion
+### Algorithm 2: Egalitarian Expansion
 
@@
 
-## Connectivity
+### Connectivity
 

### 2023-04-24 README.Rmd
add grise+geneidy citation

@@
 number of commuters between administrative zones --- Middle layer Super
-Output Areas (MSOA) --- by mode of travel. MSOAs have an average
+Output Areas (MSOA) --- by mode of travel (this is similar to the approach 
+used by @grise2018if). 
+MSOAs have an average
 population of just over 8000 [@ofn2018population].^[ See
@@
 Given that we are proposing an approach which can accommodate any cycling uptake functions, this is an acceptable limitation.
+The network prioritization is based strongly on potential cycling demand calculations.
+While this provides a solid supplementary tool for network planning, it would be more useful as part of a multi-criteria approach that considers on-the ground information from existing cyclists, road geometries, and traffic analysis.
+@grise2018if propose a multi-criteria approach that includes the first of these points.
 

### 2023-04-24 README.Rmd
add wysling+purves citation

@@
 without cycling infrastructure by a fixed impedance factor
-[@mauttone2017bicycle], or by assigning different weights to the road
-segments proportional to the investment cost of bringing them to an
-acceptable level of stress for cycling [@duthie2014optimization].
+[@mauttone2017bicycle], by assigning different weights to the road
+segments that are (a) proportional to the investment cost of bringing them to an
+acceptable level of stress for cycling [@duthie2014optimization], or (b) inversely proportional to 
+perceived level of stress experienced by cyclists [@wysling2022improve].
 Perceived stress for cyclists has been found to be a function of vehicular traffic
@@
 favouring roads with existing cycling infrastructure. 
+It is also different to the weighting profile used by @wysling2022improve as it adds a larger
+weight on trunk and primary roads then other roads without cycling infrastructure.
 

### 2023-04-24 README.Rmd
Tweak prose in intro

--- 2023-04-24 README.Rmd
@@
 including electric vehicles. But new vehicle technologies can only go so
-far and do not tackle parallel problems such as congestion, road traffic
+far and fail to tackle parallel problems such as congestion, road traffic
 casualties and physical inactivity [@brand_climate_2020].
 
-In this context, interest and investment in active modes are growing.
-The benefits extend beyond congestion and the environment, as active
+In this context, interest and investment in active modes has grown substantially, especially with the growing realisation that the benefits extend beyond congestion and the environment, as active
 travel can also help alleviate what is referred to as the pandemic of
@@
 lower risk of disease, including cancer and cardiovascular disease
-[@celis2017association; @jarrett2012effect;
-@patterson2020associations]. In the wake of the Covid-19 pandemic, and
+[@celis2017association; @jarrett2012effect; @patterson2020associations]. In the wake of the Covid-19 pandemic, and
 the resulting reduced capacity of public transport, the UK government
@@
 
-The funding on its own is therefore no guarantee of a change in
+Funding alone is therefore no guarantee of a change in
 commuting across the country; it must be used to design adequate cycling
@@
 
-# Method and Application
-
-## Method overview
-
 This research attempts to propose an egalitarian framework for cycling
@@
 
+<!-- EDIT! Add more to limitations -->
+
 Our work builds on these community finding approaches by proposing a
@@
 demand. It then introduces the network expansion algorithms, and compares an approach grounded in
-'egalitarianism' to one grounded in 'utilitarianism'. As mentioned above, Manchester is used as a case study. 
-An overview of the steps involved is shown in Figure \@ref(fig:methodflowchart).
-
-```{r methodflowchart, fig.cap="Flowchart of proposed methodology", out.width="80%", fig.show='hold', fig.align = "center", echo=FALSE}
-knitr::include_graphics("extras/methodology_flowchart.png")
-```
-
-## Calculating Potential Cycling Demand
+'egalitarianism' to one grounded in 'utilitarianism'.
+
+# Calculating Potential Cycling Demand
 
@@
 
-## Routing
+# Routing
 
@@
 
-## Road Segment Prioritization
+# Road Segment Prioritization
 
@@
 
-### Community Detection
+## Community Detection
 
@@
 
-### Algorithm 1: Utilitarian Expansion
+## Algorithm 1: Utilitarian Expansion
 
@@
 
-### Algorithm 2: Egalitarian Expansion
+## Algorithm 2: Egalitarian Expansion
 
@@
 
-### Connectivity
+## Connectivity
 

### 2023-04-24 README.Rmd
Tighten up prose around choice of Manchester

@@
 infrastructure that is based on motivators and deterrents to cycling. In this paper we 
-propose an approach for planning a connected cycling network that is meant to 
-be applicable to any city (as shown in the Appendix), and we use the city of Manchester as a case study. Greater 
-Manchester aims to become the first region to reach a carbon neutral transport network; 
-its local authorities have developed plans for a 1800 mile walking and cycling network 
-(Bee Network) that will be delivered over 10 years [@manchestercyclingbee2020]. While schemes 
+propose an approach for planning connected cycling networks that can be generalised to any city with appropriate input data (as shown in the Appendix).
+We used the city of Manchester as a case study due to its ambitions to become the first UK region with a carbon neutral transport network, supported by ambitious plans for a 1800 mile walking and cycling network
+(the 'Bee Network') that will be delivered over 10 years [@manchestercyclingbee2020]. While schemes 
 have already been put in place for around a quarter of this network, the majority is yet to be built.

### 2023-04-24 README.Rmd
Minor tweaks

@@
 
-Much research has been done to understand what gets people to cycle.
-Segregated cycling infrastructure^[*Segregated cycling infrastructure*
-refers to road space that is allocated to cyclists only, with physical
-separation to protect cyclists from other modes of transport.] has been
-shown to increase cycling uptake [@aldred2019impacts; @goodman2014new;
+Much research has expores the question of what gets people to cycle.
+Segregated cycling infrastructure, which refers to road space that is
+allocated to cyclists only with physical separation to protect cyclists,
+is associated with cycling uptake [@aldred2019impacts; @goodman2014new;
 @marques2015infrastructure], with the separation from motorized

### 2023-04-24 README.Rmd
Shorten and simplify prose

@@
 Safeguarding against increased car traffic may be counter-productive if the goal is to create a mode-shift, as research has shown that reducing road space for cars leads to less cars on the road, a phenomenon referred to as "traffic evaporation" [@nello2020environmental].
-On the other hand, @mauttone2017bicycle developed an optimization framework aiming at minimizing the total
+
+@mauttone2017bicycle developed an optimization framework aiming at minimizing the total
 user cost of cycling on the network without considering car usage. The aggregate flow^[*flow* is
@@
 penalty to prioritize connected road segments [@mauttone2017bicycle].
-
 In this paper, continuity is analysed by looking at the connectivity of
@@
 filtering to identify the minimum flow at which the whole city is
-connected by a giant component. While these approaches deal with
+connected by a single component.
+
+While these approaches deal with
 continuity better, they look at the network as a whole when attempting

### 2023-04-24 README.Rmd
Try sentence breaks for easier editing

@@
   markdown:
-    wrap: 72
+    wrap: sentence
 # elsevier style for referencing (from https://github.com/citation-style-language/styles)
@@
 
-The 2015 Paris agreement [@agreement2015paris] acknowledged that
-fundamental changes to societies and economies are necessary to mitigate
-climate change. Like other sectors, transport is under substantial
-pressure to decarbonise, resulting in a number of technical innovations
-including electric vehicles. But new vehicle technologies can only go so
-far and fail to tackle parallel problems such as congestion, road traffic
-casualties and physical inactivity [@brand_climate_2020].
+The 2015 Paris agreement [@agreement2015paris] acknowledged that fundamental changes to societies and economies are necessary to mitigate climate change.
+Like other sectors, transport is under substantial pressure to decarbonise, resulting in a number of technical innovations including electric vehicles.
+But new vehicle technologies can only go so far and fail to tackle parallel problems such as congestion, road traffic casualties and physical inactivity [@brand_climate_2020].
 
-In this context, interest and investment in active modes has grown substantially, especially with the growing realisation that the benefits extend beyond congestion and the environment, as active
-travel can also help alleviate what is referred to as the pandemic of
-global inactivity; physical inactivity is on the rise and has become the
-4th highest cause of death globally [@kohl2012pandemic]. Various
-studies have documented the association between active transport and
-lower risk of disease, including cancer and cardiovascular disease
-[@celis2017association; @jarrett2012effect; @patterson2020associations]. In the wake of the Covid-19 pandemic, and
-the resulting reduced capacity of public transport, the UK government
-has pledged to invest billions of pounds to improve walking and cycling
-infrastructure across the country. While this unprecedented sum is an
-opportunity to reshape cities in a way that improves the well-being of
-citizens, it does come with a warning:
+In this context, interest and investment in active modes has grown substantially, especially with the growing realisation that the benefits extend beyond congestion and the environment, as active travel can also help alleviate what is referred to as the pandemic of global inactivity; physical inactivity is on the rise and has become the 4th highest cause of death globally [@kohl2012pandemic].
+Various studies have documented the association between active transport and lower risk of disease, including cancer and cardiovascular disease [@celis2017association; @jarrett2012effect; @patterson2020associations].
+In the wake of the Covid-19 pandemic, and the resulting reduced capacity of public transport, the UK government has pledged to invest billions of pounds to improve walking and cycling infrastructure across the country.
+While this unprecedented sum is an opportunity to reshape cities in a way that improves the well-being of citizens, it does come with a warning:
 
-> "Inadequate cycling infrastructure discourages cycling and wastes
- public money. Much cycling infrastructure in this country is
- inadequate. It reflects a belief, conscious or otherwise, that hardly
- anyone cycles, that cycling is unimportant and that cycles must take
- no meaningful space from more important road users, such as motor
- vehicles and pedestrians" [@departmentgearchange2020].
+> "Inadequate cycling infrastructure discourages cycling and wastes public money.
+Much cycling infrastructure in this country is inadequate.
+It reflects a belief, conscious or otherwise, that hardly anyone cycles, that cycling is unimportant and that cycles must take no meaningful space from more important road users, such as motor vehicles and pedestrians" [@departmentgearchange2020].
 
-Funding alone is therefore no guarantee of a change in
-commuting across the country; it must be used to design adequate cycling
-infrastructure that is based on motivators and deterrents to cycling. In this paper we 
-propose an approach for planning connected cycling networks that can be generalised to any city with appropriate input data (as shown in the Appendix).
-We used the city of Manchester as a case study due to its ambitions to become the first UK region with a carbon neutral transport network, supported by ambitious plans for a 1800 mile walking and cycling network
-(the 'Bee Network') that will be delivered over 10 years [@manchestercyclingbee2020]. While schemes 
-have already been put in place for around a quarter of this network, the majority is yet to be built.
+Funding alone is therefore no guarantee of a change in commuting across the country; it must be used to design adequate cycling infrastructure that is based on motivators and deterrents to cycling.
+In this paper we propose an approach for planning connected cycling networks that can be generalised to any city with appropriate input data (as shown in the Appendix).
+We used the city of Manchester as a case study due to its ambitions to become the first UK region with a carbon neutral transport network, supported by ambitious plans for a 1800 mile walking and cycling network (the 'Bee Network') that will be delivered over 10 years [@manchestercyclingbee2020].
+While schemes have already been put in place for around a quarter of this network, the majority is yet to be built.
 
@@
 Much research has expores the question of what gets people to cycle.
-Segregated cycling infrastructure, which refers to road space that is
-allocated to cyclists only with physical separation to protect cyclists,
-is associated with cycling uptake [@aldred2019impacts; @goodman2014new;
-@marques2015infrastructure], with the separation from motorized
-vehicles being key [@winters2011motivators]. Revealed preference of
-cyclists shows that they are willing to deviate from the most efficient
-routes in order to commute on safer roads [@crane2017longitudinal].
-However, such deviations are only considered if they do not considerably
-increase route circuitry; behaviour studies have found that the
-probability of choosing a route decreases in proportion to its length
-relative to the shortest route [@broach2011bicycle; @winters2010far].
-Another defining feature for cycling infrastructure is how well
-connected it is. Cyclists prefer cohesive infrastructure, particularly
-when cycling on arterial roads with high levels of motorized traffic
-[@stinson2003commuter]. The lack of well-connected cycling
-infrastructure is one of the main obstacles to increasing cycling uptake
-[@caulfield2012determining]. While direct and cohesive cycling networks
-have been shown to positively impact cycling rates, density^[making an
-area's bicycle network denser means adding more cycling routes in the
-area and thereby giving cyclists more route options.] of the cycling
-network is also vital [@schoner2014missing].
+Segregated cycling infrastructure, which refers to road space that is allocated to cyclists only with physical separation to protect cyclists, is associated with cycling uptake [@aldred2019impacts; @goodman2014new; @marques2015infrastructure], with the separation from motorized vehicles being key [@winters2011motivators].
+Revealed preference of cyclists shows that they are willing to deviate from the most efficient routes in order to commute on safer roads [@crane2017longitudinal].
+However, such deviations are only considered if they do not considerably increase route circuitry; behaviour studies have found that the probability of choosing a route decreases in proportion to its length relative to the shortest route [@broach2011bicycle; @winters2010far].
+Another defining feature for cycling infrastructure is how well connected it is.
+Cyclists prefer cohesive infrastructure, particularly when cycling on arterial roads with high levels of motorized traffic [@stinson2003commuter].
+The lack of well-connected cycling infrastructure is one of the main obstacles to increasing cycling uptake [@caulfield2012determining].
+While direct and cohesive cycling networks have been shown to positively impact cycling rates, density[^1] of the cycling network is also vital [@schoner2014missing].
+
+[^1]: making an area's bicycle network denser means adding more cycling routes in the area and thereby giving cyclists more route options.
 
@@
 
-The studies outlined above lay out the fundamentals for designing
-cycling networks that generate significant cycling uptake, but they do
-not propose network-level interventions. In this section we outline
-methods used in past studies, namely optimization and network
-analysis methods, such as connected components and community
-detection, and examine how they are leveraged to suggest cycling network
-designs. We compare the effectiveness of these network-level studies in
-incorporating the fundamentals outlined above. 
-Our proposed approach is inspired by these methods, but it attempts to add to them
-by ensuring that all of the outlined fundamentals are accounted for. It
-also goes further by attempting to factor in ethical considerations
-related to the distribution of investment.
+The studies outlined above lay out the fundamentals for designing cycling networks that generate significant cycling uptake, but they do not propose network-level interventions.
+In this section we outline methods used in past studies, namely optimization and network analysis methods, such as connected components and community detection, and examine how they are leveraged to suggest cycling network designs.
+We compare the effectiveness of these network-level studies in incorporating the fundamentals outlined above.
+Our proposed approach is inspired by these methods, but it attempts to add to them by ensuring that all of the outlined fundamentals are accounted for.
+It also goes further by attempting to factor in ethical considerations related to the distribution of investment.
 
-*Optimization* methods have been used to propose improvements to
-cycling networks. @mesbah2012bilevel propose a bi-level formulation to
-optimize allocation of cycling lanes to the network without exceeding a
-set budget. They account for the effect of cycling lanes on car traffic,
-and attempt to maximize utilization of said lanes with minimal impact on
-car travel times.
+*Optimization* methods have been used to propose improvements to cycling networks.
+@mesbah2012bilevel propose a bi-level formulation to optimize allocation of cycling lanes to the network without exceeding a set budget.
+They account for the effect of cycling lanes on car traffic, and attempt to maximize utilization of said lanes with minimal impact on car travel times.
 Safeguarding against increased car traffic may be counter-productive if the goal is to create a mode-shift, as research has shown that reducing road space for cars leads to less cars on the road, a phenomenon referred to as "traffic evaporation" [@nello2020environmental].
 
-@mauttone2017bicycle developed an optimization framework aiming at minimizing the total
-user cost of cycling on the network without considering car usage. The aggregate flow^[*flow* is
-used throughout this research to refer to the cycling demand when it is
-routed onto the road network. The flow on any road segment is the
-cumulative demand on it, resulting from cyclists commuting between
-various Origin-Destination (OD) pairs.] on road segments (links) is obtained by using
-shortest paths to route existing cycling demand onto the road network. The solution is a proposed set of links where cycling infrastructure
-should be added in order to minimize the overall travel cost of cyclists
-across the network. The cost of traversing a link is given as a function
-of its length and whether or not it has cycling infrastructure. The
-problem has also been solved by attempting to find the minimum cost of
-improving roadway links to meet a desired level of service (LOS)
-[@duthie2014optimization]. In this formulation, all origin-destination
-(OD) pairs need to be connected by roads that meet the desired LOS. A
-limitation of these approaches is that they do not explicitly solve for
-continuity. Continuity is addressed using either a constraint specifying
-that each link with a bike lane should be connected to at least one
-destination [@mesbah2012bilevel], a constraint on maximum deviation
-from shortest paths [@duthie2014optimization], or a discontinuity
-penalty to prioritize connected road segments [@mauttone2017bicycle].
-In this paper, continuity is analysed by looking at the connectivity of
-the network through the graph-theoretic concept of *connected
-components*. @orozco2020data study the existing cycling network in terms of its disconnected
-components.
+@mauttone2017bicycle developed an optimization framework aiming at minimizing the total user cost of cycling on the network without considering car usage.
+The aggregate flow[^2] on road segments (links) is obtained by using shortest paths to route existing cycling demand onto the road network.
+The solution is a proposed set of links where cycling infrastructure should be added in order to minimize the overall travel cost of cyclists across the network.
+The cost of traversing a link is given as a function of its length and whether or not it has cycling infrastructure.
+The problem has also been solved by attempting to find the minimum cost of improving roadway links to meet a desired level of service (LOS) [@duthie2014optimization].
+In this formulation, all origin-destination (OD) pairs need to be connected by roads that meet the desired LOS.
+A limitation of these approaches is that they do not explicitly solve for continuity.
+Continuity is addressed using either a constraint specifying that each link with a bike lane should be connected to at least one destination [@mesbah2012bilevel], a constraint on maximum deviation from shortest paths [@duthie2014optimization], or a discontinuity penalty to prioritize connected road segments [@mauttone2017bicycle].
+In this paper, continuity is analysed by looking at the connectivity of the network through the graph-theoretic concept of *connected components*.
+@orozco2020data study the existing cycling network in terms of its disconnected components.
 They propose two different algorithms to connect disconnected components; one connects the two largest components, and the other connects the largest component to its nearest neighbor.
 By measuring the growth of the largest connected component as a function of the kilometers of network added, they determine that these approaches are more effective at improving cycling network connectivity than random allocation of cycling infrastructure.
-The concept of connected components is
-also at the core of the methodology proposed by @olmos2020data. After
-routing the cycling demand onto the network links, they use percolation
-theory to filter out the links based on the aggregate flow passing through them, varying the flow threshold for
-filtering to identify the minimum flow at which the whole city is
-connected by a single component.
+The concept of connected components is also at the core of the methodology proposed by @olmos2020data.
+After routing the cycling demand onto the network links, they use percolation theory to filter out the links based on the aggregate flow passing through them, varying the flow threshold for filtering to identify the minimum flow at which the whole city is connected by a single component.
 
-While these approaches deal with
-continuity better, they look at the network as a whole when attempting
-to improve it, and in doing so fail to account for equitable
-distribution of infrastructure.
+[^2]: *flow* is used throughout this research to refer to the cycling demand when it is routed onto the road network.
+    The flow on any road segment is the cumulative demand on it, resulting from cyclists commuting between various Origin-Destination (OD) pairs.
+
+While these approaches deal with continuity better, they look at the network as a whole when attempting to improve it, and in doing so fail to account for equitable distribution of infrastructure.
 
@@
 
-All of these network-level methodologies are underpinned by ethical
-principles, even though these principles are not explicitly acknowledged
-by the authors. This is important since different ethical principles
-constitute different problem formulations and targets. Broadly speaking,
-transport appraisal can be based on either utilitarian or egalitarian
-principles. The former seeks to maximize the overall benefit, while the
-latter is concerned with a fair distribution of benefits
-[@jafino2020transport]. 
-<!-- @nahmias2017integrating criticize the -->
-<!-- utilitarian approach that has been historically popular in the -->
-<!-- evaluation of transport investments, explaining how the maximization of -->
-<!-- overall benefit fails to account for the distribution of that benefit -->
-<!-- among communities or individuals. @lucas2016method explain how -->
-<!-- transport studies have traditionally looked at the bigger picture -->
-<!-- without studying the distribution of investments on the different parts -->
-<!-- of the study area, and go on to propose an egalitarian approach that -->
-<!-- ensures the disaggregation of transport policy benefits across the study -->
-<!-- area.  -->
-The utilitarian approach, historically popular in transport planning, has been criticised for focusing on the bigger picture and failing to account for the distribution of investments on the different communities of the study area [@lucas2016method; @nahmias2017integrating].
-@pereira2017distributive emphasize the need for a more
-egalitarian approach to transport planning. They highlight accessibility
-as a cornerstone of distributive justice, and contend that policies
-should aim to distribute investments in a way that minimizes spatial
-variations in accessibility.
+All of these network-level methodologies are underpinned by ethical principles, even though these principles are not explicitly acknowledged by the authors.
+This is important since different ethical principles constitute different problem formulations and targets.
+Broadly speaking, transport appraisal can be based on either utilitarian or egalitarian principles.
+The former seeks to maximize the overall benefit, while the latter is concerned with a fair distribution of benefits [@jafino2020transport].
+<!-- @nahmias2017integrating criticize the --> <!-- utilitarian approach that has been historically popular in the --> <!-- evaluation of transport investments, explaining how the maximization of --> <!-- overall benefit fails to account for the distribution of that benefit --> <!-- among communities or individuals.
+@lucas2016method explain how --> <!-- transport studies have traditionally looked at the bigger picture --> <!-- without studying the distribution of investments on the different parts --> <!-- of the study area, and go on to propose an egalitarian approach that --> <!-- ensures the disaggregation of transport policy benefits across the study --> <!-- area.
+--> The utilitarian approach, historically popular in transport planning, has been criticised for focusing on the bigger picture and failing to account for the distribution of investments on the different communities of the study area [@lucas2016method; @nahmias2017integrating].
+@pereira2017distributive emphasize the need for a more egalitarian approach to transport planning.
+They highlight accessibility as a cornerstone of distributive justice, and contend that policies should aim to distribute investments in a way that minimizes spatial variations in accessibility.
 
-This research attempts to propose an egalitarian framework for cycling
-network design. This is done by identifying the different sub-networks
-that exist within the larger network, and ensuring that each gets a fair
-share of investment. Trip patterns in a city are not uniformly
-distributed geographically, and *community finding* methods have been
-used to partition study areas into localized areas that experience a
-disproportionate number of trips within them. @akbarzadeh2018designing
-use a modularity maximization approach [@blondel2008fast] on taxi trip
-data to identify seven different communities in the city of Isfahan, Iran.
-An optimization problem is then formulated to connect nodes within each
-community with cycling infrastructure, with the emphasis being on
-connectivity within the communities, not between them. @bao2017planning
-adopt a similar methodology, first identifying communities and then
-using a greedy network expansion algorithm to simultaneously add links
-to each community. The link with the highest benefit-cost ratio in each
-community is selected, and the network is grown by adding neighboring
-links to the solution until a budget limit is met. The benefit is the
-flow on the link, and each link is assigned a cost based on current road
-conditions.
+This research attempts to propose an egalitarian framework for cycling network design.
+This is done by identifying the different sub-networks that exist within the larger network, and ensuring that each gets a fair share of investment.
+Trip patterns in a city are not uniformly distributed geographically, and *community finding* methods have been used to partition study areas into localized areas that experience a disproportionate number of trips within them.
+@akbarzadeh2018designing use a modularity maximization approach [@blondel2008fast] on taxi trip data to identify seven different communities in the city of Isfahan, Iran.
+An optimization problem is then formulated to connect nodes within each community with cycling infrastructure, with the emphasis being on connectivity within the communities, not between them.
+@bao2017planning adopt a similar methodology, first identifying communities and then using a greedy network expansion algorithm to simultaneously add links to each community.
+The link with the highest benefit-cost ratio in each community is selected, and the network is grown by adding neighboring links to the solution until a budget limit is met.
+The benefit is the flow on the link, and each link is assigned a cost based on current road conditions.
 
@@
 
-Our work builds on these community finding approaches by proposing a
-similar greedy network expansion algorithm for cycle network expansion
-*within communities*. We incorporate community finding methods for study
-area partitioning with weighted routing to avoid links that are
-stressful to cycle on. In doing so, we propose an approach that ensures a better distribution 
-investments across the study area while also accounting for motivators and deterrents to cycling. 
-We propose three sub-methods that address some of the limitations of previous
-studies. These limitations include (a) an inherent bias when basing network
-design solely on existing cycling demand, (b) the proposal of routes that may
-not correspond to studies on cyclist preference and government policies,
-and (c) an insufficient consideration of the ethical principles underlying
-the analysis. The current work addresses these limitations through the following organisation. Section \@ref(calculating-potential-cycling-demand)
-focuses on calculating potential cycling demand. Section \@ref(routing)
-focuses on routing the demand onto the road network while accounting for
-cyclist preferences and government priorities. Section
-\@ref(road-segment-prioritization) outlines a method for partitioning the
-study area based on a community finding algorithm and routed cycling
-demand. It then introduces the network expansion algorithms, and compares an approach grounded in
-'egalitarianism' to one grounded in 'utilitarianism'.
+Our work builds on these community finding approaches by proposing a similar greedy network expansion algorithm for cycle network expansion *within communities*.
+We incorporate community finding methods for study area partitioning with weighted routing to avoid links that are stressful to cycle on.
+In doing so, we propose an approach that ensures a better distribution investments across the study area while also accounting for motivators and deterrents to cycling.
+We propose three sub-methods that address some of the limitations of previous studies.
+These limitations include (a) an inherent bias when basing network design solely on existing cycling demand, (b) the proposal of routes that may not correspond to studies on cyclist preference and government policies, and (c) an insufficient consideration of the ethical principles underlying the analysis.
+The current work addresses these limitations through the following organisation.
+Section \@ref(calculating-potential-cycling-demand) focuses on calculating potential cycling demand.
+Section \@ref(routing) focuses on routing the demand onto the road network while accounting for cyclist preferences and government priorities.
+Section \@ref(road-segment-prioritization) outlines a method for partitioning the study area based on a community finding algorithm and routed cycling demand.
+It then introduces the network expansion algorithms, and compares an approach grounded in 'egalitarianism' to one grounded in 'utilitarianism'.
 
@@
 Some use existing cycling demand, some calculate potential cycling demand, and others ignore demand completely.
-@duthie2014optimization note that relying only on existing cycling activity to prioritize cycling
-infrastructure can reinforce existing cycling patterns and ignore potential cycling demand that could be satisfied by a connected network.
-To avoid this issue they choose to ignore existing demand completely, and focus on creating a network that connects the
-entire study area. 
-@olmos2020data opt to calculate potential demand instead; they obtain the distance distribution of cyclists using a smartphone-based bicycle GPS data, and then use a rejection-sampling algorithm on the OD data of the study area to match
-the potential demand distribution to the distribution obtained from GPS data.
+@duthie2014optimization note that relying only on existing cycling activity to prioritize cycling infrastructure can reinforce existing cycling patterns and ignore potential cycling demand that could be satisfied by a connected network.
+To avoid this issue they choose to ignore existing demand completely, and focus on creating a network that connects the entire study area.
+@olmos2020data opt to calculate potential demand instead; they obtain the distance distribution of cyclists using a smartphone-based bicycle GPS data, and then use a rejection-sampling algorithm on the OD data of the study area to match the potential demand distribution to the distribution obtained from GPS data.
 
-*OD* data can be obtained from a range of sources,
-including GPS data, household travels surveys and census data on work
-locations. In areas where observed OD data is unavailable, modelling
-techniques such as spatial interaction models can be used to estimate
-travel volumes between zones [@black1995spatial; @martinez2013new;
-@wilson1971family]. In this paper we use open access data from the UK
-census [@ONS2011flowdata], which contains aggregate statistics on
-number of commuters between administrative zones --- Middle layer Super
-Output Areas (MSOA) --- by mode of travel. MSOAs have an average
-population of just over 8000 [@ofn2018population].^[ See
-<https://wicid.ukdataservice.ac.uk/> for open access to the OD data.]
-Figures \@ref(fig:potdemhistograms) and \@ref(fig:desirefacetcycling)
-illustrate the proportion of trips cycled by distance and the
-geographic extent of the input OD dataset used in this paper.
+*OD* data can be obtained from a range of sources, including GPS data, household travels surveys and census data on work locations.
+In areas where observed OD data is unavailable, modelling techniques such as spatial interaction models can be used to estimate travel volumes between zones [@black1995spatial; @martinez2013new; @wilson1971family].
+In this paper we use open access data from the UK census [@ONS2011flowdata], which contains aggregate statistics on number of commuters between administrative zones --- Middle layer Super Output Areas (MSOA) --- by mode of travel.
+MSOAs have an average population of just over 8000 [@ofn2018population].[^3]
+Figures \@ref(fig:potdemhistograms) and \@ref(fig:desirefacetcycling) illustrate the proportion of trips cycled by distance and the geographic extent of the input OD dataset used in this paper.
 
-For our purposes, we use a logistic regression model to calculate
-potential cycling demand. The model was adapted from the Propensity to
-Cycle Tool (PCT), which estimates the proportion of trips
-($\boldsymbol{C_{p}}$) for each OD pair that would cycle under
-different scenarios of change as a
-function of distance and hilliness [@lovelace2017propensity]. We used the Government Target scenario,
-indicating a nationwide target of doubling cycling by 2025. The
-logistic regression model used to calculate $\boldsymbol{C_{p}}$ has
-the following parameters:
+[^3]: See <https://wicid.ukdataservice.ac.uk/> for open access to the OD data.
+
+For our purposes, we use a logistic regression model to calculate potential cycling demand.
+The model was adapted from the Propensity to Cycle Tool (PCT), which estimates the proportion of trips ($\boldsymbol{C_{p}}$) for each OD pair that would cycle under different scenarios of change as a function of distance and hilliness [@lovelace2017propensity].
+We used the Government Target scenario, indicating a nationwide target of doubling cycling by 2025.
+The logistic regression model used to calculate $\boldsymbol{C_{p}}$ has the following parameters:
 
@@
 ```
+\noindent where $d$ and $s$ are the distance and slope associated with each OD pair, and $\alpha$ and all $\beta$s are parameters calculated by a regression model on the training data.
+The square and square-root distance terms "capture the non-linear impact of distance on the likelihood of cycling", and interaction terms capture the combined effect of slope and distance [@lovelace2017propensity].
+Alternative cycling uptake models could be 'plugged in' to our approach for different contexts or scenarios of change.
 
-\noindent where $d$ and $s$ are the distance and slope associated with each
-OD pair, and $\alpha$ and all $\beta$s are parameters calculated by a regression
-model on the training data. The square and square-root distance terms
-"capture the non-linear impact of distance on the likelihood of
-cycling", and interaction terms capture the combined effect of slope
-and distance [@lovelace2017propensity]. Alternative cycling uptake
-models could be 'plugged in' to our approach for different contexts or scenarios of change.
-
-The potential demand calculations show that the current and potential
-number of cyclists both follow a bell-shaped distribution, with the
-number of trips peaking around the 3-4km commuting distance and then
-going back down for longer distances (see Figures
-\@ref(fig:potdemhistograms) and \@ref(fig:desirefacetcycling)).
+The potential demand calculations show that the current and potential number of cyclists both follow a bell-shaped distribution, with the number of trips peaking around the 3-4km commuting distance and then going back down for longer distances (see Figures \@ref(fig:potdemhistograms) and \@ref(fig:desirefacetcycling)).
 
@@
 <!-- ```{r potdemhistograms, fig.cap="Distribution of potential cycling demand", out.width="30%", fig.show='hold', echo=FALSE} -->
+
 <!-- knitr::include_graphics(c( -->
+
 <!--   "data/Manchester/Plots/histogram_distance_all_vs_cycling.png", -->
+
 <!--   "data/Manchester/Plots/histogram_distance_all_vs_cycling_potential.png", -->
+
 <!--   "data/Manchester/Plots/histogram_distance_cycling_potential_vs_current.png" -->
+
 <!-- )) -->
+
 <!-- ``` -->
@@
 
-The uptake model used in this paper assumes uptake of traditional bicycles,
-which are affected by topography and distance due to physical effort.
-As discussed in Section \@ref(discussion-and-conclusions), using adapted uptake
-models could enable the approach to help plan for solutions such as e-bikes
-that enable trips covering longer distances and hillier roads.
+The uptake model used in this paper assumes uptake of traditional bicycles, which are affected by topography and distance due to physical effort.
+As discussed in Section \@ref(discussion-and-conclusions), using adapted uptake models could enable the approach to help plan for solutions such as e-bikes that enable trips covering longer distances and hillier roads.
 
@@
 
-The next step in our approach is to route the potential cycling demand
-($\boldsymbol{C_{p}}$) between all OD pairs onto the road network. We
-choose not to use the PCT approach for routing as it is reliant on
-external and therefore inflexible routing services.
+The next step in our approach is to route the potential cycling demand ($\boldsymbol{C_{p}}$) between all OD pairs onto the road network.
+We choose not to use the PCT approach for routing as it is reliant on external and therefore inflexible routing services.
 <!-- This expands on the work of @mauttone2017bicycle, by going beyond simply favoring roads with existing cycling infrastructure to creating a hierarchy of road preference. -->
 
+```{=tex}
 \begin{longtable}{lll}
@@
 \end{longtable}
-
+```
 To conduct routing, the following is considered:
 
-1.  **Cyclist Preference**: Work done by @dill2013four on examining
-    cyclist typologies determined that around 60% of Portland residents
-    fit under the *interested but concerned* category. These were people
-    that enjoyed cycling but avoided it due safety concerns. The key to
-    encouraging this group was to create a low-stress cycling network,
-    not only though segregated infrastructure but also by planning
-    routes that passed through residential streets.
-2.  **Low-Traffic Neighbourhoods (LTNs)**: The UK Department for Transport is
-    allocating funding to local authorities to invest in Active
-    Transport, partially through the creation of LTNs
-    [@departmentgearchange2020]. This includes closing off residential
-    streets to motorized traffic.
-3.  **Existing Cycling Infrastructure**: Utilizing existing cycling
-    infrastructure makes economic sense, as small investments may lead
-    to large connectivity gains as the disconnected cycling
-    infrastructure gets joined together.
+1.
+**Cyclist Preference**: Work done by @dill2013four on examining cyclist typologies determined that around 60% of Portland residents fit under the *interested but concerned* category.
+These were people that enjoyed cycling but avoided it due safety concerns.
+The key to encouraging this group was to create a low-stress cycling network, not only though segregated infrastructure but also by planning routes that passed through residential streets.
+2.
+**Low-Traffic Neighbourhoods (LTNs)**: The UK Department for Transport is allocating funding to local authorities to invest in Active Transport, partially through the creation of LTNs [@departmentgearchange2020].
+This includes closing off residential streets to motorized traffic.
+3.
+**Existing Cycling Infrastructure**: Utilizing existing cycling infrastructure makes economic sense, as small investments may lead to large connectivity gains as the disconnected cycling infrastructure gets joined together.
 
-The above points are accounted for by using a weighted road network for
-routing. This has been previously done by multiplying all road segments
-without cycling infrastructure by a fixed impedance factor
-[@mauttone2017bicycle], or by assigning different weights to the road
-segments proportional to the investment cost of bringing them to an
-acceptable level of stress for cycling [@duthie2014optimization].
-Perceived stress for cyclists has been found to be a function of vehicular traffic
-volume and speed [@sorton1994bicycle], both of which vary predictably
-with road type. @gehrke2020cycling build on this information to use
-routing impedance factors for road segments that are based on road type
-and the existence of cycling infrastructure.
+The above points are accounted for by using a weighted road network for routing.
+This has been previously done by multiplying all road segments without cycling infrastructure by a fixed impedance factor [@mauttone2017bicycle], or by assigning different weights to the road segments proportional to the investment cost of bringing them to an acceptable level of stress for cycling [@duthie2014optimization].
+Perceived stress for cyclists has been found to be a function of vehicular traffic volume and speed [@sorton1994bicycle], both of which vary predictably with road type.
+@gehrke2020cycling build on this information to use routing impedance factors for road segments that are based on road type and the existence of cycling infrastructure.
 
-For the purposes of this paper we created a similar weighting profile
-that is adjusted to favor less stressful roads (based on the definition
-by @sorton1994bicycle and information from Table
-\@ref(table:osmroadtypes)) and roads with existing cycling
-infrastructure. We believe this to be more appropriate than the simplistic
-approach adopted by @mauttone2017bicycle as it makes use of a hierarchy
-of road preference based on perceived stress levels, going beyond simply
-favouring roads with existing cycling infrastructure.
+For the purposes of this paper we created a similar weighting profile that is adjusted to favor less stressful roads (based on the definition by @sorton1994bicycle and information from Table \@ref(table:osmroadtypes)) and roads with existing cycling infrastructure.
+We believe this to be more appropriate than the simplistic approach adopted by @mauttone2017bicycle as it makes use of a hierarchy of road preference based on perceived stress levels, going beyond simply favouring roads with existing cycling infrastructure.
 
@@
 
+```{=tex}
 \begin{longtable}{lrrr}
@@
 \end{longtable}
+```
+\noindent A weighted distance $d_{w}$ for each road segment is calculated as follows [^4]:
 
-\noindent A weighted distance $d_{w}$ for each road
-segment is calculated as follows ^[The **dodgr** r package
-[@padgham2019dodgr] is used to route cycling demand onto the road
-network. The package uses the OpenStreetMaps (OSM) road network and
-allows the user to assign weights to roads based on their type. The
-routing is done based on weighted shortest paths, with the distance
-along each road segment being divided by a factor to obtain the weighted
-distance for routing. It is more intuitive to multiply when weighting a
-network, but the dodgr package divides by numbers between 0 and 1, which
-achieves the same result (road types with a weight of 0 are not routed on). 
-For the sake of reproducibility, we stick to
-the convention used in the package.]:
+[^4]: The **dodgr** r package [@padgham2019dodgr] is used to route cycling demand onto the road network.
+    The package uses the OpenStreetMaps (OSM) road network and allows the user to assign weights to roads based on their type.
+    The routing is done based on weighted shortest paths, with the distance along each road segment being divided by a factor to obtain the weighted distance for routing.
+    It is more intuitive to multiply when weighting a network, but the dodgr package divides by numbers between 0 and 1, which achieves the same result (road types with a weight of 0 are not routed on).
+    For the sake of reproducibility, we stick to the convention used in the package.
 
+```{=tex}
 \begin{equation}\label{eq:weight_distance}
@@
 \end{equation}
+```
+\noindent where $d_{unw}$ is the unweighted distance and $W$ is the weight from Table \ref{table:weightprofiles}.
 
-\noindent where $d_{unw}$ is the unweighted distance and
-$W$ is the weight from Table \ref{table:weightprofiles}.
+All weights are between 0 and 1, and the values in the \textit{Weighted} profile are chosen to be inversely proportional to the stress level experienced by cyclists on them.
+The \textit{Unweighted} weighting profile is used to compare increases in route length resulting from two different approaches:
 
-All weights are between 0 and 1, and the values in the
-\textit{Weighted} profile are chosen to be inversely proportional to
-the stress level experienced by cyclists on them. The
-\textit{Unweighted} weighting profile is used to compare increases in
-route length resulting from two different approaches:
-
-1.  **Weighted**: Relatively high impedance on Primary and Trunk roads
-    (to minimize cycling on them).
+1.  **Weighted**: Relatively high impedance on Primary and Trunk roads (to minimize cycling on them).
 2.  **Weighted_2**: Avoiding Primary and Trunk Roads completely.
 
-Comparing the cycling demand routed on the weighted and unweighted road
-network allows us to get a better understanding of the importance of
-different road types. In the case of Manchester, trunk roads bisect the
-city and are a major part of unweighted shortest paths (Figure
-\@ref(fig:flowsfacetunweighted)). On the other hand, cycleways are not
-part of unweighted shortest paths, and so very little of the cycling
-demand is routed through them. In the weighted network, cycleways are
-much better utilized, and the majority of the cycling demand passes
-through tertiary roads, as expected.
+Comparing the cycling demand routed on the weighted and unweighted road network allows us to get a better understanding of the importance of different road types.
+In the case of Manchester, trunk roads bisect the city and are a major part of unweighted shortest paths (Figure \@ref(fig:flowsfacetunweighted)).
+On the other hand, cycleways are not part of unweighted shortest paths, and so very little of the cycling demand is routed through them.
+In the weighted network, cycleways are much better utilized, and the majority of the cycling demand passes through tertiary roads, as expected.
 
@@
 
-The results of routing potential cycling demand on the weighted and
-unweighted networks are understandably quite different. From Figure
-\@ref(fig:flowsfacetunweighted) we can see that trunk and primary roads
-are the most efficient means of traversing the road network of
-Manchester. Both of these road types are classified as Primary A roads
-according to the UK Department for Transport's road classification
-(Table \@ref(table:osmroadtypes)), and are therefore part of the Primary
-Route Network (PRN) [@department2012guidance]. The PRN has the widest,
-most direct roads on the network, and carries most of the through
-traffic. This includes freight, with all roads in the PRN being required
-by law to provide unrestricted access to trucks up to 40 tonnes
-[@department2012guidance].
+The results of routing potential cycling demand on the weighted and unweighted networks are understandably quite different.
+From Figure \@ref(fig:flowsfacetunweighted) we can see that trunk and primary roads are the most efficient means of traversing the road network of Manchester.
+Both of these road types are classified as Primary A roads according to the UK Department for Transport's road classification (Table \@ref(table:osmroadtypes)), and are therefore part of the Primary Route Network (PRN) [@department2012guidance].
+The PRN has the widest, most direct roads on the network, and carries most of the through traffic.
+This includes freight, with all roads in the PRN being required by law to provide unrestricted access to trucks up to 40 tonnes [@department2012guidance].
 
-We choose to avoid routing the potential cycling demand on Primary A
-Roads for the following 2 reasons:
+We choose to avoid routing the potential cycling demand on Primary A Roads for the following 2 reasons:
 
-1.  **Logistical Difficulty**: Changes on these roads need to be agreed
-    upon by all affected authorities [@department2012guidance], which
-    may prove to be difficult.
-2.  **Low Traffic Neighborhoods (LTNs)**: The UK government is aiming to
-    restrict access to motorized vehicles on residential roads to create
-    LTNs [@departmentgearchange2020]. This is part of a policy to
-    prevent automobile rat-running and make streets more accessible to
-    cyclists and pedestrians. Under such a policy, Primary A roads would
-    become even more essential for motorized traffic and it would be
-    more difficult to reallocate road space on these roads to cyclists.
+1.  **Logistical Difficulty**: Changes on these roads need to be agreed upon by all affected authorities [@department2012guidance], which may prove to be difficult.
+2.
+**Low Traffic Neighborhoods (LTNs)**: The UK government is aiming to restrict access to motorized vehicles on residential roads to create LTNs [@departmentgearchange2020].
+This is part of a policy to prevent automobile rat-running and make streets more accessible to cyclists and pedestrians.
+Under such a policy, Primary A roads would become even more essential for motorized traffic and it would be more difficult to reallocate road space on these roads to cyclists.
 
-Figure \@ref(fig:flowsfacetweighted) shows that routing on the weighted
-network significantly reduces flow on the trunk and primary roads, but
-does not eliminate it completely. This is intentional, as the impedance
-on these roads is only slightly higher than the remaining road types (See
-Table \@ref(table:weightprofiles)). Potential cycling demand is only
-routed on these roads if there are no routes through other roads that
-offer comparable directness.
+Figure \@ref(fig:flowsfacetweighted) shows that routing on the weighted network significantly reduces flow on the trunk and primary roads, but does not eliminate it completely.
+This is intentional, as the impedance on these roads is only slightly higher than the remaining road types (See Table \@ref(table:weightprofiles)).
+Potential cycling demand is only routed on these roads if there are no routes through other roads that offer comparable directness.
 
-Banning cycling flow completely on trunk and primary roads may result in
-excessively circuitous paths, as seen in Figure
-\@ref(fig:boxplotcircuity). When routing using the *weighted* weighting
-profile in Table \@ref(table:weightprofiles), we see that shortest paths
-increase by less than 5% on average from unweighted shortest paths, with
-the largest increases still below 30%. When routing on primary and trunk
-roads is banned (*weighted_2* profile in Table
-\@ref(table:weightprofiles)), the average increase relative to
-unweighted shortest paths rises to 10%, with certain locations
-experiencing more significant negative effects on accessibility. Given
-that cyclists will only deviate from shortest paths by a certain amount
-to access better cycling infrastructure (as explained in Section
-\@ref(introduction)), allowing flow on some stretches of trunk and
-primary roads is necessary to insure cycling uptake and equitable access
-to cycling infrastructure. In its new vision for walking and cycling,
-the Department for Transport acknowledges that minimal segregated
-stretches of bicycle lanes on main roads will be necessary to avoid
-circuitous cycling networks [@departmentgearchange2020].
+Banning cycling flow completely on trunk and primary roads may result in excessively circuitous paths, as seen in Figure \@ref(fig:boxplotcircuity).
+When routing using the *weighted* weighting profile in Table \@ref(table:weightprofiles), we see that shortest paths increase by less than 5% on average from unweighted shortest paths, with the largest increases still below 30%.
+When routing on primary and trunk roads is banned (*weighted_2* profile in Table \@ref(table:weightprofiles)), the average increase relative to unweighted shortest paths rises to 10%, with certain locations experiencing more significant negative effects on accessibility.
+Given that cyclists will only deviate from shortest paths by a certain amount to access better cycling infrastructure (as explained in Section \@ref(introduction)), allowing flow on some stretches of trunk and primary roads is necessary to insure cycling uptake and equitable access to cycling infrastructure.
+In its new vision for walking and cycling, the Department for Transport acknowledges that minimal segregated stretches of bicycle lanes on main roads will be necessary to avoid circuitous cycling networks [@departmentgearchange2020].
 
@@
 
-Weighting the road network also allows us to better utilize existing
-cycling infrastructure, as can be seen by the higher flow on cycleways
-in Figure \@ref(fig:flowsfacetweighted). Again, the small differences in
-impedance between cycleways and other road types mean that cycleways
-that require significant deviation are not routed on.
+Weighting the road network also allows us to better utilize existing cycling infrastructure, as can be seen by the higher flow on cycleways in Figure \@ref(fig:flowsfacetweighted).
+Again, the small differences in impedance between cycleways and other road types mean that cycleways that require significant deviation are not routed on.
 
-It should be reiterated that the weighting profile used for routing has
-been developed for the purposes of this study. It creates a hierarchy of
-road preference that is grounded in cyclist preferences and government
-plans to create LTNs. A wide range of weighting profiles could be used
-to represent different types of cyclists and road environments, as
-described in @gehrke2020cycling and @furth2016network. A sensitivity
-analysis could be done to determine an optimal weighting profile, but
-given the variation in city road
-networks<!-- (Figure \ref{fig:perc_person-km}) -->, this would probably
-require calibration to the specific city. More accurate routing could be
-carried out given the availability of road-level data. In such cases we
-would add additional impedance to specific roads, giving more useful
-routing results than the current approach which considers all roads
-of the same type to be equivalent.
+It should be reiterated that the weighting profile used for routing has been developed for the purposes of this study.
+It creates a hierarchy of road preference that is grounded in cyclist preferences and government plans to create LTNs.
+A wide range of weighting profiles could be used to represent different types of cyclists and road environments, as described in @gehrke2020cycling and @furth2016network.
+A sensitivity analysis could be done to determine an optimal weighting profile, but given the variation in city road networks<!-- (Figure \ref{fig:perc_person-km}) -->, this would probably require calibration to the specific city.
+More accurate routing could be carried out given the availability of road-level data.
+In such cases we would add additional impedance to specific roads, giving more useful routing results than the current approach which considers all roads of the same type to be equivalent.
 
-One use-case of such granular data would be to identify roads that serve
-schools. The Department of Transport notes that the number of school
-children being driven to school has trebled over the past 40 years
-[@departmentgearchange2020], and so having cycling infrastructure
-serving schools is key to achieving the government target of getting
-more children to cycle. This would not be difficult, as over 75% of
-children in the UK live within a 15 minute cycle from their school
-[@departmentcycleinfradesign2020]. @goodman2019scenarios show that if
-dutch levels of cycling were achieved in the UK, the % of children
-cycling to school could increase from 1.8% to 41%. In their typology of
-cyclists, @dill2013four found that a majority of people who say they
-would never cycle had never cycled to school, whereas confident cyclists
-were those most likely to have cycled to school. Getting people to cycle
-from a young age is therefore key to achieving societal change in
-commuting habits.
+One use-case of such granular data would be to identify roads that serve schools.
+The Department of Transport notes that the number of school children being driven to school has trebled over the past 40 years [@departmentgearchange2020], and so having cycling infrastructure serving schools is key to achieving the government target of getting more children to cycle.
+This would not be difficult, as over 75% of children in the UK live within a 15 minute cycle from their school [@departmentcycleinfradesign2020].
+@goodman2019scenarios show that if dutch levels of cycling were achieved in the UK, the % of children cycling to school could increase from 1.8% to 41%.
+In their typology of cyclists, @dill2013four found that a majority of people who say they would never cycle had never cycled to school, whereas confident cyclists were those most likely to have cycled to school.
+Getting people to cycle from a young age is therefore key to achieving societal change in commuting habits.
 
@@
 
-After routing the potential cycling demand onto the road network using
-weighted shortest paths, we have estimates for the cumulative potential
-cycling demand passing through all road segments. This cumulative demand
-(referred to as *flow*) is then used as a basis for identifying segments
-that are in most need of investment in segregated cycling
-infrastructure. In doing so, we must account for the motivations and
-deterrents for cycling identified in Section \@ref(introduction), namely
-direct and well-connected routes.
+After routing the potential cycling demand onto the road network using weighted shortest paths, we have estimates for the cumulative potential cycling demand passing through all road segments.
+This cumulative demand (referred to as *flow*) is then used as a basis for identifying segments that are in most need of investment in segregated cycling infrastructure.
+In doing so, we must account for the motivations and deterrents for cycling identified in Section \@ref(introduction), namely direct and well-connected routes.
 
-A range of algorithms could be used for prioritization. Because policy
-priorities vary, we present two algorithms. Both utilize existing
-infrastructure from the beginning and allow us to compare a solution
-that focuses on utilitarianism to one that focuses on egalitarianism. In
-both algorithms, links are selected iteratively and the iteration at
-which each link is added to the solution is recorded. Investments in
-cycling infrastructure can be limited by budget constraints, so it can
-be useful to see where best to allocate a defined length of segregated
-infrastructure. In order to incorporate egalitarian principles in our
-approach, we use community detection to partition the study area and
-distribute investment over the different subdivisions.
+A range of algorithms could be used for prioritization.
+Because policy priorities vary, we present two algorithms.
+Both utilize existing infrastructure from the beginning and allow us to compare a solution that focuses on utilitarianism to one that focuses on egalitarianism.
+In both algorithms, links are selected iteratively and the iteration at which each link is added to the solution is recorded.
+Investments in cycling infrastructure can be limited by budget constraints, so it can be useful to see where best to allocate a defined length of segregated infrastructure.
+In order to incorporate egalitarian principles in our approach, we use community detection to partition the study area and distribute investment over the different subdivisions.
 
@@
 
-As explained in Section
-\@ref(ethical-underpinnings-and-proposed-approach), a major challenge
-facing 'top-down' planning approaches is how to incorporate egalitarian
-principles by fairly distributing investments in cycling infrastructure.
-One way of quantifying this is to split up the city into smaller
-geospatial areas and target equal investment in each of those areas.
-This approach could also help ensure that on-the-ground surveys are made
-by local stakeholders, an important component of the planning process
-[@parkin2018designing]. Community detection offers us a way to
-delineate such a split; cyclists are limited in their commuting distance
-(see Figure \@ref(fig:cyclinghistmanc)), and so trip attractors are more
-likely to have a local catchment area of cyclists.
+As explained in Section \@ref(ethical-underpinnings-and-proposed-approach), a major challenge facing 'top-down' planning approaches is how to incorporate egalitarian principles by fairly distributing investments in cycling infrastructure.
+One way of quantifying this is to split up the city into smaller geospatial areas and target equal investment in each of those areas.
+This approach could also help ensure that on-the-ground surveys are made by local stakeholders, an important component of the planning process [@parkin2018designing].
+Community detection offers us a way to delineate such a split; cyclists are limited in their commuting distance (see Figure \@ref(fig:cyclinghistmanc)), and so trip attractors are more likely to have a local catchment area of cyclists.
 
@@
 
-In our case, the network is the city; the nodes are the
-population-weighted MSOA centroids and the links connecting each MSOA
-pair are weighted by the potential cycling demand between them. The
-Louvain method [@blondel2008fast] is used to separate MSOAs into
-communities. Potential cycling demand is used since we assume that this
-is what the cycling demand will be once the cycling infrastructure is
-added. To assign road links to communities, the following steps are
-carried out:
+In our case, the network is the city; the nodes are the population-weighted MSOA centroids and the links connecting each MSOA pair are weighted by the potential cycling demand between them.
+The Louvain method [@blondel2008fast] is used to separate MSOAs into communities.
+Potential cycling demand is used since we assume that this is what the cycling demand will be once the cycling infrastructure is added.
+To assign road links to communities, the following steps are carried out:
 
+```{=html}
 <!-- 1. Create links between MSOA centroids and weigh these links by potential cycling demand between them.
@@
 3. Assign each road link to the same community as the closest MSOA centroid to it. -->
-
+```
 ```{r, eval = FALSE}
@@
 
-The results show that Manchester can be split into four large
-communities and one small one (Figure \@ref(fig:communitiesmanchester)).
+The results show that Manchester can be split into four large communities and one small one (Figure \@ref(fig:communitiesmanchester)).
 
@@
 ```
-\  
+
  <!-- this prevents text from floating above figure -->
@@
 
+```{=html}
 <!-- 1. Identify all links that have segregated cycling infrastructure and add them to the initial solution
@@
 4. Repeat steps 2 \& 3 until all flow is satisfied or investment threshold is met -->
-
+```
 ```{r, eval = FALSE, results='asis'}
@@
 
-This algorithm ensures that the resulting network is connected. It also
-satisfies the directness criteria, since links on the weighted shortest
-paths are those that have the highest flow passing through them (this is
-a result of the routing in Section \@ref(routing).
+This algorithm ensures that the resulting network is connected.
+It also satisfies the directness criteria, since links on the weighted shortest paths are those that have the highest flow passing through them (this is a result of the routing in Section \@ref(routing).
 
@@
 
-The first algorithm focuses on connectivity and directness, but not on
-fairly distributing investment. The latter is not a requirement for
-increasing cycling uptake, but it is fundamental for spatial equity, as
-explained in Section \@ref(ethical-underpinnings-and-proposed-approach). This algorithm incorporates
-the ideal of fair distribution by ensuring that investment is distributed between the defined communities. 
+The first algorithm focuses on connectivity and directness, but not on fairly distributing investment.
+The latter is not a requirement for increasing cycling uptake, but it is fundamental for spatial equity, as explained in Section \@ref(ethical-underpinnings-and-proposed-approach).
+This algorithm incorporates the ideal of fair distribution by ensuring that investment is distributed between the defined communities.
 This is done using the following logic:
@@
 <!-- 1. Identify all links that have segregated cycling infrastructure and add them to the initial solution -->
+
 <!-- 2. Identify all links that neighbor links in the current solution -->
+
 <!-- 3. Select *from each community* one neighboring link with highest flow and add it to the solution -->
+
 <!-- 4. If there are no more neighboring links in a community, select the link with the highest flow in that community, regardless of connectivity, and add it to the solution -->
+
 <!-- 5. Repeat steps 2, 3 \& 4 until all flow is satisfied or investment threshold is met  -->
@@
 
-Even though we may end up with a more disconnected network, we will have
-separate connected networks in each community. Given that communities
-are defined by having more internal flow than external flow, this is a
-satisfactory solution.
+Even though we may end up with a more disconnected network, we will have separate connected networks in each community.
+Given that communities are defined by having more internal flow than external flow, this is a satisfactory solution.
 
-The results of the community detection are used to evaluate the
-algorithms. This is done by looking at the *person-km satisfied* as
-cycling infrastructure is added. Person-km is a measure of the total km
-cycled on a road segment, so it is the product of the number of
-potential commuters cycling on that road segment ($flow$) and the length
-of the segment in km ($l$). For each road segment, the person-km is
-equal to $flow * l$. In the case of Manchester, Table
-\@ref(tab:personkmtable) shows that almost half of the person-km is in
-community 1, while only 0.5% of total person-km on the network is in
-community 5.
+The results of the community detection are used to evaluate the algorithms.
+This is done by looking at the *person-km satisfied* as cycling infrastructure is added.
+Person-km is a measure of the total km cycled on a road segment, so it is the product of the number of potential commuters cycling on that road segment ($flow$) and the length of the segment in km ($l$).
+For each road segment, the person-km is equal to $flow * l$.
+In the case of Manchester, Table \@ref(tab:personkmtable) shows that almost half of the person-km is in community 1, while only 0.5% of total person-km on the network is in community 5.
 
@@
 
-Looking at the person-km satisfied (Figure \@ref(fig:growthtotal)), we
-see that the incremental addition of cycling infrastructure is better
-distributed between communities using Algorithm 2; equal distribution of
-investment results in the gain in % of person km satisfied in each
-community being inversely correlated with the size of the community. In
-addition, we find that the restrictions imposed by Algorithm 2 on the
-network expansion do not seem to have a noticeable effect on the
-city-wide % of person-km satisfied. Comparing both algorithms, we can
-see that Algorithm 1 provides only marginally quicker city-wide gains
-than Algorithm 2.
+Looking at the person-km satisfied (Figure \@ref(fig:growthtotal)), we see that the incremental addition of cycling infrastructure is better distributed between communities using Algorithm 2; equal distribution of investment results in the gain in % of person km satisfied in each community being inversely correlated with the size of the community.
+In addition, we find that the restrictions imposed by Algorithm 2 on the network expansion do not seem to have a noticeable effect on the city-wide % of person-km satisfied.
+Comparing both algorithms, we can see that Algorithm 1 provides only marginally quicker city-wide gains than Algorithm 2.
 
@@
 
-Figure \@ref(fig:growth3MapandBar) gives us a geographic representation
-of the results from Algorithm 2; it shows when each link was added to
-the solution (first 100km, second 100km, etc). We can see that,
-generally, road segments around cycling infrastructure are prioritized,
-except for those neighboring cycling infrastructure on the very
-periphery. The first 100km is also spatially distributed across the
-city, with no apparent bias towards a particular area.
+Figure \@ref(fig:growth3MapandBar) gives us a geographic representation of the results from Algorithm 2; it shows when each link was added to the solution (first 100km, second 100km, etc).
+We can see that, generally, road segments around cycling infrastructure are prioritized, except for those neighboring cycling infrastructure on the very periphery.
+The first 100km is also spatially distributed across the city, with no apparent bias towards a particular area.
 
-It is also important to understand how the different highway types
-contribute to the proposed network. Figure \@ref(fig:growth3MapandBar)
-shows that most of the flow will be on residential and tertiary roads,
-as expected from the weighting profile defined in Table
-\@ref(table:weightprofiles).
+It is also important to understand how the different highway types contribute to the proposed network.
+Figure \@ref(fig:growth3MapandBar) shows that most of the flow will be on residential and tertiary roads, as expected from the weighting profile defined in Table \@ref(table:weightprofiles).
 
@@
 
-Existing cycling infrastructure is made up of many disconnected
-components. Both Algorithm 1 and 2 start with all existing segregated
-cycling infrastructure and aim to create an efficient, connected
-network. Figure \@ref(fig:componentsandGCC) shows that both algorithms
-gradually reduce the number of components as more infrastructure is
-added, but Algorithm 2 is able to provide better connectivity with less
-investment.
+Existing cycling infrastructure is made up of many disconnected components.
+Both Algorithm 1 and 2 start with all existing segregated cycling infrastructure and aim to create an efficient, connected network.
+Figure \@ref(fig:componentsandGCC) shows that both algorithms gradually reduce the number of components as more infrastructure is added, but Algorithm 2 is able to provide better connectivity with less investment.
 
-Consistent growth can also be seen for the size of the Largest Connected
-Component in the proposed bicycle network (Figure
-\@ref(fig:componentsandGCC)). Here, however, we find that there is little
-difference between both algorithms.
+Consistent growth can also be seen for the size of the Largest Connected Component in the proposed bicycle network (Figure \@ref(fig:componentsandGCC)).
+Here, however, we find that there is little difference between both algorithms.
 
@@
 
-Overall, the improved distribution of investment shown by Algorithm 2 does not seem to come at the cost 
-of poorer connectivity or less city-wide gains when compared to Algorithm 1. 
+Overall, the improved distribution of investment shown by Algorithm 2 does not seem to come at the cost of poorer connectivity or less city-wide gains when compared to Algorithm 1.
 The results therefore advocate for the incorporation of egalitarian principles in bicycle network planning.
@@
 
-This paper demonstrates an approach for prioritizing
-investments in city-wide cycleway networks. The approach aims to respect
-both the needs of the users and the ambitions of stakeholders working at
-local or regional levels. The results, including detailed route network
-maps based on current travel behaviour derived from OD data, can provide
-an evidence-base for designing direct, connected,
-and low-stress networks. A core contribution of this approach is that it insures 
-that investments in cycling infrastructure are distributed fairly between different 
-communities in a city.
+This paper demonstrates an approach for prioritizing investments in city-wide cycleway networks.
+The approach aims to respect both the needs of the users and the ambitions of stakeholders working at local or regional levels.
+The results, including detailed route network maps based on current travel behaviour derived from OD data, can provide an evidence-base for designing direct, connected, and low-stress networks.
+A core contribution of this approach is that it insures that investments in cycling infrastructure are distributed fairly between different communities in a city.
 
-Given that the "most essential activity entailed in the design of
-cycle-friendly infrastructure is developing a cycle network"
-[@parkin2018designing], we believe that the approach has great
-potential to inform investment in cities such as Manchester where there
-is political will to invest in cycling long-term. A benefit of the
-approach is that it has relatively modest data requirements: only the
-road network (from OSM), its topography (from satellite imagery), and
-OD data (from the national census in this case) are
-needed, opening up the possibility of deploying the approach in other
-cities^[The results are easily reproducible for all UK cities, and can also be reproduced for cities elsewhere given the availability of commuter data.
-Instructions for reproducing the results are detailed in the **README** file in the R folder of the .zip file that accompanies this paper (Link to public github repo to be added when paper is published)].
+Given that the "most essential activity entailed in the design of cycle-friendly infrastructure is developing a cycle network" [@parkin2018designing], we believe that the approach has great potential to inform investment in cities such as Manchester where there is political will to invest in cycling long-term.
+A benefit of the approach is that it has relatively modest data requirements: only the road network (from OSM), its topography (from satellite imagery), and OD data (from the national census in this case) are needed, opening up the possibility of deploying the approach in other cities[^5].
+
+[^5]: The results are easily reproducible for all UK cities, and can also be reproduced for cities elsewhere given the availability of commuter data.
+    Instructions for reproducing the results are detailed in the **README** file in the R folder of the .zip file that accompanies this paper (Link to public github repo to be added when paper is published)
 
@@
 <!-- [The results are easily reproducible for all UK cities, and can -->
+
 <!-- also be reproduced for cities elsewhere given the availability of -->
+
 <!-- commuter data. The appendix shows the results for 3 other UK cities: Birmingham, Leeds and Nottingham.  -->
+
 <!-- For instructions on reproducing the results shown in this -->
+
 <!-- paper, see the **README** file in the R folder of the .zip file that -->
+
 <!-- accompanies this paper (to be released as open source software when the -->
+
 <!-- paper is published).]. -->
 
+```{=html}
 <!-- For after publishing
@@
 -->
+```
+The approach can identify not only where there is high potential for cycling but also trade-offs between stress levels (related to motor traffic) and directness.
+The results can therefore be used as a basis for recommendations on road space reallocation *and* new infrastructure to unlock potential cycling demand.
+The approach encourages consideration of a wider range of preferences and needs than previous approaches that focus only on absolute potential.
+Moreover, the inclusion of egalitarian principles in scenarios of change encourages investment in cycling infrastructure that addresses geographical and social inequalities.
+This is done by explicitly ensuring that infrastructure improvements are pursued in different parts of the city simultaneously.
+This approach compares well with one that focuses exclusively on maximizing the number of potential cycling trips that benefit from investment, showing that a fair distribution can be achieved without compromising on overall benefit.
+This ability to address inequalities in network prioritization is particularly important given research showing substantial *inequalities* around transport provision in general and cycling uptake and investment in particular [@lucas2016method; @vidaltortosa_infrastructure_2020].
 
-The approach can identify not only where there is high
-potential for cycling but also trade-offs between
-stress levels (related to motor traffic) and directness.
-The results can therefore be used as a basis for recommendations on
-road space reallocation *and* new infrastructure to unlock potential cycling demand.
-The approach encourages consideration of a wider range of preferences
-and needs than previous approaches that focus only on absolute potential.
-Moreover, the inclusion of egalitarian principles in scenarios of change
-encourages investment in cycling infrastructure that addresses geographical and 
-social inequalities. This is done by explicitly ensuring that infrastructure improvements 
-are pursued in different parts of the city simultaneously. This approach compares well with one 
-that focuses exclusively on maximizing the number of potential cycling trips that 
-benefit from investment, showing that a fair distribution can be achieved without 
-compromising on overall benefit. 
-This ability to address inequalities in network prioritization is particularly
-important given research showing substantial *inequalities*
-around transport provision in general and cycling uptake and investment
-in particular [@lucas2016method; @vidaltortosa_infrastructure_2020].
-
-The approach is not without limitations. The level of detail is only as
-good as the granularity of the available data (in this case relatively
-coarse zones). @iacono2010measuring note that such large travel zones
-are not ideal for understanding route choice behaviour of cyclists and
-pedestrians. They also give rise to an \`ecological fallacy' whereby
-average characteristics are assumed to apply to all residents of the
-aggregated geographical area, suggesting a need for applying the methods
-to more granular OD data (and for governments and other
-data-collecting organisations to make OD data more
-readily available).
-The potential demand calculation is also based on cycling in the traditional sense, and does not consider the 
-effect of micro-mobility on reducing topology-related impedance to cycling.
+The approach is not without limitations.
+The level of detail is only as good as the granularity of the available data (in this case relatively coarse zones).
+@iacono2010measuring note that such large travel zones are not ideal for understanding route choice behaviour of cyclists and pedestrians.
+They also give rise to an \`ecological fallacy' whereby average characteristics are assumed to apply to all residents of the aggregated geographical area, suggesting a need for applying the methods to more granular OD data (and for governments and other data-collecting organisations to make OD data more readily available).
+The potential demand calculation is also based on cycling in the traditional sense, and does not consider the effect of micro-mobility on reducing topology-related impedance to cycling.
 Given that we are proposing an approach which can accommodate any cycling uptake functions, this is an acceptable limitation.
 
-The approach is also focused solely on the allocation of cycling
-infrastructure, and does not consider the larger political and
-regulatory environment that needs to exist to promote cycling; while
-segregated, connected, and direct cycling infrastructure is key to
-achieving high levels of cycling, research has shown that it cannot
-exist in a vacuum. @wardman2007factors developed a mode choice model
-for the UK and their results showed that improved cycling infrastructure
-on its own only had modest impacts on mode shift, and even the unlikely
-scenario of all urban routes being serviced by segregated bike lanes was
-forecast to increase cycling mode share by only 3%. International
-research shows that cities that invest in more comprehensive cycling
-projects have a more significant increase in the number of cyclists as
-well as the cycling mode share [@parkin2018designing;
-@pucher2010infrastructure]. These cities do not just focus on
-infrastructure, but on general policies as well as restricting car use.
-Evaluation of policies in Denmark and Germany and the Netherlands has
-shown that their high cycling mode share is down to a broader set of
-soft and hard policies. Hard policies include traffic calming, filtered permeability
-interventions, cycling rights of way, bike parking, integration with the
-public transport network, and making driving cars both expensive and
-inconvenient, while soft policies include marketing and awareness campaigns [@gossling2013urban; @pucher2008making].
-While these policies are outside the scope of this research, it is important to recognize their key role in
-bringing about an increase in levels of cycling.
+The approach is also focused solely on the allocation of cycling infrastructure, and does not consider the larger political and regulatory environment that needs to exist to promote cycling; while segregated, connected, and direct cycling infrastructure is key to achieving high levels of cycling, research has shown that it cannot exist in a vacuum.
+@wardman2007factors developed a mode choice model for the UK and their results showed that improved cycling infrastructure on its own only had modest impacts on mode shift, and even the unlikely scenario of all urban routes being serviced by segregated bike lanes was forecast to increase cycling mode share by only 3%.
+International research shows that cities that invest in more comprehensive cycling projects have a more significant increase in the number of cyclists as well as the cycling mode share [@parkin2018designing; @pucher2010infrastructure].
+These cities do not just focus on infrastructure, but on general policies as well as restricting car use.
+Evaluation of policies in Denmark and Germany and the Netherlands has shown that their high cycling mode share is down to a broader set of soft and hard policies.
+Hard policies include traffic calming, filtered permeability interventions, cycling rights of way, bike parking, integration with the public transport network, and making driving cars both expensive and inconvenient, while soft policies include marketing and awareness campaigns [@gossling2013urban; @pucher2008making].
+While these policies are outside the scope of this research, it is important to recognize their key role in bringing about an increase in levels of cycling.
 
 Consideration of these limitations suggest future directions of research.
-New datasets and continuously evolving computational (hardware and software)
-capabilities should enable the data related issues to be overcome as open
-datasets, and our ability to process them, improve.
-Plugging-in alternative uptake models could help address the relatively narrow
-definition of 'potential' used in this paper, to consider broader social factors.
-A promising area of future research in this direction could be 
-to integrate a wider range of modes, including e-bikes and e-scooters, into the
-analysis. Currently scope for progress in this direction is
-partially restricted by the lack of data on the
-proliferation of these modes, raising the point that surveys could
-replace the broad category of `Bicycle' with a range of small modes such as
-'pedal cycles' (including tricycles and recumbent cycles), e-bikes
-and other forms of micro-mobility.
+New datasets and continuously evolving computational (hardware and software) capabilities should enable the data related issues to be overcome as open datasets, and our ability to process them, improve.
+Plugging-in alternative uptake models could help address the relatively narrow definition of 'potential' used in this paper, to consider broader social factors.
+A promising area of future research in this direction could be to integrate a wider range of modes, including e-bikes and e-scooters, into the analysis.
+Currently scope for progress in this direction is partially restricted by the lack of data on the proliferation of these modes, raising the point that surveys could replace the broad category of \`Bicycle' with a range of small modes such as 'pedal cycles' (including tricycles and recumbent cycles), e-bikes and other forms of micro-mobility.
 
-Despite the rapid growth of these alternative modes, there is little doubt that
-the humble bicycle is a key ingredient in the ongoing shift towards active transport.
-Recognition of the importance of this shift for improved health and well-being
-of citizens, and the wider challenge to decarbonise the global economy,
-has grown with pressures on health and public transport systems during the pandemic.
-The success of policies to accelerate cycling uptake depends on a range of
-factors including, vitally, the design of the
-cycling network for potential cyclists.
-The approach presented in this paper provides a strong evidence base,
-that considers both cycling potential and social equity,
-for designing joined-up and cost-effective strategic cycle networks.
+Despite the rapid growth of these alternative modes, there is little doubt that the humble bicycle is a key ingredient in the ongoing shift towards active transport.
+Recognition of the importance of this shift for improved health and well-being of citizens, and the wider challenge to decarbonise the global economy, has grown with pressures on health and public transport systems during the pandemic.
+The success of policies to accelerate cycling uptake depends on a range of factors including, vitally, the design of the cycling network for potential cyclists.
+The approach presented in this paper provides a strong evidence base, that considers both cycling potential and social equity, for designing joined-up and cost-effective strategic cycle networks.
 <!-- In tandem with policies to reduce excessive reliance on private cars, evidence-based investment in strategic cycle networks is key to. -->

### 2023-04-24 README.tex
plot for paper

--- 2022-09-03 README.tex
@@
 A key benefit of the approach from a policy perspective is its ability to support egalitarian and cost-effective strategic cycle network planning.
-
-\par
-
-\textbf{Keywords:} cycling networks, low-traffic neighborhoods, routing, transport equity
+¶ \par ¶ \textbf{Keywords:} cycling networks, low-traffic neighborhoods, routing, transport equity
 \end{abstract}
@@
 
+\hypertarget{method-and-application}{%
+\section{Method and Application}\label{method-and-application}}
+
+\hypertarget{method-overview}{%
+\subsection{Method overview}\label{method-overview}}
+
 This research attempts to propose an egalitarian framework for cycling
@@
 demand. It then introduces the network expansion algorithms, and compares an approach grounded in
-`egalitarianism' to one grounded in `utilitarianism'.
+`egalitarianism' to one grounded in `utilitarianism'. As mentioned above, Manchester is used as a case study.
+An overview of the steps involved is shown in Figure \ref{fig:methodflowchart}.
+
+\begin{figure}
+
+{\centering \includegraphics[width=0.8\linewidth]{extras/methodology_flowchart}
+
+}
+
+\caption{Flowchart of proposed methodology}\label{fig:methodflowchart}
+\end{figure}
 
 \hypertarget{calculating-potential-cycling-demand}{%
-\section{Calculating Potential Cycling Demand}\label{calculating-potential-cycling-demand}}
+\subsection{Calculating Potential Cycling Demand}\label{calculating-potential-cycling-demand}}
 
@@
 number of commuters between administrative zones --- Middle layer Super
-Output Areas (MSOA) --- by mode of travel. MSOAs have an average
+Output Areas (MSOA) --- by mode of travel (this is similar to the approach
+used by Grisé and El-Geneidy (2018)).
+MSOAs have an average
 population of just over 8000 (ONS, 2018).\footnote{ See
@@
 \hypertarget{routing}{%
-\section{Routing}\label{routing}}
+\subsection{Routing}\label{routing}}
 
@@
 without cycling infrastructure by a fixed impedance factor
-(Mauttone et al., 2017), or by assigning different weights to the road
-segments proportional to the investment cost of bringing them to an
-acceptable level of stress for cycling (Duthie and Unnikrishnan, 2014).
+(Mauttone et al., 2017), by assigning different weights to the road
+segments that are (a) proportional to the investment cost of bringing them to an
+acceptable level of stress for cycling (Duthie and Unnikrishnan, 2014), or (b) inversely proportional to
+perceived level of stress experienced by cyclists (Wysling and Purves, 2022).
 Perceived stress for cyclists has been found to be a function of vehicular traffic
@@
 favouring roads with existing cycling infrastructure.
+It is also different to the weighting profile used by Wysling and Purves (2022) as it adds a larger
+weight on trunk and primary roads then other roads without cycling infrastructure.
 
@@
 \hypertarget{road-segment-prioritization}{%
-\section{Road Segment Prioritization}\label{road-segment-prioritization}}
+\subsection{Road Segment Prioritization}\label{road-segment-prioritization}}
 
@@
 \hypertarget{community-detection}{%
-\subsection{Community Detection}\label{community-detection}}
+\subsubsection{Community Detection}\label{community-detection}}
 
@@
 \hypertarget{algorithm-1-utilitarian-expansion}{%
-\subsection{Algorithm 1: Utilitarian Expansion}\label{algorithm-1-utilitarian-expansion}}
+\subsubsection{Algorithm 1: Utilitarian Expansion}\label{algorithm-1-utilitarian-expansion}}
 
@@
 \hypertarget{algorithm-2-egalitarian-expansion}{%
-\subsection{Algorithm 2: Egalitarian Expansion}\label{algorithm-2-egalitarian-expansion}}
+\subsubsection{Algorithm 2: Egalitarian Expansion}\label{algorithm-2-egalitarian-expansion}}
 
@@
 \hypertarget{connectivity}{%
-\subsection{Connectivity}\label{connectivity}}
+\subsubsection{Connectivity}\label{connectivity}}
 
@@
 Given that we are proposing an approach which can accommodate any cycling uptake functions, this is an acceptable limitation.
+The network prioritization is based strongly on potential cycling demand calculations.
+While this provides a solid supplementary tool for network planning, it would be more useful as part of a multi-criteria approach that considers on-the ground information from existing cyclists, road geometries, and traffic analysis.
+Grisé and El-Geneidy (2018) propose a multi-criteria approach that includes the first of these points.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-grise2018if}{}}%
+Grisé, E., El-Geneidy, A., 2018.
+If we build it, who will benefit?
+A multi-criteria approach for the prioritization of new bicycle lanes in quebec city, canada.
+Journal of transport and land use 11, 217--235.
+
 \leavevmode\vadjust pre{\hypertarget{ref-iacono2010measuring}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-wysling2022improve}{}}%
+Wysling, L., Purves, R.S., 2022.
+Where to improve cycling infrastructure?
+Assessing bicycle suitability and bikeability with open data in the city of paris.
+Transportation research interdisciplinary perspectives 15, 100648.
+
 \end{CSLReferences}

### 2023-04-24 README.Rmd

@@
 always_allow_html: yes
+bibliography: references.bib
+bibliography: references.bib
 ---
@@
 
-Much research has expores the question of what gets people to cycle.
-Segregated cycling infrastructure, which refers to road space that is allocated to cyclists only with physical separation to protect cyclists, is associated with cycling uptake [@aldred2019impacts; @goodman2014new; @marques2015infrastructure], with the separation from motorized vehicles being key [@winters2011motivators].
+Much research has explored the question of what gets people cycling.
+Segregated cycling infrastructure --- road space that is allocated to cyclists only with physical separation to protect cyclists --- is associated with cycling uptake [@aldred2019impacts; @goodman2014new; @marques2015infrastructure].
+Separation from motorized vehicles has found to be key [@winters2011motivators].
 Revealed preference of cyclists shows that they are willing to deviate from the most efficient routes in order to commute on safer roads [@crane2017longitudinal].
-However, such deviations are only considered if they do not considerably increase route circuitry; behaviour studies have found that the probability of choosing a route decreases in proportion to its length relative to the shortest route [@broach2011bicycle; @winters2010far].
+However, such deviations are only considered if they do not considerably increase route circuitry; behavior studies have found that the probability of choosing a route decreases in proportion to its length relative to the shortest route [@broach2011bicycle; @winters2010far].
 Another defining feature for cycling infrastructure is how well connected it is.
@@
 In this paper, continuity is analysed by looking at the connectivity of the network through the graph-theoretic concept of *connected components*.
+
+[^2]: *flow* is used throughout this research to refer to the cycling demand when it is routed onto the road network.
+    The flow on any road segment is the cumulative demand on it, resulting from cyclists commuting between various Origin-Destination (OD) pairs.
+
 @orozco2020data study the existing cycling network in terms of its disconnected components.
@@
 After routing the cycling demand onto the network links, they use percolation theory to filter out the links based on the aggregate flow passing through them, varying the flow threshold for filtering to identify the minimum flow at which the whole city is connected by a single component.
+While these approaches deal with continuity better, they look at the network as a whole when attempting to improve it, and in doing so fail to account for equitable distribution of infrastructure.
 
-[^2]: *flow* is used throughout this research to refer to the cycling demand when it is routed onto the road network.
-    The flow on any road segment is the cumulative demand on it, resulting from cyclists commuting between various Origin-Destination (OD) pairs.
+Other approaches have focused on characteristics of route segments as the basis of analysis to help prioritize investment in cycle infrastructure where it is most needed.
+@grisé2018 developed a multi-criteria approach, combined with estimated levels of usage, to highlight routes most in need of investment in Quebec City.
+@wysling2022 present a method that used open data from Paris to generate cycling suitability ratings based on 'bicycle suitability multipliers' associated with different types of road across the city, to support more effective interventions.
 
-While these approaches deal with continuity better, they look at the network as a whole when attempting to improve it, and in doing so fail to account for equitable distribution of infrastructure.
+None of the aforementioned studies explore the equity impacts of different network prioritization approaches.
+The following section highlights the importance of considering (often unstated) ethical assumptions underlying network optimisation and prioritization approaches of the type outlined above.
 

### 2023-04-24 README.Rmd

@@
 output:
- # bookdown::github_document2:
-   bookdown::pdf_document2:
+ bookdown::github_document2:
+   # bookdown::pdf_document2:
  # bookdown::word_document2:
@@
 always_allow_html: yes
-bibliography: references.bib
-bibliography: references.bib
 ---
@@
 Other approaches have focused on characteristics of route segments as the basis of analysis to help prioritize investment in cycle infrastructure where it is most needed.
-@grisé2018 developed a multi-criteria approach, combined with estimated levels of usage, to highlight routes most in need of investment in Quebec City.
-@wysling2022 present a method that used open data from Paris to generate cycling suitability ratings based on 'bicycle suitability multipliers' associated with different types of road across the city, to support more effective interventions.
+@grise2018if developed a multi-criteria approach, combined with estimated levels of usage, to highlight routes most in need of investment in Quebec City.
+@wysling2022improve present a method that used open data from Paris to generate cycling suitability ratings based on 'bicycle suitability multipliers' associated with different types of road across the city, to support more effective interventions.
 
@@
 
+# Method and Application
+
+## Method overview
+
 This research attempts to propose an egalitarian framework for cycling network design.
@@
 The benefit is the flow on the link, and each link is assigned a cost based on current road conditions.
-
-<!-- EDIT! Add more to limitations -->
 
@@
 Section \@ref(road-segment-prioritization) outlines a method for partitioning the study area based on a community finding algorithm and routed cycling demand.
-It then introduces the network expansion algorithms, and compares an approach grounded in 'egalitarianism' to one grounded in 'utilitarianism'.
+It then introduces the network expansion algorithms, and compares an approach grounded in 'egalitarianism' to one grounded in 'utilitarianism'. ithms, and compares an approach grounded in
+'egalitarianism' to one grounded in 'utilitarianism'. As mentioned above, Manchester is used as a case study. 
+An overview of the steps involved is shown in Figure \@ref(fig:methodflowchart).
 
-# Calculating Potential Cycling Demand
+```{r methodflowchart, fig.cap="Flowchart of proposed methodology", out.width="80%", fig.show='hold', fig.align = "center", echo=FALSE}
+knitr::include_graphics("extras/methodology_flowchart.png")
+```
+
+## Calculating Potential Cycling Demand
 
@@
 In areas where observed OD data is unavailable, modelling techniques such as spatial interaction models can be used to estimate travel volumes between zones [@black1995spatial; @martinez2013new; @wilson1971family].
-In this paper we use open access data from the UK census [@ONS2011flowdata], which contains aggregate statistics on number of commuters between administrative zones --- Middle layer Super Output Areas (MSOA) --- by mode of travel.
+In this paper we use open access data from the UK census [@ONS2011flowdata], which contains aggregate statistics on number of commuters between administrative zones --- Middle layer Super Output Areas (MSOA) --- by mode of travel (this is similar to the approach used by @grise2018if).
 MSOAs have an average population of just over 8000 [@ofn2018population].[^3]
@@
 The above points are accounted for by using a weighted road network for routing.
-This has been previously done by multiplying all road segments without cycling infrastructure by a fixed impedance factor [@mauttone2017bicycle], or by assigning different weights to the road segments proportional to the investment cost of bringing them to an acceptable level of stress for cycling [@duthie2014optimization].
+This has been previously done by multiplying all road segments without cycling infrastructure by a fixed impedance factor [@mauttone2017bicycle], or by assigning different weights to the road segments that are (a) proportional to the investment cost of bringing them to an acceptable level of stress for cycling [@duthie2014optimization], or (b) inversely proportional to perceived level of stress experienced by cyclists [@wysling2022improve].
 Perceived stress for cyclists has been found to be a function of vehicular traffic volume and speed [@sorton1994bicycle], both of which vary predictably with road type.
@@
 We believe this to be more appropriate than the simplistic approach adopted by @mauttone2017bicycle as it makes use of a hierarchy of road preference based on perceived stress levels, going beyond simply favouring roads with existing cycling infrastructure.
+It is also different to the weighting profile used by @wysling2022improve as it adds a larger weight on trunk and primary roads then other roads without cycling infrastructure.
 

### 2023-04-24 README.tex

@@
 
-The 2015 Paris agreement (UN, 2015) acknowledged that
-fundamental changes to societies and economies are necessary to mitigate
-climate change. Like other sectors, transport is under substantial
-pressure to decarbonise, resulting in a number of technical innovations
-including electric vehicles. But new vehicle technologies can only go so
-far and do not tackle parallel problems such as congestion, road traffic
-casualties and physical inactivity (Brand et al., 2020).
-
-In this context, interest and investment in active modes are growing.
-The benefits extend beyond congestion and the environment, as active
-travel can also help alleviate what is referred to as the pandemic of
-global inactivity; physical inactivity is on the rise and has become the
-4th highest cause of death globally (Kohl 3rd et al., 2012). Various
-studies have documented the association between active transport and
-lower risk of disease, including cancer and cardiovascular disease
-(Celis-Morales et al., 2017; Jarrett et al., 2012; Patterson et al., 2020). In the wake of the Covid-19 pandemic, and
-the resulting reduced capacity of public transport, the UK government
-has pledged to invest billions of pounds to improve walking and cycling
-infrastructure across the country. While this unprecedented sum is an
-opportunity to reshape cities in a way that improves the well-being of
-citizens, it does come with a warning:
+The 2015 Paris agreement (UN, 2015) acknowledged that fundamental changes to societies and economies are necessary to mitigate climate change.
+Like other sectors, transport is under substantial pressure to decarbonise, resulting in a number of technical innovations including electric vehicles.
+But new vehicle technologies can only go so far and fail to tackle parallel problems such as congestion, road traffic casualties and physical inactivity (Brand et al., 2020).
+
+In this context, interest and investment in active modes has grown substantially, especially with the growing realisation that the benefits extend beyond congestion and the environment, as active travel can also help alleviate what is referred to as the pandemic of global inactivity; physical inactivity is on the rise and has become the 4th highest cause of death globally (Kohl 3rd et al., 2012).
+Various studies have documented the association between active transport and lower risk of disease, including cancer and cardiovascular disease (Celis-Morales et al., 2017; Jarrett et al., 2012; Patterson et al., 2020).
+In the wake of the Covid-19 pandemic, and the resulting reduced capacity of public transport, the UK government has pledged to invest billions of pounds to improve walking and cycling infrastructure across the country.
+While this unprecedented sum is an opportunity to reshape cities in a way that improves the well-being of citizens, it does come with a warning:
 
 \begin{quote}
-``Inadequate cycling infrastructure discourages cycling and wastes
-public money. Much cycling infrastructure in this country is
-inadequate. It reflects a belief, conscious or otherwise, that hardly
-anyone cycles, that cycling is unimportant and that cycles must take
-no meaningful space from more important road users, such as motor
-vehicles and pedestrians'' (DfT, 2020a).
+``Inadequate cycling infrastructure discourages cycling and wastes public money.
+Much cycling infrastructure in this country is inadequate.
+It reflects a belief, conscious or otherwise, that hardly anyone cycles, that cycling is unimportant and that cycles must take no meaningful space from more important road users, such as motor vehicles and pedestrians'' (DfT, 2020a).
 \end{quote}
 
-The funding on its own is therefore no guarantee of a change in
-commuting across the country; it must be used to design adequate cycling
-infrastructure that is based on motivators and deterrents to cycling. In this paper we
-propose an approach for planning a connected cycling network that is meant to
-be applicable to any city (as shown in the Appendix), and we use the city of Manchester as a case study. Greater
-Manchester aims to become the first region to reach a carbon neutral transport network;
-its local authorities have developed plans for a 1800 mile walking and cycling network
-(Bee Network) that will be delivered over 10 years (Greater Manchester, 2020). While schemes
-have already been put in place for around a quarter of this network, the majority is yet to be built.
+Funding alone is therefore no guarantee of a change in commuting across the country; it must be used to design adequate cycling infrastructure that is based on motivators and deterrents to cycling.
+In this paper we propose an approach for planning connected cycling networks that can be generalised to any city with appropriate input data (as shown in the Appendix).
+We used the city of Manchester as a case study due to its ambitions to become the first UK region with a carbon neutral transport network, supported by ambitious plans for a 1800 mile walking and cycling network (the `Bee Network') that will be delivered over 10 years (Greater Manchester, 2020).
+While schemes have already been put in place for around a quarter of this network, the majority is yet to be built.
 
@@
 
-Much research has been done to understand what gets people to cycle.
-Segregated cycling infrastructure\footnote{\emph{Segregated cycling infrastructure}
-  refers to road space that is allocated to cyclists only, with physical
-  separation to protect cyclists from other modes of transport.} has been
-shown to increase cycling uptake (Aldred et al., 2019; Goodman et al., 2014; Marqués et al., 2015), with the separation from motorized
-vehicles being key (Winters et al., 2011). Revealed preference of
-cyclists shows that they are willing to deviate from the most efficient
-routes in order to commute on safer roads (Crane et al., 2017).
-However, such deviations are only considered if they do not considerably
-increase route circuitry; behaviour studies have found that the
-probability of choosing a route decreases in proportion to its length
-relative to the shortest route (Broach et al., 2011; Winters et al., 2010).
-Another defining feature for cycling infrastructure is how well
-connected it is. Cyclists prefer cohesive infrastructure, particularly
-when cycling on arterial roads with high levels of motorized traffic
-(Stinson and Bhat, 2003). The lack of well-connected cycling
-infrastructure is one of the main obstacles to increasing cycling uptake
-(Caulfield et al., 2012). While direct and cohesive cycling networks
-have been shown to positively impact cycling rates, density\footnote{making an
-  area's bicycle network denser means adding more cycling routes in the
-  area and thereby giving cyclists more route options.} of the cycling
-network is also vital (Schoner and Levinson, 2014).
+Much research has explored the question of what gets people cycling.
+Segregated cycling infrastructure --- road space that is allocated to cyclists only with physical separation to protect cyclists --- is associated with cycling uptake (Aldred et al., 2019; Goodman et al., 2014; Marqués et al., 2015).
+Separation from motorized vehicles has found to be key (Winters et al., 2011).
+Revealed preference of cyclists shows that they are willing to deviate from the most efficient routes in order to commute on safer roads (Crane et al., 2017).
+However, such deviations are only considered if they do not considerably increase route circuitry; behavior studies have found that the probability of choosing a route decreases in proportion to its length relative to the shortest route (Broach et al., 2011; Winters et al., 2010).
+Another defining feature for cycling infrastructure is how well connected it is.
+Cyclists prefer cohesive infrastructure, particularly when cycling on arterial roads with high levels of motorized traffic (Stinson and Bhat, 2003).
+The lack of well-connected cycling infrastructure is one of the main obstacles to increasing cycling uptake (Caulfield et al., 2012).
+While direct and cohesive cycling networks have been shown to positively impact cycling rates, density\footnote{making an area's bicycle network denser means adding more cycling routes in the area and thereby giving cyclists more route options.} of the cycling network is also vital (Schoner and Levinson, 2014).
 
@@
 
-The studies outlined above lay out the fundamentals for designing
-cycling networks that generate significant cycling uptake, but they do
-not propose network-level interventions. In this section we outline
-methods used in past studies, namely optimization and network
-analysis methods, such as connected components and community
-detection, and examine how they are leveraged to suggest cycling network
-designs. We compare the effectiveness of these network-level studies in
-incorporating the fundamentals outlined above.
-Our proposed approach is inspired by these methods, but it attempts to add to them
-by ensuring that all of the outlined fundamentals are accounted for. It
-also goes further by attempting to factor in ethical considerations
-related to the distribution of investment.
-
-\emph{Optimization} methods have been used to propose improvements to
-cycling networks. Mesbah et al. (2012) propose a bi-level formulation to
-optimize allocation of cycling lanes to the network without exceeding a
-set budget. They account for the effect of cycling lanes on car traffic,
-and attempt to maximize utilization of said lanes with minimal impact on
-car travel times.
+The studies outlined above lay out the fundamentals for designing cycling networks that generate significant cycling uptake, but they do not propose network-level interventions.
+In this section we outline methods used in past studies, namely optimization and network analysis methods, such as connected components and community detection, and examine how they are leveraged to suggest cycling network designs.
+We compare the effectiveness of these network-level studies in incorporating the fundamentals outlined above.
+Our proposed approach is inspired by these methods, but it attempts to add to them by ensuring that all of the outlined fundamentals are accounted for.
+It also goes further by attempting to factor in ethical considerations related to the distribution of investment.
+
+\emph{Optimization} methods have been used to propose improvements to cycling networks.
+Mesbah et al. (2012) propose a bi-level formulation to optimize allocation of cycling lanes to the network without exceeding a set budget.
+They account for the effect of cycling lanes on car traffic, and attempt to maximize utilization of said lanes with minimal impact on car travel times.
 Safeguarding against increased car traffic may be counter-productive if the goal is to create a mode-shift, as research has shown that reducing road space for cars leads to less cars on the road, a phenomenon referred to as ``traffic evaporation'' (Nello-Deakin, 2020).
-On the other hand, Mauttone et al. (2017) developed an optimization framework aiming at minimizing the total
-user cost of cycling on the network without considering car usage. The aggregate flow\footnote{\emph{flow} is
-  used throughout this research to refer to the cycling demand when it is
-  routed onto the road network. The flow on any road segment is the
-  cumulative demand on it, resulting from cyclists commuting between
-  various Origin-Destination (OD) pairs.} on road segments (links) is obtained by using
-shortest paths to route existing cycling demand onto the road network. The solution is a proposed set of links where cycling infrastructure
-should be added in order to minimize the overall travel cost of cyclists
-across the network. The cost of traversing a link is given as a function
-of its length and whether or not it has cycling infrastructure. The
-problem has also been solved by attempting to find the minimum cost of
-improving roadway links to meet a desired level of service (LOS)
-(Duthie and Unnikrishnan, 2014). In this formulation, all origin-destination
-(OD) pairs need to be connected by roads that meet the desired LOS. A
-limitation of these approaches is that they do not explicitly solve for
-continuity. Continuity is addressed using either a constraint specifying
-that each link with a bike lane should be connected to at least one
-destination (Mesbah et al., 2012), a constraint on maximum deviation
-from shortest paths (Duthie and Unnikrishnan, 2014), or a discontinuity
-penalty to prioritize connected road segments (Mauttone et al., 2017).
-
-In this paper, continuity is analysed by looking at the connectivity of
-the network through the graph-theoretic concept of \emph{connected
-components}. Orozco et al. (2020) study the existing cycling network in terms of its disconnected
-components.
+
+Mauttone et al. (2017) developed an optimization framework aiming at minimizing the total user cost of cycling on the network without considering car usage.
+The aggregate flow\footnote{\emph{flow} is used throughout this research to refer to the cycling demand when it is routed onto the road network.
+  The flow on any road segment is the cumulative demand on it, resulting from cyclists commuting between various Origin-Destination (OD) pairs.} on road segments (links) is obtained by using shortest paths to route existing cycling demand onto the road network.
+The solution is a proposed set of links where cycling infrastructure should be added in order to minimize the overall travel cost of cyclists across the network.
+The cost of traversing a link is given as a function of its length and whether or not it has cycling infrastructure.
+The problem has also been solved by attempting to find the minimum cost of improving roadway links to meet a desired level of service (LOS) (Duthie and Unnikrishnan, 2014).
+In this formulation, all origin-destination (OD) pairs need to be connected by roads that meet the desired LOS.
+A limitation of these approaches is that they do not explicitly solve for continuity.
+Continuity is addressed using either a constraint specifying that each link with a bike lane should be connected to at least one destination (Mesbah et al., 2012), a constraint on maximum deviation from shortest paths (Duthie and Unnikrishnan, 2014), or a discontinuity penalty to prioritize connected road segments (Mauttone et al., 2017).
+In this paper, continuity is analysed by looking at the connectivity of the network through the graph-theoretic concept of \emph{connected components}.
+
+Orozco et al. (2020) study the existing cycling network in terms of its disconnected components.
 They propose two different algorithms to connect disconnected components; one connects the two largest components, and the other connects the largest component to its nearest neighbor.
 By measuring the growth of the largest connected component as a function of the kilometers of network added, they determine that these approaches are more effective at improving cycling network connectivity than random allocation of cycling infrastructure.
-The concept of connected components is
-also at the core of the methodology proposed by Olmos et al. (2020). After
-routing the cycling demand onto the network links, they use percolation
-theory to filter out the links based on the aggregate flow passing through them, varying the flow threshold for
-filtering to identify the minimum flow at which the whole city is
-connected by a giant component. While these approaches deal with
-continuity better, they look at the network as a whole when attempting
-to improve it, and in doing so fail to account for equitable
-distribution of infrastructure.
+The concept of connected components is also at the core of the methodology proposed by Olmos et al. (2020).
+After routing the cycling demand onto the network links, they use percolation theory to filter out the links based on the aggregate flow passing through them, varying the flow threshold for filtering to identify the minimum flow at which the whole city is connected by a single component.
+While these approaches deal with continuity better, they look at the network as a whole when attempting to improve it, and in doing so fail to account for equitable distribution of infrastructure.
+
+Other approaches have focused on characteristics of route segments as the basis of analysis to help prioritize investment in cycle infrastructure where it is most needed.
+Grisé and El-Geneidy (2018) developed a multi-criteria approach, combined with estimated levels of usage, to highlight routes most in need of investment in Quebec City.
+Wysling and Purves (2022) present a method that used open data from Paris to generate cycling suitability ratings based on `bicycle suitability multipliers' associated with different types of road across the city, to support more effective interventions.
+
+None of the aforementioned studies explore the equity impacts of different network prioritization approaches.
+The following section highlights the importance of considering (often unstated) ethical assumptions underlying network optimisation and prioritization approaches of the type outlined above.
 
@@
 
-All of these network-level methodologies are underpinned by ethical
-principles, even though these principles are not explicitly acknowledged
-by the authors. This is important since different ethical principles
-constitute different problem formulations and targets. Broadly speaking,
-transport appraisal can be based on either utilitarian or egalitarian
-principles. The former seeks to maximize the overall benefit, while the
-latter is concerned with a fair distribution of benefits
-(Jafino et al., 2020).
+All of these network-level methodologies are underpinned by ethical principles, even though these principles are not explicitly acknowledged by the authors.
+This is important since different ethical principles constitute different problem formulations and targets.
+Broadly speaking, transport appraisal can be based on either utilitarian or egalitarian principles.
+The former seeks to maximize the overall benefit, while the latter is concerned with a fair distribution of benefits (Jafino et al., 2020).
 The utilitarian approach, historically popular in transport planning, has been criticised for focusing on the bigger picture and failing to account for the distribution of investments on the different communities of the study area (Lucas et al., 2016; Nahmias-Biran et al., 2017).
-Pereira et al. (2017) emphasize the need for a more
-egalitarian approach to transport planning. They highlight accessibility
-as a cornerstone of distributive justice, and contend that policies
-should aim to distribute investments in a way that minimizes spatial
-variations in accessibility.
+Pereira et al. (2017) emphasize the need for a more egalitarian approach to transport planning.
+They highlight accessibility as a cornerstone of distributive justice, and contend that policies should aim to distribute investments in a way that minimizes spatial variations in accessibility.
 
@@
 
-This research attempts to propose an egalitarian framework for cycling
-network design. This is done by identifying the different sub-networks
-that exist within the larger network, and ensuring that each gets a fair
-share of investment. Trip patterns in a city are not uniformly
-distributed geographically, and \emph{community finding} methods have been
-used to partition study areas into localized areas that experience a
-disproportionate number of trips within them. Akbarzadeh et al. (2018)
-use a modularity maximization approach (Blondel et al., 2008) on taxi trip
-data to identify seven different communities in the city of Isfahan, Iran.
-An optimization problem is then formulated to connect nodes within each
-community with cycling infrastructure, with the emphasis being on
-connectivity within the communities, not between them. Bao et al. (2017)
-adopt a similar methodology, first identifying communities and then
-using a greedy network expansion algorithm to simultaneously add links
-to each community. The link with the highest benefit-cost ratio in each
-community is selected, and the network is grown by adding neighboring
-links to the solution until a budget limit is met. The benefit is the
-flow on the link, and each link is assigned a cost based on current road
-conditions.
-
-Our work builds on these community finding approaches by proposing a
-similar greedy network expansion algorithm for cycle network expansion
-\emph{within communities}. We incorporate community finding methods for study
-area partitioning with weighted routing to avoid links that are
-stressful to cycle on. In doing so, we propose an approach that ensures a better distribution
-investments across the study area while also accounting for motivators and deterrents to cycling.
-We propose three sub-methods that address some of the limitations of previous
-studies. These limitations include (a) an inherent bias when basing network
-design solely on existing cycling demand, (b) the proposal of routes that may
-not correspond to studies on cyclist preference and government policies,
-and (c) an insufficient consideration of the ethical principles underlying
-the analysis. The current work addresses these limitations through the following organisation. Section \ref{calculating-potential-cycling-demand}
-focuses on calculating potential cycling demand. Section \ref{routing}
-focuses on routing the demand onto the road network while accounting for
-cyclist preferences and government priorities. Section
-\ref{road-segment-prioritization} outlines a method for partitioning the
-study area based on a community finding algorithm and routed cycling
-demand. It then introduces the network expansion algorithms, and compares an approach grounded in
+This research attempts to propose an egalitarian framework for cycling network design.
+This is done by identifying the different sub-networks that exist within the larger network, and ensuring that each gets a fair share of investment.
+Trip patterns in a city are not uniformly distributed geographically, and \emph{community finding} methods have been used to partition study areas into localized areas that experience a disproportionate number of trips within them.
+Akbarzadeh et al. (2018) use a modularity maximization approach (Blondel et al., 2008) on taxi trip data to identify seven different communities in the city of Isfahan, Iran.
+An optimization problem is then formulated to connect nodes within each community with cycling infrastructure, with the emphasis being on connectivity within the communities, not between them.
+Bao et al. (2017) adopt a similar methodology, first identifying communities and then using a greedy network expansion algorithm to simultaneously add links to each community.
+The link with the highest benefit-cost ratio in each community is selected, and the network is grown by adding neighboring links to the solution until a budget limit is met.
+The benefit is the flow on the link, and each link is assigned a cost based on current road conditions.
+
+Our work builds on these community finding approaches by proposing a similar greedy network expansion algorithm for cycle network expansion \emph{within communities}.
+We incorporate community finding methods for study area partitioning with weighted routing to avoid links that are stressful to cycle on.
+In doing so, we propose an approach that ensures a better distribution investments across the study area while also accounting for motivators and deterrents to cycling.
+We propose three sub-methods that address some of the limitations of previous studies.
+These limitations include (a) an inherent bias when basing network design solely on existing cycling demand, (b) the proposal of routes that may not correspond to studies on cyclist preference and government policies, and (c) an insufficient consideration of the ethical principles underlying the analysis.
+The current work addresses these limitations through the following organisation.
+Section \ref{calculating-potential-cycling-demand} focuses on calculating potential cycling demand.
+Section \ref{routing} focuses on routing the demand onto the road network while accounting for cyclist preferences and government priorities.
+Section \ref{road-segment-prioritization} outlines a method for partitioning the study area based on a community finding algorithm and routed cycling demand.
+It then introduces the network expansion algorithms, and compares an approach grounded in `egalitarianism' to one grounded in `utilitarianism'. ithms, and compares an approach grounded in
 `egalitarianism' to one grounded in `utilitarianism'. As mentioned above, Manchester is used as a case study.
@@
 Some use existing cycling demand, some calculate potential cycling demand, and others ignore demand completely.
-Duthie and Unnikrishnan (2014) note that relying only on existing cycling activity to prioritize cycling
-infrastructure can reinforce existing cycling patterns and ignore potential cycling demand that could be satisfied by a connected network.
-To avoid this issue they choose to ignore existing demand completely, and focus on creating a network that connects the
-entire study area.
+Duthie and Unnikrishnan (2014) note that relying only on existing cycling activity to prioritize cycling infrastructure can reinforce existing cycling patterns and ignore potential cycling demand that could be satisfied by a connected network.
+To avoid this issue they choose to ignore existing demand completely, and focus on creating a network that connects the entire study area.
 Olmos et al.
-(2020) opt to calculate potential demand instead; they obtain the distance distribution of cyclists using a smartphone-based bicycle GPS data, and then use a rejection-sampling algorithm on the OD data of the study area to match
-the potential demand distribution to the distribution obtained from GPS data.
-
-\emph{OD} data can be obtained from a range of sources,
-including GPS data, household travels surveys and census data on work
-locations. In areas where observed OD data is unavailable, modelling
-techniques such as spatial interaction models can be used to estimate
-travel volumes between zones (Black, 1995; Martínez and Viegas, 2013; Wilson, 1971). In this paper we use open access data from the UK
-census (ONS, 2011), which contains aggregate statistics on
-number of commuters between administrative zones --- Middle layer Super
-Output Areas (MSOA) --- by mode of travel (this is similar to the approach
-used by Grisé and El-Geneidy (2018)).
-MSOAs have an average
-population of just over 8000 (ONS, 2018).\footnote{ See
-  \url{https://wicid.ukdataservice.ac.uk/} for open access to the OD data.}
-Figures \ref{fig:potdemhistograms} and \ref{fig:desirefacetcycling}
-illustrate the proportion of trips cycled by distance and the
-geographic extent of the input OD dataset used in this paper.
-
-For our purposes, we use a logistic regression model to calculate
-potential cycling demand. The model was adapted from the Propensity to
-Cycle Tool (PCT), which estimates the proportion of trips
-(\(\boldsymbol{C_{p}}\)) for each OD pair that would cycle under
-different scenarios of change as a
-function of distance and hilliness (Lovelace et al., 2017). We used the Government Target scenario,
-indicating a nationwide target of doubling cycling by 2025. The
-logistic regression model used to calculate \(\boldsymbol{C_{p}}\) has
-the following parameters:
+(2020) opt to calculate potential demand instead; they obtain the distance distribution of cyclists using a smartphone-based bicycle GPS data, and then use a rejection-sampling algorithm on the OD data of the study area to match the potential demand distribution to the distribution obtained from GPS data.
+
+\emph{OD} data can be obtained from a range of sources, including GPS data, household travels surveys and census data on work locations.
+In areas where observed OD data is unavailable, modelling techniques such as spatial interaction models can be used to estimate travel volumes between zones (Black, 1995; Martínez and Viegas, 2013; Wilson, 1971).
+In this paper we use open access data from the UK census (ONS, 2011), which contains aggregate statistics on number of commuters between administrative zones --- Middle layer Super Output Areas (MSOA) --- by mode of travel (this is similar to the approach used by Grisé and El-Geneidy (2018)).
+MSOAs have an average population of just over 8000 (ONS, 2018).\footnote{See \url{https://wicid.ukdataservice.ac.uk/} for open access to the OD data.}
+Figures \ref{fig:potdemhistograms} and \ref{fig:desirefacetcycling} illustrate the proportion of trips cycled by distance and the geographic extent of the input OD dataset used in this paper.
+
+For our purposes, we use a logistic regression model to calculate potential cycling demand.
+The model was adapted from the Propensity to Cycle Tool (PCT), which estimates the proportion of trips (\(\boldsymbol{C_{p}}\)) for each OD pair that would cycle under different scenarios of change as a function of distance and hilliness (Lovelace et al., 2017).
+We used the Government Target scenario, indicating a nationwide target of doubling cycling by 2025.
+The logistic regression model used to calculate \(\boldsymbol{C_{p}}\) has the following parameters:
 
@@
 
-\noindent where \(d\) and \(s\) are the distance and slope associated with each
-OD pair, and \(\alpha\) and all \(\beta\)s are parameters calculated by a regression
-model on the training data. The square and square-root distance terms
-``capture the non-linear impact of distance on the likelihood of
-cycling'', and interaction terms capture the combined effect of slope
-and distance (Lovelace et al., 2017). Alternative cycling uptake
-models could be `plugged in' to our approach for different contexts or scenarios of change.
-
-The potential demand calculations show that the current and potential
-number of cyclists both follow a bell-shaped distribution, with the
-number of trips peaking around the 3-4km commuting distance and then
-going back down for longer distances (see Figures
-\ref{fig:potdemhistograms} and \ref{fig:desirefacetcycling}).
+\noindent where \(d\) and \(s\) are the distance and slope associated with each OD pair, and \(\alpha\) and all \(\beta\)s are parameters calculated by a regression model on the training data.
+The square and square-root distance terms ``capture the non-linear impact of distance on the likelihood of cycling'', and interaction terms capture the combined effect of slope and distance (Lovelace et al., 2017).
+Alternative cycling uptake models could be `plugged in' to our approach for different contexts or scenarios of change.
+
+The potential demand calculations show that the current and potential number of cyclists both follow a bell-shaped distribution, with the number of trips peaking around the 3-4km commuting distance and then going back down for longer distances (see Figures \ref{fig:potdemhistograms} and \ref{fig:desirefacetcycling}).
 
@@
 
-The uptake model used in this paper assumes uptake of traditional bicycles,
-which are affected by topography and distance due to physical effort.
-As discussed in Section \ref{discussion-and-conclusions}, using adapted uptake
-models could enable the approach to help plan for solutions such as e-bikes
-that enable trips covering longer distances and hillier roads.
+The uptake model used in this paper assumes uptake of traditional bicycles, which are affected by topography and distance due to physical effort.
+As discussed in Section \ref{discussion-and-conclusions}, using adapted uptake models could enable the approach to help plan for solutions such as e-bikes that enable trips covering longer distances and hillier roads.
 
@@
 
-The next step in our approach is to route the potential cycling demand
-(\(\boldsymbol{C_{p}}\)) between all OD pairs onto the road network. We
-choose not to use the PCT approach for routing as it is reliant on
-external and therefore inflexible routing services.
+The next step in our approach is to route the potential cycling demand (\(\boldsymbol{C_{p}}\)) between all OD pairs onto the road network.
+We choose not to use the PCT approach for routing as it is reliant on external and therefore inflexible routing services.
 
@@
 \item
-  \textbf{Cyclist Preference}: Work done by Dill and McNeil (2013) on examining
-  cyclist typologies determined that around 60\% of Portland residents
-  fit under the \emph{interested but concerned} category. These were people
-  that enjoyed cycling but avoided it due safety concerns. The key to
-  encouraging this group was to create a low-stress cycling network,
-  not only though segregated infrastructure but also by planning
-  routes that passed through residential streets.
+  \textbf{Cyclist Preference}: Work done by Dill and McNeil (2013) on examining cyclist typologies determined that around 60\% of Portland residents fit under the \emph{interested but concerned} category.
+These were people that enjoyed cycling but avoided it due safety concerns.
+The key to encouraging this group was to create a low-stress cycling network, not only though segregated infrastructure but also by planning routes that passed through residential streets.
 \item
-  \textbf{Low-Traffic Neighbourhoods (LTNs)}: The UK Department for Transport is
-  allocating funding to local authorities to invest in Active
-  Transport, partially through the creation of LTNs
-  (DfT, 2020a). This includes closing off residential
-  streets to motorized traffic.
+  \textbf{Low-Traffic Neighbourhoods (LTNs)}: The UK Department for Transport is allocating funding to local authorities to invest in Active Transport, partially through the creation of LTNs (DfT, 2020a).
+This includes closing off residential streets to motorized traffic.
 \item
-  \textbf{Existing Cycling Infrastructure}: Utilizing existing cycling
-  infrastructure makes economic sense, as small investments may lead
-  to large connectivity gains as the disconnected cycling
-  infrastructure gets joined together.
+  \textbf{Existing Cycling Infrastructure}: Utilizing existing cycling infrastructure makes economic sense, as small investments may lead to large connectivity gains as the disconnected cycling infrastructure gets joined together.
 \end{enumerate}
 
-The above points are accounted for by using a weighted road network for
-routing. This has been previously done by multiplying all road segments
-without cycling infrastructure by a fixed impedance factor
-(Mauttone et al., 2017), by assigning different weights to the road
-segments that are (a) proportional to the investment cost of bringing them to an
-acceptable level of stress for cycling (Duthie and Unnikrishnan, 2014), or (b) inversely proportional to
-perceived level of stress experienced by cyclists (Wysling and Purves, 2022).
-Perceived stress for cyclists has been found to be a function of vehicular traffic
-volume and speed (Sorton and Walsh, 1994), both of which vary predictably
-with road type. Gehrke et al. (2020) build on this information to use
-routing impedance factors for road segments that are based on road type
-and the existence of cycling infrastructure.
-
-For the purposes of this paper we created a similar weighting profile
-that is adjusted to favor less stressful roads (based on the definition
-by Sorton and Walsh (1994) and information from Table
-\ref{table:osmroadtypes}) and roads with existing cycling
-infrastructure. We believe this to be more appropriate than the simplistic
-approach adopted by Mauttone et al. (2017) as it makes use of a hierarchy
-of road preference based on perceived stress levels, going beyond simply
-favouring roads with existing cycling infrastructure.
-It is also different to the weighting profile used by Wysling and Purves (2022) as it adds a larger
-weight on trunk and primary roads then other roads without cycling infrastructure.
+The above points are accounted for by using a weighted road network for routing.
+This has been previously done by multiplying all road segments without cycling infrastructure by a fixed impedance factor (Mauttone et al., 2017), or by assigning different weights to the road segments that are (a) proportional to the investment cost of bringing them to an acceptable level of stress for cycling (Duthie and Unnikrishnan, 2014), or (b) inversely proportional to perceived level of stress experienced by cyclists (Wysling and Purves, 2022).
+Perceived stress for cyclists has been found to be a function of vehicular traffic volume and speed (Sorton and Walsh, 1994), both of which vary predictably with road type.
+Gehrke et al. (2020) build on this information to use routing impedance factors for road segments that are based on road type and the existence of cycling infrastructure.
+
+For the purposes of this paper we created a similar weighting profile that is adjusted to favor less stressful roads (based on the definition by Sorton and Walsh (1994) and information from Table \ref{table:osmroadtypes}) and roads with existing cycling infrastructure.
+We believe this to be more appropriate than the simplistic approach adopted by Mauttone et al.
+(2017) as it makes use of a hierarchy of road preference based on perceived stress levels, going beyond simply favouring roads with existing cycling infrastructure.
+It is also different to the weighting profile used by Wysling and Purves (2022) as it adds a larger weight on trunk and primary roads then other roads without cycling infrastructure.
 
@@
 
-\noindent A weighted distance \(d_{w}\) for each road
-segment is calculated as follows \footnote{The \textbf{dodgr} r package
-  (Padgham, 2019) is used to route cycling demand onto the road
-  network. The package uses the OpenStreetMaps (OSM) road network and
-  allows the user to assign weights to roads based on their type. The
-  routing is done based on weighted shortest paths, with the distance
-  along each road segment being divided by a factor to obtain the weighted
-  distance for routing. It is more intuitive to multiply when weighting a
-  network, but the dodgr package divides by numbers between 0 and 1, which
-  achieves the same result (road types with a weight of 0 are not routed on).
-  For the sake of reproducibility, we stick to
-  the convention used in the package.}:
+\noindent A weighted distance \(d_{w}\) for each road segment is calculated as follows \footnote{The \textbf{dodgr} r package (Padgham, 2019) is used to route cycling demand onto the road network.
+  The package uses the OpenStreetMaps (OSM) road network and allows the user to assign weights to roads based on their type.
+  The routing is done based on weighted shortest paths, with the distance along each road segment being divided by a factor to obtain the weighted distance for routing.
+  It is more intuitive to multiply when weighting a network, but the dodgr package divides by numbers between 0 and 1, which achieves the same result (road types with a weight of 0 are not routed on).
+  For the sake of reproducibility, we stick to the convention used in the package.}:
 
@@
 
-\noindent where \(d_{unw}\) is the unweighted distance and
-\(W\) is the weight from Table \ref{table:weightprofiles}.
-
-All weights are between 0 and 1, and the values in the
-\textit{Weighted} profile are chosen to be inversely proportional to
-the stress level experienced by cyclists on them. The
-\textit{Unweighted} weighting profile is used to compare increases in
-route length resulting from two different approaches:
+\noindent where \(d_{unw}\) is the unweighted distance and \(W\) is the weight from Table \ref{table:weightprofiles}.
+
+All weights are between 0 and 1, and the values in the \textit{Weighted} profile are chosen to be inversely proportional to the stress level experienced by cyclists on them.
+The \textit{Unweighted} weighting profile is used to compare increases in route length resulting from two different approaches:
 
@@
 \item
-  \textbf{Weighted}: Relatively high impedance on Primary and Trunk roads
-  (to minimize cycling on them).
+  \textbf{Weighted}: Relatively high impedance on Primary and Trunk roads (to minimize cycling on them).
 \item
@@
 
-Comparing the cycling demand routed on the weighted and unweighted road
-network allows us to get a better understanding of the importance of
-different road types. In the case of Manchester, trunk roads bisect the
-city and are a major part of unweighted shortest paths (Figure
-\ref{fig:flowsfacetunweighted}). On the other hand, cycleways are not
-part of unweighted shortest paths, and so very little of the cycling
-demand is routed through them. In the weighted network, cycleways are
-much better utilized, and the majority of the cycling demand passes
-through tertiary roads, as expected.
+Comparing the cycling demand routed on the weighted and unweighted road network allows us to get a better understanding of the importance of different road types.
+In the case of Manchester, trunk roads bisect the city and are a major part of unweighted shortest paths (Figure \ref{fig:flowsfacetunweighted}).
+On the other hand, cycleways are not part of unweighted shortest paths, and so very little of the cycling demand is routed through them.
+In the weighted network, cycleways are much better utilized, and the majority of the cycling demand passes through tertiary roads, as expected.
 
@@
 
-The results of routing potential cycling demand on the weighted and
-unweighted networks are understandably quite different. From Figure
-\ref{fig:flowsfacetunweighted} we can see that trunk and primary roads
-are the most efficient means of traversing the road network of
-Manchester. Both of these road types are classified as Primary A roads
-according to the UK Department for Transport's road classification
-(Table \ref{table:osmroadtypes}), and are therefore part of the Primary
-Route Network (PRN) (DfT, 2012). The PRN has the widest,
-most direct roads on the network, and carries most of the through
-traffic. This includes freight, with all roads in the PRN being required
-by law to provide unrestricted access to trucks up to 40 tonnes
-(DfT, 2012).
-
-We choose to avoid routing the potential cycling demand on Primary A
-Roads for the following 2 reasons:
+The results of routing potential cycling demand on the weighted and unweighted networks are understandably quite different.
+From Figure \ref{fig:flowsfacetunweighted} we can see that trunk and primary roads are the most efficient means of traversing the road network of Manchester.
+Both of these road types are classified as Primary A roads according to the UK Department for Transport's road classification (Table \ref{table:osmroadtypes}), and are therefore part of the Primary Route Network (PRN) (DfT, 2012).
+The PRN has the widest, most direct roads on the network, and carries most of the through traffic.
+This includes freight, with all roads in the PRN being required by law to provide unrestricted access to trucks up to 40 tonnes (DfT, 2012).
+
+We choose to avoid routing the potential cycling demand on Primary A Roads for the following 2 reasons:
 
@@
 \item
-  \textbf{Logistical Difficulty}: Changes on these roads need to be agreed
-  upon by all affected authorities (DfT, 2012), which
-  may prove to be difficult.
+  \textbf{Logistical Difficulty}: Changes on these roads need to be agreed upon by all affected authorities (DfT, 2012), which may prove to be difficult.
 \item
-  \textbf{Low Traffic Neighborhoods (LTNs)}: The UK government is aiming to
-  restrict access to motorized vehicles on residential roads to create
-  LTNs (DfT, 2020a). This is part of a policy to
-  prevent automobile rat-running and make streets more accessible to
-  cyclists and pedestrians. Under such a policy, Primary A roads would
-  become even more essential for motorized traffic and it would be
-  more difficult to reallocate road space on these roads to cyclists.
+  \textbf{Low Traffic Neighborhoods (LTNs)}: The UK government is aiming to restrict access to motorized vehicles on residential roads to create LTNs (DfT, 2020a).
+This is part of a policy to prevent automobile rat-running and make streets more accessible to cyclists and pedestrians.
+Under such a policy, Primary A roads would become even more essential for motorized traffic and it would be more difficult to reallocate road space on these roads to cyclists.
 \end{enumerate}
 
-Figure \ref{fig:flowsfacetweighted} shows that routing on the weighted
-network significantly reduces flow on the trunk and primary roads, but
-does not eliminate it completely. This is intentional, as the impedance
-on these roads is only slightly higher than the remaining road types (See
-Table \ref{table:weightprofiles}). Potential cycling demand is only
-routed on these roads if there are no routes through other roads that
-offer comparable directness.
-
-Banning cycling flow completely on trunk and primary roads may result in
-excessively circuitous paths, as seen in Figure
-\ref{fig:boxplotcircuity}. When routing using the \emph{weighted} weighting
-profile in Table \ref{table:weightprofiles}, we see that shortest paths
-increase by less than 5\% on average from unweighted shortest paths, with
-the largest increases still below 30\%. When routing on primary and trunk
-roads is banned (\emph{weighted\_2} profile in Table
-\ref{table:weightprofiles}), the average increase relative to
-unweighted shortest paths rises to 10\%, with certain locations
-experiencing more significant negative effects on accessibility. Given
-that cyclists will only deviate from shortest paths by a certain amount
-to access better cycling infrastructure (as explained in Section
-\ref{introduction}), allowing flow on some stretches of trunk and
-primary roads is necessary to insure cycling uptake and equitable access
-to cycling infrastructure. In its new vision for walking and cycling,
-the Department for Transport acknowledges that minimal segregated
-stretches of bicycle lanes on main roads will be necessary to avoid
-circuitous cycling networks (DfT, 2020a).
+Figure \ref{fig:flowsfacetweighted} shows that routing on the weighted network significantly reduces flow on the trunk and primary roads, but does not eliminate it completely.
+This is intentional, as the impedance on these roads is only slightly higher than the remaining road types (See Table \ref{table:weightprofiles}).
+Potential cycling demand is only routed on these roads if there are no routes through other roads that offer comparable directness.
+
+Banning cycling flow completely on trunk and primary roads may result in excessively circuitous paths, as seen in Figure \ref{fig:boxplotcircuity}.
+When routing using the \emph{weighted} weighting profile in Table \ref{table:weightprofiles}, we see that shortest paths increase by less than 5\% on average from unweighted shortest paths, with the largest increases still below 30\%.
+When routing on primary and trunk roads is banned (\emph{weighted\_2} profile in Table \ref{table:weightprofiles}), the average increase relative to unweighted shortest paths rises to 10\%, with certain locations experiencing more significant negative effects on accessibility.
+Given that cyclists will only deviate from shortest paths by a certain amount to access better cycling infrastructure (as explained in Section \ref{introduction}), allowing flow on some stretches of trunk and primary roads is necessary to insure cycling uptake and equitable access to cycling infrastructure.
+In its new vision for walking and cycling, the Department for Transport acknowledges that minimal segregated stretches of bicycle lanes on main roads will be necessary to avoid circuitous cycling networks (DfT, 2020a).
 
@@
 
-Weighting the road network also allows us to better utilize existing
-cycling infrastructure, as can be seen by the higher flow on cycleways
-in Figure \ref{fig:flowsfacetweighted}. Again, the small differences in
-impedance between cycleways and other road types mean that cycleways
-that require significant deviation are not routed on.
-
-It should be reiterated that the weighting profile used for routing has
-been developed for the purposes of this study. It creates a hierarchy of
-road preference that is grounded in cyclist preferences and government
-plans to create LTNs. A wide range of weighting profiles could be used
-to represent different types of cyclists and road environments, as
-described in Gehrke et al. (2020) and Furth et al. (2016). A sensitivity
-analysis could be done to determine an optimal weighting profile, but
-given the variation in city road
-networks, this would probably
-require calibration to the specific city. More accurate routing could be
-carried out given the availability of road-level data. In such cases we
-would add additional impedance to specific roads, giving more useful
-routing results than the current approach which considers all roads
-of the same type to be equivalent.
-
-One use-case of such granular data would be to identify roads that serve
-schools. The Department of Transport notes that the number of school
-children being driven to school has trebled over the past 40 years
-(DfT, 2020a), and so having cycling infrastructure
-serving schools is key to achieving the government target of getting
-more children to cycle. This would not be difficult, as over 75\% of
-children in the UK live within a 15 minute cycle from their school
-(DfT, 2020b). Goodman et al. (2019) show that if
-dutch levels of cycling were achieved in the UK, the \% of children
-cycling to school could increase from 1.8\% to 41\%. In their typology of
-cyclists, Dill and McNeil (2013) found that a majority of people who say they
-would never cycle had never cycled to school, whereas confident cyclists
-were those most likely to have cycled to school. Getting people to cycle
-from a young age is therefore key to achieving societal change in
-commuting habits.
+Weighting the road network also allows us to better utilize existing cycling infrastructure, as can be seen by the higher flow on cycleways in Figure \ref{fig:flowsfacetweighted}.
+Again, the small differences in impedance between cycleways and other road types mean that cycleways that require significant deviation are not routed on.
+
+It should be reiterated that the weighting profile used for routing has been developed for the purposes of this study.
+It creates a hierarchy of road preference that is grounded in cyclist preferences and government plans to create LTNs.
+A wide range of weighting profiles could be used to represent different types of cyclists and road environments, as described in Gehrke et al. (2020) and Furth et al. (2016).
+A sensitivity analysis could be done to determine an optimal weighting profile, but given the variation in city road networks, this would probably require calibration to the specific city.
+More accurate routing could be carried out given the availability of road-level data.
+In such cases we would add additional impedance to specific roads, giving more useful routing results than the current approach which considers all roads of the same type to be equivalent.
+
+One use-case of such granular data would be to identify roads that serve schools.
+The Department of Transport notes that the number of school children being driven to school has trebled over the past 40 years (DfT, 2020a), and so having cycling infrastructure serving schools is key to achieving the government target of getting more children to cycle.
+This would not be difficult, as over 75\% of children in the UK live within a 15 minute cycle from their school (DfT, 2020b).
+Goodman et al. (2019) show that if dutch levels of cycling were achieved in the UK, the \% of children cycling to school could increase from 1.8\% to 41\%.
+In their typology of cyclists, Dill and McNeil (2013) found that a majority of people who say they would never cycle had never cycled to school, whereas confident cyclists were those most likely to have cycled to school.
+Getting people to cycle from a young age is therefore key to achieving societal change in commuting habits.
 
@@
 
-After routing the potential cycling demand onto the road network using
-weighted shortest paths, we have estimates for the cumulative potential
-cycling demand passing through all road segments. This cumulative demand
-(referred to as \emph{flow}) is then used as a basis for identifying segments
-that are in most need of investment in segregated cycling
-infrastructure. In doing so, we must account for the motivations and
-deterrents for cycling identified in Section \ref{introduction}, namely
-direct and well-connected routes.
-
-A range of algorithms could be used for prioritization. Because policy
-priorities vary, we present two algorithms. Both utilize existing
-infrastructure from the beginning and allow us to compare a solution
-that focuses on utilitarianism to one that focuses on egalitarianism. In
-both algorithms, links are selected iteratively and the iteration at
-which each link is added to the solution is recorded. Investments in
-cycling infrastructure can be limited by budget constraints, so it can
-be useful to see where best to allocate a defined length of segregated
-infrastructure. In order to incorporate egalitarian principles in our
-approach, we use community detection to partition the study area and
-distribute investment over the different subdivisions.
+After routing the potential cycling demand onto the road network using weighted shortest paths, we have estimates for the cumulative potential cycling demand passing through all road segments.
+This cumulative demand (referred to as \emph{flow}) is then used as a basis for identifying segments that are in most need of investment in segregated cycling infrastructure.
+In doing so, we must account for the motivations and deterrents for cycling identified in Section \ref{introduction}, namely direct and well-connected routes.
+
+A range of algorithms could be used for prioritization.
+Because policy priorities vary, we present two algorithms.
+Both utilize existing infrastructure from the beginning and allow us to compare a solution that focuses on utilitarianism to one that focuses on egalitarianism.
+In both algorithms, links are selected iteratively and the iteration at which each link is added to the solution is recorded.
+Investments in cycling infrastructure can be limited by budget constraints, so it can be useful to see where best to allocate a defined length of segregated infrastructure.
+In order to incorporate egalitarian principles in our approach, we use community detection to partition the study area and distribute investment over the different subdivisions.
 
@@
 
-As explained in Section
-\ref{ethical-underpinnings-and-proposed-approach}, a major challenge
-facing `top-down' planning approaches is how to incorporate egalitarian
-principles by fairly distributing investments in cycling infrastructure.
-One way of quantifying this is to split up the city into smaller
-geospatial areas and target equal investment in each of those areas.
-This approach could also help ensure that on-the-ground surveys are made
-by local stakeholders, an important component of the planning process
-(Parkin, 2018). Community detection offers us a way to
-delineate such a split; cyclists are limited in their commuting distance
-(see Figure \ref{fig:cyclinghistmanc}), and so trip attractors are more
-likely to have a local catchment area of cyclists.
+As explained in Section \ref{ethical-underpinnings-and-proposed-approach}, a major challenge facing `top-down' planning approaches is how to incorporate egalitarian principles by fairly distributing investments in cycling infrastructure.
+One way of quantifying this is to split up the city into smaller geospatial areas and target equal investment in each of those areas.
+This approach could also help ensure that on-the-ground surveys are made by local stakeholders, an important component of the planning process (Parkin, 2018).
+Community detection offers us a way to delineate such a split; cyclists are limited in their commuting distance (see Figure \ref{fig:cyclinghistmanc}), and so trip attractors are more likely to have a local catchment area of cyclists.
 
@@
 
-In our case, the network is the city; the nodes are the
-population-weighted MSOA centroids and the links connecting each MSOA
-pair are weighted by the potential cycling demand between them. The
-Louvain method (Blondel et al., 2008) is used to separate MSOAs into
-communities. Potential cycling demand is used since we assume that this
-is what the cycling demand will be once the cycling infrastructure is
-added. To assign road links to communities, the following steps are
-carried out:
+In our case, the network is the city; the nodes are the population-weighted MSOA centroids and the links connecting each MSOA pair are weighted by the potential cycling demand between them.
+The Louvain method (Blondel et al., 2008) is used to separate MSOAs into communities.
+Potential cycling demand is used since we assume that this is what the cycling demand will be once the cycling infrastructure is added.
+To assign road links to communities, the following steps are carried out:
 
@@
 
-The results show that Manchester can be split into four large
-communities and one small one (Figure \ref{fig:communitiesmanchester}).
+The results show that Manchester can be split into four large communities and one small one (Figure \ref{fig:communitiesmanchester}).
 
@@
 
-This algorithm ensures that the resulting network is connected. It also
-satisfies the directness criteria, since links on the weighted shortest
-paths are those that have the highest flow passing through them (this is
-a result of the routing in Section \ref{routing}.
+This algorithm ensures that the resulting network is connected.
+It also satisfies the directness criteria, since links on the weighted shortest paths are those that have the highest flow passing through them (this is a result of the routing in Section \ref{routing}.
 
@@
 
-The first algorithm focuses on connectivity and directness, but not on
-fairly distributing investment. The latter is not a requirement for
-increasing cycling uptake, but it is fundamental for spatial equity, as
-explained in Section \ref{ethical-underpinnings-and-proposed-approach}. This algorithm incorporates
-the ideal of fair distribution by ensuring that investment is distributed between the defined communities.
+The first algorithm focuses on connectivity and directness, but not on fairly distributing investment.
+The latter is not a requirement for increasing cycling uptake, but it is fundamental for spatial equity, as explained in Section \ref{ethical-underpinnings-and-proposed-approach}.
+This algorithm incorporates the ideal of fair distribution by ensuring that investment is distributed between the defined communities.
 This is done using the following logic:
@@
 
-Even though we may end up with a more disconnected network, we will have
-separate connected networks in each community. Given that communities
-are defined by having more internal flow than external flow, this is a
-satisfactory solution.
-
-The results of the community detection are used to evaluate the
-algorithms. This is done by looking at the \emph{person-km satisfied} as
-cycling infrastructure is added. Person-km is a measure of the total km
-cycled on a road segment, so it is the product of the number of
-potential commuters cycling on that road segment (\(flow\)) and the length
-of the segment in km (\(l\)). For each road segment, the person-km is
-equal to \(flow * l\). In the case of Manchester, Table
-\ref{tab:personkmtable} shows that almost half of the person-km is in
-community 1, while only 0.5\% of total person-km on the network is in
-community 5.
+Even though we may end up with a more disconnected network, we will have separate connected networks in each community.
+Given that communities are defined by having more internal flow than external flow, this is a satisfactory solution.
+
+The results of the community detection are used to evaluate the algorithms.
+This is done by looking at the \emph{person-km satisfied} as cycling infrastructure is added.
+Person-km is a measure of the total km cycled on a road segment, so it is the product of the number of potential commuters cycling on that road segment (\(flow\)) and the length of the segment in km (\(l\)).
+For each road segment, the person-km is equal to \(flow * l\).
+In the case of Manchester, Table \ref{tab:personkmtable} shows that almost half of the person-km is in community 1, while only 0.5\% of total person-km on the network is in community 5.
 
@@
 
-Looking at the person-km satisfied (Figure \ref{fig:growthtotal}), we
-see that the incremental addition of cycling infrastructure is better
-distributed between communities using Algorithm 2; equal distribution of
-investment results in the gain in \% of person km satisfied in each
-community being inversely correlated with the size of the community. In
-addition, we find that the restrictions imposed by Algorithm 2 on the
-network expansion do not seem to have a noticeable effect on the
-city-wide \% of person-km satisfied. Comparing both algorithms, we can
-see that Algorithm 1 provides only marginally quicker city-wide gains
-than Algorithm 2.
+Looking at the person-km satisfied (Figure \ref{fig:growthtotal}), we see that the incremental addition of cycling infrastructure is better distributed between communities using Algorithm 2; equal distribution of investment results in the gain in \% of person km satisfied in each community being inversely correlated with the size of the community.
+In addition, we find that the restrictions imposed by Algorithm 2 on the network expansion do not seem to have a noticeable effect on the city-wide \% of person-km satisfied.
+Comparing both algorithms, we can see that Algorithm 1 provides only marginally quicker city-wide gains than Algorithm 2.
 
@@
 
-Figure \ref{fig:growth3MapandBar} gives us a geographic representation
-of the results from Algorithm 2; it shows when each link was added to
-the solution (first 100km, second 100km, etc). We can see that,
-generally, road segments around cycling infrastructure are prioritized,
-except for those neighboring cycling infrastructure on the very
-periphery. The first 100km is also spatially distributed across the
-city, with no apparent bias towards a particular area.
-
-It is also important to understand how the different highway types
-contribute to the proposed network. Figure \ref{fig:growth3MapandBar}
-shows that most of the flow will be on residential and tertiary roads,
-as expected from the weighting profile defined in Table
-\ref{table:weightprofiles}.
+Figure \ref{fig:growth3MapandBar} gives us a geographic representation of the results from Algorithm 2; it shows when each link was added to the solution (first 100km, second 100km, etc).
+We can see that, generally, road segments around cycling infrastructure are prioritized, except for those neighboring cycling infrastructure on the very periphery.
+The first 100km is also spatially distributed across the city, with no apparent bias towards a particular area.
+
+It is also important to understand how the different highway types contribute to the proposed network.
+Figure \ref{fig:growth3MapandBar} shows that most of the flow will be on residential and tertiary roads, as expected from the weighting profile defined in Table \ref{table:weightprofiles}.
 
@@
 
-Existing cycling infrastructure is made up of many disconnected
-components. Both Algorithm 1 and 2 start with all existing segregated
-cycling infrastructure and aim to create an efficient, connected
-network. Figure \ref{fig:componentsandGCC} shows that both algorithms
-gradually reduce the number of components as more infrastructure is
-added, but Algorithm 2 is able to provide better connectivity with less
-investment.
-
-Consistent growth can also be seen for the size of the Largest Connected
-Component in the proposed bicycle network (Figure
-\ref{fig:componentsandGCC}). Here, however, we find that there is little
-difference between both algorithms.
+Existing cycling infrastructure is made up of many disconnected components.
+Both Algorithm 1 and 2 start with all existing segregated cycling infrastructure and aim to create an efficient, connected network.
+Figure \ref{fig:componentsandGCC} shows that both algorithms gradually reduce the number of components as more infrastructure is added, but Algorithm 2 is able to provide better connectivity with less investment.
+
+Consistent growth can also be seen for the size of the Largest Connected Component in the proposed bicycle network (Figure \ref{fig:componentsandGCC}).
+Here, however, we find that there is little difference between both algorithms.
 
@@
 
-Overall, the improved distribution of investment shown by Algorithm 2 does not seem to come at the cost
-of poorer connectivity or less city-wide gains when compared to Algorithm 1.
+Overall, the improved distribution of investment shown by Algorithm 2 does not seem to come at the cost of poorer connectivity or less city-wide gains when compared to Algorithm 1.
 The results therefore advocate for the incorporation of egalitarian principles in bicycle network planning.
@@
 
-This paper demonstrates an approach for prioritizing
-investments in city-wide cycleway networks. The approach aims to respect
-both the needs of the users and the ambitions of stakeholders working at
-local or regional levels. The results, including detailed route network
-maps based on current travel behaviour derived from OD data, can provide
-an evidence-base for designing direct, connected,
-and low-stress networks. A core contribution of this approach is that it insures
-that investments in cycling infrastructure are distributed fairly between different
-communities in a city.
-
-Given that the ``most essential activity entailed in the design of
-cycle-friendly infrastructure is developing a cycle network''
-(Parkin, 2018), we believe that the approach has great
-potential to inform investment in cities such as Manchester where there
-is political will to invest in cycling long-term. A benefit of the
-approach is that it has relatively modest data requirements: only the
-road network (from OSM), its topography (from satellite imagery), and
-OD data (from the national census in this case) are
-needed, opening up the possibility of deploying the approach in other
-cities\footnote{The results are easily reproducible for all UK cities, and can also be reproduced for cities elsewhere given the availability of commuter data.
+This paper demonstrates an approach for prioritizing investments in city-wide cycleway networks.
+The approach aims to respect both the needs of the users and the ambitions of stakeholders working at local or regional levels.
+The results, including detailed route network maps based on current travel behaviour derived from OD data, can provide an evidence-base for designing direct, connected, and low-stress networks.
+A core contribution of this approach is that it insures that investments in cycling infrastructure are distributed fairly between different communities in a city.
+
+Given that the ``most essential activity entailed in the design of cycle-friendly infrastructure is developing a cycle network'' (Parkin, 2018), we believe that the approach has great potential to inform investment in cities such as Manchester where there is political will to invest in cycling long-term.
+A benefit of the approach is that it has relatively modest data requirements: only the road network (from OSM), its topography (from satellite imagery), and OD data (from the national census in this case) are needed, opening up the possibility of deploying the approach in other cities\footnote{The results are easily reproducible for all UK cities, and can also be reproduced for cities elsewhere given the availability of commuter data.
   Instructions for reproducing the results are detailed in the \textbf{README} file in the R folder of the .zip file that accompanies this paper (Link to public github repo to be added when paper is published)}.
 
-The approach can identify not only where there is high
-potential for cycling but also trade-offs between
-stress levels (related to motor traffic) and directness.
-The results can therefore be used as a basis for recommendations on
-road space reallocation \emph{and} new infrastructure to unlock potential cycling demand.
-The approach encourages consideration of a wider range of preferences
-and needs than previous approaches that focus only on absolute potential.
-Moreover, the inclusion of egalitarian principles in scenarios of change
-encourages investment in cycling infrastructure that addresses geographical and
-social inequalities. This is done by explicitly ensuring that infrastructure improvements
-are pursued in different parts of the city simultaneously. This approach compares well with one
-that focuses exclusively on maximizing the number of potential cycling trips that
-benefit from investment, showing that a fair distribution can be achieved without
-compromising on overall benefit.
-This ability to address inequalities in network prioritization is particularly
-important given research showing substantial \emph{inequalities}
-around transport provision in general and cycling uptake and investment
-in particular (Lucas et al., 2016; Vidal Tortosa et al., 2021).
-
-The approach is not without limitations. The level of detail is only as
-good as the granularity of the available data (in this case relatively
-coarse zones). Iacono et al. (2010) note that such large travel zones
-are not ideal for understanding route choice behaviour of cyclists and
-pedestrians. They also give rise to an `ecological fallacy' whereby
-average characteristics are assumed to apply to all residents of the
-aggregated geographical area, suggesting a need for applying the methods
-to more granular OD data (and for governments and other
-data-collecting organisations to make OD data more
-readily available).
-The potential demand calculation is also based on cycling in the traditional sense, and does not consider the
-effect of micro-mobility on reducing topology-related impedance to cycling.
+The approach can identify not only where there is high potential for cycling but also trade-offs between stress levels (related to motor traffic) and directness.
+The results can therefore be used as a basis for recommendations on road space reallocation \emph{and} new infrastructure to unlock potential cycling demand.
+The approach encourages consideration of a wider range of preferences and needs than previous approaches that focus only on absolute potential.
+Moreover, the inclusion of egalitarian principles in scenarios of change encourages investment in cycling infrastructure that addresses geographical and social inequalities.
+This is done by explicitly ensuring that infrastructure improvements are pursued in different parts of the city simultaneously.
+This approach compares well with one that focuses exclusively on maximizing the number of potential cycling trips that benefit from investment, showing that a fair distribution can be achieved without compromising on overall benefit.
+This ability to address inequalities in network prioritization is particularly important given research showing substantial \emph{inequalities} around transport provision in general and cycling uptake and investment in particular (Lucas et al., 2016; Vidal Tortosa et al., 2021).
+
+The approach is not without limitations.
+The level of detail is only as good as the granularity of the available data (in this case relatively coarse zones).
+Iacono et al. (2010) note that such large travel zones are not ideal for understanding route choice behaviour of cyclists and pedestrians.
+They also give rise to an `ecological fallacy' whereby average characteristics are assumed to apply to all residents of the aggregated geographical area, suggesting a need for applying the methods to more granular OD data (and for governments and other data-collecting organisations to make OD data more readily available).
+The potential demand calculation is also based on cycling in the traditional sense, and does not consider the effect of micro-mobility on reducing topology-related impedance to cycling.
 Given that we are proposing an approach which can accommodate any cycling uptake functions, this is an acceptable limitation.
@@
 
-The approach is also focused solely on the allocation of cycling
-infrastructure, and does not consider the larger political and
-regulatory environment that needs to exist to promote cycling; while
-segregated, connected, and direct cycling infrastructure is key to
-achieving high levels of cycling, research has shown that it cannot
-exist in a vacuum. Wardman et al. (2007) developed a mode choice model
-for the UK and their results showed that improved cycling infrastructure
-on its own only had modest impacts on mode shift, and even the unlikely
-scenario of all urban routes being serviced by segregated bike lanes was
-forecast to increase cycling mode share by only 3\%. International
-research shows that cities that invest in more comprehensive cycling
-projects have a more significant increase in the number of cyclists as
-well as the cycling mode share (Parkin, 2018; Pucher et al., 2010). These cities do not just focus on
-infrastructure, but on general policies as well as restricting car use.
-Evaluation of policies in Denmark and Germany and the Netherlands has
-shown that their high cycling mode share is down to a broader set of
-soft and hard policies. Hard policies include traffic calming, filtered permeability
-interventions, cycling rights of way, bike parking, integration with the
-public transport network, and making driving cars both expensive and
-inconvenient, while soft policies include marketing and awareness campaigns (Gössling, 2013; Pucher and Buehler, 2008).
-While these policies are outside the scope of this research, it is important to recognize their key role in
-bringing about an increase in levels of cycling.
+The approach is also focused solely on the allocation of cycling infrastructure, and does not consider the larger political and regulatory environment that needs to exist to promote cycling; while segregated, connected, and direct cycling infrastructure is key to achieving high levels of cycling, research has shown that it cannot exist in a vacuum.
+Wardman et al.
+(2007) developed a mode choice model for the UK and their results showed that improved cycling infrastructure on its own only had modest impacts on mode shift, and even the unlikely scenario of all urban routes being serviced by segregated bike lanes was forecast to increase cycling mode share by only 3\%.
+International research shows that cities that invest in more comprehensive cycling projects have a more significant increase in the number of cyclists as well as the cycling mode share (Parkin, 2018; Pucher et al., 2010).
+These cities do not just focus on infrastructure, but on general policies as well as restricting car use.
+Evaluation of policies in Denmark and Germany and the Netherlands has shown that their high cycling mode share is down to a broader set of soft and hard policies.
+Hard policies include traffic calming, filtered permeability interventions, cycling rights of way, bike parking, integration with the public transport network, and making driving cars both expensive and inconvenient, while soft policies include marketing and awareness campaigns (Gössling, 2013; Pucher and Buehler, 2008).
+While these policies are outside the scope of this research, it is important to recognize their key role in bringing about an increase in levels of cycling.
 
 Consideration of these limitations suggest future directions of research.
-New datasets and continuously evolving computational (hardware and software)
-capabilities should enable the data related issues to be overcome as open
-datasets, and our ability to process them, improve.
-Plugging-in alternative uptake models could help address the relatively narrow
-definition of `potential' used in this paper, to consider broader social factors.
-A promising area of future research in this direction could be
-to integrate a wider range of modes, including e-bikes and e-scooters, into the
-analysis. Currently scope for progress in this direction is
-partially restricted by the lack of data on the
-proliferation of these modes, raising the point that surveys could
-replace the broad category of `Bicycle' with a range of small modes such as
-`pedal cycles' (including tricycles and recumbent cycles), e-bikes
-and other forms of micro-mobility.
-
-Despite the rapid growth of these alternative modes, there is little doubt that
-the humble bicycle is a key ingredient in the ongoing shift towards active transport.
-Recognition of the importance of this shift for improved health and well-being
-of citizens, and the wider challenge to decarbonise the global economy,
-has grown with pressures on health and public transport systems during the pandemic.
-The success of policies to accelerate cycling uptake depends on a range of
-factors including, vitally, the design of the
-cycling network for potential cyclists.
-The approach presented in this paper provides a strong evidence base,
-that considers both cycling potential and social equity,
-for designing joined-up and cost-effective strategic cycle networks.
+New datasets and continuously evolving computational (hardware and software) capabilities should enable the data related issues to be overcome as open datasets, and our ability to process them, improve.
+Plugging-in alternative uptake models could help address the relatively narrow definition of `potential' used in this paper, to consider broader social factors.
+A promising area of future research in this direction could be to integrate a wider range of modes, including e-bikes and e-scooters, into the analysis.
+Currently scope for progress in this direction is partially restricted by the lack of data on the proliferation of these modes, raising the point that surveys could replace the broad category of `Bicycle' with a range of small modes such as `pedal cycles' (including tricycles and recumbent cycles), e-bikes and other forms of micro-mobility.
+
+Despite the rapid growth of these alternative modes, there is little doubt that the humble bicycle is a key ingredient in the ongoing shift towards active transport.
+Recognition of the importance of this shift for improved health and well-being of citizens, and the wider challenge to decarbonise the global economy, has grown with pressures on health and public transport systems during the pandemic.
+The success of policies to accelerate cycling uptake depends on a range of factors including, vitally, the design of the cycling network for potential cyclists.
+The approach presented in this paper provides a strong evidence base, that considers both cycling potential and social equity, for designing joined-up and cost-effective strategic cycle networks.
 

### 2023-04-24 README.Rmd
cleaning up typos

@@
 output:
- bookdown::github_document2:
-   # bookdown::pdf_document2:
+ # bookdown::github_document2:
+   bookdown::pdf_document2:
  # bookdown::word_document2:
@@
 Section \@ref(road-segment-prioritization) outlines a method for partitioning the study area based on a community finding algorithm and routed cycling demand.
-It then introduces the network expansion algorithms, and compares an approach grounded in 'egalitarianism' to one grounded in 'utilitarianism'. ithms, and compares an approach grounded in
-'egalitarianism' to one grounded in 'utilitarianism'. As mentioned above, Manchester is used as a case study. 
+It then introduces the network expansion algorithms, and compares an approach grounded in 'egalitarianism' to one grounded in 'utilitarianism'. 
+As mentioned above, Manchester is used as a case study. 
 An overview of the steps involved is shown in Figure \@ref(fig:methodflowchart).

### 2023-05-01 README.tex
minor edit: equity literature

@@
 Grisé and El-Geneidy (2018) developed a multi-criteria approach, combined with estimated levels of usage, to highlight routes most in need of investment in Quebec City.
-Wysling and Purves (2022) present a method that used open data from Paris to generate cycling suitability ratings based on `bicycle suitability multipliers' associated with different types of road across the city, to support more effective interventions.
+Wysling and Purves (2022) present a method that used open data from Paris to generate cycling suitability ratings, based on `bicycle suitability multipliers' associated with different types of road across the city, to support more effective interventions.
 
-None of the aforementioned studies explore the equity impacts of different network prioritization approaches.
-The following section highlights the importance of considering (often unstated) ethical assumptions underlying network optimisation and prioritization approaches of the type outlined above.
+With the exception of Grisé and El-Geneidy (2018), the aforementioned studies do not explore the equity impacts of different network prioritization approaches.
+The following section highlights the importance of considering (often unstated) ethical assumptions underlying network optimization and prioritization approaches of the type outlined above.
 
@@
 Section \ref{road-segment-prioritization} outlines a method for partitioning the study area based on a community finding algorithm and routed cycling demand.
-It then introduces the network expansion algorithms, and compares an approach grounded in `egalitarianism' to one grounded in `utilitarianism'. ithms, and compares an approach grounded in
-`egalitarianism' to one grounded in `utilitarianism'. As mentioned above, Manchester is used as a case study.
+It then introduces the network expansion algorithms, and compares an approach grounded in `egalitarianism' to one grounded in `utilitarianism'.
+As mentioned above, Manchester is used as a case study.
 An overview of the steps involved is shown in Figure \ref{fig:methodflowchart}.

### 2023-09-22 README.Rmd
Add demographic data and surroundings

--- 2023-04-24 README.Rmd
@@
 
+```{r, include=FALSE}
+library(tidyverse)
+library(tmap)
+knitr::opts_chunk$set(warning = FALSE)
+```
+
 <!-- Highlights -->
@@
 Segregated cycling infrastructure --- road space that is allocated to cyclists only with physical separation to protect cyclists --- is associated with cycling uptake [@aldred2019impacts; @goodman2014new; @marques2015infrastructure].
-Separation from motorized vehicles has found to be key [@winters2011motivators].
+Separation from motorized vehicles has been found to be key [@winters2011motivators].
 Revealed preference of cyclists shows that they are willing to deviate from the most efficient routes in order to commute on safer roads [@crane2017longitudinal].
@@
 @grise2018if developed a multi-criteria approach, combined with estimated levels of usage, to highlight routes most in need of investment in Quebec City.
-@wysling2022improve present a method that used open data from Paris to generate cycling suitability ratings based on 'bicycle suitability multipliers' associated with different types of road across the city, to support more effective interventions.
+@wysling2022improve present a method that used open data from Paris to generate cycling suitability ratings, based on 'bicycle suitability multipliers' associated with different types of road across the city, to support more effective interventions.
 
-None of the aforementioned studies explore the equity impacts of different network prioritization approaches.
-The following section highlights the importance of considering (often unstated) ethical assumptions underlying network optimisation and prioritization approaches of the type outlined above.
+With the exception of @grise2018if, the aforementioned studies do not explore the equity impacts of different network prioritization approaches.
+The following section highlights the importance of considering (often unstated) ethical assumptions underlying network optimization and prioritization approaches of the type outlined above.
 
@@
 Section \@ref(road-segment-prioritization) outlines a method for partitioning the study area based on a community finding algorithm and routed cycling demand.
-It then introduces the network expansion algorithms, and compares an approach grounded in 'egalitarianism' to one grounded in 'utilitarianism'. ithms, and compares an approach grounded in
-'egalitarianism' to one grounded in 'utilitarianism'. As mentioned above, Manchester is used as a case study. 
+It then introduces the network expansion algorithms, and compares an approach grounded in 'egalitarianism' to one grounded in 'utilitarianism'. 
 An overview of the steps involved is shown in Figure \@ref(fig:methodflowchart).
@@
 knitr::include_graphics("extras/methodology_flowchart.png")
+```
+
+## Case study area
+
+Manchester is a city in the North West of England.
+"Manchester" can be used to refer to multiple things:
+
+- Manchester City Council, which is a local government district that includes the city centre and surrounding areas.
+- The Greater Manchester region, which is a metropolitan county that includes the city of Manchester and 9 other local government districts.
+- The city of Manchester, which is used by visitors to refer to the city centre and surrounding areas.
+This definition if formalised in the Major Towns and Cities (TCITY) dataset provided by the Office for National Statistics.
+
+We use the TCITY definition of Manchester, which was "developed specifically for the production and analysis of statistics" and which is shown in Figure \@ref(fig:manchestermap).
+Manchester is a cosmopolitan city with a large university population.
+As shown in Figure \@ref(fig:manchestermap) (right), the high density of the central area allows high levels of cycling, although the surrounding areas have low levels of cycling.
+Manchester and its surrounding have many deprived areas.
+
+```{r manchestermap, echo=FALSE, fig.cap="Maps of Greater Manchester (top) and the case study area (City of Manchester)"}
+greater_manchester = pct::get_pct_zones(region = "greater-manchester", geography = "msoa")
+greater_manchester = greater_manchester |>
+  mutate(pcycle = bicycle / all * 100)
+# names(greater_manchester)
+msoas = sf::read_sf("./data/Manchester/msoas_geometry.shp")
+msoas_union = sf::st_union(msoas)
+greater_manchester_lads = greater_manchester |>
+  group_by(lad_name) |>
+  summarise()
+?tm_scale_bar
+m1 = tm_shape(greater_manchester_lads) +
+  tm_polygons(col = "lad_name", palette = "Set1", alpha = 0.2, title = "") +
+  tm_layout(legend.position = c("left", "bottom")) +
+  tm_shape(msoas_union) +
+  tm_borders(col = "black", lwd = 2) +
+  tm_scale_bar(breaks = (0:3)*5)
+msoas_centroids = sf::st_centroid(msoas)
+pct_zones_in_manchester = greater_manchester[msoas_centroids, ] 
+m2 = tm_shape(greater_manchester, bb = msoas) +
+  tm_fill("pcycle", palette = "Blues", title = "% cycling to work\n(2011 Census)") +
+  tm_layout(legend.position = c("left", "bottom"), legend.bg.color = "white", legend.bg.alpha = 0.5) +
+  tm_shape(msoas_union) +
+  tm_borders(col = "black", lwd = 2) +
+  tm_scale_bar(breaks = 0:5)
+tmap_arrange(m1, m2, nrow = 1)
 ```

### 2023-09-22 README.Rmd

@@
 tmap_arrange(m1, m2, nrow = 1)
+```
+
+In the context of other cities defined in the TCITY dataset, Manchester is a large city, with a mid-level of cycling (see Figure \@ref(fig:manchestercyclinglevels)).
+
+```{r manchestercyclinglevels, echo=FALSE, fig.cap="Cycling levels in the case study area (City of Manchester)"}
+knitr::include_graphics("./extras/uk_cities_mode_share.png")
 ```

### 2023-09-22 README.tex

@@
 \usepackage{amsmath,amssymb}
-\usepackage{lmodern}
 \usepackage{iftex}
@@
 \else % if luatex or xetex
-  \usepackage{unicode-math}
+  \usepackage{unicode-math} % this also loads fontspec
   \defaultfontfeatures{Scale=MatchLowercase}
   \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+\usepackage{lmodern}
+\ifPDFTeX\else
+  % xetex/luatex font selection
 \fi
@@
 \IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\urlstyle{same} % disable monospaced font for URLs
+\urlstyle{same}
 \hypersetup{
@@
 The approach, which can be applied to other cities, is demonstrated in a case study of Manchester, resulting in cycling networks that balance directness against the need for safe and stress-free routes under different investment scenarios.
-A key benefit of the approach from a policy perspective is its ability to support egalitarian and cost-effective strategic cycle network planning.
-¶ \par ¶ \textbf{Keywords:} cycling networks, low-traffic neighborhoods, routing, transport equity
+A key benefit of the approach from a policy perspective is its ability to support egalitarian and cost-effective strategic cycle network planning.\\
+\par\\
+\textbf{Keywords:} cycling networks, low-traffic neighborhoods, routing, transport equity
 \end{abstract}
@@
 Segregated cycling infrastructure --- road space that is allocated to cyclists only with physical separation to protect cyclists --- is associated with cycling uptake (Aldred et al., 2019; Goodman et al., 2014; Marqués et al., 2015).
-Separation from motorized vehicles has found to be key (Winters et al., 2011).
+Separation from motorized vehicles has been found to be key (Winters et al., 2011).
 Revealed preference of cyclists shows that they are willing to deviate from the most efficient routes in order to commute on safer roads (Crane et al., 2017).
@@
 It then introduces the network expansion algorithms, and compares an approach grounded in `egalitarianism' to one grounded in `utilitarianism'.
+
 As mentioned above, Manchester is used as a case study.
@@
 \end{figure}
+
+\hypertarget{case-study-area}{%
+\subsection{Case study area}\label{case-study-area}}
+
+Manchester is a city in the North West of England.
+``Manchester'' can be used to refer to multiple things:
+
+\begin{itemize}
+\tightlist
+\item
+  Manchester City Council, which is a local government district that includes the city centre and surrounding areas.
+\item
+  The Greater Manchester region, which is a metropolitan county that includes the city of Manchester and 9 other local government districts.
+\item
+  The city of Manchester, which is used by visitors to refer to the city centre and surrounding areas.
+This definition if formalised in the Major Towns and Cities (TCITY) dataset provided by the Office for National Statistics.
+\end{itemize}
+
+We use the TCITY definition of Manchester, which was ``developed specifically for the production and analysis of statistics'' and which is shown in Figure \ref{fig:manchestermap}.
+
+\begin{verbatim}
+##   [1] "geo_code"                "geo_name"               
+##   [3] "lad11cd"                 "lad_name"               
+##   [5] "all"                     "bicycle"                
+##   [7] "foot"                    "car_driver"             
+##   [9] "car_passenger"           "motorbike"              
+##  [11] "train_tube"              "bus"                    
+##  [13] "taxi_other"              "govtarget_slc"          
+##  [15] "govtarget_sic"           "govtarget_slw"          
+##  [17] "govtarget_siw"           "govtarget_sld"          
+##  [19] "govtarget_sid"           "govtarget_slp"          
+##  [21] "govtarget_sip"           "govtarget_slm"          
+##  [23] "govtarget_sim"           "govtarget_slpt"         
+##  [25] "govtarget_sipt"          "govnearmkt_slc"         
+##  [27] "govnearmkt_sic"          "govnearmkt_slw"         
+##  [29] "govnearmkt_siw"          "govnearmkt_sld"         
+##  [31] "govnearmkt_sid"          "govnearmkt_slp"         
+##  [33] "govnearmkt_sip"          "govnearmkt_slm"         
+##  [35] "govnearmkt_sim"          "govnearmkt_slpt"        
+##  [37] "govnearmkt_sipt"         "gendereq_slc"           
+##  [39] "gendereq_sic"            "gendereq_slw"           
+##  [41] "gendereq_siw"            "gendereq_sld"           
+##  [43] "gendereq_sid"            "gendereq_slp"           
+##  [45] "gendereq_sip"            "gendereq_slm"           
+##  [47] "gendereq_sim"            "gendereq_slpt"          
+##  [49] "gendereq_sipt"           "dutch_slc"              
+##  [51] "dutch_sic"               "dutch_slw"              
+##  [53] "dutch_siw"               "dutch_sld"              
+##  [55] "dutch_sid"               "dutch_slp"              
+##  [57] "dutch_sip"               "dutch_slm"              
+##  [59] "dutch_sim"               "dutch_slpt"             
+##  [61] "dutch_sipt"              "ebike_slc"              
+##  [63] "ebike_sic"               "ebike_slw"              
+##  [65] "ebike_siw"               "ebike_sld"              
+##  [67] "ebike_sid"               "ebike_slp"              
+##  [69] "ebike_sip"               "ebike_slm"              
+##  [71] "ebike_sim"               "ebike_slpt"             
+##  [73] "ebike_sipt"              "base_slcyclehours"      
+##  [75] "govtarget_sicyclehours"  "govnearmkt_sicyclehours"
+##  [77] "gendereq_sicyclehours"   "dutch_sicyclehours"     
+##  [79] "ebike_sicyclehours"      "base_sldeath"           
+##  [81] "base_slyll"              "base_slvalueyll"        
+##  [83] "base_slsickdays"         "base_slvaluesick"       
+##  [85] "base_slvaluecomb"        "govtarget_sideath"      
+##  [87] "govtarget_siyll"         "govtarget_sivalueyll"   
+##  [89] "govtarget_sisickdays"    "govtarget_sivaluesick"  
+##  [91] "govtarget_sivaluecomb"   "govnearmkt_sideath"     
+##  [93] "govnearmkt_siyll"        "govnearmkt_sivalueyll"  
+##  [95] "govnearmkt_sisickdays"   "govnearmkt_sivaluesick" 
+##  [97] "govnearmkt_sivaluecomb"  "gendereq_sideath"       
+##  [99] "gendereq_siyll"          "gendereq_sivalueyll"    
+## [101] "gendereq_sisickdays"     "gendereq_sivaluesick"   
+## [103] "gendereq_sivaluecomb"    "dutch_sideath"          
+## [105] "dutch_siyll"             "dutch_sivalueyll"       
+## [107] "dutch_sisickdays"        "dutch_sivaluesick"      
+## [109] "dutch_sivaluecomb"       "ebike_sideath"          
+## [111] "ebike_siyll"             "ebike_sivalueyll"       
+## [113] "ebike_sisickdays"        "ebike_sivaluesick"      
+## [115] "ebike_sivaluecomb"       "base_slcarkm"           
+## [117] "base_slco2"              "govtarget_sicarkm"      
+## [119] "govtarget_sico2"         "govnearmkt_sicarkm"     
+## [121] "govnearmkt_sico2"        "gendereq_sicarkm"       
+## [123] "gendereq_sico2"          "dutch_sicarkm"          
+## [125] "dutch_sico2"             "ebike_sicarkm"          
+## [127] "ebike_sico2"             "perc_rf_dist_u10km"     
+## [129] "avslope_perc_u10km"      "geometry"               
+## [131] "pcycle"
+\end{verbatim}
+
+\begin{verbatim}
+## Warning: st_centroid assumes attributes are constant over geometries
+\end{verbatim}
+
+\includegraphics{README_files/figure-latex/unnamed-chunk-4-1.pdf}
 

### 2023-09-22 latexdiff.R
flowchart

@@ -1,12 +1,27 @@
 # Aim: generate diff for peer review
 
+# ROUND 1: Robin
 
+# download.file(
+#   "https://github.com/Hussein-Mahfouz/cycle-networks/raw/ce4827fc708d510bead10d4998ce7ba875286d0d/README.Rmd",
+#   "paper-original-submission.Rmd"
+#   )
+# file.edit("paper-original-submission.Rmd")
+# rmarkdown::render("paper-original-submission.Rmd")
+# system("latexdiff paper-original-submission.tex README.tex > tracked-changes.tex")
+# tinytex::pdflatex("tracked-changes.tex")
+# browseURL("tracked-changes.pdf")
+
+# ROUND 2: Hussein
+
+# download paper version submitted after round 1 of reviews -> save it as "paper_round_1"
 download.file(
-  "https://github.com/Hussein-Mahfouz/cycle-networks/raw/ce4827fc708d510bead10d4998ce7ba875286d0d/README.Rmd",
-  "paper-original-submission.Rmd"
-  )
-file.edit("paper-original-submission.Rmd")
-rmarkdown::render("paper-original-submission.Rmd")
-system("latexdiff paper-original-submission.tex README.tex > tracked-changes.tex")
-tinytex::pdflatex("tracked-changes.tex")
-browseURL("tracked-changes.pdf")
+  "https://github.com/Hussein-Mahfouz/cycle-networks/blob/c3b74611e92b4092af487c1bc179187c9e1608bb/README.Rmd",
+  "paper_round_1.Rmd"
+)
+file.edit("paper_round_1.Rmd")
+rmarkdown::render("paper_round_1.Rmd")
+# compare with current version and create tracked changes
+system("latexdiff paper_round_1.tex README.tex > tracked-changes_2.tex")
+tinytex::pdflatex("tracked-changes_2.tex")
+browseURL("tracked-changes_2.pdf")

### 2023-09-23 README.Rmd
edits for submission 3

--- 2023-09-22 README.Rmd
@@
 output:
-   bookdown::github_document2:
+   # bookdown::github_document2:
 #    bookdown::pdf_document2:
- # bookdown::word_document2:
+  bookdown::word_document2:
       highlight: null   # to remove code highlighting in r chunks
@@
 In this paper we propose an approach for planning connected cycling networks that can be generalised to any city with appropriate input data (as shown in the Appendix).
-We used the city of Manchester as a case study due to its ambitions to become the first UK region with a carbon neutral transport network, supported by ambitious plans for a 1800 mile walking and cycling network (the 'Bee Network') that will be delivered over 10 years [@manchestercyclingbee2020].
+We used the city of Manchester as a case study.
+The city is situated in the region of Greater Manchester, which has ambitions to become the first UK region with a carbon neutral transport network.
+These ambitions are supported by plans for a 1800 mile walking and cycling network (as part of the 'Bee Network') that will be delivered over 10 years [@manchestercyclingbee2020].
 While schemes have already been put in place for around a quarter of this network, the majority is yet to be built.
@@
 
-Manchester is a city in the North West of England.
+Manchester is a city in the North West of England, and is the most populous of the 10 metropolitan boroughs that make up Greater Manchester (\@ref(fig:manchestermap)).
 "Manchester" can be used to refer to multiple things:
@@
 -   The city of Manchester, which is used by visitors to refer to the city centre and surrounding areas.
-This definition if formalised in the Major Towns and Cities (TCITY) dataset provided by the Office for National Statistics.
+This definition is formalised in the Major Towns and Cities (TCITY) dataset provided by the Office for National Statistics [@ons2015cities].
 
@@
 Manchester is a cosmopolitan city with a large university population.
-As shown in Figure \@ref(fig:manchestermap) (right), the high density of the central area allows high levels of cycling, although the surrounding areas have low levels of cycling.
-Manchester and its surrounding have many deprived areas.
+As shown in Figure \@ref(fig:manchestermap) (right), the high density of the central area allows relatively high levels of cycling compared to the rest of the city.
 
-```{r manchestermap, echo=FALSE, fig.cap="Maps of Greater Manchester (top) and the case study area (City of Manchester)"}
+```{r manchestermap, echo=FALSE, fig.cap="Maps of Greater Manchester (left) and the case study area - City of Manchester (right)", fig.show='hold', fig.align="center"}
 greater_manchester = pct::get_pct_zones(region = "greater-manchester", geography = "msoa")
@@
 tmap_arrange(m1, m2, nrow = 1)
+```
+
+In the context of other cities defined in the TCITY dataset, Manchester is a large city, with a mid-level of cycling (Figure \@ref(fig:manchestercyclinglevels)).
+If the ambitions of the Bee Network are realized in the coming years, this mode share could go up significantly.
+
+```{r manchestercyclinglevels, echo=FALSE, fig.cap="Cycling levels in cities across the country", out.width="85%", fig.show='hold', fig.align="center"}
+
+knitr::include_graphics("./data/uk_cities_mode_share_2021.png")
 ```

### 2023-09-23 paper_drafts/paper_round_1.tex
edits for submission 3

--- 2023-05-01 README.tex
@@
 
-The 2015 Paris agreement (UN, 2015) acknowledged that fundamental changes to societies and economies are necessary to mitigate climate change.
-Like other sectors, transport is under substantial pressure to decarbonise, resulting in a number of technical innovations including electric vehicles.
-But new vehicle technologies can only go so far and fail to tackle parallel problems such as congestion, road traffic casualties and physical inactivity (Brand et al., 2020).
-
-In this context, interest and investment in active modes has grown substantially, especially with the growing realisation that the benefits extend beyond congestion and the environment, as active travel can also help alleviate what is referred to as the pandemic of global inactivity; physical inactivity is on the rise and has become the 4th highest cause of death globally (Kohl 3rd et al., 2012).
-Various studies have documented the association between active transport and lower risk of disease, including cancer and cardiovascular disease (Celis-Morales et al., 2017; Jarrett et al., 2012; Patterson et al., 2020).
-In the wake of the Covid-19 pandemic, and the resulting reduced capacity of public transport, the UK government has pledged to invest billions of pounds to improve walking and cycling infrastructure across the country.
-While this unprecedented sum is an opportunity to reshape cities in a way that improves the well-being of citizens, it does come with a warning:
+The 2015 Paris agreement (UN, 2015) acknowledged that
+fundamental changes to societies and economies are necessary to mitigate
+climate change. Like other sectors, transport is under substantial
+pressure to decarbonise, resulting in a number of technical innovations
+including electric vehicles. But new vehicle technologies can only go so
+far and do not tackle parallel problems such as congestion, road traffic
+casualties and physical inactivity (Brand et al., 2020).
+
+In this context, interest and investment in active modes are growing.
+The benefits extend beyond congestion and the environment, as active
+travel can also help alleviate what is referred to as the pandemic of
+global inactivity; physical inactivity is on the rise and has become the
+4th highest cause of death globally (Kohl 3rd et al., 2012). Various
+studies have documented the association between active transport and
+lower risk of disease, including cancer and cardiovascular disease
+(Celis-Morales et al., 2017; Jarrett et al., 2012; Patterson et al., 2020). In the wake of the Covid-19 pandemic, and
+the resulting reduced capacity of public transport, the UK government
+has pledged to invest billions of pounds to improve walking and cycling
+infrastructure across the country. While this unprecedented sum is an
+opportunity to reshape cities in a way that improves the well-being of
+citizens, it does come with a warning:
 
 \begin{quote}
-``Inadequate cycling infrastructure discourages cycling and wastes public money.
-Much cycling infrastructure in this country is inadequate.
-It reflects a belief, conscious or otherwise, that hardly anyone cycles, that cycling is unimportant and that cycles must take no meaningful space from more important road users, such as motor vehicles and pedestrians'' (DfT, 2020a).
+``Inadequate cycling infrastructure discourages cycling and wastes
+public money. Much cycling infrastructure in this country is
+inadequate. It reflects a belief, conscious or otherwise, that hardly
+anyone cycles, that cycling is unimportant and that cycles must take
+no meaningful space from more important road users, such as motor
+vehicles and pedestrians'' (DfT, 2020a).
 \end{quote}
 
-Funding alone is therefore no guarantee of a change in commuting across the country; it must be used to design adequate cycling infrastructure that is based on motivators and deterrents to cycling.
-In this paper we propose an approach for planning connected cycling networks that can be generalised to any city with appropriate input data (as shown in the Appendix).
-We used the city of Manchester as a case study due to its ambitions to become the first UK region with a carbon neutral transport network, supported by ambitious plans for a 1800 mile walking and cycling network (the `Bee Network') that will be delivered over 10 years (Greater Manchester, 2020).
-While schemes have already been put in place for around a quarter of this network, the majority is yet to be built.
+The funding on its own is therefore no guarantee of a change in
+commuting across the country; it must be used to design adequate cycling
+infrastructure that is based on motivators and deterrents to cycling. In this paper we
+propose an approach for planning a connected cycling network that is meant to
+be applicable to any city (as shown in the Appendix), and we use the city of Manchester as a case study. Greater
+Manchester aims to become the first region to reach a carbon neutral transport network;
+its local authorities have developed plans for a 1800 mile walking and cycling network
+(Bee Network) that will be delivered over 10 years (Greater Manchester, 2020). While schemes
+have already been put in place for around a quarter of this network, the majority is yet to be built.
 
@@
 
-Much research has explored the question of what gets people cycling.
-Segregated cycling infrastructure --- road space that is allocated to cyclists only with physical separation to protect cyclists --- is associated with cycling uptake (Aldred et al., 2019; Goodman et al., 2014; Marqués et al., 2015).
-Separation from motorized vehicles has found to be key (Winters et al., 2011).
-Revealed preference of cyclists shows that they are willing to deviate from the most efficient routes in order to commute on safer roads (Crane et al., 2017).
-However, such deviations are only considered if they do not considerably increase route circuitry; behavior studies have found that the probability of choosing a route decreases in proportion to its length relative to the shortest route (Broach et al., 2011; Winters et al., 2010).
-Another defining feature for cycling infrastructure is how well connected it is.
-Cyclists prefer cohesive infrastructure, particularly when cycling on arterial roads with high levels of motorized traffic (Stinson and Bhat, 2003).
-The lack of well-connected cycling infrastructure is one of the main obstacles to increasing cycling uptake (Caulfield et al., 2012).
-While direct and cohesive cycling networks have been shown to positively impact cycling rates, density\footnote{making an area's bicycle network denser means adding more cycling routes in the area and thereby giving cyclists more route options.} of the cycling network is also vital (Schoner and Levinson, 2014).
+Much research has been done to understand what gets people to cycle.
+Segregated cycling infrastructure\footnote{\emph{Segregated cycling infrastructure}
+  refers to road space that is allocated to cyclists only, with physical
+  separation to protect cyclists from other modes of transport.} has been
+shown to increase cycling uptake (Aldred et al., 2019; Goodman et al., 2014; Marqués et al., 2015), with the separation from motorized
+vehicles being key (Winters et al., 2011). Revealed preference of
+cyclists shows that they are willing to deviate from the most efficient
+routes in order to commute on safer roads (Crane et al., 2017).
+However, such deviations are only considered if they do not considerably
+increase route circuitry; behaviour studies have found that the
+probability of choosing a route decreases in proportion to its length
+relative to the shortest route (Broach et al., 2011; Winters et al., 2010).
+Another defining feature for cycling infrastructure is how well
+connected it is. Cyclists prefer cohesive infrastructure, particularly
+when cycling on arterial roads with high levels of motorized traffic
+(Stinson and Bhat, 2003). The lack of well-connected cycling
+infrastructure is one of the main obstacles to increasing cycling uptake
+(Caulfield et al., 2012). While direct and cohesive cycling networks
+have been shown to positively impact cycling rates, density\footnote{making an
+  area's bicycle network denser means adding more cycling routes in the
+  area and thereby giving cyclists more route options.} of the cycling
+network is also vital (Schoner and Levinson, 2014).
 
@@
 
-The studies outlined above lay out the fundamentals for designing cycling networks that generate significant cycling uptake, but they do not propose network-level interventions.
-In this section we outline methods used in past studies, namely optimization and network analysis methods, such as connected components and community detection, and examine how they are leveraged to suggest cycling network designs.
-We compare the effectiveness of these network-level studies in incorporating the fundamentals outlined above.
-Our proposed approach is inspired by these methods, but it attempts to add to them by ensuring that all of the outlined fundamentals are accounted for.
-It also goes further by attempting to factor in ethical considerations related to the distribution of investment.
-
-\emph{Optimization} methods have been used to propose improvements to cycling networks.
-Mesbah et al. (2012) propose a bi-level formulation to optimize allocation of cycling lanes to the network without exceeding a set budget.
-They account for the effect of cycling lanes on car traffic, and attempt to maximize utilization of said lanes with minimal impact on car travel times.
+The studies outlined above lay out the fundamentals for designing
+cycling networks that generate significant cycling uptake, but they do
+not propose network-level interventions. In this section we outline
+methods used in past studies, namely optimization and network
+analysis methods, such as connected components and community
+detection, and examine how they are leveraged to suggest cycling network
+designs. We compare the effectiveness of these network-level studies in
+incorporating the fundamentals outlined above.
+Our proposed approach is inspired by these methods, but it attempts to add to them
+by ensuring that all of the outlined fundamentals are accounted for. It
+also goes further by attempting to factor in ethical considerations
+related to the distribution of investment.
+
+\emph{Optimization} methods have been used to propose improvements to
+cycling networks. Mesbah et al. (2012) propose a bi-level formulation to
+optimize allocation of cycling lanes to the network without exceeding a
+set budget. They account for the effect of cycling lanes on car traffic,
+and attempt to maximize utilization of said lanes with minimal impact on
+car travel times.
 Safeguarding against increased car traffic may be counter-productive if the goal is to create a mode-shift, as research has shown that reducing road space for cars leads to less cars on the road, a phenomenon referred to as ``traffic evaporation'' (Nello-Deakin, 2020).
-
-Mauttone et al. (2017) developed an optimization framework aiming at minimizing the total user cost of cycling on the network without considering car usage.
-The aggregate flow\footnote{\emph{flow} is used throughout this research to refer to the cycling demand when it is routed onto the road network.
-  The flow on any road segment is the cumulative demand on it, resulting from cyclists commuting between various Origin-Destination (OD) pairs.} on road segments (links) is obtained by using shortest paths to route existing cycling demand onto the road network.
-The solution is a proposed set of links where cycling infrastructure should be added in order to minimize the overall travel cost of cyclists across the network.
-The cost of traversing a link is given as a function of its length and whether or not it has cycling infrastructure.
-The problem has also been solved by attempting to find the minimum cost of improving roadway links to meet a desired level of service (LOS) (Duthie and Unnikrishnan, 2014).
-In this formulation, all origin-destination (OD) pairs need to be connected by roads that meet the desired LOS.
-A limitation of these approaches is that they do not explicitly solve for continuity.
-Continuity is addressed using either a constraint specifying that each link with a bike lane should be connected to at least one destination (Mesbah et al., 2012), a constraint on maximum deviation from shortest paths (Duthie and Unnikrishnan, 2014), or a discontinuity penalty to prioritize connected road segments (Mauttone et al., 2017).
-In this paper, continuity is analysed by looking at the connectivity of the network through the graph-theoretic concept of \emph{connected components}.
-
-Orozco et al. (2020) study the existing cycling network in terms of its disconnected components.
+On the other hand, Mauttone et al. (2017) developed an optimization framework aiming at minimizing the total
+user cost of cycling on the network without considering car usage. The aggregate flow\footnote{\emph{flow} is
+  used throughout this research to refer to the cycling demand when it is
+  routed onto the road network. The flow on any road segment is the
+  cumulative demand on it, resulting from cyclists commuting between
+  various Origin-Destination (OD) pairs.} on road segments (links) is obtained by using
+shortest paths to route existing cycling demand onto the road network. The solution is a proposed set of links where cycling infrastructure
+should be added in order to minimize the overall travel cost of cyclists
+across the network. The cost of traversing a link is given as a function
+of its length and whether or not it has cycling infrastructure. The
+problem has also been solved by attempting to find the minimum cost of
+improving roadway links to meet a desired level of service (LOS)
+(Duthie and Unnikrishnan, 2014). In this formulation, all origin-destination
+(OD) pairs need to be connected by roads that meet the desired LOS. A
+limitation of these approaches is that they do not explicitly solve for
+continuity. Continuity is addressed using either a constraint specifying
+that each link with a bike lane should be connected to at least one
+destination (Mesbah et al., 2012), a constraint on maximum deviation
+from shortest paths (Duthie and Unnikrishnan, 2014), or a discontinuity
+penalty to prioritize connected road segments (Mauttone et al., 2017).
+
+In this paper, continuity is analysed by looking at the connectivity of
+the network through the graph-theoretic concept of \emph{connected
+components}. Orozco et al. (2020) study the existing cycling network in terms of its disconnected
+components.
 They propose two different algorithms to connect disconnected components; one connects the two largest components, and the other connects the largest component to its nearest neighbor.
 By measuring the growth of the largest connected component as a function of the kilometers of network added, they determine that these approaches are more effective at improving cycling network connectivity than random allocation of cycling infrastructure.
-The concept of connected components is also at the core of the methodology proposed by Olmos et al. (2020).
-After routing the cycling demand onto the network links, they use percolation theory to filter out the links based on the aggregate flow passing through them, varying the flow threshold for filtering to identify the minimum flow at which the whole city is connected by a single component.
-While these approaches deal with continuity better, they look at the network as a whole when attempting to improve it, and in doing so fail to account for equitable distribution of infrastructure.
-
-Other approaches have focused on characteristics of route segments as the basis of analysis to help prioritize investment in cycle infrastructure where it is most needed.
-Grisé and El-Geneidy (2018) developed a multi-criteria approach, combined with estimated levels of usage, to highlight routes most in need of investment in Quebec City.
-Wysling and Purves (2022) present a method that used open data from Paris to generate cycling suitability ratings, based on `bicycle suitability multipliers' associated with different types of road across the city, to support more effective interventions.
-
-With the exception of Grisé and El-Geneidy (2018), the aforementioned studies do not explore the equity impacts of different network prioritization approaches.
-The following section highlights the importance of considering (often unstated) ethical assumptions underlying network optimization and prioritization approaches of the type outlined above.
+The concept of connected components is
+also at the core of the methodology proposed by Olmos et al. (2020). After
+routing the cycling demand onto the network links, they use percolation
+theory to filter out the links based on the aggregate flow passing through them, varying the flow threshold for
+filtering to identify the minimum flow at which the whole city is
+connected by a giant component. While these approaches deal with
+continuity better, they look at the network as a whole when attempting
+to improve it, and in doing so fail to account for equitable
+distribution of infrastructure.
 
@@
 
-All of these network-level methodologies are underpinned by ethical principles, even though these principles are not explicitly acknowledged by the authors.
-This is important since different ethical principles constitute different problem formulations and targets.
-Broadly speaking, transport appraisal can be based on either utilitarian or egalitarian principles.
-The former seeks to maximize the overall benefit, while the latter is concerned with a fair distribution of benefits (Jafino et al., 2020).
+All of these network-level methodologies are underpinned by ethical
+principles, even though these principles are not explicitly acknowledged
+by the authors. This is important since different ethical principles
+constitute different problem formulations and targets. Broadly speaking,
+transport appraisal can be based on either utilitarian or egalitarian
+principles. The former seeks to maximize the overall benefit, while the
+latter is concerned with a fair distribution of benefits
+(Jafino et al., 2020).
 The utilitarian approach, historically popular in transport planning, has been criticised for focusing on the bigger picture and failing to account for the distribution of investments on the different communities of the study area (Lucas et al., 2016; Nahmias-Biran et al., 2017).
-Pereira et al. (2017) emphasize the need for a more egalitarian approach to transport planning.
-They highlight accessibility as a cornerstone of distributive justice, and contend that policies should aim to distribute investments in a way that minimizes spatial variations in accessibility.
-
-\hypertarget{method-and-application}{%
-\section{Method and Application}\label{method-and-application}}
-
-\hypertarget{method-overview}{%
-\subsection{Method overview}\label{method-overview}}
-
-This research attempts to propose an egalitarian framework for cycling network design.
-This is done by identifying the different sub-networks that exist within the larger network, and ensuring that each gets a fair share of investment.
-Trip patterns in a city are not uniformly distributed geographically, and \emph{community finding} methods have been used to partition study areas into localized areas that experience a disproportionate number of trips within them.
-Akbarzadeh et al. (2018) use a modularity maximization approach (Blondel et al., 2008) on taxi trip data to identify seven different communities in the city of Isfahan, Iran.
-An optimization problem is then formulated to connect nodes within each community with cycling infrastructure, with the emphasis being on connectivity within the communities, not between them.
-Bao et al. (2017) adopt a similar methodology, first identifying communities and then using a greedy network expansion algorithm to simultaneously add links to each community.
-The link with the highest benefit-cost ratio in each community is selected, and the network is grown by adding neighboring links to the solution until a budget limit is met.
-The benefit is the flow on the link, and each link is assigned a cost based on current road conditions.
-
-Our work builds on these community finding approaches by proposing a similar greedy network expansion algorithm for cycle network expansion \emph{within communities}.
-We incorporate community finding methods for study area partitioning with weighted routing to avoid links that are stressful to cycle on.
-In doing so, we propose an approach that ensures a better distribution investments across the study area while also accounting for motivators and deterrents to cycling.
-We propose three sub-methods that address some of the limitations of previous studies.
-These limitations include (a) an inherent bias when basing network design solely on existing cycling demand, (b) the proposal of routes that may not correspond to studies on cyclist preference and government policies, and (c) an insufficient consideration of the ethical principles underlying the analysis.
-The current work addresses these limitations through the following organisation.
-Section \ref{calculating-potential-cycling-demand} focuses on calculating potential cycling demand.
-Section \ref{routing} focuses on routing the demand onto the road network while accounting for cyclist preferences and government priorities.
-Section \ref{road-segment-prioritization} outlines a method for partitioning the study area based on a community finding algorithm and routed cycling demand.
-It then introduces the network expansion algorithms, and compares an approach grounded in `egalitarianism' to one grounded in `utilitarianism'.
-As mentioned above, Manchester is used as a case study.
-An overview of the steps involved is shown in Figure \ref{fig:methodflowchart}.
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.8\linewidth]{extras/methodology_flowchart}
-
-}
-
-\caption{Flowchart of proposed methodology}\label{fig:methodflowchart}
-\end{figure}
+Pereira et al. (2017) emphasize the need for a more
+egalitarian approach to transport planning. They highlight accessibility
+as a cornerstone of distributive justice, and contend that policies
+should aim to distribute investments in a way that minimizes spatial
+variations in accessibility.
+
+This research attempts to propose an egalitarian framework for cycling
+network design. This is done by identifying the different sub-networks
+that exist within the larger network, and ensuring that each gets a fair
+share of investment. Trip patterns in a city are not uniformly
+distributed geographically, and \emph{community finding} methods have been
+used to partition study areas into localized areas that experience a
+disproportionate number of trips within them. Akbarzadeh et al. (2018)
+use a modularity maximization approach (Blondel et al., 2008) on taxi trip
+data to identify seven different communities in the city of Isfahan, Iran.
+An optimization problem is then formulated to connect nodes within each
+community with cycling infrastructure, with the emphasis being on
+connectivity within the communities, not between them. Bao et al. (2017)
+adopt a similar methodology, first identifying communities and then
+using a greedy network expansion algorithm to simultaneously add links
+to each community. The link with the highest benefit-cost ratio in each
+community is selected, and the network is grown by adding neighboring
+links to the solution until a budget limit is met. The benefit is the
+flow on the link, and each link is assigned a cost based on current road
+conditions.
+
+Our work builds on these community finding approaches by proposing a
+similar greedy network expansion algorithm for cycle network expansion
+\emph{within communities}. We incorporate community finding methods for study
+area partitioning with weighted routing to avoid links that are
+stressful to cycle on. In doing so, we propose an approach that ensures a better distribution
+investments across the study area while also accounting for motivators and deterrents to cycling.
+We propose three sub-methods that address some of the limitations of previous
+studies. These limitations include (a) an inherent bias when basing network
+design solely on existing cycling demand, (b) the proposal of routes that may
+not correspond to studies on cyclist preference and government policies,
+and (c) an insufficient consideration of the ethical principles underlying
+the analysis. The current work addresses these limitations through the following organisation. Section \ref{calculating-potential-cycling-demand}
+focuses on calculating potential cycling demand. Section \ref{routing}
+focuses on routing the demand onto the road network while accounting for
+cyclist preferences and government priorities. Section
+\ref{road-segment-prioritization} outlines a method for partitioning the
+study area based on a community finding algorithm and routed cycling
+demand. It then introduces the network expansion algorithms, and compares an approach grounded in
+`egalitarianism' to one grounded in `utilitarianism'.
 
 \hypertarget{calculating-potential-cycling-demand}{%
-\subsection{Calculating Potential Cycling Demand}\label{calculating-potential-cycling-demand}}
+\section{Calculating Potential Cycling Demand}\label{calculating-potential-cycling-demand}}
 
@@
 Some use existing cycling demand, some calculate potential cycling demand, and others ignore demand completely.
-Duthie and Unnikrishnan (2014) note that relying only on existing cycling activity to prioritize cycling infrastructure can reinforce existing cycling patterns and ignore potential cycling demand that could be satisfied by a connected network.
-To avoid this issue they choose to ignore existing demand completely, and focus on creating a network that connects the entire study area.
+Duthie and Unnikrishnan (2014) note that relying only on existing cycling activity to prioritize cycling
+infrastructure can reinforce existing cycling patterns and ignore potential cycling demand that could be satisfied by a connected network.
+To avoid this issue they choose to ignore existing demand completely, and focus on creating a network that connects the
+entire study area.
 Olmos et al.
-(2020) opt to calculate potential demand instead; they obtain the distance distribution of cyclists using a smartphone-based bicycle GPS data, and then use a rejection-sampling algorithm on the OD data of the study area to match the potential demand distribution to the distribution obtained from GPS data.
-
-\emph{OD} data can be obtained from a range of sources, including GPS data, household travels surveys and census data on work locations.
-In areas where observed OD data is unavailable, modelling techniques such as spatial interaction models can be used to estimate travel volumes between zones (Black, 1995; Martínez and Viegas, 2013; Wilson, 1971).
-In this paper we use open access data from the UK census (ONS, 2011), which contains aggregate statistics on number of commuters between administrative zones --- Middle layer Super Output Areas (MSOA) --- by mode of travel (this is similar to the approach used by Grisé and El-Geneidy (2018)).
-MSOAs have an average population of just over 8000 (ONS, 2018).\footnote{See \url{https://wicid.ukdataservice.ac.uk/} for open access to the OD data.}
-Figures \ref{fig:potdemhistograms} and \ref{fig:desirefacetcycling} illustrate the proportion of trips cycled by distance and the geographic extent of the input OD dataset used in this paper.
-
-For our purposes, we use a logistic regression model to calculate potential cycling demand.
-The model was adapted from the Propensity to Cycle Tool (PCT), which estimates the proportion of trips (\(\boldsymbol{C_{p}}\)) for each OD pair that would cycle under different scenarios of change as a function of distance and hilliness (Lovelace et al., 2017).
-We used the Government Target scenario, indicating a nationwide target of doubling cycling by 2025.
-The logistic regression model used to calculate \(\boldsymbol{C_{p}}\) has the following parameters:
+(2020) opt to calculate potential demand instead; they obtain the distance distribution of cyclists using a smartphone-based bicycle GPS data, and then use a rejection-sampling algorithm on the OD data of the study area to match
+the potential demand distribution to the distribution obtained from GPS data.
+
+\emph{OD} data can be obtained from a range of sources,
+including GPS data, household travels surveys and census data on work
+locations. In areas where observed OD data is unavailable, modelling
+techniques such as spatial interaction models can be used to estimate
+travel volumes between zones (Black, 1995; Martínez and Viegas, 2013; Wilson, 1971). In this paper we use open access data from the UK
+census (ONS, 2011), which contains aggregate statistics on
+number of commuters between administrative zones --- Middle layer Super
+Output Areas (MSOA) --- by mode of travel. MSOAs have an average
+population of just over 8000 (ONS, 2018).\footnote{ See
+  \url{https://wicid.ukdataservice.ac.uk/} for open access to the OD data.}
+Figures \ref{fig:potdemhistograms} and \ref{fig:desirefacetcycling}
+illustrate the proportion of trips cycled by distance and the
+geographic extent of the input OD dataset used in this paper.
+
+For our purposes, we use a logistic regression model to calculate
+potential cycling demand. The model was adapted from the Propensity to
+Cycle Tool (PCT), which estimates the proportion of trips
+(\(\boldsymbol{C_{p}}\)) for each OD pair that would cycle under
+different scenarios of change as a
+function of distance and hilliness (Lovelace et al., 2017). We used the Government Target scenario,
+indicating a nationwide target of doubling cycling by 2025. The
+logistic regression model used to calculate \(\boldsymbol{C_{p}}\) has
+the following parameters:
 
@@
 
-\noindent where \(d\) and \(s\) are the distance and slope associated with each OD pair, and \(\alpha\) and all \(\beta\)s are parameters calculated by a regression model on the training data.
-The square and square-root distance terms ``capture the non-linear impact of distance on the likelihood of cycling'', and interaction terms capture the combined effect of slope and distance (Lovelace et al., 2017).
-Alternative cycling uptake models could be `plugged in' to our approach for different contexts or scenarios of change.
-
-The potential demand calculations show that the current and potential number of cyclists both follow a bell-shaped distribution, with the number of trips peaking around the 3-4km commuting distance and then going back down for longer distances (see Figures \ref{fig:potdemhistograms} and \ref{fig:desirefacetcycling}).
+\noindent where \(d\) and \(s\) are the distance and slope associated with each
+OD pair, and \(\alpha\) and all \(\beta\)s are parameters calculated by a regression
+model on the training data. The square and square-root distance terms
+``capture the non-linear impact of distance on the likelihood of
+cycling'', and interaction terms capture the combined effect of slope
+and distance (Lovelace et al., 2017). Alternative cycling uptake
+models could be `plugged in' to our approach for different contexts or scenarios of change.
+
+The potential demand calculations show that the current and potential
+number of cyclists both follow a bell-shaped distribution, with the
+number of trips peaking around the 3-4km commuting distance and then
+going back down for longer distances (see Figures
+\ref{fig:potdemhistograms} and \ref{fig:desirefacetcycling}).
 
@@
 
-The uptake model used in this paper assumes uptake of traditional bicycles, which are affected by topography and distance due to physical effort.
-As discussed in Section \ref{discussion-and-conclusions}, using adapted uptake models could enable the approach to help plan for solutions such as e-bikes that enable trips covering longer distances and hillier roads.
+The uptake model used in this paper assumes uptake of traditional bicycles,
+which are affected by topography and distance due to physical effort.
+As discussed in Section \ref{discussion-and-conclusions}, using adapted uptake
+models could enable the approach to help plan for solutions such as e-bikes
+that enable trips covering longer distances and hillier roads.
 
 \hypertarget{routing}{%
-\subsection{Routing}\label{routing}}
-
-The next step in our approach is to route the potential cycling demand (\(\boldsymbol{C_{p}}\)) between all OD pairs onto the road network.
-We choose not to use the PCT approach for routing as it is reliant on external and therefore inflexible routing services.
+\section{Routing}\label{routing}}
+
+The next step in our approach is to route the potential cycling demand
+(\(\boldsymbol{C_{p}}\)) between all OD pairs onto the road network. We
+choose not to use the PCT approach for routing as it is reliant on
+external and therefore inflexible routing services.
 
@@
 \item
-  \textbf{Cyclist Preference}: Work done by Dill and McNeil (2013) on examining cyclist typologies determined that around 60\% of Portland residents fit under the \emph{interested but concerned} category.
-These were people that enjoyed cycling but avoided it due safety concerns.
-The key to encouraging this group was to create a low-stress cycling network, not only though segregated infrastructure but also by planning routes that passed through residential streets.
+  \textbf{Cyclist Preference}: Work done by Dill and McNeil (2013) on examining
+  cyclist typologies determined that around 60\% of Portland residents
+  fit under the \emph{interested but concerned} category. These were people
+  that enjoyed cycling but avoided it due safety concerns. The key to
+  encouraging this group was to create a low-stress cycling network,
+  not only though segregated infrastructure but also by planning
+  routes that passed through residential streets.
 \item
-  \textbf{Low-Traffic Neighbourhoods (LTNs)}: The UK Department for Transport is allocating funding to local authorities to invest in Active Transport, partially through the creation of LTNs (DfT, 2020a).
-This includes closing off residential streets to motorized traffic.
+  \textbf{Low-Traffic Neighbourhoods (LTNs)}: The UK Department for Transport is
+  allocating funding to local authorities to invest in Active
+  Transport, partially through the creation of LTNs
+  (DfT, 2020a). This includes closing off residential
+  streets to motorized traffic.
 \item
-  \textbf{Existing Cycling Infrastructure}: Utilizing existing cycling infrastructure makes economic sense, as small investments may lead to large connectivity gains as the disconnected cycling infrastructure gets joined together.
+  \textbf{Existing Cycling Infrastructure}: Utilizing existing cycling
+  infrastructure makes economic sense, as small investments may lead
+  to large connectivity gains as the disconnected cycling
+  infrastructure gets joined together.
 \end{enumerate}
 
-The above points are accounted for by using a weighted road network for routing.
-This has been previously done by multiplying all road segments without cycling infrastructure by a fixed impedance factor (Mauttone et al., 2017), or by assigning different weights to the road segments that are (a) proportional to the investment cost of bringing them to an acceptable level of stress for cycling (Duthie and Unnikrishnan, 2014), or (b) inversely proportional to perceived level of stress experienced by cyclists (Wysling and Purves, 2022).
-Perceived stress for cyclists has been found to be a function of vehicular traffic volume and speed (Sorton and Walsh, 1994), both of which vary predictably with road type.
-Gehrke et al. (2020) build on this information to use routing impedance factors for road segments that are based on road type and the existence of cycling infrastructure.
-
-For the purposes of this paper we created a similar weighting profile that is adjusted to favor less stressful roads (based on the definition by Sorton and Walsh (1994) and information from Table \ref{table:osmroadtypes}) and roads with existing cycling infrastructure.
-We believe this to be more appropriate than the simplistic approach adopted by Mauttone et al.
-(2017) as it makes use of a hierarchy of road preference based on perceived stress levels, going beyond simply favouring roads with existing cycling infrastructure.
-It is also different to the weighting profile used by Wysling and Purves (2022) as it adds a larger weight on trunk and primary roads then other roads without cycling infrastructure.
+The above points are accounted for by using a weighted road network for
+routing. This has been previously done by multiplying all road segments
+without cycling infrastructure by a fixed impedance factor
+(Mauttone et al., 2017), or by assigning different weights to the road
+segments proportional to the investment cost of bringing them to an
+acceptable level of stress for cycling (Duthie and Unnikrishnan, 2014).
+Perceived stress for cyclists has been found to be a function of vehicular traffic
+volume and speed (Sorton and Walsh, 1994), both of which vary predictably
+with road type. Gehrke et al. (2020) build on this information to use
+routing impedance factors for road segments that are based on road type
+and the existence of cycling infrastructure.
+
+For the purposes of this paper we created a similar weighting profile
+that is adjusted to favor less stressful roads (based on the definition
+by Sorton and Walsh (1994) and information from Table
+\ref{table:osmroadtypes}) and roads with existing cycling
+infrastructure. We believe this to be more appropriate than the simplistic
+approach adopted by Mauttone et al. (2017) as it makes use of a hierarchy
+of road preference based on perceived stress levels, going beyond simply
+favouring roads with existing cycling infrastructure.
 
@@
 
-\noindent A weighted distance \(d_{w}\) for each road segment is calculated as follows \footnote{The \textbf{dodgr} r package (Padgham, 2019) is used to route cycling demand onto the road network.
-  The package uses the OpenStreetMaps (OSM) road network and allows the user to assign weights to roads based on their type.
-  The routing is done based on weighted shortest paths, with the distance along each road segment being divided by a factor to obtain the weighted distance for routing.
-  It is more intuitive to multiply when weighting a network, but the dodgr package divides by numbers between 0 and 1, which achieves the same result (road types with a weight of 0 are not routed on).
-  For the sake of reproducibility, we stick to the convention used in the package.}:
+\noindent A weighted distance \(d_{w}\) for each road
+segment is calculated as follows \footnote{The \textbf{dodgr} r package
+  (Padgham, 2019) is used to route cycling demand onto the road
+  network. The package uses the OpenStreetMaps (OSM) road network and
+  allows the user to assign weights to roads based on their type. The
+  routing is done based on weighted shortest paths, with the distance
+  along each road segment being divided by a factor to obtain the weighted
+  distance for routing. It is more intuitive to multiply when weighting a
+  network, but the dodgr package divides by numbers between 0 and 1, which
+  achieves the same result (road types with a weight of 0 are not routed on).
+  For the sake of reproducibility, we stick to
+  the convention used in the package.}:
 
@@
 
-\noindent where \(d_{unw}\) is the unweighted distance and \(W\) is the weight from Table \ref{table:weightprofiles}.
-
-All weights are between 0 and 1, and the values in the \textit{Weighted} profile are chosen to be inversely proportional to the stress level experienced by cyclists on them.
-The \textit{Unweighted} weighting profile is used to compare increases in route length resulting from two different approaches:
+\noindent where \(d_{unw}\) is the unweighted distance and
+\(W\) is the weight from Table \ref{table:weightprofiles}.
+
+All weights are between 0 and 1, and the values in the
+\textit{Weighted} profile are chosen to be inversely proportional to
+the stress level experienced by cyclists on them. The
+\textit{Unweighted} weighting profile is used to compare increases in
+route length resulting from two different approaches:
 
@@
 \item
-  \textbf{Weighted}: Relatively high impedance on Primary and Trunk roads (to minimize cycling on them).
+  \textbf{Weighted}: Relatively high impedance on Primary and Trunk roads
+  (to minimize cycling on them).
 \item
@@
 
-Comparing the cycling demand routed on the weighted and unweighted road network allows us to get a better understanding of the importance of different road types.
-In the case of Manchester, trunk roads bisect the city and are a major part of unweighted shortest paths (Figure \ref{fig:flowsfacetunweighted}).
-On the other hand, cycleways are not part of unweighted shortest paths, and so very little of the cycling demand is routed through them.
-In the weighted network, cycleways are much better utilized, and the majority of the cycling demand passes through tertiary roads, as expected.
+Comparing the cycling demand routed on the weighted and unweighted road
+network allows us to get a better understanding of the importance of
+different road types. In the case of Manchester, trunk roads bisect the
+city and are a major part of unweighted shortest paths (Figure
+\ref{fig:flowsfacetunweighted}). On the other hand, cycleways are not
+part of unweighted shortest paths, and so very little of the cycling
+demand is routed through them. In the weighted network, cycleways are
+much better utilized, and the majority of the cycling demand passes
+through tertiary roads, as expected.
 
@@
 
-The results of routing potential cycling demand on the weighted and unweighted networks are understandably quite different.
-From Figure \ref{fig:flowsfacetunweighted} we can see that trunk and primary roads are the most efficient means of traversing the road network of Manchester.
-Both of these road types are classified as Primary A roads according to the UK Department for Transport's road classification (Table \ref{table:osmroadtypes}), and are therefore part of the Primary Route Network (PRN) (DfT, 2012).
-The PRN has the widest, most direct roads on the network, and carries most of the through traffic.
-This includes freight, with all roads in the PRN being required by law to provide unrestricted access to trucks up to 40 tonnes (DfT, 2012).
-
-We choose to avoid routing the potential cycling demand on Primary A Roads for the following 2 reasons:
+The results of routing potential cycling demand on the weighted and
+unweighted networks are understandably quite different. From Figure
+\ref{fig:flowsfacetunweighted} we can see that trunk and primary roads
+are the most efficient means of traversing the road network of
+Manchester. Both of these road types are classified as Primary A roads
+according to the UK Department for Transport's road classification
+(Table \ref{table:osmroadtypes}), and are therefore part of the Primary
+Route Network (PRN) (DfT, 2012). The PRN has the widest,
+most direct roads on the network, and carries most of the through
+traffic. This includes freight, with all roads in the PRN being required
+by law to provide unrestricted access to trucks up to 40 tonnes
+(DfT, 2012).
+
+We choose to avoid routing the potential cycling demand on Primary A
+Roads for the following 2 reasons:
 
@@
 \item
-  \textbf{Logistical Difficulty}: Changes on these roads need to be agreed upon by all affected authorities (DfT, 2012), which may prove to be difficult.
+  \textbf{Logistical Difficulty}: Changes on these roads need to be agreed
+  upon by all affected authorities (DfT, 2012), which
+  may prove to be difficult.
 \item
-  \textbf{Low Traffic Neighborhoods (LTNs)}: The UK government is aiming to restrict access to motorized vehicles on residential roads to create LTNs (DfT, 2020a).
-This is part of a policy to prevent automobile rat-running and make streets more accessible to cyclists and pedestrians.
-Under such a policy, Primary A roads would become even more essential for motorized traffic and it would be more difficult to reallocate road space on these roads to cyclists.
+  \textbf{Low Traffic Neighborhoods (LTNs)}: The UK government is aiming to
+  restrict access to motorized vehicles on residential roads to create
+  LTNs (DfT, 2020a). This is part of a policy to
+  prevent automobile rat-running and make streets more accessible to
+  cyclists and pedestrians. Under such a policy, Primary A roads would
+  become even more essential for motorized traffic and it would be
+  more difficult to reallocate road space on these roads to cyclists.
 \end{enumerate}
 
-Figure \ref{fig:flowsfacetweighted} shows that routing on the weighted network significantly reduces flow on the trunk and primary roads, but does not eliminate it completely.
-This is intentional, as the impedance on these roads is only slightly higher than the remaining road types (See Table \ref{table:weightprofiles}).
-Potential cycling demand is only routed on these roads if there are no routes through other roads that offer comparable directness.
-
-Banning cycling flow completely on trunk and primary roads may result in excessively circuitous paths, as seen in Figure \ref{fig:boxplotcircuity}.
-When routing using the \emph{weighted} weighting profile in Table \ref{table:weightprofiles}, we see that shortest paths increase by less than 5\% on average from unweighted shortest paths, with the largest increases still below 30\%.
-When routing on primary and trunk roads is banned (\emph{weighted\_2} profile in Table \ref{table:weightprofiles}), the average increase relative to unweighted shortest paths rises to 10\%, with certain locations experiencing more significant negative effects on accessibility.
-Given that cyclists will only deviate from shortest paths by a certain amount to access better cycling infrastructure (as explained in Section \ref{introduction}), allowing flow on some stretches of trunk and primary roads is necessary to insure cycling uptake and equitable access to cycling infrastructure.
-In its new vision for walking and cycling, the Department for Transport acknowledges that minimal segregated stretches of bicycle lanes on main roads will be necessary to avoid circuitous cycling networks (DfT, 2020a).
+Figure \ref{fig:flowsfacetweighted} shows that routing on the weighted
+network significantly reduces flow on the trunk and primary roads, but
+does not eliminate it completely. This is intentional, as the impedance
+on these roads is only slightly higher than the remaining road types (See
+Table \ref{table:weightprofiles}). Potential cycling demand is only
+routed on these roads if there are no routes through other roads that
+offer comparable directness.
+
+Banning cycling flow completely on trunk and primary roads may result in
+excessively circuitous paths, as seen in Figure
+\ref{fig:boxplotcircuity}. When routing using the \emph{weighted} weighting
+profile in Table \ref{table:weightprofiles}, we see that shortest paths
+increase by less than 5\% on average from unweighted shortest paths, with
+the largest increases still below 30\%. When routing on primary and trunk
+roads is banned (\emph{weighted\_2} profile in Table
+\ref{table:weightprofiles}), the average increase relative to
+unweighted shortest paths rises to 10\%, with certain locations
+experiencing more significant negative effects on accessibility. Given
+that cyclists will only deviate from shortest paths by a certain amount
+to access better cycling infrastructure (as explained in Section
+\ref{introduction}), allowing flow on some stretches of trunk and
+primary roads is necessary to insure cycling uptake and equitable access
+to cycling infrastructure. In its new vision for walking and cycling,
+the Department for Transport acknowledges that minimal segregated
+stretches of bicycle lanes on main roads will be necessary to avoid
+circuitous cycling networks (DfT, 2020a).
 
@@
 
-Weighting the road network also allows us to better utilize existing cycling infrastructure, as can be seen by the higher flow on cycleways in Figure \ref{fig:flowsfacetweighted}.
-Again, the small differences in impedance between cycleways and other road types mean that cycleways that require significant deviation are not routed on.
-
-It should be reiterated that the weighting profile used for routing has been developed for the purposes of this study.
-It creates a hierarchy of road preference that is grounded in cyclist preferences and government plans to create LTNs.
-A wide range of weighting profiles could be used to represent different types of cyclists and road environments, as described in Gehrke et al. (2020) and Furth et al. (2016).
-A sensitivity analysis could be done to determine an optimal weighting profile, but given the variation in city road networks, this would probably require calibration to the specific city.
-More accurate routing could be carried out given the availability of road-level data.
-In such cases we would add additional impedance to specific roads, giving more useful routing results than the current approach which considers all roads of the same type to be equivalent.
-
-One use-case of such granular data would be to identify roads that serve schools.
-The Department of Transport notes that the number of school children being driven to school has trebled over the past 40 years (DfT, 2020a), and so having cycling infrastructure serving schools is key to achieving the government target of getting more children to cycle.
-This would not be difficult, as over 75\% of children in the UK live within a 15 minute cycle from their school (DfT, 2020b).
-Goodman et al. (2019) show that if dutch levels of cycling were achieved in the UK, the \% of children cycling to school could increase from 1.8\% to 41\%.
-In their typology of cyclists, Dill and McNeil (2013) found that a majority of people who say they would never cycle had never cycled to school, whereas confident cyclists were those most likely to have cycled to school.
-Getting people to cycle from a young age is therefore key to achieving societal change in commuting habits.
+Weighting the road network also allows us to better utilize existing
+cycling infrastructure, as can be seen by the higher flow on cycleways
+in Figure \ref{fig:flowsfacetweighted}. Again, the small differences in
+impedance between cycleways and other road types mean that cycleways
+that require significant deviation are not routed on.
+
+It should be reiterated that the weighting profile used for routing has
+been developed for the purposes of this study. It creates a hierarchy of
+road preference that is grounded in cyclist preferences and government
+plans to create LTNs. A wide range of weighting profiles could be used
+to represent different types of cyclists and road environments, as
+described in Gehrke et al. (2020) and Furth et al. (2016). A sensitivity
+analysis could be done to determine an optimal weighting profile, but
+given the variation in city road
+networks, this would probably
+require calibration to the specific city. More accurate routing could be
+carried out given the availability of road-level data. In such cases we
+would add additional impedance to specific roads, giving more useful
+routing results than the current approach which considers all roads
+of the same type to be equivalent.
+
+One use-case of such granular data would be to identify roads that serve
+schools. The Department of Transport notes that the number of school
+children being driven to school has trebled over the past 40 years
+(DfT, 2020a), and so having cycling infrastructure
+serving schools is key to achieving the government target of getting
+more children to cycle. This would not be difficult, as over 75\% of
+children in the UK live within a 15 minute cycle from their school
+(DfT, 2020b). Goodman et al. (2019) show that if
+dutch levels of cycling were achieved in the UK, the \% of children
+cycling to school could increase from 1.8\% to 41\%. In their typology of
+cyclists, Dill and McNeil (2013) found that a majority of people who say they
+would never cycle had never cycled to school, whereas confident cyclists
+were those most likely to have cycled to school. Getting people to cycle
+from a young age is therefore key to achieving societal change in
+commuting habits.
 
 \hypertarget{road-segment-prioritization}{%
-\subsection{Road Segment Prioritization}\label{road-segment-prioritization}}
-
-After routing the potential cycling demand onto the road network using weighted shortest paths, we have estimates for the cumulative potential cycling demand passing through all road segments.
-This cumulative demand (referred to as \emph{flow}) is then used as a basis for identifying segments that are in most need of investment in segregated cycling infrastructure.
-In doing so, we must account for the motivations and deterrents for cycling identified in Section \ref{introduction}, namely direct and well-connected routes.
-
-A range of algorithms could be used for prioritization.
-Because policy priorities vary, we present two algorithms.
-Both utilize existing infrastructure from the beginning and allow us to compare a solution that focuses on utilitarianism to one that focuses on egalitarianism.
-In both algorithms, links are selected iteratively and the iteration at which each link is added to the solution is recorded.
-Investments in cycling infrastructure can be limited by budget constraints, so it can be useful to see where best to allocate a defined length of segregated infrastructure.
-In order to incorporate egalitarian principles in our approach, we use community detection to partition the study area and distribute investment over the different subdivisions.
+\section{Road Segment Prioritization}\label{road-segment-prioritization}}
+
+After routing the potential cycling demand onto the road network using
+weighted shortest paths, we have estimates for the cumulative potential
+cycling demand passing through all road segments. This cumulative demand
+(referred to as \emph{flow}) is then used as a basis for identifying segments
+that are in most need of investment in segregated cycling
+infrastructure. In doing so, we must account for the motivations and
+deterrents for cycling identified in Section \ref{introduction}, namely
+direct and well-connected routes.
+
+A range of algorithms could be used for prioritization. Because policy
+priorities vary, we present two algorithms. Both utilize existing
+infrastructure from the beginning and allow us to compare a solution
+that focuses on utilitarianism to one that focuses on egalitarianism. In
+both algorithms, links are selected iteratively and the iteration at
+which each link is added to the solution is recorded. Investments in
+cycling infrastructure can be limited by budget constraints, so it can
+be useful to see where best to allocate a defined length of segregated
+infrastructure. In order to incorporate egalitarian principles in our
+approach, we use community detection to partition the study area and
+distribute investment over the different subdivisions.
 
 \hypertarget{community-detection}{%
-\subsubsection{Community Detection}\label{community-detection}}
-
-As explained in Section \ref{ethical-underpinnings-and-proposed-approach}, a major challenge facing `top-down' planning approaches is how to incorporate egalitarian principles by fairly distributing investments in cycling infrastructure.
-One way of quantifying this is to split up the city into smaller geospatial areas and target equal investment in each of those areas.
-This approach could also help ensure that on-the-ground surveys are made by local stakeholders, an important component of the planning process (Parkin, 2018).
-Community detection offers us a way to delineate such a split; cyclists are limited in their commuting distance (see Figure \ref{fig:cyclinghistmanc}), and so trip attractors are more likely to have a local catchment area of cyclists.
+\subsection{Community Detection}\label{community-detection}}
+
+As explained in Section
+\ref{ethical-underpinnings-and-proposed-approach}, a major challenge
+facing `top-down' planning approaches is how to incorporate egalitarian
+principles by fairly distributing investments in cycling infrastructure.
+One way of quantifying this is to split up the city into smaller
+geospatial areas and target equal investment in each of those areas.
+This approach could also help ensure that on-the-ground surveys are made
+by local stakeholders, an important component of the planning process
+(Parkin, 2018). Community detection offers us a way to
+delineate such a split; cyclists are limited in their commuting distance
+(see Figure \ref{fig:cyclinghistmanc}), and so trip attractors are more
+likely to have a local catchment area of cyclists.
 
@@
 
-In our case, the network is the city; the nodes are the population-weighted MSOA centroids and the links connecting each MSOA pair are weighted by the potential cycling demand between them.
-The Louvain method (Blondel et al., 2008) is used to separate MSOAs into communities.
-Potential cycling demand is used since we assume that this is what the cycling demand will be once the cycling infrastructure is added.
-To assign road links to communities, the following steps are carried out:
+In our case, the network is the city; the nodes are the
+population-weighted MSOA centroids and the links connecting each MSOA
+pair are weighted by the potential cycling demand between them. The
+Louvain method (Blondel et al., 2008) is used to separate MSOAs into
+communities. Potential cycling demand is used since we assume that this
+is what the cycling demand will be once the cycling infrastructure is
+added. To assign road links to communities, the following steps are
+carried out:
 
@@
 
-The results show that Manchester can be split into four large communities and one small one (Figure \ref{fig:communitiesmanchester}).
+The results show that Manchester can be split into four large
+communities and one small one (Figure \ref{fig:communitiesmanchester}).
 
@@
 \hypertarget{algorithm-1-utilitarian-expansion}{%
-\subsubsection{Algorithm 1: Utilitarian Expansion}\label{algorithm-1-utilitarian-expansion}}
+\subsection{Algorithm 1: Utilitarian Expansion}\label{algorithm-1-utilitarian-expansion}}
 
@@
 
-This algorithm ensures that the resulting network is connected.
-It also satisfies the directness criteria, since links on the weighted shortest paths are those that have the highest flow passing through them (this is a result of the routing in Section \ref{routing}.
+This algorithm ensures that the resulting network is connected. It also
+satisfies the directness criteria, since links on the weighted shortest
+paths are those that have the highest flow passing through them (this is
+a result of the routing in Section \ref{routing}.
 
 \hypertarget{algorithm-2-egalitarian-expansion}{%
-\subsubsection{Algorithm 2: Egalitarian Expansion}\label{algorithm-2-egalitarian-expansion}}
-
-The first algorithm focuses on connectivity and directness, but not on fairly distributing investment.
-The latter is not a requirement for increasing cycling uptake, but it is fundamental for spatial equity, as explained in Section \ref{ethical-underpinnings-and-proposed-approach}.
-This algorithm incorporates the ideal of fair distribution by ensuring that investment is distributed between the defined communities.
+\subsection{Algorithm 2: Egalitarian Expansion}\label{algorithm-2-egalitarian-expansion}}
+
+The first algorithm focuses on connectivity and directness, but not on
+fairly distributing investment. The latter is not a requirement for
+increasing cycling uptake, but it is fundamental for spatial equity, as
+explained in Section \ref{ethical-underpinnings-and-proposed-approach}. This algorithm incorporates
+the ideal of fair distribution by ensuring that investment is distributed between the defined communities.
 This is done using the following logic:
@@
 
-Even though we may end up with a more disconnected network, we will have separate connected networks in each community.
-Given that communities are defined by having more internal flow than external flow, this is a satisfactory solution.
-
-The results of the community detection are used to evaluate the algorithms.
-This is done by looking at the \emph{person-km satisfied} as cycling infrastructure is added.
-Person-km is a measure of the total km cycled on a road segment, so it is the product of the number of potential commuters cycling on that road segment (\(flow\)) and the length of the segment in km (\(l\)).
-For each road segment, the person-km is equal to \(flow * l\).
-In the case of Manchester, Table \ref{tab:personkmtable} shows that almost half of the person-km is in community 1, while only 0.5\% of total person-km on the network is in community 5.
+Even though we may end up with a more disconnected network, we will have
+separate connected networks in each community. Given that communities
+are defined by having more internal flow than external flow, this is a
+satisfactory solution.
+
+The results of the community detection are used to evaluate the
+algorithms. This is done by looking at the \emph{person-km satisfied} as
+cycling infrastructure is added. Person-km is a measure of the total km
+cycled on a road segment, so it is the product of the number of
+potential commuters cycling on that road segment (\(flow\)) and the length
+of the segment in km (\(l\)). For each road segment, the person-km is
+equal to \(flow * l\). In the case of Manchester, Table
+\ref{tab:personkmtable} shows that almost half of the person-km is in
+community 1, while only 0.5\% of total person-km on the network is in
+community 5.
 
@@
 
-Looking at the person-km satisfied (Figure \ref{fig:growthtotal}), we see that the incremental addition of cycling infrastructure is better distributed between communities using Algorithm 2; equal distribution of investment results in the gain in \% of person km satisfied in each community being inversely correlated with the size of the community.
-In addition, we find that the restrictions imposed by Algorithm 2 on the network expansion do not seem to have a noticeable effect on the city-wide \% of person-km satisfied.
-Comparing both algorithms, we can see that Algorithm 1 provides only marginally quicker city-wide gains than Algorithm 2.
+Looking at the person-km satisfied (Figure \ref{fig:growthtotal}), we
+see that the incremental addition of cycling infrastructure is better
+distributed between communities using Algorithm 2; equal distribution of
+investment results in the gain in \% of person km satisfied in each
+community being inversely correlated with the size of the community. In
+addition, we find that the restrictions imposed by Algorithm 2 on the
+network expansion do not seem to have a noticeable effect on the
+city-wide \% of person-km satisfied. Comparing both algorithms, we can
+see that Algorithm 1 provides only marginally quicker city-wide gains
+than Algorithm 2.
 
@@
 
-Figure \ref{fig:growth3MapandBar} gives us a geographic representation of the results from Algorithm 2; it shows when each link was added to the solution (first 100km, second 100km, etc).
-We can see that, generally, road segments around cycling infrastructure are prioritized, except for those neighboring cycling infrastructure on the very periphery.
-The first 100km is also spatially distributed across the city, with no apparent bias towards a particular area.
-
-It is also important to understand how the different highway types contribute to the proposed network.
-Figure \ref{fig:growth3MapandBar} shows that most of the flow will be on residential and tertiary roads, as expected from the weighting profile defined in Table \ref{table:weightprofiles}.
+Figure \ref{fig:growth3MapandBar} gives us a geographic representation
+of the results from Algorithm 2; it shows when each link was added to
+the solution (first 100km, second 100km, etc). We can see that,
+generally, road segments around cycling infrastructure are prioritized,
+except for those neighboring cycling infrastructure on the very
+periphery. The first 100km is also spatially distributed across the
+city, with no apparent bias towards a particular area.
+
+It is also important to understand how the different highway types
+contribute to the proposed network. Figure \ref{fig:growth3MapandBar}
+shows that most of the flow will be on residential and tertiary roads,
+as expected from the weighting profile defined in Table
+\ref{table:weightprofiles}.
 
@@
 \hypertarget{connectivity}{%
-\subsubsection{Connectivity}\label{connectivity}}
-
-Existing cycling infrastructure is made up of many disconnected components.
-Both Algorithm 1 and 2 start with all existing segregated cycling infrastructure and aim to create an efficient, connected network.
-Figure \ref{fig:componentsandGCC} shows that both algorithms gradually reduce the number of components as more infrastructure is added, but Algorithm 2 is able to provide better connectivity with less investment.
-
-Consistent growth can also be seen for the size of the Largest Connected Component in the proposed bicycle network (Figure \ref{fig:componentsandGCC}).
-Here, however, we find that there is little difference between both algorithms.
+\subsection{Connectivity}\label{connectivity}}
+
+Existing cycling infrastructure is made up of many disconnected
+components. Both Algorithm 1 and 2 start with all existing segregated
+cycling infrastructure and aim to create an efficient, connected
+network. Figure \ref{fig:componentsandGCC} shows that both algorithms
+gradually reduce the number of components as more infrastructure is
+added, but Algorithm 2 is able to provide better connectivity with less
+investment.
+
+Consistent growth can also be seen for the size of the Largest Connected
+Component in the proposed bicycle network (Figure
+\ref{fig:componentsandGCC}). Here, however, we find that there is little
+difference between both algorithms.
 
@@
 
-Overall, the improved distribution of investment shown by Algorithm 2 does not seem to come at the cost of poorer connectivity or less city-wide gains when compared to Algorithm 1.
+Overall, the improved distribution of investment shown by Algorithm 2 does not seem to come at the cost
+of poorer connectivity or less city-wide gains when compared to Algorithm 1.
 The results therefore advocate for the incorporation of egalitarian principles in bicycle network planning.
@@
 
-This paper demonstrates an approach for prioritizing investments in city-wide cycleway networks.
-The approach aims to respect both the needs of the users and the ambitions of stakeholders working at local or regional levels.
-The results, including detailed route network maps based on current travel behaviour derived from OD data, can provide an evidence-base for designing direct, connected, and low-stress networks.
-A core contribution of this approach is that it insures that investments in cycling infrastructure are distributed fairly between different communities in a city.
-
-Given that the ``most essential activity entailed in the design of cycle-friendly infrastructure is developing a cycle network'' (Parkin, 2018), we believe that the approach has great potential to inform investment in cities such as Manchester where there is political will to invest in cycling long-term.
-A benefit of the approach is that it has relatively modest data requirements: only the road network (from OSM), its topography (from satellite imagery), and OD data (from the national census in this case) are needed, opening up the possibility of deploying the approach in other cities\footnote{The results are easily reproducible for all UK cities, and can also be reproduced for cities elsewhere given the availability of commuter data.
+This paper demonstrates an approach for prioritizing
+investments in city-wide cycleway networks. The approach aims to respect
+both the needs of the users and the ambitions of stakeholders working at
+local or regional levels. The results, including detailed route network
+maps based on current travel behaviour derived from OD data, can provide
+an evidence-base for designing direct, connected,
+and low-stress networks. A core contribution of this approach is that it insures
+that investments in cycling infrastructure are distributed fairly between different
+communities in a city.
+
+Given that the ``most essential activity entailed in the design of
+cycle-friendly infrastructure is developing a cycle network''
+(Parkin, 2018), we believe that the approach has great
+potential to inform investment in cities such as Manchester where there
+is political will to invest in cycling long-term. A benefit of the
+approach is that it has relatively modest data requirements: only the
+road network (from OSM), its topography (from satellite imagery), and
+OD data (from the national census in this case) are
+needed, opening up the possibility of deploying the approach in other
+cities\footnote{The results are easily reproducible for all UK cities, and can also be reproduced for cities elsewhere given the availability of commuter data.
 Instructions for reproducing the results are detailed in the \textbf{README} file in the R folder of the .zip file that accompanies this paper (Link to public github repo to be added when paper is published)}.
 
-The approach can identify not only where there is high potential for cycling but also trade-offs between stress levels (related to motor traffic) and directness.
-The results can therefore be used as a basis for recommendations on road space reallocation \emph{and} new infrastructure to unlock potential cycling demand.
-The approach encourages consideration of a wider range of preferences and needs than previous approaches that focus only on absolute potential.
-Moreover, the inclusion of egalitarian principles in scenarios of change encourages investment in cycling infrastructure that addresses geographical and social inequalities.
-This is done by explicitly ensuring that infrastructure improvements are pursued in different parts of the city simultaneously.
-This approach compares well with one that focuses exclusively on maximizing the number of potential cycling trips that benefit from investment, showing that a fair distribution can be achieved without compromising on overall benefit.
-This ability to address inequalities in network prioritization is particularly important given research showing substantial \emph{inequalities} around transport provision in general and cycling uptake and investment in particular (Lucas et al., 2016; Vidal Tortosa et al., 2021).
-
-The approach is not without limitations.
-The level of detail is only as good as the granularity of the available data (in this case relatively coarse zones).
-Iacono et al. (2010) note that such large travel zones are not ideal for understanding route choice behaviour of cyclists and pedestrians.
-They also give rise to an `ecological fallacy' whereby average characteristics are assumed to apply to all residents of the aggregated geographical area, suggesting a need for applying the methods to more granular OD data (and for governments and other data-collecting organisations to make OD data more readily available).
-The potential demand calculation is also based on cycling in the traditional sense, and does not consider the effect of micro-mobility on reducing topology-related impedance to cycling.
+The approach can identify not only where there is high
+potential for cycling but also trade-offs between
+stress levels (related to motor traffic) and directness.
+The results can therefore be used as a basis for recommendations on
+road space reallocation \emph{and} new infrastructure to unlock potential cycling demand.
+The approach encourages consideration of a wider range of preferences
+and needs than previous approaches that focus only on absolute potential.
+Moreover, the inclusion of egalitarian principles in scenarios of change
+encourages investment in cycling infrastructure that addresses geographical and
+social inequalities. This is done by explicitly ensuring that infrastructure improvements
+are pursued in different parts of the city simultaneously. This approach compares well with one
+that focuses exclusively on maximizing the number of potential cycling trips that
+benefit from investment, showing that a fair distribution can be achieved without
+compromising on overall benefit.
+This ability to address inequalities in network prioritization is particularly
+important given research showing substantial \emph{inequalities}
+around transport provision in general and cycling uptake and investment
+in particular (Lucas et al., 2016; Vidal Tortosa et al., 2021).
+
+The approach is not without limitations. The level of detail is only as
+good as the granularity of the available data (in this case relatively
+coarse zones). Iacono et al. (2010) note that such large travel zones
+are not ideal for understanding route choice behaviour of cyclists and
+pedestrians. They also give rise to an `ecological fallacy' whereby
+average characteristics are assumed to apply to all residents of the
+aggregated geographical area, suggesting a need for applying the methods
+to more granular OD data (and for governments and other
+data-collecting organisations to make OD data more
+readily available).
+The potential demand calculation is also based on cycling in the traditional sense, and does not consider the
+effect of micro-mobility on reducing topology-related impedance to cycling.
 Given that we are proposing an approach which can accommodate any cycling uptake functions, this is an acceptable limitation.
-The network prioritization is based strongly on potential cycling demand calculations.
-While this provides a solid supplementary tool for network planning, it would be more useful as part of a multi-criteria approach that considers on-the ground information from existing cyclists, road geometries, and traffic analysis.
-Grisé and El-Geneidy (2018) propose a multi-criteria approach that includes the first of these points.
-
-The approach is also focused solely on the allocation of cycling infrastructure, and does not consider the larger political and regulatory environment that needs to exist to promote cycling; while segregated, connected, and direct cycling infrastructure is key to achieving high levels of cycling, research has shown that it cannot exist in a vacuum.
-Wardman et al.
-(2007) developed a mode choice model for the UK and their results showed that improved cycling infrastructure on its own only had modest impacts on mode shift, and even the unlikely scenario of all urban routes being serviced by segregated bike lanes was forecast to increase cycling mode share by only 3\%.
-International research shows that cities that invest in more comprehensive cycling projects have a more significant increase in the number of cyclists as well as the cycling mode share (Parkin, 2018; Pucher et al., 2010).
-These cities do not just focus on infrastructure, but on general policies as well as restricting car use.
-Evaluation of policies in Denmark and Germany and the Netherlands has shown that their high cycling mode share is down to a broader set of soft and hard policies.
-Hard policies include traffic calming, filtered permeability interventions, cycling rights of way, bike parking, integration with the public transport network, and making driving cars both expensive and inconvenient, while soft policies include marketing and awareness campaigns (Gössling, 2013; Pucher and Buehler, 2008).
-While these policies are outside the scope of this research, it is important to recognize their key role in bringing about an increase in levels of cycling.
+
+The approach is also focused solely on the allocation of cycling
+infrastructure, and does not consider the larger political and
+regulatory environment that needs to exist to promote cycling; while
+segregated, connected, and direct cycling infrastructure is key to
+achieving high levels of cycling, research has shown that it cannot
+exist in a vacuum. Wardman et al. (2007) developed a mode choice model
+for the UK and their results showed that improved cycling infrastructure
+on its own only had modest impacts on mode shift, and even the unlikely
+scenario of all urban routes being serviced by segregated bike lanes was
+forecast to increase cycling mode share by only 3\%. International
+research shows that cities that invest in more comprehensive cycling
+projects have a more significant increase in the number of cyclists as
+well as the cycling mode share (Parkin, 2018; Pucher et al., 2010). These cities do not just focus on
+infrastructure, but on general policies as well as restricting car use.
+Evaluation of policies in Denmark and Germany and the Netherlands has
+shown that their high cycling mode share is down to a broader set of
+soft and hard policies. Hard policies include traffic calming, filtered permeability
+interventions, cycling rights of way, bike parking, integration with the
+public transport network, and making driving cars both expensive and
+inconvenient, while soft policies include marketing and awareness campaigns (Gössling, 2013; Pucher and Buehler, 2008).
+While these policies are outside the scope of this research, it is important to recognize their key role in
+bringing about an increase in levels of cycling.
 
 Consideration of these limitations suggest future directions of research.
-New datasets and continuously evolving computational (hardware and software) capabilities should enable the data related issues to be overcome as open datasets, and our ability to process them, improve.
-Plugging-in alternative uptake models could help address the relatively narrow definition of `potential' used in this paper, to consider broader social factors.
-A promising area of future research in this direction could be to integrate a wider range of modes, including e-bikes and e-scooters, into the analysis.
-Currently scope for progress in this direction is partially restricted by the lack of data on the proliferation of these modes, raising the point that surveys could replace the broad category of `Bicycle' with a range of small modes such as `pedal cycles' (including tricycles and recumbent cycles), e-bikes and other forms of micro-mobility.
-
-Despite the rapid growth of these alternative modes, there is little doubt that the humble bicycle is a key ingredient in the ongoing shift towards active transport.
-Recognition of the importance of this shift for improved health and well-being of citizens, and the wider challenge to decarbonise the global economy, has grown with pressures on health and public transport systems during the pandemic.
-The success of policies to accelerate cycling uptake depends on a range of factors including, vitally, the design of the cycling network for potential cyclists.
-The approach presented in this paper provides a strong evidence base, that considers both cycling potential and social equity, for designing joined-up and cost-effective strategic cycle networks.
+New datasets and continuously evolving computational (hardware and software)
+capabilities should enable the data related issues to be overcome as open
+datasets, and our ability to process them, improve.
+Plugging-in alternative uptake models could help address the relatively narrow
+definition of `potential' used in this paper, to consider broader social factors.
+A promising area of future research in this direction could be
+to integrate a wider range of modes, including e-bikes and e-scooters, into the
+analysis. Currently scope for progress in this direction is
+partially restricted by the lack of data on the
+proliferation of these modes, raising the point that surveys could
+replace the broad category of `Bicycle' with a range of small modes such as
+`pedal cycles' (including tricycles and recumbent cycles), e-bikes
+and other forms of micro-mobility.
+
+Despite the rapid growth of these alternative modes, there is little doubt that
+the humble bicycle is a key ingredient in the ongoing shift towards active transport.
+Recognition of the importance of this shift for improved health and well-being
+of citizens, and the wider challenge to decarbonise the global economy,
+has grown with pressures on health and public transport systems during the pandemic.
+The success of policies to accelerate cycling uptake depends on a range of
+factors including, vitally, the design of the
+cycling network for potential cyclists.
+The approach presented in this paper provides a strong evidence base,
+that considers both cycling potential and social equity,
+for designing joined-up and cost-effective strategic cycle networks.
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-grise2018if}{}}%
-Grisé, E., El-Geneidy, A., 2018.
-If we build it, who will benefit?
-A multi-criteria approach for the prioritization of new bicycle lanes in quebec city, canada.
-Journal of transport and land use 11, 217--235.
-
 \leavevmode\vadjust pre{\hypertarget{ref-iacono2010measuring}{}}%
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-wysling2022improve}{}}%
-Wysling, L., Purves, R.S., 2022.
-Where to improve cycling infrastructure?
-Assessing bicycle suitability and bikeability with open data in the city of paris.
-Transportation research interdisciplinary perspectives 15, 100648.
-
 \end{CSLReferences}

### 2023-09-23 paper_drafts/paper_round_2.Rmd
edits for submission 3

--- 2023-09-22 README.Rmd
@@
 output:
-   bookdown::github_document2:
-#    bookdown::pdf_document2:
+ # bookdown::github_document2:
+   bookdown::pdf_document2:
  # bookdown::word_document2:
@@
 file.rename("README.pdf", "cycle-network-paper.pdf")
-```
-
-```{r, include=FALSE}
-library(tidyverse)
-library(tmap)
-knitr::opts_chunk$set(warning = FALSE)
 ```
@@
 It then introduces the network expansion algorithms, and compares an approach grounded in 'egalitarianism' to one grounded in 'utilitarianism'. 
+As mentioned above, Manchester is used as a case study. 
 An overview of the steps involved is shown in Figure \@ref(fig:methodflowchart).
@@
 knitr::include_graphics("extras/methodology_flowchart.png")
-```
-
-## Case study area
-
-Manchester is a city in the North West of England.
-"Manchester" can be used to refer to multiple things:
-
-- Manchester City Council, which is a local government district that includes the city centre and surrounding areas.
-- The Greater Manchester region, which is a metropolitan county that includes the city of Manchester and 9 other local government districts.
-- The city of Manchester, which is used by visitors to refer to the city centre and surrounding areas.
-This definition if formalised in the Major Towns and Cities (TCITY) dataset provided by the Office for National Statistics.
-
-We use the TCITY definition of Manchester, which was "developed specifically for the production and analysis of statistics" and which is shown in Figure \@ref(fig:manchestermap).
-Manchester is a cosmopolitan city with a large university population.
-As shown in Figure \@ref(fig:manchestermap) (right), the high density of the central area allows high levels of cycling, although the surrounding areas have low levels of cycling.
-Manchester and its surrounding have many deprived areas.
-
-```{r manchestermap, echo=FALSE, fig.cap="Maps of Greater Manchester (top) and the case study area (City of Manchester)"}
-greater_manchester = pct::get_pct_zones(region = "greater-manchester", geography = "msoa")
-greater_manchester = greater_manchester |>
-  mutate(pcycle = bicycle / all * 100)
-# names(greater_manchester)
-msoas = sf::read_sf("./data/Manchester/msoas_geometry.shp")
-msoas_union = sf::st_union(msoas)
-greater_manchester_lads = greater_manchester |>
-  group_by(lad_name) |>
-  summarise()
-?tm_scale_bar
-m1 = tm_shape(greater_manchester_lads) +
-  tm_polygons(col = "lad_name", palette = "Set1", alpha = 0.2, title = "") +
-  tm_layout(legend.position = c("left", "bottom")) +
-  tm_shape(msoas_union) +
-  tm_borders(col = "black", lwd = 2) +
-  tm_scale_bar(breaks = (0:3)*5)
-msoas_centroids = sf::st_centroid(msoas)
-pct_zones_in_manchester = greater_manchester[msoas_centroids, ] 
-m2 = tm_shape(greater_manchester, bb = msoas) +
-  tm_fill("pcycle", palette = "Blues", title = "% cycling to work\n(2011 Census)") +
-  tm_layout(legend.position = c("left", "bottom"), legend.bg.color = "white", legend.bg.alpha = 0.5) +
-  tm_shape(msoas_union) +
-  tm_borders(col = "black", lwd = 2) +
-  tm_scale_bar(breaks = 0:5)
-tmap_arrange(m1, m2, nrow = 1)
 ```

### 2023-09-23 R/archived/mode_share_2021.R

@@ -1,0 +1,60 @@
+library(tidyverse)
+library(ggrepel)
+library(sf)
+
+# towns and cities dataset 
+cities <- read_csv('data-raw/MSOA_2011_to_Major_Towns_and_Cities.csv') 
+
+# census 2021 commuting mode share
+census_2021_modes <- read_csv('data-raw/census_2021_method_travel_work.csv', , skip = 6) %>%
+  rename(MSOA21CD = mnemonic)
+  
+
+# msoa 2011 to msoa 2021
+msoa_lookup <- read_csv('data-raw/MSOA_2011_to_MSOA_2021.csv') %>%
+  select(MSOA11CD, MSOA21CD)
+
+
+# match census 2021 data onto 2011 msoa codes
+census_2021_modes <- census_2021_modes %>%
+  left_join(msoa_lookup, by = "MSOA21CD") %>%
+  relocate(MSOA11CD, .after = MSOA21CD)
+
+# total commuters
+
+
+census_2021_modes <- census_2021_modes %>%
+  mutate(total_commuters = rowSums(across(`Underground, metro, light rail, tram`:`Other method of travel to work`))) 
+  
+
+# ----- join cities dataset and get cycling % per city 
+
+census_2021_cities <- census_2021_modes %>% 
+  left_join(cities, by = "MSOA11CD")
+
+# get cycling mode share
+census_2021_cities <- census_2021_cities %>%
+  group_by(TCITY15NM) %>%
+  summarise(Bicycle = sum(Bicycle), total_commuters = sum(total_commuters)) %>%
+  ungroup() %>%
+  mutate(cycling_perc = round((Bicycle / total_commuters) * 100, 1))
+
+census_2021_cities <- census_2021_cities %>%
+  mutate(manc = ifelse(TCITY15NM == "Manchester", "yes", "no"))
+
+# ----- plot cycling 
+
+
+census_2021_cities %>% #filter(city_origin != 'London') %>%
+  ggplot(aes(total_commuters, cycling_perc)) +
+  geom_point(aes(alpha = sqrt(cycling_perc), col = manc, size = total_commuters), show.legend = FALSE) +
+  # some filtering labels for aesthetic purposes. Add some cities explicitly as they are in my case study
[…]

### 2023-09-23 latexdiff.R

@@ -25,3 +25,13 @@ rmarkdown::render("paper_round_1.Rmd")
 system("latexdiff paper_round_1.tex README.tex > tracked-changes_2.tex")
 tinytex::pdflatex("tracked-changes_2.tex")
 browseURL("tracked-changes_2.pdf")
+
+
+# ROUND 3: Hussein
+
+file.edit("paper_drafts/paper_round_2.Rmd")
+rmarkdown::render("paper_drafts/paper_round_2.Rmd")
+# compare with current version and create tracked changes
+system("latexdiff paper_round_2.tex README.tex > tracked-changes_3.tex")
+tinytex::pdflatex("tracked-changes_3.tex")
+browseURL("tracked-changes_3.pdf")
